### 2024-04-19 notes.tex
init

% Please compile this document using LuaLaTeX.
% XeLaTeX and PDFLaTeX will not work.

\newif\iffull\fulltrue
\newif\ifacm\acmfalse

%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
\ifacm
    \documentclass[acmsmall,screen]{acmart}
    %% For double-blind review submission, w/ CCS and ACM Reference
    %\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
    %% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
    %\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
    %% For single-blind review submission, w/ CCS and ACM Reference
    %\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
    %% For final camera-ready submission, w/ required CCS and ACM Reference
    %\documentclass[acmsmall]{acmart}\settopmatter{}
\else
    \documentclass{extarticle}
\fi

\ifacm
    %%% If you see 'ACMUNKNOWN' in the 'setcopyright' statement below,
    %%% please first submit your publishing-rights agreement with ACM (follow link on submission page).
    %%% Then please update our instructions page and copy-and-paste the NEW commands into your article.
    %%% Please contact us in case of questions; allow up to 10 min for the system to propagate the information.
    %%%
    %%% The following is specific to POPL '22 and the paper
    %%% 'On Incorrectness Logic and Kleene Algebra with Top and Tests'
    %%% by Cheng Zhang, Arthur Azevedo de Amorim, and Marco Gaboardi.
    %%%
    \acmPrice{}
    \acmDOI{10.1145/3498690}
    \acmYear{2022}
    \copyrightyear{2022}
    \acmSubmissionID{popl22main-p160-p}
    \acmJournal{PACMPL}
    \acmVolume{6}
    \acmNumber{POPL}
    \acmArticle{29}
    \acmMonth{1}
    \startPage{1}

    %% Copyright information
    %% Supplied to authors (based on authors' rights management selection;
    %% see authors.acm.org) by publisher for camera-ready submission;
    %% use 'none' for review submission.
    \setcopyright{rightsretained}
    %\setcopyright{acmcopyright}
    %\setcopyright{acmlicensed}
    %\setcopyright{rightsretained}
    %\copyrightyear{2018}           %% If different from \acmYear

    %% Bibliography style
    \bibliographystyle{ACM-Reference-Format}
    %% Citation style
    %% Note: author/year citations are required for papers published as an
    %% issue of PACMPL.
    \citestyle{acmauthoryear}   %% For author/year citations

\else

    %% Bibliography style
    \bibliographystyle{plainnat}

\fi

%%%%% PACKAGES

\ifacm 
\else 
    \usepackage{amsmath}
    \usepackage{amssymb}
    \usepackage{amsthm}
    \usepackage[numbers]{natbib}
    \usepackage[margin=3cm]{geometry}
\fi

% For adding inline comments in the text.
\usepackage[margin=false,inline=true]{fixme}
\FXRegisterAuthor{aaa}{anaaa}{\color{cyan}AAA}
\FXRegisterAuthor{mg}{anmg}{\color{red}MG}
\FXRegisterAuthor{cz}{ancz}{\color{orange}CZ}
% \newcommand{\aaa}[1]{\aaanote{#1}}
% \newcommand{\mg}[1]{\mgnote{#1}}
% \newcommand{\cz}[1]{\cznote{#1}}
\newcommand{\aaa}[1]{}
\newcommand{\mg}[1]{}
\newcommand{\cz}[1]{}

\usepackage{stmaryrd}

\usepackage{stackengine}
\usepackage{mathrsfs}
\usepackage{braket}
\usepackage{annotate-equations}
\usepackage{scalerel}

% graphs
\usepackage{tikz}
\usetikzlibrary{arrows,automata,positioning}

% commutative diagram
\usepackage{tikz-cd}

% inference rule
\usepackage{mathpartir}

% ref
\usepackage{hyperref}
\usepackage{cleveref}

% item spacing
\usepackage{enumitem}

% for code
\usepackage{listings}

% for better table
\usepackage{booktabs}

% subcaption
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption

% unicode math symbols
\usepackage{unicode-math}
% support for hat, overline, underline, vec, and sim combining charactors
\protected\def\afteracc{\directlua{
    local nest = tex.nest[tex.nest.ptr]
    local last = nest.tail
    if not (last and last.id == 18) then
      error'I can only put accents on simple noads.'
    end
    if last.sub or last.sup then
      error'If you want accents on a superscript or subscript, please use braces.'
    end
    local acc = node.new(21, 1)
    acc.nucleus = last.nucleus
    last.nucleus = nil
    local is_bottom = token.scan_keyword'bot' and 'bot_accent' or 'accent'
    acc[is_bottom] = node.new(23)
    acc[is_bottom].fam, acc[is_bottom].char = 0, token.scan_int()
    nest.head = node.insert_after(node.remove(nest.head, last), nil, acc)
    nest.tail = acc
    node.flush_node(last)
  }}
\AtBeginDocument{
\begingroup
  \def\UnicodeMathSymbol#1#2#3#4{%
    \ifx#3\mathaccent
      \def\mytmpmacro{\afteracc#1 }%
      \global\letcharcode#1=\mytmpmacro
      \global\mathcode#1="8000
    \else\ifx#3\mathbotaccentwide
      \def\mytmpmacro{\afteracc bot#1 }%
      \global\letcharcode#1=\mytmpmacro
      \global\mathcode#1="8000
    \fi\fi
  }
  \input{unicode-math-table}
\endgroup
}

% math font, this is needed to render \setminus command
\setmathfont{latinmodern-math}
\setmathfont[range=\setminus]{STIX Two Math}
\setmathfont[range=\similarrightarrow]{STIX Two Math}

% move proof to the end
\iffull
    \usepackage[conf={restate}]{proof-at-the-end}
\else
    \usepackage[conf={restate, no link to proof}]{proof-at-the-end}
\fi
% enable the following to debug proofs
% \renewenvironment{proofEnd}{%
%     \begin{proof}%
% }{%
%     \end{proof}%
% }

\newtheorem{corollary}{Corollary}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}

\newtheorem{definition}{Definition}
\newtheorem*{definition*}{Definition}
\newtheorem{example}{Example}
\newtheorem{remark}{Remark}

%%%% Macros %%%%%

% Math?
\newcommand{\true}{\mathrm{true}}
\newcommand{\false}{\mathrm{false}}
\newcommand{\At}{\mathbf{At}}

% operators
\newcommand{\dom}[1]{\mathrm{dom}(#1)}
\newcommand{\cod}[1]{\mathrm{cod}(#1)}
\DeclareMathOperator{\post}{\mathrm{post}}
\newcommand{\reject}{\mathinner{\mathtt{Reject}}}
\newcommand{\accept}{\mathinner{\mathtt{Accept}}}
\DeclareMathOperator*{\bigplus}{\scalerel*{+}{\sum}}

% commands 
\newcommand{\command}[1]{{\mathtt{#1}}}
\newcommand{\comSkip}{\command{skip}}
\newcommand{\comDiverge}{\command{diverge}}
\newcommand{\comFork}{\command{fork}}
\newcommand{\comJoin}{\command{join}}
\newcommand{\comError}{\command{error ()}}
\newcommand{\comAssume}[1]{\command{assume}~#1}
\newcommand{\comITE}[3]{\command{if}~#1~\command{then}~#2~\command{else}~#3}
\newcommand{\comWhile}[2]{\command{while}~#1~\command{do}~#2}
\newcommand{\comAssign}[2]{#1:= #2}
\newcommand{\comLoop}[1]{\command{loop}~#1}
\newcommand{\comIter}[1]{{#1}^{\star}}
\newcommand{\comBreakN}[1]{\command{break} ~ #1}
\newcommand{\comBreak}{\command{break}}

% texts
\newcommand{\tand}{\text{ and }}
\newcommand{\totherwise}{\text{otherwise}}
\newcommand{\tif}{\text{if }}
\newcommand{\tor}{\text{ or }}
\newcommand{\twhere}{\text{ where }}
\newcommand{\tst}{\text{ s.t.\ }}

% typeset
\newcommand{\itemTitle}[1]{\textbf{#1}}
\newcommand{\keyword}[1]{\emph{#1}}

% set of models
\newcommand{\theoryOf}[1]{\ensuremath{\mathsf{#1}}}
\newcommand{\KA}{\theoryOf{KA}}
\newcommand{\KAT}{\theoryOf{KAT}}
\newcommand{\GKAT}{\theoryOf{GKAT}}
\newcommand{\BExp}{\theoryOf{BExp}}
\DeclareMathOperator{\GS}{\mathrm{GS}}

\newcommand\altxrightarrow[2][0pt]{\mathrel{\ensurestackMath{\stackengine%
  {\dimexpr#1-7.5pt}{\xrightarrow{\phantom{#2}}}{\scriptstyle\!#2\,}%
  {O}{c}{F}{F}{S}}}}

\newcommand{\transvia}[1]{
    \mathrel{\raisebox{-2px}{\(\altxrightarrow[-2px]{#1}\)}}
}

\begin{document}

%% Title information
\title{A Symbolic Decision Procedure for GKAT}         
% \title{Incorrectness Logic and Kleene Algebra With Top and Test}         
                                        %% [Short Title] is optional;
                                        %% when present, will be used in
                                        %% header instead of Full Title.
% \titlenote{}                            %% \titlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'
%% \subtitle is optional

% \subtitlenote{with subtitle note}       %% \subtitlenote is optional;
%                                         %% can be repeated if necessary;
%                                         %% contents suppressed with 'anonymous'

%% Author information
%% Contents and number of authors suppressed with 'anonymous'.
%% Each author should be introduced by \author, followed by
%% \authornote (optional), \orcid (optional), \affiliation, and
%% \email.
%% An author may have multiple affiliations and/or emails; repeat the
%% appropriate command.
%% Many elements are not rendered, but should be provided for metadata
%% extraction tools.
\ifacm
    \author{Cheng Zhang}
    \orcid{0000-0002-8197-6181}            
    \email{czhang03@bu.edu}

    % \author{Arthur Azevedo de Amorim}
    % \orcid{}
    % \email{arthur.aa@gmail.com}

    % \author{Marco Gaboardi}
    % \orcid{}
    % \email{gaboardi@bu.edu}

    \affiliation{
    % \position{Position1}
    \department{Department of Computer Science}              %% \department is recommended
    \institution{Boston University}            %% \institution is required
    \city{Boston}
    \state{MA}
    \postcode{02215}
    \country{USA}                    %% \country is recommended
    }
\else
    \author{
        Cheng Zhang \\
        % \and
        % Arthur Azevedo de Amorim \\
        % \and
        % Marco Gaboardi \\
    }
    \date{}
\fi

\ifacm 
\else 
    % In non-ACM format \maketitle needs to happen before abstract
    \maketitle
\fi

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
\end{abstract}

\ifacm 
%% 2012 ACM Computing Classification System (CSS) concepts
%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
\begin{CCSXML}
<ccs2012>
    <concept>
        <concept_id>10003752.10010124.10010138.10010140</concept_id>
        <concept_desc>Theory of computation~Program specifications</concept_desc>
        <concept_significance>500</concept_significance>
        </concept>
    <concept>
        <concept_id>10003752.10010124.10010138.10010141</concept_id>
        <concept_desc>Theory of computation~Pre- and post-conditions</concept_desc>
        <concept_significance>500</concept_significance>
        </concept>
    <concept>
        <concept_id>10003752.10010124.10010138.10011119</concept_id>
        <concept_desc>Theory of computation~Abstraction</concept_desc>
        <concept_significance>500</concept_significance>
        </concept>
    </ccs2012>
\end{CCSXML}

\ccsdesc[500]{Theory of computation~Program specifications}
\ccsdesc[500]{Theory of computation~Pre- and post-conditions}
\ccsdesc[500]{Theory of computation~Abstraction}

%% Keywords
%% comma separated list
\keywords{Program Reasoning,
  Incorrectness Logic, 
  Hoare Logic,
Kleene Algebra with Tests,
}  %% \keywords are mandatory in final camera-ready submission

\fi

\ifacm 
%% \maketitle
%% Note: in ACM format \maketitle command must come after title commands, author
%% commands, abstract environment, Computing Classification System
%% environment and commands, and keywords command.
\maketitle
\fi

\section{Symbolic Derivatives}

Given a finite set \(K\), a \emph{symbolic Guarded Kleene Coalgebra with Tests} (sGKCT) \(𝒮 ≜ ⟨S, ϵ, δ⟩\) over a boolean algebra \(ℬ\) consists of a state set \(S\) consists of a accepting boolean \(ϵ\) and a transition function \(δ\):
\begin{mathpar}
    ϵ: S → ℬ, \and
    δ: S → S → K → ℬ,
\end{mathpar}
where for all states \(s ∈ S\), all the booleans are ``disjoint''; namely the conjunction of any two expression from the set \(ϵ(s) + \{δ(s, s', p) ∣ s' ∈ S, p ∈ K\}\) are false.
The ordering on \(𝒮\) is defined pointwise, namely 
\[⟨ϵ₁, δ₁⟩ ≤ ⟨ϵ₂, δ₂⟩ \text{ when } ∀ s, s' ∈ S, ∀ p ∈ K, ϵ₁(s) ≤ ϵ₂(s) \text{ and } δ₁(s, s', p) ≤ δ₂(s, s', p),\]
where the ordering on the right hand side is defined in the boolean algebra.

Intuitively, the elements of boolean is treated as the ``symbolic'' transitions, which can be think of as a set of labels by the classical stone's representation theorem. 
In particular, when the boolean algebra is the free boolean algebra \(\BExp_B\) generated by a finite set \(B\), the labels can be considered as atoms, since \(\BExp_B ≅ 2^{\At_B}\). 
Then we can explain the intuition for the definition of GKCT
\begin{itemize}
    \item \(ϵ(s)\) can be thought of as a set of labels that is accepted by the state \(s\).
    \item \(δ(s, s', p)\) denotes all the labels that can transition from \(s\) to \(s'\) while executing \(p\).
\end{itemize}
Then the labels that is not in either operations are implicitly rejecting.

Given a sGKCT \(𝒮 ≜ ⟨S, ϵ, δ⟩\) and a state \(s ∈ S\), we use the notation \(s ⇒ b\) to denote that \(ϵ(s) ≥ b\), and we use the notation \(s \transvia{b ∣ p} s'\) to denote that \(δ(s)₃(s', p) ≥ b\),
where the ordering is the typical ordering in boolean algebra.

The symbolic derivatives for GKAT forms a GKCT where the states are represented by GKAT expression \(\GKAT_{K, B}\), and the boolean algebra is the free boolean algebra \(\BExp_B\) over the test alphabet \(B\):
\begin{mathpar}
    ϵ: \GKAT_{K, B} → \BExp_B; \and  
    δ: \GKAT_{K, B} → \GKAT_{K, B} → K → \BExp_B.
\end{mathpar} 
the accepting map \(ϵ\) and transition map \(δ\) can be defined by the smallest maps that satisfy the following rules:
\begin{mathpar}
    \inferrule{\\}{b ⇒ b} \and 
    \inferrule{\\}{p \transvia{1 ∣ p} 1} \and 
    \inferrule{e ⇒ a}{e +_b f ⇒ b a} \and 
    \inferrule{f ⇒ a}{e +_b f ⇒ b̄ a} \and 
    \inferrule{e \transvia{a ∣ p} e'}{e +_b f \transvia{b ∧ a ∣ p} e'} \and 
    \inferrule{f \transvia{a ∣ p} f'}{e +_b f \transvia{b̄ ∧ a ∣ p} f'} \and 
    \inferrule{e ⇒ a \\ f ⇒ b}{e ⋅ f ⇒ a ∧ b} \and 
    \inferrule{e ⇒ a \\ f \transvia{b ∣ p} f'}{e ⋅ f \transvia{a ∧ b ∣ p} f'} \and 
    \inferrule{e \transvia{b ∣ p} e'}{e ⋅ f \transvia{b ∣ p} e' ⋅ f} \and 
    \inferrule{\\}{e^{(b)} ⇒ b̄} \and 
    \inferrule{e \transvia{a ∣ p} e'}{e^{(b)} \transvia{b ∧ a ∣ p} e' ⋅ {e}^{(b)}}
\end{mathpar}

\section{Implementing derivatives}

Since the rules listed above is monotonic, that is if we enlarge the boolean in the premise, we also enlarge the boolean in the result. 
So we can simply pick the largest boolean on the premise, which is either \(ϵ(e)\) or \(δ(e, e', p)\).
Since the disjunction of two boolean \(a ∨ b\) is the smallest boolean that is greater than both \(a\) and \(b\), thus we can implement \(ϵ\) and \(δ\) by iteration through all the rules, and take the disjunction of all the possible booleans.

\begin{align*}
    ϵ(b) & ≜ b 
        &  ϵ(e +_b f) & ≜ (b ∧ ϵ(e)) ∨ (b̄ ∧ ϵ(f))\\
    ϵ(q) & ≜ 0
        & ϵ(e ⋅ f) & ≜ ϵ(e) ∧ ϵ(f)\\  
    ϵ(e^{(b)}) & ≜ b̄ 
\end{align*}

However, the derivative function \[δ: \GKAT_{K, B} → \GKAT_{K, B} → K → \BExp_B,\] poses challenges in the implementation, since \(\GKAT_{K, B}\) is infinite, so it will be impractical to search through all the expressions.
However, we can treat the function \(δ\) extensionally: \[\GKAT_{K, B} → K → \BExp_B ⊆ 2^{\GKAT_{K, B} × K × \BExp_B},\]
and then because the boolean expression don't overlap for each input, hence the boolean expression are necessarily unequal; 
we can model the set by a partial map, thus all the \(δ\) can be represented as a function to the following partial maps.
\begin{align*}
    δ & : \GKAT_{K, B} → (\BExp_B ↛ \GKAT_{K, B} × K) \\
    δ(b) & ≜ \{\} \\
    δ(q) & ≜ \{1 ↦ (1, p)\}\\  
    δ(e +_b f) & ≜ \{b ∧ a ↦ (e', p) ∣ a ↦ (e', p) ∈ δ(e)\} 
        ∪ \{b̄ ∧ a ↦ (f', p) ∣ a ↦ (f', p) ∈ δ(f)\}\\
    δ(e ⋅ f) & ≜ \{b ↦ (e' ⋅ f, p) ∣ b ↦ (e, p) ∈ δ(e)\}
        ∪ \{ϵ(e) ∧ b ↦ (f', p) ∣ b ↦ (f', p) ∈ δ(f)\}\\
    δ(e^{(b)}) & ≜ \{b ∧ a ↦ (e' ⋅ e^{(b)}, p) ∣ a ↦ (e', p) ∈ δ(e)\}.
\end{align*}
FIXME: we still need to think about what representation do we want to go with, with this representation, you can also have duplicated K. We need a good example here of why don't we go with map. One good argument is that this representation is complex.

\section{On the Fly Normalization}

Normalization of a GKAT automaton will remove all the dead state, which are states that cannot lead to an accepting transitions. 
However, this process is hard to do on-the-fly, mainly because of the difficulty of handling infinite loops.
\begin{example}
    We can consider a simple extension to a BFS algorithm, where we will mark a state as either live, dead, or searching. 
    A state is marked as ``searching'' when it has been explored but not yet known to be dead or alive yet,
    and when a state loops back to a searching state, then we can declare the detection of a infinite loop, hence marking the transition dead.
    We consider the following coalgebra over a two element set \(\{1, 2\}\):
    \[
    \begin{tikzcd}[row sep=15px]
        1 \ar[bend right, swap]{r}{b ∣ p} \ar[Rightarrow, swap]{d}{b̄}& 2 \ar[bend right, swap]{l}{b ∣ p}\\  
        \ &  
    \end{tikzcd}
    \]
    It is clear that both states \(1\) and \(2\) are live,
    since \(1\) is accepting and \(2\) have a transition to \(1\).
    However, if we run our search algorithm: to determine whether \(1\) is live, we will need to compute the backtrack of both of its transitions. 
    We will compute the backtrack from \(b\) first, which computes the liveness of \(2\). Thus we will mark \(1\) as searching. Yet state \(2\) only have one transition to \(1\), because \(1\) is marked as searching, the algorithm will think a infinite loop is discovered and mark \(2\) as dead, which doesn't align with the reality.
\end{example}

Thus, it is crucial to resolve these loops when we a designing a on-the-fly algorithm.
In fact, we can utilize a common graph theory technique called \emph{graph condensation} to remove the loops, where it treat all thee strong connected component as a vertex as a new graph, and the edges between the components are preserved. 
The resulting graph is a directed acyclic graph (DAG), then we can utilize a simple search to find the live state.

\begin{definition}[Graph For sGKCT]
    Given a sGKCT \(𝒮 ≜ ⟨S, δ, ϵ⟩\), we can construct a (potentially infinite) graph, where the vertices are \(S\) and the edge \((s, s')\) is in the graph when  connected there exists a \(p ∈ K\), s.t. \(δ(s, s', p) ≠ 0\)
\end{definition}

We will also recall the definition of strongly connected components 
\begin{definition}[Walk and strongly connected components]
    A walk from vertex \(s₁\) to vertex \(s₂\) in a graph is a sequence of vertices \([s₁, s₂, …, sₙ]\), s.t. for all \(i\), \((sᵢ, s_{i + 1})\) are edges in the graph.
    A strongly connected component is a equivalence class on the vertices, where two vertices \(s₁\) and \(s₂\) are equivalent when there is a walk from \(s₁\) to \(s₂\) and there is a walk from \(s₂\) to \(s₁\).
\end{definition}

And in fact strongly connected components has strong connections with liveness of states.

\begin{lemma}
    By induction on the length of the walk, if there is a walk from state \(s₁\) and \(s₂\) then if \(s₂\) is live and \(s₁\) is live. 
    Thus by definition of strongly connected component, a strongly connected component is live if and only if any of the state in the component is live. 
    And consequently, all the states in a strongly connected component are either all dead or all live.
\end{lemma}

To identify the strongly connected components, we can run the Tarjan's algorithm along side the normal Hopcroft and Karp's bisimulation algorithm.
The Tarjan's algorithm is a one pass DFS based algorithm where it assigns each state a \emph{low-link value}, at the end of the algorithm, all the algorithm with the same low-link value will be in the same strongly connected components. 
Thus instead of marking each state with dead or live, we will mark each state with a low-link value, and each low-link value is marked with either dead or alive.
This kind of mappings are enabled by the fact that all states in a strongly connected components are either all dead or all live.

\section{The Algorithm}

FIXME: proof a theorem that if the transformation shares boolean with the original program, then the decision procedure is nearly linear time.

% \section{Definitions}

% A deterministic symbolic Kleene Coalgebra with Tests (dsKCT) \(𝒮\) over a finite alphabet \(K\) and a boolean algebra \(ℬ\) is a coalgebra with with the following operation:
% \[δ: 𝒮 → 𝒮 → ℬ.\]
% Following notations in previous works,
% we will sometimes write \(δ_{p}: 𝒮 → 𝒮 → \BExp(B)\) for \(p ∈ K\).

% The boolean term algebra over \(B\) formed by all the boolean terms over \(B\) defined by 
% \[e ≜ b ∈ B ∣ e₁ ∨ e₂ ∣ e₁ ∧ e₂ ∣ ē,\]
% modulo boolean algebra equalities. 
% The operations in the boolean term algebra is syntactical,
% for example, the logical conjunction of term \(e₁\) and \(e₂\)
% in the term algebra is the term \(e₁ ∧ e₂\).
% The boolean term algebra is a free boolean algebra, 
% this means that all the set map \(Î: B → ℬ\) for any boolean algebra \(ℬ\)
% can be lifted to a unique boolean algebra homomorphism \(I: \BExp(B) → ℬ\),
% where \[∀ b ∈ B, I(b) = Î(b).\]

% A sKCT homomorphism is a map \(h: 𝒮 → 𝒮'\) where it preserves both operations:
% \[
%     \begin{tikzcd}
%         𝒮 \ar{d}{ϵ} \ar{r}{h} & 𝒮' \ar{dl}{ϵ} \\  
%         \BExp(B)
%     \end{tikzcd} \qquad \qquad
%     ∀ p ∈ K, 
%     \begin{tikzcd}[column sep=1.5cm]
%         𝒮 \ar{d}{δ_{p}} \ar{r}{h} & 𝒮' \ar{d}{δ_{p}} \\  
%         \BExp(B)^𝒮 & \BExp(B)^{𝒮'} \ar{l}{\BExp(B)^{h}}
%     \end{tikzcd}    
% \]
% where \(\BExp(B)^{h}\) is defined as 
% \begin{align*}
%     \BExp(B)^{h} & : (𝒮' → \BExp(B)) → (𝒮 → \BExp(B)) \\
%     \BExp(B)^{h} & (f) ≜ f ∘ h.
% \end{align*}
% Notice the function \(\BExp(B)^{h}\) has the type \(\BExp(B)^{𝒮'} → \BExp(B)^𝒮\)
% instead of the conventional \(\BExp(B)^{𝒮} → \BExp(B)^{𝒮'}\),
% this is because \(\BExp(B)^{(-)}\) is a contravariant functor.

% Similarly, sKCT bisimulation \(≡\) between \(𝒮\) and \(𝒮'\) is a 
% relation \({≡} ⊆ 𝒮 × 𝒮'\) that makes the following diagrams commutes:
% \[
%     \begin{tikzcd}
%         𝒮 \ar[swap]{dr}{ϵ} & 
%             ≡ \ar[swap]{l}{π₁} \ar{r}{π₂}& 
%             𝒮' \ar{dl}{ϵ} \\  
%         & \BExp(B) & 
%     \end{tikzcd} \qquad \qquad
%     ∀ p ∈ K, 
%     \begin{tikzcd}[column sep=1.5cm]
%         𝒮 \ar{d}{δ_{p}} 
%             & ≡ \ar[swap]{l}{π₁} \ar{r}{π₂}
%             & 𝒮' \ar{d}{δ_{p}} \\  
%         \BExp(B)^𝒮 \ar{r}{\BExp(B)^{(π₁)}}
%         & \BExp(B)^{(≡)} & \BExp(B)^{𝒮'} \ar[swap]{l}{\BExp(B)^{(π₂)}}
%     \end{tikzcd}    
% \]

% \begin{example}
%     Consider the following two sKCT
%     \[
%     \begin{tikzpicture}
%         \node (0) [state] {0};
%         \node (1) [state, above right = of 0] {1};
%         \node (2) [state, below right = of 0] {2};
%         \path[->] (0) edge node[above]{\(b p\)} (1);
%         \path[->] (0) edge node[above]{\(b̄ p\)} (2);   
%         \draw[->, double] (0) -- +(0,1) node[left]{\(1\)};
%         \draw[->, double] (1) -- +(1,0) node[above]{\(1\)};
%         \draw[->, double] (2) -- +(1,0) node[above]{\(1\)};
%     \end{tikzpicture}
%     \qquad \qquad
%     \begin{tikzpicture}
%         \node (0) [state] {0};
%         \node (1) [state, above right = of 0] {1};
%         \node (2) [state, below right = of 0] {2};
%         \path[->] (0) edge node[above]{\(c p\)} (1);
%         \path[->] (0) edge node[above]{\(c̄ p\)} (2);   
%         \draw[->, double] (0) -- +(0,1) node[left]{\(1\)};
%         \draw[->, double] (1) -- +(1,0) node[above]{\(1\)};
%         \draw[->, double] (2) -- +(1,0) node[above]{\(1\)};
%     \end{tikzpicture}
%     \]
%     They are bisimilar
% \end{example}

% \begin{proof}
%     It is clean that both of these sKCT unfolds to the same KCT,
%     and they both accepts languages \(\{α p β ∣ α, β ∈ \At\}\). 
%     However, their transition has some overlaps, and some distinction.

%     we consider the following bisimulation \(\{(0,0)\} ∪ \{(a, b) ∣ a, b ∈ \{1,2\}\}\)
% \end{proof}

% \section{Semantics And Relation with KCT}

% There are several representations of free boolean algebra 
% and all of them are isomorphic. 
% For example a well-known representation of free boolean algebra is \(2^\At\),
% Notice 
% \begin{align}
%     ⋁ \{α ∣ α ≤ ϕ\} & = ϕ & \text{for } ϕ ∈ \BExp(B), 
%         \label{the: lor of atom}\\  
%     \{α ∣ α ≤ ⋁ A\} & = A & \text{for } A ∈ 2^{\At}.
%         \label{the: atom of lor}
% \end{align}
% Then the isomorphism \(f: \BExp(B) ⥲ 2^\At\) is defined by 
% the following function \(at(ϕ) ≜ \{α ∣ α ≤ ϕ\}\) with the inverse \(at^{-1}(A) ≜ ⋁ A\).

% The classical way to provide semantics to a certain coalgebra is by finding the final coalgebra.
% \begin{definition}[Final sKCT]
%     The final sKCT is formed by set of guarded strings \(2^{\GS}\), 
%     with the operations defined as follows:
%     \begin{align*}
%         ϵ(l) & ≜ ⋁ \{α ∣ α ∈ l\} & 
%         δₚ(l)(l') & ≜ ⋁ \{α ∣ ∃ w ∈ l', α p w ∈ l\}
%     \end{align*}
% \end{definition}

% \begin{proof}
%     We need to show that for all sKCT \(𝒮\), there exists a unique homomorphism \(L\):
%     \[
%         \begin{tikzcd}
%             𝒮 \ar{d}{ϵ} \ar[dashed]{r}{L} & 2^{\GS} \ar{dl}{ϵ} \\  
%             \BExp(B)
%         \end{tikzcd} \qquad \qquad
%         ∀ p ∈ K, 
%         \begin{tikzcd}[column sep=1.5cm]
%             𝒮 \ar{d}{δ_{p}} \ar[dashed]{r}{L} & 2^{GS} \ar{d}{δ_{p}} \\  
%             \BExp(B)^𝒮 & \BExp(B)^{2^{GS}} \ar[dashed]{l}{\BExp(B)^{L}}
%         \end{tikzcd}
%     \]
%     We can simplify these two diagrams as follows
%     \begin{itemize}
%         \item If the first diagram commutes, then for all \(s ∈ 𝒮\), 
%             \(ϵ(s) = ϵ(L(s)) = ⋁ \{α ∣ α ∈ L(s)\}\), which implies
%             \[α ≤ ϵ(s) ⟺ α ≤ ⋁ \{α ∣ α ∈ L(s)\} ⟺ α ∈ L(s).\]
%             Furthermore, if \(α ≤ ϵ(s) ⟺ α ∈ L(s)\), we have 
%             \[ϵ(L(s)) = ⋁ \{α ∣ α ∈ L(s)\} = ⋁ \{α ∣ α ≤ ϵ(s)\} = ϵ(s);\]
%             thus the first diagram commutes.
%             Therefore, the first diagram commutes if and only if the following equivalence holds:
%             \[α ≤ ϵ(s) ⟺ α ∈ L(s).\]
%         \item If the second diagram commutes, then for all \(s, s' ∈ 𝒮\), 
%             \[δₚ(s)(s') = \BExp(B)^{L}(δₚ(L(s)))(s') 
%                 = δₚ(L(s))(L(s')) 
%                 = ⋁ \{α ∣ ∃ w ∈ L(s'), α p w ∈ L(s)\},\]
%             which would imply 
%             \[α p w ∈ L(s) ⟺ ∃ s' ∈ 𝒮, w ∈ L(s') ∧ α ≤ δₚ(s)(s').\]
%     \end{itemize}

%     where the language \(L(s)\) for each state \(s\) is coinductively defined as follows:
%     \[L(s) ≜ \{α ∣ α ≤ ϵ(s)\} ∪ \{α p w ∣ p ∈ K, s' ∈ 𝒮, α ≤ δₚ(s)(s'), w ∈ L(s')\}.\]
%     Notice the following equivalence can be observed from the definition
%     \begin{align*}
%         α ∈ L(s) & ⟺ α ≤ ϵ(s), \\
%         α p w ∈ L(s) & ⟺ ∃ s' ∈ 𝒮, w ∈ L(s') ∧ α ≤ δₚ(s)(s').
%     \end{align*}

%     We first show that the diagram commutes:
%     \begin{align*}
%         ϵ(L(s)) & = ⋁ \{α ∣ α ∈ L(s)\} \\
%             & = ⋁ \{α ∣ α ≤ ϵ(s)\} 
%             = ϵ(s);\\   
%         \BExp(B)^{L}(δₚ(L(s)))(s')
%             & = δₚ(L(s))(L(s')) \\
%             & = ⋁ \{α ∣ ∃ w ∈ L(s'), α p w ∈ L(s)\} \\  
%             & = ⋁ \{α ∣ α ≤ δₚ(s)(s')\} 
%             = δₚ(s)(s').
%     \end{align*}

%     We then show that such a homomorphism is unique,
%     assume that \(L': 𝒮 → 2^{\GS}\), 
%     we will show that given any guarded word \(w ∈ 2^{\GS}\), and state \(s ∈ 𝒮\)
%     \[w ∈ L'(s) ⟺ w ∈ L(s),\]
%     by induction on \(w\):
%     \begin{itemize}
%         \item If \(w = α\) for some atom \(α\),
%             then \(w ∈ L'(s) \)
%     \end{itemize}
% \end{proof}

% Since \(\BExp(B) ≅ 2^\At\) in set, 
% we can develop the following natural isomorphism:
% \begin{align*}
%     ϵ & : 𝒮 → \BExp(B) ≅ 𝒮 → \At → 2; \\  
%     δ & : 𝒮 → K → 𝒮 → \BExp(B) ≅ 𝒮 → K → 𝒮 → \At → 2 
%         ≅ 𝒮 → \At → K → 2^{𝒮}.
% \end{align*} 
% In other word, a symbolic Kleene Coalgebra with Tests is a ``non-deterministic'' KCT,
% in the sense that given a atom and a primitive action, it can transition 
% to multiple states non-deterministically.
% There we can apply the following determination procedure to it:
% \begin{align*}
%     ϵ̂ &:  2^𝒮 → \At → 2 & δ̂ &: 2^𝒮 → \At → K → 2^{𝒮} \\  
%     ϵ̂_{α} & (S) ≜ ∃ s ∈ S, ϵ(s) = 1, & 
%     δ̂_{α p} & (S) ≜ \{s' ∣ δ_{α p}(s) = s', s ∈ S\}.
% \end{align*}
% In other words, given a sKCT \(𝒮\), then we can construct a KCT \(2^{𝒮}\).

\end{document}

### 2024-05-10 notes.tex
FInally, almost, finished normalized bisimulation

@@
 \usepackage{listings}
+\lstset{basicstyle=\footnotesize\ttfamily,breaklines=true}
+\lstset{framextopmargin=50pt,frame=bottomline}
+\lstset{language=caml, escapeinside={[*}{*]}}
 
@@
 \DeclareMathOperator{\post}{\mathrm{post}}
+\DeclareMathOperator{\Img}{\mathrm{\mathbf{Img}}}
 \newcommand{\reject}{\mathinner{\mathtt{Reject}}}
@@
 \DeclareMathOperator*{\bigplus}{\scalerel*{+}{\sum}}
+\newcommand{\closSim}{\mathrel{\overline{\sim}}}
+\DeclareMathOperator{\norm}{\mathrm{norm}}
 
@@
 \newcommand{\comBreak}{\command{break}}
-
-% texts
-\newcommand{\tand}{\text{ and }}
-\newcommand{\totherwise}{\text{otherwise}}
-\newcommand{\tif}{\text{if }}
-\newcommand{\tor}{\text{ or }}
-\newcommand{\twhere}{\text{ where }}
-\newcommand{\tst}{\text{ s.t.\ }}
-
-% typeset
-\newcommand{\itemTitle}[1]{\textbf{#1}}
-\newcommand{\keyword}[1]{\emph{#1}}
 
@@
     \author{
-        Cheng Zhang \\
-        % \and
-        % Arthur Azevedo de Amorim \\
-        % \and
-        % Marco Gaboardi \\
     }
@@
 
-Given a finite set \(K\), a \emph{symbolic Guarded Kleene Coalgebra with Tests} (sGKCT) \(𝒮 ≜ ⟨S, ϵ, δ⟩\) over a boolean algebra \(ℬ\) consists of a state set \(S\) consists of a accepting boolean \(ϵ\) and a transition function \(δ\):
+Given a finite set \(K\), a \emph{symbolic Deterministic Kleene Coalgebra with Tests} (sDKCT) \(𝒮 ≜ ⟨S, ϵ, δ⟩\) over a boolean algebra \(ℬ\) consists of a state set \(S\) consists of a accepting boolean \(ϵ\) and a transition function \(δ\):
 \begin{mathpar}
@@
 In particular, when the boolean algebra is the free boolean algebra \(\BExp_B\) generated by a finite set \(B\), the labels can be considered as atoms, since \(\BExp_B ≅ 2^{\At_B}\). 
-Then we can explain the intuition for the definition of GKCT
+Then we can explain the intuition for the definition of DKCT
 \begin{itemize}
@@
 
-Given a sGKCT \(𝒮 ≜ ⟨S, ϵ, δ⟩\) and a state \(s ∈ S\), we use the notation \(s ⇒ b\) to denote that \(ϵ(s) ≥ b\), and we use the notation \(s \transvia{b ∣ p} s'\) to denote that \(δ(s)₃(s', p) ≥ b\),
-where the ordering is the typical ordering in boolean algebra.
+Given a sDKCT \(𝒮 ≜ ⟨S, ϵ, δ⟩\) and a state \(s ∈ S\), we use the notation \(s ⇒ b\) to denote that \(ϵ(s) ≥ b\), and we use the notation \(s \transvia{b ∣ p} s'\) to denote that \(δ(s)₃(s', p) ≥ b\), where the ordering is the typical ordering in boolean algebra.
 
-The symbolic derivatives for GKAT forms a GKCT where the states are represented by GKAT expression \(\GKAT_{K, B}\), and the boolean algebra is the free boolean algebra \(\BExp_B\) over the test alphabet \(B\):
+The symbolic derivatives for GKAT forms a DKCT where the states are represented by GKAT expression \(\GKAT_{K, B}\), and the boolean algebra is the free boolean algebra \(\BExp_B\) over the test alphabet \(B\):
 \begin{mathpar}
@@
 
-\section{On the Fly Normalization}
+% \section{On the Fly Normalization}
 
-Normalization of a GKAT automaton will remove all the dead state, which are states that cannot lead to an accepting transitions. 
-However, this process is hard to do on-the-fly, mainly because of the difficulty of handling infinite loops.
-\begin{example}
-    We can consider a simple extension to a BFS algorithm, where we will mark a state as either live, dead, or searching. 
-    A state is marked as ``searching'' when it has been explored but not yet known to be dead or alive yet,
-    and when a state loops back to a searching state, then we can declare the detection of a infinite loop, hence marking the transition dead.
-    We consider the following coalgebra over a two element set \(\{1, 2\}\):
+% Normalization of a GKAT automaton will remove all the dead state, which are states that cannot lead to an accepting transitions. 
+% However, this process is hard to do on-the-fly, mainly because of the difficulty of handling infinite loops.
+% \begin{example}
+%     We can consider a simple extension to a BFS algorithm, where we will mark a state as either live, dead, or searching. 
+%     A state is marked as ``searching'' when it has been explored but not yet known to be dead or alive yet,
+%     and when a state loops back to a searching state, then we can declare the detection of a infinite loop, hence marking the transition dead.
+%     We consider the following coalgebra over a two element set \(\{1, 2\}\):
+%     \[
+%     \begin{tikzcd}[row sep=15px]
+%         1 \ar[bend right, swap]{r}{b ∣ p} \ar[Rightarrow, swap]{d}{b̄}& 2 \ar[bend right, swap]{l}{b ∣ p}\\  
+%         \ &  
+%     \end{tikzcd}
+%     \]
+%     It is clear that both states \(1\) and \(2\) are live,
+%     since \(1\) is accepting and \(2\) have a transition to \(1\).
+%     However, if we run our search algorithm: to determine whether \(1\) is live, we will need to compute the backtrack of both of its transitions. 
+%     We will compute the backtrack from \(b\) first, which computes the liveness of \(2\).
+Thus we will mark \(1\) as searching.
+Yet state \(2\) only have one transition to \(1\), because \(1\) is marked as searching, the algorithm will think a infinite loop is discovered and mark \(2\) as dead, which doesn't align with the reality.
+% \end{example}
+
+% Thus, it is crucial to resolve these loops when we a designing a on-the-fly algorithm.
+% In fact, we can utilize a common graph theory technique called \emph{graph condensation} to remove the loops, where it treat all thee strong connected component as a vertex as a new graph, and the edges between the components are preserved.
+% The resulting graph is a directed acyclic graph (DAG), then we can utilize a simple search to find the live state.
+
+% \begin{definition}[Graph For sDKCT]
+%     Given a sDKCT \(𝒮 ≜ ⟨S, δ, ϵ⟩\), we can construct a (potentially infinite) graph, where the vertices are \(S\) and the edge \((s, s')\) is in the graph when  connected there exists a \(p ∈ K\), s.t.
+\(δ(s, s', p) ≠ 0\)
+% \end{definition}
+
+% We will also recall the definition of strongly connected components 
+% \begin{definition}[Walk and strongly connected components]
+%     A walk from vertex \(s₁\) to vertex \(s₂\) in a graph is a sequence of vertices \([s₁, s₂, …, sₙ]\), s.t. for all \(i\), \((sᵢ, s_{i + 1})\) are edges in the graph.
+%     A strongly connected component is a equivalence class on the vertices, where two vertices \(s₁\) and \(s₂\) are equivalent when there is a walk from \(s₁\) to \(s₂\) and there is a walk from \(s₂\) to \(s₁\).
+% \end{definition}
+
+% And in fact strongly connected components has strong connections with liveness of states.
+
+% \begin{lemma}
+%     By induction on the length of the walk, if there is a walk from state \(s₁\) and \(s₂\) then if \(s₂\) is live and \(s₁\) is live. 
+%     Thus by definition of strongly connected component, a strongly connected component is live if and only if any of the state in the component is live. 
+%     And consequently, all the states in a strongly connected component are either all dead or all live.
+% \end{lemma}
+
+% To identify the strongly connected components, we can run the Tarjan's algorithm along side the normal Hopcroft and Karp's bisimulation algorithm.
+% The Tarjan's algorithm is a one pass DFS based algorithm where it assigns each state a \emph{low-link value}, at the end of the algorithm, all the algorithm with the same low-link value will be in the same strongly connected components.
+% Thus instead of marking each state with dead or live, we will mark each state with a low-link value, and each low-link value is marked with either dead or alive.
+% This kind of mappings are enabled by the fact that all states in a strongly connected components are either all dead or all live.
+
+\section{Normalized Bisimulation}
+
+Bisimilarity implies language sematnical equivalence \[s ∼ t ⟹ ⟦s⟧ = ⟦t⟧\], however the converse doesn't hold unless we are working in a \emph{normal} GKAT coalgebra.
+It is well known that two states in a normal GKAT coalgebra if and only if they have the same language semantics \[s ∼ t ⟺ ⟦s⟧ = ⟦t⟧.\]
+However, instead of normalizing automatons, we can in fact, check for normality on the fly, this is important to design a derivatives based decision algorithm.
+
+We will use a notion of normalized bisimulation, where we will enlarge the bisimulation by dead states, since all the dead states are trace equivalent.
+Operationally, when we find a discrepancy during the bisimulation algorithm we will check whether the states causing the discrepancy are dead.
+We can prove the correctness of this algorithm by proving that they are equivalent to normalize and then computing bisimulation.
+Let \(G\) be the functor defining GKAT coalgebra: \[G ≜ (2 + (-) × K)^{\At}.\]
+Given two DKCT \(𝒮, 𝒯\), and a relation on these two expression \({∼} ⊆ S × T\), we can define \({\closSim} ⊆ (2 + S × K) × (2 + T × K)\) as the smallest relation generated by the following rules: for all \(p, q ∈ K\) and states \(s' ∈ S\), \(t' ∈ T\):
+\begin{mathpar}
+    \inferrule[Bisim-0]{\\}{0 \closSim 0}
+    \and 
+    \inferrule[Bisim-1]{\\}{1 \closSim 1}
+    \and 
+    \inferrule[Bisim-Trans]{s' ∼ t'}{(s', p) \closSim (t', p)}
+    \and
+    \inferrule[Norm-Left]{s' \text{ is dead}}{(s', p) \closSim 0}
+    \and 
+    \inferrule[Norm-Right]{t' \text{ is dead}}{0 \closSim (t', p)}
+    \and
+    \inferrule[Norm-Both]{s', t' \text{ are both dead}}{(s', p) \closSim (t', q)}
+\end{mathpar}
+Intuitively, \({\closSim}\) enlarges the relation \(∼\) for \emph{results} of a transition. for example, it allows transition to dead states to equivalent to rejection, and so on.
+This extension enables ``on-the-fly normalization'', where to determine whether \(s \closSim t\), we can simply first check for bisimulation using the bisimulation rules; and if all failed, we can check the normalization rules.
+
+We call the bisimulation algorithm with on-the fly normalization ``\emph{normalized bisimulation}''.
+This concept can be summed up nicely using diagrams.
+For a relation \(∼\), then there is a straight forward embedding \((2 + {∼} × K) ↪ {\closSim}\):
+This embedding generates all the elements that is formed by the bisimulation rules of \(\closSim\),
+also note that \(∼\) is not restricted to live states, but dead states as well; this means that if two dead states follows the bisimulation rules, then we don't need to fall back to the normalization rules.
+Since detecting whether a state is dead requires iterating through all its predecessors, we can use this property to forgo dead state detection until all the bisimulation rules failed.
+
+The aforementioned embedding \((2 + {∼} × K) ↪ {\closSim}\) can be point-wise lifted to \(G(∼) ↪ (\At → {\closSim})\), thus our normalized algorithm can be seen as finding a normalized bisimulation \(∼\) that satisfy the following diagram:
 \[
-    \begin{tikzcd}[row sep=15px]
-        1 \ar[bend right, swap]{r}{b ∣ p} \ar[Rightarrow, swap]{d}{b̄}& 2 \ar[bend right, swap]{l}{b ∣ p}\\  
-        \ &  
+    \begin{tikzcd}[column sep=1.5cm, row sep=1cm]
+        S \ar[swap]{d}{δ_𝒮}
+            & ∼  \ar[dashed, swap]{d}{∃} \ar[swap]{l}{π₁} \ar{r}{π₂}
+            & T \ar{d}{δ_𝒯} \\  
+        (2 + S × K)^\At 
+            & ({\closSim}^\At) 
+                \ar[swap]{l}{{π₁}^\At} \ar{r}{{π₂}^\At}
+            & (2 + T × K)^\At
     \end{tikzcd}
 \]
-    It is clear that both states \(1\) and \(2\) are live,
-    since \(1\) is accepting and \(2\) have a transition to \(1\).
-    However, if we run our search algorithm: to determine whether \(1\) is live, we will need to compute the backtrack of both of its transitions. 
-    We will compute the backtrack from \(b\) first, which computes the liveness of \(2\).
-Thus we will mark \(1\) as searching.
-Yet state \(2\) only have one transition to \(1\), because \(1\) is marked as searching, the algorithm will think a infinite loop is discovered and mark \(2\) as dead, which doesn't align with the reality.
+Notice, unlike \(G(∼)\), the map \({∼} ↦ (\closSim)^\At\) is not a functor,
+since it is only defined on relations.
+
+\begin{example}
+    Let's consider a example with a single primitive test \(B ≜ \{b\}\),
+    thus atoms are defined as \(\At_B = \{\{b\}, ∅\}\).
+    We focus on two states in the bisimulation \(s\) and \(t\).
+    with their respective transition defined as follows 
+    \begin{align*}
+            δ_𝒮(s, \{b\}) & ≜ 0   
+                & δ_𝒮(t, \{b\}) & ≜ (t', p) \\  
+            δ_𝒮(s, ∅) & ≜ (s', p) 
+                & δ_𝒮(t, ∅) & ≜ (t', p) 
+    \end{align*}
+    It is quite clear \(s\) and \(t\) will not be bisimular, however, there can be a normalized bisimulation relation between them, provided that \(s' ∼ t'\) and \(t'\) is dead: 
+    \[
+    \begin{tikzcd}[column sep=1.5cm, row sep=1cm]
+        s \ar[swap, mapsto]{d}{δ_𝒮}
+            & (s, t)  \ar[dashed, swap, mapsto]{d}{∃} \ar[swap, mapsto]{l}{π₁} \ar[mapsto]{r}{π₂}
+            & t \ar[mapsto]{d}{δ_𝒯} \\  
+        \begin{gathered}
+            \{b\}↦ 0 \\  
+            ∅ ↦ (s', p)
+        \end{gathered} 
+            & \begin{gathered}
+                \{b\}↦ 0, (t', p) \\  
+                ∅ ↦ (s', p), (t', p)
+            \end{gathered} 
+                \ar[swap, mapsto]{l}{{π₁}^\At} \ar[mapsto]{r}{{π₂}^\At}
+            & \begin{gathered}
+                \{b\} ↦ (t', p) \\  
+                ∅ ↦(t', p)
+            \end{gathered}
+    \end{tikzcd}
+    \]
+    Notice that \(0 \closSim (t', p)\) because \(t'\) is dead, 
+    and \((s', p) \closSim (t', p)\) because we assume that \(s' ∼ t'\). 
+    Furthermore, if both \(s'\) and \(t'\) are dead, we can establish \(s' \closSim t'\) using either the \textnormal{\textsc{Bisim-Trans}} rule or \textnormal{\textsc{Norm-Both}} rule, and we are free to pick the more efficient check in our implementation.
 \end{example}
 
-Thus, it is crucial to resolve these loops when we a designing a on-the-fly algorithm.
-In fact, we can utilize a common graph theory technique called \emph{graph condensation} to remove the loops, where it treat all thee strong connected component as a vertex as a new graph, and the edges between the components are preserved.
-The resulting graph is a directed acyclic graph (DAG), then we can utilize a simple search to find the live state.
+\begin{theorem}[Universiality]
+    Given a normalized bisimulation relation \({∼} ⊆ S × T\)
+    \[
+        \begin{tikzcd}[column sep=1.5cm, row sep=1cm]
+            S \ar[swap]{d}{δ_𝒮}
+                & ∼  \ar[dashed, swap]{d}{h} \ar[swap]{l}{π₁} \ar{r}{π₂}
+                & T \ar{d}{δ_𝒯} \\  
+            (2 + S × K)^\At 
+                & (\closSim)^\At 
+                    \ar[swap]{l}{{π₁}^\At} \ar{r}{{π₂}^\At}
+                & (2 + T × K)^\At
+        \end{tikzcd}
+    \] 
+    Then \(h\) is unique, and defined as follows: 
+    \begin{align*}
+        h & : {∼} → {\closSim}^\At \\
+        h & (s, t, α) ≜ (δ_𝒮(s, α), δ_𝒯(s, α))
+    \end{align*}
+\end{theorem}
 
-\begin{definition}[Graph For sGKCT]
-    Given a sGKCT \(𝒮 ≜ ⟨S, δ, ϵ⟩\), we can construct a (potentially infinite) graph, where the vertices are \(S\) and the edge \((s, s')\) is in the graph when  connected there exists a \(p ∈ K\), s.t.
-\(δ(s, s', p) ≠ 0\)
+\begin{proof}
+    The commutativity can be verified by computation. We only need to show uniqueness.
+
+    The above commutative diagram implies the commutativity of the following diagram:
+    \[
+        \begin{tikzcd}
+            & {∼} \ar{d}{h} \ar[swap]{ddl}{δ_𝒮 ∘ π₁} \ar{ddr}{δ_𝒯 ∘ π₂}& \\
+            & (\closSim)^\At \ar[hook]{d}{i} & \\  
+            (2 + S × K)^\At & 
+                (2 + S × K)^\At × (2 + T × K)^\At
+                    \ar{l}{π₁^\At} \ar[swap]{r}{π₂^\At}
+                & (2 + T × K)^\At
+        \end{tikzcd}
+    \] 
+    Since \((-)^{\At}\) is a right adjoint and right adjoints preserve limits, \((2 + S × K)^\At × (2 + T × K)^\At\) is the product of \((2 + T × K)^\At\) and \((2 + T × K)^\At\). 
+    Hence, \(h ∘ i\) is the unique function to make the above diagram commute. 
+    Assume we can replace \(h\) with \(g\) while preserving the commutativity of the diagram: 
+    \begin{align*}
+        & i ∘ h  = i ∘ g & \text{\((2 + S × K)^\At × (2 + T × K)^\At\) is a product.} \\  
+        ⟹ {}& h = g & \text{\(i\) is a embedding, hence left cancelable.}
+    \end{align*}
+\end{proof}
+
+\begin{definition}
+    Given a DKCT \(𝒮 = ⟨S, δ⟩\), a \emph{sub-DKCT} \(𝒮' = ⟨S', δ|_{S'}⟩\) is a coagebra induced by \(S' ⊆ S\), a sub-DKCT can be characterized as a pullback square:
+    \[
+        \begin{tikzcd}
+            S' \ar[hook]{r}{i} \ar{d}{δ|_{S'}} & S \ar{d}{δ} \\  
+            G(S') \ar[hook]{r}{i} & G(S)
+        \end{tikzcd}
+    \]
+    By the commutativity of this diagram \(i\) is also a DKCT homomorphism.
+    A \emph{principle sub-DKCT} of state \(s ∈ S\), denoted \(⟨s⟩\) is the smallest sub-DKCT that contains \(s\). 
+    We can show that the principle DKCT of \(s\) exactly characterize the DKCT induced by the reachable state of \(s\).
 \end{definition}
 
-We will also recall the definition of strongly connected components 
-\begin{definition}[Walk and strongly connected components]
-    A walk from vertex \(s₁\) to vertex \(s₂\) in a graph is a sequence of vertices \([s₁, s₂, …, sₙ]\), s.t. for all \(i\), \((sᵢ, s_{i + 1})\) are edges in the graph.
-    A strongly connected component is a equivalence class on the vertices, where two vertices \(s₁\) and \(s₂\) are equivalent when there is a walk from \(s₁\) to \(s₂\) and there is a walk from \(s₂\) to \(s₁\).
-\end{definition}
+%%%% NOTE: The universiality is not true, since 𝒯 will need to have the same transition structure as ⟨s⟩. 
+%%%%
+% \begin{definition}
+%     a \emph{principle sub-DKCT} of \(𝒮\) is a ``free'' sub-DKCT generated by one element \(s\), namely, given any sub-DKCT \(𝒯\) and a element \(t ∈ 𝒯\), there is a unique DKCT homomorphism \(h: ⟨s⟩ → 𝒯\), s.t.
+\(h(s) = t\).
 
-And in fact strongly connected components has strong connections with liveness of states.
+%     We denote the principle sub-DKCT of \(⟨s⟩_{𝒮}\), and we will omit the subscript \(𝒮\) if it is clear from context.
+% \end{definition}
+
+% \begin{lemma}
+%     The principle sub-DKCT \(⟨s⟩_𝒮\) always exists, and it is the smallest sub-DKCT that contain \(s\).
+% \end{lemma}
+
+% \begin{proof}
+%     \(⟨s⟩_𝒮\) can be constructed as the set of all the reachable state of \(s\) in \(𝒮\), specifically it is the smallest set that is generated by the following rules:
+%     \begin{mathpar}
+%         \inferrule{\\}{s ∈ ⟨s⟩_𝒮} \and
+%         \inferrule{s₁ ∈ ⟨s⟩_𝒮 \\ ∃ α, δ(s₁, α) = (s₂, p)}{s₂ ∈ ⟨s⟩_𝒮}
+%     \end{mathpar}
+%     then the homomorphism \(h: ⟨s⟩ → 𝒯\) can be inductively defined as follows:
+%     \begin{align*}
+%         h(s) & ≜ t \\ 
+%         h(s₂) & ≜ h(δ(s₁)) 
+%     \end{align*}
+% \end{proof}
 
 \begin{lemma}
-    By induction on the length of the walk, if there is a walk from state \(s₁\) and \(s₂\) then if \(s₂\) is live and \(s₁\) is live. 
-    Thus by definition of strongly connected component, a strongly connected component is live if and only if any of the state in the component is live. 
-    And consequently, all the states in a strongly connected component are either all dead or all live.
+    A state \(s\) is live if and only if there exists a accepting state in \(⟨s⟩\);  
+    and a state \(s\) is dead if and only if there is no accepting state in \(⟨s⟩\).
 \end{lemma}
 
-To identify the strongly connected components, we can run the Tarjan's algorithm along side the normal Hopcroft and Karp's bisimulation algorithm.
-The Tarjan's algorithm is a one pass DFS based algorithm where it assigns each state a \emph{low-link value}, at the end of the algorithm, all the algorithm with the same low-link value will be in the same strongly connected components.
-Thus instead of marking each state with dead or live, we will mark each state with a low-link value, and each low-link value is marked with either dead or alive.
-This kind of mappings are enabled by the fact that all states in a strongly connected components are either all dead or all live.
+\begin{corollary}
+    \(s\) is dead if and only if \(⟨s⟩\) is all dead.
+\end{corollary}
 
-\section{The Algorithm}
+\begin{proof}
+    Take any state \(s' ∈ ⟨s⟩\), we can construct the DKCT \(⟨s'⟩\).
+    Since \(⟨s'⟩\) is the smallest sub-DKCT that contains \(s'\), and \(s' ∈ ⟨s⟩\), thus \(⟨s'⟩ ⊆ ⟨s⟩\).
+    Finally, because there is no accepting state in \(⟨s⟩\), there cannot be any accepting state in \(⟨s'⟩\) thus \(s'\) is also dead.
+\end{proof}
 
-FIXME: proof a theorem that if the transformation shares boolean with the original program, then the decision procedure is nearly linear time.
+In fact the normalization operation commutes with principle sub-DKCT.
+This result is intuitive as taking all the reachable states and then normalize is the same as normalize and then take all the reachable state.
+
+\begin{lemma}
+    sub-DKCT preserves principle sub-DKCT, that is, given \(𝒮' ⊑ 𝒮\) and \(s ∈ 𝒮'\), then \[⟨s⟩_{𝒮'} = ⟨s⟩_{𝒮}.\]
+\end{lemma}
+
+\begin{proof}
+    First by definition, \(s ∈ ⟨s⟩_{𝒮}\),
+    then we show that \(⟨s⟩_{𝒮}\) is the smallest sub-DKCT of \(𝒮'\), that contains \(s\) in \(𝒮\):
+
+    Because \(s ∈ ⟨s⟩_{𝒮'}\) and \(⟨s⟩_{𝒮'} ⊑ 𝒮' ⊑ 𝒮\), 
+    by minimality of \(⟨s⟩_{𝒮}\): \[⟨s⟩_{𝒮} ⊑ ⟨s⟩_{𝒮'} ⊑ 𝒮',\] that is, \(⟨s⟩_{𝒮}\) is a sub-DKCT of \(𝒮'\).
+
+    Given any \(𝒯  ⊑ 𝒮'\) and \(s ∈ 𝒯\), we have \(𝒯 ⊑ 𝒮' ⊑ 𝒮\),
+    thus by minimality of \(⟨s⟩_{𝒮}\): \(⟨s⟩_{𝒮} ⊑ 𝒯\).
+    In other word, \(⟨s⟩_{𝒮}\) is the the minimal sub-algebra of \(𝒮'\) that contains \(s\), reaching desired conclusion.
+\end{proof}
+
+\begin{lemma}
+    sub-DKCT preserves liveness, that is given \(𝒮' ⊑ 𝒮\) and \(s ∈ 𝒮'\), 
+    \[s \text{ is live in \(𝒮'\)} ⟺ s \text{ is live in \(𝒮\)}\]
+\end{lemma}
+\begin{proof}
+    recall that sub-DKCT preserves principle sub-DKCT, namely \(⟨s⟩_{𝒮'} = ⟨s⟩_{𝒮}\), thus 
+    \begin{align*}
+        s \text{ is live in \(𝒮'\)} 
+        & ⟺ ⟨s⟩_{𝒮'} \text{ contains an accepting state} \\
+        & ⟺ ⟨s⟩_{𝒮} \text{ contains an accepting state} \\
+        & ⟺ s \text{ is live in \(𝒮\)} 
+    \end{align*}
+\end{proof}
+
+\begin{lemma}
+    \(\norm\) is a functor, and it preserves inclusion, that is 
+    \[𝒮 ⊑ 𝒯 ⟹ \norm(𝒮) ⊑ \norm(𝒯).\]
+\end{lemma}
+
+\begin{lemma}
+    Given a DKCT \(𝒮\), and two sub-DKCT \(𝒮₁ ⊑ 𝒮\) and \(𝒮₂ ⊑ 𝒮\), \(𝒮₁ = 𝒮₂\) if and only if their carrier set is equal.
+\end{lemma}
+
+\begin{lemma}
+    Given a DKCT \(𝒮\), its dead states induces a sub-DKCT.
+\end{lemma}
+
+\begin{theorem}
+    Given a DKCT \(𝒮\) and a live state \(s ∈ 𝒮\):
+    \[⟨s⟩_{\norm(𝒮)} = \norm(⟨s⟩_{𝒮}).\]
+\end{theorem}
+
+\begin{proof}
+    We need to show \(\norm(⟨s⟩_{𝒮})\) is the smallest sub-DKCT in \(\norm(𝒮)\) that contains \(s\).
+
+    First, \(\norm(⟨s⟩_{𝒮})\) contains \(s\) because \(s\) is live. 
+    Second, take any \(𝒯 ⊑ \norm(𝒮)\), we will use \(\overline{𝒯}\) to denote the smallest sub-DKCT of \(𝒮\) that contains all the states in \(𝒯\). 
+    Since \(𝒯\) is a sub-DKCT of \(\norm(𝒮)\), \(\overline{𝒯} ∖ 𝒯\) can only contain dead states, thus \(\norm(\overline{𝒯}) = 𝒯\).
+    \begin{align*}
+        & ⟨s⟩_{𝒮} ⊑ \overline{𝒯} & \text{\(\overline{𝒯}\) contains \(s\)} \\  
+        ⟹ {}& \norm(⟨s⟩_{𝒮}) ⊑ \norm(\overline{𝒯}) & \text{monotonicity of \(\norm\)} \\  
+        ⟹ {}& \norm(⟨s⟩_{𝒮}) ⊑ 𝒯 & \norm(\overline{𝒯}) = 𝒯
+    \end{align*}
+\end{proof}
+
+We will show this by establishing the equivalence between ``normalized bisimulation'' and ``bisimulation in normalized coalgebra''.
+Given a GKAT coalgebra \(𝒮\), we can normalize it via removing the dead states and reroute all the transitions to dead state as direct rejections.
+Similarly, any relation \({∼} ⊆ 𝒮 × 𝒯\) can be normalized by filtering out all the states that are live:
+\[\norm(∼) ≜ \{(s, t) ∣ \text{both \(s\) and \(t\) are live states}\}.\]
+Thus there is a natural embedding \(\norm(∼) ↪ {∼}\) by inclusion. 
+Similarly, there is a embedding \(G(\norm(∼)) → (\closSim)^\At\) by lifting the following embedding point-wise via \((-)^\At\):
+\begin{align*}
+    2 + {\norm(∼)} × K & ↪ {\closSim} \\  
+    0 & ↦ (0, 0) \\
+    1 & ↦ (1, 1) \\
+    (s, t), p & ↦ (s, p), (t, p) 
+\end{align*}
+These two embeddings allows us to construct the following pullback square:
+\[
+    \begin{tikzcd}
+        \norm(∼) \ar[hook]{r}{i} \ar[dashed, swap]{d}{\norm(h)} & ∼ \ar{d}{h}\\  
+        G(\norm(∼)) \ar[hook]{r}{i} & (\closSim)^\At
+    \end{tikzcd}
+\]
+
+\begin{theorem}
+    A normalized bisimulation induces a bisimulation in the normal coalgebras.
+    That is 
+    \[
+        \begin{tikzcd}[column sep=1cm, row sep=1cm]
+            S \ar[swap]{d}{δ_𝒮}
+                & ∼  \ar[dashed, swap]{d}{(δ_𝒮, δ_𝒯)} \ar[swap]{l}{π₁} \ar{r}{π₂}
+                & T \ar{d}{δ_𝒯} \\  
+            G(S) 
+                & {\closSim}^\At 
+                    \ar[swap]{l}{{π₁}^\At} \ar{r}{{π₂}^\At}
+                & G(T)
+        \end{tikzcd}
+        ⟹
+        \begin{tikzcd}[column sep=1cm, row sep=1cm]
+            \norm(S) \ar[swap]{d}{δ_{\norm(𝒮)}}
+                & \norm(∼)  \ar[dashed, swap]{d}{\norm(h)} \ar[swap]{l}{π₁} \ar{r}{π₂}
+                & \norm(T) \ar{d}{δ_{\norm(𝒯)}} \\  
+            G(\norm(S)) 
+                & G(\norm(∼)) 
+                    \ar[swap]{l}{G(π₁)} \ar{r}{G(π₂)}
+                & G(\norm(T))
+        \end{tikzcd}
+    \]
+\end{theorem}
+
+\begin{proof}
+    We can 
+\end{proof}
+
+\begin{lemma}
+    Given a normalized bisimulation \({∼} ⊆ 𝒮 × 𝒯\), and a pair of state \(s ∼ t\), then the range of \(π₁: {∼} → S\), will cover all the live states in \(⟨s⟩\);
+    similarly for the state \(t ∈ 𝒯\).
+\end{lemma}
+
+\begin{proof}
+    let \(\norm(∼)|_{⟨s⟩}\) be all the live elements of \(∼\) restricted to \(⟨s⟩\):
+    \[{\norm(∼)|_{⟨s⟩}} ≜ \{(s', t') ∈ {∼}∣ s' ∈ ⟨s⟩, \text{both \(s'\) and \(t'\) are live}\}.\]
+    since all the live states in \(⟨s⟩\) is the carrier of \(\norm(⟨s⟩)\),
+    we only need to show \(π₁: {\norm(∼)|_{⟨s⟩}} → ⟨s⟩\) is surjective.
+
+    Since \(∼\) is a normalize bi-simulation, then \(\norm(∼)\) is a bisimulation,
+    and the following diagram commute by computation:
+    \[
+    \begin{tikzcd}
+        \norm(∼)|_{⟨s⟩} \ar{r}{π₁} \ar[swap]{d}{h} 
+            & \norm(⟨s⟩) \ar{d}{δ_{\norm(𝒮)}}\\  
+        G(\norm(∼)|_{⟨s⟩}) \ar{r}{G(π₁)} & G(\norm(⟨s⟩))
+    \end{tikzcd}
+    \]
+    Because of above diagram, \(\Img(π₁) ⊆ \norm(⟨s⟩)\), the range of \(π₁\), also forms a sub-DKCT of \(\norm(𝒮)\) with the transition: \(δ_{\norm(𝒮)}: \Img(π₁) → G(\Img(π₁))\)
+    then there are two cases:
+    \begin{itemize}
+        \item if \(s\) is dead, then \(\norm(⟨s⟩)\) is empty, thus \(π₁\) is surjective,
+        \item if \(s\) is live, then \(\norm(⟨s⟩) = ⟨s⟩_{\norm(𝒮)}\), because \(\Img(π₁)\) is a sub-DKCT and contains \(s\), thus \[\norm(⟨s⟩) = ⟨s⟩_{\norm(𝒮)} ⊑ \Img(π₁).\] 
+        Hence \(\Img(π₁)\) covers \(\norm(⟨s⟩)\) and \(π₁\) is surjective.
+    \end{itemize}
+\end{proof}
+
+% \begin{theorem}
+%     Given a normalized bisimulation \({∼} ⊆ 𝒮 × 𝒯\), and two states \(s ∼ t\), then \(∼\) can be restricted to the principle sub-DKCT \(⟨s⟩\) and \(⟨t⟩\): 
+%     \[
+%     \begin{tikzcd}
+%         ⟨s⟩ \ar[swap]{d}{δ_𝒮 |_{⟨s⟩}}
+%             & ∼|_{⟨s⟩ × ⟨t⟩} \ar[swap]{l}{π₁} \ar{r}{π₂} \ar{d}{h}
+%             & ⟨t⟩ \ar{d}{δ_𝒮 |_{⟨s⟩}} \\  
+%         G(⟨s⟩)
+%             & (\overline{∼|_{⟨s⟩ × ⟨t⟩}})^{\At} \ar[swap]{l}{(π₁)^\At} \ar{r}{(π₂)^\At} 
+%             & G(⟨t⟩)
+%     \end{tikzcd}
+%     \]
+%     In this case, \(π₁\) and \(π₂\) is surjective.
+% \end{theorem}
+
+% \begin{proof}
+%     Take any element \(s' ∈ \Img(π₁)\), we can show that \(δ_𝒮(s') ∈ G(Img(π))\),
+%     in other word, \(\Img(π₁)\) from a sub-DKCT of \(𝒮\).
+%     NTS \(δ_𝒮(s', α) = π₁(h(s', α)) ∈ 2 + Img(π) × K\).
+
+%     Note that because since \(s ∼ t\), \(\Img(π₁)\) contains \(s\) and \(\Img(π₁)\) contains \(t\).
+%     Because \(⟨s⟩\) and \(⟨t⟩\) are the smallest sub-DKCTs that contain \(s\) and \(t\) respectively, thus the image of \(π₁\) is \(⟨s⟩\) and the image of \(π₂\) is \(⟨t⟩\), hence both \(π₁\) and \(π₂\) are surjective.
+% \end{proof}
+
+\begin{theorem}
+    Given a dead state \(s\) and a live state \(t\), there doesn't exists a normalized bisimulation \(∼\) s.t. \(s ∼ t\) or \(t ∼ s\).
+\end{theorem}
+
+\begin{proof}
+    WLOG we will only prove that \(s ∼ t\) will lead to a contradiction.
+    We can restrict the simulation diagram of \(∼\) to \(⟨s⟩\) and \(\norm(⟨t⟩)\),
+    let \[∼_{s, t} = \{s', t' ∣ s' ∼ t', s' ∈ ⟨s⟩, t' ∈ \norm(⟨t⟩)\}.\]
+    Note that only \(𝒯\) is restricted to \(\norm(⟨t⟩)\); while \(𝒮\) is restricted to \(⟨s⟩\), without \(\norm\). 
+    This discrepancy is because \(s\) is dead, thus \(\norm(⟨s⟩)\) is empty. 
+    If we restrict \(∼\) to \(\norm(⟨t⟩) × \norm(⟨s⟩)\), then there will be no element in the restriction, since \(\norm(⟨t⟩) × \norm(⟨s⟩)\) is empty.
+
+    We can show that \(∼_{s, t}\) form a normalized bisimulation between \(⟨s⟩\) and \(\norm(⟨t⟩)\).
+    \[
+        \begin{tikzcd}[column sep=1cm, row sep=1cm]
+            ⟨s⟩ \ar[swap]{d}{δ_𝒮}
+                & ∼_{s, t} \ar{d}{(δ_𝒮, \norm(δ_𝒯))} \ar[swap]{l}{π₁} \ar{r}{π₂}
+                & \norm(⟨t⟩) \ar{d}{\norm(δ_𝒯)} \\  
+            G(⟨s⟩) 
+                & {\overline{∼_{s, t}}}^\At 
+                    \ar[swap]{l}{{π₁}^\At} \ar{r}{{π₂}^\At}
+                & G(\norm(⟨t⟩))
+        \end{tikzcd}
+    \]
+    Then we can perform epi-mono factorization on the right square to obtain the range of \(π₂\):
+    \[
+        \begin{tikzcd}[column sep=1cm, row sep=1cm]
+            ∼_{s, t} \ar[swap]{d}{(δ_𝒮, \norm(δ_𝒯))} \ar{r}{π₂}
+                & \Img(π₂) \ar[hook]{r}{i} \ar{d}{\norm(δ_𝒯)}
+                & \norm(⟨t⟩) \ar{d}{\norm(δ_𝒯)} \\  
+            {\overline{∼_{s, t}}}^\At  \ar{r}{π₂^{\At}}
+                & G(\Img(π₂)) \ar[hook]{r}{G(i)}
+                & G(\norm(⟨t⟩))
+        \end{tikzcd}
+    \]
+    Thus, \[\Img(π₂) ⊑ \norm(⟨t⟩) ⊑ \norm(𝒯).\]
+    Recall that \(\norm(⟨t⟩_{𝒯}) = ⟨t⟩_{\norm(𝒯)}\), hence \(\norm(⟨t⟩)\) is the smallest sub-DKCT of \(\norm(𝒯)\),  since \(s ∼ t\), thus \(t ∈ \Img(π₂)\).
+    Hence \(\Img(π₂) ⊒ \norm(⟨t⟩)\), thus \(π₂\) is surjective.
+
+    Since \(t\) is a live state, hence there exists an accepting state \(t'\) in \(\norm(⟨t⟩)\); 
+    because \(s\) is a dead state, thus all the states in \(⟨s⟩\) is dead.
+    Because \(π₂: {∼_{s, t}} → T\) is surjective, there exists a dead state \(s' ∈ ⟨s⟩\) s.t. \(s' ∼ t'\).
+    However, a dead state cannot be normally bisimilar to a accepting state, since \(((s', p), 1) ∉ \overline{∼_{s, t}}\) for all dead state \(s'\) and action \(p\).
+\end{proof}
+
+\begin{theorem}[soundness and completenss]
+    Given a normalized bisimulation \({∼} ⊆ 𝒮 × 𝒯\), then for all state \(s ∈ 𝒮\) and \(t ∈ 𝒯\), \(∼\) is sound and complete with respect to trace equivalence
+    \[s ∼ t \text{ or both \(s\) and \(t\) are dead} ⟺ ⟦s⟧ = ⟦t⟧.\]
+\end{theorem}
+
+\begin{proof}
+    The \(⟹\) direction can be shown by case analysis on the liveness of \(s\) and \(t\):
+    \begin{itemize}
+        \item If both \(s\) and \(t\) are live, 
+            \[s ∼ t ⟹ (s, t) ∈ \norm(∼) ⟹ ⟦s⟧ = ⟦t⟧.\]
+        \item If one of \(s\) and \(t\) is live, and the other is dead, then \(s ∼ t\) cannot hold.
+        \item If both \(s\) and \(t\) are dead, since dead states don't have any trace:
+            \[⟦s⟧ = ⟦t⟧ = ∅.\]
+    \end{itemize}
+
+    The \(⟸\) direction can also be shown using similar case analysis:
+    \begin{itemize}
+        \item If both \(s\) and \(t\) are live, then by definition of \(\norm(∼)\)
+            \[⟦s⟧ = ⟦t⟧ ⟹ (s, t) ∈ \norm(∼) ⟹ s ∼ t.\]
+        \item If one of \(s\) and \(t\) is live, and the other is dead, then one would contain traces, yet the other don't. Hence their trace semantics cannot equal.
+        \item If both \(s\) and \(t\) are dead, we have reached our conclusion.
+    \end{itemize}
+\end{proof}
+
+\begin{corollary}
+    The trace semantics of a state is preserved in sub-DKCTs, that is given \(𝒮 '⊑ 𝒮\) and a state \(s ∈ 𝒮'\), \(⟦s⟧_{𝒮} = ⟦s⟧_{𝒮'}\)
+\end{corollary}
+
+\begin{proof}
+    We only need to exhibit a normalized bisimulation between \(𝒮'\) and \(𝒮\):
+    \[s ∼ t \text{ when \(s = t\) or \(s\) and \(t\) are both dead}.\]
+    To show that the relation \(∼\) is a normalized bisimulation, we only need to check the commutativity of the following diagram:
+    \[
+    \begin{tikzcd}[column sep=1cm, row sep=1cm]
+        𝒮' \ar[swap]{d}{δ_𝒮}
+            & ∼ \ar{d}{(δ_𝒮, δ_𝒮)} \ar[swap]{l}{π₁} \ar{r}{π₂}
+            & 𝒮 \ar{d}{δ_𝒮} \\  
+        G(𝒮') 
+            & {\overline{∼}}^\At 
+                \ar[swap]{l}{{π₁}^\At} \ar{r}{{π₂}^\At}
+            & G(𝒮)
+    \end{tikzcd}
+    \]
+    The commutativity is apparent by computation. The function \((δ_𝒮, δ_𝒮): {∼} → {\overline{∼}}^\At\) is well-defined, by case analysis on the input:
+    \begin{itemize}
+        \item if \(s = t\), then \(δ_𝒮(s, α) = δ_𝒮(t, α)\), hence will be in \(\overline{∼}\) for all \(α\).
+        \item if both \(s\) and \(t\) are dead, then for all \(α ∈ \At\) both \(δ_𝒮(s, α)\) and \(δ_𝒮(t, α)\) are will either transition to a dead state or reject, thus the results will also be in \(\overline{∼}\).
+    \end{itemize}
+\end{proof}
+
+Finally, since sub-DKCT preserves semantics, we can restrict \(∼\) to \(⟨s⟩\) and \(⟨t⟩\) to minimize the search space. 
+We denote the restricted normalized bisimulation as \(∼_{s, t}\):
+\[
+    \begin{tikzcd}[column sep=1cm, row sep=1cm]
+        ⟨s⟩ \ar[swap]{d}{δ_𝒮}
+            & ∼_{s, t} \ar{d}{(δ_𝒮, δ_𝒯)} \ar[swap]{l}{π₁} \ar{r}{π₂}
+            & ⟨t⟩ \ar{d}{δ_𝒯} \\  
+        G(⟨s⟩) 
+            & {\overline{∼_{s, t}}}^\At 
+                \ar[swap]{l}{{π₁}^\At} \ar{r}{{π₂}^\At}
+            & G(⟨t⟩)
+    \end{tikzcd}
+\]
+The because sub-DKCT preserves trace semantics, thus the completeness result still holds: 
+\[s ∼_{s, t} t \text{ or both \(s\) and \(t\) are dead} ⟺ ⟦s⟧ = ⟦t⟧,\] 
+Our algorithm will focus on finding \(∼\) in a single DKCT with transition function \(δ\). 
+This approach is general enough, when given \(𝒮\) and \(𝒯\), we can merge them via coproduct \(𝒮 + 𝒯\), where both \(𝒮\) and \(𝒯\) are sub-DKCT of \(𝒮 + 𝒯\), thus preserving the semantics of all the states.
+
+Given a state \(s ∈ 𝒮\) we will run a simple depth-first search in \(⟨s⟩\) to determine whether \(s\) is dead.
+Specifically recall that \(s\) is live if and only if there exists an accepting state in \(s\), and if \(s\) is dead, then all of \(⟨s⟩\) is dead
+\begin{lstlisting}
+(** return whether the state is accepting*)
+let is_accept (s: state): bool = 
+    [*\(∃ α ∈ \At, δ(s, α) = \accept\)*]
+
+(** a mutable set to keep track of states 
+that has been explored by check_dead function*)
+let explored = [*\(∅\)*]
+(** return a all the states in [*\(⟨s⟩\)*] if s is dead, 
+otherwise return None*)
+let check_dead (s: state): state set option =
+    (* [*\(⟨s⟩\)*] has all be explored *)
+    if s [*\(∈\)*] explored then [*\(∅\)*]
+
+    (* exploring [*\(⟨s⟩\)*] *)
+    else for [*\(α ∈ \At\)*]:
+        if [*\(δ(s, α) = \accept\)*] then None 
+        else if [*\(δ(s, α) = (s', p)\)*] then 
+            if check_dead s' = None then None 
+            else explored := explored [*\(∪\)*] check_dead s'
+        (* [*\(δ(s, α) = \reject\)*] case, skip *)
+
+    (* finished exploring [*\(⟨s⟩\)*] *)
+    explored [*\(∪\)*] {s}
+\end{lstlisting}
+To achieve better efficiency, we can cache all the dead state we found:
+\begin{lstlisting}
+(** a mutable set to keep track of all the found dead states *)
+let dead_states = [*\(∅\)*]
+
+(** a function to check whether a state is dead *)
+let is_dead (s: state): bool =
+    if s [*\(∈\)*] dead_states then true 
+    else if check_dead s = None then false 
+    else 
+        dead_states := dead_states [*\(∪\)*] check_dead s
+        return true
+\end{lstlisting}
+Then we can design our algorithm to construct \(∼\) from two input...  
+TODO: think a bit about the best way to check for both dead.
 

### 2024-05-13 notes.tex
finish normalized bisimulation

@@
 \begin{lemma}
+    Given two sub-DKCT \(𝒮₁ ⊑ 𝒮\) and \(𝒮₂ ⊑ 𝒯\), then \[𝒮₁ ⊑ 𝒮₂ ⟺ 𝒮₁ ⊆ 𝒮₂\]
+\end{lemma}
+
+\begin{proof}
+    because their transition function are both restriction of \(δ_𝒮\).
+\end{proof}
+
+\begin{lemma}
+    Homomorphism preserves sub-DKCT, given \(h: 𝒮 → 𝒯\):
+    \[𝒮' ⊑ 𝒮 ⟹ h(𝒮') ⊑ h(𝒮) ⊑ 𝒯\]
+\end{lemma}
+
+\begin{proof}
+    consequence of previous lemma.
+\end{proof}
+
+\begin{lemma}
+    Homomorphism preserves principle sub-DKCT. Specifically, given a homomorphism \(h: 𝒮 → 𝒯\), 
+    \[h(⟨s⟩_{𝒮}) = ⟨h(s)⟩_{𝒯},\]
+    where \(h(⟨s⟩_{𝒮})\) is the image of \(⟨s⟩_{𝒮}\) under \(h\).
+\end{lemma}
+
+\begin{proof}
+    Specifically, we will need to show that \(h(⟨s⟩_{𝒮})\) is the smallest sub-DKCT of \(𝒯\) that contain \(h(s)\).
+First by definition of image, \(h(s) ∈ h(⟨s⟩_{𝒮})\), second because \(h\) is a homomorphism \(𝒮 → 𝒯\), and \(⟨s⟩_{𝒮} ⊑ 𝒮\), thus \(h(⟨s⟩_{𝒮}) ⊑ 𝒯\).
+
+    Finally, take any \(𝒯' ⊑ 𝒯\), and \(h(s) ∈ 𝒯'\). We take the preimage of \(𝒯'\) under \(h\),
+    since \(s ∈ h^{-1}(𝒯)\), thus \(⟨s⟩_{𝒮} ⊑ h^{-1}(𝒯)\), which implies \(h(⟨s⟩_{𝒮}) ⊑ 𝒯\).
+\end{proof}
+
+\begin{corollary}
+    Homomorphism preserves live/dead/accepting states.
+    Given a homomorphism \(h: 𝒮 → 𝒯\) then \(s ∈ 𝒮\) is a live/dead/accepting state if and only if \(h(s) ∈ 𝒯\) is a live/dead/accepting state, respectively.
+\end{corollary}
+
+\begin{proof}
+    First, homomorphism preserves accepting state by definition.
+    Then because homomorphism preserves principle sub-DKCT and a state \(s\) is live if and only if there exists an accepting state \(⟨s⟩\), thus homomorphism will also preserve the liveness of \(s\).
+\end{proof}
+
+\begin{corollary}
     sub-DKCT preserves principle sub-DKCT, that is, given \(𝒮' ⊑ 𝒮\) and \(s ∈ 𝒮'\), then \[⟨s⟩_{𝒮'} = ⟨s⟩_{𝒮}.\]
-\end{lemma}
-
-\begin{proof}
-    First by definition, \(s ∈ ⟨s⟩_{𝒮}\),
-    then we show that \(⟨s⟩_{𝒮}\) is the smallest sub-DKCT of \(𝒮'\), that contains \(s\) in \(𝒮\):
-
-    Because \(s ∈ ⟨s⟩_{𝒮'}\) and \(⟨s⟩_{𝒮'} ⊑ 𝒮' ⊑ 𝒮\), 
-    by minimality of \(⟨s⟩_{𝒮}\): \[⟨s⟩_{𝒮} ⊑ ⟨s⟩_{𝒮'} ⊑ 𝒮',\] that is, \(⟨s⟩_{𝒮}\) is a sub-DKCT of \(𝒮'\).
-
-    Given any \(𝒯  ⊑ 𝒮'\) and \(s ∈ 𝒯\), we have \(𝒯 ⊑ 𝒮' ⊑ 𝒮\),
-    thus by minimality of \(⟨s⟩_{𝒮}\): \(⟨s⟩_{𝒮} ⊑ 𝒯\).
-    In other word, \(⟨s⟩_{𝒮}\) is the the minimal sub-algebra of \(𝒮'\) that contains \(s\), reaching desired conclusion.
-\end{proof}
-
-\begin{lemma}
+\end{corollary}
+
+\begin{proof}
+    Because inclusion \(i: 𝒮' → 𝒮\) is a homomorphism, thus \[⟨s⟩_{𝒮'} = i(⟨s⟩_{𝒮'}) = ⟨i(s)⟩_𝒮 = ⟨s⟩_{𝒮}. \qedhere\]
+\end{proof}
+
+\begin{corollary}
     sub-DKCT preserves liveness, that is given \(𝒮' ⊑ 𝒮\) and \(s ∈ 𝒮'\), 
     \[s \text{ is live in \(𝒮'\)} ⟺ s \text{ is live in \(𝒮\)}\]
-\end{lemma}
-\begin{proof}
-    recall that sub-DKCT preserves principle sub-DKCT, namely \(⟨s⟩_{𝒮'} = ⟨s⟩_{𝒮}\), thus 
-    \begin{align*}
-        s \text{ is live in \(𝒮'\)} 
-        & ⟺ ⟨s⟩_{𝒮'} \text{ contains an accepting state} \\
-        & ⟺ ⟨s⟩_{𝒮} \text{ contains an accepting state} \\
-        & ⟺ s \text{ is live in \(𝒮\)} 
-    \end{align*}
-\end{proof}
+\end{corollary}
 
@@
 \begin{proof}
-    We can 
+    By computation.
 \end{proof}
@@
     We can restrict the simulation diagram of \(∼\) to \(⟨s⟩\) and \(\norm(⟨t⟩)\),
-    let \[∼_{s, t} = \{s', t' ∣ s' ∼ t', s' ∈ ⟨s⟩, t' ∈ \norm(⟨t⟩)\}.\]
+    let \[∼_{s, t̂} = \{s', t' ∣ s' ∼ t', s' ∈ ⟨s⟩, t' ∈ \norm(⟨t⟩)\}.\]
     Note that only \(𝒯\) is restricted to \(\norm(⟨t⟩)\); while \(𝒮\) is restricted to \(⟨s⟩\), without \(\norm\). 
@@
 
-    We can show that \(∼_{s, t}\) form a normalized bisimulation between \(⟨s⟩\) and \(\norm(⟨t⟩)\).
+    We can show that \(∼_{s, t̂}\) form a normalized bisimulation between \(⟨s⟩\) and \(\norm(⟨t⟩)\).
     \[
@@
             ⟨s⟩ \ar[swap]{d}{δ_𝒮}
-                & ∼_{s, t} \ar{d}{(δ_𝒮, \norm(δ_𝒯))} \ar[swap]{l}{π₁} \ar{r}{π₂}
+                & ∼_{s, t̂} \ar{d}{(δ_𝒮, \norm(δ_𝒯))} \ar[swap]{l}{π₁} \ar{r}{π₂}
                 & \norm(⟨t⟩) \ar{d}{\norm(δ_𝒯)} \\  
             G(⟨s⟩) 
-                & {\overline{∼_{s, t}}}^\At 
+                & {\overline{∼_{s, t̂}}}^\At 
                     \ar[swap]{l}{{π₁}^\At} \ar{r}{{π₂}^\At}
@@
         \begin{tikzcd}[column sep=1cm, row sep=1cm]
-            ∼_{s, t} \ar[swap]{d}{(δ_𝒮, \norm(δ_𝒯))} \ar{r}{π₂}
+            ∼_{s, t̂} \ar[swap]{d}{(δ_𝒮, \norm(δ_𝒯))} \ar{r}{π₂}
                 & \Img(π₂) \ar[hook]{r}{i} \ar{d}{\norm(δ_𝒯)}
                 & \norm(⟨t⟩) \ar{d}{\norm(δ_𝒯)} \\  
-            {\overline{∼_{s, t}}}^\At  \ar{r}{π₂^{\At}}
+            {\overline{∼_{s, t̂}}}^\At  \ar{r}{π₂^{\At}}
                 & G(\Img(π₂)) \ar[hook]{r}{G(i)}
@@
     because \(s\) is a dead state, thus all the states in \(⟨s⟩\) is dead.
-    Because \(π₂: {∼_{s, t}} → T\) is surjective, there exists a dead state \(s' ∈ ⟨s⟩\) s.t. \(s' ∼ t'\).
-    However, a dead state cannot be normally bisimilar to a accepting state, since \(((s', p), 1) ∉ \overline{∼_{s, t}}\) for all dead state \(s'\) and action \(p\).
-\end{proof}
-
-\begin{theorem}[soundness and completenss]
-    Given a normalized bisimulation \({∼} ⊆ 𝒮 × 𝒯\), then for all state \(s ∈ 𝒮\) and \(t ∈ 𝒯\), \(∼\) is sound and complete with respect to trace equivalence
-    \[s ∼ t \text{ or both \(s\) and \(t\) are dead} ⟺ ⟦s⟧ = ⟦t⟧.\]
+    Because \(π₂: {∼_{s, t̂}} → T\) is surjective, there exists a dead state \(s' ∈ ⟨s⟩\) s.t. \(s' ∼ t'\).
+    However, a dead state cannot be normally bisimilar to a accepting state, since \(((s', p), 1) ∉ \overline{∼_{s, t̂}}\) for all dead state \(s'\) and action \(p\).
+\end{proof}
+
+\begin{theorem}
+    Homomorphism reflects normalized bisimulation. Let \(h: 𝒮₁ → 𝒮₂\), and a normalized bisimulation \({∼} ⊆ 𝒮₂ × 𝒯\), then there exists a normalized bisimulation \({∼ₕ}: 𝒮₁ × 𝒯\).
+    And similarly for \({∼} ⊆ 𝒯 × 𝒮₂\).
 \end{theorem}
@@
 \begin{proof}
-    The \(⟹\) direction can be shown by case analysis on the liveness of \(s\) and \(t\):
+    \(∼ₕ\) can be defined as follows: 
+    \[{∼ₕ} ≜ \{(s, t) ∣ h(s) ∼ t\}.\]
+    Then the function \(h × id: {∼ₕ} → {∼}\), can be lifted to \({{\overline{∼ₕ}}^\At} → {\closSim^\At}\):
+    \[
+        \begin{tikzcd}[column sep=2.5cm]
+            ∼ \ar[swap]{d}{(δ_{𝒮₂}, δ_𝒯)}
+                & ∼ₕ \ar[swap]{l}{h × id} \ar{d}{(δ_{𝒮₁}, δ_𝒯)}   \\ 
+            \closSim^\At
+                & {\overline{∼ₕ}}^\At \ar[dashed]{l}{((id + h × id) × id)^\At}
+        \end{tikzcd}
+    \]
+    The generated function \((id + h × id) × id: {\overline{∼ₕ}} → {\closSim}\) is indeed well-defined by case analysis on the input in \(∼ₕ\), and also by the fact that homomorphism like \(h\) preserves liveness.
+\end{proof}
+
+\begin{definition}
+    A guarded language \(L\) over alphabet \(K, B\) is deterministic when given a list of atoms \(α₀, α₁, α₂, …, αₙ\),
+    there is at most one word in the language \(w ∈ L\) that has the form:
+    \[α₀ p₁ α₁ … pₙ αₙ \text{ for some } pᵢ ∈ K\]
+\end{definition}
+
+\begin{theorem}
+    the set of deterministic guarded language over the alphabet \(K, B\): \(𝒢_{K, B}\) is a DKCT;  
+    and the trace semantics \(⟦-⟧: 𝒮 → 𝒢\) is a homomorphism.
+\end{theorem}
+
+\begin{theorem}[soundness and completeness]
+    the trace of two state \(s ∈ 𝒮\) and \(t ∈ 𝒯\) are trace equivalent if and only if there exists a normalized bisimulation between them
+    \[⟦s⟧ = ⟦t⟧ ⟺ \text{there exists a normalized bisimulation \(∼\) s.t. } s ∼ t.\]
+\end{theorem}
+
+\begin{proof}
+    The \(⟸\) direction can be shown by case analysis on the liveness of \(s\) and \(t\):
     \begin{itemize}
@@
 
-    The \(⟸\) direction can also be shown using similar case analysis:
-    \begin{itemize}
-        \item If both \(s\) and \(t\) are live, then by definition of \(\norm(∼)\)
-            \[⟦s⟧ = ⟦t⟧ ⟹ (s, t) ∈ \norm(∼) ⟹ s ∼ t.\]
-        \item If one of \(s\) and \(t\) is live, and the other is dead, then one would contain traces, yet the other don't. Hence their trace semantics cannot equal.
-        \item If both \(s\) and \(t\) are dead, we have reached our conclusion.
-    \end{itemize}
+    \(⟹\) direction can be shown by constructing the identity normalized bisimulation on languages: \({∼_{id}} ⊆ 𝒢 × 𝒢\) and then reflect both side by the language semantics \(⟦-⟧: 𝒮 → 𝒢\) and \(⟦-⟧: 𝒯 → 𝒢\).
+Thus we obtain a normalized bisimulation on \(𝒮\) and \(𝒯\).
+\end{proof}
+
+\begin{corollary}
+    The language equivalence is the maximal normalized bisimulation.  
+\end{corollary}
+
+\begin{corollary}
+    There exists a normalized bisimulation s.t. \(s ∼ t\) if and only if there exists a normalized bisimulation equivalence \(≡\) s.t. \(s ≡ t\).
+\end{corollary}
+
+\begin{proof}
+    take the \(≡\) to be language equivalence.
 \end{proof}
@@
 
-Finally, since sub-DKCT preserves semantics, we can restrict \(∼\) to \(⟨s⟩\) and \(⟨t⟩\) to minimize the search space. 
-We denote the restricted normalized bisimulation as \(∼_{s, t}\):
+Since sub-DKCT preserves semantics, we only need to exhibit a normalized bisimulation \({∼} ⊆ ⟨s⟩ × ⟨t⟩\), to shrink the size of the bisimulation:
 \[
@@
 The because sub-DKCT preserves trace semantics, thus the completeness result still holds: 
-\[s ∼_{s, t} t \text{ or both \(s\) and \(t\) are dead} ⟺ ⟦s⟧ = ⟦t⟧,\] 
-Our algorithm will focus on finding \(∼\) in a single DKCT with transition function \(δ\). 
-This approach is general enough, when given \(𝒮\) and \(𝒯\), we can merge them via coproduct \(𝒮 + 𝒯\), where both \(𝒮\) and \(𝒯\) are sub-DKCT of \(𝒮 + 𝒯\), thus preserving the semantics of all the states.
-
-Given a state \(s ∈ 𝒮\) we will run a simple depth-first search in \(⟨s⟩\) to determine whether \(s\) is dead.
+\[⟦s⟧_𝒮 = ⟦t⟧_𝒯 ⟺ ⟦s⟧_{⟨s⟩} = ⟦t⟧_{⟨t⟩} ⟺ \text{there exists } {∼} ⊆ ⟨s⟩ × ⟨t⟩ \text{ such that } s ∼ t.\]
+
+Finally, to construct a normalized bisimulation using programs, we will need the following inductive construction theorem
+\begin{theorem}[Inductive Construction]
+    Given two DKCT \(𝒮\) and \(𝒯\), and two of their elements \(s ∈ 𝒮\) and \(t ∈ 𝒯\),
+    there exists a normalized bisimulation \(s ∼ t\) between \(⟨s⟩\) and \(⟨t⟩\), if and only if either both \(s\) and \(t\) are dead or for all \(α ∈ \At\), all of the following holds:
+    \begin{enumerate}
+        \item \(δ_{𝒮}(s, α) = \accept ⟺ δ_{𝒯}(t, α) = \accept\);
+        \item\label{itm: transition condition} If \(δ_{𝒮}(s, α) = (s', p)\) and \(δ_{𝒯}(t, α) = (t', p)\), then and there exists a normalized bisimulation \({∼_{s',t'}}\) on \(⟨s'⟩\) and \(⟨t'⟩\), s.t.
+\(s' ∼_{s',t'} t'\);
+        \item If \(δ_{𝒮}(s, α) = (s', p)\) and \(δ_{𝒯}(t, α) = (t', q)\), s.t. \(p ≠ q\), then both \(s'\) and \(t'\) are dead;
+        \item \(s\) reject \(α\) or transition to a dead state via \(α\) if and only if \(t\) rejects \(α\) or transition to a dead state via \(α\).
+    \end{enumerate}
+\end{theorem}
+
+\begin{proof}
+    We first prove the \(⟹\) direction: if there exists a normalized bisimulation \(∼\), then for all \(α ∈ \At\), \(δ_{𝒮}(s, α) \closSim δ_{𝒯}(t, α)\).
+And by unfolding the definition of \(\closSim\), we can obtain all the conditions we desired.
+
+    For the \(⟸\) direction, we can explicitly construct \(∼\) as the union of all the normalized bisimulations \(∼_{s', t'} ⊆ ⟨s'⟩ × ⟨t'⟩\): 
+    \[{∼} = \{(s, t)\} ∪ \{{∼_{s', t'}} ∣ ∃ α ∈ \At, p ∈ K, δ_𝒮(s, α) = (s', p) \text{ and } δ_𝒯(t, α) = (t', p)\}.\]
+    Then, we will need to show \(∼\) is indeed a normalized bisimulation on \(⟨s⟩\) and \(⟨t⟩\), that is for all \(s' ∈ ⟨s⟩\) and \(t' ∈ ⟨t⟩\):
+    \[s' ∼ t' ⟹ δ_𝒮(s', α) \closSim δ_𝒯(t', α).\]
+    We only need to show the above implication for \(s ∼ t\), as other element is already in some normalized bisimulation between sub-DKCTs of \(𝒮\) and \(𝒯\).
+    And the implication \(s ∼ t ⟹ δ_𝒮(s', α) \closSim δ_𝒯(t', α)\) can be proven by unfolding the definition of \(\closSim\).
+\end{proof}
+
+Given a state \(s ∈ 𝒮\) we can run a simple depth-first search in \(⟨s⟩\) to determine whether \(s\) is dead.
 Specifically recall that \(s\) is live if and only if there exists an accepting state in \(s\), and if \(s\) is dead, then all of \(⟨s⟩\) is dead
@@
 \end{lstlisting}
-Then we can design our algorithm to construct \(∼\) from two input...  
-TODO: think a bit about the best way to check for both dead.
+
+Thus our algorithm can be directly obtained from previous theorem, with couple modifications
+\begin{itemize}
+    \item  we will finding normalized bisimulation equivalence \(∼\) in a DKCT with transition function \(δ\), as oppose to in \(𝒮\) and \(𝒯\). 
+    This approach is general enough: when given \(𝒮\) and \(𝒯\), we can merge them via coproduct \(𝒮 + 𝒯\), where both \(𝒮\) and \(𝒯\) are sub-DKCT of \(𝒮 + 𝒯\), thus preserving the semantics of all the states.
+    \item we will find a bisimulation equivalence instead of bisimulation relation, this allows us to use union-find structure to keep track of related states.
+In particular, the \texttt{union} function will mark two states as equivalent, and \texttt{eq} function will return whether two states has the same representative in the union-find.
+\end{itemize}
+
+\begin{lstlisting}
+(** computing the normalized bisimulation equivalence using union find *)
+let equiv (s: state) (t: state): bool = 
+    (* if s and t are already in the normalized bisimulation equivalence *)
+    if eq s t then true else
+
+    (* checking if they are both dead *)
+    if s [*\(∈\)*] dead_states then is_dead t else 
+    if t [*\(∈\)*] dead_states then is_dead s else
+
+    (* all of the following condition needs to be true for all atoms *)
+    forall ([*\(α ∈ \At\)*]): 
+        match [*\(δ(s, α), δ(t, α)\)*] with 
+            [*\(\accept, \accept\)*] -> true 
+            [*\((s', p), (t', q)\)*] -> 
+                if p = q 
+                    then equiv s' t' 
+                    else is_dead s' && is_dead t'
+            [*\((s', p), \reject\)*] -> is_dead s'
+            [*\(\reject, (t', p)\)*] -> is_dead t'
+            [*\(\reject, \reject\)*] -> true 
+            _ -> false
+\end{lstlisting}
+TODO: I think I still need to use one of those algorithm package...
+
+The correctness of this algorithm can be exhibited by the following induction invariant:
+\[\texttt{equiv s t} ⟺ \text{there exists a normalized bisimulation equivalence \(∼\) s.t. \(s ∼ t\)}\]
+Then by the soundness and completeness of normalized bisimulation:
+\[\texttt{equiv s t} ⟺ ⟦s⟧ = ⟦t⟧.\]
 

### 2024-05-14 notes.tex
typo

@@
                 if p = q 
-                    then equiv s' t' 
+                    (* mark s and t as bisimular and check the rest*)
+                    then union s t; equiv s' t' 
                     else is_dead s' && is_dead t'

### 2024-05-21 notes.tex
minor update that doesn't matter that much

@@
 
-Bisimilarity implies language sematnical equivalence \[s ∼ t ⟹ ⟦s⟧ = ⟦t⟧\], however the converse doesn't hold unless we are working in a \emph{normal} GKAT coalgebra.
+Bisimilarity implies language semantical equivalence \[s ∼ t ⟹ ⟦s⟧ = ⟦t⟧\], however the converse doesn't hold unless we are working in a \emph{normal} GKAT coalgebra.
 It is well known that two states in a normal GKAT coalgebra if and only if they have the same language semantics \[s ∼ t ⟺ ⟦s⟧ = ⟦t⟧.\]
@@
     Given two DKCT \(𝒮\) and \(𝒯\), and two of their elements \(s ∈ 𝒮\) and \(t ∈ 𝒯\),
-    there exists a normalized bisimulation \(s ∼ t\) between \(⟨s⟩\) and \(⟨t⟩\), if and only if either both \(s\) and \(t\) are dead or for all \(α ∈ \At\), all of the following holds:
+    there exists a normalized bisimulation equivalence \({∼} ⊆ ⟨s⟩ × ⟨t⟩\) s.t. \(s ∼ t\), if and only if either both \(s\) and \(t\) are dead or for all \(α ∈ \At\), all of the following holds:
     \begin{enumerate}
@@
 \end{proof}
+
+\begin{corollary}
+    when constructing \(∼\) notice that 
+\end{corollary}
 
@@
 
+\section{Symbolic Coalgebra}
+
+\begin{theorem}
+    Given two states \(s, t\) there exists a normalized bisimulation \({∼} ⊆ ⟨s⟩ × ⟨t⟩\) if and only if the following is true 
+    \begin{itemize}
+        \item \(ϵ(s) = ϵ(t)\)
+        \item for all \((b, s', p) ∈ δ(s)\), \(b ∧ \)
+    \end{itemize}
+\end{theorem}
+
 \end{document}

### 2024-09-23 main.tex
add main

% Please compile this document using LuaLaTeX.
% XeLaTeX and PDFLaTeX will not work.

\newif\iffull\fulltrue

\documentclass{extarticle}

%%%%% PACKAGES
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage[margin=2.5cm]{geometry}

% Biblatex
\usepackage[style=numeric]{biblatex}
\addbibresource{refs.bib}

% For adding inline comments in the text.
\usepackage[margin=false,inline=true]{fixme}
\FXRegisterAuthor{aaa}{anaaa}{\color{cyan}AAA}
\FXRegisterAuthor{mg}{anmg}{\color{red}MG}
\FXRegisterAuthor{cz}{ancz}{\color{orange}CZ}
% \newcommand{\aaa}[1]{\aaanote{#1}}
% \newcommand{\mg}[1]{\mgnote{#1}}
% \newcommand{\cz}[1]{\cznote{#1}}
\newcommand{\aaa}[1]{}
\newcommand{\mg}[1]{}
\newcommand{\cz}[1]{}

\usepackage{stmaryrd}

\usepackage{stackengine}
\usepackage{mathrsfs}
\usepackage{braket}
\usepackage{annotate-equations}
\usepackage{scalerel}

% graphs
\usepackage{tikz}
\usetikzlibrary{arrows,automata,positioning}

% commutative diagram
\usepackage{tikz-cd}

% ref
\usepackage{hyperref}
\usepackage{cleveref}

% inference rule
\usepackage{mathpartir}
% cross-referencing infer rule
% based on https://tex.stackexchange.com/questions/340788/cross-referencing-inference-rules
\makeatletter
\let\originferrule\inferrule
\DeclareDocumentCommand \inferrule { s O {} m m}{%
  \IfBooleanTF{#1}%
  {%
    \mpr@inferstar[#2]{#3}{#4}%
  }{%
    \mpr@inferrule[#2]{#3}{#4}%
  }%
  \IfValueT{#2}%
  {%
    \my@name@inferrule{#2}%
  }%
}
\NewDocumentCommand \my@name@inferrule { m }{%
  \def\@currentlabelname{\textsc{#1}}%
}
\makeatother

% item spacing
\usepackage{enumitem}

% for code
\usepackage{listings}
\lstset{basicstyle=\footnotesize\ttfamily,breaklines=true}
\lstset{framextopmargin=50pt,frame=bottomline}
\lstset{language=caml, escapeinside={[*}{*]}}

% for algorithm
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
% switch statement for pattern matching
\algnewcommand\algorithmicmatch{\textbf{match}}
\algnewcommand\algorithmicwith{\textbf{with}}
\algnewcommand\algorithmiccase{\textbf{case}}
\algnewcommand\algorithmicdefault{\textbf{default}}
\algnewcommand\Skip{\textbf{skip}}
\algdef{SE}[MATCH]{Match}{EndMatch}[1]{\algorithmicmatch\ #1\ \algorithmicwith}{\algorithmicend\ \algorithmicmatch}%
\algdef{SE}[CASE]{Case}{EndCase}[1]{\algorithmiccase\ #1 \algorithmicthen}{\algorithmicend\ \algorithmiccase}%
\algdef{SE}[DEFAULT]{Default}{EndDefault}{\algorithmicdefault\ }{\algorithmicend\ \algorithmicdefault}%
\algtext*{EndMatch}%
\algtext*{EndCase}%
\algtext*{EndDefault}%

% for better table
\usepackage{booktabs}

% subcaption
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption

% unicode math symbols
\usepackage{unicode-math}
% support for hat, overline, underline, vec, and sim combining charactors
\protected\def\afteracc{\directlua{
    local nest = tex.nest[tex.nest.ptr]
    local last = nest.tail
    if not (last and last.id == 18) then
      error'I can only put accents on simple noads.'
    end
    if last.sub or last.sup then
      error'If you want accents on a superscript or subscript, please use braces.'
    end
    local acc = node.new(21, 1)
    acc.nucleus = last.nucleus
    last.nucleus = nil
    local is_bottom = token.scan_keyword'bot' and 'bot_accent' or 'accent'
    acc[is_bottom] = node.new(23)
    acc[is_bottom].fam, acc[is_bottom].char = 0, token.scan_int()
    nest.head = node.insert_after(node.remove(nest.head, last), nil, acc)
    nest.tail = acc
    node.flush_node(last)
  }}
\AtBeginDocument{
\begingroup
  \def\UnicodeMathSymbol#1#2#3#4{%
    \ifx#3\mathaccent
      \def\mytmpmacro{\afteracc#1 }%
      \global\letcharcode#1=\mytmpmacro
      \global\mathcode#1="8000
    \else\ifx#3\mathbotaccentwide
      \def\mytmpmacro{\afteracc bot#1 }%
      \global\letcharcode#1=\mytmpmacro
      \global\mathcode#1="8000
    \fi\fi
  }
  \input{unicode-math-table}
\endgroup
}

% math font, this is needed to render \setminus command
\setmathfont{latinmodern-math}
\setmathfont[range=\setminus]{STIX Two Math}
\setmathfont[range=\similarrightarrow]{STIX Two Math}

\newtheorem{theorem}{Theorem}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}

\newtheorem{definition}{Definition}
\newtheorem*{definition*}{Definition}
\newtheorem{example}[definition]{Example}
\newtheorem{remark}[definition]{Remark}

%%%% Macros %%%%%

% Math?
\newcommand{\true}{\mathrm{true}}
\newcommand{\false}{\mathrm{false}}
\newcommand{\At}{\mathbf{At}}

% operators
\newcommand{\dom}[1]{\mathrm{dom}(#1)}
\newcommand{\cod}[1]{\mathrm{cod}(#1)}
\DeclareMathOperator{\post}{\mathrm{post}}
\newcommand{\reject}{\mathinner{\mathrm{rej}}}
\newcommand{\accept}{\mathinner{\mathrm{acc}}}
\DeclareMathOperator*{\bigplus}{\scalerel*{+}{\sum}}
\newcommand{\clos}[1]{\mathrel{\overline{#1}}}
\DeclareMathOperator{\norm}{\mathrm{norm}}
\DeclareMathOperator{\dead}{\mathrm{dead}}
\DeclareMathOperator{\symb}{\mathrm{symb}}
\DeclareMathOperator{\unsymb}{\symb^{-1}}

% commands 
\newcommand{\command}[1]{{\mathtt{#1}}}
\newcommand{\comAssume}[1]{\command{assume}~#1}
\newcommand{\comITE}[3]{\command{if}~#1~\command{then}~#2~\command{else}~#3}
\newcommand{\comWhile}[2]{\command{while}~#1~\command{do}~#2}

% set of models
\newcommand{\theoryOf}[1]{\ensuremath{\mathsf{#1}}}
\newcommand{\Exp}{\theoryOf{Exp}}
\newcommand{\GKAT}{\theoryOf{GKAT}}
\newcommand{\Bool}{\theoryOf{Bool}}
\DeclareMathOperator{\GS}{\mathrm{GS}}

\newcommand\altxrightarrow[2][0pt]{\mathrel{\ensurestackMath{\stackengine%
  {\dimexpr#1-7.5pt}{\xrightarrow{\phantom{#2}}}{\scriptstyle\!#2\,}%
  {O}{c}{F}{F}{S}}}}
\newcommand{\transvia}[1]{
    \mathrel{\raisebox{-2px}{\(\altxrightarrow[-2px]{#1}\)}}
}

\begin{document}

%% Title information
\title{One-The-Fly Decision Procedures For GKAT}         
% \title{Incorrectness Logic and Kleene Algebra With Top and Test}         
                                        %% [Short Title] is optional;
                                        %% when present, will be used in
                                        %% header instead of Full Title.
% \titlenote{}                            %% \titlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'
%% \subtitle is optional

% \subtitlenote{with subtitle note}       %% \subtitlenote is optional;
%                                         %% can be repeated if necessary;
%                                         %% contents suppressed with 'anonymous'

%% Author information
%% Contents and number of authors suppressed with 'anonymous'.
%% Each author should be introduced by \author, followed by
%% \authornote (optional), \orcid (optional), \affiliation, and
%% \email.
%% An author may have multiple affiliations and/or emails; repeat the
%% appropriate command.
%% Many elements are not rendered, but should be provided for metadata
%% extraction tools.
\author{
    Cheng Zhang \and 
    Hang Ji \and 
    Ines Santacruz \and 
    Marco Gaboardi
}
\date{}

% In non-ACM format \maketitle needs to happen before abstract
\maketitle

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
\end{abstract}

\section{Introduction}

\paragraph{Notation: } In this paper, we will use un-curried notation to apply curried functions, for example, given a function \(δ: X → Y → Z\), we will write the function applications as follow \(δ(x): Y → Z\) and \(δ(x, y): Z\). And when drawing commutative diagram, we will leave function restriction implicit. Specifically given \(A' ⊆ A\), and a function \(h: A → B\), we will draw:
\[
    \begin{tikzcd}
        A' \ar{r}{h} & B
    \end{tikzcd}
\]
where the function \(h\) is implicitly restricted to \(A'\).
For bifunctors like \((-) × (-)\), we will write function lifting by applying the bifunctors on these functions: for example, given \(h₁: A₁ → B₁\) and \(h₂: A₂ → B₂\), we will use \[h₁ × h₂: A₁ × A₂ → B₁ × B₂\] to denote the bifunctorial lift of \(h₁\) and \(h₂\) via product \((-) × (-)\).

\section{Preliminary}

\subsection{Concepts in Universal Coalgebra}

Our notion of normalized bisimulation is inherently coalgebraic, thus it is empirical for us to recall some notions and theorems in coalgebra.
Given a functor \(F\) on the category of set and functions, a \emph{coalgebra over \(F\)} or \emph{\(F\)-coalgebra} consists of a set \(S\) and a function \(δ_S: S → F(S)\).
We typically call \(F\) the \emph{signature} of the coalgebra, \(S\) the states of the coalgebra, and \(δ_S\) the transition function of the coalgebra.
We will sometimes use the states \(S\) to denote the coalgebra, when no ambiguity can arise.

A homomorphism between two \(F\)-coalgebra \(S\) and \(T\) is a map \(h: S → T\) that preserves the transition function; diagrammatical, the following diagram commutes:
\[
    \begin{tikzcd}
        S \ar{r}{h} \ar[swap]{d}{δ_S} & T \ar{d}{δ_T} \\  
        F(S) \ar{r}{F(h)} & F(T)
    \end{tikzcd}    
\]

When we can restrict the homomorphism map into a inclusion map \(i: S' → S\) for \(S' ⊆ S\), then we say that \(S'\) is a \emph{sub-coalgebra} of \(S\), denoted as \(S' ⊑ S\). Specifically, the following diagram commutes when \(S' ⊑ S\):
\[
    \begin{tikzcd}
        S' \ar[hook]{r}{i} \ar[swap]{d}{δ_{S'}} & S \ar{d}{δ_S} \\  
        F(S') \ar[hook]{r}{F(i)} & F(S)
    \end{tikzcd}    
\]
In fact, the transition function \(δ_{S'}\) is uniquely determined by the states \(S'\)~\cite[Proposition 6.1]{rutten_UniversalCoalgebraTheory_2000}.

The sub-coalgebras are preserved under homomorphic images and pre-images: 
\begin{lemma}[Theorem 6.3~\cite{rutten_UniversalCoalgebraTheory_2000}]\label{thm:hom-(pre)img-preserve-sub-coalg}
    given a homomorphism \(h: S → T\), and sub-coalgebras \(S' ⊑ S\) and \(T' ⊑ T\), then 
    \[h(S') ⊑ T \text{ and } h^{-1}(T') ⊑ S.\]
\end{lemma}

One particularly important sub-coalgebra of and coalgebra \(S\) is the least coalgebra generated by a single element \(s\), we will typically denote it as \(⟨s⟩_{S}\), and call it \emph{principle sub-coalgebra} generated by \(s\). We will omit the subscript \(S\) when it can be inferred from context or irrelevant.
As we will see later, principle sub-coalgebra \(⟨s⟩_S\) is important because it models all the reachable state of \(s\) in \(S\).

Principle sub-coalgebra always exists, because sub-coalgebra of any coalgebra forms a complete lattice~\cite[theorem 6.4]{rutten_UniversalCoalgebraTheory_2000}.
And similar to sub-coalgebra, principle sub-coalgebra is also preserved under homomorphic image:
\begin{theorem}\label{thm:homo-img-preserve-principle-sub-coalg}
    Homomorphic image preserves principle sub-GKAT coalgebra. Specifically, given a homomorphism \(h: S → T\):
    \[h(⟨S⟩_S) = ⟨h(s)⟩_T\]
\end{theorem}

\begin{proof}
    We will need to show that \(h(⟨s⟩_{S})\) is the smallest sub-GKAT coalgebra of \(S\) that contain \(h(s)\). First by definition of image, \(h(s) ∈ h(⟨s⟩_{S})\); second by \cref{thm:hom-(pre)img-preserve-sub-coalg}, \(h(⟨S⟩_S) ⊑ T\).

    Finally, take any \(T' ⊑ T\) and \(h(s) ∈ T'\), recall that by \cref{thm:hom-(pre)img-preserve-sub-coalg}, \(h^{-1}(T') ⊑ S\). We can then derive that \(h(⟨s⟩_S) ⊑ T'\): 
    \begin{align*}
        h(s) ∈ T' 
        & ⟹ s ∈ h^{-1}(T') \\  
        & ⟹ ⟨s⟩_S ⊑ h^{-1}(T') & \text{definition of \(⟨s⟩_S\)}\\  
        & ⟹ h(⟨s⟩_S) ⊑ T' & \text{take the image of \(h\) and \cref{thm:hom-(pre)img-preserve-sub-coalg}}
    \end{align*}
    Hence \(h(⟨s⟩_S)\) is the smallest sub-GKAT coalgebra of \(T\) that contains \(h(s)\).
\end{proof}

% bisimulation

A \emph{final coalgebra} \(ℱ\) over a signature \(F\), sometimes called the \emph{behavior} of coalgebras over \(F\), is a \(F\)-coalgebra s.t. for all \(F\)-coalgebra \(S\), there exists a unique homomorphism \(⟦-⟧_S: S → ℱ\).

Given two \(F\)-coalgebra \(S\) and \(T\), the \emph{behavioral equivalence} between states in \(S\) and \(T\) can be computed by a notion called \emph{bisimulation}.
A relation \({∼} ⊆ S × T\) is called a \emph{bisimulation relation} if it forms a \(F\)-coalgebra: \[δ_{∼}: {∼} → F(∼),\] 
And its projection functions \(π₁: S × T → S\) and \(π₂: S × T → T\) are both homomorphisms:
\[
    \begin{tikzcd}
        S \ar[swap]{d}{δ_S}
            & ∼ \ar[swap]{l}{π₁} \ar{r}{π₂} \ar{d}{δ_{∼}} 
            & T \ar{d}{δ_T}\\  
        F(S) & F(∼) \ar{l}{F(π₁)} \ar[swap]{r}{F(π₂)} & F(T)
    \end{tikzcd}
\]
As we will see later, bisimulation or variation of bisimulation indeed correspond to important semantical equivalences.

\subsection{Guarded Kleene Algebra With Tests}

\subsection{GKAT Coalgebra}

% definition
GKAT coalgebra~\cite{smolka_GuardedKleeneAlgebra_2020,schmid_GuardedKleeneAlgebra_2021}, is a coalgebraic systems for GKAT.
Specifically GKAT coalgebras over a alphabet \(K, B\) are coalgebras over the following functor:
\[G(S) ≜ (2 + S × K)^{\At_B},\] 
where \(2 ≜ \{\accept, \reject\}\). Intuitively given a state \(s ∈ S\) and an atom \(α ∈ \At\), \(δ(s, α)\) will deterministically execute one of the following: reject \(α\), denoted as \(δ(s, α) = \reject\); accept \(α\), denoted \(δ(s, α) = \accept\); or transition to a state \(s' ∈ S\) and execute action \(p ∈ K\), denoted as \(δ(s, α) = (s', p)\).

This deterministic behavior contrast that of Kleene coalgebra with tests~\cite{kozen_CoalgebraicTheoryKleene_2017}, where for each atom, the state can accept or  reject the atom (but not both), yet the state can also non-deterministically transition to multiple different state via the same atom, while executing different actions.
% TODO: I think this belongs in the intro
As we will see later, the deterministic behavior of GKAT coalgebra not only enables a further optimized symbolic algorithm than KCT~\cite{pous_SymbolicAlgorithmsLanguage_2015}, but also present challenges. Specifically, GKAT coalgebra requires normalization to compute finite trace equivalences~\cite{smolka_GuardedKleeneAlgebra_2020}, where we will remove all the state that cannot lead to acceptance. 
It seems like we need to traverse the entire automaton to identify these ``dead states'', however we have shown that these dead state detection can be invoked lazily, only when discrepancy between bisimulation are found.

\subsection{Liveness and Sub-GKAT coalgebras}

% liveness

Traditionally, live and dead states are defined by whether they can reach an accepting state~\cite{smolka_GuardedKleeneAlgebra_2020}. However, recall that principle sub-coalgebra \(⟨s⟩_S\) models reachable states of \(s\) in coalgebra \(S\). Thus the classical definition is equivalent to the following:
\begin{definition}[liveness of states]\label{def:liveness-of-states}
    A state \(s\) is \emph{accepting} if there exists a \(α ∈ \At\) s.t. \(δ(s, α) = \accept\). A state \(s'\) is \emph{live} if there exists an accepting state \(s' ∈ ⟨s⟩\). A state \(s'\) is \emph{dead} if there is no accepting state in \(⟨s⟩\).
\end{definition}
This alternative liveness definition can help us formally prove important theorems regarding reachability and liveness without performing induction on traces. We can show the following lemmas as examples:
\begin{lemma}\label{thm:dead-iff-all-reachable-dead}
    A state \(s\) is dead if and only if all elements in \(⟨s⟩\) is dead.
\end{lemma}
\begin{proof}
    \(⟸\) direction is true, because \(s ∈ ⟨s⟩\): if all \(⟨s⟩\) is dead, then \(s\) is dead. 
    \(⟹\) direction can be proven as follows.
    Take \(s' ∈ ⟨s⟩\), then \(⟨s'⟩ ⊑ ⟨s⟩\) by definition. 
    Since there is no accepting state in \(⟨s⟩\), thus there cannot be any accepting state in \(⟨s'⟩\), hence \(⟨s'⟩\) is also dead.
\end{proof}

\begin{theorem}[homomorphism perserves liveness]\label{thm:hom-preserve-liveness}
    Given a homomorphism \(h: S → T\) and a state \(s ∈ S\):
    \[\text{\(s\) is live} ⟺ \text{\(h(s)\) is live}\]
\end{theorem}

\begin{proof}
    Because homomorphic image preserves principle sub-GKAT coalgebra~\cref{thm:homo-img-preserve-principle-sub-coalg}
    \[h(⟨s⟩_S) = ⟨h(s)⟩_T;\]
    therefore for any state \(s' ∈ S\):
    \[s' ∈ ⟨s⟩_S ⟺ h(s') ∈ h(⟨s⟩_S) ⟺ h(s') ∈ ⟨h(s)⟩_T.\]
    And because \(s'\) is accepting if and only if \(h(s')\) accepting by definition of homomorphism; then \(⟨s⟩_S\) contains an accepting state if and only if \(h(⟨s⟩_S) = ⟨h(s)⟩_T\) contains an accepting state. 
    Therefore \(s\) is live in \(S\) if and only if \(h(s)\) is live in \(T\).
\end{proof}

\begin{corollary}[sub-coalgebra perserves liveness]\label{thm:sub-coalg-preserve-liveness}
    Given a sub-coalgebra \(S' ⊑ S\), then for all states \(s ∈ S'\),
    \[\text{\(s\) is live in \(S'\)} ⟺ \text{\(s\) is live in \(S\)}.\]
\end{corollary}

\begin{proof}
    take the homomorphism \(h\) in \cref{thm:hom-preserve-liveness} to be the inclusion homomorphism \(i: S' → S\).
\end{proof}

\begin{corollary}[bisimulation preserves liveness]\label{thm:bisim-preserve-liveness}
    If there exists a bisimulation \(∼\) between GKAT coalgebra \(S\) and \(T\) s.t. \(s ∼ t\) for some states \(s ∈ S\) and \(t ∈ T\), then \(s\) and \(t\) has to be either both accepting, both live or both dead.
\end{corollary}

\begin{proof}
    Because for a \(∼\) is a bisimulation when both \(π₁: {∼} → S\) and \(π₂: {∼} → T\) are homomorphisms.
    Therefore, 
    \begin{align*}
        s \text{ is live in } S 
        & ⟺ π₁((s, t)) \text{ is live in } S 
            & π₁((s, t)) = s\\  
        & ⟺ (s, t) \text{ is live in } {∼} 
            & \text{\(π₁\) is a homomorphism}\\  
        & ⟺ π₂((s, t)) \text{ is live in } T 
            & \text{\(π₂\) is a homomorphism} \\
        & ⟺ t \text{ is live in } T 
            & π₂((s, t)) = t
    \end{align*}
\end{proof}

% The signature of GKAT coalgebra \(G\) is a simple polynomial functor~\cite[Definition 2.2.1]{jacobs_IntroductionCoalgebraMathematics_2016}, implying that it preserves the inclusion function:
% \begin{lemma}%used in the comments below
%     Given a inclusion \(i: S' → S\) for \(S' ⊆ S\), then \(G(i): G(S') → G(S)\) is also a inclusion in the sense that:
%     \[∀ m ∈ G(S'), G(i)(m) = m.\]
% \end{lemma}
% This implies that for a sub-GKAT coalgebra \(S' ⊑ S\), then the restriction of transition function will satisfy the the diagram for sub-coalgebra.
% \[
%     \begin{tikzcd}
%         S' \ar{d}{δ_S} \ar[hook]{r}{i} & S \ar{d}{δ_S} \\  
%         G(S') \ar[hook]{r}{G(i)} & G(S)
%     \end{tikzcd}
% \]
% Then by the uniqueness of transition for sub-coalgebra~\cite[Proposition 6.1]{rutten_UniversalCoalgebraTheory_2000}, the transition of \(S'\) is necessarily the restriction of \(δ_S\). A nice consequence of this observation is that the ordering between sub-GKAT coalgebra is exactly the subset ordering.
% \begin{lemma}[ordering in sub-GKAT coalgebra]\label{thm:sub-GKAT coalgebra-is-subset}
%     Given two sub-GKAT coalgebra \(S₁ ⊑ S\) and \(S₂ ⊑ S\), then 
%     \[S₁ ⊑ S₂ ⟺ S₁ ⊆ S₂.\]
% \end{lemma}

% \begin{proof}
%     \(⟹\) direction is true by definition, and \(⟸\) direction is true because both \(S₁\) and \(S₂\) are sub-GKAT coalgebra, thus their transition functions are both restrictions of \(δ_S\). Hence the commutativity of the following diagram can be verified by computation:
%     \[
%     \begin{tikzcd}
%         S₁ \ar[hook]{r}{i} \ar{d}{δ_S} & S₂ \ar[hook]{r}{i} \ar{d}{δ_S} & S \ar{d}{δ} \\  
%         G(S₁) \ar[hook]{r}{G(i)} & G(S₂) \ar[hook]{r}{G(i)} & G(S)
%     \end{tikzcd}
% \]
% \end{proof}

% trace semantics
\subsection{Normalization And Trace Semantics}

(Possibly infinite) trace model \(𝒢_ω\) is the natural semantics of states in a GKAT coalgebra, specifically it forms the final coalgebra of GKAT coalgebras~\cite{schmid_GuardedKleeneAlgebra_2021}.
The finality of the model means that every state in any GKAT coalgebra \(S\) can be assigned a semantics under the unique homomorphism \(⟦-⟧^{ω}_{S}: S → 𝒢_ω\); and such semantical equivalences can indeed be identified by bisimulation~\cite{schmid_GuardedKleeneAlgebra_2021}:
\[⟦s⟧^{ω}_{S} = ⟦t⟧^{ω}_{T} ⟺ \text{exists a bisimulation \({∼} ⊆ S × T\), s.t. \(s ∼ t\)}.\]

The infinite trace equivalences is relatively easy to compute efficiently, as bisimulation is, in general, compatible with derivative based computation on-the-fly algorithm~\cite{kozen_CoalgebraicTheoryKleene_2017,almeida_DecidingKATHoare_2012,pous_SymbolicAlgorithmsLanguage_2015}. 
However, the \emph{finite} trace model \(𝒢\) is only the final coalgebra of GKAT coalgebras without dead states, which we call \emph{normal GKAT coalgebra}~\cite{smolka_GuardedKleeneAlgebra_2020}. Fortunately every GKAT coalgebra can be normalized by rerouting all the transition from dead states to rejection
\begin{align*}
    \norm(δ_S) & : S → G(S) \\
    \norm(δ_S)(s, α) & ≜ \begin{cases}
        \reject & \text{if } δ_S(s, α) = (s', p) \text{ and \(s'\) is dead}\\
        δ_S(s, α) & \text{otherwise}
    \end{cases}
\end{align*}
We denote the normalized coalgebra \((S, \norm(δ_S))\) as \(\norm(S)\).
This means that the finite trace semantics \(⟦-⟧\) is the unique coalgebra homomorphism \(\norm(S) → \norm(𝒢)\). Hence the finite trace equivalence between \(s ∈ S\) and \(t ∈ T\) can be computed by first normalizing \(S\) and \(T\), then decide whether there is a bisimulation on \(\norm(S)\) and \(\norm(T)\) that includes \((s, t)\).
For a more intuitive account for the trace semantics, we refer the reader to the work of \Citeauthor{smolka_GuardedKleeneAlgebra_2020}~\cite{smolka_GuardedKleeneAlgebra_2020}.

Then by the finality of GKAT colagebra, the soundness and completeness of bisimulation is a simple corollary in universal coalgebra~\cite{smolka_GuardedKleeneAlgebra_2020,jacobs_IntroductionCoalgebraMathematics_2016,rutten_UniversalCoalgebraTheory_2000}. No matter through explicit construction~\cite{smolka_GuardedKleeneAlgebra_2020} or (weak) pullback~\cite{jacobs_IntroductionCoalgebraMathematics_2016,rutten_UniversalCoalgebraTheory_2000}, both completeness proof shows that the language equivalence: \[{≡} ≜ \{(s,t) ∣ ⟦s⟧_S = ⟦t⟧_T\}\] is indeed a bisimulation. 
This result, together with soundness result \(s ∼ t ⟹ ⟦s⟧_S = ⟦t⟧_T\) means that the language equivalence \(≡\) is indeed the largest bisimulation, which allows us to work with bisimulation equivalence instead of bisimulation.

\begin{definition}
    A \emph{bisimulation equivalence} in \(S\) is a bisimulation between \(S\) and itself, and it is also an equivalence relation.
\end{definition}

\begin{theorem}\label{thm:bisim-iff-bisim-equiv}
    Given two states \(s, t ∈ S\), then there exists a bisimulation \(∼\) s.t. \(s ∼ t\) if and only if there exists a bisimulation equivalence \(≃\) s.t. \(s ≃ t\).
\end{theorem}

\begin{proof}
    The \(⟹\) direction can just take \(≃\) to be the language equivalence \(≡\), which is a bisimulation equivalence, and because \(≡\) is maximal, therefore \({∼} ⊑ {≡}\), and \((s, t) ∈ {∼} ⊆ {≡}\).

    The \(⟸\) direction is true because all bisimulation is a bisimulation equivalence, thus we can take \(∼\) to just be the given bisimulation equivalence \(≃\).
\end{proof}

% Although \(\norm(S)\) might not be a homomorphic image of \(S\), normalization does satisfy many nice properties of homomorphic images. For example, it is monotonic and it preserves principle sub-GKAT coalgebra, like in \cref{thm:homo-img-preserve-principle-sub-coalg} 
% \begin{lemma}[monotonicity]\label{thm:monotonicity-norm}
%     For all GKAT coalgebra \(S\) and \(T\):
%     \[S ⊑ T ⟹ \norm(S) ⊑ \norm(T).\]
% \end{lemma}

% \begin{theorem}[\(\norm\) preserves principle sub coalgebra]\label{thm:norm-preserve-principle-sub-coalg}
%     Given a GKAT coalgebra \(S\) and a live state \(s ∈ S\):
%     \[⟨s⟩_{\norm(S)} = \norm(⟨s⟩_{S}).\]
% \end{theorem}

% \begin{proof}
%     We need to show \(\norm(⟨s⟩_{S})\) is the smallest sub-GKAT coalgebra in \(\norm(S)\) that contains \(s\).

%     First, \(\norm(⟨s⟩_{S})\) contains \(s\) because \(s\) is live. 
%     Second, take any \(T ⊑ \norm(S)\), we will use \(⟨T⟩_S\) to denote the smallest sub-GKAT coalgebra in \(S\) that contains all the states in \(T\). 
%     Since \(T\) is a sub-GKAT coalgebra of \(\norm(S)\), \(⟨T⟩_S ∖ T\) can only contain dead states, thus \(\norm(⟨T⟩_S) = T\).
%     \begin{align*}
%         & ⟨s⟩_{S} ⊑ ⟨T⟩_S & \text{\(⟨T⟩_S\) contains \(s\)} \\  
%         ⟹ {}& \norm(⟨s⟩_{S}) ⊑ \norm(⟨T⟩_S) & \text{monotonicity: \cref{thm:monotonicity-norm}} \\  
%         ⟹ {}& \norm(⟨s⟩_{S}) ⊑ T & \norm(⟨T⟩_S) = T
%     \end{align*}
% \end{proof}

\section{On-The-Fly Bisimulation}

The original algorithm for deciding GKAT equivalences~\cite{smolka_GuardedKleeneAlgebra_2020} requires the entire automaton to be known prior to the execution of the bisimulation algorithm; specifically, in order to compute the liveness of a state \(s\), it is necessary iterate through all its reachable states \(⟨s⟩\) to see if there are any accepting states within.
This limitation poses challenges to design an efficient on-the-fly algorithm for GKAT.
In order to make the decision procedure scalable, we will need to merge the normalization and bisimulation procedure, so that our algorithm can normalized the automaton only when we need to.

In this section, we introduce an algorithm that merges bisimulation and normalization where we only need to test the liveness of the state when a disparity in the bisimulation has been found.
For example, when one automaton leads to reject where the other transition to a state, then we will need to verify whether that state is dead or not.

This on-the-fly algorithm inherits the efficiency of the original algorithm~\cite{smolka_GuardedKleeneAlgebra_2020}, where the worst case will require two passes of the automaton, where one pass will try to establish a bisimulation, when failed the other pass will kick in and compute whether the failed states are dead.
In some special case, the on-the-fly algorithm can even out perform the original algorithm; for example, when the two input automata are bisimular (even when they are not normal), the on-the-fly algorithm can skip the liveness checking, only performing the bisimulation.

\begin{theorem}[sub-coalgebra perserve bisimulation]\label{thm:sub-coalg-preserve-bisim}
    Given any sub-coalgebra \(S' ⊑ S\) and \(T' ⊑ T\),
    \begin{itemize}
        \item Given a bisimulation \(∼\) between \(S'\) and \(T'\), then \(∼\) is also a bisimulation between \(S\) and \(T\);
        \item if there exists a bisimulation \(∼\) between \(S\) and \(T\), then the restriction 
        \begin{mathpar}
            ∼_{S', T'} ≜ \{(s, t) ∣ s ∈ S', t ∈ T', s ∼ t\}
        \end{mathpar}
        forms a bisimulation between \(S'\) and \(T'\).
    \end{itemize}
\end{theorem}

\begin{proof}
    To prove that bisimulation \(∼\) between \(S'\) and \(T'\) is also a bisimulation of \(S\) and \(T\), we can simply enlarge the diagram by the inclusion homomorphism
    \[
        \begin{tikzcd}
            S \ar{d}{δ_S} & S' \ar[hook',swap]{l}{i} \ar{d}{δ_S}
            & ∼ \ar[swap]{l}{π₁} \ar{r}{π₂} \ar{d}{δ_∼}
            & T' \ar[hook]{r}{i} \ar{d}{δ_S} & T \ar{d}{δ_T}\\  
            G(S) & G(S') \ar[hook',swap]{l}{G(i)} 
            & G(∼) \ar[swap]{l}{G(π₁)} \ar{r}{G(π₂)} & T' \ar[hook]{r}{G(i)} & T \\  
        \end{tikzcd}
    \]
    Because the inclusion homomorphism \(i\) doesn't change the input thus, we have:
    \begin{mathpar}
        {∼} \xrightarrow{π₁} S' \xrightarrow{i} S = {∼} \xrightarrow{π₁} S \and 
        {∼} \xrightarrow{π₂} T' \xrightarrow{i} T = {∼} \xrightarrow{π₂} T
    \end{mathpar}

    To prove that the bisimulation can be restricted, we first realize that \(∼_{S', T'}\) is a pre-image of the maximal bisimulation \(≣_{S', T'}\) along the inclusion homomorphism \(i: {∼} → {≡_{S, T}}\).
    This means that \(∼_{S', T'}\) can be formed by a pullback square:
    \[
        \begin{tikzcd}
            ∼_{S', T'} \ar{r}{i} \ar[swap]{d}{i} \ar[phantom, very near start]{dr}{\scalebox{1.5}{\(\lrcorner\)}} & ≣_{S', T'} \ar{d}{i}\\ 
            {∼} \ar[swap]{r}{i} & {≡_{S, T}}
        \end{tikzcd}
    \]
    Recall that elementary polynomial functor~\cite{jacobs_IntroductionCoalgebraMathematics_2016} like \(G\) preserves pullback, hence the pullback also uniquely generates a GKAT coalgebra~\cite{rutten_UniversalCoalgebraTheory_2000}
\end{proof}

\begin{lemma}[bisimulation between dead states]\label{thm:bisim-between-dead}
    Given two dead states \(s ∈ S\) and \(t ∈ T\), then the singleton bisimulation 
    \begin{mathpar}
        {∼} ≜ \{(s,t)\} \and 
        δ_{∼}((s,t), α) ≜ \reject
    \end{mathpar}
    is a bisimulation between \(S\) and \(T\).
\end{lemma}

\begin{proof}
    By computation
\end{proof}

\begin{theorem}[inductive construction]\label{thm:inductive-construction}
    Given two GKAT coalgebra \(S\) and \(T\), and two of their elements \(s ∈ S\) and \(t ∈ T\),
    there exists a bisimulation \({∼} ⊆ ⟨s⟩ × ⟨t⟩\) s.t. \(s ∼ t\), if and only if all of the following holds:
    \begin{enumerate}
        \item\label{itm:acc-condition} for all \(α ∈ \At\), \(δ_{S}(s, α) = \accept ⟺ δ_{T}(t, α) = \accept\);
        \item\label{itm:transition-bisim} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{T}(t, α) = (t', p)\), then there exists a bisimulation \({∼_{s',t'}}\) on \(⟨s'⟩\) and \(⟨t'⟩\), s.t. \(s' ∼_{s',t'} t'\);
        \item\label{itm:transition-dead} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{T}(t, α) = (t', q)\), s.t. \(p ≠ q\), then both \(s'\) and \(t'\) are dead;
        \item\label{itm:rej-or-dead} \(s\) reject \(α\) or transition to a dead state via \(α\) if and only if \(t\) rejects \(α\) or transition to a dead state via \(α\).
    \end{enumerate}
\end{theorem}

\begin{proof}
    We first prove \(⟹\) direction, recall the definition of bisimulation:
    \[
        \begin{tikzcd}
            S \ar[swap]{d}{\norm(δ_S)} 
            & ∼ \ar{d}{δ_∼} \ar[swap]{l}{π₁} \ar{r}{π₂}
            & T \ar{d}{\norm(δ_T)}\\  
            G(S) 
            & G(∼) \ar{l}{G(π₁)} \ar[swap]{r}{G(π₂)} 
            & G(T) 
        \end{tikzcd}
    \]

    The condition \labelcref{itm:acc-condition} holds:
    \begin{align*}
        δ_S(s, α) = \accept 
        & ⟺ \norm(δ_S)(s, α) = \accept 
        ⟺ \norm(δ_∼)((s,t), α) = \accept \\ 
        & ⟺ \norm(δ_T)(t, α) = \accept 
        ⟺ δ_T(t, α) = \accept
    \end{align*}

    The condition~\labelcref{itm:transition-bisim} holds, by case analysis on the liveness of \(s'\) and \(t'\).
    First note that \(s'\) and \(t'\) has to be both live or both dead: because \(δ_S(s, α) = (s', p)\), then \(\norm(δ_S)(s', α)\) can either be rejection or \((s',p)\), and so is \(\norm(δ_T)(t', α)\). Finally because \(s ∼ t\), then
    \[s' \text{ is live} ⟺ \norm(δ_S)(s, α) = (s', p) ⟺ \norm(δ_T)(t, α) = (t', p) ⟺ t' \text{ is live}.\]
    \begin{itemize}
        \item If both \(s'\) and \(t'\) are live, then \(s' ∼ t'\), the bisimulation \(∼_{s', t'}\) is just \(∼\) restricted to \(⟨s'⟩\) and \(⟨t'⟩\).
        \item If both \(s'\) and \(t'\) are dead, then \(∼_{s', t'}\) can just be the singleton relation, according to~\cref{thm:bisim-between-dead}.
    \end{itemize}

    The condition~\labelcref{itm:transition-dead} holds: by the proof of condition~\labelcref{itm:transition-bisim}, \(s'\) and \(t'\) has to be either both live or both dead; if they are both live, then there cannot be a element in \(G(∼)\) that can project to \((s', p)\) under \(π₁\) but projects to \((t', q)\) under \(π₂\). Thus both \(s'\) and \(t'\) has to be dead.

    The condition~\labelcref{itm:rej-or-dead} holds: 
    \begin{align*}
        δ_S(s, α) \text{ rejects or transition to dead states} 
    & ⟺ \norm(δ_S)(s, α) = \reject \\
    & ⟺ \norm(δ_T)(t, α) = \reject \\
    & ⟺ δ_T(t, α) \text{ rejects or transition to dead states}.
    \end{align*}

    We then show the \(⟸\) direction, we use \(≡_{s', t'}\) to denote the maximal bisimulation between \(⟨s'⟩\) and \(⟨t'⟩\).
    \[{∼'} ≜ ⋃ \{≡_{s', t'} ∣ ∃ α ∈ \At, p ∈ K, δ_S(s, α) = (s', p) \text{ and } δ_T(t, α) = (t', p)\}.\]
    Notice because \(⟨s'⟩ ⊑ ⟨s⟩\) and \(⟨t'⟩ ⊑ ⟨t⟩\), then \(≡_{s', t'}\) is a bisimulation between \(⟨s⟩\) and \(⟨t⟩\), and because bisimulation is closed under arbitrary union~\cite{rutten_UniversalCoalgebraTheory_2000}, then \(∼'\) is a bisimulation between \(⟨s⟩\) and \(⟨t⟩\).

    We then augment \(∼'\) with \((s, t)\) to obtain the bisimulation we required:
    \begin{mathpar}
        {∼} ≜ {∼'} ∪ \{(s, t)\} \and 
        δ_{∼}((s₁, t₁), α) ≜ \begin{cases}
            δ_{∼'}((s,t), α) & (s,t) ≠ (s₁, t₁) \\  
            \accept & \norm(δ_S)(s, α) = \norm(δ_T)(s, α) = \accept \\  
            \reject & \norm(δ_S)(s, α) = \norm(δ_T)(s, α) = \reject \\  
            ((s₂, t₂), p) & \norm(δ_S)(s, α) = (s₂, p) \text{ and } \norm(δ_T)(s, α) = (t₂, p) \\  
        \end{cases}
    \end{mathpar}
    The above definition of \(δ_{∼}\) is indeed well-defined, by case analysis on the result of \(δ_S\) and \(δ_T\) using the condition above:
    \begin{itemize}
        \item If \(δ_S(s, α) = \accept\), then by condition~\labelcref{itm:acc-condition}, \(δ_T(t, α) = \accept\) therefore \[\norm(δ_S)(s, α) = \norm(δ_T)(s, α) = \accept.\]
        \item If \(δ_S(s, α)\) transitions to a dead state or reject, then by conditions~\labelcref{itm:rej-or-dead} \(δ_T(t, α)\) will also transition to a dead state or reject, then \[norm(δ_S)(s, α) = \norm(δ_T)(s, α) = \reject.\]
        \item If \(δ_S(s, α) = (s', p)\), then by condition~\labelcref{itm:acc-condition} and condition~\labelcref{itm:rej-or-dead}, \(δ_T(t, α) = (t', q)\). By the contrapositive of condition~\labelcref{itm:transition-dead}, if either \(s, t\) are live, then \(p = q\).

        Then by condition~\labelcref{itm:transition-bisim}, there exists a bisimulation \(∼_{s', t'}\) between \(⟨s'⟩\) and \(⟨t'⟩\) s.t. \(s' ∼_{s', t'} t'\). Because bisimulation preserves liveness (\cref{thm:bisim-preserve-liveness}), \(s', t'\) has to be both dead or live, the both dead case is handled by the previous item, both live case will give us the case we desired:
        \[\norm(δ_S)(s, α) = (s', p) \text{ and } \norm(δ_T)(t, α) = (t', p)\]
    \end{itemize}
    And the diagram bisimulation needing to satisfy can be verified by unfolding the definition.
\end{proof}

The above theorem already gives us a way to recursively construct a algorithm that include \(s ∼ t\), this consequently will let us decide the trace equivalence of \(s\) and \(t\): \(⟦s⟧ = ⟦t⟧\).
However, this algorithm can be further optimized, we will then derive that a dead state can never relate to live states. This means that when checking the bisimulation of states \(s\) and \(t\), if we already know one of them is dead, we only need to check whether the other is dead, instead of going through the convoluted process mentioned in~\cref{thm:inductive-construction}.

However because homomorphism preserves liveness, if we already know one of the \(s\) and \(t\) is dead, the other has to be dead.

\begin{theorem}\label{thm:bisim-one-dead}
    Given two states \(s ∈ S\) and \(t ∈ T\), if \(s\) is a dead state in \(S\), then there exists a bisimulation \(∼\) between \(S\) and \(T\) where \(s ∼ t\) if and only if \(t\) is dead. Similarly for \(t ∈ T\).
\end{theorem}

\begin{proof}
    if there exists a bisimulation \(∼\), s.t. \(s ∼ t\), because \(s\) is dead and bisimulation preserves liveness~\cref{thm:bisim-preserve-liveness}, then \(t\) is dead.

    And if both \(t\) and \(s\) is dead, then a bisimulation can by constructed by~\cref{thm:bisim-between-dead}.
\end{proof}

\section{The Algorithm}

In this section we will present the pseudo-code for our on-the-fly algorithm. 
In order to implement the the inductive construction theorem (\cref{thm:inductive-construction}), we will need to determine the liveness of the state. This can be simply computed via a DFS from the state being checked.

\begin{algorithm}
    \caption{Check whether a state \(s\) is dead}\label{alg:check-dead-main}
    \begin{algorithmic}
        \Function{isDeadLoop}{$s ∈ S$, explored}
        \If {\(s ∈\) explored} {\Return explored} 
        \Else { 
            \For{\(α ∈ \At\)}{}{
                \Match{\(δ_{S}(s, α)\)}
                \Case{\(\accept\)} {\Return none} \Comment{\(s\) transition to accept}\EndCase
                \Case{($s', p$)}{ 
                    \If{\Call{IsDead}{s'} = none} {\Return none} \Comment{\(s\) transitions to a live state \(s'\)}
                    \Else { explored \(←\) (explored \(∪\) \Call{isDeadLoop}{$s'$, explored}) } \EndIf
                } \EndCase
                \EndMatch
            }\EndFor}
        \EndIf
        \State {\Return explored}    
        \EndFunction
    \end{algorithmic}
\end{algorithm}

By~\cref{thm:dead-iff-all-reachable-dead}, if \(s\) is dead then all the reachable states of \(s\) (denoted by \(⟨s⟩\)). Then by returning all the reachable states of \(s\), we can cache these states to avoid checking them again. To encapsulate the caching, we have the following function, which we will actually use in our bisimulation algorithm.

\begin{algorithm}
    \caption{A cached algorithm to check whether a state is dead}\label{alg:is-dead}
    \begin{algorithmic}
        \State{deadStates \(← ∅\)}

        \Function{isDead}{$s ∈ S$}
        \If {\(s ∈\) deadStates} {\Return true} 
        \ElsIf {\Call{isDeadLoop}{$s, ∅$} = none} {\Return false}
        \Else 
            \State {deadStates \(←\) (deadStates \(∪\) \Call{isDeadLoop}{$s, ∅$})}
            \State {\Return {true}}
        \EndIf
        \EndFunction
    \end{algorithmic}
\end{algorithm}

Given the direct correspondence between bisimulation and bisimulation equivalence and bisimulation in sub-algebra:
\begin{align*}
    & ∃ \text{ bisimulation } {∼} ⊆ ⟨s⟩ × ⟨t⟩ \text{ s.t. } s ∼ t \\
    & ⟺ ∃ \text{ bisimulation } {∼} ⊆ (⟨s⟩∪⟨t⟩) × (⟨s⟩∪⟨t⟩) \text{ s.t. } s ∼ t 
        & \text{\cref{thm:sub-coalg-preserve-bisim}}\\  
    & ⟺ ∃ \text{ bisimulation equivalence } {≃} ⊆ (⟨s⟩∪⟨t⟩) × (⟨s⟩∪⟨t⟩) \text{ s.t. } s ≃ t 
        & \text{\cref{thm:bisim-iff-bisim-equiv}}
\end{align*}
we can safely replace the bisimulation in inductive construction (\cref{thm:inductive-construction}) with bisimulation equivalence. 
Dealing with equivalence relations allows us to leverage efficient data structures like union find in our bisimulation algorithm.

We will use \(\Call{union}{$s,t$}\) to denote the operation to equate \(s\) and \(t\) in a union-find, and use \(\Call{eq}{$s,t$}\) to check if \(s\) and \(t\) belongs to the same equivalence class, i.e. share the same representative.
Specifically, we will use the union-find structures to keep track of the equivalence classes that we are in the process of checking, hence avoiding repeatedly checking the same pair of states to remove infinite loops.

Our on-the-fly bisimulation algorithm will decide whether there exists a bisimulation relation in \(⟨s⟩ ∪ ⟨t⟩\) s.t. \(s ∼ t\). This algorithm generally reproduce the setting of inductive construction theorem~\cref{thm:inductive-construction};
except by~\cref{thm:bisim-one-dead}, in the special case where \(s\) or \(t\) is dead, then we will only need to check whether the other is dead.

\begin{algorithm}
    \caption{On-the-fly bisimulation algorithm}\label{alg:bisim}
    \begin{algorithmic}
        \Function{equiv}{$s ∈ S, t ∈ T$}
        \If {\Call{eq}{$s, t$}} {\Return true}
        \ElsIf {\(s ∈\) deadStates} {\Return \Call{isDead}{$t$}} 
        \ElsIf {\(t ∈\) deadStates} {\Return \Call{isDead}{$s$}} 
        \Else 
        \For{\(α ∈ \At\)}{}{
            \Comment{Inductive construction,~\cref{thm:inductive-construction}}
            \Match{$δ_{S}(s, α), δ_{T}(t, α)$}
            \Case{\(\accept, \accept\)} {\Skip} \EndCase
            \Case{\(\reject, \reject\)} {\Skip} \EndCase
            \Case{\(\reject, (t', q)\)} {\Call{isDead}{$t'$}} \EndCase
            \Case{\((s', p), \reject\)} {\Call{isDead}{$s'$}} \EndCase
            \Case{\((s', p), (t', q)\)} {
                \If {\(p = q\)} {\Call{union}{$s, t$}; \Call{equiv}{$s, t$}} 
                \ElsIf {\Call{isDead}{$s$} and \Call{isDead}{$s$}} {\Skip}
                \Else { \Return false }
                \EndIf
            } \EndCase
            \Default {\Return false} \Comment{the results format does not match} \EndDefault
            \EndMatch
        }\EndFor
        \EndIf
        \Return true \Comment{no mismatch found}
        \EndFunction
    \end{algorithmic}
\end{algorithm}

The soundness and completeness of this algorithm can be observed by the fact that \emph{when the algorithm terminate}, the algorithm returns true. if and only if there exists a bisimulation between \(⟨s⟩\) and \(⟨t⟩\) s.t. \(s ∼ t\), which is then logically equivalent to trace equivalence.
Such equivalence is a direct consequence of~\cref{thm:bisim-one-dead,thm:inductive-construction}.

\section{Symbolic algorithm}

\subsection{Symbolic GKAT Coalgebra}

Given an alphabet \(K, B\), a \emph{symbolic GKAT coalgebra} \(Ŝ ≜ ⟨S, ϵ̂, δ̂⟩\) consists of a state set \(S\) consists of a accepting boolean \(ϵ̂\) and a transition function \(δ̂\):
\begin{mathpar}
    ϵ̂: S → \Bool_B, \and
    δ̂: S → S → K → \Bool_B,
\end{mathpar}
where \(\Bool_B\) is the free boolean algebra over \(B\) (boolean expressions modulo boolean algebra axioms), and for all states \(s ∈ S\), all the booleans results are ``disjoint''; namely the conjunction of any two expression from the set \(ϵ̂(s) + \{δ̂(s, s', p) ∣ s' ∈ S, p ∈ K\}\) are false in \(\Bool_B\).
The ordering on two symbolic coalgebra on the same carrier \(S\) is defined pointwise, namely 
\[⟨ϵ̂₁, δ̂₁⟩ ≤ ⟨ϵ̂₂, δ₂⟩ \text{ if and only if } ∀ s, s' ∈ S, ∀ p ∈ K, \hat{ϵ₁}(s) ≤ \hat{ϵ₂}(s) \text{ and } \hat{δ₁}(s, s', p) ≤ \hat{δ₂}(s, s', p).\]

Intuitively, the expressions in \(\Bool_B\) is treated as the ``symbolic'' transitions, which can be think of as a set of atoms by the isomorphism \(\Bool_B ≅ 2^{\At_B}\). 
Then we can explain the intuition for the definition of symbolic GKAT coalgebra:
\begin{itemize}
    \item \(ϵ̂(s)\) can be thought of as all the atoms that is accepted by the state \(s\).
    \item \(δ̂(s, s', p)\) denotes all the atoms that can transition from \(s\) to \(s'\) while executing \(p\).
\end{itemize}
Then the atoms excluded by both operations are implicitly rejecting. Formally, we can define a boolean expression that contains all the atoms that will be rejected 
\[ρ̂(s) = ¬ (ϵ̂(s) ∨ ⋁ \{b ∣ ∃ s' ∈ S, p ∈ K, δ̂(s, s', p) = b\}).\]

This intuition gives us a isomorphism between between GKAT coalgebra and symbolic GKAT coalgebra. Formally, given a GKAT coalgebra \(⟨S, δ⟩\), then its corresponding symbolic GKAT coalgebra will share the same states \(S\), with the transition defined as follows:
\begin{align*}
    ϵ̂(s) & ≜ ⋁ \{α ∈ \At ∣ δ(s, α) = \accept\} & 
    δ̂(s, s', p) & ⋁ \{α ∈ \At ∣ δ(s, α) = (s', p)\}.
\end{align*}
We will denote the above map as \(\symb\), i.e. \(\symb(⟨S, δ⟩) ≜ ⟨S, ϵ̂, δ̂⟩\). Notice that we are implicitly using atoms as an element in the boolean algebra, indeed an atom can be thought about as a conjunction.

The inverse of \(\symb\) is denoted as \(\unsymb\) where given a symbolic GKAT coalgebra \(⟨S, ϵ̂, δ̂⟩\), we construct a GKAT coalgebra over states \(S\) with the following transition:
\[δ(s, α) ≜ \begin{cases}
    \accept & α ∈ ϵ̂(s) \\  
    (s', p) & α ∈ δ̂(s, s', p) \\  
    \reject & \text{otherwise}
\end{cases}\]
Notice the \(δ\) is well-defined, that is given any \(s ∈ S\) and \(α ∈ \At\), \(δ(s,α)\) can only take on one value; because symbolic GKAT coalgebra is deterministic.

\subsection{Decision Procedure Of Symbolic GKAT coalgebra}

The formulation of symbolic GKAT coalgebra is mathematically nice, yet inefficient. Between most states \(s, s' ∈ S\) and action \(p ∈ K\), \(δ(s, s', p)\) will always be false. Therefore, we can developed a set representation of the derivative, where we will typically denote using \(Δ̂\):
\[Δ̂: S → 𝒫(\Bool_B × S × K).\]
we always assume transition functions denoted this way are \emph{deterministic}; specifically, all the boolean expression in the following set are disjoint: \[\{b ∣ ∃ s' ∈ S, p ∈ K, (b, s', p) ∈ Δ̂(s)\}.\]

Intuitively, if \((b, s', p) ∈ Δ̂(s)\), then \(s\) will transition to \(s'\) for all the atoms \(α ≤ b\) while executing \(p\). Notice that the above statement does not imply that \emph{all} \(α\) that makes \(s\) transition to \(s'\) while executing \(p\) will be covered by \(b\); that is, there can exists a \((b', s', p) ∈ Δ̂(s)\), where \(b ≠ b'\).

This means that the set representation of a symbolic GKAT coalgebra no on longer uniquely correspond to a GKAT coalgebra.
\begin{example}

\end{example}

However, every set representation of a symbolic GKAT coalgebra, indeed correspond to a GKAT coalgebra, specifically given a set representation \(⟨S, ϵ̂, Δ̂⟩\), we can construct a symbolic GKAT \(⟨S, ϵ̂, δ̂⟩\), by defining the following \(δ̂\):
\[δ̂(s, s', p) = ⋁ \{b ∣ (b, s', p) ∈ Δ̂(s)\}.\]
More importantly, the map described above is surjective. This means that all the GKAT coalgebra can indeed be represented as at least one set-representation.
Similarly, set representation can also be converted to GKAT coalgebra: given a set representation \(⟨S, ϵ̂, Δ̂⟩\), we can lower it into a GKAT coalgebra \(⟨S, δ⟩\), by constructing the following \(δ\):
\[δ(s, α) ≜ \begin{cases}
    \accept & α ∈ ϵ̂(s) \\  
    (s', p) & ∃ b ≥ α, (b, s', p) ∈ Δ̂(s) \\  
    \reject & \text{otherwise}
\end{cases}\]
Indeed, lowering the set representation directly to GKAT coalgebra give the same result as converting to symbolic GKAT coalgebra and then lowering. This fact is essential for us to relate these semantics.

Finally, we can lift the inductive construction theorem (\cref{thm:inductive-construction}) to set representation of GKAT coalgebras, giving a sound and complete symbolic decision procedure. 
In the following lemmas, we will use \(S, T\) to denote some GKAT coalgebra, and \(Ŝ, T̂\) will be as some set-represented symbolic GKAT coalgebra that corresponds to \(S\). 
Formally, if we lower \(Ŝ\) to GKAT coalgebra, we will obtain exactly \(S\); similarly for \(T̂\).

\begin{lemma}
    We let the \(s ∈ S\) and \(t ∈ T\) be two states, then the following equivalence between conditions in \cref{thm:inductive-construction} hold: 
    \begin{itemize}
        \item condition~\labelcref{itm:acc-condition} in inductive construction holds if and only if \(ϵ̂_{Ŝ}(s) = ϵ̂_{T̂}(t)\).
        \item condition~\labelcref{itm:transition-bisim} holds if and only if for all \((a, s', p) ∈ δ̂_{Ŝ}(s)\) and \((b, t', q) ∈ δ̂_{T̂}(t)\):
        \[a ∧ b ≠ 0 \text{ and } p = q ⟹ \text{exists a bisimulation \(∼\) between \(S\) and \(T\), } s.t. s' ∼ t'.\]
        \item condition~\labelcref{itm:transition-dead}, holds if and only if for all \((a, s', p) ∈ δ̂_{Ŝ}(s)\) and \((b, t', q) ∈ δ̂_{T̂}(t)\):
        \[a ∧ b ≠ 0 \text{ and } p ≠ q ⟹ \text{both \(s\) and \(t\) are dead}.\]
        \item condition~\labelcref{itm:rej-or-dead} holds if and only if the following holds:
    \end{itemize}
\end{lemma}

\begin{lemma}
    Given two states \(\)
\end{lemma}

\subsection{Symbolic Derivatives and Correctness}

Given a symbolic GKAT coalgebra \(⟨S, δ̂, ϵ̂⟩\) and a state \(s ∈ S\), we use the notation \(s ⇒ b\) to denote that \(ϵ̂(s) ≥ b\), and we use the notation \(s \transvia{b ∣ p} s'\) to denote that \(δ̂(s, s', p) ≥ b\), where the ordering is the typical ordering in boolean algebra.

The symbolic derivatives for GKAT forms a DKCT where the states are represented by GKAT expression \(\GKAT_{K, B}\), and the boolean algebra is the free boolean algebra \(\Bool_B\) over the test alphabet \(B\):
\begin{mathpar}
    ϵ: \GKAT_{K, B} → \Bool_B; \and  
    δ: \GKAT_{K, B} → \GKAT_{K, B} → K → \Bool_B.
\end{mathpar} 
the accepting map \(ϵ\) and transition map \(δ\) can be defined by the smallest maps that satisfy the following rules:
\begin{mathpar}
    \inferrule{\\}{b ⇒ b} \and 
    \inferrule{\\}{p \transvia{1 ∣ p} 1} \and 
    \inferrule{e ⇒ a}{e +_b f ⇒ b a} \and 
    \inferrule{f ⇒ a}{e +_b f ⇒ b̄ a} \and 
    \inferrule{e \transvia{a ∣ p} e'}{e +_b f \transvia{b ∧ a ∣ p} e'} \and 
    \inferrule{f \transvia{a ∣ p} f'}{e +_b f \transvia{b̄ ∧ a ∣ p} f'} \and 
    \inferrule{e ⇒ a \\ f ⇒ b}{e ⋅ f ⇒ a ∧ b} \and 
    \inferrule{e ⇒ a \\ f \transvia{b ∣ p} f'}{e ⋅ f \transvia{a ∧ b ∣ p} f'} \and 
    \inferrule{e \transvia{b ∣ p} e'}{e ⋅ f \transvia{b ∣ p} e' ⋅ f} \and 
    \inferrule{\\}{e^{(b)} ⇒ b̄} \and 
    \inferrule{e \transvia{a ∣ p} e'}{e^{(b)} \transvia{b ∧ a ∣ p} e' ⋅ {e}^{(b)}}
\end{mathpar}
Notice the rules above are very similar to the derivative rules for GKAT~\cite{schmid_GuardedKleeneAlgebra_2021}, this is no coincidence, as we can establish a correspondence of these two to prove correctness of our symbolic algorithm.

Since the rules listed above is monotonic, that is if we enlarge the boolean in the premise, we also enlarge the boolean in the result. 
So we can simply pick the largest boolean on the premise, which is either \(ϵ(e)\) or \(δ(e, e', p)\).
Since the disjunction of two boolean \(a ∨ b\) is the smallest boolean that is greater than both \(a\) and \(b\), thus we can implement \(ϵ\) and \(δ\) by iteration through all the rules, and take the disjunction of all the possible booleans.

\begin{align*}
    ϵ(b) & ≜ b 
        &  ϵ(e +_b f) & ≜ (b ∧ ϵ(e)) ∨ (b̄ ∧ ϵ(f))\\
    ϵ(q) & ≜ 0
        & ϵ(e ⋅ f) & ≜ ϵ(e) ∧ ϵ(f)\\  
    ϵ(e^{(b)}) & ≜ b̄ 
\end{align*}

However, the derivative function \[δ: \GKAT_{K, B} → \GKAT_{K, B} → K → \Bool_B,\] poses challenges in the implementation, since \(\GKAT_{K, B}\) is infinite, so it will be impractical to search through all the expressions.
However, we can treat the function \(δ\) extensionally: \[\GKAT_{K, B} → K → \Bool_B ⊆ 2^{\GKAT_{K, B} × K × \Bool_B},\]
and then because the boolean expression don't overlap for each input, hence the boolean expression are necessarily unequal; 
we can model the set by a partial map, thus all the \(δ\) can be represented as a function to the following partial maps.
\begin{align*}
    δ & : \GKAT_{K, B} → (\Bool_B ↛ \GKAT_{K, B} × K) \\
    δ(b) & ≜ \{\} \\
    δ(q) & ≜ \{1 ↦ (1, p)\}\\  
    δ(e +_b f) & ≜ \{b ∧ a ↦ (e', p) ∣ a ↦ (e', p) ∈ δ(e)\} 
        ∪ \{b̄ ∧ a ↦ (f', p) ∣ a ↦ (f', p) ∈ δ(f)\}\\
    δ(e ⋅ f) & ≜ \{b ↦ (e' ⋅ f, p) ∣ b ↦ (e, p) ∈ δ(e)\}
        ∪ \{ϵ(e) ∧ b ↦ (f', p) ∣ b ↦ (f', p) ∈ δ(f)\}\\
    δ(e^{(b)}) & ≜ \{b ∧ a ↦ (e' ⋅ e^{(b)}, p) ∣ a ↦ (e', p) ∈ δ(e)\}.
\end{align*}
FIXME: we still need to think about what representation do we want to go with, with this representation, you can also have duplicated K. We need a good example here of why don't we go with map. One good argument is that this representation is complex.

\subsection{Algorithm and Complexity}

\section{Implementation and Experiments}

% \subsection{the new norm}

% The theoretical foundation of this algorithm is surprisingly simple, where we only need one minor tweak to the definition of normalization.
% Instead of normalizing both the states \(S\) and the transition function \(δ_S\), we will only normalize the function \(δ_S\) to obtain \(δ_S\).

% Intuitively, this only rejects transitions to dead states, but doesn't remove dead states completely. As we will see later, the bisimulation generated by this notion will be able to contain rejection state, this allows us to only check for liveness after ``discrepancy'' as described in the intro.

% \begin{definition}[normalized homomorphism]
%     A normalized homomorphism \(h: S → T\) is a homomorphism with normalized transition functions:
%     \[
%         \begin{tikzcd}
%             S \ar{r}{h} \ar[swap]{d}{\norm(δ_S)} & T \ar{d}{\norm(δ_T)}\\  
%             G(S) \ar[swap]{r}{G(h)} & G(T)
%         \end{tikzcd}
%     \]
% \end{definition}

% This normalized homomorphism is very close to the definition of homomorphism in normalized GKAT coalgebra, hence inherits many of the properties. 
% Specifically, the language model \(𝒢\) is the final coalgebra in here.

% \subsection{soundness and completeness}

% The definition of normalized bisimulation naturally arise from a weakened form of homomorphism.

% \begin{definition}[Normalizd homomorphism]
%     A \emph{normalized homomorphism} \(h: S → T\) is coproduct (in set) of a \emph{homomorphism} between the normalized coalgebra \(h_{\norm}: \norm(S) → \norm(T)\) and a \emph{function} \(h_{\dead}: \dead{S} → \dead(T)\). Concretely,
%     \[h(s) ≜ \begin{cases}
%         h_{\norm}(s) & \text{if } s ∈ \norm(S) \\  
%         h_{\dead}(s) & \text{if } s ∈ \dead(S)
%     \end{cases}.\]
%     We will sometimes use the coproduct notation \(h = h_{\norm} + h_{\dead}\).
% \end{definition}

% \begin{theorem}
%     GKAT coalgebras under normalized homomorphisms form a category.
% \end{theorem}

% \begin{proof}
%     Recall that the composition operation compose with coproduct:
%     \[\begin{tikzcd}[row sep=3px]
%         {A_1} & {B_1} & {C_1} \\
%         & {+} \\
%         {A_2} & {B_2} & {C_2}
%         \arrow["{x_1}", from=1-1, to=1-2]
%         \arrow["{y_1}", from=1-2, to=1-3]
%         \arrow["{x_2}", from=3-1, to=3-2]
%         \arrow["{y_2}", from=3-2, to=3-3]
%     \end{tikzcd}
%     = 
%     A₁ + A₁ \xrightarrow{x₁ + x₂} B₁ + B₂ \xrightarrow{y₁ + y₂} C₁ + C₂.\]

%     Therefore the composition of \(h₁: S → T\) and \(h₂: T → U\) can be simply defined as 
%     the composition of its components: 
%     \begin{align*}
%         (h₁ ∘ h₂)_{\norm} & : S_{\norm} → U_{\norm} \\
%         (h₁ ∘ h₂)_{\norm} & ≜ h_{1\norm} ∘ h_{2\norm} \\[5px]
%         (h₁ ∘ h₂)_{\dead} & : S_{\dead} → U_{\dead}\\
%         (h₁ ∘ h₂)_{\dead} & ≜ h_{1\dead} ∘ h_{2\dead} \\[5px]
%         h₁ ∘ h₂ & : S → U\\
%         h₁ ∘ h₂ & ≜ (h₁ ∘ h₂)_{\norm} + (h₁ ∘ h₂)_{\dead}
%     \end{align*}
%     concretely this can be written down as
%     \[h₁ ∘ h₂(s) ≜ \begin{cases}
%         h_{1\norm} ∘ h_{2\norm}(s) & \text{if } s ∈ \norm(S) \\  
%         h_{1\dead} ∘ h_{2\dead}(s) & \text{if } s ∈ \dead(S)
%     \end{cases}.\]

%     The identity morphism is indeed the coproduct of the identity homomorphism on normalized coalgebra and the identity function on the dead states.

%     The associativity and identity axiom can be verified by decomposing the homomorphism to live and dead states.
% \end{proof}

% \begin{remark}
%     Diagrammatically, normalized homomorphism is completely characterized by the following commutative square in the category of set:
%     \[
%         \begin{tikzcd}
%             S \ar{rr}{h} \ar[swap]{dd}{δ_S} && T \ar{d}{δ_T} \\  
%             && G(T) \ar{d}{[-]_{\norm}}\\
%             G(S) \ar[swap]{r}{ G(h)}
%                 & G(T) \ar[swap]{r}{[-]_{\norm}}
%                 & G(T)/{∼_{\norm}}
%         \end{tikzcd}
%     \]
%     where \({∼_{\norm}} ⊆ (2 + S × K) × (2 + S × K)\) is the minimal equivalence generated by the following relation \[\{(\reject, (s, p)) ∣ p ∈ K, \text{\(s\) is dead in \(S\)}\};\] 
%     and \(G(T)/{∼_{\norm}}\) is the point-wise quotient of \(G(T)\) by the smallest equivalence relation containing the relation \(∼_{\norm}\), and \([-]_{\norm}\) is the function that maps a function in \(G(T)\) to its corresponding equivalence class in \(G(T)/{∼_{\norm}}\).

%     The proof of this equivalence is interesting, however irrelevant to this paper, we will defer the proof to appendix, in~\cref{sec:diagram-of-norm-sim}.
% \end{remark}

% With the category of GKAT coalgebra and normalized homomorphism defined, we obtain the conventional definition of bisimulation, which is a span of projections in that category~\cite{jacobs_IntroductionCoalgebraMathematics_2016,rutten_UniversalCoalgebraTheory_2000}:
% \[
%     \begin{tikzcd}[column sep=1cm]
%         S & ∼ \ar[swap]{l}{π₁} \ar{r}{π₂} & T
%     \end{tikzcd}
% \]
% We call such bisimulation \emph{normalized bisimulation} to distinguish it from the bisimulation generated in the category of GKAT coalgebra and GKAT coalgebra homomorphism.

% Like proving the soundness and completeness of bisimulations, we can first identify that the language semantics is indeed the unique normalized homomorphism to the final GKAT coalgebra in out category:
% \begin{theorem}
%     The language GKAT \(𝒢\) is the final GKAT coalgebra in the category of GKAT coalgebra and normalized homomorphisms, with the language semantics \(⟦-⟧_S: S → 𝒢\) as the unique normalized homomorphism.
% \end{theorem}

% \begin{proof}
%     Recall that \(\norm(𝒢)\) is the final object in the category of GKAT coalgebra with homomorphisms, then there exists a unique homomorphism \(⟦-⟧_{\norm(S)}: S → 𝒢\).

%     Also recall that \(\dead(𝒢)\) is a singleton set that only contains the empty language \(∅\). Recall the singleton set is the final algebra in the category of set \(!_S: S → \{∅\}\).

%     And by observing the definition of \(⟦-⟧_S\), we can verify that
%     \[⟦-⟧_S = ⟦-⟧_{\norm(S)} + !_S\]. 
%     The above equation shows that the language semantics \(⟦-⟧_S\) is the unique normalized homomorphism from any GKAT coalgebra \(S\) to \(𝒢\), proving that \(𝒢\) is the final GKAT coalgebra in the category of GKAT coalgebra with normalized homomorphism.
% \end{proof}

% \begin{corollary}[soundness]
%     Given two states \(s ∈ S\) and \(t ∈ T\), 
%     \[s ∼ t ⟹ ⟦s⟧_S = ⟦t⟧_T\]
% \end{corollary}

% \begin{proof}
%     This proof mirrors the classical soundness proof for bisimulation with regard to behavioral equivalence~\cite{rutten_UniversalCoalgebraTheory_2000}.

%     Then by finality of \(𝒢\):
%     \[
%         \begin{tikzcd}[column sep=2cm, row sep=1cm]
%             & 𝒢 & \\
%             S \ar{ur}{⟦-⟧_S} 
%                 & ∼ \ar[u, near start, dashed, "⟦-⟧_S ∘ π₁ = ⟦-⟧_T ∘ π₂"description]\ar{l}{π₁} \ar[swap]{r}{π₂} 
%                 & T \ar[swap]{ul}{⟦-⟧_T}
%         \end{tikzcd}
%     \]
%     And take any two states \(s ∈ S\) and \(t ∈ T\):
%     \[s ∼ t ⟹ (s, t) ∈ {∼} ⟹ ⟦-⟧_S ∘ π₁(s,t) = ⟦-⟧_T ∘ π₂(s,t) ⟹ ⟦s⟧_S = ⟦t⟧_T.\]
% \end{proof}

% The completeness proof is slightly more complicated: because we use a modified notion of homomorphism, thus the classical completeness argument~\cite{rutten_UniversalCoalgebraTheory_2000} no longer applies.
% However, we can still devise a simple proof based on the completeness of bisimulation between normalized GKAT coalgebras~\cite{smolka_GuardedKleeneAlgebra_2020}.

% Given two states \(s ∈ S\) and \(t ∈ T\), we use the notation \(s ≡ t\) to denote that there exists a normalized bisimulation \({∼} ⊆ S × T\) s.t. \(s ∼ t\). 
% \begin{remark}
%     An alternative interpretation of \(≡\) is the maximal normalized bisimulation between \(S\) and \(T\). However, by using the current definition of \(≡\), we can omit the proof that all normalized bisimulations form a complete lattice.
% \end{remark}

% \begin{lemma}[normalized bisimulation between dead states]\label{thm:norm-bisim-between-dead}
%     Given two GKAT coalgebra \(S\) and \(T\), any relation \({∼} ⊆ \dead(S) × \dead(T)\) is a normalized bisimulation between \(S\) and \(T\).
% \end{lemma}

% \begin{proof}
%     From the definition, \(∼\) only contains one dead state \((s,t)\) and no live states, thus the projection from \(\norm(∼) = ∅\) is the unique map from the empty set \(π₁: ∅ → \norm(S)\) and \(π₂: ∅ → \norm(T)\), which are indeed a GKAT homomorphism.

%     The projection from the only dead state \((s, t)\) will also reach \(s\) and \(t\), which are dead states in their respective GKAT coalgebra.

%     Thus the projection from live states from a homomorphism in normalized coalgebra, and projections from dead states from a function to dead states. Therefore both projections are normalized homomorphism, and \(∼\) is a bisimulation.
% \end{proof}

% \begin{lemma}\label{thm:norm-bisim-from-bisim-on-norm}
%     Given a bisimulation between normalized GKAT coalgebras \({∼} ⊆ \norm(S) × \norm(T)\), \(∼\) is also a normalized bisimulation between \(S\) and \(T\).
% \end{lemma}

% \begin{proof}
%     TODO
% \end{proof}

% \begin{theorem}[completeness]
%     Given two states \(s ∈ S\) and \(t ∈ T\), 
%     \[⟦s⟧ = ⟦t⟧ ⟹ s ≡ t.\]
% \end{theorem}

% \begin{proof}
%     We do this by case analysis on the whether \(⟦s⟧\) and \(⟦t⟧\) is empty.

%     Recall that the empty language is a dead state in the language model \(𝒢\), and because \(⟦-⟧\) is a normalized homomorphism, which preserves liveness by definition. Hence, when \(⟦s⟧ = ⟦t⟧ = ∅\), both \(s\) and \(t\) has to be dead. Therefore by \cref{thm:norm-bisim-between-dead}, we can simply construct a bisimulation \(∼ ≜ \{(s, t)\}\).

%     If both \(⟦s⟧\) and \(⟦t⟧\) are non-empty, then by the same reasoning, both \(s\) and \(t\) are live, then by the completeness of bisimulation on normalized GKAT coalgebra~\cite{smolka_GuardedKleeneAlgebra_2020}, there exists a bisimulation \({∼} ⊆ \norm(S) × \norm(T)\), s.t. \(s ∼ t\).
% \end{proof}

% \begin{lemma}\label{thm:norm-bisim-preserve-liveness}
%     Given a normalized bisimulation \({∼} ⊆ S × T\), then by definition \(∼\) form a GKAT coalgebra, and the following statements hold:
%     \begin{itemize}
%         \item if \(s ∼ t\) then either \(s\) and \(t\) are both live or \(s\) and \(t\) are both dead.
%         \item the live states in \(∼\) are exactly the states where both components are live: \[\norm(∼) ≜ \{(s, t) ∣ s ∼ t, s ∈ \norm(S), t ∈ \norm(T)\}.\]
%     \end{itemize}
% \end{lemma}

% \begin{proof}
%     Because \(π₁: {∼} → S\) and \(π₂: {∼} → T\) are normalized homomorphisms, then by definition, they will map live states to live states and dead states to dead states.
%     Thus if we consider \((s, t) ∈ {∼}\): \[π₁(s, t) = s \text{ is live in } S ⟺ (s, t) \text{ is live in } {∼} ⟺ π₂(s, t) = t \text{ is live in } T.\]
%     Thus \(∼\) only relates states with the same liveness.

%     By the above equivalence chain, \((s, t) ∈ {∼}\) is live if and only if both \(s, t\) are live, thus 
%     \[\norm(∼) ≜ \{(s, t) ∣ s ∼ t, s ∈ \norm(S), t ∈ \norm(T)\}.\]
% \end{proof}

% This observation allows us to express a strong connection between normalized bisimulation and bisimulation on normalized coalgebra.

% \begin{theorem}\label{thm:bisim-on-norm<=norm-bisim}
%     Given two GKAT coalgebra, if \({∼} ⊆ S × T\) then \(\norm(∼) ⊆ \norm(S) × \norm(T)\) is a bisimulation. 
% \end{theorem}

% \begin{proof}
%     because \(π₁\) and \(π₂\) are normalized homomorphisms, then they can be decomposed into a homomorphism on normalized coalgebra and a function on dead states: 
%     \[π₁ = π_{1\norm} + π_{1\dead} \text{ and } π₂ = π_{2\norm} + π_{2\dead},\] 
%     where \(π_{1\norm}\) and \(π_{2\norm}\) are projection homomorphism to the normalized GKAT coalgebra \(\norm(S)\) and \(\norm(T)\).
%     Therefore, \(\norm(∼)\) form a span of \(\norm(S)\) and \(\norm(T)\) with these projection homomorphisms. 
%     Diagrammatically, this span depicts the following commutative diagram in set:
%     \[
%         \begin{tikzcd}[column sep=2cm]
%             \norm(S) \ar[swap]{d}{\norm(δ_S)}& 
%             \norm(∼) \ar[swap]{l}{π_{1\norm}} \ar{r}{π_{2\norm}} \ar{d}{\norm(δ_∼)}
%             & \norm(T) \ar{d}{\norm(δ_T)}\\  
%             G(\norm(S)) & G(\norm(∼)) \ar[swap]{l}{G(π_{1\norm})} \ar{r}{G(π_{2\norm})} & G(\norm(T))
%         \end{tikzcd}
%     \]
%     hence \(\norm(∼)\) is indeed a bisimulation between \(\norm(S)\) and \(\norm(T)\).
% \end{proof}

% \begin{theorem}\label{thm:bisim-on-norm=>live}
%     Given a bisimulation \({∼} ⊆ \norm(S) × \norm(T)\), then by definition \({∼}\) is also a normalized bisimulation between \(S\) and \(T\).
% \end{theorem}

% \begin{proof}
%     because \(∼\) is a bisimulation, then the projection \(π₁: {∼} → \norm(S)\) and \(π₂: {∼} → \norm(T)\) are indeed  GKAT homomorphisms. 
%     By~\cref{thm:hom-preserve-liveness}, \(\dead(∼)\) is the empty set, thus have unique function 
% \end{proof}

% With the connection between normalization and normalized bisimulation, we can utilize the soundness and completeness of bisimulation between normal GKAT coalgebra to derive the soundness and completeness of normalized bisimulation. 
% Given two states \(s ∈ S\) and \(t ∈ T\), we use the notation \(s ≡ t\) to denote that there exists a normalized bisimulation \({∼} ⊆ S × T\) s.t. \(s ∼ t\). 
% \begin{remark}
%     An alternative interpretation of \(≡\) is the maximal normalized bisimulation between \(S\) and \(T\). However, by using the current definition of \(≡\), we can omit the proof that all normalized bisimulation form a complete lattice, which is a corollary that bisimulation forms a complete lattice.
% \end{remark}

% \begin{theorem}[soundness and completeness]
%     Given two states \(s ∈ S\) and \(t ∈ T\):
%     \[s ≡ t ⟺ ⟦s⟧ = ⟦t⟧.\]
% \end{theorem}

% \begin{proof}
%     We first show the \(⟹\) direction, because \(s ≡ t\), then there exists \(∼\) s.t. \(s ∼ t\). And by~\cref{thm:bisim-on-norm<=norm-bisim}, \(\norm(∼)\) is a bisimulation between 
% \end{proof}

% \subsection{Construction}

% \begin{lemma}
%     Given a sub-GKAT coalgebra \(S' ⊑ S\), the inclusion homomorphism \(i: S' → S\) is also a normalized homomorphism.
% \end{lemma}
% \begin{proof}
%     TODO.
% \end{proof}

% \begin{theorem}[sub-coalgebra perserves normalized bisimulation]\label{thm:sub-coalg-preserve-norm-bisim}
%     Given two sub-GKAT colagebra \(S' ⊑ S\) and \(T' ⊑ T\), then 
%     \begin{itemize}
%         \item a normalized bisimulation \(∼_{S', T'}\) between \(S'\) and \(T'\) is also a normalized bisimulation between \(S\) and \(T\)
%         \item a normalized bisimulation \(∼\) between \(S\) and \(T\), can also be restricted to \(S'\) and \(T'\) in the following manner, by restricting both the carrier and the transition:
%         \begin{align*}
%             {∼_{S', T'}} & ≜ \{(s,t) ∣ s ∼ t, s ∈ S', t ∈ T'\} \\
%             δ_{∼_{S', T'}} & ≜ δ_{∼} |_{∼_{S', T'}}
%         \end{align*}
%     \end{itemize}
% \end{theorem}

% \begin{proof}
%     TODO.
% \end{proof}

% \begin{lemma}[normalized bisimulation closed under arbitrary union]\label{thm:norm-bisim-close-under-union}
%     Given a collection of normalized bisimulation \(∼ᵢ ⊆ S × T\) indexed by an arbitrary set \(i ∈ I\), then there exists a normalized bisimulation \(∼\) that includes all the pairs in \(∼ᵢ\) for all \(i\).
% \end{lemma}

% \begin{theorem}[inductive construction]\label{thm:inductive-construction}
%     Given two GKAT coalgebra \(S\) and \(T\), and two of their elements \(s ∈ S\) and \(t ∈ T\),
%     there exists a normalized bisimulation \({∼} ⊆ ⟨s⟩ × ⟨t⟩\) s.t. \(s ∼ t\), if and only if all of the following holds:
%     \begin{enumerate}
%         \item\label{itm:acc-condition} \(δ_{S}(s, α) = \accept ⟺ δ_{T}(t, α) = \accept\);
%         \item\label{itm:transition-bisim} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{T}(t, α) = (t', p)\), then there exists a normalized bisimulation \({∼_{s',t'}}\) on \(⟨s'⟩\) and \(⟨t'⟩\), s.t. \(s' ∼_{s',t'} t'\);
%         \item\label{itm:transition-dead} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{T}(t, α) = (t', q)\), s.t. \(p ≠ q\), then both \(s'\) and \(t'\) are dead;
%         \item\label{itm:rej-or-dead} \(s\) reject \(α\) or transition to a dead state via \(α\) if and only if \(t\) rejects \(α\) or transition to a dead state via \(α\).
%     \end{enumerate}
% \end{theorem}

% \begin{proof}
%     We first prove \(⟹\) direction, recall that \(∼\) is a GKAT coalgebra,
%     where the projection from the normalized version \(π₁ : \norm(∼) → \norm(S)\) and \(π₂: \norm(∼) → \norm(T)\), are GKAT homomorphisms:
%     \[
%         \begin{tikzcd}
%             \norm(S) \ar[swap]{d}{\norm(δ_S)} 
%             & \norm(∼) \ar{d}{\norm(δ_∼)} \ar[swap]{l}{π₁} \ar{r}{π₂}
%             & \norm(T) \ar{d}{\norm(δ_T)}\\  
%             G(\norm(S)) 
%             & G(\norm(∼)) \ar{l}{G(π₁)} \ar[swap]{r}{G(π₂)} 
%             & G(\norm(T)) 
%         \end{tikzcd}
%     \]

%     First realize that in order for either \(s\) or \(t\) to have an accepting transition, it has to be a live state. Because \(s ∼ t\) and normalized bisimulation preserves liveness (\Cref{thm:norm-bisim-preserve-liveness}), both \(s\) and \(t\) has to be live. Therefore, condition \labelcref{itm:acc-condition} holds:
%     \begin{align*}
%         δ_S(s, α) = \accept 
%         & ⟺ \norm(δ_S)(s, α) = \accept 
%         ⟺ \norm(δ_∼)((s,t), α) = \accept \\ 
%         & ⟺ \norm(δ_T)(t, α) = \accept 
%         ⟺ δ_T(t, α) = \accept
%     \end{align*}

%     Notice that if \(s ∼ t\), \(s\) transition to \(s'\), and \(t\) transition to \(t'\), then \(s'\) and \(t'\) are either both dead or both live: 
%     \begin{itemize}
%         \item if \(s, t\) are dead, then \(s'\) and \(t'\) has to both be dead, because reachable state of \(s\) and \(t\) has to be dead (\cref{thm:dead-iff-all-reachable-dead}).
%         \item if \(s, t\) are live, then \((s, t) ∈ \norm(∼)\). Because \(\norm(∼)\) is a bisimulation and by its equation, \(s'\) and \(t\) has to be both live or dead.
%     \end{itemize}
%     Condition~\labelcref{itm:transition-bisim} can be proven by case analysis on the liveness of \(s'\) and \(t'\):
%     \begin{itemize}
%         \item if both \(s'\) and \(t'\) are live, then both \(s\) and \(t\) are live:
%         \[(s, t) ∈ \norm(∼) ⟹ (s', t') ∈ \norm(∼)\]
%         Then, \(∼_{s',t'}\) can be constructed by restricting \(∼\) to \(⟨s'⟩\) and \(⟨t'⟩\), like stated in~\cref{thm:sub-coalg-preserve-norm-bisim}.
%         \item if both \(s'\) and \(t'\) are dead, then we can construct \(∼_{s',t'}\) as described in~\cref{thm:norm-bisim-between-dead}.
%     \end{itemize}
%     Thus, concluding our proof of condition~\labelcref{itm:transition-bisim}.

%     Then for condition~\labelcref{itm:transition-dead}, if both \(s\) and \(t\) are live, then this transition does not satisfy the condition of bisimulation, therefore \(s'\) and \(t'\) has to be live. If both \(s\) and \(t\) are dead, then all the reachable state from \(s\) and \(t\) has to be dead (\cref{thm:dead-iff-all-reachable-dead}), therefore \(s'\) and \(t'\) will also need to be dead.

%     Finally the proof of condition~\labelcref{itm:rej-or-dead} is similar to the proof of condition~\labelcref{itm:transition-dead}, which is proven by case analysis on the liveness of \(s\) and \(t\): 
%     \begin{itemize}
%         \item if both \(s\) and \(t\) are dead, by \cref{thm:dead-iff-all-reachable-dead} and definition, they can only transition to dead states or reject.
%         \item if both \(s\) and \(t\) are live, then \((s, t) ∈ \norm(∼)\), and
%         \begin{align*}
%             & δ_S(s, α) = \reject \text{or transition to dead states} \\
%             & ⟺ \norm(δ_S)(s, α) = \reject  
%             ⟺ \norm(δ_∼)((s,t), α) = \reject 
%             ⟺ \norm(δ_T)(t, α) = \reject \\
%             & ⟺ δ_T(t, α) = \reject \text{or transition to dead states}
%         \end{align*}
%     \end{itemize}

%     Then we show the \(⟸\) direction. We use \(∼_{s',t'}\) to denote the maximal bisimulation between \(⟨s'⟩\) and \(⟨t'⟩\):
%     \[∼' ≜ ⋃ \{{∼_{s', t'}} ∣ ∃ α ∈ \At, p ∈ K, δ_S(s, α) = (s', p), δ_T = (t', p)\}\]
%     Consider all the \(s'\) and \(t'\) mentioned in the above definition, because they are reachable from \(s\) and \(t\) respectively, then \(⟨s'⟩ ⊑ ⟨s⟩\) and \(⟨t'⟩ ⊑ ⟨t⟩\). And by~\cref{thm:sub-coalg-preserve-norm-bisim}, \(∼_{s',t'}\) in above definitions are all bisimulation between \(⟨s⟩\) and \(⟨t⟩\). Finally, because union of normalized bisimulation is still a normalized bisimulation (\cref{thm:norm-bisim-close-under-union}), \(∼'\) is a normalized bisimulation between \(S\) and \(T\).

%     Finally we will need to show that the addition of \((s, t)\) also gives a normalized bisimulation.
% \end{proof}

% \section{Normalized Bisimulation .bak}

% The normalization procedure for GKAT requires the entire automaton to be known prior to the execution of the bisimulation algorithm; specifically, in order to compute the liveness of a state \(s\), it is necessary iterate through all its reachable states \(⟨s⟩\) to see if there are any accepting states within.
% This limitation poses challenges to design an efficient on-the-fly algorithm for GKAT.
% In order to make the decision procedure scalable, we will need to merge the normalization and bisimulation procedure, so that our algorithm can normalized the automaton only when we need to.

% \subsection{the new norm}

% normalization is a important procedure, as it identifies the trace model \(𝒢\) as the final coalgebra in normalized GKAT coalgebra, enabling its computation via bisimulation~\cite{smolka_GuardedKleeneAlgebra_2020}.
% Specifically, given a GKAT coalgebra \(S\), we can normalized it by taking out all of its live state, denoted as \(\norm(S)\), and reroute all the transitions to dead state into reject:
% \begin{align*}
%     \norm(δ_S) & : \norm(S) → G(\norm(S))\\\
%     \norm(δ_S)(s) & = \begin{cases}
%         \reject & \text{if } δ_S(s) = (s', p) \text{ and \(s'\) is dead}\\
%         δ_S(s) & \text{otherwise}
%     \end{cases}
% \end{align*}
% Similarly, we will denote the dead states of \(S\) as \(\dead(S)\), and recall that \(S\) can be partitioned into its dead and live states \(S = \norm(S) + \dead(S)\), where the \(+\) is the coproduct operation in set.

% Unfortunately, this style of normalization~\cite{smolka_GuardedKleeneAlgebra_2020} does not naturally induce a on-the-fly decision algorithm, as removing the dead state necessitates detecting dead states prior to execution of bisimulation. In here, we introduce a new notion of normalization, which is implemented by relaxing the equation of coalgebra homomorphism.

% We first define the relation \(∼_{\norm}\) on results of transitions, specifically for any GKAT coalgebra \(S\):
% \[{∼_{\norm}} ⊆ (2 + S × K) × (2 + S × K)\]
% is the minimal equivalence generated by the following relation
% \[\{(\reject, (s, p)) ∣ p ∈ K, \text{\(s\) is dead in \(S\)}\}.\]
% Intuitively, \({∼_{\norm}}\) will relate all the transition to dead state with rejection, while other results, like acceptance or transitions to the live states, are only related to themselves.

% A function \(h: S → T\) is a \emph{normalized homomorphism} if the following diagram commute:
% \[
%     \begin{tikzcd}
%         S \ar{rr}{h} \ar[swap]{dd}{δ_S} && T \ar{d}{δ_T} \\  
%         && G(T) \ar{d}{[-]_{\norm}}\\
%         G(S) \ar[swap]{r}{ G(h)}
%             & G(T) \ar[swap]{r}{[-]_{\norm}}
%             & G(T)/{∼_{\norm}}
%     \end{tikzcd}
% \]
% where \(G(T)/{∼_{\norm}}\) is the point-wise quotient of \(G(T)\) by the smallest equivalence relation containing the relation \(∼_{\norm}\), and \([-]_{\norm}\) is the function that maps a element in \(G(T)\) to its corresponding equivalence class in \(G(T)/{∼_{\norm}}\).
% For brevity, we sometimes leave the function \([-]_{\norm}\) implicit and draw the commutative diagram as follows:
% \[
%     \begin{tikzcd}
%         S \ar{r}{h} \ar[swap]{d}{δ_S} & T \ar{d}{δ_T} \\  
%         G(S) \ar[swap]{r}{G(h)} & G(T)/{∼_{\norm}}
%     \end{tikzcd}
% \]

% \begin{example}

% \end{example}

% The notion of normalized homomorphism indeed have connection with existing noting of normalization

% %%%%% NOTE: this proof is incorrect, we cannot verify the function 
% %%%%% h : \norm(S) -> \norm(T) is well defined
% %
% % \begin{theorem}
% %     Normalized homomorphism induces a homomorphism on normalized coalgebra:
% %     \[
% %         \begin{tikzcd}
% %             S \ar{r}{h} \ar[swap]{d}{δ_S} & T \ar{d}{δ_T} \\  
% %             G(S) \ar[swap]{r}{G(h)} & G(T)/{∼_{\norm}}
% %         \end{tikzcd}
% %         ⟹ 
% %         \begin{tikzcd}
% %             \norm(S) \ar{r}{h} \ar[swap]{d}{\norm(δ_S)} 
% %                 & \norm(T) \ar{d}{\norm(δ_T)} \\  
% %             G(\norm(S)) \ar[swap]{r}{G(h)} & G(\norm(T))
% %         \end{tikzcd}
% %     \]
% % \end{theorem}

% % \begin{proof}
% %     The commutative triangle in the premise can be written as the following equation:
% %     \[G(h) ∘ δ_S ∼_{\norm} δ_T ∘ h.\] 
% %     Take any live state \(s ∈ S\), we will show the above implication by case analysis on the value on the bottom right corner \((G(h) ∘ δ_S)(s, α)\) for an arbitrary \(α ∈ \At\):
% %     \begin{itemize}
% %         \item if \((G(h) ∘ δ_S)(s, α)\) transitions to dead state or reject, then \((δ_T ∘ h)(s, α)\) will also transition to dead state or reject. Therefore 
% %         \[(\norm(δ_T) ∘ h)(s, α) = \reject.\]
% %         We then case on the result of \(δ_S(s, α) ∈ G(S)\):
% %         \begin{itemize}
% %             \item If \(δ_S(s, α)\) transitions to dead state or reject, then 
% %             \[\norm(δ_S)(s, a) = \reject;\] 
% %             and 
% %             \begin{align*}
% %                  (G(h) ∘ \norm(δ_S))(s, α) = \reject = (\norm(δ_T) ∘ h)(s, α).
% %             \end{align*}
% %             \item \(δ_S(s, α)\) transition to live state or accept, then by the definition of norms:
% %             \[\norm(δ_S)(s, a) = δ_S(s, α);\]
% %             which implies 
% %             \[(G(h) ∘ \norm(δ_S))(s, α) = (G(h) ∘ δ_S)(s, α)\]
% %             Finally, because \((G(h) ∘ δ_S)(s, α)\) reject or transition to dead states, yet dead states are not in \(\norm(T)\).
% %             \[(G(h) ∘ \norm(δ_S))(s, α) = \reject = (\norm(δ_T) ∘ h)(s, α).\]
% %         \end{itemize}
% %         \item if 
% %     \end{itemize}
% % \end{proof}

% \begin{lemma}\label{thm:norm-hom-to-norm-is-hom}
%     If given a normal GKAT coalgebra \(T\), then a normalized homomorphism \(S → T\) is a GKAT coalgebra homomorphism from \(S → T\).
% \end{lemma}

% \begin{proof}
%     Since there is no dead state in \(T\), then \(∼_{\norm}\) is just the identity relation, thus \(G(T)/∼_{\norm} = G(T)\).
%     Hence the following diagram commute for \(h\):
%     \[
%         \begin{tikzcd}
%             S \ar{r}{h} \ar[swap]{d}{δ_S} & T \ar{d}{δ_T} \\  
%             G(S) \ar[swap]{r}{G(h)} & G(T)
%         \end{tikzcd}
%     \]
%     Therefore \(h\) is a GKAT coalgebra homomorphism.
% \end{proof}

% \begin{lemma}
%     Given a normal GKAT coalgebra \(T\) and a normalized bisimulation \(h: S → T\), then for all \(s ∈ S\), the principle sub-domain \(⟨h(s)⟩_T\) is in the range of \(h\).
% \end{lemma}

% \begin{proof}
%     Because \(T\) is normal and by~\cref{thm:norm-hom-to-norm-is-hom}, \(h\) is a normalized homomorphism, thus we can construct a epi-mono factorization~\cite[Theorem 7.1]{rutten_UniversalCoalgebraTheory_2000}:
%     \[
%         \begin{tikzcd}
%             S \ar{r}{h} \ar[swap]{d}{δ_S} 
%                 & h(S) \ar[hook]{r}{i} \ar{d}{δ_T}
%                 & T \ar{d}{δ_T} \\  
%             G(S) \ar[swap]{r}{G(h)} 
%                 & G(h(S)) \ar[hook]{r}{i}
%                 & G(T)
%         \end{tikzcd}
%     \]
%     Thus \(h(S)\) is a sub-GKAT coalgebra of \(T\). Because \(h(s) ∈ h(S)\) and \(⟨h(s)⟩_T\) is the smallest sub-GKAT coalgebra that contains \(h(S)\), \[⟨h(s)⟩_T ⊑ h(S) ⟹ ⟨h(s)⟩_T ⊆ h(S).\]
% \end{proof}

% \begin{theorem}[normalized homomorphism preserves liveness]
%     Given a normalized homomorphism \(h: S → T\), then 
%     \begin{align*}
%         s \text{ is accepting} & ⟺ h(s) \text{ is accepting} \\
%         s \text{ is live} & ⟺ h(s) \text{ is live} \\
%         s \text{ is dead} & ⟺ h(s) \text{ is dead}
%     \end{align*}
% \end{theorem}

% \begin{theorem}[decomposition]\label{thm:norm-hom-decomp}
%     Let \(S\) and \(T\) be GKAT coalgebras, a function \(h: S → T\) is a normalized homomorphism if and only if \(h = h_{\norm} + h_{\dead}\) where \(h_{\norm} : \norm(S) → \norm(T)\) is a \emph{homomorphism}, and \(h_{\dead}: \dead(S) → \dead(T)\) is a function.
% \end{theorem}
% \begin{proof}
%     The \(⟹\) direction can be shown by constructing \(h_{\norm}\) and \(h_{\dead}\) as restrictions of \(h\):
%     \begin{mathpar}
%         h_{\norm} ≜ h|_{\norm(S)}
%         \and 
%         h_{\dead} ≜ h|_{\dead(S)}
%     \end{mathpar}
%     and because 
% \end{proof}

% \begin{lemma}[homomorphic pre-image preserves normalized bisimulation]
%     Given two GKAT coalgebra homomorphism \(h₁: S₁ → T₁\) and \(h₂: S₁ → T₁\), with a normalized bisimulation \(∼\), then the homomorphic pre-image of \(∼\) with respect to \(h₁ × h₂\):
%     \[(h₁ × h₂)^{-1}(∼) ⊆ S₁ × S₂,\]
%     is a normalized bisimulation between \(S₁\) and \(S₂\).
% \end{lemma}
% \begin{proof}
%     Recall that pre-image in the category of set forms a (weak) pullback, and because \(G\) is a elementary polynomial functor~\cite[Definition 2.2.1]{jacobs_IntroductionCoalgebraMathematics_2016} hence preserves weak pullback~\cite[Lemma 3.3.6]{jacobs_IntroductionCoalgebraMathematics_2016}. Therefore we obtain the following pullback squares: we denote \((h₁ × h₂)^{-1}(∼)\) as \(∼_S\),
%     \begin{mathpar}
%         \begin{tikzcd}
%             ∼_S \ar[hook]{r}{i} \ar[swap]{d}{h₁ × h₂} & S₁ × S₂ \ar{d}{h₁ × h₂} \\  
%             ∼ \ar[hook]{r}{i} & T₁ × T₂ \\
%         \end{tikzcd}
%         \and 
%         \begin{tikzcd}
%             G(∼_S) \ar[hook]{r}{G(i)} \ar[swap]{d}{G(h₁ × h₂)} & G(S₁ × S₂) \ar{d}{G(h₁ × h₂)}\\  
%             G(∼) \ar[hook]{r}{G(i)} & G(T₁ × T₂) \\
%         \end{tikzcd}
%     \end{mathpar}

% \end{proof}

% \begin{theorem}[completeness]
%     Given two GKAT coalgebra \(S, T\) and two states \(s ∈ S\) and \(t ∈ T\), 
%     \[⟦s⟧ = ⟦t⟧ ⟺ \text{there exists a bisimulation \({∼} ⊆ S × T\) s.t. } s ∼ t.\]
% \end{theorem}
% \begin{proof}

% \end{proof}

% \subsection{definitions and construction}

% Notice that the finite trace semantics \(⟦-⟧: S → 𝒢\) is not a GKAT coalgebra homomorphism; we will present a counter example:
% \begin{example}
%     We consider a empty set of primitive boolean \(B = ∅\), then the only atom is the constant true formula. And the example consists of the following transition function \(δ₂: 2 → \{\true\}→ G(2)\), where the states \(2 ≜ \{0, 1\}\):
%     \begin{mathpar}
%         δ₂(1, \true) = (0, p) \and δ₂(0, \true) = \reject
%     \end{mathpar}
%     Because both state \(0\) and \(1\) are dead, then \(⟦0⟧ = ⟦1⟧ = ∅\).
%     In order for \(⟦-⟧\) to be homomorphism, \(δ_𝒢 ∘ ⟦-⟧ = G(⟦-⟧) ∘ δ₂\) needs to be satisfied, however
%     \begin{align*}
%         (δ_𝒢 ∘ ⟦-⟧)(1) & = δ_𝒢(∅) = \{\true ↦ \reject\}.\\[10px]
%         (G(⟦-⟧) ∘ δ₂)(1) & = G(⟦-⟧)\{\true ↦ (0, p)\} \\
%             & = \{\true ↦ (⟦0⟧, p)\} = \{\true ↦ (∅, p)\}.
%     \end{align*}
%     Thus proving that \(⟦-⟧\) is not a GKAT coalgebra homomorphism.
% \end{example}

% Normalized homomorphism inherits many important properties of homomorphism; for example when composing two normalized homomorphism, we will obtain a normalized homomorphism. And the identity homomorphism is the identity of normalized homomorphism under composition.
% This notion gives rise to a

% \begin{definition}[normalized closure]\label{def:norm-closure}
%     Given two GKAT coalgebra \(S, T\), and a relation on these two expression \({∼} ⊆ S × T\), we can define \({\clos{∼}} ⊆ (2 + S × K) × (2 + T × K)\) as the smallest relation generated by the following rules: for all \(p, q ∈ K\) and states \(s' ∈ S\), \(t' ∈ T\):
%     \begin{mathpar}
%         \inferrule[Bisim-Acc]{\\}{\accept \clos{∼} \accept}\label{rule:bisim-acc}
%         \and 
%         \inferrule[Bisim-Rej]{\\}{\reject \clos{∼} \reject}\label{rule:bisim-rej}
%         \and 
%         \inferrule[Bisim-Trans]{s' ∼ t'}{(s', p) \clos{∼} (t', p)}\label{rule:bisim-trans}
%         \and
%         \inferrule[Norm-Left]{s' \text{ is dead}}{(s', p) \clos{∼} \reject}\label{rule:norm-left}
%         \and 
%         \inferrule[Norm-Right]{t' \text{ is dead}}{\reject \clos{∼} (t', p)}\label{rule:norm-right}
%         \and
%         \inferrule[Norm-Both]{s', t' \text{ are both dead}}{(s', p) \clos{∼} (t', q)}\label{rule:norm-both}
%     \end{mathpar}
% \end{definition}
% Intuitively, \({\clos{∼}}\) enlarges the relation \(∼\) for \emph{results} of a transition. For example, it allows transition to dead states to equivalent to rejection, and so on.

% This extension enables ``on-the-fly normalization'', where to determine whether \(s \clos{∼} t\), we can simply first check for bisimulation using the bisimulation rules; and if all failed, we can check the normalization rules.

% \begin{definition}
%     Given two GKAT coalgebra \(S\) and \(T\), a relation \({∼} ⊆ S × T\) is a normalized bisimulation when there exists a function \({∼} → {\clos{∼}}^\At\), such that the following diagram commutes:
%     \[
%         \begin{tikzcd}[column sep=1.5cm, row sep=1cm]
%             S \ar[swap]{d}{δ_S}
%                 & ∼  \ar[dashed, swap]{d}{∃} \ar[swap]{l}{π₁} \ar{r}{π₂}
%                 & T \ar{d}{δ_T} \\  
%             (2 + S × K)^\At 
%                 & ({\clos{∼}}^\At) 
%                     \ar[swap]{l}{{π₁}^\At} \ar{r}{{π₂}^\At}
%                 & (2 + T × K)^\At
%         \end{tikzcd}
%     \]
% \end{definition}
% Notice, unlike \(G\), \(\clos{(-)}^\At\) is not a functor, since it is only defined on relations.

% \begin{example}
%     Let's consider a example with a single primitive test \(B ≜ \{b\}\),
%     thus atoms are defined as \(\At_B = \{\{b\}, ∅\}\).
%     We focus on two states in the bisimulation \(s\) and \(t\).
%     with their respective transition defined as follows 
%     \begin{align*}
%             δ_S(s, \{b\}) & ≜ 0   
%                 & δ_S(t, \{b\}) & ≜ (t', p) \\  
%             δ_S(s, ∅) & ≜ (s', p) 
%                 & δ_S(t, ∅) & ≜ (t', p) 
%     \end{align*}
%     It is quite clear \(s\) and \(t\) will not be bisimular, however, there can be a normalized bisimulation relation between them, provided that \(s' ∼ t'\) and \(t'\) is dead: 
%     \[
%     \begin{tikzcd}[column sep=1.5cm, row sep=1cm]
%         s \ar[swap, mapsto]{d}{δ_S}
%             & (s, t)  \ar[dashed, swap, mapsto]{d}{∃} \ar[swap, mapsto]{l}{π₁} \ar[mapsto]{r}{π₂}
%             & t \ar[mapsto]{d}{δ_T} \\  
%         \begin{gathered}
%             \{b\}↦ 0 \\  
%             ∅ ↦ (s', p)
%         \end{gathered} 
%             & \begin{gathered}
%                 \{b\}↦ 0, (t', p) \\  
%                 ∅ ↦ (s', p), (t', p)
%             \end{gathered} 
%                 \ar[swap, mapsto]{l}{{π₁}^\At} \ar[mapsto]{r}{{π₂}^\At}
%             & \begin{gathered}
%                 \{b\} ↦ (t', p) \\  
%                 ∅ ↦(t', p)
%             \end{gathered}
%     \end{tikzcd}
%     \]
%     Notice that \(0 \clos{∼} (t', p)\) because \(t'\) is dead, 
%     and \((s', p) \clos{∼} (t', p)\) because we assume that \(s' ∼ t'\). 
%     Furthermore, if both \(s'\) and \(t'\) are dead, we can establish \(s' \clos{∼} t'\) using either the \textnormal{\textsc{Bisim-Trans}} rule or \textnormal{\textsc{Norm-Both}} rule, and we are free to pick the more efficient check in our implementation.
% \end{example}

% \begin{theorem}[Universiality]
%     Given a normalized bisimulation relation \({∼} ⊆ S × T\)
%     \[
%         \begin{tikzcd}[column sep=1.5cm, row sep=1cm]
%             S \ar[swap]{d}{δ_S}
%                 & ∼  \ar[dashed, swap]{d}{h} \ar[swap]{l}{π₁} \ar{r}{π₂}
%                 & T \ar{d}{δ_T} \\  
%             (2 + S × K)^\At 
%                 & (\clos{∼})^\At 
%                     \ar[swap]{l}{{π₁}^\At} \ar{r}{{π₂}^\At}
%                 & (2 + T × K)^\At
%         \end{tikzcd}
%     \] 
%     Then \(h\) is unique, and defined as \(δ_{S}(s) × δ_{T}(t)\).
% \end{theorem}

% \begin{proof}
%     The commutativity can be verified by computation; and the uniqueness can be observed using the property of pairs. We will show an alternative diagrammatical proof for reference.

%     The above commutative diagram implies the commutativity of the following diagram:
%     \[
%         \begin{tikzcd}
%             & {∼} \ar{d}{h} \ar[swap]{ddl}{δ_S ∘ π₁} \ar{ddr}{δ_T ∘ π₂}& \\
%             & (\clos{∼})^\At \ar[hook]{d}{i} & \\  
%             (2 + S × K)^\At & 
%                 (2 + S × K)^\At × (2 + T × K)^\At
%                     \ar{l}{π₁^\At} \ar[swap]{r}{π₂^\At}
%                 & (2 + T × K)^\At
%         \end{tikzcd}
%     \] 
%     Since \((-)^{\At}\) is a right adjoint and right adjoints preserve limits, \((2 + S × K)^\At × (2 + T × K)^\At\) is the product of \((2 + T × K)^\At\) and \((2 + T × K)^\At\). 
%     Hence, \(h ∘ i\) is the unique function to make the above diagram commute. 
%     Assume we can replace \(h\) with \(g\) while preserving the commutativity of the diagram: 
%     \begin{align*}
%         & i ∘ h  = i ∘ g & \text{\((2 + S × K)^\At × (2 + T × K)^\At\) is a product.} \\  
%         ⟹ {}& h = g & \text{\(i\) is a embedding, hence left cancelable.}
%     \end{align*}
% \end{proof}

% The above theorem is particularly important, as in order to check whether a relation \(∼\) is a bisimulation, we only need to check whether 
% \[s ∼ t ⟹ ∀ α ∈ \At, δ_S(s, α) \clos{∼} δ_T(t, α).\]

% A easy normalized bisimulation we can construct is between two dead states
% \begin{lemma}[normalized bisimulation between dead states]\label{thm:norm-bisim-dead-states}
%     Given two dead states \(s ∈ S\) and \(t ∈ T\), then the singleton \({∼} ≜ \{(s, t)\}\) is a normalized bisimulation between \(s\) and \(t\).
% \end{lemma}
% \begin{proof}
%     By~\cref{thm:dead-iff-all-reachable-dead,def:liveness-of-states}, dead states can only transition to either dead states or rejection. 
%     Thus, for all \(α ∈ \At\), \(δ_S(s, α) \clos{∼} δ_T(t, α)\) by \nameref{rule:norm-left}, \nameref{rule:norm-right}, or \nameref{rule:norm-both}.
% \end{proof}

% \begin{lemma}\label{thm:sub-coalg-preserve-norm-bisim}
%     Normalized bisimulation is preserved under sub-GKAT coalgebra. Given two sub-GKAT coalgebra \(S' ⊑ S\) and \(T' ⊑ T\) and a normalized bisimulation \(∼\) between \(S\) and \(T\), then the restriction 
%     \[∼_{S', T'} ≜ \{(s, t) ∣ s ∈ S', t ∈ T', s ∼ t\}\]
%     is a normalized bisimulation on between \(S'\) and \(T'\).
%     Similarly, every normalized bisimulation \(∼_{S', T'}\) between \(S'\) and \(T'\) is also a normalized bisimulation between \(S\) and \(T\).
% \end{lemma}

% Thus, this gives us a way to algorithmically construct a normalized bisimulation on principle sub-GKAT coalgebra.
% This normalized bisimulation can also be extended to a normalized bisimulation on the entire coalgebras.

% \begin{theorem}[inductive construction]\label{thm:inductive-construction}
%     Given two GKAT coalgebra \(S\) and \(T\), and two of their elements \(s ∈ S\) and \(t ∈ T\),
%     there exists a normalized bisimulation \({∼} ⊆ ⟨s⟩ × ⟨t⟩\) s.t. \(s ∼ t\), if and only if either both \(s\) and \(t\) are dead or for all \(α ∈ \At\), all of the following holds:
%     \begin{enumerate}
%         \item \(δ_{S}(s, α) = \accept ⟺ δ_{T}(t, α) = \accept\);
%         \item\label{itm:transition-condition} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{T}(t, α) = (t', p)\), then there exists a normalized bisimulation \({∼_{s',t'}}\) on \(⟨s'⟩\) and \(⟨t'⟩\), s.t. \(s' ∼_{s',t'} t'\);
%         \item If \(δ_{S}(s, α) = (s', p)\) and \(δ_{T}(t, α) = (t', q)\), s.t. \(p ≠ q\), then both \(s'\) and \(t'\) are dead;
%         \item \(s\) reject \(α\) or transition to a dead state via \(α\) if and only if \(t\) rejects \(α\) or transition to a dead state via \(α\).
%     \end{enumerate}
% \end{theorem}

% \begin{proof}
%     We first prove the \(⟹\) direction: if there exists a normalized bisimulation \(∼\), then for all \(α ∈ \At\), \(δ_{S}(s, α) \clos{∼} δ_{T}(t, α)\). And by unfolding the definition of \(\clos{∼}\), we can obtain all the conditions we desired. 
%     We will show condition \labelcref{itm:transition-condition} as an example:

%     Because \(∼\) is a normalized bisimulation: \[(s', p) \clos{∼} (t', p).\]
%     This relation can be obtained by two rules: \nameref{rule:bisim-trans} and \nameref{rule:norm-both}:
%     \begin{itemize}
%         \item if \((s', p) \clos{∼} (t', p)\) is obtained from \nameref{rule:bisim-trans}, then \(s' ∼ t'\); hence the bisimulation \(∼_{s',t'}\) can be obtained from restricting \(∼\) to \(⟨s'⟩\) and \(⟨t'⟩\)
%         \item if \((s', p) \clos{∼} (t', p)\) is obtained from \nameref{rule:norm-both}, then by \cref{thm:norm-bisim-dead-states}, there is indeed a noramlized bisimulation where \(⟨s'⟩ ∼_{s',t'} ⟨t'⟩\).
%     \end{itemize}

%     For the \(⟸\) direction, we can explicitly construct \(∼\) as the union of all the normalized bisimulations \(∼_{s', t'} ⊆ ⟨s'⟩ × ⟨t'⟩\), for all reachable states pair \(s', t'\) from state \(s\) and \(t\): 
%     \[{∼} = \{(s, t)\} ∪ \{{∼_{s', t'}} ∣ ∃ α ∈ \At, p ∈ K, δ_S(s, α) = (s', p) \text{ and } δ_T(t, α) = (t', p)\}.\]
%     Then, we will need to show \(∼\) is indeed a normalized bisimulation on \(⟨s⟩\) and \(⟨t⟩\), that is for all \(s' ∈ ⟨s⟩\) and \(t' ∈ ⟨t⟩\):
%     \[s' ∼ t' ⟹ δ_S(s', α) \clos{∼} δ_T(t', α).\]
%     We only need to show the above implication for \(s ∼ t\), as other element is already in some normalized bisimulation between sub-GKAT coalgebras of \(S\) and \(T\).
%     And the implication \(s ∼ t ⟹ δ_S(s', α) \clos{∼} δ_T(t', α)\) can be proven by unfolding the definition of \(\clos{∼}\).
% \end{proof}

% \Cref{thm:inductive-construction} provides a way to inductively construct a normalized bisimulation. At this point, we can already write down a algorithm for normalized bisimulation; however, with more properties developed in the next section, we will be able to devise a more efficient algorithm.

% \subsection{soundness and completeness}

% In this section we will further investigate the properties of normalized bisimulation. Not only will we prove the soundness and completeness of normalized bisimulation between and we will also identify several special cases where the normalized bisimulation algorithm can be further optimized.

% We can constraint both \(s\) and \(t\) to live states, and obtain a bisimulation on normalized coalgebra \(\norm(S)\) and \(\norm(T)\).
% With a slight abuse of notation, we can define the bisimulation \(\norm(∼) ⊆ \norm(S) × \norm(T)\) as follows:
% \begin{align*}
%     \norm(∼) & ⊆ \norm(S) × \norm(T) \\
%     \norm(∼) & ≜ \{(s, t) ∣ s ∼ t, s ∈ \norm(S), t ∈ \norm(T)\}
% \end{align*}
% and because \(∼\) is a normalized bisimulation, we have
% \[s ∼ t ⟹ ∀ α ∈ \At, δ(s, α) \clos{∼} δ(t, α).\]
% We proceed to define the transition \(\norm(δ_∼): \norm(∼) → G(\norm(∼))\) by case analysis on the elements of \(\clos{∼}\) (\Cref{def:norm-closure}):
% \begin{itemize}
%     \item If both \(δ(s, α)\) and \(δ(t, α)\) accepts, then \[\norm(δ_∼)((s,t), α) = \accept.\]
%     \item If both \(δ(s, α)\) and \(δ(t, α)\) either reject or transition to dead states, then \[\norm(δ_∼)((s,t), α) = \reject.\]
%     \item If both \(δ(s, α)\) and \(δ(t, α)\) transition to live states \(s, t\) respectively executing the same action \(p\), then \[\norm(δ_∼)((s,t), α) = ((s', t'), p).\]
% \end{itemize}
% Then the following theorem can be proven by computation.

% \begin{theorem}\label{thm:norm-bisim-to-bisim-on-norm-coalg}
%     For all normalized bisimulation \({∼} ⊆ S × T\), \(\norm(∼)\) is a bisimulation between \(\norm(S)\) and \(\norm(T)\). 
%     This means that for all live states \(s ∈ S\) and \(t ∈ T\),
%     \[s ∼ t ⟹ ⟦s⟧_S = ⟦t⟧_T\]
% \end{theorem}

% \Cref{thm:norm-bisim-to-bisim-on-norm-coalg} establishes the soundness for live state, and the soundness of normalized bisimulation on dead states are trivial, as all dead states have empty traces, hence are finite trace equivalent by definition.

% The final case for soundness is that \(s ∼ t\), where one of \(s\) and \(t\) is dead, and the other is live. We will show that this case is impossible.

% \begin{lemma}\label{thm:surj-on-norm-princ-coalg}
%     Let \(T\) be a normal GKAT coalgebra, and \(S\) be any GKAT coalgebra,
%     then a normalized bisimulation \({∼} ⊆ S × T\) will cover all the elements of \(⟨t⟩\) if it covers \(t\), for any \(t ∈ T\).
%     Formally, 
%     \[t ∈ π₂(∼) ⟹ ⟨t⟩ ⊆ π₂(∼)\]
% \end{lemma}

% \begin{proof}
%     Because normalized bisimulation, we can obtain the following diagram:
%     \[
%         \begin{tikzcd}[column sep=1cm, row sep=1cm]
%             S \ar[swap]{d}{δ_S}
%                 & ∼ \ar{d}{(δ_S, δ_T)} \ar[swap]{l}{π₁} \ar{r}{π₂}
%                 & T \ar{d}{δ_T} \\  
%             G(S) 
%                 & {\clos{∼}}^\At 
%                     \ar[swap]{l}{{π₁}^\At} \ar{r}{{π₂}^\At}
%                 & G(T)
%         \end{tikzcd}
%     \]
%     We then perform the following epi-mono factorization on the right square:
%     \[
%         \begin{tikzcd}[column sep=1cm, row sep=1cm]
%             ∼ \ar[swap]{d}{(δ_S, δ_T)} \ar[twoheadrightarrow]{r}{π₂}
%                 & π₂(∼) \ar[hook]{r}{i} \ar{d}{δ_T}
%                 & T \ar{d}{δ_T} \\  
%             {\clos{∼}}^\At \ar[twoheadrightarrow]{r}{π₂^{\At}}
%                 & G(π₂(∼)) \ar[hook]{r}{G(i)}
%                 & G(T) 
%         \end{tikzcd}
%     \]
%     The above diagram is indeed well-defined, specifically it suffice to show that \[∀ (r₁, r₂) ∈ {\clos{∼}}, π₂(r₁, r₂) = r₂ ∈ 2 + π₂(∼) × K.\] 
%     The above inclusion can be proven by case analysis on the input in \(\clos{∼}\). Specifically, given that \(T\) is a normal GKAT coalgebra, thus elements in \(\clos{∼}\) can only be obtained by \nameref{rule:bisim-acc}, \nameref{rule:bisim-rej}, \nameref{rule:bisim-trans}, and \nameref{rule:norm-left} rule.
%     \begin{itemize}
%         \item If the input \((r₁, r₂)\) of \(π₂\) is obtained by \nameref{rule:bisim-acc}, then \[r₂ = \accept ∈ 2 + π₂(∼) × K.\]
%         \item If the input \((r₁, r₂)\) of \(π₂\) is obtained by \nameref{rule:bisim-rej} or \nameref{rule:norm-left}, then \[r₂ = \reject ∈ 2 + π₂(∼) × K.\]
%         \item If the input \((r₁, r₂)\) of \(π₂\) is obtained by \nameref{rule:bisim-trans}, then \(r₁\) and \(r₂\) can be expressed as \((s, p)\) and \((t, p)\) respectively, and \(s ∼ t\). Thus we have \(t ∈ π₂(∼)\) and \[r₂ = (t, p) ∈ 2 + π₂(∼) × K.\]
%     \end{itemize}

%     Then we can lift \(π₂: {\clos{∼}} → 2 + π₂(∼) × K\) over functor \((-)^{\At}\) to obtain the well-defined function \(π₂^{\At}: {\clos{∼}^{\At}} → G(π₂(∼))\), and \(δ_T: π₂(∼) → G(π₂(∼))\) is well-defined by the commutativity of the diagram.

%     Finally, the above diagram tells us that \(π₂(∼)\) forms a sub-GKAT coalgebra of \(T\). Since \(⟨t⟩_T\) is the smallest sub-GKAT coalgebra tha contains \(t\), 
%     \[t ∈ π₂(∼) ⟹ t ⊑ π₂(∼) ⟹ t ⊆ π₂(∼).\]
% \end{proof}

% \begin{lemma}\label{thm:norm-preserve-norm-bisim}
%     Given a normalized bisimulation \({∼} ⊆ S × T\), the following restricted relation \(\norm ₗ(∼) ⊆ S × \norm(T)\) is also a normalized bisimulation:
%     \[\norm ₗ(∼) = \{s, t ∣ s ∼ t, s ∈ S, t ∈ \norm(T)\}.\]
% \end{lemma}

% \begin{proof}
%     Because \(∼\) is a normalized bisimulation, and \(\norm ₗ(∼) ⊆ {∼}\),
%     \[(s, t) ∈ \norm ₗ(∼) ⟹ s ∼ t ⟹ ∀ α ∈ \At, δ_S(s, α) ∼ δ_T(t, α).\]
%     And we can prove that 
%     \[δ_S(s, α) ∼ δ_T(t, α) ⟹ (δ_S(s, α), \norm(δ_T)(t, α)) ∈ \norm ₗ(∼):\]
%     \begin{itemize}
%         \item If \(δ_T(t, α)\) transition to a dead state, then \(δ_S(s, α)\) will either reject or transition to a dead state, and \(\norm(δ_T)(t, α)\) will reject, thus: 
%         \[(δ_S(s, α), \norm(δ_T)(t, α)) = (δ_S(s, α), \reject) ∈ \norm ₗ(∼).\]
%         \item If \(δ_T(t', α)\) doesn't transition to a dead state, then \( \norm(δ_T)(t', α) = δ_T(t', α)\), and 
%         \[(δ_S(s, α), \norm(δ_T)(t, α)) = (δ_S(s, α), δ_T(t, α)) ∈ \norm ₗ(∼).\]
%     \end{itemize}
%     Combining the two implications above, we have 
%     \[(s, t) ∈ \norm ₗ(∼) ⟹ ∀ α ∈ \At, (δ_S(s, α), \norm(δ_T)(t, α)) ∈ \norm ₗ(∼).\]
% \end{proof}

% \begin{theorem}\label{thm:dead-live-don't-norm-bisim}
%     Given a dead state \(s\) and a live state \(t\), there doesn't exists a normalized bisimulation \(∼\) s.t. \(s ∼ t\) or \(t ∼ s\).
% \end{theorem}

% \begin{proof}
%     Because \({∼} ⊆ S × T\) is a normalized bisimulation, then we can construct a restricted normalized bisimulation \(∼_{s, t} ⊆ ⟨s⟩ × ⟨t⟩\)by~\Cref{thm:sub-coalg-preserve-norm-bisim}; then we can normalize \(⟨t⟩\) to obtain the normalized bisimulation \(\norm ₗ(∼_{s, t}) ⊆ ⟨s⟩ × \norm(⟨t⟩)\) by~\Cref{thm:norm-preserve-norm-bisim}.

%     Since \(t\) is a live state, hence there exists an accepting state \(t'\) in \(\norm(⟨t⟩)\); because; \(s\) is a dead state, thus all the states in \(⟨s⟩\) is dead.

%     Because \(t\) is live, thus \[(s, t) ∈ \norm ₗ(∼_{s, t}),\]
%     and by~\Cref*{thm:surj-on-norm-princ-coalg}, every state in \(⟨t⟩\) is covered by \(\norm ₗ(∼_{s, t})\).
%     In other word, there exists a there exists a state \(s' ∈ ⟨s⟩\) that is normally bisimilar to the accepting state \(s' ∼ t'\).
%     However, recall that all states in \(⟨s⟩\) is dead (\Cref{thm:dead-iff-all-reachable-dead}) and a dead state cannot be normally bisimilar to a accepting state, since none of the transition (reject or transition to a dead state) of a dead state can be related to acceptance in \(\clos{\norm ₗ(∼_{s, t})}\).
% \end{proof}

% Thus we are able to show our soundness theorem:

% \begin{theorem}[soundness]
%     For all GKAT coalgebra \(S\) and \(T\), if there exists a  normalized bisimulation \({∼} ⊆ S × T\) s.t. \(s ∼ t\), then \(⟦s⟧_S = ⟦t⟧_T\).
% \end{theorem}

% \begin{proof}
%     We prove this theorem by case analysis on the liveness of \(s\) and \(t\), we have indeed covered all the cases by previous theorems:
%     \begin{itemize}
%         \item if both \(s\) and \(t\) are live then the implication holds by~\Cref{thm:norm-bisim-to-bisim-on-norm-coalg}; 
%         \item if one of \(s\) and \(t\) is live and the other is dead, then \(s ∼ t\) is necessarily false by~\cref{thm:dead-live-don't-norm-bisim}, thus the implication trivially holds;
%         \item if both \(s\) and \(t\) are dead, then they are finite trace equivalent, thus the implication holds.
%     \end{itemize}
% \end{proof}

% We then prove the completeness, which is the converse of the soundness statement, stating that the trace equivalence is ``completely'' characterized by normalized bisimulation. 
% Formally, when two states are finite trace equivalent, then there will be a normalized bisimulation between them.

% Our proof of completeness shares the core idea of as the corresponding completeness theorem~\cite{smolka_GuardedKleeneAlgebra_2020}, which is to construct a bisimulation from trace equivalence. 
% However, we streamline our proof using the language of category theory, in the process, we were also able to prove a more general result, that is, homomorphism reflects normalized bisimulation:

% \begin{theorem}[homomorphism reflects normalized bisimulation]\label{thm:hom-refl-norm-bisim}
%     Given a pair of homomorphism \(h_S: S → S'\) and \(h_T: T → T'\), and a normalized bisimulation \({∼'} ⊆ S' × T'\), then there exists a normalized bisimulation \({∼} ⊆ S × T\), s.t.
%     \[h_S(s) ∼' h_T(t) ⟺ s ∼ t.\]
% \end{theorem}

% \begin{proof}
%     We can define \(∼\) as the pre-image of \(∼'\) under \(h_S × h_T\):
%     \[{∼} ≜ \{(s, t) ∣ h_S(s) ∼' h_T(t)\}.\]
%     Similar characterization also apply to their closure counterpart:
%     \[{\clos{∼}} = \{(r₁, r₂) ∣ G(h_S)(r₁) \clos{∼'} G(h_S)(r₂)\}.\]
%     The above equation can be verified by case analysis on the elements of \(\clos{∼'}\) and \(\clos{∼}\):
%     \begin{itemize}
%         \item by \nameref{rule:bisim-acc} and \nameref{rule:bisim-rej}, \((\accept, \accept)\) and \((\reject, \reject)\) are in both \(\clos{∼}\) and \(\clos{∼'}\).
%         \item if the element is introduced by \nameref{rule:bisim-trans}, then 
%         \[(h(s), p) \clos{∼'} (h(t), p) 
%         ⟺ h(s) \clos{∼'} h(t) 
%         ⟺ s \clos{∼} t 
%         ⟺ (s, p) \clos{∼} (t, p).
%         \]
%         \item if the element is introduced by \nameref{rule:norm-left}, \nameref{rule:norm-right}, or \nameref{rule:norm-both}, we will recall that \(s\) is dead if and only if \(h(s)\) is dead (\Cref{thm:hom-preserve-liveness}). The reasoning is similar to previous case, we will show \nameref{rule:norm-left} case as an example:
%         \[(h(s), p) \clos{∼'} \reject
%         ⟺ h(s) \text{ is dead}
%         ⟺ s \text{ is dead}
%         ⟺ (s, p) \clos{∼} \reject.\]
%     \end{itemize}
%     by above equation about \(\clos{∼}\), 
%     we will be able to verify that \(∼\) is indeed a normalized bisimulation, specifically, given any two states \(s ∼ t\):
%     \begin{align*}
%         & s ∼ t \\ 
%         & ⟹ h_S(s) ∼' h_T(t) \\ 
%         & ⟹ ∀ α, δ_{S'}(h_S(s), α) \clos{∼'} δ_{T'}(h_T(t), α) \\  
%         & ⟹ ∀ α, (G(h_S) ∘ δ_S)(s, α) \clos{∼'} (G(h_T) ∘ δ_T)(t, α)
%             & \text{\(h\) is homomorphism} \\
%         & ⟹ ∀ α, δ_S(s, α) \clos{∼} δ_T(t, α) 
%             & \text{above equation about \(\clos{∼}\)}
%     \end{align*}
% \end{proof}

% \begin{lemma}
%     Given any bisimulation \(S\), the identity relation \(=\) on \(S\) is a a normalized bisimulation.
% \end{lemma}

% \begin{proof}
%     Because \(s = s' ⟹ ∀ α ∈ \At, δ(s, α) = δ(t, α)\), and the equality \(=\) on \(2 + S × K\) is a sub-relation of its normalized closure \(\clos{=}\), therefore \[s = s' ⟹ ∀ α ∈ \At, δ(s, α) \clos{=} δ(t, α)\]
% \end{proof}

% \begin{theorem}
%     Finite trace equivalence \({∼} ≜ \{(s, t) ∣ ⟦s⟧ = ⟦t⟧\}\) is a normalized bisimulation.
% \end{theorem}

% \begin{proof}
%     Take the identity relation \(=\) on the finite
% \end{proof}

\printbibliography

\newpage
\appendix

\section{diagrammatical characterization of normalized homomorphism}\label{sec:diagram-of-norm-sim}

\end{document}

### 2024-11-29 main.tex
update

@@
     Cheng Zhang \and 
+    Qiancheng Fu \and
     Hang Ji \and 
@@
 \end{abstract}
+
+TODO: change all the \(e_b\) to \(b\).
 
@@
 
-The soundness and completeness of this algorithm can be observed by the fact that \emph{when the algorithm terminate}, the algorithm returns true.
-if and only if there exists a bisimulation between \(⟨s⟩\) and \(⟨t⟩\) s.t.
+The soundness and completeness of~\cref{alg:bisim} can be observed by the fact that \emph{when the algorithm terminate}, the algorithm returns true if and only if there exists a bisimulation between \(⟨s⟩\) and \(⟨t⟩\) s.t.
 \(s ∼ t\), which is then logically equivalent to trace equivalence.
@@
 
-\section{Symbolic algorithm}
-
-\subsection{Symbolic GKAT Coalgebra}
-
-Given an alphabet \(K, B\), a \emph{symbolic GKAT coalgebra} \(Ŝ ≜ ⟨S, ϵ̂, δ̂⟩\) consists of a state set \(S\) consists of a accepting boolean \(ϵ̂\) and a transition function \(δ̂\):
+\begin{remark}
+    The caching of dead state and the shortcut to check whether \(s\) is dead when \(t\) is dead and vise versa, is not essential to the soundness and completeness of algorithm, they are here to trade speed with memory.
+    In a memory-constraint situation, the ``\textnormal{deadStates}'' variable can be cleared periodically to save memory.
+\end{remark}
+
+\section{Symbolic Algorithm}
+
+Given the alphabet \(K, B\), a \emph{symbolic GKAT coalgebra} \(Ŝ ≜ ⟨S, ϵ̂, δ̂⟩\) consists of a state set \(S\) and a accepting function \(ϵ̂\) and a transition function \(δ̂\):
 \begin{mathpar}
     ϵ̂: S → \Bool_B, \and
-    δ̂: S → S → K → \Bool_B,
+    δ̂: S → \Bool_B × S × K,
 \end{mathpar}
-where \(\Bool_B\) is the free boolean algebra over \(B\) (boolean expressions modulo boolean algebra axioms), and for all states \(s ∈ S\), all the booleans results are ``disjoint''; namely the conjunction of any two expression from the set \(ϵ̂(s) + \{δ̂(s, s', p) ∣ s' ∈ S, p ∈ K\}\) are false in \(\Bool_B\).
-The ordering on two symbolic coalgebra on the same carrier \(S\) is defined pointwise, namely 
-\[⟨ϵ̂₁, δ̂₁⟩ ≤ ⟨ϵ̂₂, δ₂⟩ \text{ if and only if } ∀ s, s' ∈ S, ∀ p ∈ K, \hat{ϵ₁}(s) ≤ \hat{ϵ₂}(s) \text{ and } \hat{δ₁}(s, s', p) ≤ \hat{δ₂}(s, s', p).\]
-
-Intuitively, the expressions in \(\Bool_B\) is treated as the ``symbolic'' transitions, which can be think of as a set of atoms by the isomorphism \(\Bool_B ≅ 2^{\At_B}\). 
-Then we can explain the intuition for the definition of symbolic GKAT coalgebra:
+where \(\Bool_B\) is the free boolean algebra over \(B\) (boolean expressions modulo boolean algebra axioms); for all states \(s ∈ S\), all the booleans are ``disjoint''; namely the conjunction of any two expression from the set \(\{ϵ̂(s)\} ∪ \{b ∣ ∃ (b, s', p) ∈ δ(s)\}\) are false.
+We will then use \(ρ̂(s): \Bool_B\) to denote the boolean expressions that contain all the atoms that the state \(s\) rejects, and \(ρ̂(s)\) can be computed as follows:
+\[ρ̂(s) ≜ ¬ ϵ̂(s) ∨ ¬ \left( ⋁_{(b, s', p) ∈ δ(s)} b \right)\]
+
+Instead of modeling each atom individually in the automata, we group them into boolean expressions, this leads to a much more space efficient automata, and enables efficient bisimulation algorithms using off-the-shelf SAT solvers.
+
+With the above intuition in mind, a symbolic GKAT coalgebra \(Ŝ ≜ ⟨S, ϵ̂, δ̂⟩\) can be lowered into a GKAT coalgebra \(⟨S, δ⟩\) in the following manner:
+\[
+δ(s, α) ≜ \begin{cases}
+    \accept & α ≤ ϵ̂(s) \\  
+    (s', p) & ∃ b ∈ \Bool_B, α ≤ b \text{ and } δ(s, b) = (s', p) \\  
+    \reject & \text{otherwise}
+\end{cases}
+\]
+This is well-defined, i.e. no more than one clause can be satisfied precisely because the boolean expressions appear in \(ϵ̂\) and \(δ̂\) are disjoint.
+The trace semantics of a GKAT coalgebra \(⟨S, ϵ̂, δ̂⟩\) is then defined as the trace semantics of its lowering \(⟨S, δ⟩\).
+
+\begin{remark}[Canonicity]
+    Notice that symbolic GKAT coalgebra is not canonical, i.e. there exists two different symbolic GKAT colagebra with the same lowering, consider the state set \(S ≜ \{*\}\):
+    \[{δ̂}₁(*) ≜ \{b ↦ (*, p), ¬ b ↦ (*, p)\} \qquad 
+    {δ̂}₂(*) ≜ \{⊤ ↦ (*, p)\},\] 
+    and both \(ϵ̂\) will return constant \(⊥\).
+    These two symbolic GKAT coalgebra obviously have the same lowering hence behavior, yet, they are different.
+    There are other symbolic representation that will satisfy canonicity, yet we opt to use our current representation for ease of construction and computational efficiency.
+\end{remark}
+
+We can then migrate the normalized bisimulation algorithm to the symbolic setting, we will first prove an inductive construction theorem like~\cref{thm:inductive-construction}.
+
+\begin{theorem}[Symbolic Inductive Construction]\label{thm:symb-inductive-construction}
+    Given two symbolic GKAT coalgebra \(Ŝ = ⟨S, ϵ̂_S, δ̂_S⟩\) and \(T̂ = ⟨T, ϵ̂_T, δ̂_T⟩\) and two states \(s ∈ S\) and \(t ∈ T\), there exists a normalized bisimulation on the lowered coalgebra \({∼} ⊆ S × T\) s.t.
+\(s ∼ t\) if and only if all the following holds:
     \begin{itemize}
-    \item \(ϵ̂(s)\) can be thought of as all the atoms that is accepted by the state \(s\).
-    \item \(δ̂(s, s', p)\) denotes all the atoms that can transition from \(s\) to \(s'\) while executing \(p\).
+        \item \(ϵ̂_S(s) ≡ ϵ̂_T(t)\);
+        \item for all \((b, s', p) ∈ δ̂_S(s)\) and \((c, t', q) ∈ δ̂_T(t)\), if \(b ∧ c ≢ 0\) and \(p = q\) then here exists a normalized bisimulation \({∼_{s',t'}} ⊆ S × T\) s.t. \(s' ∼_{s',t'} t'\);
+        \item for all \((b, s', p) ∈ δ̂_S(s)\) and \((c, t', q) ∈ δ̂_T(t)\), if \(b ∧ c ≢ 0\) and \(p ≠ q\) then both \(s'\) and \(t'\) is dead;  
+        \item for all \((b, s', p) ∈ δ̂_S(s)\) and \(c ∈ ρ̂_T(t)\), if \(b ∧ c ≢ 0\), then \(s'\) is dead;
+        \item for all \(b ∈ ρ̂_S(s)\) and \((c, t', q) ∈ δ̂_T(t)\), if \(b ∧ c ≢ 0\), then \(t'\) is dead;
     \end{itemize}
-Then the atoms excluded by both operations are implicitly rejecting. Formally, we can define a boolean expression that contains all the atoms that will be rejected 
-\[ρ̂(s) = ¬ (ϵ̂(s) ∨ ⋁ \{b ∣ ∃ s' ∈ S, p ∈ K, δ̂(s, s', p) = b\}).\]
-
-This intuition gives us a isomorphism between between GKAT coalgebra and symbolic GKAT coalgebra.
+\end{theorem}
+
+\begin{proof}
+    Reduces to~\cref{thm:inductive-construction} i.e. all the above condition holds if and only if all the condition in~\cref{thm:inductive-construction} holds in the lowered coalgebra.
+\end{proof}
+
+% \section{Symbolic algorithm BAK}
+
+% \subsection{Symbolic GKAT Coalgebra}
+
+% Given an alphabet \(K, B\), a \emph{symbolic GKAT coalgebra} \(Ŝ ≜ ⟨S, ϵ̂, δ̂⟩\) consists of a state set \(S\) consists of a accepting boolean \(ϵ̂\) and a transition function \(δ̂\):
+% \begin{mathpar}
+%     ϵ̂: S → \Bool_B, \and
+%     δ̂: S → S → K → \Bool_B,
+% \end{mathpar}
+% where \(\Bool_B\) is the free boolean algebra over \(B\) (boolean expressions modulo boolean algebra axioms), and for all states \(s ∈ S\), all the booleans results are ``disjoint''; namely the conjunction of any two expression from the set \(ϵ̂(s) + \{δ̂(s, s', p) ∣ s' ∈ S, p ∈ K\}\) are false in \(\Bool_B\).
+% The ordering on two symbolic coalgebra on the same carrier \(S\) is defined pointwise, namely 
+% \[⟨ϵ̂₁, δ̂₁⟩ ≤ ⟨ϵ̂₂, δ₂⟩ \text{ if and only if } ∀ s, s' ∈ S, ∀ p ∈ K, \hat{ϵ₁}(s) ≤ \hat{ϵ₂}(s) \text{ and } \hat{δ₁}(s, s', p) ≤ \hat{δ₂}(s, s', p).\]
+
+% Intuitively, the expressions in \(\Bool_B\) is treated as the ``symbolic'' transitions, which can be think of as a set of atoms by the isomorphism \(\Bool_B ≅ 2^{\At_B}\). 
+% Then we can explain the intuition for the definition of symbolic GKAT coalgebra:
+% \begin{itemize}
+%     \item \(ϵ̂(s)\) can be thought of as all the atoms that is accepted by the state \(s\).
+%     \item \(δ̂(s, s', p)\) denotes all the atoms that can transition from \(s\) to \(s'\) while executing \(p\).
+% \end{itemize}
+% Then the atoms excluded by both operations are implicitly rejecting. Formally, we can define a boolean expression that contains all the atoms that will be rejected 
+% \[ρ̂(s) = ¬ (ϵ̂(s) ∨ ⋁ \{b ∣ ∃ s' ∈ S, p ∈ K, δ̂(s, s', p) = b\}).\]
+
+% This intuition gives us a isomorphism between between GKAT coalgebra and symbolic GKAT coalgebra.
 Formally, given a GKAT coalgebra \(⟨S, δ⟩\), then its corresponding symbolic GKAT coalgebra will share the same states \(S\), with the transition defined as follows:
-\begin{align*}
-    ϵ̂(s) & ≜ ⋁ \{α ∈ \At ∣ δ(s, α) = \accept\} & 
-    δ̂(s, s', p) & ⋁ \{α ∈ \At ∣ δ(s, α) = (s', p)\}.
-\end{align*}
-We will denote the above map as \(\symb\), i.e.
+% \begin{align*}
+%     ϵ̂(s) & ≜ ⋁ \{α ∈ \At ∣ δ(s, α) = \accept\} & 
+%     δ̂(s, s', p) & ⋁ \{α ∈ \At ∣ δ(s, α) = (s', p)\}.
+% \end{align*}
+% We will denote the above map as \(\symb\), i.e.
 \(\symb(⟨S, δ⟩) ≜ ⟨S, ϵ̂, δ̂⟩\).
@@
 
-The inverse of \(\symb\) is denoted as \(\unsymb\) where given a symbolic GKAT coalgebra \(⟨S, ϵ̂, δ̂⟩\), we construct a GKAT coalgebra over states \(S\) with the following transition:
-\[δ(s, α) ≜ \begin{cases}
-    \accept & α ∈ ϵ̂(s) \\  
-    (s', p) & α ∈ δ̂(s, s', p) \\  
-    \reject & \text{otherwise}
-\end{cases}\]
-Notice the \(δ\) is well-defined, that is given any \(s ∈ S\) and \(α ∈ \At\), \(δ(s,α)\) can only take on one value; because symbolic GKAT coalgebra is deterministic.
-
-\subsection{Decision Procedure Of Symbolic GKAT coalgebra}
-
-The formulation of symbolic GKAT coalgebra is mathematically nice, yet inefficient.
+% The inverse of \(\symb\) is denoted as \(\unsymb\) where given a symbolic GKAT coalgebra \(⟨S, ϵ̂, δ̂⟩\), we construct a GKAT coalgebra over states \(S\) with the following transition:
+% \[δ(s, α) ≜ \begin{cases}
+%     \accept & α ∈ ϵ̂(s) \\  
+%     (s', p) & α ∈ δ̂(s, s', p) \\  
+%     \reject & \text{otherwise}
+% \end{cases}\]
+% Notice the \(δ\) is well-defined, that is given any \(s ∈ S\) and \(α ∈ \At\), \(δ(s,α)\) can only take on one value; because symbolic GKAT coalgebra is deterministic.
+
+% \subsection{Decision Procedure Of Symbolic GKAT coalgebra}
+
+% The formulation of symbolic GKAT coalgebra is mathematically nice, yet inefficient.
 Between most states \(s, s' ∈ S\) and action \(p ∈ K\), \(δ(s, s', p)\) will always be false.
 Therefore, we can developed a set representation of the derivative, where we will typically denote using \(Δ̂\):
-\[Δ̂: S → 𝒫(\Bool_B × S × K).\]
-we always assume transition functions denoted this way are \emph{deterministic}; specifically, all the boolean expression in the following set are disjoint: \[\{b ∣ ∃ s' ∈ S, p ∈ K, (b, s', p) ∈ Δ̂(s)\}.\]
-
-Intuitively, if \((b, s', p) ∈ Δ̂(s)\), then \(s\) will transition to \(s'\) for all the atoms \(α ≤ b\) while executing \(p\).
+% \[Δ̂: S → 𝒫(\Bool_B × S × K).\]
+% we always assume transition functions denoted this way are \emph{deterministic}; specifically, all the boolean expression in the following set are disjoint: \[\{b ∣ ∃ s' ∈ S, p ∈ K, (b, s', p) ∈ Δ̂(s)\}.\]
+
+% Intuitively, if \((b, s', p) ∈ Δ̂(s)\), then \(s\) will transition to \(s'\) for all the atoms \(α ≤ b\) while executing \(p\).
 Notice that the above statement does not imply that \emph{all} \(α\) that makes \(s\) transition to \(s'\) while executing \(p\) will be covered by \(b\); that is, there can exists a \((b', s', p) ∈ Δ̂(s)\), where \(b ≠ b'\).
 
-This means that the set representation of a symbolic GKAT coalgebra no on longer uniquely correspond to a GKAT coalgebra.
-\begin{example}
-
-\end{example}
-
-However, every set representation of a symbolic GKAT coalgebra, indeed correspond to a GKAT coalgebra, specifically given a set representation \(⟨S, ϵ̂, Δ̂⟩\), we can construct a symbolic GKAT \(⟨S, ϵ̂, δ̂⟩\), by defining the following \(δ̂\):
-\[δ̂(s, s', p) = ⋁ \{b ∣ (b, s', p) ∈ Δ̂(s)\}.\]
-More importantly, the map described above is surjective. This means that all the GKAT coalgebra can indeed be represented as at least one set-representation.
-Similarly, set representation can also be converted to GKAT coalgebra: given a set representation \(⟨S, ϵ̂, Δ̂⟩\), we can lower it into a GKAT coalgebra \(⟨S, δ⟩\), by constructing the following \(δ\):
-\[δ(s, α) ≜ \begin{cases}
-    \accept & α ∈ ϵ̂(s) \\  
-    (s', p) & ∃ b ≥ α, (b, s', p) ∈ Δ̂(s) \\  
-    \reject & \text{otherwise}
-\end{cases}\]
-Indeed, lowering the set representation directly to GKAT coalgebra give the same result as converting to symbolic GKAT coalgebra and then lowering.
+% This means that the set representation of a symbolic GKAT coalgebra no on longer uniquely correspond to a GKAT coalgebra.
+% \begin{example}
+
+% \end{example}
+
+% However, every set representation of a symbolic GKAT coalgebra, indeed correspond to a GKAT coalgebra, specifically given a set representation \(⟨S, ϵ̂, Δ̂⟩\), we can construct a symbolic GKAT \(⟨S, ϵ̂, δ̂⟩\), by defining the following \(δ̂\):
+% \[δ̂(s, s', p) = ⋁ \{b ∣ (b, s', p) ∈ Δ̂(s)\}.\]
+% More importantly, the map described above is surjective. This means that all the GKAT coalgebra can indeed be represented as at least one set-representation.
+% Similarly, set representation can also be converted to GKAT coalgebra: given a set representation \(⟨S, ϵ̂, Δ̂⟩\), we can lower it into a GKAT coalgebra \(⟨S, δ⟩\), by constructing the following \(δ\):
+% \[δ(s, α) ≜ \begin{cases}
+%     \accept & α ∈ ϵ̂(s) \\  
+%     (s', p) & ∃ b ≥ α, (b, s', p) ∈ Δ̂(s) \\  
+%     \reject & \text{otherwise}
+% \end{cases}\]
+% Indeed, lowering the set representation directly to GKAT coalgebra give the same result as converting to symbolic GKAT coalgebra and then lowering.
 This fact is essential for us to relate these semantics.
 
-Finally, we can lift the inductive construction theorem (\cref{thm:inductive-construction}) to set representation of GKAT coalgebras, giving a sound and complete symbolic decision procedure. 
-In the following lemmas, we will use \(S, T\) to denote some GKAT coalgebra, and \(Ŝ, T̂\) will be as some set-represented symbolic GKAT coalgebra that corresponds to \(S\). 
-Formally, if we lower \(Ŝ\) to GKAT coalgebra, we will obtain exactly \(S\); similarly for \(T̂\).
-
-\begin{lemma}
-    We let the \(s ∈ S\) and \(t ∈ T\) be two states, then the following equivalence between conditions in \cref{thm:inductive-construction} hold: 
-    \begin{itemize}
-        \item condition~\labelcref{itm:acc-condition} in inductive construction holds if and only if \(ϵ̂_{Ŝ}(s) = ϵ̂_{T̂}(t)\).
-        \item condition~\labelcref{itm:transition-bisim} holds if and only if for all \((a, s', p) ∈ δ̂_{Ŝ}(s)\) and \((b, t', q) ∈ δ̂_{T̂}(t)\):
-        \[a ∧ b ≠ 0 \text{ and } p = q ⟹ \text{exists a bisimulation \(∼\) between \(S\) and \(T\), } s.t. s' ∼ t'.\]
-        \item condition~\labelcref{itm:transition-dead}, holds if and only if for all \((a, s', p) ∈ δ̂_{Ŝ}(s)\) and \((b, t', q) ∈ δ̂_{T̂}(t)\):
-        \[a ∧ b ≠ 0 \text{ and } p ≠ q ⟹ \text{both \(s\) and \(t\) are dead}.\]
-        \item condition~\labelcref{itm:rej-or-dead} holds if and only if the following holds:
-    \end{itemize}
-\end{lemma}
-
-\begin{lemma}
-    Given two states \(\)
-\end{lemma}
-
-\subsection{Symbolic Derivatives and Correctness}
-
-Given a symbolic GKAT coalgebra \(⟨S, δ̂, ϵ̂⟩\) and a state \(s ∈ S\), we use the notation \(s ⇒ b\) to denote that \(ϵ̂(s) ≥ b\), and we use the notation \(s \transvia{b ∣ p} s'\) to denote that \(δ̂(s, s', p) ≥ b\), where the ordering is the typical ordering in boolean algebra.
-
-The symbolic derivatives for GKAT forms a DKCT where the states are represented by GKAT expression \(\GKAT_{K, B}\), and the boolean algebra is the free boolean algebra \(\Bool_B\) over the test alphabet \(B\):
-\begin{mathpar}
-    ϵ: \GKAT_{K, B} → \Bool_B; \and  
-    δ: \GKAT_{K, B} → \GKAT_{K, B} → K → \Bool_B.
-\end{mathpar} 
-the accepting map \(ϵ\) and transition map \(δ\) can be defined by the smallest maps that satisfy the following rules:
-\begin{mathpar}
-    \inferrule{\\}{b ⇒ b} \and 
-    \inferrule{\\}{p \transvia{1 ∣ p} 1} \and 
-    \inferrule{e ⇒ a}{e +_b f ⇒ b a} \and 
-    \inferrule{f ⇒ a}{e +_b f ⇒ b̄ a} \and 
-    \inferrule{e \transvia{a ∣ p} e'}{e +_b f \transvia{b ∧ a ∣ p} e'} \and 
-    \inferrule{f \transvia{a ∣ p} f'}{e +_b f \transvia{b̄ ∧ a ∣ p} f'} \and 
-    \inferrule{e ⇒ a \\ f ⇒ b}{e ⋅ f ⇒ a ∧ b} \and 
-    \inferrule{e ⇒ a \\ f \transvia{b ∣ p} f'}{e ⋅ f \transvia{a ∧ b ∣ p} f'} \and 
-    \inferrule{e \transvia{b ∣ p} e'}{e ⋅ f \transvia{b ∣ p} e' ⋅ f} \and 
-    \inferrule{\\}{e^{(b)} ⇒ b̄} \and 
-    \inferrule{e \transvia{a ∣ p} e'}{e^{(b)} \transvia{b ∧ a ∣ p} e' ⋅ {e}^{(b)}}
-\end{mathpar}
-Notice the rules above are very similar to the derivative rules for GKAT~\cite{schmid_GuardedKleeneAlgebra_2021}, this is no coincidence, as we can establish a correspondence of these two to prove correctness of our symbolic algorithm.
-
-Since the rules listed above is monotonic, that is if we enlarge the boolean in the premise, we also enlarge the boolean in the result. 
-So we can simply pick the largest boolean on the premise, which is either \(ϵ(e)\) or \(δ(e, e', p)\).
-Since the disjunction of two boolean \(a ∨ b\) is the smallest boolean that is greater than both \(a\) and \(b\), thus we can implement \(ϵ\) and \(δ\) by iteration through all the rules, and take the disjunction of all the possible booleans.
-
-\begin{align*}
-    ϵ(b) & ≜ b 
-        &  ϵ(e +_b f) & ≜ (b ∧ ϵ(e)) ∨ (b̄ ∧ ϵ(f))\\
-    ϵ(q) & ≜ 0
-        & ϵ(e ⋅ f) & ≜ ϵ(e) ∧ ϵ(f)\\  
-    ϵ(e^{(b)}) & ≜ b̄ 
-\end{align*}
-
-However, the derivative function \[δ: \GKAT_{K, B} → \GKAT_{K, B} → K → \Bool_B,\] poses challenges in the implementation, since \(\GKAT_{K, B}\) is infinite, so it will be impractical to search through all the expressions.
-However, we can treat the function \(δ\) extensionally: \[\GKAT_{K, B} → K → \Bool_B ⊆ 2^{\GKAT_{K, B} × K × \Bool_B},\]
-and then because the boolean expression don't overlap for each input, hence the boolean expression are necessarily unequal; 
-we can model the set by a partial map, thus all the \(δ\) can be represented as a function to the following partial maps.
-\begin{align*}
-    δ & : \GKAT_{K, B} → (\Bool_B ↛ \GKAT_{K, B} × K) \\
-    δ(b) & ≜ \{\} \\
-    δ(q) & ≜ \{1 ↦ (1, p)\}\\  
-    δ(e +_b f) & ≜ \{b ∧ a ↦ (e', p) ∣ a ↦ (e', p) ∈ δ(e)\} 
-        ∪ \{b̄ ∧ a ↦ (f', p) ∣ a ↦ (f', p) ∈ δ(f)\}\\
-    δ(e ⋅ f) & ≜ \{b ↦ (e' ⋅ f, p) ∣ b ↦ (e, p) ∈ δ(e)\}
-        ∪ \{ϵ(e) ∧ b ↦ (f', p) ∣ b ↦ (f', p) ∈ δ(f)\}\\
-    δ(e^{(b)}) & ≜ \{b ∧ a ↦ (e' ⋅ e^{(b)}, p) ∣ a ↦ (e', p) ∈ δ(e)\}.
-\end{align*}
-FIXME: we still need to think about what representation do we want to go with, with this representation, you can also have duplicated K.
+% Finally, we can lift the inductive construction theorem (\cref{thm:inductive-construction}) to set representation of GKAT coalgebras, giving a sound and complete symbolic decision procedure. 
+% In the following lemmas, we will use \(S, T\) to denote some GKAT coalgebra, and \(Ŝ, T̂\) will be as some set-represented symbolic GKAT coalgebra that corresponds to \(S\). 
+% Formally, if we lower \(Ŝ\) to GKAT coalgebra, we will obtain exactly \(S\); similarly for \(T̂\).
+
+% \begin{lemma}
+%     We let the \(s ∈ S\) and \(t ∈ T\) be two states, then the following equivalence between conditions in \cref{thm:inductive-construction} hold: 
+%     \begin{itemize}
+%         \item condition~\labelcref{itm:acc-condition} in inductive construction holds if and only if \(ϵ̂_{Ŝ}(s) = ϵ̂_{T̂}(t)\).
+%         \item condition~\labelcref{itm:transition-bisim} holds if and only if for all \((a, s', p) ∈ δ̂_{Ŝ}(s)\) and \((b, t', q) ∈ δ̂_{T̂}(t)\):
+%         \[a ∧ b ≠ 0 \text{ and } p = q ⟹ \text{exists a bisimulation \(∼\) between \(S\) and \(T\), } s.t. s' ∼ t'.\]
+%         \item condition~\labelcref{itm:transition-dead}, holds if and only if for all \((a, s', p) ∈ δ̂_{Ŝ}(s)\) and \((b, t', q) ∈ δ̂_{T̂}(t)\):
+%         \[a ∧ b ≠ 0 \text{ and } p ≠ q ⟹ \text{both \(s\) and \(t\) are dead}.\]
+%         \item condition~\labelcref{itm:rej-or-dead} holds if and only if the following holds:
+%     \end{itemize}
+% \end{lemma}
+
+% \begin{lemma}
+%     Given two states \(\)
+% \end{lemma}
+
+% \subsection{Symbolic Derivatives and Correctness}
+
+% Given a symbolic GKAT coalgebra \(⟨S, δ̂, ϵ̂⟩\) and a state \(s ∈ S\), we use the notation \(s ⇒ b\) to denote that \(ϵ̂(s) ≥ b\), and we use the notation \(s \transvia{b ∣ p} s'\) to denote that \(δ̂(s, s', p) ≥ b\), where the ordering is the typical ordering in boolean algebra.
+
+% The symbolic derivatives for GKAT forms a DKCT where the states are represented by GKAT expression \(\GKAT_{K, B}\), and the boolean algebra is the free boolean algebra \(\Bool_B\) over the test alphabet \(B\):
+% \begin{mathpar}
+%     ϵ: \GKAT_{K, B} → \Bool_B; \and  
+%     δ: \GKAT_{K, B} → \GKAT_{K, B} → K → \Bool_B.
+% \end{mathpar} 
+% the accepting map \(ϵ\) and transition map \(δ\) can be defined by the smallest maps that satisfy the following rules:
+% \begin{mathpar}
+%     \inferrule{\\}{b ⇒ b} \and 
+%     \inferrule{\\}{p \transvia{1 ∣ p} 1} \and 
+%     \inferrule{e ⇒ a}{e +_b f ⇒ b a} \and 
+%     \inferrule{f ⇒ a}{e +_b f ⇒ b̄ a} \and 
+%     \inferrule{e \transvia{a ∣ p} e'}{e +_b f \transvia{b ∧ a ∣ p} e'} \and 
+%     \inferrule{f \transvia{a ∣ p} f'}{e +_b f \transvia{b̄ ∧ a ∣ p} f'} \and 
+%     \inferrule{e ⇒ a \\ f ⇒ b}{e ⋅ f ⇒ a ∧ b} \and 
+%     \inferrule{e ⇒ a \\ f \transvia{b ∣ p} f'}{e ⋅ f \transvia{a ∧ b ∣ p} f'} \and 
+%     \inferrule{e \transvia{b ∣ p} e'}{e ⋅ f \transvia{b ∣ p} e' ⋅ f} \and 
+%     \inferrule{\\}{e^{(b)} ⇒ b̄} \and 
+%     \inferrule{e \transvia{a ∣ p} e'}{e^{(b)} \transvia{b ∧ a ∣ p} e' ⋅ {e}^{(b)}}
+% \end{mathpar}
+% Notice the rules above are very similar to the derivative rules for GKAT~\cite{schmid_GuardedKleeneAlgebra_2021}, this is no coincidence, as we can establish a correspondence of these two to prove correctness of our symbolic algorithm.
+
+% Since the rules listed above is monotonic, that is if we enlarge the boolean in the premise, we also enlarge the boolean in the result. 
+% So we can simply pick the largest boolean on the premise, which is either \(ϵ(e)\) or \(δ(e, e', p)\).
+% Since the disjunction of two boolean \(a ∨ b\) is the smallest boolean that is greater than both \(a\) and \(b\), thus we can implement \(ϵ\) and \(δ\) by iteration through all the rules, and take the disjunction of all the possible booleans.
+
+% \begin{align*}
+%     ϵ(b) & ≜ b 
+%         &  ϵ(e +_b f) & ≜ (b ∧ ϵ(e)) ∨ (b̄ ∧ ϵ(f))\\
+%     ϵ(q) & ≜ 0
+%         & ϵ(e ⋅ f) & ≜ ϵ(e) ∧ ϵ(f)\\  
+%     ϵ(e^{(b)}) & ≜ b̄ 
+% \end{align*}
+
+% However, the derivative function \[δ: \GKAT_{K, B} → \GKAT_{K, B} → K → \Bool_B,\] poses challenges in the implementation, since \(\GKAT_{K, B}\) is infinite, so it will be impractical to search through all the expressions.
+% However, we can treat the function \(δ\) extensionally: \[\GKAT_{K, B} → K → \Bool_B ⊆ 2^{\GKAT_{K, B} × K × \Bool_B},\]
+% and then because the boolean expression don't overlap for each input, hence the boolean expression are necessarily unequal; 
+% we can model the set by a partial map, thus all the \(δ\) can be represented as a function to the following partial maps.
+% \begin{align*}
+%     δ & : \GKAT_{K, B} → (\Bool_B ↛ \GKAT_{K, B} × K) \\
+%     δ(b) & ≜ \{\} \\
+%     δ(q) & ≜ \{1 ↦ (1, p)\}\\  
+%     δ(e +_b f) & ≜ \{b ∧ a ↦ (e', p) ∣ a ↦ (e', p) ∈ δ(e)\} 
+%         ∪ \{b̄ ∧ a ↦ (f', p) ∣ a ↦ (f', p) ∈ δ(f)\}\\
+%     δ(e ⋅ f) & ≜ \{b ↦ (e' ⋅ f, p) ∣ b ↦ (e, p) ∈ δ(e)\}
+%         ∪ \{ϵ(e) ∧ b ↦ (f', p) ∣ b ↦ (f', p) ∈ δ(f)\}\\
+%     δ(e^{(b)}) & ≜ \{b ∧ a ↦ (e' ⋅ e^{(b)}, p) ∣ a ↦ (e', p) ∈ δ(e)\}.
+% \end{align*}
+% FIXME: we still need to think about what representation do we want to go with, with this representation, you can also have duplicated K.
 We need a good example here of why don't we go with map.
@@
 
-\subsection{Algorithm and Complexity}
-
-\section{Implementation and Experiments}
+% \subsection{Algorithm and Complexity}
+
+% \section{Implementation and Experiments}
 

### 2024-12-02 main.tex
symbolic algorithm

@@
 
+TODO: we should merge the two so that it is easier to 
 \begin{algorithm}
@@
                 \Match{\(δ_{S}(s, α)\)}
-                \Case{\(\accept\)} {\Return none} \Comment{\(s\) transition to accept}\EndCase
+                \Case{\(\accept\)} {\Return none} \Comment{\(s\) transition to accept}
+                \EndCase
+                \Case{\(\reject\)} {\texttt{continue}}
+                \Comment{skip if \(s\) transition to reject}
+                \EndCase
                 \Case{($s', p$)}{ 
-                    \If{\Call{IsDead}{s'} = none} {\Return none} \Comment{\(s\) transitions to a live state \(s'\)}
+                    \If{\Call{IsDeadLoop}{$s'$} = none} {\Return none} \Comment{\(s\) transitions to a live state \(s'\)}
                     \Else { explored \(←\) (explored \(∪\) \Call{isDeadLoop}{$s'$, explored}) } \EndIf
@@
         \If {\Call{eq}{$s, t$}} {\Return true}
-        \ElsIf {\(s ∈\) deadStates} {\Return \Call{isDead}{$t$}} 
-        \ElsIf {\(t ∈\) deadStates} {\Return \Call{isDead}{$s$}} 
+        \ElsIf {\(s ∈\) deadStates\(_S\)} {\Return \Call{isDead\(_T\)}{$t$}} 
+        \ElsIf {\(t ∈\) deadStates\(_T\)} {\Return \Call{isDead\(_S\)}{$s$}} 
         \Else 
@@
 
+Because the dead state detection algorithm is coalgebra-specific, we use a subscript on ``deadStates'' and ``\textsc{IsDead}'' to indicate the coalgebra. 
 The soundness and completeness of~\cref{alg:bisim} can be observed by the fact that \emph{when the algorithm terminate}, the algorithm returns true if and only if there exists a bisimulation between \(⟨s⟩\) and \(⟨t⟩\) s.t.
@@
 \end{proof}
+
+Then for the algorithm, we can just recursively check all the conditions in~\cref{thm:symb-inductive-construction}.
+
+\begin{algorithm}
+    \caption{On-the-fly bisimulation algorithm}\label{alg:bisim}
+    \begin{algorithmic}
+        \Function{equiv}{$s ∈ S, t ∈ T$}
+        \If {\Call{eq}{$s, t$}} {\Return true}
+        \ElsIf {\(s ∈\) deadStates\(_S\)} {\Return \Call{isDead\(_T\)}{$t$}} 
+        \ElsIf {\(t ∈\) deadStates\(_T\)} {\Return \Call{isDead\(_S\)}{$s$}} 
+        \Else {}
+        \Return {
+            \Comment{conditions of ~\cref{thm:symb-inductive-construction}}
+            \\\vspace{5px}
+            \(\qquad
+            \begin{aligned}
+                & ϵ̂_S(s) ≡ ϵ̂_T(t) \mathrel{\&\!\&} \\  
+                & \forall (b, s', p) ∈ δ̂_S(s), (c, t', q) ∈ δ̂_T(t), (b ∧ c) ≢ ⊥ ⟹ 
+                \begin{cases}
+                    \text{\Call{IsDead$_S$}{$s$}} ∧ \text{\Call{IsDead$_T$}{$t$}} & \text{if \(p ≠ q\)} \\
+                    \text{\Call{Union}{$s$, $t$}}; \text{\Call{Equiv}{$s', t'$}} & \text{if \(p = q\)}
+                \end{cases} \mathrel{\&\!\&}\\
+                & \forall (b, s', p) ∈ δ̂_S(s), c ∈ ρ̂_T(t), (b ∧ c) ≢ ⊥ ⟹ \text{\Call{IsDead$_S$}{$s'$}} \mathrel{\&\!\&}\\
+                & \forall b ∈ ρ̂_S(s), (c, t', q) ∈ ρ̂_T(t), (b ∧ c) ≢ ⊥ ⟹ \text{\Call{IsDead$_T$}{$t'$}}
+            \end{aligned}\)
+        }
+        \EndIf
+        \EndFunction
+    \end{algorithmic}
+\end{algorithm}
+
+Inspired by the syntax of Ocaml, the \(\mathrel{\&\!\&}\) is the logical-and operator on the language level, specifying that all four conditions in the return statements must be satisfied to return true.
+Notice just like the non-symbolic case, this algorithm can be modified to perform symbolic bisimulation of (non-normalized) GKAT automaton, which coincides with infinite trace equivalence~\cite{schmid_GuardedKleeneAlgebra_2021}, by letting \textsc{IsDead} always return false and keep deadStates empty.
+
+\section{Construction of Symbolic GKAT Automata}
+
+We will go through two ways of constructing symbolic GKAT automata, one is Thompson's construction, the other is derivative.
+However, in order to exhibit the relation between these two construction, specifically, the derivative construction always produces a smaller automaton than the Thompson's construction, we opt to use to produce "weak" symbolic GKAT automata: symbolic GKAT automata where the transition no longer need to perform an action.
+
+\subsection{Weak Symbolic GKAT Automata And Tightening}
+
+Weak Symbolic GKAT automata is a symbolic GKAT automata that allow internal transitions via a boolean expression:
+\[⟨S, 
+ϵ̂: S → \Bool, 
+δ̂: S → (\Bool × S) + (\Bool × S × K)⟩.\]
+We also require the deterministic criteria, all the boolean transitions of \(s\), i.e. all the boolean expressions in the following set:
+\[\{ϵ̂(s)\} ∪ 
+\{b ∣ ∃ s' ∈ S, (b, s') ∈ δ̂(s)\} ∪ 
+\{b ∣ ∃ s' ∈ S, p ∈ K, (b, s', p) ∈ δ̂(s)\},\]
+are disjoint.
+Notice that our notion of
 

### 2024-12-17 main.tex
update

@@
 }
+\newcommand{\transAcc}[2]{⇒_{#1} #2}
 
@@
 As we will see later, bisimulation or variation of bisimulation indeed correspond to important semantical equivalences.
+
+\begin{theorem}
+    Given a functionr \(F\) on set, two \(F\)-colagebra \(S\) and \(T\), if there exists a surjective homomorphism \(h: S → T\), then \(S\) and \(T\) are bisimular.
+\end{theorem}
 
@@
 \begin{proof}
-    Because homomorphic image preserves principle sub-GKAT coalgebra~\cref{thm:homo-img-preserve-principle-sub-coalg}
+    Because homomorphic image preserves principle sub-GKAT coalgebra (\cref{thm:homo-img-preserve-principle-sub-coalg})
     \[h(⟨s⟩_S) = ⟨h(s)⟩_T;\]
@@
     \begin{itemize}
-        \item If both \(s'\) and \(t'\) are live, then \(s' ∼ t'\), the bisimulation \(∼_{s', t'}\) is just \(∼\) restricted to \(⟨s'⟩\) and \(⟨t'⟩\).
+        \item If both \(s'\) and \(t'\) are live, then \(s' ∼ t'\). By~\cref{thm:sub-coalg-preserve-bisim}, the bisimulation \(∼_{s', t'}\) is just \(∼\) restricted to \(⟨s'⟩\) and \(⟨t'⟩\).
         \item If both \(s'\) and \(t'\) are dead, then \(∼_{s', t'}\) can just be the singleton relation, according to~\cref{thm:bisim-between-dead}.
@@
 \begin{mathpar}
-    ϵ̂: S → \Bool_B, \and
-    δ̂: S → \Bool_B × S × K,
+    ϵ̂: S → 𝒫(\Bool_B), \and
+    δ̂: S → 𝒫(\Bool_B × S × K),
 \end{mathpar}
@@
 δ(s, α) ≜ \begin{cases}
-    \accept & α ≤ ϵ̂(s) \\  
+    \accept & ∃ b ∈ ϵ̂(s), α ≤ b \\  
     (s', p) & ∃ b ∈ \Bool_B, α ≤ b \text{ and } δ(s, b) = (s', p) \\  
@@
     \begin{itemize}
-        \item \(ϵ̂_S(s) ≡ ϵ̂_T(t)\);
+        \item \(⋁ ϵ̂_S(s) ≡ ⋁ ϵ̂_T(t)\);
         \item for all \((b, s', p) ∈ δ̂_S(s)\) and \((c, t', q) ∈ δ̂_T(t)\), if \(b ∧ c ≢ 0\) and \(p = q\) then here exists a normalized bisimulation \({∼_{s',t'}} ⊆ S × T\) s.t. \(s' ∼_{s',t'} t'\);
@@
             \begin{aligned}
-                & ϵ̂_S(s) ≡ ϵ̂_T(t) \mathrel{\&\!\&} \\  
+                & ⋁ ϵ̂_S(s) ≡ ⋁ ϵ̂_T(t) \mathrel{\&\!\&} \\  
                 & \forall (b, s', p) ∈ δ̂_S(s), (c, t', q) ∈ δ̂_T(t), (b ∧ c) ≢ ⊥ ⟹ 
@@
                 & \forall (b, s', p) ∈ δ̂_S(s), c ∈ ρ̂_T(t), (b ∧ c) ≢ ⊥ ⟹ \text{\Call{IsDead$_S$}{$s'$}} \mathrel{\&\!\&}\\
-                & \forall b ∈ ρ̂_S(s), (c, t', q) ∈ ρ̂_T(t), (b ∧ c) ≢ ⊥ ⟹ \text{\Call{IsDead$_T$}{$t'$}}
+                & \forall b ∈ ρ̂_S(s), (c, t', q) ∈ δ̂_T(t), (b ∧ c) ≢ ⊥ ⟹ \text{\Call{IsDead$_T$}{$t'$}}
             \end{aligned}\)
@@
 
-\subsection{Weak Symbolic GKAT Automata And Tightening}
+\subsection{Weak Symbolic GKAT Automata, Star System, and Tightening}
 
@@
 \[⟨S, 
-ϵ̂: S → \Bool, 
-δ̂: S → (\Bool × S) + (\Bool × S × K)⟩.\]
+ϵ̂: S → 𝒫(\Bool), 
+δ̂: S → 𝒫(\Bool × S + \Bool × S × K)⟩.\]
 We also require the deterministic criteria, all the boolean transitions of \(s\), i.e. all the boolean expressions in the following set:
-\[\{ϵ̂(s)\} ∪ 
+\[ϵ̂(s) ∪ 
 \{b ∣ ∃ s' ∈ S, (b, s') ∈ δ̂(s)\} ∪ 
@@
 are disjoint.
-Notice that our notion of
+To ease the notation, in a symbolic coalgebra \(Ŝ\) we will write 
+\begin{itemize}[nosep]
+    \item \(s \transAcc{S}{b}\) to mean that \(b ∈ ϵ̂(s)\).
+    \item \(s \transvia{b}_{Ŝ} s'\) to mean that \((b, s') ∈ δ̂(s)\);  
+    \item and \(s \transvia{b ∣ p}_{Ŝ} s'\) to mean that \((b, s', p) ∈ δ̂(s)\).
+\end{itemize}
+We will call transitions (to states) that do not emit actions \emph{unproductive}, and similarly, ones that emit actions will be named \emph{productive}.
+Specifically, transitions of the form \(s \transvia{b}_{Ŝ} s'\) are unproductive, and \(s \transvia{b ∣ p}_{Ŝ} s'\) represents productive transitions.
+Notice that our notion of empty transitions are different from previous work~\cite{silva_CoalgebraicViewETransitions_2013a,bonchi_KillingEpsilonsDagger_2015a,sokolova_WeakBisimulationActionType_2005}, the empty transitions, i.e.
+\(δ̂(s) ∈ \Bool × S\), although do not execute any action, but will require a boolean formula to transition.
+Furthermore, the boolean formula has a natural meet semi-lattice structure given by the logical and operator.
+In this section, we will design a specific ``unproductive transition removal'' procedure.
+
+% \begin{definition}[Star Systems]
+%     Given a weak symbolic GKAT coalgebra \(Ŝ ≜ ⟨S, ϵ̂, δ̂⟩\) a star system \(Ŝ^* ≜ ⟨S, ϵ̂^*, δ̂^*⟩\) is a system that contains exactly the transitions generated by the following rules:
+%     \begin{mathpar}
+%         \inferrule[Acc]
+%         {s \transAcc{Ŝ}{b}}
+%         {s \transAcc{Ŝ^*}{b}}
+%         \and
+%         \inferrule[StarAcc]
+%         {s₁ \transvia{b}_{Ŝ} s₂ \\ s₂ \transAcc{Ŝ^*}{c}}
+%         {s₁ \transAcc{Ŝ^*}{b ∧ c}}
+%         \and
+%         \inferrule[Unprod]
+%         {s₁ \transvia{b}_{Ŝ} s₂}
+%         {s₁ \transvia{b}_{Ŝ^*} s₂}
+%         \and 
+%         \inferrule[StarUnprod]
+%         {s₁ \transvia{b}_{Ŝ} s₂ \\ s₂ \transvia{d}_{Ŝ^*} s₃}
+%         {s₁ \transvia{b ∧ d}_{Ŝ^*} s₂}
+%         \and 
+%         \inferrule[Prod]
+%         {s₁ \transvia{b ∣ p}_{Ŝ} s₂}
+%         {s₁ \transvia{b ∣ p}_{Ŝ^*} s₂}
+%         \and 
+%         \inferrule[StarProd]
+%         {s₁ \transvia{b}_{Ŝ} s₂ \\ s₂ \transvia{c ∣ p}_{Ŝ^*} s₃}
+%         {s₁ \transvia{b ∧ c ∣ p}_{Ŝ^*} s₂}
+%     \end{mathpar}
+
+%     The objects in \emph{the category of star system} are weak symbolic GKAT coalgebra and the morphisms are homomorphisms between star systems.
+% \end{definition}
+
+% Intuitively, star systems contain \emph{all} the possible transitions both productive and unproductive from a state, until an action is reached.
+% Notice that a star system will not necessarily be a weak symbolic GKAT coalgebra, because it will no longer be deterministic.
+% One exception is symbolic GKAT coalgebra, i.e.also weak symbolic systems that doesn't have unproductive transitions, where their star system will just be themselves, because only the \textsc{Acc} and \textsc{Prod} rules can be applied.
+% Hence the homomorphisms between symbolic GKAT coalgebra in the category of star systems are exactly the standard coalgebra homomorphism of symbolic GKAT coalgebra.
+
+% \begin{theorem}[Universal Property]
+%     Take two  GKAT coalgebra \(Ŝ\) and \(Û\), then 
+% \end{theorem}
+
+% \begin{proof}
+%     Assume \(s \transAcc{Ŝ^*}{b}\), then 
+% \end{proof}
+
+% \begin{definition}[Tightening]
+%     For a weak symbolic GKAT coalgebra \(Ŝ\), its tightening \(Ŝ\!↓\) is \(Ŝ^*\) with all the unproductive transitions removed.
+Formally let \(Ŝ^* ≜ ⟨S, ϵ̂^*, δ̂^*⟩\), then \(Ŝ\!↓ ≜ ⟨S, ϵ̂\!↓, δ̂\!↓⟩\) is defined as follows:
+%     \begin{mathpar}
+%         ϵ̂\!↓(s) ≜ ϵ̂^*(s) \and
+%         δ̂\!↓(s) ≜ \{(b, s', p) ∣ (b, s', p) ∈ δ̂^*(s)\} 
+%     \end{mathpar}
+% \end{definition}
+
+% \begin{example}
+
+% \end{example}
+
+% \begin{lemma}
+%     The tightening \(Ŝ\!↓\) of \(Ŝ\) contains exactly the transitions formed by appling \textsc{Acc}, \textsc{StarAcc}, \textsc{Prod}, and \textsc{StarProd} rules to transitions of \(Ŝ\).
+% \end{lemma}
+
+% \begin{proof}
+%     First all the transitions formed by the four rules will not contain non-productive transition, thus will be contained in the tightening.
+
+%     Second, consider the derivation tree for a productive transition in \(Ŝ^*\) like \(s₁ \transvia{b ∣ p}_{Ŝ^*} s₂\), the \textsc{Unprod} and \textsc{StarUnprod} cannot be the last derivation, since they do not produce a productive transition; and their result cannot be used by any other rules.
+Thus, \textsc{Unprod} and \textsc{StarUnprod} rule cannot be applied to derive a productive transition.
+
+%     Hence, all the transitions of \(Ŝ^*\), i.e. all the productive transitions of \(Ŝ^*\), are generated exactly by \textsc{Acc}, \textsc{StarAcc}, \textsc{Prod}, and \textsc{StarProd} rules.
+% \end{proof}
+
+% \begin{theorem}
+%     For any weak symbolic GKAT coalgebra \(Ŝ\), its tightening \(Ŝ\!↓\) is a symbolic GKAT coalgebra.
+% \end{theorem}
+
+% \begin{proof}
+%     By definition, the tightened system \(Ŝ\!↓\) only contains productive transition. Thus, we only need to verify the disjointness condition.
+
+% \end{proof}
+
+% We define the \emph{category of star systems} where the objects are weak symbolic GKAT coalgebras, and the homomorphisms between them are standard coalgebra homomorphism between their star systems.
+
+% \begin{theorem}[Universal Property]\label{thm:tight-universal-prop}
+%     For all weak symbolic GKAT coalgebra \(Ŝ\), and symbolic GKAT coalgebra \(Û\!↓\), following universal property holds in the category of star system:
+%     \[
+%     \begin{tikzcd}[sep=1.25cm]
+%         Û\!↓ \ar[dashed,swap]{d}{f} \ar{dr}{f}& \\  
+%         Ŝ\!↓ \ar[swap]{r}{γ_{Ŝ}} & Ŝ
+%     \end{tikzcd}
+%     \]
+%     where \(γ_{Ŝ}\) maps each state in \(Ŝ\) to itself.
+% \end{theorem}
+
+% \begin{proof}
+%     We will need to prove three items: first \(γ_{Ŝ}\) is a star system homomorphism, then \(h\) exists and is unique.
+
+%     By definition, tightening contains all the productive and accepting transitions of the star system,
+%     therefore if \(s \transAcc{Ŝ\!↓}{b}\) then \(s \transAcc{Ŝ^*}{b}\);
+%     similarly, \(s₁ \transvia{b ∣ p}_{Ŝ\!↓} s₂\) implies \(s₁ \transvia{b ∣ p}_{Ŝ^*} s₂\).
+%     Thus, \(γ_{Ŝ}\), which maps each state to itself, is a coalgebra homomorphism the star systems.
+
+%     Because \(γ_{Ŝ}\) maps every state to itself, thus \(f\) is the unique function that can make the diagram commute. 
+%     We just need to verify that \(f: Û\!↓ → Ŝ\!↓\) satisfy the rules of a star-system homomorphism.
+%     \begin{align*}
+%         & u₁ \transAcc{Û\!↓}{b} 
+%         ⟹ u₁ \transAcc{Û^*}{b} 
+%         ⟹ f(u₁) \transAcc{Ŝ^*}{b}
+%         ⟹ f(u₁) \transAcc{Ŝ\!↓}{b} \\   
+%         & u₁ \transvia{b ∣ p}_{Û\!↓} u₂
+%         ⟹ u₁ \transvia{b ∣ p}_{Û^*} u₂
+%         ⟹ f(u₁) \transvia{b ∣ p}_{Ŝ^*} u₂
+%         ⟹ f(u₁) \transvia{b ∣ p}_{Ŝ\!↓} u₂
+%     \end{align*}
+% \end{proof}
+
+% Finally, because the star system of a symbolic GKAT coalgebra is just itself, then the coalgebra homomorphisms between them are all symbolic GKAT coalgebra homomorphisms, therefore we obtain the following adjunction theorem.
+
+% \begin{corollary}
+%     The tightening \((-)\!↓\) is a functor and is right adjoint to the embedding from symbolic GKAT coalgebra to star systems.
+% \end{corollary}
+
+% \begin{proof}
+%     This is mostly a standard result of category theory, but for the sake of completeness, we will recreate the result.
+%     Given a weak symbolic GKAT coalgebra homomorphism \(f: Û → Ŝ\), it is also a homomorphism of their star systems by definition, then \(f\!↓: Û\!↓ → Ŝ\!↓\) is the unique arrow that satisfy the following diagram:
+%     \[
+%     \begin{tikzcd}[sep=1cm]
+%         Û\!↓ \ar{r}{γ_U} \ar[dashed,swap]{d}{f\!↓} & Û \ar{d}{f}\\  
+%         Ŝ\!↓ \ar{r}{γ_S} & Ŝ 
+%     \end{tikzcd}
+%     \]
+%     Because both \(γ_U\) and \(γ_S\) maps states to itself, therefore \(f\!↓\) is necessarily point-wise equal to \(f\); and the preservation of identity and composition can be verified by the universal property.
+%     Finally, the adjunction is a concequence of~\Cref{thm:tight-universal-prop}.
+% \end{proof}
+
+% The functoriality of tightening is particularly important as it not only means all the homomorphism between the weak coalgebra is also a homomorphism between their tightened version, but also means that it will preserve injectivity and surjectivity as these maps are functions between set.
+% In the next section, we will prove the derivatives are more ``efficient'' than Thomposon's construction by proving there exists a surjective homomorphism from the automata created by derivarives to the automata created by Thompson's construction.
+% Such a homomorphism will also be a witness of the bisimularity of the derivatives and Thompson's construction, aiding us in the proof of correctness.
+
+\begin{figure*}
+    \begin{mathpar}
+        \inferrule[Acc]
+        {s₁ \transAcc{Ŝ}{b}}
+        {s₁ \transAcc{Ŝ↓}{b}}
+        \and
+        \inferrule[TightenAcc]
+        {s₁ \transvia{b}_{Ŝ} s₂ \\ s₂ \transAcc{Ŝ↓}{c}}
+        {s₁ s₂ \transAcc{Ŝ↓}{b ∧ c}}
+        \and
+        \inferrule[Trans]
+        {s₁ \transvia{b ∣ p}_{Ŝ} s₂}
+        {s₁ \transvia{b ∣ p}_{Ŝ↓} s₂}
+        \and
+        \inferrule[TightenTrans]
+        {s₁ \transvia{b}_{Ŝ} s₂ \\ s₂ \transvia{c ∣ p}_{Ŝ↓} s₃}
+        {s₁ \transvia{b ∧ c ∣ p}_{Ŝ↓} s₃}
+    \end{mathpar}
+    \caption{Tightening Rules}\label{fig:tightening-rules}
+\end{figure*}
+
+\begin{definition}
+    Given a weak symbolic GKAT coalgebra \(Ŝ ≜ ⟨S, ϵ̂, δ̂⟩\), its tightening \(Ŝ\!↓ ≜ ⟨S, ϵ̂\!↓, δ̂\!↓⟩\) is a symbolic GKAT coalgebra s.t.
+\(ϵ̂\!↓\) and \(δ̂\!↓\) are the point-wise least (ordered by set inclusion) function that satisfy the rules in~\cref{fig:tightening-rules} for all \(s₁, s₂, s₃ ∈ S\),
+\end{definition}
+
+In particular, infinite loop of unproductive transition will be resolved to rejection, by the definition above.
+\begin{example}
+
+\end{example}
+
+One of the helpful observation is that tightening is a functor from the category of weak symbolic GKAT coalgebra into the category of symbolic GKAT colagebra.
+The functoriality us to establish isomorphism on the weak case and automatically transfer it to symbolic GKAT coalgebra.
+
+\begin{theorem}[Functoriality]\label{thm:tight-functor-induction}
+    Given two weak symbolic GKAT coalgebra \(Ŝ, T̂\) and a homomorphism between them \(h: Ŝ → T̂\), then \(h\) is also a homomorphism on the tightening \(h: Ŝ\!↓ → T̂\!↓\).
+\end{theorem}
+
+\begin{proof}
+    We will need to show that \(h\) preserves both \(ϵ̂\) and \(δ̂\!↓\). 
+    Since the tightening do not change \(ϵ̂\), thus the perseverance of \(ϵ̂\) is trivial.
+    Then \(h\) preserves \(δ̂\!↓\) can be shown in two case: consider a transition \(s \transvia{b ∣ p}_{Ŝ\!↓} s'\),
+    then there exists a path from \(s\) to \(s'\) in \(Ŝ\) which produced this transition.
+    Thus, we can prove the perseverance by induction on the length of the path.
+\end{proof}
+
+% \subsection{An Aside, The Categorical Structure of Tightening}
+
+% In this section we will explore the categorical structure of tightening, where we will show the categorical structure of tightening. 
+% In particular, we will not only show that \((-)\!↓\) is a functor, but a right adjoint in the appropriate category, without using induction on path as in~\cref{thm:tight-functor-induction}.
+% To exhibit the connection between tightened coalgebra and the weak coalgebra, we will need a weaker version of homomorphism. 
+% However, before that, we will need to define an alternative way to manipulate weak symbolic colagebra
+
+% \begin{definition}[Star Systems]
+%     Given a weak symbolic GKAT coalgebra \(Ŝ ≜ ⟨S, ϵ̂, δ̂⟩\) a star system \(Ŝ^* ≜ ⟨S, ϵ̂^*, δ̂^*⟩\) is a system that contains exactly the transitions generated by the following rules:
+%     \begin{mathpar}
+%         \inferrule[Acc]
+%         {s \transAcc{Ŝ}{b}}
+%         {s \transAcc{Ŝ^*}{b}}
+%         \and
+%         \inferrule[StarAcc]
+%         {s₁ \transvia{b}_{Ŝ} s₂ \\ s₂ \transAcc{Ŝ^*}{b}}
+%         {s₁ \transAcc{Ŝ^*}{b}}
+%         \and
+%         \inferrule[Unprod]
+%         {s₁ \transvia{b}_{Ŝ} s₂}
+%         {s₁ \transvia{b}_{Ŝ^*} s₂}
+%         \and 
+%         \inferrule[StarUnprod]
+%         {s₁ \transvia{b}_{Ŝ} s₂ \\ s₂ \transvia{d}_{Ŝ^*} s₃}
+%         {s₁ \transvia{b ∧ d}_{Ŝ^*} s₂}
+%         \and 
+%         \inferrule[Prod]
+%         {s₁ \transvia{b ∣ p}_{Ŝ} s₂}
+%         {s₁ \transvia{b ∣ p}_{Ŝ^*} s₂}
+%         \and 
+%         \inferrule[StarProd]
+%         {s₁ \transvia{b}_{Ŝ} s₂ \\ s₂ \transvia{c ∣ p}_{Ŝ^*} s₃}
+%         {s₁ \transvia{b ∧ c ∣ p}_{Ŝ^*} s₂}
+%     \end{mathpar}
+% \end{definition}
+
+% Intuitively, star systems contain \emph{all} the possible transitions both productive and unproductive from a state, until an action is reached.
+% Notice that a star system will not necessarily be a weak symbolic GKAT coalgebra, because it will no longer be deterministic.
+% One exception to these are symbolic GKAT coalgebra, i.e.
+weak symbolic systems that doesn't have unproductive transitions, where their star system will just be themselves, because only the \textsc{Acc} and \textsc{Prod} rules can be applied.
+% Hence the homomorphisms between symbolic GKAT coalgebra in the category of star systems are exactly the standard coalgebra homomorphism of symbolic GKAT coalgebra
+
+% We define the \emph{category of star systems} where the objects are weak symbolic GKAT coalgebras, and the homomorphisms between them are standard coalgebra homomorphism between their star systems.
+
+% \begin{example}
+%     some example about \(γ_{Ŝ}\)
+% \end{example}
+
+% \begin{theorem}[Universal Property]
+%     For all weak symbolic GKAT coalgebra \(Ŝ\), and symbolic GKAT coalgebra \(T̂\!↓\), following universal property holds in the category of star system:
+%     \[
+%     \begin{tikzcd}[sep=1.25cm]
+%         T̂\!↓ \ar[dashed,swap]{d}{f} \ar{dr}{f}& \\  
+%         Ŝ\!↓ \ar[swap]{r}{γ_{Ŝ}} & Ŝ
+%     \end{tikzcd}
+%     \]
+%     where \(γ_{Ŝ}\) maps each state in \(Ŝ\) to itself.
+% \end{theorem}
+
+% \begin{proof}
+%     We will need to prove three items: first \(γ_{Ŝ}\) is a star system homomorphism, then \(h\) exists and is unique.
+
+%     Notice that every the proof rules of tightening is also a proof rule for star system. 
+%     Therefore, if \(s \transAcc{Ŝ\!↓}{b}\) can be derived using the rules of tightening, the same trasition can also be derived using the proof rules of the star system.
+%     In other word, \(s \transAcc{Ŝ\!↓}{b}\) implies \(s \transAcc{Ŝ^*}{b}\).
+%     Similarly \(s₁ \transvia{b ∣ p}_{Ŝ\!↓} s₂\) implies \(s₁ \transvia{b ∣ p}_{Ŝ^*} s₂\).
+%     Thus, \(γ_{Ŝ}\), which maps each state to itself, is a morphism in the category of star systems.
+
+%     Because \(γ_{Ŝ}\) maps every state to itself, thus \(f\) is the unique function that can make the diagram commute. 
+%     We just need to verify that \(f\) indeed is a star-system homomorphism.
+
+% \end{proof}
+
+% \begin{corollary}
+%     The tightening \((-)\!↓\) is a functor and is right adjoint to the embedding from symbolic GKAT coalgebra to star systems.
+% \end{corollary}
+
+\subsection{Derivative, Thomposon's Construction, and Their Relation}
+
+In this section, we will demonstrate the construction of Derivative and Thompson's construction in a symbolic setting.
+In addition to the construction itself, we will show there is a surjective homomorphism from Thompson's construction to derivative.
+The existence of such surjective homomorphism, especially its lowering by functoriality of lowering, serves several purposes: first, it proves that the semantics of Thompson's construction and derivatives coincides, allowing us to only prove correctness for one of the semantics, instead of both;
+second, it shows that derivative construction produces a smaller automaton than Thompson's construction; finally, it allow us to establish the finiteness of derivatives with the finiteness of Thompson's construction.
+
+However, instead of the usual Derivative and Thomposon's construction, we will construct their weak version, i.e. with unproductive transitions.
+This approach offers a unique advantage, that is the state of the derivatives from an expression \(e\) will all be structurally related to \(e\). 
+Specifically, they will be sub-expressions of \(e\) except the while loop case.
+This structural rigidity allows us to establish a connection between Derivative and Thompson's construction, then propagate this connection down to symbolic GKAT coalgebra and GKAT coalgebra using the functoriality of tightening and lowering.
+
+The coalgebra generated by derivatives is denoted as \(D\), and its states are expressions, with the transitions in \cref{fig:weak-derivarives}.
+\begin{figure*}
+    \begin{mathpar}
+        p \transvia{1 ∣ p}_D 1
+        \and  
+        b \transAcc{D}{b}
+        \and  
+        e +_b f \transvia{b}_D e
+        \and
+        e +_b f \transvia{\overline{b}}_D f
+        \and
+        \inferrule[]
+        {e \transvia{b}_D e'}
+        {e; f \transvia{b}_D e'; f}
+        \and
+        \inferrule[]
+        {e \transvia{b ∣ p}_D e'}
+        {e; f \transvia{b ∣ p}_D e'; f}
+        \and
+        \inferrule[]
+        {e \transAcc{D}{b}}
+        {e; f \transvia{b}_D f}
+        \and
+        e^{(b)} \transvia{b}_D e; e^{(b)} 
+        \and 
+        e^{(b)} \transAcc{D}{\overline{b}}
+    \end{mathpar}
+    \caption{derivative rules}\label{fig:weak-derivarives}
+\end{figure*}
+We call the principle GKAT sub-coalgebra generated by an expression \(e\), the \emph{derivative automaton of \(e\)}, and denote it \(⟨e⟩_D\).
+
+We use \((S_e, ŝ_e)\) to denote the \emph{Thompson's automaton of \(e\)}, which is a weak symbolic GKAT coalgebra \(S_e\) with a \emph{start state} \(ŝ_e ∈ S_e\).
+The Thompson's construction can be inductively defined on expressions as in~\cref{tab:weak-thompson}
+\begin{table}
+    \begin{tabular}{r||c|c|l|l}
+        Expression & State & Start & Accepts & Transitions \\
+        \hline  
+        \(p\) 
+        & \(\{ŝ, s_{acc}\}\) 
+        & \(ŝ\)
+        & \(ϵ(s) ≜ \begin{cases}
+            ∅ & s = ŝ \\  
+            \{1\} & s = s_{acc}
+        \end{cases}\) 
+        & \(δ(s) ≜ \begin{cases}
+            \{(1, s_{acc}, p)\} & s = ŝ \\  
+            ∅ & s = s_{acc}
+        \end{cases}\) \\
+        \(b\)
+        & \(\{ŝ\}\)
+        & \(ŝ\)
+        & \(ϵ(ŝ) ≜ \{b\}\)
+        & \(δ(ŝ) ≜ ∅\) \\
+        \(e +_b f\) 
+        & \(\{ŝ\} + S_e + S_f\) 
+        & \(ŝ\) 
+        & \(ϵ(s) ≜ \begin{cases}
+            ∅ & s = ŝ \\  
+            ϵ_e(s) & s ∈ S_e \\  
+            ϵ_f(s) & s ∈ S_f
+        \end{cases}\)
+        & \(δ(s) ≜ \begin{cases}
+            \{(b, ŝ_e), (\overline{b}, ŝ_f)\} & s = ŝ \\  
+            δ_e(s) & s ∈ S_e \\  
+            δ_f(s) & s ∈ S_f
+        \end{cases}\) \\
+        \(e; f\) 
+        & \(S_e + S_f\) 
+        & \(ŝ_e\)
+        & \(ϵ(s) ≜ \begin{cases}
+            ∅ & s ∈ S_e \\
+            ϵ_f(s) & s ∈ S_f
+        \end{cases}\)
+        & \(δ(s) ≜ \begin{cases}
+            \{(b, ŝ_f) ∣ b ∈ ϵ_e(s)\} ∪
+            δ_e(s) & s ∈ S_e \\
+            δ_f(s) & s ∈ S_f \\  
+        \end{cases}\)\\
+        \(e^{(b)}\) 
+        & \(\{ŝ\} + S_e\) 
+        & \(ŝ\) 
+        & \(ϵ(s) ≜ \begin{cases}
+            \{\overline{b}\} & s = ŝ \\
+            ∅ & \text{otherwise}
+        \end{cases}\)
+        & \(δ(s) ≜ \begin{cases}
+            \{(b, ŝ_e)\} & s = ŝ \\
+            \{(c, ŝ) ∣ c ∈ ϵ_e(s)\} ∪ δ_e(s) & s ≠ ŝ 
+        \end{cases}\)
+    \end{tabular}
+    \caption{rules for thompson's construction}\label{tab:weak-thompson}
+\end{table}
+
+\begin{theorem}\label{thm:hom-thompson-to-deriv}
+    For each expression \(e\), there exists a homomorphism \(h_e: S_e → D\), s.t. \(h_e(ŝ_e) = e\).
+\end{theorem}
+
+\begin{proof}
+    By induction on the structure of \(e\).
+
+    The map \(h_p\) for a primitive action \(p\) can be defined as follows,
+    \begin{mathpar}
+        h_p(ŝ) ≜ p \and h_p(s_{acc}) ≜ 1
+    \end{mathpar}
+    the required conditions, i.e. homomorphism, surjectivity, and \(h_e(ŝ_e) = e\) by unfolding the definition. 
+    Similarly, for a boolean expression \(b\), the map \(h_b\) is defined as \(h_b(ŝ) ≜ b\).
+
+    The map \(h_{e +_b f}\) is defined as follows 
+    \begin{mathpar}
+        h_{e +_b f}(ŝ) ≜ e +_b f \and  
+        h_{e +_b f}(s) ≜ h_e(s) \text{ when } s ∈ S_e \and  
+        h_{e +_b f}(s) ≜ h_f(s) \text{ when } s ∈ S_f
+    \end{mathpar}
+    Then the fact that \(h_{e +_b f}\) is a homomorphism can be verified by case analysis:
+    \begin{align*}
+        \text{when \(s = ŝ\)} 
+        &&& ϵ_{e +_b f}(ŝ) 
+        = ∅ 
+        = ϵ_D(e +_b f) = ϵ_D(h_{e +_b f}(ŝ)) \\
+        &&& δ_D(h_{e +_b f}(ŝ))
+        =  δ_D(e +_b f)
+        = \{(b, e), (\overline{b}, f)\} 
+        = \{(b, h_{e +_b f}(ŝ_e)), (\overline{b}, h_{e +_b f}(ŝ_f))\} \\  
+        \text{when \(s ∈ S_e\)} 
+        &&& ϵ_{e +_b f}(s) 
+        = ϵ_{S_e}(s)
+        = ϵ_D(h_e(s)) = ϵ_D(h_{e +_b f}(s)) \\
+        &&& δ_D(h_{e +_b f}(s))
+        =  δ_D(h_{e}(s))\\
+        &&&
+        = \{(b, h_{e}(s)) ∣ (b, s) ∈ δ_S(s)\}\\ 
+        \text{when \(s ∈ S_f\)} 
+        &&& ϵ_{e +_b f}(s) 
+        = ϵ_{S_f}(s)
+        = ϵ_D(h_f(s)) = ϵ_D(h_{e +_b f}(s)) \\  
+    \end{align*}
+\end{proof}
+
+\begin{corollary}\label{thm:deriv-subcoalgebra-thomp}
+    The coalgebra generated by the derivarives of \(e\), i.e. \(⟨e⟩_D\) is a sub-coalgebra of the image of \(h_e: S_e → D\) in~\cref{thm:hom-thompson-to-deriv}.
+\end{corollary}
+
+\begin{proof}
+    Since \(h_e\) is a coalgebra homomorphism, its imagem \(h_e(S_e)\) is a sub-coalgebra of \(D\). 
+    Because \(⟨e⟩_D\) is the smallest sub-coalgebra that contains \(D\), then \(⟨e⟩_D ⊑ h_e(S_e)\).
+\end{proof}
+
+Intuitively, \Cref{thm:deriv-subcoalgebra-thomp} means generated from derivative \(⟨e⟩_D\) will always have ``less states'' than the coalgebra generated by Thomposon's construction \(S_e\). 
+Specifically, the number of state in \(S_e\) is linear to the size of \(e\), therefore the size of the derivative \(⟨e⟩_D\) is not only finite, but also linear to the size of \(e\).
+
+\begin{theorem}
+    Given any weak symbolic GKAT coalgebra \(S\) and a state \(s ∈ S\), the \(⟨s⟩_{S\!↓}\) is a subalgebra of \((⟨s⟩_{S})\!↓\).
+\end{theorem}
+
+In particular, \(⟨e⟩_{D\!↓}\) is a sub-coalgebra of \((⟨e⟩_{D})\!↓\), which has the same number of states as \(⟨e⟩_{D}\). 
+Recall that the number of state in \(⟨e⟩_{D}\) is at most linear to the size of \(e\). Thus, the automata generated by tightened derivative \(⟨e⟩_{D↓}\) is at most linear to \(e\).
+This gives us a much better complexity bound for derivative-based decision procedure than the traditional closure techniques.
+
+\begin{figure*}
+    \begin{mathpar}
+        \inferrule[]{\\}
+        {p \transvia{1 ∣ p}_{D\!↓} 1} \and  
+        \inferrule[]{\\}
+        {b \transAcc{D\!↓}{b}} \and  
+        \inferrule[]
+        {e \transvia{c ∣ p}_{D\!↓} e'}
+        {e +_b f \transvia{b ∧ c ∣ p}_{D\!↓} e'} 
+        \and
+        \inferrule[]
+        {e \transAcc{D\!↓}{c}}
+        {e +_b f \transAcc{D\!↓}{b ∧ c}}
+        \and
+        \inferrule[]
+        {f \transvia{c ∣ p}_{D\!↓} f'}
+        {e +_b f \transvia{\overline{b} ∧ c ∣ p}_{D\!↓} f'}
+        \and
+        \inferrule[]
+        {f \transAcc{D\!↓}{c}}
+        {e +_b f \transAcc{D\!↓}{\overline{b} ∧ c}}
+        \and
+        \inferrule[]
+        {e \transAcc{D\!↓}{b} \\ f \transAcc{D\!↓}{c}}
+        {e; f \transAcc{D\!↓}{b ∧ c}}
+        \and 
+        \inferrule[]
+        {e \transvia{b ∣ p} e'}
+        {e; f \transvia{b ∣ p} e'; f}
+        \and
+        \inferrule[]
+        {e \transAcc{D\!↓}{b} \\ f \transvia{d ∣ p}_{D\!↓} f'}
+        {e; f \transvia{b ∧ d ∣ p}_{D\!↓} f'}
+        \and  
+        \inferrule[]
+        {\\}
+        {e^{(b)} \transAcc{D\!↓}{\overline{b}}}  
+        \and  
+        \inferrule[]
+        {e \transvia{b ∣ p} e'}
+        {e^{(b)} \transvia{b ∣ p}_{D\!↓} e'; e^{(b)}}
+    \end{mathpar}
+    \caption{Tightened Derivarive rules}\label{fig:tight-derivatives}
+\end{figure*}
+
+\begin{theorem}[Tightened Derivative]
+    Every transition in the tightened derivatives coalgebra \(D\!↓\) can be derived from rules in~\Cref{fig:tight-derivatives}.
+\end{theorem}
+
+\begin{proof}
+    We need to show any transition can be derivable from tightened derivative rules (\Cref{fig:tight-derivatives}) if and only if the transition is derivable from the weak derivative rules (\Cref{fig:weak-derivarives}) and tightening rule (\Cref{fig:tightening-rules}).
+    We will show this by case analysis on the type of the transtion i.e.
+whether \(e \transvia{b ∣ p}_{D\!↓} e'\) or \(e \transAcc{D\!↓}{b}\), and the start of the transition, which is denoted \(e\) in previous example.
+
+    Taken as an example, if the transition is an accepting transition starting with a sequencing statement, i.e. \(e; f \transAcc{D\!↓}{b}\) for some boolean expression \(e\).
+    By observing the weak derivatives rules and the tightening rules, the only rules that can reach \(e; f \transAcc{D\!↓}{b}\) is one of the tightening rules for some expression \(g\):
+    \begin{mathpar}
+        \inferrule[]
+        {e; f \transvia{b}_{D} g \\ g \transAcc{D\!↓}{c}}
+        {e; f \transAcc{D\!↓}{b ∧ c}}
+    \end{mathpar}
+    Then observe the only proof rules that can lead to condition \(e; f \transvia{b}_{D} g\) for some \(g\) are the following derivatives rules
+    \begin{mathpar}
+        \inferrule[]
+        {e \transvia{b}_D e'}
+        {e; f \transvia{b}_D e'; f}
+        \and  
+        \inferrule[]
+        {e \transAcc{D}{b}}
+        {e; f \transvia{b} f}
+    \end{mathpar}
+\end{proof}
+
+\begin{corollary}[Correctness]
+
+\end{corollary}
+
+\begin{theorem}[Tightened Thomposon's Construction]
+
+\end{theorem}
+
+\section{Future Work}
+
+Can weak symbolic coalgebra leads to a simpler completeness proof.
 
@@
 
-\section{diagrammatical characterization of normalized homomorphism}\label{sec:diagram-of-norm-sim}
-
 \end{document}

### 2024-12-29 main.tex
big proof done.

--- 2024-12-02 main.tex
@@
 With the above intuition in mind, a symbolic GKAT coalgebra \(Ŝ ≜ ⟨S, ϵ̂, δ̂⟩\) can be lowered into a GKAT coalgebra \(⟨S, δ⟩\) in the following manner:
-\[
+\begin{align}\label{cons:lowering}
 δ(s, α) ≜ \begin{cases}
-    \accept & α ≤ ϵ̂(s) \\  
+    \accept & ∃ b ∈ ϵ̂(s), α ≤ b \\  
     (s', p) & ∃ b ∈ \Bool_B, α ≤ b \text{ and } δ(s, b) = (s', p) \\  
@@
 \end{cases}
-\]
+\end{align}
 This is well-defined, i.e. no more than one clause can be satisfied precisely because the boolean expressions appear in \(ϵ̂\) and \(δ̂\) are disjoint.
@@
 
+\begin{theorem}[Functoriality]
+    The lowering operation is a functor, given a symbolic GKAT coalgebra homomorphism \(h: Ŝ → Û\), then \(h\) is also a homomorphism \(h: S → U\).
+\end{theorem}
+
+\begin{proof}
+
+\end{proof}
+
 We can then migrate the normalized bisimulation algorithm to the symbolic setting, we will first prove an inductive construction theorem like~\cref{thm:inductive-construction}.
@@
 
-We will go through two ways of constructing symbolic GKAT automata, one is Thompson's construction, the other is derivative.
-However, in order to exhibit the relation between these two construction, specifically, the derivative construction always produces a smaller automaton than the Thompson's construction, we opt to use to produce "weak" symbolic GKAT automata: symbolic GKAT automata where the transition no longer need to perform an action.
-
-\subsection{Weak Symbolic GKAT Automata And Tightening}
-
-Weak Symbolic GKAT automata is a symbolic GKAT automata that allow internal transitions via a boolean expression:
-\[⟨S, 
-ϵ̂: S → \Bool, 
-δ̂: S → (\Bool × S) + (\Bool × S × K)⟩.\]
-We also require the deterministic criteria, all the boolean transitions of \(s\), i.e. all the boolean expressions in the following set:
-\[\{ϵ̂(s)\} ∪ 
-\{b ∣ ∃ s' ∈ S, (b, s') ∈ δ̂(s)\} ∪ 
-\{b ∣ ∃ s' ∈ S, p ∈ K, (b, s', p) ∈ δ̂(s)\},\]
-are disjoint.
-Notice that our notion of
-
-% \section{Symbolic algorithm BAK}
-
-% \subsection{Symbolic GKAT Coalgebra}
-
-% Given an alphabet \(K, B\), a \emph{symbolic GKAT coalgebra} \(Ŝ ≜ ⟨S, ϵ̂, δ̂⟩\) consists of a state set \(S\) consists of a accepting boolean \(ϵ̂\) and a transition function \(δ̂\):
-% \begin{mathpar}
-%     ϵ̂: S → \Bool_B, \and
-%     δ̂: S → S → K → \Bool_B,
-% \end{mathpar}
-% where \(\Bool_B\) is the free boolean algebra over \(B\) (boolean expressions modulo boolean algebra axioms), and for all states \(s ∈ S\), all the booleans results are ``disjoint''; namely the conjunction of any two expression from the set \(ϵ̂(s) + \{δ̂(s, s', p) ∣ s' ∈ S, p ∈ K\}\) are false in \(\Bool_B\).
-% The ordering on two symbolic coalgebra on the same carrier \(S\) is defined pointwise, namely 
-% \[⟨ϵ̂₁, δ̂₁⟩ ≤ ⟨ϵ̂₂, δ₂⟩ \text{ if and only if } ∀ s, s' ∈ S, ∀ p ∈ K, \hat{ϵ₁}(s) ≤ \hat{ϵ₂}(s) \text{ and } \hat{δ₁}(s, s', p) ≤ \hat{δ₂}(s, s', p).\]
-
-% Intuitively, the expressions in \(\Bool_B\) is treated as the ``symbolic'' transitions, which can be think of as a set of atoms by the isomorphism \(\Bool_B ≅ 2^{\At_B}\). 
-% Then we can explain the intuition for the definition of symbolic GKAT coalgebra:
-% \begin{itemize}
-%     \item \(ϵ̂(s)\) can be thought of as all the atoms that is accepted by the state \(s\).
-%     \item \(δ̂(s, s', p)\) denotes all the atoms that can transition from \(s\) to \(s'\) while executing \(p\).
-% \end{itemize}
-% Then the atoms excluded by both operations are implicitly rejecting. Formally, we can define a boolean expression that contains all the atoms that will be rejected 
-% \[ρ̂(s) = ¬ (ϵ̂(s) ∨ ⋁ \{b ∣ ∃ s' ∈ S, p ∈ K, δ̂(s, s', p) = b\}).\]
-
-% This intuition gives us a isomorphism between between GKAT coalgebra and symbolic GKAT coalgebra.
-Formally, given a GKAT coalgebra \(⟨S, δ⟩\), then its corresponding symbolic GKAT coalgebra will share the same states \(S\), with the transition defined as follows:
-% \begin{align*}
-%     ϵ̂(s) & ≜ ⋁ \{α ∈ \At ∣ δ(s, α) = \accept\} & 
-%     δ̂(s, s', p) & ⋁ \{α ∈ \At ∣ δ(s, α) = (s', p)\}.
-% \end{align*}
-% We will denote the above map as \(\symb\), i.e.
-\(\symb(⟨S, δ⟩) ≜ ⟨S, ϵ̂, δ̂⟩\).
-Notice that we are implicitly using atoms as an element in the boolean algebra, indeed an atom can be thought about as a conjunction.
-
-% The inverse of \(\symb\) is denoted as \(\unsymb\) where given a symbolic GKAT coalgebra \(⟨S, ϵ̂, δ̂⟩\), we construct a GKAT coalgebra over states \(S\) with the following transition:
-% \[δ(s, α) ≜ \begin{cases}
-%     \accept & α ∈ ϵ̂(s) \\  
-%     (s', p) & α ∈ δ̂(s, s', p) \\  
-%     \reject & \text{otherwise}
-% \end{cases}\]
-% Notice the \(δ\) is well-defined, that is given any \(s ∈ S\) and \(α ∈ \At\), \(δ(s,α)\) can only take on one value; because symbolic GKAT coalgebra is deterministic.
-
-% \subsection{Decision Procedure Of Symbolic GKAT coalgebra}
-
-% The formulation of symbolic GKAT coalgebra is mathematically nice, yet inefficient.
-Between most states \(s, s' ∈ S\) and action \(p ∈ K\), \(δ(s, s', p)\) will always be false.
-Therefore, we can developed a set representation of the derivative, where we will typically denote using \(Δ̂\):
-% \[Δ̂: S → 𝒫(\Bool_B × S × K).\]
-% we always assume transition functions denoted this way are \emph{deterministic}; specifically, all the boolean expression in the following set are disjoint: \[\{b ∣ ∃ s' ∈ S, p ∈ K, (b, s', p) ∈ Δ̂(s)\}.\]
-
-% Intuitively, if \((b, s', p) ∈ Δ̂(s)\), then \(s\) will transition to \(s'\) for all the atoms \(α ≤ b\) while executing \(p\).
-Notice that the above statement does not imply that \emph{all} \(α\) that makes \(s\) transition to \(s'\) while executing \(p\) will be covered by \(b\); that is, there can exists a \((b', s', p) ∈ Δ̂(s)\), where \(b ≠ b'\).
-
-% This means that the set representation of a symbolic GKAT coalgebra no on longer uniquely correspond to a GKAT coalgebra.
-% \begin{example}
-
-% \end{example}
-
-% However, every set representation of a symbolic GKAT coalgebra, indeed correspond to a GKAT coalgebra, specifically given a set representation \(⟨S, ϵ̂, Δ̂⟩\), we can construct a symbolic GKAT \(⟨S, ϵ̂, δ̂⟩\), by defining the following \(δ̂\):
-% \[δ̂(s, s', p) = ⋁ \{b ∣ (b, s', p) ∈ Δ̂(s)\}.\]
-% More importantly, the map described above is surjective. This means that all the GKAT coalgebra can indeed be represented as at least one set-representation.
-% Similarly, set representation can also be converted to GKAT coalgebra: given a set representation \(⟨S, ϵ̂, Δ̂⟩\), we can lower it into a GKAT coalgebra \(⟨S, δ⟩\), by constructing the following \(δ\):
-% \[δ(s, α) ≜ \begin{cases}
-%     \accept & α ∈ ϵ̂(s) \\  
-%     (s', p) & ∃ b ≥ α, (b, s', p) ∈ Δ̂(s) \\  
-%     \reject & \text{otherwise}
-% \end{cases}\]
-% Indeed, lowering the set representation directly to GKAT coalgebra give the same result as converting to symbolic GKAT coalgebra and then lowering.
-This fact is essential for us to relate these semantics.
-
-% Finally, we can lift the inductive construction theorem (\cref{thm:inductive-construction}) to set representation of GKAT coalgebras, giving a sound and complete symbolic decision procedure. 
-% In the following lemmas, we will use \(S, T\) to denote some GKAT coalgebra, and \(Ŝ, T̂\) will be as some set-represented symbolic GKAT coalgebra that corresponds to \(S\). 
-% Formally, if we lower \(Ŝ\) to GKAT coalgebra, we will obtain exactly \(S\); similarly for \(T̂\).
-
-% \begin{lemma}
-%     We let the \(s ∈ S\) and \(t ∈ T\) be two states, then the following equivalence between conditions in \cref{thm:inductive-construction} hold: 
-%     \begin{itemize}
-%         \item condition~\labelcref{itm:acc-condition} in inductive construction holds if and only if \(ϵ̂_{Ŝ}(s) = ϵ̂_{T̂}(t)\).
-%         \item condition~\labelcref{itm:transition-bisim} holds if and only if for all \((a, s', p) ∈ δ̂_{Ŝ}(s)\) and \((b, t', q) ∈ δ̂_{T̂}(t)\):
-%         \[a ∧ b ≠ 0 \text{ and } p = q ⟹ \text{exists a bisimulation \(∼\) between \(S\) and \(T\), } s.t. s' ∼ t'.\]
-%         \item condition~\labelcref{itm:transition-dead}, holds if and only if for all \((a, s', p) ∈ δ̂_{Ŝ}(s)\) and \((b, t', q) ∈ δ̂_{T̂}(t)\):
-%         \[a ∧ b ≠ 0 \text{ and } p ≠ q ⟹ \text{both \(s\) and \(t\) are dead}.\]
-%         \item condition~\labelcref{itm:rej-or-dead} holds if and only if the following holds:
-%     \end{itemize}
-% \end{lemma}
-
-% \begin{lemma}
-%     Given two states \(\)
-% \end{lemma}
-
-% \subsection{Symbolic Derivatives and Correctness}
-
-% Given a symbolic GKAT coalgebra \(⟨S, δ̂, ϵ̂⟩\) and a state \(s ∈ S\), we use the notation \(s ⇒ b\) to denote that \(ϵ̂(s) ≥ b\), and we use the notation \(s \transvia{b ∣ p} s'\) to denote that \(δ̂(s, s', p) ≥ b\), where the ordering is the typical ordering in boolean algebra.
-
-% The symbolic derivatives for GKAT forms a DKCT where the states are represented by GKAT expression \(\GKAT_{K, B}\), and the boolean algebra is the free boolean algebra \(\Bool_B\) over the test alphabet \(B\):
-% \begin{mathpar}
-%     ϵ: \GKAT_{K, B} → \Bool_B; \and  
-%     δ: \GKAT_{K, B} → \GKAT_{K, B} → K → \Bool_B.
-% \end{mathpar} 
-% the accepting map \(ϵ\) and transition map \(δ\) can be defined by the smallest maps that satisfy the following rules:
-% \begin{mathpar}
-%     \inferrule{\\}{b ⇒ b} \and 
-%     \inferrule{\\}{p \transvia{1 ∣ p} 1} \and 
-%     \inferrule{e ⇒ a}{e +_b f ⇒ b a} \and 
-%     \inferrule{f ⇒ a}{e +_b f ⇒ b̄ a} \and 
-%     \inferrule{e \transvia{a ∣ p} e'}{e +_b f \transvia{b ∧ a ∣ p} e'} \and 
-%     \inferrule{f \transvia{a ∣ p} f'}{e +_b f \transvia{b̄ ∧ a ∣ p} f'} \and 
-%     \inferrule{e ⇒ a \\ f ⇒ b}{e ⋅ f ⇒ a ∧ b} \and 
-%     \inferrule{e ⇒ a \\ f \transvia{b ∣ p} f'}{e ⋅ f \transvia{a ∧ b ∣ p} f'} \and 
-%     \inferrule{e \transvia{b ∣ p} e'}{e ⋅ f \transvia{b ∣ p} e' ⋅ f} \and 
-%     \inferrule{\\}{e^{(b)} ⇒ b̄} \and 
-%     \inferrule{e \transvia{a ∣ p} e'}{e^{(b)} \transvia{b ∧ a ∣ p} e' ⋅ {e}^{(b)}}
-% \end{mathpar}
-% Notice the rules above are very similar to the derivative rules for GKAT~\cite{schmid_GuardedKleeneAlgebra_2021}, this is no coincidence, as we can establish a correspondence of these two to prove correctness of our symbolic algorithm.
-
-% Since the rules listed above is monotonic, that is if we enlarge the boolean in the premise, we also enlarge the boolean in the result. 
-% So we can simply pick the largest boolean on the premise, which is either \(ϵ(e)\) or \(δ(e, e', p)\).
-% Since the disjunction of two boolean \(a ∨ b\) is the smallest boolean that is greater than both \(a\) and \(b\), thus we can implement \(ϵ\) and \(δ\) by iteration through all the rules, and take the disjunction of all the possible booleans.
-
-% \begin{align*}
-%     ϵ(b) & ≜ b 
-%         &  ϵ(e +_b f) & ≜ (b ∧ ϵ(e)) ∨ (b̄ ∧ ϵ(f))\\
-%     ϵ(q) & ≜ 0
-%         & ϵ(e ⋅ f) & ≜ ϵ(e) ∧ ϵ(f)\\  
-%     ϵ(e^{(b)}) & ≜ b̄ 
-% \end{align*}
-
-% However, the derivative function \[δ: \GKAT_{K, B} → \GKAT_{K, B} → K → \Bool_B,\] poses challenges in the implementation, since \(\GKAT_{K, B}\) is infinite, so it will be impractical to search through all the expressions.
-% However, we can treat the function \(δ\) extensionally: \[\GKAT_{K, B} → K → \Bool_B ⊆ 2^{\GKAT_{K, B} × K × \Bool_B},\]
-% and then because the boolean expression don't overlap for each input, hence the boolean expression are necessarily unequal; 
-% we can model the set by a partial map, thus all the \(δ\) can be represented as a function to the following partial maps.
-% \begin{align*}
-%     δ & : \GKAT_{K, B} → (\Bool_B ↛ \GKAT_{K, B} × K) \\
-%     δ(b) & ≜ \{\} \\
-%     δ(q) & ≜ \{1 ↦ (1, p)\}\\  
-%     δ(e +_b f) & ≜ \{b ∧ a ↦ (e', p) ∣ a ↦ (e', p) ∈ δ(e)\} 
-%         ∪ \{b̄ ∧ a ↦ (f', p) ∣ a ↦ (f', p) ∈ δ(f)\}\\
-%     δ(e ⋅ f) & ≜ \{b ↦ (e' ⋅ f, p) ∣ b ↦ (e, p) ∈ δ(e)\}
-%         ∪ \{ϵ(e) ∧ b ↦ (f', p) ∣ b ↦ (f', p) ∈ δ(f)\}\\
-%     δ(e^{(b)}) & ≜ \{b ∧ a ↦ (e' ⋅ e^{(b)}, p) ∣ a ↦ (e', p) ∈ δ(e)\}.
-% \end{align*}
-% FIXME: we still need to think about what representation do we want to go with, with this representation, you can also have duplicated K.
-We need a good example here of why don't we go with map.
-One good argument is that this representation is complex.
-
-% \subsection{Algorithm and Complexity}
-
-% \section{Implementation and Experiments}
-
-% \subsection{the new norm}
-
-% The theoretical foundation of this algorithm is surprisingly simple, where we only need one minor tweak to the definition of normalization.
-% Instead of normalizing both the states \(S\) and the transition function \(δ_S\), we will only normalize the function \(δ_S\) to obtain \(δ_S\).
-
-% Intuitively, this only rejects transitions to dead states, but doesn't remove dead states completely.
-As we will see later, the bisimulation generated by this notion will be able to contain rejection state, this allows us to only check for liveness after ``discrepancy'' as described in the intro.
-
-% \begin{definition}[normalized homomorphism]
-%     A normalized homomorphism \(h: S → T\) is a homomorphism with normalized transition functions:
-%     \[
-%         \begin{tikzcd}
-%             S \ar{r}{h} \ar[swap]{d}{\norm(δ_S)} & T \ar{d}{\norm(δ_T)}\\  
-%             G(S) \ar[swap]{r}{G(h)} & G(T)
-%         \end{tikzcd}
-%     \]
-% \end{definition}
-
-% This normalized homomorphism is very close to the definition of homomorphism in normalized GKAT coalgebra, hence inherits many of the properties. 
-% Specifically, the language model \(𝒢\) is the final coalgebra in here.
-
-% \subsection{soundness and completeness}
-
-% The definition of normalized bisimulation naturally arise from a weakened form of homomorphism.
-
-% \begin{definition}[Normalizd homomorphism]
-%     A \emph{normalized homomorphism} \(h: S → T\) is coproduct (in set) of a \emph{homomorphism} between the normalized coalgebra \(h_{\norm}: \norm(S) → \norm(T)\) and a \emph{function} \(h_{\dead}: \dead{S} → \dead(T)\).
-Concretely,
-%     \[h(s) ≜ \begin{cases}
-%         h_{\norm}(s) & \text{if } s ∈ \norm(S) \\  
-%         h_{\dead}(s) & \text{if } s ∈ \dead(S)
-%     \end{cases}.\]
-%     We will sometimes use the coproduct notation \(h = h_{\norm} + h_{\dead}\).
-% \end{definition}
-
-% \begin{theorem}
-%     GKAT coalgebras under normalized homomorphisms form a category.
+The final piece of the puzzle is to convert any given expression into an ``equivalent'' Symbolic GKAT Automata.
+This goal can be achieved by lifting existent constructions like derivatives and Thompson's construction~\cite{schmid_GuardedKleeneAlgebra_2021,smolka_GuardedKleeneAlgebra_2020}.
+The correctness of these conversions is a consequence of  correctness of their non-symbolic counter-part, i.e.
+we will prove that the lowering as shown in~\labelcref{cons:lowering} of these constructions will yield the conventional derivative and Thompson's construction.
+
+\begin{figure*}
+    \begin{mathpar}
+        \inferrule[]{\\}
+        {p \transvia{1 ∣ p}_{D̂} 1} \and  
+        \inferrule[]{\\}
+        {b \transAcc{D̂}{b}} \and  
+        \inferrule[]
+        {e \transvia{c ∣ p}_{D̂} e'}
+        {e +_b f \transvia{b ∧ c ∣ p}_{D̂} e'} 
+        \and
+        \inferrule[]
+        {e \transAcc{D̂}{c}}
+        {e +_b f \transAcc{D̂}{b ∧ c}}
+        \and
+        \inferrule[]
+        {f \transvia{c ∣ p}_{D̂} f'}
+        {e +_b f \transvia{\overline{b} ∧ c ∣ p}_{D̂} f'}
+        \and
+        \inferrule[]
+        {f \transAcc{D̂}{c}}
+        {e +_b f \transAcc{D̂}{\overline{b} ∧ c}}
+        \and
+        \inferrule[]
+        {e \transAcc{D̂}{b} \\ f \transAcc{D̂}{c}}
+        {e; f \transAcc{D̂}{b ∧ c}}
+        \and 
+        \inferrule[]
+        {e \transvia{b ∣ p}_{D̂} e'}
+        {e; f \transvia{b ∣ p}_{D̂} e'; f}
+        \and
+        \inferrule[]
+        {e \transAcc{D̂}{b} \\ f \transvia{c ∣ p}_{D̂} f'}
+        {e; f \transvia{b ∧ c ∣ p}_{D̂} f'}
+        \and  
+        \inferrule[]
+        {\\}
+        {e^{(b)} \transAcc{D̂}{\overline{b}}}  
+        \and  
+        \inferrule[]
+        {e \transvia{c ∣ p} e'}
+        {e^{(b)} \transvia{b ∧ c ∣ p}_{D̂} e'; e^{(b)}}
+    \end{mathpar}
+    \caption{Symbolic Derivative of GKAT Automata.}\label{fig:derivatives-rules}
+\end{figure*}
+
+The symbolic derivative coalgebra \(D̂\), with expressions as states, is the least symbolic GKAT coalgebra (ordered by point-wise subset ordering on \(ϵ̂\) and \(δ̂\)) that satisfy the rules in~\Cref{fig:derivatives-rules}.
+Notice that the rules listed on~\Cref{fig:derivatives-rules} is very close to that of Schmid et al.~\cite{schmid_GuardedKleeneAlgebra_2021}.
+This is no coincidence, as our definition exactly lowers to the definition of theirs.
+This fact can be proven by case analysis on the shape of the source expression, and forms a basis on our correctness argument.
+
+\begin{theorem}[Correctness]
+    The lowering of \(D̂\) is exactly the derivative defined by Schmid et. al.~\cite{schmid_GuardedKleeneAlgebra_2021}.
+    TODO: unfold the statement.
+\end{theorem}
+
+% \begin{table*}
+%     \centering
+%     \begin{tabular}{c||c|c|c|c|c|c}
+%         Exp & \(S\) 
+%         & \(ϵ^*\) & \(δ^*\) 
+%         & \(ϵ̂(s)\) & \(δ̂(s)\) \\\hline
+%         \(b\) & \(∅\) 
+%         & \{b\} & \(∅\) 
+%         & N/A & N/A \\  
+%         \(p\) & \(\{*\}\) 
+%         & \(∅\) & \(\{(1, *, p)\}\) 
+%         & \(\{1\}\) & \(∅\)\\  
+%         \(e₁ +_b e₂\) & \(S₁ + S₂\) &
+%         \(⟨\{b\}|~ϵ₁^* ∪ ⟨\{\overline{b}\}|~ϵ₂^*\) &
+%         \(⟨\{b\}|~δ₁^* + ⟨\{\overline{b}\}|~δ₂^*\) & 
+%         \(\begin{cases}
+%             ϵ̂₁(s) & s ∈ S₁\\
+%             ϵ̂₂(s) & s ∈ S₂\\
+%         \end{cases}\) & 
+%         \(\begin{cases}
+%             δ̂₁(s) & s ∈ S₁\\
+%             δ̂₂(s) & s ∈ S₂\\
+%         \end{cases}\) \\  
+%         \(e₁ ; e₂\) & \(S₁ + S₂\) & 
+%         \(⟨ϵ₁^*|~ϵ₂^*\) & \(δ₁^* + ⟨ϵ₁^*|~δ₂^*\) & 
+%         \(\begin{cases}
+%             ⟨ϵ̂₁(s)| ~ ϵ₂^*& s ∈ S₁ \\  
+%             ϵ̂₂(s) & s ∈ S₂
+%         \end{cases}\)& 
+%         \(\begin{cases}
+%             δ̂₁(s) ∪ ⟨ϵ̂(s)|~δ₂^* & s ∈ S₁ \\  
+%             δ̂2(s) & s ∈ S₂
+%         \end{cases}\) \\  
+%         \(e₁^{(b)}\) & \(S₁\) & 
+%         \(\{\overline{b}\}\) & \(⟨\{b\}|~δ₁^*\) & 
+%         \(⟨\{\overline{b}\}|~ϵ̂1(s)\) & 
+%         \( ⟨\{b\}|⟨ϵ̂1(s)|~δ₁^* ∪ δ₁(s)\)
+%     \end{tabular}
+%     \caption{Symbolic Thompson's Construction}\label{tab:symb-thomposon-construction}
+% \end{table*}
+
+\begin{table*}
+    \centering
+    \begin{tabular}{c||c|c|l|l}
+        Exp & \(S\) & \(s^*\)  
+        & \(ϵ̂(s)\) & \(δ̂(s)\) \\\hline
+        \(b\) & \(\{s^*\}\) 
+        & \(s^*\)
+        & \(\{b\}\) & \(∅\) \\  
+        \(p\) & \(\{s^*, s₁\}\) 
+        & \(s^*\)
+        & \(\begin{cases}
+           ∅ & s = s^* \\  
+           \{1\} & s = s₁ 
+        \end{cases}\) 
+        & \(\begin{cases}
+            \{(1, s₁, 0)\} & s = s^* \\  
+            ∅ & s = s₁
+        \end{cases}\)\\  
+        \(e₁ +_b e₂\) & \(\{s^*\} + S₁ + S₂\) &
+        \(s^*\) &
+        \(\begin{cases}
+            ⟨\{b\}| ϵ̂₁(s₁^*) ∪ ⟨\{b\}| ϵ̂₂(s₂^*) & s = s^* \\
+            ϵ̂₁(s) & s ∈ S₁\\
+            ϵ̂₂(s) & s ∈ S₂\\
+        \end{cases}\) & 
+        \(\begin{cases}
+            ⟨\{b\}| δ̂₁(s₁^*) + ⟨\{b\}| δ̂₂(s₂^*) & s = s^* \\
+            δ̂₁(s) & s ∈ S₁\\
+            δ̂₂(s) & s ∈ S₂\\
+        \end{cases}\) \\  
+        \(e₁ ; e₂\) & \(S₁ + S₂\) & 
+        \(s₁^*\) & 
+        \(\begin{cases}
+            ⟨ϵ̂₁(s)| ϵ̂₂(s₂^*)& s ∈ S₁ \\  
+            ϵ̂₂(s) & s ∈ S₂
+        \end{cases}\)& 
+        \(\begin{cases}
+            δ̂₁(s) + ⟨ϵ̂(s)| δ̂₂(s₂^*) & s ∈ S₁ \\  
+            δ̂2(s) & s ∈ S₂
+        \end{cases}\) \\  
+        \(e₁^{(b)}\) & \(\{s^*\} + S₁\) & 
+        \(s^*\) &
+        \(\begin{cases}
+            \{\overline{b}\} & s = s^*\\
+            ⟨\{\overline{b}\}| ϵ̂1(s) & s ∈ S₁
+        \end{cases}\) & 
+        \(\begin{cases}
+            ⟨\{b\}| δ̂₁(s₁^*) & s = s^* \\
+            δ₁(s) ∪ ⟨\{b\}|⟨ϵ̂1(s)| δ̂₁(s₁^*) & s ∈ S₁
+        \end{cases}\)
+    \end{tabular}
+    \caption{Symbolic Thompson's Construction}\label{tab:symb-thomposon-construction}
+\end{table*}
+
+Another way to construct an automaton is via Thompson's construction, we lift the original construction to the symbolic setting.
+A common expression to construct is a guard operation, denoted by \(⟨B|\), where \(B\) is a set of boolean expressions.
+TODO: define transition dynamics and accepting dynamics earlier.
+Concretely, this guard can be defined on both accepting dynamics and transition dynamics:
+\begin{align*}
+    ⟨B|ϵ̂(s) & ≜ \{b ∧ c ∣ b ∈ B, c ∈ ϵ(s)\}; \\
+    ⟨B|δ̂(s) & ≜ \{(b ∧ c, s', p) ∣ b ∈ B, (c, s', p) ∈ δ(s)\}.
+\end{align*}
+Notably, besides guarding transition and acceptance with different conditions, like in if statements, the guard expression can be used to simulate uniform continuation.
+Specifically, we can use \(⟨ϵ̂(s)|~δ(s)\) to connecting all the accepting state of \(s\) to the dynamic \(δ(s)\).
+
+With these definitions in mind, we can define symbolic Thompson's construction inductively as in~\Cref{tab:symb-thomposon-construction}, where we let \((S₁, ϵ̂₁, δ̂₁)\) and \((S₂, ϵ̂₂, δ̂₂)\) to be result of Thompson's construction for \(e₁\) and \(e₂\) respectively.
+The \(S₁ + S₂\) is the disjoint union of \(S₁\) and \(S₂\), and for any two transition dynamics \(δ₁(s₁): 𝒫(\Bool × S₁ × K)\) and \(δ₂(s₂): 𝒫(\Bool × S₂ × K)\), then \(δ₁^*(s₁) + δ₂^*(s₂): 𝒫(\Bool × (S₁ + S₂) × K)\) is the bifunctorial lift via \(+\).
+
+One notable difference between the original construction~\cite{smolka_GuardedKleeneAlgebra_2020} and our construction is that we use a start state \(s^* ∈ S\), instead of a start dynamics (or pseudo-state).
+This choice will make the proof slightly easier. 
+However, in~\Cref{sec:optimization-implementation}, we will explain that our implementation uses start dynamics instead of start state, to avoid unnecessary lookups and unreachable states.
+
+We would like to explore several desirable theoretical properties of both derivatives and Thompson's construction.
+Specifically, the correctness, i.e.
+the semantics of the ``start state'' the both construction have the same preserves the trace semantics of the expression; finiteness, i.e.
+the coalgebra generated is always finite, which means that our equivalence algorithm will eventually terminate; and finally, how does the number of reachable state relate to the size of the expression, so that we can estimate the complexity of the equivalence checking algorithm.
+Turns out all of these questions can be answered by a connection by a homomorphism from symbolic Thompson's construction to the symbolic derivatives.
+
+\begin{theorem}\label{thm:hom-thompson-derivative}
+    Given any GKAT expression \(e\), the resulting symbolic GKAT coalgebra from Thompson's construction \(Ŝ_e\) have a homomorphism to derivatives \(h: Ŝ_e → D̂\), s.t.
+for the start state \(s^* ∈ S, h(s^*) = e\).
+\end{theorem}
+
+\begin{proof}
+    By induction on the structure of \(e\).
+We will recall that \(h: Ŝ_e → ⟨e⟩_D\) is a symbolic GKAT coalgebra homomorphism when the following two conditions are true: \(s \transAcc{Sₑ}{b}\) if and only if \(h(s) \transAcc{D̂}{b}\); and \(s \transvia{b ∣ p}_{Sₑ} s'\) if and only if \(h(s) \transvia{b ∣ p}_{D̂} h(s')\).
+
+    When \(e ≜ b\) for some tests \(b\), then the function \(h\) is defined as \(\{s^* ↦ b\}\).
+    When \(e ≜ p\) for some primitive action \(p\), then the function \(h\) is defined as \(\{s^* ↦ p, * ↦ 1\}\).
+    The homomorphism condition can then be verified by unfolding the definition.
+
+    When \(e ≜ e₁ +_b e₂\), by induction hypothesis, we have homomorphisms \(h₁: Ŝ_{e₁} → ⟨e₁⟩_D\) and \(h₂: Ŝ_{e₂} → ⟨e₂⟩_D\).
+    Then we define the homomorphism 
+    \[h(s) ≜ \begin{cases}
+        e₁ +_b e₂ & s = s^* \\  
+        h₁(s) & s ∈ Ŝ_{e₁} \\  
+        h₂(s) & s ∈ Ŝ_{e₂}
+    \end{cases}\]
+    We show that \(h\) is a homomorphism. Because \(Ŝₑ\) preserves the transition and acceptance of \(Ŝ_{e₁}\) and \(Ŝ_{e₂}\), then for all \(s ∈ Ŝ_{e₁} ∩ Ŝ_{e}\), we have
+    \begin{align*}
+        & s \transAcc{Ŝ_{e}}{c} \text{ iff }
+        s \transAcc{Ŝ_{e₁}}{c} \text{ iff }
+        h₁(s) \transAcc{D̂}{c} \text{ iff }
+        h(s) \transAcc{D̂}{c} \\ 
+        & s \transvia{c ∣ p}_{Ŝ_{e}} s' \text{ iff }
+        s \transvia{c ∣ p}_{Ŝ_{e₁}} s' \text{ iff }
+        h₁(s) \transvia{c ∣ p}_{D̂} h₁(s') \text{ iff }
+        h(s) \transvia{c ∣ p}_{D̂} h(s')  
+    \end{align*}
+    And similarly for \(s ∈ Ŝ_{e₂} ∩ Ŝ_{e}\).
+    So we only need to show the homomorphic condition for the start state \(s^*\):
+    \begin{align*}
+        & s^* \transAcc{Ŝ_{e}}{c} \\
+        \text{iff } &
+        (∃ a, b ∧ a = c \text{ and } 
+        s₁^* \transAcc{Ŝ_{e₁}}{a})
+        \text{ or } 
+        (∃ a, \overline{b} ∧ a = c \text{ and } 
+        s₂^* \transAcc{Ŝ_{e₂}}{a}) \\
+        \text{iff } & 
+        (∃ a, b ∧ a = c \text{ and } 
+        h₁(s₁^*) \transAcc{D̂}{a})
+        \text{ or } 
+        (∃ a, \overline{b} ∧ a = c \text{ and } 
+        h₂(s₂^*) \transAcc{D̂}{a}) \\  
+        \text{iff } & 
+        (∃ a, b ∧ a = c \text{ and } 
+        e₁ \transAcc{D̂}{a})
+        \text{ or } 
+        (∃ a, \overline{b} ∧ a = c \text{ and } 
+        e₂ \transAcc{D̂}{a}) \\  
+        \text{iff }& e₁ +_b e₂ \transAcc{D̂}{c}\\
+        \text{iff }& h(s^*) \transAcc{D̂}{c}. \\[5px]
+        & s^* \transvia{a ∣ p}_{Ŝ_{e}} s' \\
+        \text{iff } &
+        (∃ a, b ∧ a = c \text{ and } 
+        s₁^* \transvia{a ∣ p}_{Ŝ_{e₁}} s')
+        \text{ or } 
+        (∃ a, \overline{b} ∧ a = c \text{ and } 
+        s₂^* \transvia{a ∣ p}_{Ŝ_{e₂}} s') \\
+        \text{iff } & 
+        (∃ a, b ∧ a = c \text{ and } 
+        h₁(s₁^*) \transvia{a ∣ p}_{D̂} h(s'))
+        \text{ or } 
+        (∃ a, \overline{b} ∧ a = c \text{ and } 
+        h₂(s₂^*) \transvia{a ∣ p}_{D̂} h(s')) \\  
+        \text{iff } & 
+        (∃ a, b ∧ a = c \text{ and } 
+        e₁ \transvia{a ∣ p}_{D̂} h(s'))
+        \text{ or } 
+        (∃ a, \overline{b} ∧ a = c \text{ and } 
+        e₂ \transvia{a ∣ p}_{D̂} h(s')) \\  
+        \text{iff }& e₁ +_b e₂ \transvia{a ∣ p}_{D̂} h(s') \\ 
+        \text{iff }& h(s^*) \transvia{a ∣ p}_{D̂} h(s')
+    \end{align*}
+
+    When \(e ≜ e₁; e₂\), by induction hypothesis, we have two homomorphisms \(h₁: Ŝ_{e₁} → D̂\) and \(h₂: Ŝ_{e₂} → D̂\).
+    We define \(h\) as follows:
+    \[h(s) ≜ \begin{cases}
+        h₁(s); e₂ & s ∈ Ŝ_{e₁} \\  
+        h₂(s) & s ∈ Ŝ_{e₂}
+    \end{cases}\]
+    Then we can prove that \(h\) is a homomorphism by case analysis on \(s\). 
+    First case is that \(s ∈ Ŝ_{e₁}\):
+    \begin{align*}
+        s \transAcc{Ŝ_{e}}{c} 
+        \text{ iff } & 
+            ∃ a, b, a ∧ b = c, 
+                s \transAcc{Ŝ_{e₁}}{a} 
+                \text{ and } 
+                s₂^* \transAcc{Ŝ_{e₂}}{b} \\
+        \text{ iff } & 
+            ∃ a, b, a ∧ b = c, 
+                h₁(s) \transAcc{D̂}{a} 
+                \text{ and } 
+                f \transAcc{D̂}{b} \\
+        \text{ iff } & h₁(s); f \transAcc{D̂}{c} 
+        \text{ iff } h(s) \transAcc{D̂}{c}.\\[5px]
+        s \transvia{c ∣ p}_{S_{e}} s'
+        \text{ iff } & 
+            (s \transvia{c ∣ p}_{Ŝ_{e₁}} s')
+            \text{ or }
+            (∃ a, b, a ∧ b = c
+                \text{ and }
+                s \transAcc{Ŝ_{e₁}}{a}
+                \text{ and }
+                s₂^* \transvia{b ∣ p}_{Ŝ_{e₂}} s') \\  
+        \text{ iff } & 
+            (h₁(s) \transvia{c ∣ p}_{D̂} h₁(s')) 
+            \text{ or }
+            (∃ a, b, a ∧ b = c
+                \text{ and }
+                h₁(s) \transAcc{D̂}{a}
+                \text{ and }
+                e₂ \transvia{b ∣ p}_{D̂} h₂(s')) \\
+        \text{ iff } & h₁(s) \transvia{c ∣ p} h(s') 
+        \text{ iff } h(s) \transvia{c ∣ p} h(s').
+    \end{align*}
+    The case where \(s₂ ∈ Ŝ_{e₂}\) is straightforward, as \(Ŝ_{e}\) preserves the transitions of \(Ŝ_{e₂}\):
+    \begin{align*}
+        & s \transAcc{Ŝ_{e}}{c} \text{ iff }
+        s \transAcc{Ŝ_{e₂}}{c} \text{ iff }
+        h₂(s) \transAcc{D̂}{c} \text{ iff }
+        h(s) \transAcc{D̂}{c}, \\ 
+        & s \transvia{c ∣ p}_{Ŝ_{e}} s' \text{ iff }
+        s \transvia{c ∣ p}_{Ŝ_{e₂}} s' \text{ iff }
+        h₂(s) \transvia{c ∣ p}_{D̂} h₂(s') \text{ iff }
+        h(s) \transvia{c ∣ p}_{D̂} h(s').
+    \end{align*}
+
+    When \(e ≜ {e₁}^{(b)}\), by induction hypothesis, we have a homomorphism \(h₁: Ŝ_{e₁} → D̂\); the homomorphism \(h\) can be defined as follows: 
+    \[h(s) ≜ \begin{cases}
+        {e₁}^{(b)} & s ≜ s^* \\  
+        h₁(s); e₁^{(b)} & s ∈ Ŝ_{e₁}
+    \end{cases}\]
+    We prove the homomorphism condition by case analysis on \(s\). First case is that \(s = s^*\), then:
+    \begin{align*}
+        (s^* \transAcc{Ŝ_e}{c})
+        \text{ iff } & (s^* \transAcc{Ŝ_e}{c} \text{ and } c = \overline{b})
+        \text{ iff } ({e₁}^{(b)} \transAcc{D̂}{c} \text{ and } c = \overline{b})
+        \text{ iff } (h(s^*) \transAcc{D̂}{c}) \\
+        (s^* \transvia{c ∣ p}_{Ŝ_e} s')
+        \text{ iff } & (∃ a, b ∧ a = c \text{ and } s₁^* \transvia{a ∣ p}_{Ŝ_{e₁}} s') \\
+        \text{ iff } & (∃ a, b ∧ a = c \text{ and } e₁ \transvia{a ∣ p}_{D̂} h₁(s')) \\ 
+        \text{ iff } & {e₁}^{(b)} \transvia{a ∣ p}_{D̂} h₁(s') 
+        \text{ iff } h(s^*) \transvia{a ∣ p}_{D̂} h(s')
+    \end{align*}
+    The second case is when \(s ∈ Ŝ_{e₁}\), then:
+    \begin{align*}
+        & s \transAcc{Ŝ_e}{c}\\
+        \text{ iff } & (∃ a, \overline{b} ∧ a = c \text{ and } s \transAcc{Ŝ_{e₁}}{a}) \\  
+        \text{ iff } & (∃ a, \overline{b} ∧ a = c \text{ and } h₁(s) \transAcc{D̂}{a}) \\
+        \text{ iff } & h₁(s);e₁^{(b)} \transAcc{D̂}{c} 
+        \text{ iff } h(s) \transAcc{D̂}{c} \\[5px]
+        & s \transvia{c ∣ p}_{Ŝ_e} s' \\
+        \text{ iff } & 
+            (s \transvia{c ∣ p}_{Ŝ_{e₁}} s' 
+            \text{ or } 
+            ∃ a₁, a₂, 
+                b ∧ a₁ ∧ a₂ = c 
+                \text{ and } 
+                s \transAcc{Ŝ_{e₁}}{a₁} 
+                \text{ and } 
+                s₁^* \transvia{a₂ ∣ p}_{Ŝ_{e₁}} s') \\
+        \text{ iff } & 
+            (h₁(s) \transvia{c ∣ p}_{D̂} h₁(s') 
+            \text{ or } 
+            ∃ a₁, a₂, b ∧ a₁ ∧ a₂ = c 
+                \text{ and } 
+                h₁(s) \transAcc{D̂}{a₁} 
+                \text{ and } 
+                e₁ \transvia{a₂ ∣ p}_{D̂} h₁(s'))\\
+        \text{ iff } & 
+            (h₁(s) \transvia{c ∣ p}_{D̂} h₁(s') 
+            \text{ or } 
+            ∃ a₁, a₂, b ∧ a₁ ∧ a₂ = c 
+                \text{ and } 
+                h₁(s) \transAcc{D̂}{a₁} 
+                \text{ and } 
+                e₁^{(b)} \transvia{b ∧ a₂ ∣ p}_{D̂} h₁(s');e₁^{(b)}) \\ 
+        \text{ iff } &
+            (h₁(s); e₁^{(b)} \transvia{c ∣ p}_{D̂} h₁(s'); e₁^{(b)}) 
+        \text{ iff } h(s) \transvia{c ∣ p}_{D̂} h(s')
+        \qedhere
+    \end{align*}
+\end{proof}
+
+% However, when considering designing a practical algorithm, we not just cares about the correctness of the derivative construction, we are also interested in the size of the generated automata. 
+% Specifically, we are interested in all the reachable states (or expressions) from any expression \(e\), which is exactly the principle sub-coalgebra \(⟨e⟩_{D̂}\).
+
+% \begin{theorem}[Complexity]
+%     The number of state in the principle sub-coalgebra \(⟨e⟩_{D̂}\) is linear to the number of primitive actions in \(e\).
 % \end{theorem}
@@
 % \begin{proof}
-%     Recall that the composition operation compose with coproduct:
-%     \[\begin{tikzcd}[row sep=3px]
-%         {A_1} & {B_1} & {C_1} \\
-%         & {+} \\
-%         {A_2} & {B_2} & {C_2}
-%         \arrow["{x_1}", from=1-1, to=1-2]
-%         \arrow["{y_1}", from=1-2, to=1-3]
-%         \arrow["{x_2}", from=3-1, to=3-2]
-%         \arrow["{y_2}", from=3-2, to=3-3]
-%     \end{tikzcd}
-%     = 
-%     A₁ + A₁ \xrightarrow{x₁ + x₂} B₁ + B₂ \xrightarrow{y₁ + y₂} C₁ + C₂.\]
-
-%     Therefore the composition of \(h₁: S → T\) and \(h₂: T → U\) can be simply defined as 
-%     the composition of its components: 
-%     \begin{align*}
-%         (h₁ ∘ h₂)_{\norm} & : S_{\norm} → U_{\norm} \\
-%         (h₁ ∘ h₂)_{\norm} & ≜ h_{1\norm} ∘ h_{2\norm} \\[5px]
-%         (h₁ ∘ h₂)_{\dead} & : S_{\dead} → U_{\dead}\\
-%         (h₁ ∘ h₂)_{\dead} & ≜ h_{1\dead} ∘ h_{2\dead} \\[5px]
-%         h₁ ∘ h₂ & : S → U\\
-%         h₁ ∘ h₂ & ≜ (h₁ ∘ h₂)_{\norm} + (h₁ ∘ h₂)_{\dead}
-%     \end{align*}
-%     concretely this can be written down as
-%     \[h₁ ∘ h₂(s) ≜ \begin{cases}
-%         h_{1\norm} ∘ h_{2\norm}(s) & \text{if } s ∈ \norm(S) \\  
-%         h_{1\dead} ∘ h_{2\dead}(s) & \text{if } s ∈ \dead(S)
-%     \end{cases}.\]
-
-%     The identity morphism is indeed the coproduct of the identity homomorphism on normalized coalgebra and the identity function on the dead states.
-
-%     The associativity and identity axiom can be verified by decomposing the homomorphism to live and dead states.
+%     To prove this, we prove the following two theorem:
+
 % \end{proof}
 
-% \begin{remark}
-%     Diagrammatically, normalized homomorphism is completely characterized by the following commutative square in the category of set:
-%     \[
-%         \begin{tikzcd}
-%             S \ar{rr}{h} \ar[swap]{dd}{δ_S} && T \ar{d}{δ_T} \\  
-%             && G(T) \ar{d}{[-]_{\norm}}\\
-%             G(S) \ar[swap]{r}{ G(h)}
-%                 & G(T) \ar[swap]{r}{[-]_{\norm}}
-%                 & G(T)/{∼_{\norm}}
-%         \end{tikzcd}
-%     \]
-%     where \({∼_{\norm}} ⊆ (2 + S × K) × (2 + S × K)\) is the minimal equivalence generated by the following relation \[\{(\reject, (s, p)) ∣ p ∈ K, \text{\(s\) is dead in \(S\)}\};\] 
-%     and \(G(T)/{∼_{\norm}}\) is the point-wise quotient of \(G(T)\) by the smallest equivalence relation containing the relation \(∼_{\norm}\), and \([-]_{\norm}\) is the function that maps a function in \(G(T)\) to its corresponding equivalence class in \(G(T)/{∼_{\norm}}\).
-
-%     The proof of this equivalence is interesting, however irrelevant to this paper, we will defer the proof to appendix, in~\cref{sec:diagram-of-norm-sim}.
-% \end{remark}
-
-% With the category of GKAT coalgebra and normalized homomorphism defined, we obtain the conventional definition of bisimulation, which is a span of projections in that category~\cite{jacobs_IntroductionCoalgebraMathematics_2016,rutten_UniversalCoalgebraTheory_2000}:
-% \[
-%     \begin{tikzcd}[column sep=1cm]
-%         S & ∼ \ar[swap]{l}{π₁} \ar{r}{π₂} & T
-%     \end{tikzcd}
-% \]
-% We call such bisimulation \emph{normalized bisimulation} to distinguish it from the bisimulation generated in the category of GKAT coalgebra and GKAT coalgebra homomorphism.
-
-% Like proving the soundness and completeness of bisimulations, we can first identify that the language semantics is indeed the unique normalized homomorphism to the final GKAT coalgebra in out category:
-% \begin{theorem}
-%     The language GKAT \(𝒢\) is the final GKAT coalgebra in the category of GKAT coalgebra and normalized homomorphisms, with the language semantics \(⟦-⟧_S: S → 𝒢\) as the unique normalized homomorphism.
-% \end{theorem}
-
-% \begin{proof}
-%     Recall that \(\norm(𝒢)\) is the final object in the category of GKAT coalgebra with homomorphisms, then there exists a unique homomorphism \(⟦-⟧_{\norm(S)}: S → 𝒢\).
-
-%     Also recall that \(\dead(𝒢)\) is a singleton set that only contains the empty language \(∅\). Recall the singleton set is the final algebra in the category of set \(!_S: S → \{∅\}\).
-
-%     And by observing the definition of \(⟦-⟧_S\), we can verify that
-%     \[⟦-⟧_S = ⟦-⟧_{\norm(S)} + !_S\]. 
-%     The above equation shows that the language semantics \(⟦-⟧_S\) is the unique normalized homomorphism from any GKAT coalgebra \(S\) to \(𝒢\), proving that \(𝒢\) is the final GKAT coalgebra in the category of GKAT coalgebra with normalized homomorphism.
-% \end{proof}
-
-% \begin{corollary}[soundness]
-%     Given two states \(s ∈ S\) and \(t ∈ T\), 
-%     \[s ∼ t ⟹ ⟦s⟧_S = ⟦t⟧_T\]
-% \end{corollary}
-
-% \begin{proof}
-%     This proof mirrors the classical soundness proof for bisimulation with regard to behavioral equivalence~\cite{rutten_UniversalCoalgebraTheory_2000}.
-
-%     Then by finality of \(𝒢\):
-%     \[
-%         \begin{tikzcd}[column sep=2cm, row sep=1cm]
-%             & 𝒢 & \\
-%             S \ar{ur}{⟦-⟧_S} 
-%                 & ∼ \ar[u, near start, dashed, "⟦-⟧_S ∘ π₁ = ⟦-⟧_T ∘ π₂"description]\ar{l}{π₁} \ar[swap]{r}{π₂} 
-%                 & T \ar[swap]{ul}{⟦-⟧_T}
-%         \end{tikzcd}
-%     \]
-%     And take any two states \(s ∈ S\) and \(t ∈ T\):
-%     \[s ∼ t ⟹ (s, t) ∈ {∼} ⟹ ⟦-⟧_S ∘ π₁(s,t) = ⟦-⟧_T ∘ π₂(s,t) ⟹ ⟦s⟧_S = ⟦t⟧_T.\]
-% \end{proof}
-
-% The completeness proof is slightly more complicated: because we use a modified notion of homomorphism, thus the classical completeness argument~\cite{rutten_UniversalCoalgebraTheory_2000} no longer applies.
-% However, we can still devise a simple proof based on the completeness of bisimulation between normalized GKAT coalgebras~\cite{smolka_GuardedKleeneAlgebra_2020}.
-
-% Given two states \(s ∈ S\) and \(t ∈ T\), we use the notation \(s ≡ t\) to denote that there exists a normalized bisimulation \({∼} ⊆ S × T\) s.t. \(s ∼ t\). 
-% \begin{remark}
-%     An alternative interpretation of \(≡\) is the maximal normalized bisimulation between \(S\) and \(T\).
-However, by using the current definition of \(≡\), we can omit the proof that all normalized bisimulations form a complete lattice.
-% \end{remark}
-
-% \begin{lemma}[normalized bisimulation between dead states]\label{thm:norm-bisim-between-dead}
-%     Given two GKAT coalgebra \(S\) and \(T\), any relation \({∼} ⊆ \dead(S) × \dead(T)\) is a normalized bisimulation between \(S\) and \(T\).
-% \end{lemma}
-
-% \begin{proof}
-%     From the definition, \(∼\) only contains one dead state \((s,t)\) and no live states, thus the projection from \(\norm(∼) = ∅\) is the unique map from the empty set \(π₁: ∅ → \norm(S)\) and \(π₂: ∅ → \norm(T)\), which are indeed a GKAT homomorphism.
-
-%     The projection from the only dead state \((s, t)\) will also reach \(s\) and \(t\), which are dead states in their respective GKAT coalgebra.
-
-%     Thus the projection from live states from a homomorphism in normalized coalgebra, and projections from dead states from a function to dead states.
-Therefore both projections are normalized homomorphism, and \(∼\) is a bisimulation.
-% \end{proof}
-
-% \begin{lemma}\label{thm:norm-bisim-from-bisim-on-norm}
-%     Given a bisimulation between normalized GKAT coalgebras \({∼} ⊆ \norm(S) × \norm(T)\), \(∼\) is also a normalized bisimulation between \(S\) and \(T\).
-% \end{lemma}
-
-% \begin{proof}
-%     TODO
-% \end{proof}
-
-% \begin{theorem}[completeness]
-%     Given two states \(s ∈ S\) and \(t ∈ T\), 
-%     \[⟦s⟧ = ⟦t⟧ ⟹ s ≡ t.\]
-% \end{theorem}
-
-% \begin{proof}
-%     We do this by case analysis on the whether \(⟦s⟧\) and \(⟦t⟧\) is empty.
-
-%     Recall that the empty language is a dead state in the language model \(𝒢\), and because \(⟦-⟧\) is a normalized homomorphism, which preserves liveness by definition.
-Hence, when \(⟦s⟧ = ⟦t⟧ = ∅\), both \(s\) and \(t\) has to be dead.
-Therefore by \cref{thm:norm-bisim-between-dead}, we can simply construct a bisimulation \(∼ ≜ \{(s, t)\}\).
-
-%     If both \(⟦s⟧\) and \(⟦t⟧\) are non-empty, then by the same reasoning, both \(s\) and \(t\) are live, then by the completeness of bisimulation on normalized GKAT coalgebra~\cite{smolka_GuardedKleeneAlgebra_2020}, there exists a bisimulation \({∼} ⊆ \norm(S) × \norm(T)\), s.t.
-\(s ∼ t\).
-% \end{proof}
-
-% \begin{lemma}\label{thm:norm-bisim-preserve-liveness}
-%     Given a normalized bisimulation \({∼} ⊆ S × T\), then by definition \(∼\) form a GKAT coalgebra, and the following statements hold:
-%     \begin{itemize}
-%         \item if \(s ∼ t\) then either \(s\) and \(t\) are both live or \(s\) and \(t\) are both dead.
-%         \item the live states in \(∼\) are exactly the states where both components are live: \[\norm(∼) ≜ \{(s, t) ∣ s ∼ t, s ∈ \norm(S), t ∈ \norm(T)\}.\]
-%     \end{itemize}
-% \end{lemma}
-
-% \begin{proof}
-%     Because \(π₁: {∼} → S\) and \(π₂: {∼} → T\) are normalized homomorphisms, then by definition, they will map live states to live states and dead states to dead states.
-%     Thus if we consider \((s, t) ∈ {∼}\): \[π₁(s, t) = s \text{ is live in } S ⟺ (s, t) \text{ is live in } {∼} ⟺ π₂(s, t) = t \text{ is live in } T.\]
-%     Thus \(∼\) only relates states with the same liveness.
-
-%     By the above equivalence chain, \((s, t) ∈ {∼}\) is live if and only if both \(s, t\) are live, thus 
-%     \[\norm(∼) ≜ \{(s, t) ∣ s ∼ t, s ∈ \norm(S), t ∈ \norm(T)\}.\]
-% \end{proof}
-
-% This observation allows us to express a strong connection between normalized bisimulation and bisimulation on normalized coalgebra.
-
-% \begin{theorem}\label{thm:bisim-on-norm<=norm-bisim}
-%     Given two GKAT coalgebra, if \({∼} ⊆ S × T\) then \(\norm(∼) ⊆ \norm(S) × \norm(T)\) is a bisimulation. 
-% \end{theorem}
-
-% \begin{proof}
-%     because \(π₁\) and \(π₂\) are normalized homomorphisms, then they can be decomposed into a homomorphism on normalized coalgebra and a function on dead states: 
-%     \[π₁ = π_{1\norm} + π_{1\dead} \text{ and } π₂ = π_{2\norm} + π_{2\dead},\] 
-%     where \(π_{1\norm}\) and \(π_{2\norm}\) are projection homomorphism to the normalized GKAT coalgebra \(\norm(S)\) and \(\norm(T)\).
-%     Therefore, \(\norm(∼)\) form a span of \(\norm(S)\) and \(\norm(T)\) with these projection homomorphisms. 
-%     Diagrammatically, this span depicts the following commutative diagram in set:
-%     \[
-%         \begin{tikzcd}[column sep=2cm]
-%             \norm(S) \ar[swap]{d}{\norm(δ_S)}& 
-%             \norm(∼) \ar[swap]{l}{π_{1\norm}} \ar{r}{π_{2\norm}} \ar{d}{\norm(δ_∼)}
-%             & \norm(T) \ar{d}{\norm(δ_T)}\\  
-%             G(\norm(S)) & G(\norm(∼)) \ar[swap]{l}{G(π_{1\norm})} \ar{r}{G(π_{2\norm})} & G(\norm(T))
-%         \end{tikzcd}
-%     \]
-%     hence \(\norm(∼)\) is indeed a bisimulation between \(\norm(S)\) and \(\norm(T)\).
-% \end{proof}
-
-% \begin{theorem}\label{thm:bisim-on-norm=>live}
-%     Given a bisimulation \({∼} ⊆ \norm(S) × \norm(T)\), then by definition \({∼}\) is also a normalized bisimulation between \(S\) and \(T\).
-% \end{theorem}
-
-% \begin{proof}
-%     because \(∼\) is a bisimulation, then the projection \(π₁: {∼} → \norm(S)\) and \(π₂: {∼} → \norm(T)\) are indeed  GKAT homomorphisms. 
-%     By~\cref{thm:hom-preserve-liveness}, \(\dead(∼)\) is the empty set, thus have unique function 
-% \end{proof}
-
-% With the connection between normalization and normalized bisimulation, we can utilize the soundness and completeness of bisimulation between normal GKAT coalgebra to derive the soundness and completeness of normalized bisimulation.
-% Given two states \(s ∈ S\) and \(t ∈ T\), we use the notation \(s ≡ t\) to denote that there exists a normalized bisimulation \({∼} ⊆ S × T\) s.t. \(s ∼ t\). 
-% \begin{remark}
-%     An alternative interpretation of \(≡\) is the maximal normalized bisimulation between \(S\) and \(T\).
-However, by using the current definition of \(≡\), we can omit the proof that all normalized bisimulation form a complete lattice, which is a corollary that bisimulation forms a complete lattice.
-% \end{remark}
-
-% \begin{theorem}[soundness and completeness]
-%     Given two states \(s ∈ S\) and \(t ∈ T\):
-%     \[s ≡ t ⟺ ⟦s⟧ = ⟦t⟧.\]
-% \end{theorem}
-
-% \begin{proof}
-%     We first show the \(⟹\) direction, because \(s ≡ t\), then there exists \(∼\) s.t. \(s ∼ t\). And by~\cref{thm:bisim-on-norm<=norm-bisim}, \(\norm(∼)\) is a bisimulation between 
-% \end{proof}
-
-% \subsection{Construction}
-
-% \begin{lemma}
-%     Given a sub-GKAT coalgebra \(S' ⊑ S\), the inclusion homomorphism \(i: S' → S\) is also a normalized homomorphism.
-% \end{lemma}
-% \begin{proof}
-%     TODO.
-% \end{proof}
-
-% \begin{theorem}[sub-coalgebra perserves normalized bisimulation]\label{thm:sub-coalg-preserve-norm-bisim}
-%     Given two sub-GKAT colagebra \(S' ⊑ S\) and \(T' ⊑ T\), then 
-%     \begin{itemize}
-%         \item a normalized bisimulation \(∼_{S', T'}\) between \(S'\) and \(T'\) is also a normalized bisimulation between \(S\) and \(T\)
-%         \item a normalized bisimulation \(∼\) between \(S\) and \(T\), can also be restricted to \(S'\) and \(T'\) in the following manner, by restricting both the carrier and the transition:
-%         \begin{align*}
-%             {∼_{S', T'}} & ≜ \{(s,t) ∣ s ∼ t, s ∈ S', t ∈ T'\} \\
-%             δ_{∼_{S', T'}} & ≜ δ_{∼} |_{∼_{S', T'}}
-%         \end{align*}
-%     \end{itemize}
-% \end{theorem}
-
-% \begin{proof}
-%     TODO.
-% \end{proof}
-
-% \begin{lemma}[normalized bisimulation closed under arbitrary union]\label{thm:norm-bisim-close-under-union}
-%     Given a collection of normalized bisimulation \(∼ᵢ ⊆ S × T\) indexed by an arbitrary set \(i ∈ I\), then there exists a normalized bisimulation \(∼\) that includes all the pairs in \(∼ᵢ\) for all \(i\).
-% \end{lemma}
-
-% \begin{theorem}[inductive construction]\label{thm:inductive-construction}
-%     Given two GKAT coalgebra \(S\) and \(T\), and two of their elements \(s ∈ S\) and \(t ∈ T\),
-%     there exists a normalized bisimulation \({∼} ⊆ ⟨s⟩ × ⟨t⟩\) s.t. \(s ∼ t\), if and only if all of the following holds:
-%     \begin{enumerate}
-%         \item\label{itm:acc-condition} \(δ_{S}(s, α) = \accept ⟺ δ_{T}(t, α) = \accept\);
-%         \item\label{itm:transition-bisim} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{T}(t, α) = (t', p)\), then there exists a normalized bisimulation \({∼_{s',t'}}\) on \(⟨s'⟩\) and \(⟨t'⟩\), s.t.
-\(s' ∼_{s',t'} t'\);
-%         \item\label{itm:transition-dead} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{T}(t, α) = (t', q)\), s.t. \(p ≠ q\), then both \(s'\) and \(t'\) are dead;
-%         \item\label{itm:rej-or-dead} \(s\) reject \(α\) or transition to a dead state via \(α\) if and only if \(t\) rejects \(α\) or transition to a dead state via \(α\).
-%     \end{enumerate}
-% \end{theorem}
-
-% \begin{proof}
-%     We first prove \(⟹\) direction, recall that \(∼\) is a GKAT coalgebra,
-%     where the projection from the normalized version \(π₁ : \norm(∼) → \norm(S)\) and \(π₂: \norm(∼) → \norm(T)\), are GKAT homomorphisms:
-%     \[
-%         \begin{tikzcd}
-%             \norm(S) \ar[swap]{d}{\norm(δ_S)} 
-%             & \norm(∼) \ar{d}{\norm(δ_∼)} \ar[swap]{l}{π₁} \ar{r}{π₂}
-%             & \norm(T) \ar{d}{\norm(δ_T)}\\  
-%             G(\norm(S)) 
-%             & G(\norm(∼)) \ar{l}{G(π₁)} \ar[swap]{r}{G(π₂)} 
-%             & G(\norm(T)) 
-%         \end{tikzcd}
-%     \]
-
-%     First realize that in order for either \(s\) or \(t\) to have an accepting transition, it has to be a live state.
-Because \(s ∼ t\) and normalized bisimulation preserves liveness (\Cref{thm:norm-bisim-preserve-liveness}), both \(s\) and \(t\) has to be live.
-Therefore, condition \labelcref{itm:acc-condition} holds:
-%     \begin{align*}
-%         δ_S(s, α) = \accept 
-%         & ⟺ \norm(δ_S)(s, α) = \accept 
-%         ⟺ \norm(δ_∼)((s,t), α) = \accept \\ 
-%         & ⟺ \norm(δ_T)(t, α) = \accept 
-%         ⟺ δ_T(t, α) = \accept
-%     \end{align*}
-
-%     Notice that if \(s ∼ t\), \(s\) transition to \(s'\), and \(t\) transition to \(t'\), then \(s'\) and \(t'\) are either both dead or both live: 
-%     \begin{itemize}
-%         \item if \(s, t\) are dead, then \(s'\) and \(t'\) has to both be dead, because reachable state of \(s\) and \(t\) has to be dead (\cref{thm:dead-iff-all-reachable-dead}).
-%         \item if \(s, t\) are live, then \((s, t) ∈ \norm(∼)\). Because \(\norm(∼)\) is a bisimulation and by its equation, \(s'\) and \(t\) has to be both live or dead.
-%     \end{itemize}
-%     Condition~\labelcref{itm:transition-bisim} can be proven by case analysis on the liveness of \(s'\) and \(t'\):
-%     \begin{itemize}
-%         \item if both \(s'\) and \(t'\) are live, then both \(s\) and \(t\) are live:
-%         \[(s, t) ∈ \norm(∼) ⟹ (s', t') ∈ \norm(∼)\]
-%         Then, \(∼_{s',t'}\) can be constructed by restricting \(∼\) to \(⟨s'⟩\) and \(⟨t'⟩\), like stated in~\cref{thm:sub-coalg-preserve-norm-bisim}.
-%         \item if both \(s'\) and \(t'\) are dead, then we can construct \(∼_{s',t'}\) as described in~\cref{thm:norm-bisim-between-dead}.
-%     \end{itemize}
-%     Thus, concluding our proof of condition~\labelcref{itm:transition-bisim}.
-
-%     Then for condition~\labelcref{itm:transition-dead}, if both \(s\) and \(t\) are live, then this transition does not satisfy the condition of bisimulation, therefore \(s'\) and \(t'\) has to be live.
-If both \(s\) and \(t\) are dead, then all the reachable state from \(s\) and \(t\) has to be dead (\cref{thm:dead-iff-all-reachable-dead}), therefore \(s'\) and \(t'\) will also need to be dead.
-
-%     Finally the proof of condition~\labelcref{itm:rej-or-dead} is similar to the proof of condition~\labelcref{itm:transition-dead}, which is proven by case analysis on the liveness of \(s\) and \(t\):
-%     \begin{itemize}
-%         \item if both \(s\) and \(t\) are dead, by \cref{thm:dead-iff-all-reachable-dead} and definition, they can only transition to dead states or reject.
-%         \item if both \(s\) and \(t\) are live, then \((s, t) ∈ \norm(∼)\), and
-%         \begin{align*}
-%             & δ_S(s, α) = \reject \text{or transition to dead states} \\
-%             & ⟺ \norm(δ_S)(s, α) = \reject  
-%             ⟺ \norm(δ_∼)((s,t), α) = \reject 
-%             ⟺ \norm(δ_T)(t, α) = \reject \\
-%             & ⟺ δ_T(t, α) = \reject \text{or transition to dead states}
-%         \end{align*}
-%     \end{itemize}
-
-%     Then we show the \(⟸\) direction. We use \(∼_{s',t'}\) to denote the maximal bisimulation between \(⟨s'⟩\) and \(⟨t'⟩\):
-%     \[∼' ≜ ⋃ \{{∼_{s', t'}} ∣ ∃ α ∈ \At, p ∈ K, δ_S(s, α) = (s', p), δ_T = (t', p)\}\]
-%     Consider all the \(s'\) and \(t'\) mentioned in the above definition, because they are reachable from \(s\) and \(t\) respectively, then \(⟨s'⟩ ⊑ ⟨s⟩\) and \(⟨t'⟩ ⊑ ⟨t⟩\).
-And by~\cref{thm:sub-coalg-preserve-norm-bisim}, \(∼_{s',t'}\) in above definitions are all bisimulation between \(⟨s⟩\) and \(⟨t⟩\).
-Finally, because union of normalized bisimulation is still a normalized bisimulation (\cref{thm:norm-bisim-close-under-union}), \(∼'\) is a normalized bisimulation between \(S\) and \(T\).
-
-%     Finally we will need to show that the addition of \((s, t)\) also gives a normalized bisimulation.
-% \end{proof}
-
-% \section{Normalized Bisimulation .bak}
-
-% The normalization procedure for GKAT requires the entire automaton to be known prior to the execution of the bisimulation algorithm; specifically, in order to compute the liveness of a state \(s\), it is necessary iterate through all its reachable states \(⟨s⟩\) to see if there are any accepting states within.
-% This limitation poses challenges to design an efficient on-the-fly algorithm for GKAT.
-% In order to make the decision procedure scalable, we will need to merge the normalization and bisimulation procedure, so that our algorithm can normalized the automaton only when we need to.
-
-% \subsection{the new norm}
-
-% normalization is a important procedure, as it identifies the trace model \(𝒢\) as the final coalgebra in normalized GKAT coalgebra, enabling its computation via bisimulation~\cite{smolka_GuardedKleeneAlgebra_2020}.
-% Specifically, given a GKAT coalgebra \(S\), we can normalized it by taking out all of its live state, denoted as \(\norm(S)\), and reroute all the transitions to dead state into reject:
-% \begin{align*}
-%     \norm(δ_S) & : \norm(S) → G(\norm(S))\\\
-%     \norm(δ_S)(s) & = \begin{cases}
-%         \reject & \text{if } δ_S(s) = (s', p) \text{ and \(s'\) is dead}\\
-%         δ_S(s) & \text{otherwise}
-%     \end{cases}
-% \end{align*}
-% Similarly, we will denote the dead states of \(S\) as \(\dead(S)\), and recall that \(S\) can be partitioned into its dead and live states \(S = \norm(S) + \dead(S)\), where the \(+\) is the coproduct operation in set.
-
-% Unfortunately, this style of normalization~\cite{smolka_GuardedKleeneAlgebra_2020} does not naturally induce a on-the-fly decision algorithm, as removing the dead state necessitates detecting dead states prior to execution of bisimulation.
-In here, we introduce a new notion of normalization, which is implemented by relaxing the equation of coalgebra homomorphism.
-
-% We first define the relation \(∼_{\norm}\) on results of transitions, specifically for any GKAT coalgebra \(S\):
-% \[{∼_{\norm}} ⊆ (2 + S × K) × (2 + S × K)\]
-% is the minimal equivalence generated by the following relation
-% \[\{(\reject, (s, p)) ∣ p ∈ K, \text{\(s\) is dead in \(S\)}\}.\]
-% Intuitively, \({∼_{\norm}}\) will relate all the transition to dead state with rejection, while other results, like acceptance or transitions to the live states, are only related to themselves.
-
-% A function \(h: S → T\) is a \emph{normalized homomorphism} if the following diagram commute:
-% \[
-%     \begin{tikzcd}
-%         S \ar{rr}{h} \ar[swap]{dd}{δ_S} && T \ar{d}{δ_T} \\  
-%         && G(T) \ar{d}{[-]_{\norm}}\\
-%         G(S) \ar[swap]{r}{ G(h)}
-%             & G(T) \ar[swap]{r}{[-]_{\norm}}
-%             & G(T)/{∼_{\norm}}
-%     \end{tikzcd}
-% \]
-% where \(G(T)/{∼_{\norm}}\) is the point-wise quotient of \(G(T)\) by the smallest equivalence relation containing the relation \(∼_{\norm}\), and \([-]_{\norm}\) is the function that maps a element in \(G(T)\) to its corresponding equivalence class in \(G(T)/{∼_{\norm}}\).
-% For brevity, we sometimes leave the function \([-]_{\norm}\) implicit and draw the commutative diagram as follows:
-% \[
-%     \begin{tikzcd}
-%         S \ar{r}{h} \ar[swap]{d}{δ_S} & T \ar{d}{δ_T} \\  
-%         G(S) \ar[swap]{r}{G(h)} & G(T)/{∼_{\norm}}
-%     \end{tikzcd}
-% \]
-
-% \begin{example}
-
-% \end{example}
-
-% The notion of normalized homomorphism indeed have connection with existing noting of normalization
-
-% %%%%% NOTE: this proof is incorrect, we cannot verify the function 
-% %%%%% h : \norm(S) -> \norm(T) is well defined
-% %
-% % \begin{theorem}
-% %     Normalized homomorphism induces a homomorphism on normalized coalgebra:
-% %     \[
-% %         \begin{tikzcd}
-% %             S \ar{r}{h} \ar[swap]{d}{δ_S} & T \ar{d}{δ_T} \\  
-% %             G(S) \ar[swap]{r}{G(h)} & G(T)/{∼_{\norm}}
-% %         \end{tikzcd}
-% %         ⟹ 
-% %         \begin{tikzcd}
-% %             \norm(S) \ar{r}{h} \ar[swap]{d}{\norm(δ_S)} 
-% %                 & \norm(T) \ar{d}{\norm(δ_T)} \\  
-% %             G(\norm(S)) \ar[swap]{r}{G(h)} & G(\norm(T))
-% %         \end{tikzcd}
-% %     \]
-% % \end{theorem}
-
-% % \begin{proof}
-% %     The commutative triangle in the premise can be written as the following equation:
-% %     \[G(h) ∘ δ_S ∼_{\norm} δ_T ∘ h.\] 
-% %     Take any live state \(s ∈ S\), we will show the above implication by case analysis on the value on the bottom right corner \((G(h) ∘ δ_S)(s, α)\) for an arbitrary \(α ∈ \At\):
-% %     \begin{itemize}
-% %         \item if \((G(h) ∘ δ_S)(s, α)\) transitions to dead state or reject, then \((δ_T ∘ h)(s, α)\) will also transition to dead state or reject. Therefore 
-% %         \[(\norm(δ_T) ∘ h)(s, α) = \reject.\]
-% %         We then case on the result of \(δ_S(s, α) ∈ G(S)\):
-% %         \begin{itemize}
-% %             \item If \(δ_S(s, α)\) transitions to dead state or reject, then 
-% %             \[\norm(δ_S)(s, a) = \reject;\] 
-% %             and 
-% %             \begin{align*}
-% %                  (G(h) ∘ \norm(δ_S))(s, α) = \reject = (\norm(δ_T) ∘ h)(s, α).
-% %             \end{align*}
-% %             \item \(δ_S(s, α)\) transition to live state or accept, then by the definition of norms:
-% %             \[\norm(δ_S)(s, a) = δ_S(s, α);\]
-% %             which implies 
-% %             \[(G(h) ∘ \norm(δ_S))(s, α) = (G(h) ∘ δ_S)(s, α)\]
-% %             Finally, because \((G(h) ∘ δ_S)(s, α)\) reject or transition to dead states, yet dead states are not in \(\norm(T)\).
-% %             \[(G(h) ∘ \norm(δ_S))(s, α) = \reject = (\norm(δ_T) ∘ h)(s, α).\]
-% %         \end{itemize}
-% %         \item if 
-% %     \end{itemize}
-% % \end{proof}
-
-% \begin{lemma}\label{thm:norm-hom-to-norm-is-hom}
-%     If given a normal GKAT coalgebra \(T\), then a normalized homomorphism \(S → T\) is a GKAT coalgebra homomorphism from \(S → T\).
-% \end{lemma}
-
-% \begin{proof}
-%     Since there is no dead state in \(T\), then \(∼_{\norm}\) is just the identity relation, thus \(G(T)/∼_{\norm} = G(T)\).
-%     Hence the following diagram commute for \(h\):
-%     \[
-%         \begin{tikzcd}
-%             S \ar{r}{h} \ar[swap]{d}{δ_S} & T \ar{d}{δ_T} \\  
-%             G(S) \ar[swap]{r}{G(h)} & G(T)
-%         \end{tikzcd}
-%     \]
-%     Therefore \(h\) is a GKAT coalgebra homomorphism.
-% \end{proof}
-
-% \begin{lemma}
-%     Given a normal GKAT coalgebra \(T\) and a normalized bisimulation \(h: S → T\), then for all \(s ∈ S\), the principle sub-domain \(⟨h(s)⟩_T\) is in the range of \(h\).
-% \end{lemma}
-
-% \begin{proof}
-%     Because \(T\) is normal and by~\cref{thm:norm-hom-to-norm-is-hom}, \(h\) is a normalized homomorphism, thus we can construct a epi-mono factorization~\cite[Theorem 7.1]{rutten_UniversalCoalgebraTheory_2000}:
-%     \[
-%         \begin{tikzcd}
-%             S \ar{r}{h} \ar[swap]{d}{δ_S} 
-%                 & h(S) \ar[hook]{r}{i} \ar{d}{δ_T}
-%                 & T \ar{d}{δ_T} \\  
-%             G(S) \ar[swap]{r}{G(h)} 
-%                 & G(h(S)) \ar[hook]{r}{i}
-%                 & G(T)
-%         \end{tikzcd}
-%     \]
-%     Thus \(h(S)\) is a sub-GKAT coalgebra of \(T\). Because \(h(s) ∈ h(S)\) and \(⟨h(s)⟩_T\) is the smallest sub-GKAT coalgebra that contains \(h(S)\), \[⟨h(s)⟩_T ⊑ h(S) ⟹ ⟨h(s)⟩_T ⊆ h(S).\]
-% \end{proof}
-
-% \begin{theorem}[normalized homomorphism preserves liveness]
-%     Given a normalized homomorphism \(h: S → T\), then 
-%     \begin{align*}
-%         s \text{ is accepting} & ⟺ h(s) \text{ is accepting} \\
-%         s \text{ is live} & ⟺ h(s) \text{ is live} \\
-%         s \text{ is dead} & ⟺ h(s) \text{ is dead}
-%     \end{align*}
-% \end{theorem}
-
-% \begin{theorem}[decomposition]\label{thm:norm-hom-decomp}
-%     Let \(S\) and \(T\) be GKAT coalgebras, a function \(h: S → T\) is a normalized homomorphism if and only if \(h = h_{\norm} + h_{\dead}\) where \(h_{\norm} : \norm(S) → \norm(T)\) is a \emph{homomorphism}, and \(h_{\dead}: \dead(S) → \dead(T)\) is a function.
-% \end{theorem}
-% \begin{proof}
-%     The \(⟹\) direction can be shown by constructing \(h_{\norm}\) and \(h_{\dead}\) as restrictions of \(h\):
-%     \begin{mathpar}
-%         h_{\norm} ≜ h|_{\norm(S)}
-%         \and 
-%         h_{\dead} ≜ h|_{\dead(S)}
-%     \end{mathpar}
-%     and because 
-% \end{proof}
-
-% \begin{lemma}[homomorphic pre-image preserves normalized bisimulation]
-%     Given two GKAT coalgebra homomorphism \(h₁: S₁ → T₁\) and \(h₂: S₁ → T₁\), with a normalized bisimulation \(∼\), then the homomorphic pre-image of \(∼\) with respect to \(h₁ × h₂\):
-%     \[(h₁ × h₂)^{-1}(∼) ⊆ S₁ × S₂,\]
-%     is a normalized bisimulation between \(S₁\) and \(S₂\).
-% \end{lemma}
-% \begin{proof}
-%     Recall that pre-image in the category of set forms a (weak) pullback, and because \(G\) is a elementary polynomial functor~\cite[Definition 2.2.1]{jacobs_IntroductionCoalgebraMathematics_2016} hence preserves weak pullback~\cite[Lemma 3.3.6]{jacobs_IntroductionCoalgebraMathematics_2016}.
-Therefore we obtain the following pullback squares: we denote \((h₁ × h₂)^{-1}(∼)\) as \(∼_S\),
-%     \begin{mathpar}
-%         \begin{tikzcd}
-%             ∼_S \ar[hook]{r}{i} \ar[swap]{d}{h₁ × h₂} & S₁ × S₂ \ar{d}{h₁ × h₂} \\  
-%             ∼ \ar[hook]{r}{i} & T₁ × T₂ \\
-%         \end{tikzcd}
-%         \and 
-%         \begin{tikzcd}
-%             G(∼_S) \ar[hook]{r}{G(i)} \ar[swap]{d}{G(h₁ × h₂)} & G(S₁ × S₂) \ar{d}{G(h₁ × h₂)}\\  
-%             G(∼) \ar[hook]{r}{G(i)} & G(T₁ × T₂) \\
-%         \end{tikzcd}
-%     \end{mathpar}
-
-% \end{proof}
-
-% \begin{theorem}[completeness]
-%     Given two GKAT coalgebra \(S, T\) and two states \(s ∈ S\) and \(t ∈ T\), 
-%     \[⟦s⟧ = ⟦t⟧ ⟺ \text{there exists a bisimulation \({∼} ⊆ S × T\) s.t. } s ∼ t.\]
-% \end{theorem}
-% \begin{proof}
-
-% \end{proof}
-
-% \subsection{definitions and construction}
-
-% Notice that the finite trace semantics \(⟦-⟧: S → 𝒢\) is not a GKAT coalgebra homomorphism; we will present a counter example:
-% \begin{example}
-%     We consider a empty set of primitive boolean \(B = ∅\), then the only atom is the constant true formula.
-And the example consists of the following transition function \(δ₂: 2 → \{\true\}→ G(2)\), where the states \(2 ≜ \{0, 1\}\):
-%     \begin{mathpar}
-%         δ₂(1, \true) = (0, p) \and δ₂(0, \true) = \reject
-%     \end{mathpar}
-%     Because both state \(0\) and \(1\) are dead, then \(⟦0⟧ = ⟦1⟧ = ∅\).
-%     In order for \(⟦-⟧\) to be homomorphism, \(δ_𝒢 ∘ ⟦-⟧ = G(⟦-⟧) ∘ δ₂\) needs to be satisfied, however
-%     \begin{align*}
-%         (δ_𝒢 ∘ ⟦-⟧)(1) & = δ_𝒢(∅) = \{\true ↦ \reject\}.\\[10px]
-%         (G(⟦-⟧) ∘ δ₂)(1) & = G(⟦-⟧)\{\true ↦ (0, p)\} \\
-%             & = \{\true ↦ (⟦0⟧, p)\} = \{\true ↦ (∅, p)\}.
-%     \end{align*}
-%     Thus proving that \(⟦-⟧\) is not a GKAT coalgebra homomorphism.
-% \end{example}
-
-% Normalized homomorphism inherits many important properties of homomorphism; for example when composing two normalized homomorphism, we will obtain a normalized homomorphism.
-And the identity homomorphism is the identity of normalized homomorphism under composition.
-% This notion gives rise to a
-
-% \begin{definition}[normalized closure]\label{def:norm-closure}
-%     Given two GKAT coalgebra \(S, T\), and a relation on these two expression \({∼} ⊆ S × T\), we can define \({\clos{∼}} ⊆ (2 + S × K) × (2 + T × K)\) as the smallest relation generated by the following rules: for all \(p, q ∈ K\) and states \(s' ∈ S\), \(t' ∈ T\):
-%     \begin{mathpar}
-%         \inferrule[Bisim-Acc]{\\}{\accept \clos{∼} \accept}\label{rule:bisim-acc}
-%         \and 
-%         \inferrule[Bisim-Rej]{\\}{\reject \clos{∼} \reject}\label{rule:bisim-rej}
-%         \and 
-%         \inferrule[Bisim-Trans]{s' ∼ t'}{(s', p) \clos{∼} (t', p)}\label{rule:bisim-trans}
-%         \and
-%         \inferrule[Norm-Left]{s' \text{ is dead}}{(s', p) \clos{∼} \reject}\label{rule:norm-left}
-%         \and 
-%         \inferrule[Norm-Right]{t' \text{ is dead}}{\reject \clos{∼} (t', p)}\label{rule:norm-right}
-%         \and
-%         \inferrule[Norm-Both]{s', t' \text{ are both dead}}{(s', p) \clos{∼} (t', q)}\label{rule:norm-both}
-%     \end{mathpar}
-% \end{definition}
-% Intuitively, \({\clos{∼}}\) enlarges the relation \(∼\) for \emph{results} of a transition. For example, it allows transition to dead states to equivalent to rejection, and so on.
-
-% This extension enables ``on-the-fly normalization'', where to determine whether \(s \clos{∼} t\), we can simply first check for bisimulation using the bisimulation rules; and if all failed, we can check the normalization rules.
-
-% \begin{definition}
-%     Given two GKAT coalgebra \(S\) and \(T\), a relation \({∼} ⊆ S × T\) is a normalized bisimulation when there exists a function \({∼} → {\clos{∼}}^\At\), such that the following diagram commutes:
-%     \[
-%         \begin{tikzcd}[column sep=1.5cm, row sep=1cm]
-%             S \ar[swap]{d}{δ_S}
-%                 & ∼  \ar[dashed, swap]{d}{∃} \ar[swap]{l}{π₁} \ar{r}{π₂}
-%                 & T \ar{d}{δ_T} \\  
-%             (2 + S × K)^\At 
-%                 & ({\clos{∼}}^\At) 
-%                     \ar[swap]{l}{{π₁}^\At} \ar{r}{{π₂}^\At}
-%                 & (2 + T × K)^\At
-%         \end{tikzcd}
-%     \]
-% \end{definition}
-% Notice, unlike \(G\), \(\clos{(-)}^\At\) is not a functor, since it is only defined on relations.
-
-% \begin{example}
-%     Let's consider a example with a single primitive test \(B ≜ \{b\}\),
-%     thus atoms are defined as \(\At_B = \{\{b\}, ∅\}\).
-%     We focus on two states in the bisimulation \(s\) and \(t\).
-%     with their respective transition defined as follows 
-%     \begin{align*}
-%             δ_S(s, \{b\}) & ≜ 0   
-%                 & δ_S(t, \{b\}) & ≜ (t', p) \\  
-%             δ_S(s, ∅) & ≜ (s', p) 
-%                 & δ_S(t, ∅) & ≜ (t', p) 
-%     \end{align*}
-%     It is quite clear \(s\) and \(t\) will not be bisimular, however, there can be a normalized bisimulation relation between them, provided that \(s' ∼ t'\) and \(t'\) is dead: 
-%     \[
-%     \begin{tikzcd}[column sep=1.5cm, row sep=1cm]
-%         s \ar[swap, mapsto]{d}{δ_S}
-%             & (s, t)  \ar[dashed, swap, mapsto]{d}{∃} \ar[swap, mapsto]{l}{π₁} \ar[mapsto]{r}{π₂}
-%             & t \ar[mapsto]{d}{δ_T} \\  
-%         \begin{gathered}
-%             \{b\}↦ 0 \\  
-%             ∅ ↦ (s', p)
-%         \end{gathered} 
-%             & \begin{gathered}
-%                 \{b\}↦ 0, (t', p) \\  
-%                 ∅ ↦ (s', p), (t', p)
-%             \end{gathered} 
-%                 \ar[swap, mapsto]{l}{{π₁}^\At} \ar[mapsto]{r}{{π₂}^\At}
-%             & \begin{gathered}
-%                 \{b\} ↦ (t', p) \\  
-%                 ∅ ↦(t', p)
-%             \end{gathered}
-%     \end{tikzcd}
-%     \]
-%     Notice that \(0 \clos{∼} (t', p)\) because \(t'\) is dead, 
-%     and \((s', p) \clos{∼} (t', p)\) because we assume that \(s' ∼ t'\). 
-%     Furthermore, if both \(s'\) and \(t'\) are dead, we can establish \(s' \clos{∼} t'\) using either the \textnormal{\textsc{Bisim-Trans}} rule or \textnormal{\textsc{Norm-Both}} rule, and we are free to pick the more efficient check in our implementation.
-% \end{example}
-
-% \begin{theorem}[Universiality]
-%     Given a normalized bisimulation relation \({∼} ⊆ S × T\)
-%     \[
-%         \begin{tikzcd}[column sep=1.5cm, row sep=1cm]
-%             S \ar[swap]{d}{δ_S}
-%                 & ∼  \ar[dashed, swap]{d}{h} \ar[swap]{l}{π₁} \ar{r}{π₂}
-%                 & T \ar{d}{δ_T} \\  
-%             (2 + S × K)^\At 
-%                 & (\clos{∼})^\At 
-%                     \ar[swap]{l}{{π₁}^\At} \ar{r}{{π₂}^\At}
-%                 & (2 + T × K)^\At
-%         \end{tikzcd}
-%     \] 
-%     Then \(h\) is unique, and defined as \(δ_{S}(s) × δ_{T}(t)\).
-% \end{theorem}
-
-% \begin{proof}
-%     The commutativity can be verified by computation; and the uniqueness can be observed using the property of pairs. We will show an alternative diagrammatical proof for reference.
-
-%     The above commutative diagram implies the commutativity of the following diagram:
-%     \[
-%         \begin{tikzcd}
-%             & {∼} \ar{d}{h} \ar[swap]{ddl}{δ_S ∘ π₁} \ar{ddr}{δ_T ∘ π₂}& \\
-%             & (\clos{∼})^\At \ar[hook]{d}{i} & \\  
-%             (2 + S × K)^\At & 
-%                 (2 + S × K)^\At × (2 + T × K)^\At
-%                     \ar{l}{π₁^\At} \ar[swap]{r}{π₂^\At}
-%                 & (2 + T × K)^\At
-%         \end{tikzcd}
-%     \] 
-%     Since \((-)^{\At}\) is a right adjoint and right adjoints preserve limits, \((2 + S × K)^\At × (2 + T × K)^\At\) is the product of \((2 + T × K)^\At\) and \((2 + T × K)^\At\). 
-%     Hence, \(h ∘ i\) is the unique function to make the above diagram commute. 
-%     Assume we can replace \(h\) with \(g\) while preserving the commutativity of the diagram: 
-%     \begin{align*}
-%         & i ∘ h  = i ∘ g & \text{\((2 + S × K)^\At × (2 + T × K)^\At\) is a product.} \\  
-%         ⟹ {}& h = g & \text{\(i\) is a embedding, hence left cancelable.}
-%     \end{align*}
-% \end{proof}
-
-% The above theorem is particularly important, as in order to check whether a relation \(∼\) is a bisimulation, we only need to check whether 
-% \[s ∼ t ⟹ ∀ α ∈ \At, δ_S(s, α) \clos{∼} δ_T(t, α).\]
-
-% A easy normalized bisimulation we can construct is between two dead states
-% \begin{lemma}[normalized bisimulation between dead states]\label{thm:norm-bisim-dead-states}
-%     Given two dead states \(s ∈ S\) and \(t ∈ T\), then the singleton \({∼} ≜ \{(s, t)\}\) is a normalized bisimulation between \(s\) and \(t\).
-% \end{lemma}
-% \begin{proof}
-%     By~\cref{thm:dead-iff-all-reachable-dead,def:liveness-of-states}, dead states can only transition to either dead states or rejection. 
-%     Thus, for all \(α ∈ \At\), \(δ_S(s, α) \clos{∼} δ_T(t, α)\) by \nameref{rule:norm-left}, \nameref{rule:norm-right}, or \nameref{rule:norm-both}.
-% \end{proof}
-
-% \begin{lemma}\label{thm:sub-coalg-preserve-norm-bisim}
-%     Normalized bisimulation is preserved under sub-GKAT coalgebra.
-Given two sub-GKAT coalgebra \(S' ⊑ S\) and \(T' ⊑ T\) and a normalized bisimulation \(∼\) between \(S\) and \(T\), then the restriction
-%     \[∼_{S', T'} ≜ \{(s, t) ∣ s ∈ S', t ∈ T', s ∼ t\}\]
-%     is a normalized bisimulation on between \(S'\) and \(T'\).
-%     Similarly, every normalized bisimulation \(∼_{S', T'}\) between \(S'\) and \(T'\) is also a normalized bisimulation between \(S\) and \(T\).
-% \end{lemma}
-
-% Thus, this gives us a way to algorithmically construct a normalized bisimulation on principle sub-GKAT coalgebra.
-% This normalized bisimulation can also be extended to a normalized bisimulation on the entire coalgebras.
-
-% \begin{theorem}[inductive construction]\label{thm:inductive-construction}
-%     Given two GKAT coalgebra \(S\) and \(T\), and two of their elements \(s ∈ S\) and \(t ∈ T\),
-%     there exists a normalized bisimulation \({∼} ⊆ ⟨s⟩ × ⟨t⟩\) s.t. \(s ∼ t\), if and only if either both \(s\) and \(t\) are dead or for all \(α ∈ \At\), all of the following holds:
-%     \begin{enumerate}
-%         \item \(δ_{S}(s, α) = \accept ⟺ δ_{T}(t, α) = \accept\);
-%         \item\label{itm:transition-condition} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{T}(t, α) = (t', p)\), then there exists a normalized bisimulation \({∼_{s',t'}}\) on \(⟨s'⟩\) and \(⟨t'⟩\), s.t.
-\(s' ∼_{s',t'} t'\);
-%         \item If \(δ_{S}(s, α) = (s', p)\) and \(δ_{T}(t, α) = (t', q)\), s.t. \(p ≠ q\), then both \(s'\) and \(t'\) are dead;
-%         \item \(s\) reject \(α\) or transition to a dead state via \(α\) if and only if \(t\) rejects \(α\) or transition to a dead state via \(α\).
-%     \end{enumerate}
-% \end{theorem}
-
-% \begin{proof}
-%     We first prove the \(⟹\) direction: if there exists a normalized bisimulation \(∼\), then for all \(α ∈ \At\), \(δ_{S}(s, α) \clos{∼} δ_{T}(t, α)\).
-And by unfolding the definition of \(\clos{∼}\), we can obtain all the conditions we desired.
-%     We will show condition \labelcref{itm:transition-condition} as an example:
-
-%     Because \(∼\) is a normalized bisimulation: \[(s', p) \clos{∼} (t', p).\]
-%     This relation can be obtained by two rules: \nameref{rule:bisim-trans} and \nameref{rule:norm-both}:
-%     \begin{itemize}
-%         \item if \((s', p) \clos{∼} (t', p)\) is obtained from \nameref{rule:bisim-trans}, then \(s' ∼ t'\); hence the bisimulation \(∼_{s',t'}\) can be obtained from restricting \(∼\) to \(⟨s'⟩\) and \(⟨t'⟩\)
-%         \item if \((s', p) \clos{∼} (t', p)\) is obtained from \nameref{rule:norm-both}, then by \cref{thm:norm-bisim-dead-states}, there is indeed a noramlized bisimulation where \(⟨s'⟩ ∼_{s',t'} ⟨t'⟩\).
-%     \end{itemize}
-
-%     For the \(⟸\) direction, we can explicitly construct \(∼\) as the union of all the normalized bisimulations \(∼_{s', t'} ⊆ ⟨s'⟩ × ⟨t'⟩\), for all reachable states pair \(s', t'\) from state \(s\) and \(t\):
-%     \[{∼} = \{(s, t)\} ∪ \{{∼_{s', t'}} ∣ ∃ α ∈ \At, p ∈ K, δ_S(s, α) = (s', p) \text{ and } δ_T(t, α) = (t', p)\}.\]
-%     Then, we will need to show \(∼\) is indeed a normalized bisimulation on \(⟨s⟩\) and \(⟨t⟩\), that is for all \(s' ∈ ⟨s⟩\) and \(t' ∈ ⟨t⟩\):
-%     \[s' ∼ t' ⟹ δ_S(s', α) \clos{∼} δ_T(t', α).\]
-%     We only need to show the above implication for \(s ∼ t\), as other element is already in some normalized bisimulation between sub-GKAT coalgebras of \(S\) and \(T\).
-%     And the implication \(s ∼ t ⟹ δ_S(s', α) \clos{∼} δ_T(t', α)\) can be proven by unfolding the definition of \(\clos{∼}\).
-% \end{proof}
-
-% \Cref{thm:inductive-construction} provides a way to inductively construct a normalized bisimulation.
-At this point, we can already write down a algorithm for normalized bisimulation; however, with more properties developed in the next section, we will be able to devise a more efficient algorithm.
-
-% \subsection{soundness and completeness}
-
-% In this section we will further investigate the properties of normalized bisimulation.
-Not only will we prove the soundness and completeness of normalized bisimulation between and we will also identify several special cases where the normalized bisimulation algorithm can be further optimized.
-
-% We can constraint both \(s\) and \(t\) to live states, and obtain a bisimulation on normalized coalgebra \(\norm(S)\) and \(\norm(T)\).
-% With a slight abuse of notation, we can define the bisimulation \(\norm(∼) ⊆ \norm(S) × \norm(T)\) as follows:
-% \begin{align*}
-%     \norm(∼) & ⊆ \norm(S) × \norm(T) \\
-%     \norm(∼) & ≜ \{(s, t) ∣ s ∼ t, s ∈ \norm(S), t ∈ \norm(T)\}
-% \end{align*}
-% and because \(∼\) is a normalized bisimulation, we have
-% \[s ∼ t ⟹ ∀ α ∈ \At, δ(s, α) \clos{∼} δ(t, α).\]
-% We proceed to define the transition \(\norm(δ_∼): \norm(∼) → G(\norm(∼))\) by case analysis on the elements of \(\clos{∼}\) (\Cref{def:norm-closure}):
-% \begin{itemize}
-%     \item If both \(δ(s, α)\) and \(δ(t, α)\) accepts, then \[\norm(δ_∼)((s,t), α) = \accept.\]
-%     \item If both \(δ(s, α)\) and \(δ(t, α)\) either reject or transition to dead states, then \[\norm(δ_∼)((s,t), α) = \reject.\]
-%     \item If both \(δ(s, α)\) and \(δ(t, α)\) transition to live states \(s, t\) respectively executing the same action \(p\), then \[\norm(δ_∼)((s,t), α) = ((s', t'), p).\]
-% \end{itemize}
-% Then the following theorem can be proven by computation.
-
-% \begin{theorem}\label{thm:norm-bisim-to-bisim-on-norm-coalg}
-%     For all normalized bisimulation \({∼} ⊆ S × T\), \(\norm(∼)\) is a bisimulation between \(\norm(S)\) and \(\norm(T)\). 
-%     This means that for all live states \(s ∈ S\) and \(t ∈ T\),
-%     \[s ∼ t ⟹ ⟦s⟧_S = ⟦t⟧_T\]
-% \end{theorem}
-
-% \Cref{thm:norm-bisim-to-bisim-on-norm-coalg} establishes the soundness for live state, and the soundness of normalized bisimulation on dead states are trivial, as all dead states have empty traces, hence are finite trace equivalent by definition.
-
-% The final case for soundness is that \(s ∼ t\), where one of \(s\) and \(t\) is dead, and the other is live. We will show that this case is impossible.
-
-% \begin{lemma}\label{thm:surj-on-norm-princ-coalg}
-%     Let \(T\) be a normal GKAT coalgebra, and \(S\) be any GKAT coalgebra,
-%     then a normalized bisimulation \({∼} ⊆ S × T\) will cover all the elements of \(⟨t⟩\) if it covers \(t\), for any \(t ∈ T\).
-%     Formally, 
-%     \[t ∈ π₂(∼) ⟹ ⟨t⟩ ⊆ π₂(∼)\]
-% \end{lemma}
-
-% \begin{proof}
-%     Because normalized bisimulation, we can obtain the following diagram:
-%     \[
-%         \begin{tikzcd}[column sep=1cm, row sep=1cm]
-%             S \ar[swap]{d}{δ_S}
-%                 & ∼ \ar{d}{(δ_S, δ_T)} \ar[swap]{l}{π₁} \ar{r}{π₂}
-%                 & T \ar{d}{δ_T} \\  
-%             G(S) 
-%                 & {\clos{∼}}^\At 
-%                     \ar[swap]{l}{{π₁}^\At} \ar{r}{{π₂}^\At}
-%                 & G(T)
-%         \end{tikzcd}
-%     \]
-%     We then perform the following epi-mono factorization on the right square:
-%     \[
-%         \begin{tikzcd}[column sep=1cm, row sep=1cm]
-%             ∼ \ar[swap]{d}{(δ_S, δ_T)} \ar[twoheadrightarrow]{r}{π₂}
-%                 & π₂(∼) \ar[hook]{r}{i} \ar{d}{δ_T}
-%                 & T \ar{d}{δ_T} \\  
-%             {\clos{∼}}^\At \ar[twoheadrightarrow]{r}{π₂^{\At}}
-%                 & G(π₂(∼)) \ar[hook]{r}{G(i)}
-%                 & G(T) 
-%         \end{tikzcd}
-%     \]
-%     The above diagram is indeed well-defined, specifically it suffice to show that \[∀ (r₁, r₂) ∈ {\clos{∼}}, π₂(r₁, r₂) = r₂ ∈ 2 + π₂(∼) × K.\] 
-%     The above inclusion can be proven by case analysis on the input in \(\clos{∼}\).
-Specifically, given that \(T\) is a normal GKAT coalgebra, thus elements in \(\clos{∼}\) can only be obtained by \nameref{rule:bisim-acc}, \nameref{rule:bisim-rej}, \nameref{rule:bisim-trans}, and \nameref{rule:norm-left} rule.
-%     \begin{itemize}
-%         \item If the input \((r₁, r₂)\) of \(π₂\) is obtained by \nameref{rule:bisim-acc}, then \[r₂ = \accept ∈ 2 + π₂(∼) × K.\]
-%         \item If the input \((r₁, r₂)\) of \(π₂\) is obtained by \nameref{rule:bisim-rej} or \nameref{rule:norm-left}, then \[r₂ = \reject ∈ 2 + π₂(∼) × K.\]
-%         \item If the input \((r₁, r₂)\) of \(π₂\) is obtained by \nameref{rule:bisim-trans}, then \(r₁\) and \(r₂\) can be expressed as \((s, p)\) and \((t, p)\) respectively, and \(s ∼ t\).
-Thus we have \(t ∈ π₂(∼)\) and \[r₂ = (t, p) ∈ 2 + π₂(∼) × K.\]
-%     \end{itemize}
-
-%     Then we can lift \(π₂: {\clos{∼}} → 2 + π₂(∼) × K\) over functor \((-)^{\At}\) to obtain the well-defined function \(π₂^{\At}: {\clos{∼}^{\At}} → G(π₂(∼))\), and \(δ_T: π₂(∼) → G(π₂(∼))\) is well-defined by the commutativity of the diagram.
-
-%     Finally, the above diagram tells us that \(π₂(∼)\) forms a sub-GKAT coalgebra of \(T\). Since \(⟨t⟩_T\) is the smallest sub-GKAT coalgebra tha contains \(t\), 
-%     \[t ∈ π₂(∼) ⟹ t ⊑ π₂(∼) ⟹ t ⊆ π₂(∼).\]
-% \end{proof}
-
-% \begin{lemma}\label{thm:norm-preserve-norm-bisim}
-%     Given a normalized bisimulation \({∼} ⊆ S × T\), the following restricted relation \(\norm ₗ(∼) ⊆ S × \norm(T)\) is also a normalized bisimulation:
-%     \[\norm ₗ(∼) = \{s, t ∣ s ∼ t, s ∈ S, t ∈ \norm(T)\}.\]
-% \end{lemma}
-
-% \begin{proof}
-%     Because \(∼\) is a normalized bisimulation, and \(\norm ₗ(∼) ⊆ {∼}\),
-%     \[(s, t) ∈ \norm ₗ(∼) ⟹ s ∼ t ⟹ ∀ α ∈ \At, δ_S(s, α) ∼ δ_T(t, α).\]
-%     And we can prove that 
-%     \[δ_S(s, α) ∼ δ_T(t, α) ⟹ (δ_S(s, α), \norm(δ_T)(t, α)) ∈ \norm ₗ(∼):\]
-%     \begin{itemize}
-%         \item If \(δ_T(t, α)\) transition to a dead state, then \(δ_S(s, α)\) will either reject or transition to a dead state, and \(\norm(δ_T)(t, α)\) will reject, thus: 
-%         \[(δ_S(s, α), \norm(δ_T)(t, α)) = (δ_S(s, α), \reject) ∈ \norm ₗ(∼).\]
-%         \item If \(δ_T(t', α)\) doesn't transition to a dead state, then \( \norm(δ_T)(t', α) = δ_T(t', α)\), and 
-%         \[(δ_S(s, α), \norm(δ_T)(t, α)) = (δ_S(s, α), δ_T(t, α)) ∈ \norm ₗ(∼).\]
-%     \end{itemize}
-%     Combining the two implications above, we have 
-%     \[(s, t) ∈ \norm ₗ(∼) ⟹ ∀ α ∈ \At, (δ_S(s, α), \norm(δ_T)(t, α)) ∈ \norm ₗ(∼).\]
-% \end{proof}
-
-% \begin{theorem}\label{thm:dead-live-don't-norm-bisim}
-%     Given a dead state \(s\) and a live state \(t\), there doesn't exists a normalized bisimulation \(∼\) s.t. \(s ∼ t\) or \(t ∼ s\).
-% \end{theorem}
-
-% \begin{proof}
-%     Because \({∼} ⊆ S × T\) is a normalized bisimulation, then we can construct a restricted normalized bisimulation \(∼_{s, t} ⊆ ⟨s⟩ × ⟨t⟩\)by~\Cref{thm:sub-coalg-preserve-norm-bisim}; then we can normalize \(⟨t⟩\) to obtain the normalized bisimulation \(\norm ₗ(∼_{s, t}) ⊆ ⟨s⟩ × \norm(⟨t⟩)\) by~\Cref{thm:norm-preserve-norm-bisim}.
-
-%     Since \(t\) is a live state, hence there exists an accepting state \(t'\) in \(\norm(⟨t⟩)\); because; \(s\) is a dead state, thus all the states in \(⟨s⟩\) is dead.
-
-%     Because \(t\) is live, thus \[(s, t) ∈ \norm ₗ(∼_{s, t}),\]
-%     and by~\Cref*{thm:surj-on-norm-princ-coalg}, every state in \(⟨t⟩\) is covered by \(\norm ₗ(∼_{s, t})\).
-%     In other word, there exists a there exists a state \(s' ∈ ⟨s⟩\) that is normally bisimilar to the accepting state \(s' ∼ t'\).
-%     However, recall that all states in \(⟨s⟩\) is dead (\Cref{thm:dead-iff-all-reachable-dead}) and a dead state cannot be normally bisimilar to a accepting state, since none of the transition (reject or transition to a dead state) of a dead state can be related to acceptance in \(\clos{\norm ₗ(∼_{s, t})}\).
-% \end{proof}
-
-% Thus we are able to show our soundness theorem:
-
-% \begin{theorem}[soundness]
-%     For all GKAT coalgebra \(S\) and \(T\), if there exists a  normalized bisimulation \({∼} ⊆ S × T\) s.t. \(s ∼ t\), then \(⟦s⟧_S = ⟦t⟧_T\).
-% \end{theorem}
-
-% \begin{proof}
-%     We prove this theorem by case analysis on the liveness of \(s\) and \(t\), we have indeed covered all the cases by previous theorems:
-%     \begin{itemize}
-%         \item if both \(s\) and \(t\) are live then the implication holds by~\Cref{thm:norm-bisim-to-bisim-on-norm-coalg}; 
-%         \item if one of \(s\) and \(t\) is live and the other is dead, then \(s ∼ t\) is necessarily false by~\cref{thm:dead-live-don't-norm-bisim}, thus the implication trivially holds;
-%         \item if both \(s\) and \(t\) are dead, then they are finite trace equivalent, thus the implication holds.
-%     \end{itemize}
-% \end{proof}
-
-% We then prove the completeness, which is the converse of the soundness statement, stating that the trace equivalence is ``completely'' characterized by normalized bisimulation. 
-% Formally, when two states are finite trace equivalent, then there will be a normalized bisimulation between them.
-
-% Our proof of completeness shares the core idea of as the corresponding completeness theorem~\cite{smolka_GuardedKleeneAlgebra_2020}, which is to construct a bisimulation from trace equivalence. 
-% However, we streamline our proof using the language of category theory, in the process, we were also able to prove a more general result, that is, homomorphism reflects normalized bisimulation:
-
-% \begin{theorem}[homomorphism reflects normalized bisimulation]\label{thm:hom-refl-norm-bisim}
-%     Given a pair of homomorphism \(h_S: S → S'\) and \(h_T: T → T'\), and a normalized bisimulation \({∼'} ⊆ S' × T'\), then there exists a normalized bisimulation \({∼} ⊆ S × T\), s.t.
-%     \[h_S(s) ∼' h_T(t) ⟺ s ∼ t.\]
-% \end{theorem}
-
-% \begin{proof}
-%     We can define \(∼\) as the pre-image of \(∼'\) under \(h_S × h_T\):
-%     \[{∼} ≜ \{(s, t) ∣ h_S(s) ∼' h_T(t)\}.\]
-%     Similar characterization also apply to their closure counterpart:
-%     \[{\clos{∼}} = \{(r₁, r₂) ∣ G(h_S)(r₁) \clos{∼'} G(h_S)(r₂)\}.\]
-%     The above equation can be verified by case analysis on the elements of \(\clos{∼'}\) and \(\clos{∼}\):
-%     \begin{itemize}
-%         \item by \nameref{rule:bisim-acc} and \nameref{rule:bisim-rej}, \((\accept, \accept)\) and \((\reject, \reject)\) are in both \(\clos{∼}\) and \(\clos{∼'}\).
-%         \item if the element is introduced by \nameref{rule:bisim-trans}, then 
-%         \[(h(s), p) \clos{∼'} (h(t), p) 
-%         ⟺ h(s) \clos{∼'} h(t) 
-%         ⟺ s \clos{∼} t 
-%         ⟺ (s, p) \clos{∼} (t, p).
-%         \]
-%         \item if the element is introduced by \nameref{rule:norm-left}, \nameref{rule:norm-right}, or \nameref{rule:norm-both}, we will recall that \(s\) is dead if and only if \(h(s)\) is dead (\Cref{thm:hom-preserve-liveness}).
-The reasoning is similar to previous case, we will show \nameref{rule:norm-left} case as an example:
-%         \[(h(s), p) \clos{∼'} \reject
-%         ⟺ h(s) \text{ is dead}
-%         ⟺ s \text{ is dead}
-%         ⟺ (s, p) \clos{∼} \reject.\]
-%     \end{itemize}
-%     by above equation about \(\clos{∼}\), 
-%     we will be able to verify that \(∼\) is indeed a normalized bisimulation, specifically, given any two states \(s ∼ t\):
-%     \begin{align*}
-%         & s ∼ t \\ 
-%         & ⟹ h_S(s) ∼' h_T(t) \\ 
-%         & ⟹ ∀ α, δ_{S'}(h_S(s), α) \clos{∼'} δ_{T'}(h_T(t), α) \\  
-%         & ⟹ ∀ α, (G(h_S) ∘ δ_S)(s, α) \clos{∼'} (G(h_T) ∘ δ_T)(t, α)
-%             & \text{\(h\) is homomorphism} \\
-%         & ⟹ ∀ α, δ_S(s, α) \clos{∼} δ_T(t, α) 
-%             & \text{above equation about \(\clos{∼}\)}
-%     \end{align*}
-% \end{proof}
-
-% \begin{lemma}
-%     Given any bisimulation \(S\), the identity relation \(=\) on \(S\) is a a normalized bisimulation.
-% \end{lemma}
-
-% \begin{proof}
-%     Because \(s = s' ⟹ ∀ α ∈ \At, δ(s, α) = δ(t, α)\), and the equality \(=\) on \(2 + S × K\) is a sub-relation of its normalized closure \(\clos{=}\), therefore \[s = s' ⟹ ∀ α ∈ \At, δ(s, α) \clos{=} δ(t, α)\]
-% \end{proof}
-
-% \begin{theorem}
-%     Finite trace equivalence \({∼} ≜ \{(s, t) ∣ ⟦s⟧ = ⟦t⟧\}\) is a normalized bisimulation.
-% \end{theorem}
-
-% \begin{proof}
-%     Take the identity relation \(=\) on the finite
-% \end{proof}
+\section{Implementation}
+
+\subsection{Optimization}\label{sec:optimization-implementation}
+
+\subsection{Performance}
+
+\section{Future Work}
+
+Can weak symbolic coalgebra leads to a simpler completeness proof.
 

### 2024-12-31 main.tex
finish all the proof and uses ieee format.

@@
 % Please compile this document using LuaLaTeX.
-% XeLaTeX and PDFLaTeX will not work.
+% because of the use of unicode-math
+% XeLaTeX and PDFLaTeX will result in error.
 
 \newif\iffull\fulltrue
-
-\documentclass{extarticle}
+\documentclass[conference]{IEEEtran}
+\IEEEoverridecommandlockouts
+% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
+%Template version as of 6/27/2024
+
+% \usepackage{cite}
+\usepackage{graphicx}
+\usepackage{textcomp}
+\usepackage{xcolor}
+\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
+    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
 
@@
 
-%% Title information
-\title{One-The-Fly Decision Procedures For GKAT}         
-% \title{Incorrectness Logic and Kleene Algebra With Top and Test}         
-                                        %% [Short Title] is optional;
-                                        %% when present, will be used in
-                                        %% header instead of Full Title.
-% \titlenote{}                            %% \titlenote is optional;
-                                        %% can be repeated if necessary;
-                                        %% contents suppressed with 'anonymous'
-%% \subtitle is optional
-
-% \subtitlenote{with subtitle note}       %% \subtitlenote is optional;
-%                                         %% can be repeated if necessary;
-%                                         %% contents suppressed with 'anonymous'
-
-%% Author information
-%% Contents and number of authors suppressed with 'anonymous'.
-%% Each author should be introduced by \author, followed by
-%% \authornote (optional), \orcid (optional), \affiliation, and
-%% \email.
-%% An author may have multiple affiliations and/or emails; repeat the
-%% appropriate command.
-%% Many elements are not rendered, but should be provided for metadata
-%% extraction tools.
-\author{
-    Cheng Zhang \and 
-    Qiancheng Fu \and
-    Hang Ji \and 
-    Ines Santacruz \and 
-    Marco Gaboardi
+\title{Symbolic On-the-fly Algorithms for GKAT Equivalences
+\thanks{Identify applicable funding agency here. If none, delete this.}
 }
-\date{}
-
-% In non-ACM format \maketitle needs to happen before abstract
+
+\author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
+\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+\textit{name of organization (of Aff.)}\\
+City, Country \\
+email address or ORCID}
+\and
+\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
+\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+\textit{name of organization (of Aff.)}\\
+City, Country \\
+email address or ORCID}
+\and
+\IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
+\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+\textit{name of organization (of Aff.)}\\
+City, Country \\
+email address or ORCID}
+\and
+\IEEEauthorblockN{4\textsuperscript{th} Given Name Surname}
+\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+\textit{name of organization (of Aff.)}\\
+City, Country \\
+email address or ORCID}
+\and
+\IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
+\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+\textit{name of organization (of Aff.)}\\
+City, Country \\
+email address or ORCID}
+\and
+\IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
+\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+\textit{name of organization (of Aff.)}\\
+City, Country \\
+email address or ORCID}
+}
+
 \maketitle
 
-%% Abstract
-%% Note: \begin{abstract}...\end{abstract} environment must come
-%% before \maketitle command
 \begin{abstract}
+This document is a model and instructions for \LaTeX.
+This and the IEEEtran.cls file define the components of your paper [title, text, heads, etc.]. *CRITICAL: Do Not Use Symbols, Special Characters, Footnotes, 
+or Math in Paper Title or Abstract.
 \end{abstract}
+
+\begin{IEEEkeywords}
+component, formatting, style, styling, insert.
+\end{IEEEkeywords}
 
@@
     Homomorphic image preserves principle sub-GKAT coalgebra. Specifically, given a homomorphism \(h: S → T\):
-    \[h(⟨S⟩_S) = ⟨h(s)⟩_T\]
+    \[h(⟨s⟩_S) = ⟨h(s)⟩_T\]
 \end{theorem}
@@
 
-\begin{algorithm}
-    \caption{On-the-fly bisimulation algorithm}\label{alg:bisim}
+\begin{algorithm*}
+    \caption{On-the-fly bisimulation algorithm}\label{alg:symb-bisim}
     \begin{algorithmic}
@@
     \end{algorithmic}
-\end{algorithm}
+\end{algorithm*}
 
@@
 
-% However, when considering designing a practical algorithm, we not just cares about the correctness of the derivative construction, we are also interested in the size of the generated automata. 
-% Specifically, we are interested in all the reachable states (or expressions) from any expression \(e\), which is exactly the principle sub-coalgebra \(⟨e⟩_{D̂}\).
-
-% \begin{theorem}[Complexity]
-%     The number of state in the principle sub-coalgebra \(⟨e⟩_{D̂}\) is linear to the number of primitive actions in \(e\).
-% \end{theorem}
-
-% \begin{proof}
-%     To prove this, we prove the following two theorem:
-
-% \end{proof}
+\Cref{thm:hom-thompson-derivative} have several consequences, one of the more obvious one is that we can use the functoriality of the lowering operation to show the semantic equivalence of the start state in the thompson's construction and the expression in derivative.
+
+\begin{corollary}[Correctness]
+    Given any expression \(e\) and its Thompson's coalgebra \(Ŝ_{e}\) with a start state \(s^* ∈ Ŝ_{e}\), then the semantics of the start state is equivalent to the semantics of \(e\): \(⟦s^*⟧_{Ŝ_{e}} = ⟦e⟧.\)
+\end{corollary}
+
+A not so obvious consequence of the homomorphism in~\Cref{thm:hom-thompson-derivative}, is the complexity of the algorithm based on derivatives.
+Our bisimulation algorithm (\Cref{alg:symb-bisim}) only explores the principle sub-coalgebra of the start state, i.e.
+\(s^*\) in the Thompson's construction \(Ŝ_{e}\) or \(e\) in the derivative \(D̂\); thus, deducing an upper bound on the size of the principle sub-coalgebras \(⟨s^*⟩_{Ŝ_{e}}\) and \(⟨e⟩_{D̂}\) are crucial to our complexity analysis.
+An upper bound on \(⟨s^*⟩_{Ŝ_{e}}\) is easy to obtain, as the size of \(Ŝ_{e}\), which subsumes the states of \(⟨s^*⟩_{Ŝ_{e}}\), is linear to the size of expression \(e\); therefore \(⟨s^*⟩_{Ŝ_{e}}\) is at most linear to the size of the expression \(e\).
+On the other hand the size of \(⟨e⟩_{D̂}\) can, again, be derived from the homomorphism in~\cref{thm:hom-thompson-derivative}.
+
+\begin{corollary}\label{thm:suj-hom-thompson-derivative}
+    There exists a surjective homomorphism \(h': ⟨s^*⟩_{Ŝ_{e}} → ⟨e⟩_{D̂}\). 
+    Because the size of \(⟨s^*⟩_{Ŝ_{e}}\) is linear to \(e\), the size of \(⟨e⟩_{D̂}\) is at most linear to the size of expression \(e\).
+\end{corollary}
+
+\begin{proof}
+    We define \(h'\) to be point-wise equal to \(h\), i.e. \(h'(s) ≜ h(s)\). 
+    Then we need to show that \(h'\) is well-defined ans surjective, which is a consequence of homomorphic image preserves principle sub-coalgebra (\Cref{thm:homo-img-preserve-principle-sub-coalg}): \(h(⟨s^*⟩_{Ŝ_{e}}) = ⟨h(s)⟩_{D̂} = ⟨e⟩_{D̂}.\)
+    In other words, the image of \(h\) on \(⟨s^*⟩_{Ŝ_{e}}\) is equal to \(⟨e⟩_{D̂}\); thus, because \(h'\) is point-wise equal to \(h\) and is defined on \(⟨s^*⟩_{Ŝ_{e}}\), the range of \(h'\) contains its codomain \(⟨e⟩_{D̂}\), showing that \(h'\) is surjective.
+\end{proof}
+
+An important consequence of~\cref{thm:suj-hom-thompson-derivative} is that \(⟨s^*⟩_{Ŝ_{e}}\) will have no less state than \(⟨e⟩_{D̂}\).
+However, it is important to notice that this does not mean running bisimulation algorithm on
 
@@
 
-\subsection{Performance}
+\subsection{Performance}\label{sec:performance-implementation}
 

### 2024-12-31 main.tex
pass on universal coalgebra

@@
 
-\author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
+\author{\IEEEauthorblockN{Cheng Zhang\textsuperscript{\textsection}}
+\IEEEauthorblockA{\textit{Department of Computer Science} \\
+\textit{University College London}\\
+London, United Kingdom \\
+0000-0002-8197-6181}
+\and
+\IEEEauthorblockN{Qiancheng Fu}
+\IEEEauthorblockA{\textit{Department of Computer Science} \\
+\textit{Boston University}\\
+Boston, USA \\
 email address or ORCID}
 \and
-\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
+\IEEEauthorblockN{Hang Ji}
+\IEEEauthorblockA{\textit{Department of Computer Science} \\
+\textit{Boston University}\\
+Boston, USA \\
 email address or ORCID}
 \and
-\IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
+\IEEEauthorblockN{Ines Santacruz}
+\IEEEauthorblockA{\textit{Department of Computer Science} \\
+\textit{Boston University}\\
+Boston, USA \\
 email address or ORCID}
 \and
-\IEEEauthorblockN{4\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address or ORCID}
-\and
-\IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address or ORCID}
-\and
-\IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
+\IEEEauthorblockN{Marco Gaboardi}
+\IEEEauthorblockA{\textit{Department of Computer Science} \\
+\textit{Boston University}\\
+Boston, USA \\
 email address or ORCID}
@@
 \maketitle
+\begingroup\renewcommand\thefootnote{\textsection}
+\footnotetext{Work largely performed at Boston University}
+\endgroup
 
@@
 \end{IEEEkeywords}
-
-TODO: change all the \(e_b\) to \(b\).
 
@@
 
-Our notion of normalized bisimulation is inherently coalgebraic, thus it is empirical for us to recall some notions and theorems in coalgebra.
-Given a functor \(F\) on the category of set and functions, a \emph{coalgebra over \(F\)} or \emph{\(F\)-coalgebra} consists of a set \(S\) and a function \(δ_S: S → F(S)\).
-We typically call \(F\) the \emph{signature} of the coalgebra, \(S\) the states of the coalgebra, and \(δ_S\) the transition function of the coalgebra.
+In this paper, we will make heavy use of coalgebraic theory, thus it is empirical for us to recall some notions and useful theorems in universal coalgebra.
+Given a functor \(F\) on the category of set and functions, a \emph{coalgebra over \(F\)} or \emph{\(F\)-coalgebra} consists of a set \(S\) and a function \(σ_S: S → F(S)\).
+We typically call elements in \(S\) the \emph{states} of the coalgebra, and \(σ_S(s)\) the \emph{dynamic} of state \(s\).
 We will sometimes use the states \(S\) to denote the coalgebra, when no ambiguity can arise.
 
-A homomorphism between two \(F\)-coalgebra \(S\) and \(T\) is a map \(h: S → T\) that preserves the transition function; diagrammatical, the following diagram commutes:
+A homomorphism between two \(F\)-coalgebra \(S\) and \(U\) is a map \(h: S → U\) that preserves the function \(σ\); diagrammatically, the following diagram commutes:
 \[
     \begin{tikzcd}
-        S \ar{r}{h} \ar[swap]{d}{δ_S} & T \ar{d}{δ_T} \\  
-        F(S) \ar{r}{F(h)} & F(T)
+        S \ar{r}{h} \ar[swap]{d}{σ_S} & U \ar{d}{σ_U} \\  
+        F(S) \ar{r}{F(h)} & F(U)
     \end{tikzcd}    
@@
     \begin{tikzcd}
-        S' \ar[hook]{r}{i} \ar[swap]{d}{δ_{S'}} & S \ar{d}{δ_S} \\  
+        S' \ar[hook]{r}{i} \ar[swap]{d}{σ_{S'}} & S \ar{d}{σ_S} \\  
         F(S') \ar[hook]{r}{F(i)} & F(S)
@@
 \]
-In fact, the transition function \(δ_{S'}\) is uniquely determined by the states \(S'\)~\cite[Proposition 6.1]{rutten_UniversalCoalgebraTheory_2000}.
+In fact, the function \(σ_{S'}\) is uniquely determined by the states \(S'\)~\cite[Proposition 6.1]{rutten_UniversalCoalgebraTheory_2000}.
 
@@
 \begin{lemma}[Theorem 6.3~\cite{rutten_UniversalCoalgebraTheory_2000}]\label{thm:hom-(pre)img-preserve-sub-coalg}
-    given a homomorphism \(h: S → T\), and sub-coalgebras \(S' ⊑ S\) and \(T' ⊑ T\), then 
-    \[h(S') ⊑ T \text{ and } h^{-1}(T') ⊑ S.\]
+    given a homomorphism \(h: S → U\), and sub-coalgebras \(S' ⊑ S\) and \(U' ⊑ U\), then 
+    \[h(S') ⊑ U \text{ and } h^{-1}(U') ⊑ S.\]
 \end{lemma}
 
-One particularly important sub-coalgebra of and coalgebra \(S\) is the least coalgebra generated by a single element \(s\), we will typically denote it as \(⟨s⟩_{S}\), and call it \emph{principle sub-coalgebra} generated by \(s\).
-We will omit the subscript \(S\) when it can be inferred from context or irrelevant.
-As we will see later, principle sub-coalgebra \(⟨s⟩_S\) is important because it models all the reachable state of \(s\) in \(S\).
-
-Principle sub-coalgebra always exists, because sub-coalgebra of any coalgebra forms a complete lattice~\cite[theorem 6.4]{rutten_UniversalCoalgebraTheory_2000}.
-And similar to sub-coalgebra, principle sub-coalgebra is also preserved under homomorphic image:
+One particularly important sub-coalgebra of and coalgebra \(S\) is the least coalgebra generated by a single element \(s\). 
+We will denote this sub-coalgebra as \(⟨s⟩_{S}\), and call it \emph{principle sub-coalgebra} generated by \(s\). 
+We sometimes omit the subscript \(S\) when it can be inferred from context or irrelevant.
+Intuitively, we usually think of principle sub-coalgebra \(⟨s⟩_S\) as the sub-coalgebra that is formed by all the ``reachable state'' form the state \(s\).
+This coalgebraic characterization of reachable state can allow us to avoid induction on the length of path from \(s\) to reach another state.
+
+For all coalgebra \(S\) and a state \(s ∈ S\), principle sub-coalgebra \(⟨s⟩_S\) always exists and is unique, because sub-coalgebra of any coalgebra forms a complete lattice~\cite[theorem 6.4]{rutten_UniversalCoalgebraTheory_2000}; thus taking the meet of all the sub-coalgebra that contains \(s\) will yield \(⟨s⟩_S\).
+
+Similar to sub-coalgebra, principle sub-coalgebra is also preserved under homomorphic image:
 \begin{theorem}\label{thm:homo-img-preserve-principle-sub-coalg}
-    Homomorphic image preserves principle sub-GKAT coalgebra. Specifically, given a homomorphism \(h: S → T\):
-    \[h(⟨s⟩_S) = ⟨h(s)⟩_T\]
+    Homomorphic image preserves principle sub-GKAT coalgebra. Specifically, given a homomorphism \(h: S → U\):
+    \[h(⟨s⟩_S) = ⟨h(s)⟩_U\]
 \end{theorem}
@@
     We will need to show that \(h(⟨s⟩_{S})\) is the smallest sub-GKAT coalgebra of \(S\) that contain \(h(s)\).
-First by definition of image, \(h(s) ∈ h(⟨s⟩_{S})\); second by \cref{thm:hom-(pre)img-preserve-sub-coalg}, \(h(⟨S⟩_S) ⊑ T\).
-
-    Finally, take any \(T' ⊑ T\) and \(h(s) ∈ T'\), recall that by \cref{thm:hom-(pre)img-preserve-sub-coalg}, \(h^{-1}(T') ⊑ S\). We can then derive that \(h(⟨s⟩_S) ⊑ T'\): 
+First by definition of image, \(h(s) ∈ h(⟨s⟩_{S})\); second by \cref{thm:hom-(pre)img-preserve-sub-coalg}, \(h(⟨S⟩_S) ⊑ U\).
+
+    Finally, take any \(U' ⊑ U\) and \(h(s) ∈ U'\), recall that by \cref{thm:hom-(pre)img-preserve-sub-coalg}, \(h^{-1}(U') ⊑ S\). We can then derive that \(h(⟨s⟩_S) ⊑ U'\): 
     \begin{align*}
-        h(s) ∈ T' 
-        & ⟹ s ∈ h^{-1}(T') \\  
-        & ⟹ ⟨s⟩_S ⊑ h^{-1}(T') & \text{definition of \(⟨s⟩_S\)}\\  
-        & ⟹ h(⟨s⟩_S) ⊑ T' & \text{take the image of \(h\) and \cref{thm:hom-(pre)img-preserve-sub-coalg}}
+        h(s) ∈ U' 
+        & ⟹ s ∈ h^{-1}(U') \\  
+        & ⟹ ⟨s⟩_S ⊑ h^{-1}(U') & \text{definition of \(⟨s⟩_S\)}\\  
+        & ⟹ h(⟨s⟩_S) ⊑ U' & \text{\cref{thm:hom-(pre)img-preserve-sub-coalg}}
     \end{align*}
-    Hence \(h(⟨s⟩_S)\) is the smallest sub-GKAT coalgebra of \(T\) that contains \(h(s)\).
+    Hence \(h(⟨s⟩_S)\) is the smallest sub-GKAT coalgebra of \(U\) that contains \(h(s)\).
 \end{proof}
@@
 
-Given two \(F\)-coalgebra \(S\) and \(T\), the \emph{behavioral equivalence} between states in \(S\) and \(T\) can be computed by a notion called \emph{bisimulation}.
-A relation \({∼} ⊆ S × T\) is called a \emph{bisimulation relation} if it forms a \(F\)-coalgebra: \[δ_{∼}: {∼} → F(∼),\] 
-And its projection functions \(π₁: S × T → S\) and \(π₂: S × T → T\) are both homomorphisms:
+Given two \(F\)-coalgebra \(S\) and \(U\), the \emph{behavioral equivalence} between states in \(S\) and \(U\) can be computed by a notion called \emph{bisimulation}.
+A relation \({∼} ⊆ S × U\) is called a \emph{bisimulation relation} if it forms an \(F\)-coalgebra: \[σ_{∼}: {∼} → F(∼),\] 
+And its projections \(π₁: S × U → S\) and \(π₂: S × U → U\) are both homomorphisms:
 \[
     \begin{tikzcd}
-        S \ar[swap]{d}{δ_S}
-            & ∼ \ar[swap]{l}{π₁} \ar{r}{π₂} \ar{d}{δ_{∼}} 
-            & T \ar{d}{δ_T}\\  
+        S \ar[swap]{d}{σ_S}
+            & ∼ \ar[swap]{l}{π₁} \ar{r}{π₂} \ar{d}{σ_{∼}} 
+            & U \ar{d}{σ_T}\\  
         F(S) & F(∼) \ar{l}{F(π₁)} \ar[swap]{r}{F(π₂)} & F(T)
@@
 \]
-As we will see later, bisimulation or variation of bisimulation indeed correspond to important semantical equivalences.
-
-\begin{theorem}
-    Given a functionr \(F\) on set, two \(F\)-colagebra \(S\) and \(T\), if there exists a surjective homomorphism \(h: S → T\), then \(S\) and \(T\) are bisimular.
-\end{theorem}
+In a special case, when there exists a homomorphism \(h: S → U\), then we can simply pick \({∼} ⊆ S × U\) to be \(\{(s, h(s)) ∣ s ∈ S\}\), which gives us a bisimulation with the lift \(σ_{∼} ≜ σ_S × σ_U\).
+
+\begin{corollary}
+    Given two \(F\)-coalgebra \(S, U\) and a homomorphism \(h: S → U\), then all for all \(s\), \(⟦s⟧_{S} = ⟦h(s)⟧_{U}.\)
+\end{corollary}
 
@@
     \begin{align*}
-        δ_S(s, α) = \accept 
-        & ⟺ \norm(δ_S)(s, α) = \accept 
-        ⟺ \norm(δ_∼)((s,t), α) = \accept \\ 
-        & ⟺ \norm(δ_T)(t, α) = \accept 
-        ⟺ δ_T(t, α) = \accept
+        & δ_S(s, α) = \accept \\
+        ⟺ & \norm(δ_S)(s, α) = \accept \\
+        ⟺ & \norm(δ_∼)((s,t), α) = \accept \\ 
+        ⟺ & \norm(δ_T)(t, α) = \accept \\
+        ⟺ & δ_T(t, α) = \accept
     \end{align*}
@@
 Finally because \(s ∼ t\), then
-    \[s' \text{ is live} ⟺ \norm(δ_S)(s, α) = (s', p) ⟺ \norm(δ_T)(t, α) = (t', p) ⟺ t' \text{ is live}.\]
+    \begin{align*}
+        s' \text{ is live} 
+        & ⟺ \norm(δ_S)(s, α) = (s', p) \\
+        & ⟺ \norm(δ_T)(t, α) = (t', p) \\
+        & ⟺ t' \text{ is live}.
+    \end{align*}
     \begin{itemize}
@@
     \begin{align*}
-        δ_S(s, α) \text{ rejects or transition to dead states} 
-    & ⟺ \norm(δ_S)(s, α) = \reject \\
-    & ⟺ \norm(δ_T)(t, α) = \reject \\
-    & ⟺ δ_T(t, α) \text{ rejects or transition to dead states}.
+        & δ_S(s, α) \text{ rejects or transition to dead states} \\
+        ⟺ & \norm(δ_S)(s, α) = \reject \\
+        ⟺ & \norm(δ_T)(t, α) = \reject \\
+        ⟺ & δ_T(t, α) \text{ rejects or transition to dead states}.
     \end{align*}

### 2025-01-01 main.tex
check point

@@
 %Template version as of 6/27/2024
+
+% page number
+\pagestyle{plain}
 
@@
 
-\section{Preliminary}
+\section{Background on Coalgebra and GKAT}
 
@@
 
-\subsection{Guarded Kleene Algebra With Tests}
-
-\subsection{GKAT Coalgebra}
-
-% definition
-GKAT coalgebra~\cite{smolka_GuardedKleeneAlgebra_2020,schmid_GuardedKleeneAlgebra_2021}, is a coalgebraic systems for GKAT.
-Specifically GKAT coalgebras over a alphabet \(K, B\) are coalgebras over the following functor:
+\subsection{GKAT and Its Coalgebra}
+
+Guarded Kleene Algebra with Tests, or GKAT~\cite{smolka_GuardedKleeneAlgebra_2020}, is a deterministic fragment of Kleene Algebra with Tests. 
+The syntax of GKAT over a set of primitive actions \(K\) and a set of primitive tests \(T\) can be defined in two sorts, boolean expressions \(\Bool\) and expressions \(\Exp\):
+\begin{align*}
+    a, b, c ∈ \Bool 
+        & ≜ 1 ∣ 0 ∣ t ∈ T ∣ b ∧ c ∣ b ∨ c ∣ \overline{b} \\  
+    e, f ∈ \Exp 
+        & ≜ p ∈ K ∣ b ∈ \Bool ∣ e +_b f ∣ e ; f ∣ e^{(b)} 
+\end{align*}
+where \(e +_b f\) is the if statement with condition \(b\); \(e;f\) is the sequencing of expression \(e\) and \(f\); finally \(e^{(b)}\) is the while loop with body \(e\) and condition \(b\).
+A GKAT expression can be unfolded into a KAT expression in the usual manner~\cite{kozen_KleeneAlgebraTests_1997c}:
+\begin{align*}
+    e +_b f & ≜ b; e + \overline{b}; f &
+    e^{(b)} & ≜ (b; e)^*; \overline{b}.
+\end{align*}
+Then the semantics of each expression \(⟦e⟧\) can be computed by the semantics of Kleene Algebra with tests.
+
+However, the semantics of GKAT can also be obtained coalgebraically, specifically, with GKAT coalgebra~\cite{smolka_GuardedKleeneAlgebra_2020,schmid_GuardedKleeneAlgebra_2021}.
+Formally, GKAT coalgebras over primitive actions \(K\) and primitive tests \(T\) are coalgebras over the following functor:
 \[G(S) ≜ (2 + S × K)^{\At_B},\] 
@@
 % TODO: I think this belongs in the intro
-As we will see later, the deterministic behavior of GKAT coalgebra not only enables a further optimized symbolic algorithm than KCT~\cite{pous_SymbolicAlgorithmsLanguage_2015}, but also present challenges.
-Specifically, GKAT coalgebra requires normalization to compute finite trace equivalences~\cite{smolka_GuardedKleeneAlgebra_2020}, where we will remove all the state that cannot lead to acceptance.
-It seems like we need to traverse the entire automaton to identify these ``dead states'', however we have shown that these dead state detection can be invoked lazily, only when discrepancy between bisimulation are found.
+As we will see later, the deterministic behavior of GKAT coalgebra not only enables a more versatile symbolic algorithm than KCT~\cite{pous_SymbolicAlgorithmsLanguage_2015}, but also present challenges.
+Specifically, GKAT coalgebra requires normalization to compute finite trace equivalences~\cite{smolka_GuardedKleeneAlgebra_2020}, where we will remove all the state that cannot lead to acceptance, which are called ``dead states''.
+
+In previous works, these dead states are detected after all the necessary state and transition is computed.
+This approach requires storing all the information of the coalgebra in memory, which do not allow on-the-fly computation, which can not only terminate whenever a counter-example is found, but can also erase past states from memory.
+
+However, to truly understand the on-the-fly algorithm, we will first need to define ``dead states'', and its role in defining the coalgebraic semantic of GKAT.
 
@@
 Traditionally, live and dead states are defined by whether they can reach an accepting state~\cite{smolka_GuardedKleeneAlgebra_2020}.
-However, recall that principle sub-coalgebra \(⟨s⟩_S\) models reachable states of \(s\) in coalgebra \(S\).
-Thus the classical definition is equivalent to the following:
+However, it is straightforward to show that principle sub-coalgebra \(⟨s⟩_S\) exactly correspondents to reachable states of \(s\) in coalgebra \(S\).
+Thus, the classical definition is equivalent to the following:
 \begin{definition}[liveness of states]\label{def:liveness-of-states}
@@
 \begin{theorem}[homomorphism perserves liveness]\label{thm:hom-preserve-liveness}
-    Given a homomorphism \(h: S → T\) and a state \(s ∈ S\):
+    Given a homomorphism \(h: S → U\) and a state \(s ∈ S\):
     \[\text{\(s\) is live} ⟺ \text{\(h(s)\) is live}\]
@@
 \begin{proof}
-    Because homomorphic image preserves principle sub-GKAT coalgebra (\cref{thm:homo-img-preserve-principle-sub-coalg})
-    \[h(⟨s⟩_S) = ⟨h(s)⟩_T;\]
+    Because homomorphic image preserves principle sub-GKAT coalgebra (\Cref{thm:homo-img-preserve-principle-sub-coalg})
+    \[h(⟨s⟩_S) = ⟨h(s)⟩_U;\]
     therefore for any state \(s' ∈ S\):
-    \[s' ∈ ⟨s⟩_S ⟺ h(s') ∈ h(⟨s⟩_S) ⟺ h(s') ∈ ⟨h(s)⟩_T.\]
-    And because \(s'\) is accepting if and only if \(h(s')\) accepting by definition of homomorphism; then \(⟨s⟩_S\) contains an accepting state if and only if \(h(⟨s⟩_S) = ⟨h(s)⟩_T\) contains an accepting state.
-    Therefore \(s\) is live in \(S\) if and only if \(h(s)\) is live in \(T\).
+    \[s' ∈ ⟨s⟩_S ⟺ h(s') ∈ h(⟨s⟩_S) ⟺ h(s') ∈ ⟨h(s)⟩_U.\]
+    And because \(s'\) is accepting if and only if \(h(s')\) accepting by definition of homomorphism; then \(⟨s⟩_S\) contains an accepting state if and only if \(h(⟨s⟩_S) = ⟨h(s)⟩_U\) contains an accepting state.
+    Therefore, \(s\) is live in \(S\) if and only if \(h(s)\) is live in \(U\).
 \end{proof}
 
+The above theorem then leads to several interesting liveness preservation properties for important structures on coalgebras, like sub-coalgebra and bisimulation.
+
 \begin{corollary}[sub-coalgebra perserves liveness]\label{thm:sub-coalg-preserve-liveness}
-    Given a sub-coalgebra \(S' ⊑ S\), then for all states \(s ∈ S'\),
-    \[\text{\(s\) is live in \(S'\)} ⟺ \text{\(s\) is live in \(S\)}.\]
+    For a sub-coalgebra \(S' ⊑ S\) and a state \(s ∈ S'\), \(s\) is live in \(S'\) if and only if \(s\) is live in \(S\).
 \end{corollary}
@@
 \begin{proof}
-    take the homomorphism \(h\) in \cref{thm:hom-preserve-liveness} to be the inclusion homomorphism \(i: S' → S\).
+    Let the homomorphism \(h\) in \cref{thm:hom-preserve-liveness} be the inclusion homomorphism \(i: S' → S\).
 \end{proof}
@@
 \begin{corollary}[bisimulation preserves liveness]\label{thm:bisim-preserve-liveness}
-    If there exists a bisimulation \(∼\) between GKAT coalgebra \(S\) and \(T\) s.t.
-\(s ∼ t\) for some states \(s ∈ S\) and \(t ∈ T\), then \(s\) and \(t\) has to be either both accepting, both live or both dead.
+    If there exists a bisimulation \(∼\) between GKAT coalgebra \(S\) and \(U\) s.t.
+\(s ∼ u\) for some states \(s ∈ S\) and \(u ∈ U\), then \(s\) and \(u\) has to be either both accepting, both live or both dead.
 \end{corollary}
@@
 \begin{proof}
-    Because for a \(∼\) is a bisimulation when both \(π₁: {∼} → S\) and \(π₂: {∼} → T\) are homomorphisms.
+    Because for a \(∼\) is a bisimulation when both \(π₁: {∼} → S\) and \(π₂: {∼} → U\) are homomorphisms.
     Therefore, 
@@
         s \text{ is live in } S 
-        & ⟺ π₁((s, t)) \text{ is live in } S 
-            & π₁((s, t)) = s\\  
-        & ⟺ (s, t) \text{ is live in } {∼} 
-            & \text{\(π₁\) is a homomorphism}\\  
-        & ⟺ π₂((s, t)) \text{ is live in } T 
-            & \text{\(π₂\) is a homomorphism} \\
-        & ⟺ t \text{ is live in } T 
-            & π₂((s, t)) = t
+        & ⟺ π₁((s, u)) \text{ is live in } S \\        
+        & ⟺ (s, u) \text{ is live in } {∼} \\  
+        & ⟺ π₂((s, u)) \text{ is live in } U  \\
+        & ⟺ u \text{ is live in } U. 
+        \qedhere
     \end{align*}
@@
 
-% The signature of GKAT coalgebra \(G\) is a simple polynomial functor~\cite[Definition 2.2.1]{jacobs_IntroductionCoalgebraMathematics_2016}, implying that it preserves the inclusion function:
-% \begin{lemma}%used in the comments below
-%     Given a inclusion \(i: S' → S\) for \(S' ⊆ S\), then \(G(i): G(S') → G(S)\) is also a inclusion in the sense that:
-%     \[∀ m ∈ G(S'), G(i)(m) = m.\]
-% \end{lemma}
-% This implies that for a sub-GKAT coalgebra \(S' ⊑ S\), then the restriction of transition function will satisfy the the diagram for sub-coalgebra.
-% \[
-%     \begin{tikzcd}
-%         S' \ar{d}{δ_S} \ar[hook]{r}{i} & S \ar{d}{δ_S} \\  
-%         G(S') \ar[hook]{r}{G(i)} & G(S)
-%     \end{tikzcd}
-% \]
-% Then by the uniqueness of transition for sub-coalgebra~\cite[Proposition 6.1]{rutten_UniversalCoalgebraTheory_2000}, the transition of \(S'\) is necessarily the restriction of \(δ_S\).
-A nice consequence of this observation is that the ordering between sub-GKAT coalgebra is exactly the subset ordering.
-% \begin{lemma}[ordering in sub-GKAT coalgebra]\label{thm:sub-GKAT coalgebra-is-subset}
-%     Given two sub-GKAT coalgebra \(S₁ ⊑ S\) and \(S₂ ⊑ S\), then 
-%     \[S₁ ⊑ S₂ ⟺ S₁ ⊆ S₂.\]
-% \end{lemma}
-
-% \begin{proof}
-%     \(⟹\) direction is true by definition, and \(⟸\) direction is true because both \(S₁\) and \(S₂\) are sub-GKAT coalgebra, thus their transition functions are both restrictions of \(δ_S\).
-Hence the commutativity of the following diagram can be verified by computation:
-%     \[
-%     \begin{tikzcd}
-%         S₁ \ar[hook]{r}{i} \ar{d}{δ_S} & S₂ \ar[hook]{r}{i} \ar{d}{δ_S} & S \ar{d}{δ} \\  
-%         G(S₁) \ar[hook]{r}{G(i)} & G(S₂) \ar[hook]{r}{G(i)} & G(S)
-%     \end{tikzcd}
-% \]
-% \end{proof}
-
 % trace semantics
-\subsection{Normalization And Trace Semantics}
-
-(Possibly infinite) trace model \(𝒢_ω\) is the natural semantics of states in a GKAT coalgebra, specifically it forms the final coalgebra of GKAT coalgebras~\cite{schmid_GuardedKleeneAlgebra_2021}.
-The finality of the model means that every state in any GKAT coalgebra \(S\) can be assigned a semantics under the unique homomorphism \(⟦-⟧^{ω}_{S}: S → 𝒢_ω\); and such semantical equivalences can indeed be identified by bisimulation~\cite{schmid_GuardedKleeneAlgebra_2021}:
-\[⟦s⟧^{ω}_{S} = ⟦t⟧^{ω}_{T} ⟺ \text{exists a bisimulation \({∼} ⊆ S × T\), s.t. \(s ∼ t\)}.\]
-
-The infinite trace equivalences is relatively easy to compute efficiently, as bisimulation is, in general, compatible with derivative based computation on-the-fly algorithm~\cite{kozen_CoalgebraicTheoryKleene_2017,almeida_DecidingKATHoare_2012,pous_SymbolicAlgorithmsLanguage_2015}.
-However, the \emph{finite} trace model \(𝒢\) is only the final coalgebra of GKAT coalgebras without dead states, which we call \emph{normal GKAT coalgebra}~\cite{smolka_GuardedKleeneAlgebra_2020}.
-Fortunately every GKAT coalgebra can be normalized by rerouting all the transition from dead states to rejection
-\begin{align*}
-    \norm(δ_S) & : S → G(S) \\
-    \norm(δ_S)(s, α) & ≜ \begin{cases}
-        \reject & \text{if } δ_S(s, α) = (s', p) \text{ and \(s'\) is dead}\\
-        δ_S(s, α) & \text{otherwise}
-    \end{cases}
-\end{align*}
+\subsection{Normalization And Semantics}
+
+(Possibly infinite) trace model \(𝒢_ω\) is the final coalgebra of GKAT coalgebras~\cite{schmid_GuardedKleeneAlgebra_2021}.
+The finality of the model implies that every state in any GKAT coalgebra \(S\) can be assigned a semantics under the unique homomorphism \(⟦-⟧^{ω}_{S}: S → 𝒢_ω\); and such semantic equivalences can indeed be identified by bisimulation~\cite{schmid_GuardedKleeneAlgebra_2021}:
+\(⟦s⟧^{ω}_{S} = ⟦t⟧^{ω}_{T}\) if and only if there exists a bisimulation \({∼} ⊆ S × T\), s.t. \(s ∼ t\).
+
+The infinite trace equivalences can be directly computed with bisimulation on derivative, which supports on-the-fly algorithm as demonstrated by similar systems~\cite{kozen_CoalgebraicTheoryKleene_2017,almeida_DecidingKATHoare_2012,pous_SymbolicAlgorithmsLanguage_2015}.
+However, the \emph{finite} trace model \(𝒢\) is the final coalgebra of GKAT coalgebras without dead states, which we call \emph{normal GKAT coalgebra}~\cite{smolka_GuardedKleeneAlgebra_2020}. 
+Fortunately every GKAT coalgebra can be normalized by rerouting all the transition from dead states to rejection.
+Concretely, \(\norm(δ_S) : S → G(S)\) is defined as \(\norm(δ_S)(s, α) ≜ \reject\) when \(δ_S(s, α) = (s', p)\) and \(s'\) is dead in \(S\); and \(\norm(δ_S)(s, α) ≜ δ_S(s, α)\) otherwise.
 We denote the normalized coalgebra \((S, \norm(δ_S))\) as \(\norm(S)\).
-This means that the finite trace semantics \(⟦-⟧\) is the unique coalgebra homomorphism \(\norm(S) → \norm(𝒢)\).
-Hence the finite trace equivalence between \(s ∈ S\) and \(t ∈ T\) can be computed by first normalizing \(S\) and \(T\), then decide whether there is a bisimulation on \(\norm(S)\) and \(\norm(T)\) that includes \((s, t)\).
-For a more intuitive account for the trace semantics, we refer the reader to the work of \Citeauthor{smolka_GuardedKleeneAlgebra_2020}~\cite{smolka_GuardedKleeneAlgebra_2020}.
-
-Then by the finality of GKAT colagebra, the soundness and completeness of bisimulation is a simple corollary in universal coalgebra~\cite{smolka_GuardedKleeneAlgebra_2020,jacobs_IntroductionCoalgebraMathematics_2016,rutten_UniversalCoalgebraTheory_2000}.
+
+The finality of \(𝒢\) means that the finite trace semantics \(⟦-⟧\) is the unique coalgebra homomorphism \(\norm(S) → 𝒢\). 
+Furthermore, the finite trace equivalence between \(s ∈ S\) and \(u ∈ U\) can be computed by first normalizing \(S\) and \(U\), then decide whether there is a bisimulation on \(\norm(S)\) and \(\norm(U)\) that includes \((s, t)\).
+For more details on the finite trace semantics, we refer the reader to the work of \Citeauthor{smolka_GuardedKleeneAlgebra_2020}~\cite{smolka_GuardedKleeneAlgebra_2020}.
+
+% Then by the finality of GKAT colagebra, the soundness and completeness of bisimulation is a simple corollary in universal coalgebra~\cite{smolka_GuardedKleeneAlgebra_2020,jacobs_IntroductionCoalgebraMathematics_2016,rutten_UniversalCoalgebraTheory_2000}.
 No matter through explicit construction~\cite{smolka_GuardedKleeneAlgebra_2020} or (weak) pullback~\cite{jacobs_IntroductionCoalgebraMathematics_2016,rutten_UniversalCoalgebraTheory_2000}, both completeness proof shows that the language equivalence: \[{≡} ≜ \{(s,t) ∣ ⟦s⟧_S = ⟦t⟧_T\}\] is indeed a bisimulation.
-This result, together with soundness result \(s ∼ t ⟹ ⟦s⟧_S = ⟦t⟧_T\) means that the language equivalence \(≡\) is indeed the largest bisimulation, which allows us to work with bisimulation equivalence instead of bisimulation.
+% This result, together with soundness result \(s ∼ t ⟹ ⟦s⟧_S = ⟦t⟧_T\) means that the language equivalence \(≡\) is indeed the largest bisimulation, which allows us to work with bisimulation equivalence instead of bisimulation.
 

### 2025-01-02 main.tex
checkpoint: recursive construction

@@
 \usepackage[margin=2.5cm]{geometry}
+
+% allow page break in align environment
+\allowdisplaybreaks
 
@@
 
+Finally, we are interested in bisimulation equivalence.
+Given a coalgebra \(S\), a bisimulation \({≃} ⊆ S × S\) is a bisimulation equivalence when \(≃\) is a bisimulation and also an equivalence relation.
+As it turns out, searching for bisimulation is equivalent to bisimulation equivalence. 
+This result is very important, as it allows us to leverage more efficient data structure, like union-find, to search for equivalence relation instead of relations in general.
+
+\begin{theorem}\label{thm:bisim-iff-bisim-eq}
+    For any coalgebra \(S\), there exists a bisimulation \({∼} ⊆ S × S\) s.t. \(s₁ ∼ s₂\) if and only if there exists a bisimulation equivalence \({≃} ⊆ S × S\) s.t. \(s₁ ≃ s₂\)
+\end{theorem}
+
+\begin{proof}
+    First show the \(⟹\) direction, we consider the maximal bisimulation between \(S\) and itself \({≡} ⊆ S × S\), which is known to be a bisimulation equivalence~\cite[Corollary 5.6]{rutten_UniversalCoalgebraTheory_2000}.
+    Because \(≡\) is maximal, therefore \({∼} ⊆ {≡}\) and \(s₁ ≡ s₂\).
+
+    Then the \(⟸\) direction can be proven by setting \({∼} ≜ {≃}\).
+\end{proof}
+
 \subsection{GKAT and Its Coalgebra}
@@
 \end{align*}
-Then the semantics of each expression \(⟦e⟧\) can be computed by the semantics of Kleene Algebra with tests.
-
-However, the semantics of GKAT can also be obtained coalgebraically, specifically, with GKAT coalgebra~\cite{smolka_GuardedKleeneAlgebra_2020,schmid_GuardedKleeneAlgebra_2021}.
+Then the semantics of each expression \(⟦e⟧\) can be computed by the semantics of Kleene Algebra with tests~\cite{kozen_KleeneAlgebraTests_1997c}.
+An important construct in the semantics is \emph{atoms}, which are a string of all the primitive tests either in its positive or negative form: for \(B ≜ \{b₁, b₂, …, bₙ\}\)
+\[\At_B ≜ \{b₁' ⋅ b₂' ⋅ ⋯ ⋅ bₙ' ∣ bᵢ' ∈ \{bᵢ, \overline{bᵢ}\}\}.\]
+Follow the conventional notation, we denote an atom using \(α, β\); and we sometimes omit the subscript \(B\) when no confusing can arise.
+Intuitively, atoms are truth assignments to each primitive tests, indicating which primitive tests is satisfied in the current program states.
+
+For the sake of brevity, we omit the complete definition of GKAT and KAT semantics, we refer the reader to previous works~\cite{smolka_GuardedKleeneAlgebra_2020,schmid_GuardedKleeneAlgebra_2021,kozen_KleeneAlgebraTests_1997c}, which explains these semantics in detail.
+Our work avoids directly interact with the semantics by leveraging results in the coalgebraic theory of GKAT, which we will recap below.
+
 Formally, GKAT coalgebras over primitive actions \(K\) and primitive tests \(T\) are coalgebras over the following functor:
-\[G(S) ≜ (2 + S × K)^{\At_B},\] 
-where \(2 ≜ \{\accept, \reject\}\).
-Intuitively given a state \(s ∈ S\) and an atom \(α ∈ \At\), \(δ(s, α)\) will deterministically execute one of the following: reject \(α\), denoted as \(δ(s, α) = \reject\); accept \(α\), denoted \(δ(s, α) = \accept\); or transition to a state \(s' ∈ S\) and execute action \(p ∈ K\), denoted as \(δ(s, α) = (s', p)\).
+\[G(S) ≜ (\{\accept, \reject\} + S × K)^{\At_B}.\] 
+Intuitively, given a state \(s ∈ S\) and an atom \(α ∈ \At\), \(δ(s, α)\) will deterministically execute one of the following: reject \(α\), denoted as \(δ(s, α) = \reject\); accept \(α\), denoted \(δ(s, α) = \accept\); or transition to a state \(s' ∈ S\) and execute action \(p ∈ K\), denoted as \(δ(s, α) = (s', p)\).
 
@@
 Fortunately every GKAT coalgebra can be normalized by rerouting all the transition from dead states to rejection.
-Concretely, \(\norm(δ_S) : S → G(S)\) is defined as \(\norm(δ_S)(s, α) ≜ \reject\) when \(δ_S(s, α) = (s', p)\) and \(s'\) is dead in \(S\); and \(\norm(δ_S)(s, α) ≜ δ_S(s, α)\) otherwise.
-We denote the normalized coalgebra \((S, \norm(δ_S))\) as \(\norm(S)\).
+Concretely, given a GKAT coalgebra \(S ≜ (S, δ_S)\), we let \(δ_{\norm(S)} : S → G(S)\) is defined as \(δ_{\norm(S)}(s, α) ≜ \reject\) when \(δ_S(s, α) = (s', p)\) and \(s'\) is dead in \(S\); and \(δ_{\norm(S)}(s, α) ≜ δ_S(s, α)\) otherwise; then \((S, δ_{\norm(S)})\) is the \emph{normalized} coalgebra of \(S ≜ (S, δ_S)\) denoted as \(\norm(S)\).
 
@@
 Furthermore, the finite trace equivalence between \(s ∈ S\) and \(u ∈ U\) can be computed by first normalizing \(S\) and \(U\), then decide whether there is a bisimulation on \(\norm(S)\) and \(\norm(U)\) that includes \((s, t)\).
-For more details on the finite trace semantics, we refer the reader to the work of \Citeauthor{smolka_GuardedKleeneAlgebra_2020}~\cite{smolka_GuardedKleeneAlgebra_2020}.
-
-% Then by the finality of GKAT colagebra, the soundness and completeness of bisimulation is a simple corollary in universal coalgebra~\cite{smolka_GuardedKleeneAlgebra_2020,jacobs_IntroductionCoalgebraMathematics_2016,rutten_UniversalCoalgebraTheory_2000}.
-No matter through explicit construction~\cite{smolka_GuardedKleeneAlgebra_2020} or (weak) pullback~\cite{jacobs_IntroductionCoalgebraMathematics_2016,rutten_UniversalCoalgebraTheory_2000}, both completeness proof shows that the language equivalence: \[{≡} ≜ \{(s,t) ∣ ⟦s⟧_S = ⟦t⟧_T\}\] is indeed a bisimulation.
-% This result, together with soundness result \(s ∼ t ⟹ ⟦s⟧_S = ⟦t⟧_T\) means that the language equivalence \(≡\) is indeed the largest bisimulation, which allows us to work with bisimulation equivalence instead of bisimulation.
-
-\begin{definition}
-    A \emph{bisimulation equivalence} in \(S\) is a bisimulation between \(S\) and itself, and it is also an equivalence relation.
-\end{definition}
-
-\begin{theorem}\label{thm:bisim-iff-bisim-equiv}
-    Given two states \(s, t ∈ S\), then there exists a bisimulation \(∼\) s.t. \(s ∼ t\) if and only if there exists a bisimulation equivalence \(≃\) s.t. \(s ≃ t\).
+For a more detailed explanation on the finite trace semantics, we refer the reader to the work of \Citeauthor{smolka_GuardedKleeneAlgebra_2020}~\cite{smolka_GuardedKleeneAlgebra_2020}.
+
+Besides giving us the finite-trace semantics, the normalization operation also connects the finite and infinite trace semantics, because it is an endofunctor on the category of GKAT coalgebra.
+
+\begin{theorem}\label{thm:norm-functor}
+    \(\norm\) is an endofunctor in the category GKAT coalgebra.
+    More specifically, if \(h: S → U\) is a GKAT homomorphism, then \(h: \norm(S) → \norm(U)\) is also a homomorphism.
 \end{theorem}
@@
 \begin{proof}
-    The \(⟹\) direction can just take \(≃\) to be the language equivalence \(≡\), which is a bisimulation equivalence, and because \(≡\) is maximal, therefore \({∼} ⊑ {≡}\), and \((s, t) ∈ {∼} ⊆ {≡}\).
-
-    The \(⟸\) direction is true because all bisimulation is a bisimulation equivalence, thus we can take \(∼\) to just be the given bisimulation equivalence \(≃\).
-\end{proof}
-
-% Although \(\norm(S)\) might not be a homomorphic image of \(S\), normalization does satisfy many nice properties of homomorphic images.
-For example, it is monotonic and it preserves principle sub-GKAT coalgebra, like in \cref{thm:homo-img-preserve-principle-sub-coalg}
-% \begin{lemma}[monotonicity]\label{thm:monotonicity-norm}
-%     For all GKAT coalgebra \(S\) and \(T\):
-%     \[S ⊑ T ⟹ \norm(S) ⊑ \norm(T).\]
-% \end{lemma}
-
-% \begin{theorem}[\(\norm\) preserves principle sub coalgebra]\label{thm:norm-preserve-principle-sub-coalg}
-%     Given a GKAT coalgebra \(S\) and a live state \(s ∈ S\):
-%     \[⟨s⟩_{\norm(S)} = \norm(⟨s⟩_{S}).\]
-% \end{theorem}
-
-% \begin{proof}
-%     We need to show \(\norm(⟨s⟩_{S})\) is the smallest sub-GKAT coalgebra in \(\norm(S)\) that contains \(s\).
-
-%     First, \(\norm(⟨s⟩_{S})\) contains \(s\) because \(s\) is live. 
-%     Second, take any \(T ⊑ \norm(S)\), we will use \(⟨T⟩_S\) to denote the smallest sub-GKAT coalgebra in \(S\) that contains all the states in \(T\). 
-%     Since \(T\) is a sub-GKAT coalgebra of \(\norm(S)\), \(⟨T⟩_S ∖ T\) can only contain dead states, thus \(\norm(⟨T⟩_S) = T\).
-%     \begin{align*}
-%         & ⟨s⟩_{S} ⊑ ⟨T⟩_S & \text{\(⟨T⟩_S\) contains \(s\)} \\  
-%         ⟹ {}& \norm(⟨s⟩_{S}) ⊑ \norm(⟨T⟩_S) & \text{monotonicity: \cref{thm:monotonicity-norm}} \\  
-%         ⟹ {}& \norm(⟨s⟩_{S}) ⊑ T & \norm(⟨T⟩_S) = T
-%     \end{align*}
-% \end{proof}
+    Recall that \(h\) is a homomorphism if and only if for all \(s ∈ S\) and \(α ∈ \At\):
+    \begin{itemize}[nosep]
+        \item for a result \(r ∈ \{\reject, \accept\}\), 
+        \[δ_S(s, α) = r ⟺ δ_U(h(s), α) = r;\]
+        \item for any \(s' ∈ S\) and \(p ∈ K\), 
+        \[δ_S(s, α) = (s', p) ⟺ δ_{U}(h(s), α) = (h(s'), p).\]
+    \end{itemize}
+
+    Then we show that \(h: \norm(S) → \norm(U)\) is a homomorphism, this is a consequence of homomorphism preserves liveness (\Cref{thm:hom-preserve-liveness}): for all \(s ∈ \norm(S)\) and \(α ∈ \At\):
+    \begin{align*}
+        & δ_{\norm(S)}(s, α) = \accept \\*
+        ⟺{}& δ_S(s, α) = \accept \\*  
+        ⟺{}& δ_U(h(s), α) = \accept \\*  
+        ⟺{}& δ_{\norm(U)}(h(s), α) = \accept;\\
+        & δ_{\norm(S)}(s, α) = \reject \\*  
+        ⟺{}& δ_{S}(s, α) = \reject 
+        \text{ or } δ_{S}(s, α) = (s', p), s' \text{ is dead} \\*
+        ⟺{}& δ_{U}(h(s), α) = \reject \\*
+        & \text{ or } δ_{U}(h(s), α) = (h(s'), p), h(s') \text{ is dead} \\*  
+        ⟺{}& δ_{\norm(U)}(h(s), α) = \reject;\\
+        & δ_{\norm(S)}(s, α) = (s', p) \\*
+        ⟺{}& δ_{S}(s, α) = (s', p), s' \text{ is live} \\*  
+        ⟺{}& δ_{U}(h(s), α) = (h(s'), p), h(s') \text{ is live}\\* 
+        ⟺{}& δ_{\norm(U)}(h(s), α) = (h(s'), p).
+        \qedhere
+    \end{align*}
+\end{proof}
+
+\begin{corollary}\label{thm:norm-sub-coalg}
+    Normalization preserves sub-coalgebra, i.e. if \(S' ⊑ S\) then \(\norm(S') ⊑ \norm(S)\).
+\end{corollary}
+
+\begin{proof}
+    By letting the homomorphism in~\Cref{thm:norm-functor} to be the inclusion homomorphism \(i: S' → S\)
+\end{proof}
+
+Because of the functoriality, we can show that two states are infinite-trace equivalent implies these two states are finite-trance equivalent.
+This gives us more tool in proving semantic equivalence between two states in GKAT coalgebras: proving bisimulation in the GKAT coalgebra is already enough to obtain the semantic equivalence for two states.
+
+\begin{corollary}\label{thm:inf-trace-equiv-implies-fin-trace-equiv}
+    Given two states in two GKAT coalgebra \(s ∈ S, u ∈ U\), \(⟦s⟧^{ω}_{S} = ⟦u⟧^{ω}_{U} ⟹ ⟦s⟧_{S} = ⟦u⟧_{U}\).
+\end{corollary}
+
+\begin{proof}
+    Because \(⟦s⟧^{ω}_{S} = ⟦u⟧^{ω}_{U}\), there exists a bisimulation \({∼} ⊆ S × U\) s.t. \(s ∼ u\)~\cite{schmid_GuardedKleeneAlgebra_2021}.
+    Therefore, we have the following span in the category of GKAT coalgebra:
+    \[\begin{tikzcd}
+        S & ∼ \ar{r}{π₂} \ar[swap]{l}{π₁} & U
+    \end{tikzcd}\]
+    Then by~\Cref{thm:norm-functor}, \(\norm(∼)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\):
+    \[\begin{tikzcd}
+        \norm(S) 
+        & \norm(∼) \ar{r}{π₂} \ar[swap]{l}{π₁} 
+        & \norm(U)
+    \end{tikzcd}\]
+    And because normalization operation preserves states in \(∼\), therefore \((s, u) ∈ \norm(∼)\), and because \(\norm(∼)\) is a bisimulation between the normalization of \(S\) and \(U\), therefore \(⟦s⟧_{S} = ⟦u⟧_{U}\)
+\end{proof}
 
@@
 
-\begin{theorem}[sub-coalgebra perserve bisimulation]\label{thm:sub-coalg-preserve-bisim}
+TODO: I think we should move the next couple theorem to the background.
+
+\begin{theorem}[sub-coalgebras perserve and reflect bisimulation]\label{thm:sub-coalg-preserve-bisim}
     Given any sub-coalgebra \(S' ⊑ S\) and \(T' ⊑ T\),
@@
 
+\Cref{thm:sub-coalg-preserve-bisim} allows us to only search for bisimulation on a sub-coalgebra, speeding up our search algorithm.
+Another way to speed up the algorithm is to use efficient data structures to find a bisimulation equivalence instead of a bisimulation relation.
+This optimization is a special case of the \emph{up-to technique}~\cite{bonchi_GeneralAccountCoinduction_2017a}.
+Specifically, we will extend~\Cref{thm:bisim-iff-bisim-eq} to a setting where the bisimulation is no longer over the same coalgebra.
+
+\begin{theorem}
+    Given two states in two coalgebras \(s ∈ S, u ∈ U\), let \(S + U\) be the coproduct coalgebra of \(S\) and \(U\)~\cite{rutten_UniversalCoalgebraTheory_2000}. 
+    There exists a bisimulation \({∼} ⊆ S × U\) s.t. \(s ∼ u\) if and only if there exists a bisimulation equivalence on \({≃} ⊆ (S + U) × (S + U)\), s.t. \(s ≃ u\).
+\end{theorem}
+
+\begin{proof}
+    Notice that both \(S\) and \(U\) are sub-coalgebra of \(S + U\) witnessed by the canonical injection \(\mathrm{inj}ₗ: S → S + U\) and \(\mathrm{inj}ᵣ: U → S + U\).
+
+    Let \(∼\) be a bisimulation and \(≃\) be a bisimulation equivalence:
+    \begin{align*}
+        & ∃~ {∼} ⊆ S × U, s ∼ u \\
+        ⟺{}& ∃~ {∼} ⊆ (S+U) × (S+U), s ∼ u \\  
+        ⟺{}& ∃~ {≃} ⊆ (S+U) × (S+U), s ≃ u.
+        \qedhere
+    \end{align*}
+\end{proof}
+
+After we justify the above optimizations of the algorithm, we will show the core theorem that establishes the correctness of our algorithm.
+
 \begin{lemma}[bisimulation between dead states]\label{thm:bisim-between-dead}
-    Given two dead states \(s ∈ S\) and \(t ∈ T\), then the singleton bisimulation 
+    Given two dead states \(s ∈ S\) and \(u ∈ U\), then the singleton bisimulation \({∼} ⊆ \norm(S) × \norm(U)\):
     \begin{mathpar}
-        {∼} ≜ \{(s,t)\} \and 
-        δ_{∼}((s,t), α) ≜ \reject
+        {∼} ≜ \{(s,u)\} \and 
+        δ_{∼}((s,u), α) ≜ \reject
     \end{mathpar}
-    is a bisimulation between \(S\) and \(T\).
+    is a bisimulation between \(\norm(S)\) and \(\norm(U)\).
 \end{lemma}
@@
 
-\begin{theorem}[inductive construction]\label{thm:inductive-construction}
-    Given two GKAT coalgebra \(S\) and \(T\), and two of their elements \(s ∈ S\) and \(t ∈ T\),
-    there exists a bisimulation \({∼} ⊆ ⟨s⟩ × ⟨t⟩\) s.t. \(s ∼ t\), if and only if all of the following holds:
+\begin{theorem}[Recursive Construction]\label{thm:inductive-construction}
+    Given two GKAT coalgebra \(S\) and \(U\), and two of their elements \(s ∈ S\) and \(u ∈ U\),
+    there exists a bisimulation \({∼} ⊆ \norm(⟨s⟩) × \norm(⟨u⟩)\) s.t. \(s ∼ u\), if and only if all the following holds:
     \begin{enumerate}
-        \item\label{itm:acc-condition} for all \(α ∈ \At\), \(δ_{S}(s, α) = \accept ⟺ δ_{T}(t, α) = \accept\);
-        \item\label{itm:transition-bisim} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{T}(t, α) = (t', p)\), then there exists a bisimulation \({∼_{s',t'}}\) on \(⟨s'⟩\) and \(⟨t'⟩\), s.t.
-\(s' ∼_{s',t'} t'\);
-        \item\label{itm:transition-dead} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{T}(t, α) = (t', q)\), s.t. \(p ≠ q\), then both \(s'\) and \(t'\) are dead;
-        \item\label{itm:rej-or-dead} \(s\) reject \(α\) or transition to a dead state via \(α\) if and only if \(t\) rejects \(α\) or transition to a dead state via \(α\).
+        \item\label{itm:acc-condition} for all \(α ∈ \At\), \(δ_{S}(s, α) = \accept ⟺ δ_{U}(u, α) = \accept\);
+        \item\label{itm:rej-or-dead} \(s\) reject \(α\) or transition to a dead state via \(α\) if and only if \(u\) rejects \(α\) or transition to a dead state via \(α\);  
+        \item\label{itm:transition-bisim} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', p)\), then there exists a bisimulation \({∼'} ⊆ \norm(⟨s'⟩) × \norm(⟨u'⟩)\), s.t. \(s' ∼' u'\);
+        \item\label{itm:transition-dead} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', q)\), s.t. \(p ≠ q\), then both \(s'\) and \(t'\) are dead.
     \end{enumerate}
@@
 \begin{proof}
-    We first prove \(⟹\) direction, recall the definition of bisimulation:
-    \[
-        \begin{tikzcd}
-            S \ar[swap]{d}{\norm(δ_S)} 
-            & ∼ \ar{d}{δ_∼} \ar[swap]{l}{π₁} \ar{r}{π₂}
-            & T \ar{d}{\norm(δ_T)}\\  
-            G(S) 
-            & G(∼) \ar{l}{G(π₁)} \ar[swap]{r}{G(π₂)} 
-            & G(T) 
-        \end{tikzcd}
-    \]
+    We first prove \(⟹\) direction, recall that there exists a bisimulation \({∼} ⊆ \norm(⟨s⟩) × \norm(⟨u⟩)\) if and only if for all \(s₁ ∼ u₁\):
+    \begin{itemize}
+        \item for all results \(r ∈ \{\accept, \reject\}\): \(δ_{\norm(S)}(s₁, α) = r ⟺ δ_{\norm(U)}(u₂, α) = r\);
+        \item otherwise, let \((s₂, p) ≜ δ_{\norm(S)}(s₁, α)\) and \((u₂, q) ≜ δ_{\norm(u)}(u₁, α)\), then \(p = q\) and \(s₂ ∼ u₂\)
+    \end{itemize}
 
@@
     \begin{align*}
-        & δ_S(s, α) = \accept \\
-        ⟺ & \norm(δ_S)(s, α) = \accept \\
-        ⟺ & \norm(δ_∼)((s,t), α) = \accept \\ 
-        ⟺ & \norm(δ_T)(t, α) = \accept \\
-        ⟺ & δ_T(t, α) = \accept
-    \end{align*}
-
-    The condition~\labelcref{itm:transition-bisim} holds, by case analysis on the liveness of \(s'\) and \(t'\).
-    First note that \(s'\) and \(t'\) has to be both live or both dead: because \(δ_S(s, α) = (s', p)\), then \(\norm(δ_S)(s', α)\) can either be rejection or \((s',p)\), and so is \(\norm(δ_T)(t', α)\).
-Finally because \(s ∼ t\), then
+        δ_S(s, α) = \accept 
+        ⟺{}& δ_{\norm(S)}(s, α) = \accept \\
+        ⟺{}& δ_∼((s,u), α) = \accept \\ 
+        ⟺{}& δ_{\norm(U)}(u, α) = \accept \\
+        ⟺{}& δ_U(u, α) = \accept
+    \end{align*}
+
+    The condition~\labelcref{itm:rej-or-dead} holds: 
+    \begin{align*}
+        & δ_S(s, α) \text{ rejects or transition to dead states} \\
+        ⟺{}& δ_{\norm(S)}(s, α) = \reject \\
+        ⟺{}& δ_{\norm(U)}(u, α) = \reject \\
+        ⟺{}& δ_U(u, α) \text{ rejects or transition to dead states}.
+    \end{align*}
+
+    The condition~\labelcref{itm:transition-bisim} holds, by case analysis on the liveness of \(s'\) and \(u'\).
+    First note that \(s'\) and \(u'\) has to be both live or both dead: because \(δ_S(s, α) = (s', p)\), then \(\norm(δ_S)(s', α)\) can either be rejection or \((s',p)\), and so is \(\norm(δ_U)(u', α)\):
     \begin{align*}
         s' \text{ is live} 
-        & ⟺ \norm(δ_S)(s, α) = (s', p) \\
-        & ⟺ \norm(δ_T)(t, α) = (t', p) \\
-        & ⟺ t' \text{ is live}.
+        & ⟺ δ_{\norm(S)}(s, α) = (s', p) \\
+        & ⟺ δ_{\norm(U)}(u, α) = (u', p) \\
+        & ⟺ u' \text{ is live}.
     \end{align*}
     \begin{itemize}
-        \item If both \(s'\) and \(t'\) are live, then \(s' ∼ t'\). By~\cref{thm:sub-coalg-preserve-bisim}, the bisimulation \(∼_{s', t'}\) is just \(∼\) restricted to \(⟨s'⟩\) and \(⟨t'⟩\).
-        \item If both \(s'\) and \(t'\) are dead, then \(∼_{s', t'}\) can just be the singleton relation, according to~\cref{thm:bisim-between-dead}.
+        \item If both \(s'\) and \(u'\) are live, then \(s' ∼ u'\). By~\cref{thm:sub-coalg-preserve-bisim}, the bisimulation \(∼'\) is just \(∼\) restricted to \(⟨s'⟩\) and \(⟨u'⟩\).
+        \item If both \(s'\) and \(u'\) are dead, then \(∼'\) can just be the singleton relation, according to~\cref{thm:bisim-between-dead}.
     \end{itemize}
 
-    The condition~\labelcref{itm:transition-dead} holds: by the proof of condition~\labelcref{itm:transition-bisim}, \(s'\) and \(t'\) has to be either both live or both dead; if they are both live, then there cannot be a element in \(G(∼)\) that can project to \((s', p)\) under \(π₁\) but projects to \((t', q)\) under \(π₂\).
+    The condition~\labelcref{itm:transition-dead} holds: by the proof of condition~\labelcref{itm:transition-bisim}, \(s'\) and \(u'\) has to be either both live or both dead; if they are both live, then there cannot be a element in \(G(∼)\) that can project to \((s', p)\) under \(π₁\) but projects to \((t', q)\) under \(π₂\).
 Thus both \(s'\) and \(t'\) has to be dead.
 
-    The condition~\labelcref{itm:rej-or-dead} holds: 
-    \begin{align*}
-        & δ_S(s, α) \text{ rejects or transition to dead states} \\
-        ⟺ & \norm(δ_S)(s, α) = \reject \\
-        ⟺ & \norm(δ_T)(t, α) = \reject \\
-        ⟺ & δ_T(t, α) \text{ rejects or transition to dead states}.
-    \end{align*}
-
-    We then show the \(⟸\) direction, we use \(≡_{s', t'}\) to denote the maximal bisimulation between \(⟨s'⟩\) and \(⟨t'⟩\).
-    \[{∼'} ≜ ⋃ \{≡_{s', t'} ∣ ∃ α ∈ \At, p ∈ K, δ_S(s, α) = (s', p) \text{ and } δ_T(t, α) = (t', p)\}.\]
-    Notice because \(⟨s'⟩ ⊑ ⟨s⟩\) and \(⟨t'⟩ ⊑ ⟨t⟩\), then \(≡_{s', t'}\) is a bisimulation between \(⟨s⟩\) and \(⟨t⟩\), and because bisimulation is closed under arbitrary union~\cite{rutten_UniversalCoalgebraTheory_2000}, then \(∼'\) is a bisimulation between \(⟨s⟩\) and \(⟨t⟩\).
-
-    We then augment \(∼'\) with \((s, t)\) to obtain the bisimulation we required:
-    \begin{mathpar}
-        {∼} ≜ {∼'} ∪ \{(s, t)\} \and 
-        δ_{∼}((s₁, t₁), α) ≜ \begin{cases}
-            δ_{∼'}((s,t), α) & (s,t) ≠ (s₁, t₁) \\  
-            \accept & \norm(δ_S)(s, α) = \norm(δ_T)(s, α) = \accept \\  
-            \reject & \norm(δ_S)(s, α) = \norm(δ_T)(s, α) = \reject \\  
-            ((s₂, t₂), p) & \norm(δ_S)(s, α) = (s₂, p) \text{ and } \norm(δ_T)(s, α) = (t₂, p) \\  
-        \end{cases}
-    \end{mathpar}
-    The above definition of \(δ_{∼}\) is indeed well-defined, by case analysis on the result of \(δ_S\) and \(δ_T\) using the condition above:
+    We then show the \(⟸\) direction, for arbitrary \(s' ∈ S\) and \(u' ∈ U\), we use \(≡_{s', u'}\) to denote the maximal bisimulation between \(\norm(⟨s'⟩)\) and \(\norm(⟨u'⟩)\).
+    \begin{align*}
+        {∼'} ≜ ⋃ \{≡_{s', u'} & ∣ ∃ α ∈ \At, p ∈ K, \\*
+            & δ_{\norm(S)}(s, α) = (s', p) \\*
+            & \text{ and } δ_{\norm(U)}(u, α) = (u', p)\}.
+    \end{align*}
+    For all the \(s'\) and \(u'\) in the above definition, \(⟨s'⟩ ⊑ ⟨s⟩\) and \(⟨u'⟩ ⊑ ⟨u⟩\), therefore by~\Cref{thm:norm-sub-coalg}, \(\norm(⟨s'⟩) ⊑ \norm(⟨s⟩)\) and \(\norm(⟨u'⟩) ⊑ \norm(⟨u⟩)\). 
+    By~\Cref{thm:sub-coalg-preserve-bisim}, every \(≡_{s', u'}\) is a bisimulation between \(\norm(⟨s⟩)\) and \(\norm(⟨t⟩)\), and because bisimulation is closed under arbitrary union~\cite{rutten_UniversalCoalgebraTheory_2000}, \(∼'\) is a bisimulation between \(\norm(⟨s⟩)\) and \(\norm(⟨t⟩)\).
+
+    To obtain the desired bisimulation \({∼}\) between \(\norm(⟨s⟩)\) and \(\norm(⟨u⟩)\), we add the pair \((s, t)\) to \(∼'\), 
+    \[{∼} ≜ {∼'} ∪ \{(s, u)\},\] 
+    with the following transition \(δ_{∼}\): for all \(α ∈ \At\),
+    \begin{itemize}[nosep]
+        \item If \(δ_{\norm(S)}(s, α) = δ_{\norm(U)}(u, α) = \accept\), then \(δ_{∼}((s, u), α) ≜ \accept\);
+        \item If \(δ_{\norm(S)}(s, α) = δ_{\norm(U)}(u, α) = \reject\), then \(δ_{∼}((s, u), α) ≜ \reject\);
+        \item If \(δ_{\norm(S)}(s, α) = (s', p)\) and \(δ_{\norm(U)}(u, α) = (u', p)\), then \(δ_{∼}((s, u), α) = ((s', u'), p)\)
+        \item Finally, for all \((s', u') ∈ {∼}'\) that is not equal to \((s, u)\), we let \(δ_∼\) inherits the transition of \(δ_{∼'}\), i.e. \(δ_{∼}((s', u'), α) = δ_{∼'}((s', u'), α)\)
+    \end{itemize}
+    By checking all the requirement for bisimulation, we can verify that \emph{if \(δ_{∼}\) is well-defined}, then \(∼\) is indeed a bisimulation between \(\norm(⟨s⟩)\) and \(\norm(⟨u⟩)\).
+
+    Finally, we can verify that \(δ_∼\) is well-defined by leveraging the conditions in~\Cref{thm:inductive-construction}. 
+    Specifically, we will show that the definition of \(δ_∼\) covers all the possible cases, by case analysis on the result of \(δ_S\): for all \(α ∈ \At\),
     \begin{itemize}
-        \item If \(δ_S(s, α) = \accept\), then by condition~\labelcref{itm:acc-condition}, \(δ_T(t, α) = \accept\) therefore \[\norm(δ_S)(s, α) = \norm(δ_T)(s, α) = \accept.\]
-        \item If \(δ_S(s, α)\) transitions to a dead state or reject, then by conditions~\labelcref{itm:rej-or-dead} \(δ_T(t, α)\) will also transition to a dead state or reject, then \[norm(δ_S)(s, α) = \norm(δ_T)(s, α) = \reject.\]
-        \item If \(δ_S(s, α) = (s', p)\), then by condition~\labelcref{itm:acc-condition} and condition~\labelcref{itm:rej-or-dead}, \(δ_T(t, α) = (t', q)\).
-By the contrapositive of condition~\labelcref{itm:transition-dead}, if either \(s, t\) are live, then \(p = q\).
+        \item If \(δ_S(s, α) = \accept\), then by condition~\labelcref{itm:acc-condition}, \(δ_U(u, α) = \accept\); therefore \[(δ_{\norm(S)})(s, α) = \norm(δ_{\norm(U)})(u, α) = \accept.\]
+        \item If \(δ_S(s, α)\) transitions to a dead state or reject, then by conditions~\labelcref{itm:rej-or-dead} \(δ_U(u, α)\) will also transition to a dead state or reject, then \[δ_{\norm(S)}(s, α) = δ_{\norm(U)}(u, α) = \reject.\]
+        \item If \(δ_S(s, α) = (s', p)\) and \(s'\) is live, then by condition~\labelcref{itm:acc-condition} and condition~\labelcref{itm:rej-or-dead}, \(δ_U(u, α) = (u', q)\).
+By the contrapositive of condition~\labelcref{itm:transition-dead}, if either \(s, u\) are live, then \(p = q\).
 
@@
 \(s' ∼_{s', t'} t'\).
-Because bisimulation preserves liveness (\cref{thm:bisim-preserve-liveness}), \(s', t'\) has to be both dead or live, the both dead case is handled by the previous item, both live case will give us the case we desired:
-        \[\norm(δ_S)(s, α) = (s', p) \text{ and } \norm(δ_T)(t, α) = (t', p)\]
+Because bisimulation preserves liveness (\cref{thm:bisim-preserve-liveness}), \(s', t'\) has to be both dead or live.
+        Finally, because \(s'\) is live, therefore \(u'\) is also live, and we obtain the final case in the definition of \(δ_{∼}\): \(δ_{\norm(S)}(s, α) = (s', p)\) and \(δ_{\norm(U)}(u, α) = (u', p)\).
     \end{itemize}
-    And the diagram bisimulation needing to satisfy can be verified by unfolding the definition.
-\end{proof}
-
-The above theorem already gives us a way to recursively construct a algorithm that include \(s ∼ t\), this consequently will let us decide the trace equivalence of \(s\) and \(t\): \(⟦s⟧ = ⟦t⟧\).
+\end{proof}
+
+The above theorem already gives us a way to recursively construct an algorithm that include \(s ∼ u\), this consequently will let us decide the trace equivalence of \(s\) and \(u\): \(⟦s⟧ = ⟦u⟧\).
 However, this algorithm can be further optimized, we will then derive that a dead state can never relate to live states.
@@
 
-\begin{algorithm}
+\begin{algorithm*}
     \caption{On-the-fly bisimulation algorithm}\label{alg:bisim}
@@
 \begin{algorithm*}
-    \caption{On-the-fly bisimulation algorithm}\label{alg:symb-bisim}
+    \caption{Symbolic On-the-fly bisimulation algorithm}\label{alg:symb-bisim}
     \begin{algorithmic}

### 2025-01-02 main.tex
minor tweak

@@
     \end{itemize}
-    By checking all the requirement for bisimulation, we can verify that \emph{if \(δ_{∼}\) is well-defined}, then \(∼\) is indeed a bisimulation between \(\norm(⟨s⟩)\) and \(\norm(⟨u⟩)\).
+    By checking all the requirement for bisimulation, we can verify that \emph{if \(δ_{∼}\) is well-defined}, then \(∼\) is indeed a bisimulation between \(\norm(⟨s⟩)\) and \(\norm(⟨u⟩)\) and \(s ∼ u\).
 
@@
         Finally, because \(s'\) is live, therefore \(u'\) is also live, and we obtain the final case in the definition of \(δ_{∼}\): \(δ_{\norm(S)}(s, α) = (s', p)\) and \(δ_{\norm(U)}(u, α) = (u', p)\).
+        \qedhere
     \end{itemize}

### 2025-01-03 main.tex
minor fix in recursive construction and a dedicated correctness theorem

--- 2025-01-02 main.tex
@@
     \begin{itemize}[nosep]
-        \item If \(δ_{\norm(S)}(s, α) = δ_{\norm(U)}(u, α) = \accept\), then \(δ_{∼}((s, u), α) ≜ \accept\);
-        \item If \(δ_{\norm(S)}(s, α) = δ_{\norm(U)}(u, α) = \reject\), then \(δ_{∼}((s, u), α) ≜ \reject\);
-        \item If \(δ_{\norm(S)}(s, α) = (s', p)\) and \(δ_{\norm(U)}(u, α) = (u', p)\), then \(δ_{∼}((s, u), α) = ((s', u'), p)\)
-        \item Finally, for all \((s', u') ∈ {∼}'\) that is not equal to \((s, u)\), we let \(δ_∼\) inherits the transition of \(δ_{∼'}\), i.e. \(δ_{∼}((s', u'), α) = δ_{∼'}((s', u'), α)\)
+        \item if \(δ_{\norm(S)}(s, α) = δ_{\norm(U)}(u, α) = \accept\), then \(δ_{∼}((s, u), α) ≜ \accept\);
+        \item if \(δ_{\norm(S)}(s, α) = δ_{\norm(U)}(u, α) = \reject\), then \(δ_{∼}((s, u), α) ≜ \reject\);
+        \item if \(δ_{\norm(S)}(s, α) = (s', p)\) and \(δ_{\norm(U)}(u, α) = (u', p)\), then \(δ_{∼}((s, u), α) = ((s', u'), p)\);
+        \item for all \((s', u') ∈ {∼}'\) that is not equal to \((s, u)\), we let \(δ_∼\) inherits the transition of \(δ_{∼'}\), i.e. \(δ_{∼}((s', u'), α) = δ_{∼'}((s', u'), α)\)
     \end{itemize}
-    By checking all the requirement for bisimulation, we can verify that \emph{if \(δ_{∼}\) is well-defined}, then \(∼\) is indeed a bisimulation between \(\norm(⟨s⟩)\) and \(\norm(⟨u⟩)\).
-
-    Finally, we can verify that \(δ_∼\) is well-defined by leveraging the conditions in~\Cref{thm:inductive-construction}. 
+    \emph{if \(δ_{∼}\) is well-defined}, then we can verify that \(∼\) is indeed a bisimulation between \(\norm(⟨s⟩)\) and \(\norm(⟨u⟩)\) where \(s ∼ u\). 
+    We show the slightly more complicated case: \(δ_{\norm(S)}(s, α) = (s', p)\) and \(δ_{\norm(U)}(u, α) = (u', p)\) implies \(s' ∼ u'\) as an example. 
+    By condition~\labelcref{itm:transition-bisim}, there exists a bisimulation \(∼_{s', u'} ⊆ \norm(⟨s'⟩) × \norm(⟨u'⟩)\), and because \(≡_{s', u'}\) is the maximal bisimulation between \(\norm(⟨s'⟩)\) and \(\norm(⟨u'⟩)\),
+    \[(s', u') ∈ {∼_{s', u'}} ⊆ {≡_{s', u'}} ⊆ {∼}.\]
+
+    Finally, we demonstrate that \(δ_∼\) is well-defined by leveraging the conditions in~\Cref{thm:inductive-construction}. 
     Specifically, we will show that the definition of \(δ_∼\) covers all the possible cases, by case analysis on the result of \(δ_S\): for all \(α ∈ \At\),
@@
         \item If \(δ_S(s, α) = \accept\), then by condition~\labelcref{itm:acc-condition}, \(δ_U(u, α) = \accept\); therefore \[(δ_{\norm(S)})(s, α) = \norm(δ_{\norm(U)})(u, α) = \accept.\]
-        \item If \(δ_S(s, α)\) transitions to a dead state or reject, then by conditions~\labelcref{itm:rej-or-dead} \(δ_U(u, α)\) will also transition to a dead state or reject, then \[δ_{\norm(S)}(s, α) = δ_{\norm(U)}(u, α) = \reject.\]
-        \item If \(δ_S(s, α) = (s', p)\) and \(s'\) is live, then by condition~\labelcref{itm:acc-condition} and condition~\labelcref{itm:rej-or-dead}, \(δ_U(u, α) = (u', q)\).
-By the contrapositive of condition~\labelcref{itm:transition-dead}, if either \(s, u\) are live, then \(p = q\).
-
-        Then by condition~\labelcref{itm:transition-bisim}, there exists a bisimulation \(∼_{s', t'}\) between \(⟨s'⟩\) and \(⟨t'⟩\) s.t.
-\(s' ∼_{s', t'} t'\).
-Because bisimulation preserves liveness (\cref{thm:bisim-preserve-liveness}), \(s', t'\) has to be both dead or live.
+        \item If \(δ_S(s, α)\) transitions to a dead state or reject, then by condition~\labelcref{itm:rej-or-dead} \(δ_U(u, α)\) will also transition to a dead state or reject, then \[δ_{\norm(S)}(s, α) = δ_{\norm(U)}(u, α) = \reject.\]
+        \item If \(δ_S(s, α) = (s', p)\) and \(s'\) is live, then \(δ_U(u, α) = (u', p)\) necessarily holds, otherwise it would violate one of conditions~\labelcref{itm:acc-condition,itm:rej-or-dead,itm:transition-dead}.
+
+        By condition~\labelcref{itm:transition-bisim}, there exists a bisimulation \(∼_{s', u'}\) between \(\norm(⟨s'⟩)\) and \(\norm(⟨u'⟩)\) s.t.
+\(s' ∼_{s', u'} u'\).
+Because bisimulation preserves liveness (\Cref{thm:bisim-preserve-liveness}), \(s', u'\) has to be both dead or live.
         Finally, because \(s'\) is live, therefore \(u'\) is also live, and we obtain the final case in the definition of \(δ_{∼}\): \(δ_{\norm(S)}(s, α) = (s', p)\) and \(δ_{\norm(U)}(u, α) = (u', p)\).
+        \qedhere
     \end{itemize}
+\end{proof}
+
+\begin{corollary}[Correctness]
+    For any two states in two GKAT coalgebra \(s ∈ S, u ∈ U\),the four condition in~\Cref{thm:inductive-construction} is satisfied if and only if \(s\) and \(u\) are finite-trace equivalent
+\end{corollary}
+
+\begin{proof}
+    By the standard argument with normalization preserves sub-coalgebra (\Cref{thm:norm-sub-coalg}), we can obtain 
+    \(\norm(⟨s⟩) ⊑ \norm(S)\) and \(\norm(⟨u⟩) ⊑ \norm(U)\).
+    Because sub-coalgebra reflects and preserves bisimulation (\Cref{thm:sub-coalg-preserve-bisim}) with the correctness of bisimulation between normalized coalgebras: let \({∼}\) and \({∼'}\) be bisimulations,
+    \begin{align*}
+        & ∃~ {∼} ⊆ \norm(⟨s⟩) × \norm(⟨u⟩), s ∼ u \\
+        ⟺{}& ∃~ {∼'} ⊆ \norm(S) × \norm(U), s ∼ u \\ 
+        ⟺{}& ⟦s⟧_{S} = ⟦u⟧_{U}. \qedhere
+    \end{align*}
 \end{proof}

### 2025-01-05 main.tex
Finish the on-the-fly algorithm

@@
 \algnewcommand\algorithmicdefault{\textbf{default}}
-\algnewcommand\Skip{\textbf{skip}}
+\algnewcommand\Continue{\textbf{continue}}
 \algdef{SE}[MATCH]{Match}{EndMatch}[1]{\algorithmicmatch\ #1\ \algorithmicwith}{\algorithmicend\ \algorithmicmatch}%
 \algdef{SE}[CASE]{Case}{EndCase}[1]{\algorithmiccase\ #1 \algorithmicthen}{\algorithmicend\ \algorithmiccase}%
-\algdef{SE}[DEFAULT]{Default}{EndDefault}{\algorithmicdefault\ }{\algorithmicend\ \algorithmicdefault}%
+\algdef{SE}[DEFAULT]{Default}{EndDefault}{\hskip\algorithmicindent\algorithmicdefault}{\algorithmicend\algorithmicdefault}%
 \algtext*{EndMatch}%
@@
 
-\begin{theorem}
+\begin{theorem}\label{thm:GKAT-bisim-iff-bisim-eq}
     Given two states in two coalgebras \(s ∈ S, u ∈ U\), let \(S + U\) be the coproduct coalgebra of \(S\) and \(U\)~\cite{rutten_UniversalCoalgebraTheory_2000}. 
@@
 
-    Let \(∼\) be a bisimulation and \(≃\) be a bisimulation equivalence:
+    Let \(∼\) be a bisimulation and \(≃\) be a bisimulation equivalence, by~\Cref{thm:sub-coalg-preserve-bisim} and~\Cref{thm:bisim-iff-bisim-eq}:
     \begin{align*}
@@
 
-\begin{theorem}[Recursive Construction]\label{thm:inductive-construction}
+\begin{theorem}[Recursive Construction]\label{thm:recursive-construction}
     Given two GKAT coalgebra \(S\) and \(U\), and two of their elements \(s ∈ S\) and \(u ∈ U\),
@@
 
-    Finally, we demonstrate that \(δ_∼\) is well-defined by leveraging the conditions in~\Cref{thm:inductive-construction}. 
+    Finally, we demonstrate that \(δ_∼\) is well-defined by leveraging the conditions in~\Cref{thm:recursive-construction}. 
     Specifically, we will show that the definition of \(δ_∼\) covers all the possible cases, by case analysis on the result of \(δ_S\): for all \(α ∈ \At\),
@@
 
-\begin{corollary}[Correctness]
-    For any two states in two GKAT coalgebra \(s ∈ S, u ∈ U\),the four condition in~\Cref{thm:inductive-construction} is satisfied if and only if \(s\) and \(u\) are finite-trace equivalent
+\begin{corollary}[Correctness]\label{thm:recursive-construction-correctness}
+    For any two states in two GKAT coalgebra \(s ∈ S, u ∈ U\), \(s\) and \(u\) are finite-trace equivalent if and only if there exists a \emph{bisimulation equivalence} \(≃\) in \(\norm(⟨s⟩) + \norm(⟨u⟩)\), which is equivalent to all the following conditions hold:
+    \begin{enumerate}
+        \item for all \(α ∈ \At\), \(δ_{S}(s, α) = \accept ⟺ δ_{U}(u, α) = \accept\);
+        \item \(s\) reject \(α\) or transition to a dead state via \(α\) if and only if \(u\) rejects \(α\) or transition to a dead state via \(α\);  
+        \item if \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', p)\), then there exists a bisimulation equivalence \({≃'}\) in \(\norm(⟨s'⟩) + \norm(⟨u'⟩)\), s.t. \(s' ≃' u'\);
+        \item if \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', q)\), s.t. \(p ≠ q\), then both \(s'\) and \(t'\) are dead.
+    \end{enumerate}
+    Notice that above condition is similar to those in~\cref{thm:recursive-construction}, except bisimulation is replaced with bisimulation \emph{equivalence}.
 \end{corollary}
@@
     \(\norm(⟨s⟩) ⊑ \norm(S)\) and \(\norm(⟨u⟩) ⊑ \norm(U)\).
-    Because sub-coalgebra reflects and preserves bisimulation (\Cref{thm:sub-coalg-preserve-bisim}) with the correctness of bisimulation between normalized coalgebras: let \({∼}\) and \({∼'}\) be bisimulations,
-    \begin{align*}
-        & ∃~ {∼} ⊆ \norm(⟨s⟩) × \norm(⟨u⟩), s ∼ u \\
-        ⟺{}& ∃~ {∼'} ⊆ \norm(S) × \norm(U), s ∼ u \\ 
-        ⟺{}& ⟦s⟧_{S} = ⟦u⟧_{U}. \qedhere
-    \end{align*}
-\end{proof}
-
-The above theorem already gives us a way to recursively construct an algorithm that include \(s ∼ u\), this consequently will let us decide the trace equivalence of \(s\) and \(u\): \(⟦s⟧ = ⟦u⟧\).
-However, this algorithm can be further optimized, we will then derive that a dead state can never relate to live states.
-This means that when checking the bisimulation of states \(s\) and \(t\), if we already know one of them is dead, we only need to check whether the other is dead, instead of going through the convoluted process mentioned in~\cref{thm:inductive-construction}.
-
-However because homomorphism preserves liveness, if we already know one of the \(s\) and \(t\) is dead, the other has to be dead.
+    Because sub-coalgebra reflects and preserves bisimulation (\Cref{thm:sub-coalg-preserve-bisim}) and the correspondence between bisimulations and bisimulation equivalences in GKAT (\cref{thm:GKAT-bisim-iff-bisim-eq}): let \({∼_{s, u}} ⊆ \norm(⟨s⟩) × \norm(⟨u⟩)\) and \({∼} ⊆ \norm(S) × \norm(U)\) be bisimulations and \(≃\) be a bisimulation equivalence in \(\norm(⟨s⟩) + \norm(⟨u⟩)\):
+    \begin{align*}
+        & \text{Conditions in \cref{thm:recursive-construction} is satisfied} \\  
+        ⟺{}& \text{Conditions in current theorem is satisfied}\\
+        ⟺{}& ∃~ {≃}, s ≃ u 
+        ⟺ ∃~ {∼_{s, u}}, s ∼ u \\
+        ⟺{}& ∃~ {∼} , s ∼ u 
+        ⟺ ⟦s⟧_{S} = ⟦u⟧_{U}. \qedhere
+    \end{align*}
+\end{proof}
+
+The above theorem already gives us an algorithm to recursively decide whether \(⟦s⟧_{S} = ⟦u⟧_U\), when \(⟨s⟩_S\) and \(⟨u⟩_U\) is finite.
+However, this algorithm can be further optimized in several ways: we will then derive that a dead state can never relate to live states.
+This means that when checking the bisimulation of states \(s\) and \(t\), if we already know one of them is dead, we only need to check whether the other is dead, instead of going through all the conditions in~\cref{thm:recursive-construction}.
 
 \begin{theorem}\label{thm:bisim-one-dead}
-    Given two states \(s ∈ S\) and \(t ∈ T\), if \(s\) is a dead state in \(S\), then there exists a bisimulation \(∼\) between \(S\) and \(T\) where \(s ∼ t\) if and only if \(t\) is dead.
-Similarly for \(t ∈ T\).
+    Given two states \(s ∈ S\) and \(t ∈ T\), if \(s\) is a dead state in \(S\), then there exists a bisimulation equivalence \(≃\) in \(S + T\) where \(s ≃ t\) if and only if \(t\) is dead.
+Similarly, for \(t ∈ T\).
 \end{theorem}
@@
 \begin{proof}
-    if there exists a bisimulation \(∼\), s.t. \(s ∼ t\), because \(s\) is dead and bisimulation preserves liveness~\cref{thm:bisim-preserve-liveness}, then \(t\) is dead.
-
-    And if both \(t\) and \(s\) is dead, then a bisimulation can by constructed by~\cref{thm:bisim-between-dead}.
-\end{proof}
-
-\section{The Algorithm}
-
-In this section we will present the pseudo-code for our on-the-fly algorithm. 
-In order to implement the the inductive construction theorem (\cref{thm:inductive-construction}), we will need to determine the liveness of the state.
-This can be simply computed via a DFS from the state being checked.
-
-TODO: we should merge the two so that it is easier to 
+    Notice because of the homomorphism \(\mathrm{inj_l}: S → S + T\) and homomorphism preserves liveness (\cref{thm:hom-preserve-liveness}), therefore \(s ∈ S\) is dead if and only if \(s ∈ S + T\) is dead; similarly for \(t ∈ T\).
+
+    If there exists a bisimulation equivalence \(≃\), s.t. \(s ≃ t\), because \(s\) is dead and bisimulation preserves liveness~\cref{thm:bisim-preserve-liveness}, then \(t\) is dead.
+
+    And if both \(t\) and \(s\) is dead, then a bisimulation between \(S\) and \(T\) can be constructed in~\cref{thm:bisim-between-dead}, which induces a bisimulation equivalence via~\Cref{thm:GKAT-bisim-iff-bisim-eq}.
+\end{proof}
+
+Before we write down the optimized algorithm, we will first briefly sketch the dead state detection algorithm, which will use a DFS to check whether the there exists any accepting states in reachable states of \(s ∈ S\): if there exists any accepting state in \(⟨s⟩_S\), then the algorithm terminates immediately, and return that \(s\) is live, otherwise it would return all the states in \(⟨s⟩\), and by~\cref{thm:dead-iff-all-reachable-dead}, all the states in \(⟨s⟩\) is dead.
+
+We will cache all the known dead states from all the previous searches, we use the call \Call{knownDead\(_S\)}{$s$} to check whether the state \(s\) is in the cached dead states. 
+And \Call{isDead\(_S\)}{$s$} will first check if \(s\) is known to be dead, and invoke the depth-first-search algorithm in the coalgebra \(s\), if \(s\) is not in the cached dead states.
+Because the non-symbolic version of dead state checking is similar to the symbolic version, we only provide the symbolic version of this algorithm in appendix
+TODO: give the link to the section.
+
 \begin{algorithm}
-    \caption{Check whether a state \(s\) is dead}\label{alg:check-dead-main}
+    \caption{On-the-fly bisimulation algorithm}\label{alg:bisim}
     \begin{algorithmic}
-        \Function{isDeadLoop}{$s ∈ S$, explored}
-        \If {\(s ∈\) explored} {\Return explored} 
-        \Else { 
+        \Function{equiv}{$s ∈ S, u ∈ U$}
+        \If {\Call{eq}{$s, u$}} {\Return true} \EndIf
+        \If {\Call{knownDead\(_S\)}{s}} 
+            \State\Return \Call{isDead\(_U\)}{$u$}
+        \EndIf
+        \If {\Call{knownDead\(_U\)}{u}} 
+            \State\Return \Call{isDead\(_S\)}{$s$}
+        \EndIf
         \For{\(α ∈ \At\)}{}{
-                \Match{\(δ_{S}(s, α)\)}
-                \Case{\(\accept\)} {\Return none} \Comment{\(s\) transition to accept}
+            \Match{$δ_{S}(s, α), δ_{U}(u, α)$}
+            \Case{\(\accept, \accept\)} {\Continue} \EndCase
+            \Case{\(\reject, \reject\)} {\Continue} \EndCase
+            \Case{\(\reject, (u', q)\)} 
+                \State\Return \Call{isDead\(_U\)}{$u'$} 
             \EndCase
-                \Case{\(\reject\)} {\texttt{continue}}
-                \Comment{skip if \(s\) transition to reject}
+            \Case{\((s', p), \reject\)}
+                \State\Return\Call{isDead\(_S\)}{$s'$}
             \EndCase
-                \Case{($s', p$)}{ 
-                    \If{\Call{IsDeadLoop}{$s'$} = none} {\Return none} \Comment{\(s\) transitions to a live state \(s'\)}
-                    \Else { explored \(←\) (explored \(∪\) \Call{isDeadLoop}{$s'$, explored}) } \EndIf
+            \Case{\((s', p), (u', q)\)} {
+                \If {\(p = q\)}
+                    \State\Call{union}{$s, u$}
+                    \State\Return \Call{equiv}{$s, t$} 
+                \EndIf
+                \If{\Call{isDead\(_S\)}{$s$} and \Call{isDead\(_U\)}{$u$}}
+                    \State\Continue
+                \EndIf
+                \State\Return false
             } \EndCase
+            \Default { \Return false } \EndDefault
             \EndMatch
-            }\EndFor}
-        \EndIf
-        \State {\Return explored}    
+        }\EndFor
+        \State\Return true
         \EndFunction
@@
 
-By~\cref{thm:dead-iff-all-reachable-dead}, if \(s\) is dead then all the reachable states of \(s\) (denoted by \(⟨s⟩\)).
+TODO: I am not sure if we need to give some intuition for this algorithm here.
+
+Finally, we present our finite-trace equivalent checking algorithm as~\cref{alg:bisim}, where we search for a bisimulation equivalence \({≃}\) in \(\norm(⟨s⟩) + \norm(⟨u⟩)\), s.t.
+\(s ≃ u\), using the condition in~\Cref{thm:recursive-construction-correctness}.
+Seeking an equivalence relation allows us to organize our explored state pairs into equivalence classes, as opposed to a set of pairs, and enables us to use union-find data structures to avoid looking for two states that are already in the same equivalence class.
+Specifically, we use the call \(\Call{union}{$s, u$}\) to union the equivalence class of \(s\) and \(u\), and use \(\Call{eq}{$s, u$}\) to check whether \(s\) and \(u\) is in the same equivalence class.
+In this algorithm, we first check if one of \(s\) and \(u\) is known to be dead, if so we only need to check whether the other is dead, because of~\Cref{thm:bisim-one-dead}.
+Otherwise, we will check the conditions in~\Cref{thm:recursive-construction-correctness}, which is a minor modification of the conditions in~\cref{thm:recursive-construction}.
+
+% \section{The Algorithm}
+
+% In this section we will present the pseudo-code for our on-the-fly algorithm. 
+% In order to implement the the inductive construction theorem (\cref{thm:recursive-construction}), we will need to determine the liveness of the state.
+This can be simply computed via a DFS from the state being checked.
+
+% TODO: we should merge the two so that it is easier to 
+% \begin{algorithm}
+%     \caption{Check whether a state \(s\) is dead}\label{alg:check-dead-main}
+%     \begin{algorithmic}
+%         \Function{isDeadLoop}{$s ∈ S$, explored}
+%         \If {\(s ∈\) explored} {\Return explored} 
+%         \Else { 
+%             \For{\(α ∈ \At\)}{}{
+%                 \Match{\(δ_{S}(s, α)\)}
+%                 \Case{\(\accept\)} {\Return none} \Comment{\(s\) transition to accept}
+%                 \EndCase
+%                 \Case{\(\reject\)} {\texttt{continue}}
+%                 \Comment{skip if \(s\) transition to reject}
+%                 \EndCase
+%                 \Case{($s', p$)}{ 
+%                     \If{\Call{IsDeadLoop}{$s'$} = none} {\Return none} \Comment{\(s\) transitions to a live state \(s'\)}
+%                     \Else { explored \(←\) (explored \(∪\) \Call{isDeadLoop}{$s'$, explored}) } \EndIf
+%                 } \EndCase
+%                 \EndMatch
+%             }\EndFor}
+%         \EndIf
+%         \State {\Return explored}    
+%         \EndFunction
+%     \end{algorithmic}
+% \end{algorithm}
+
+% By~\cref{thm:dead-iff-all-reachable-dead}, if \(s\) is dead then all the reachable states of \(s\) (denoted by \(⟨s⟩\)).
 Then by returning all the reachable states of \(s\), we can cache these states to avoid checking them again.
@@
 
-\begin{algorithm}
-    \caption{A cached algorithm to check whether a state is dead}\label{alg:is-dead}
-    \begin{algorithmic}
-        \State{deadStates \(← ∅\)}
-
-        \Function{isDead}{$s ∈ S$}
-        \If {\(s ∈\) deadStates} {\Return true} 
-        \ElsIf {\Call{isDeadLoop}{$s, ∅$} = none} {\Return false}
-        \Else 
-            \State {deadStates \(←\) (deadStates \(∪\) \Call{isDeadLoop}{$s, ∅$})}
-            \State {\Return {true}}
-        \EndIf
-        \EndFunction
-    \end{algorithmic}
-\end{algorithm}
-
-Given the direct correspondence between bisimulation and bisimulation equivalence and bisimulation in sub-algebra:
-\begin{align*}
-    & ∃ \text{ bisimulation } {∼} ⊆ ⟨s⟩ × ⟨t⟩ \text{ s.t. } s ∼ t \\
-    & ⟺ ∃ \text{ bisimulation } {∼} ⊆ (⟨s⟩∪⟨t⟩) × (⟨s⟩∪⟨t⟩) \text{ s.t. } s ∼ t 
-        & \text{\cref{thm:sub-coalg-preserve-bisim}}\\  
-    & ⟺ ∃ \text{ bisimulation equivalence } {≃} ⊆ (⟨s⟩∪⟨t⟩) × (⟨s⟩∪⟨t⟩) \text{ s.t. } s ≃ t 
-        & \text{\cref{thm:bisim-iff-bisim-equiv}}
-\end{align*}
-we can safely replace the bisimulation in inductive construction (\cref{thm:inductive-construction}) with bisimulation equivalence. 
-Dealing with equivalence relations allows us to leverage efficient data structures like union find in our bisimulation algorithm.
-
-We will use \(\Call{union}{$s,t$}\) to denote the operation to equate \(s\) and \(t\) in a union-find, and use \(\Call{eq}{$s,t$}\) to check if \(s\) and \(t\) belongs to the same equivalence class, i.e.
+% \begin{algorithm}
+%     \caption{A cached algorithm to check whether a state is dead}\label{alg:is-dead}
+%     \begin{algorithmic}
+%         \State{deadStates \(← ∅\)}
+
+%         \Function{isDead}{$s ∈ S$}
+%         \If {\(s ∈\) deadStates} {\Return true} 
+%         \ElsIf {\Call{isDeadLoop}{$s, ∅$} = none} {\Return false}
+%         \Else 
+%             \State {deadStates \(←\) (deadStates \(∪\) \Call{isDeadLoop}{$s, ∅$})}
+%             \State {\Return {true}}
+%         \EndIf
+%         \EndFunction
+%     \end{algorithmic}
+% \end{algorithm}
+
+% Given the direct correspondence between bisimulation and bisimulation equivalence and bisimulation in sub-algebra:
+% \begin{align*}
+%     & ∃ \text{ bisimulation } {∼} ⊆ ⟨s⟩ × ⟨t⟩ \text{ s.t. } s ∼ t \\
+%     & ⟺ ∃ \text{ bisimulation } {∼} ⊆ (⟨s⟩∪⟨t⟩) × (⟨s⟩∪⟨t⟩) \text{ s.t. } s ∼ t 
+%         & \text{\cref{thm:sub-coalg-preserve-bisim}}\\  
+%     & ⟺ ∃ \text{ bisimulation equivalence } {≃} ⊆ (⟨s⟩∪⟨t⟩) × (⟨s⟩∪⟨t⟩) \text{ s.t. } s ≃ t 
+%         & \text{\cref{thm:bisim-iff-bisim-equiv}}
+% \end{align*}
+% we can safely replace the bisimulation in inductive construction (\cref{thm:recursive-construction}) with bisimulation equivalence. 
+% Dealing with equivalence relations allows us to leverage efficient data structures like union find in our bisimulation algorithm.
+
+% We will use \(\Call{union}{$s,t$}\) to denote the operation to equate \(s\) and \(t\) in a union-find, and use \(\Call{eq}{$s,t$}\) to check if \(s\) and \(t\) belongs to the same equivalence class, i.e.
 share the same representative.
-Specifically, we will use the union-find structures to keep track of the equivalence classes that we are in the process of checking, hence avoiding repeatedly checking the same pair of states to remove infinite loops.
-
-Our on-the-fly bisimulation algorithm will decide whether there exists a bisimulation relation in \(⟨s⟩ ∪ ⟨t⟩\) s.t.
+% Specifically, we will use the union-find structures to keep track of the equivalence classes that we are in the process of checking, hence avoiding repeatedly checking the same pair of states to remove infinite loops.
+
+% Our on-the-fly bisimulation algorithm will decide whether there exists a bisimulation relation in \(⟨s⟩ ∪ ⟨t⟩\) s.t.
 \(s ∼ t\).
-This algorithm generally reproduce the setting of inductive construction theorem~\cref{thm:inductive-construction};
-except by~\cref{thm:bisim-one-dead}, in the special case where \(s\) or \(t\) is dead, then we will only need to check whether the other is dead.
-
-\begin{algorithm*}
-    \caption{On-the-fly bisimulation algorithm}\label{alg:bisim}
-    \begin{algorithmic}
-        \Function{equiv}{$s ∈ S, t ∈ T$}
-        \If {\Call{eq}{$s, t$}} {\Return true}
-        \ElsIf {\(s ∈\) deadStates\(_S\)} {\Return \Call{isDead\(_T\)}{$t$}} 
-        \ElsIf {\(t ∈\) deadStates\(_T\)} {\Return \Call{isDead\(_S\)}{$s$}} 
-        \Else 
-        \For{\(α ∈ \At\)}{}{
-            \Comment{Inductive construction,~\cref{thm:inductive-construction}}
-            \Match{$δ_{S}(s, α), δ_{T}(t, α)$}
-            \Case{\(\accept, \accept\)} {\Skip} \EndCase
-            \Case{\(\reject, \reject\)} {\Skip} \EndCase
-            \Case{\(\reject, (t', q)\)} {\Call{isDead}{$t'$}} \EndCase
-            \Case{\((s', p), \reject\)} {\Call{isDead}{$s'$}} \EndCase
-            \Case{\((s', p), (t', q)\)} {
-                \If {\(p = q\)} {\Call{union}{$s, t$}; \Call{equiv}{$s, t$}} 
-                \ElsIf {\Call{isDead}{$s$} and \Call{isDead}{$s$}} {\Skip}
-                \Else { \Return false }
-                \EndIf
-            } \EndCase
-            \Default {\Return false} \Comment{the results format does not match} \EndDefault
-            \EndMatch
-        }\EndFor
-        \EndIf
-        \Return true \Comment{no mismatch found}
-        \EndFunction
-    \end{algorithmic}
-\end{algorithm*}
-
-Because the dead state detection algorithm is coalgebra-specific, we use a subscript on ``deadStates'' and ``\textsc{IsDead}'' to indicate the coalgebra. 
-The soundness and completeness of~\cref{alg:bisim} can be observed by the fact that \emph{when the algorithm terminate}, the algorithm returns true if and only if there exists a bisimulation between \(⟨s⟩\) and \(⟨t⟩\) s.t.
+This algorithm generally reproduce the setting of inductive construction theorem~\cref{thm:recursive-construction};
+% except by~\cref{thm:bisim-one-dead}, in the special case where \(s\) or \(t\) is dead, then we will only need to check whether the other is dead.
+
+% Because the dead state detection algorithm is coalgebra-specific, we use a subscript on ``deadStates'' and ``\textsc{IsDead}'' to indicate the coalgebra. 
+% The soundness and completeness of~\cref{alg:bisim} can be observed by the fact that \emph{when the algorithm terminate}, the algorithm returns true if and only if there exists a bisimulation between \(⟨s⟩\) and \(⟨t⟩\) s.t.
 \(s ∼ t\), which is then logically equivalent to trace equivalence.
-Such equivalence is a direct consequence of~\cref{thm:bisim-one-dead,thm:inductive-construction}.
-
-\begin{remark}
-    The caching of dead state and the shortcut to check whether \(s\) is dead when \(t\) is dead and vise versa, is not essential to the soundness and completeness of algorithm, they are here to trade speed with memory.
-    In a memory-constraint situation, the ``\textnormal{deadStates}'' variable can be cleared periodically to save memory.
-\end{remark}
+% Such equivalence is a direct consequence of~\cref{thm:bisim-one-dead,thm:recursive-construction}.
+
+% \begin{remark}
+%     The caching of dead state and the shortcut to check whether \(s\) is dead when \(t\) is dead and vise versa, is not essential to the soundness and completeness of algorithm, they are here to trade speed with memory.
+%     In a memory-constraint situation, the ``\textnormal{deadStates}'' variable can be cleared periodically to save memory.
+% \end{remark}
 
@@
 
-We can then migrate the normalized bisimulation algorithm to the symbolic setting, we will first prove an inductive construction theorem like~\cref{thm:inductive-construction}.
-
-\begin{theorem}[Symbolic Inductive Construction]\label{thm:symb-inductive-construction}
+We can then migrate the normalized bisimulation algorithm to the symbolic setting, we will first prove an inductive construction theorem like~\cref{thm:recursive-construction}.
+
+\begin{theorem}[Symbolic Inductive Construction]\label{thm:symb-recursive-construction}
     Given two symbolic GKAT coalgebra \(Ŝ = ⟨S, ϵ̂_S, δ̂_S⟩\) and \(T̂ = ⟨T, ϵ̂_T, δ̂_T⟩\) and two states \(s ∈ S\) and \(t ∈ T\), there exists a normalized bisimulation on the lowered coalgebra \({∼} ⊆ S × T\) s.t.
@@
 \begin{proof}
-    Reduces to~\cref{thm:inductive-construction} i.e. all the above condition holds if and only if all the condition in~\cref{thm:inductive-construction} holds in the lowered coalgebra.
-\end{proof}
-
-Then for the algorithm, we can just recursively check all the conditions in~\cref{thm:symb-inductive-construction}.
+    Reduces to~\cref{thm:recursive-construction} i.e. all the above condition holds if and only if all the condition in~\cref{thm:recursive-construction} holds in the lowered coalgebra.
+\end{proof}
+
+Then for the algorithm, we can just recursively check all the conditions in~\cref{thm:symb-recursive-construction}.
 
@@
         \Return {
-            \Comment{conditions of ~\cref{thm:symb-inductive-construction}}
+            \Comment{conditions of ~\cref{thm:symb-recursive-construction}}
             \\\vspace{5px}

### 2025-01-06 main.tex
use trace equivalence instead of bisimulation equivalence

@@
 \newcommand{\Exp}{\theoryOf{Exp}}
-\newcommand{\GKAT}{\theoryOf{GKAT}}
-\newcommand{\Bool}{\theoryOf{Bool}}
+\newcommand{\BExp}{\theoryOf{BExp}}
 \DeclareMathOperator{\GS}{\mathrm{GS}}
@@
     We will need to show that \(h(⟨s⟩_{S})\) is the smallest sub-GKAT coalgebra of \(S\) that contain \(h(s)\).
-First by definition of image, \(h(s) ∈ h(⟨s⟩_{S})\); second by \cref{thm:hom-(pre)img-preserve-sub-coalg}, \(h(⟨S⟩_S) ⊑ U\).
+First by definition of image, \(h(s) ∈ h(⟨s⟩_{S})\); second by \cref{thm:hom-(pre)img-preserve-sub-coalg}, \(h(⟨s⟩_S) ⊑ U\).
 
@@
 Guarded Kleene Algebra with Tests, or GKAT~\cite{smolka_GuardedKleeneAlgebra_2020}, is a deterministic fragment of Kleene Algebra with Tests. 
-The syntax of GKAT over a set of primitive actions \(K\) and a set of primitive tests \(T\) can be defined in two sorts, boolean expressions \(\Bool\) and expressions \(\Exp\):
-\begin{align*}
-    a, b, c ∈ \Bool 
+The syntax of GKAT over a set of primitive actions \(K\) and a set of primitive tests \(T\) can be defined in two sorts, boolean expressions \(\BExp\) and expressions \(\Exp\):
+\begin{align*}
+    a, b, c ∈ \BExp 
         & ≜ 1 ∣ 0 ∣ t ∈ T ∣ b ∧ c ∣ b ∨ c ∣ \overline{b} \\  
     e, f ∈ \Exp 
-        & ≜ p ∈ K ∣ b ∈ \Bool ∣ e +_b f ∣ e ; f ∣ e^{(b)} 
+        & ≜ p ∈ K ∣ b ∈ \BExp ∣ e +_b f ∣ e ; f ∣ e^{(b)} 
 \end{align*}
@@
 Then the semantics of each expression \(⟦e⟧\) can be computed by the semantics of Kleene Algebra with tests~\cite{kozen_KleeneAlgebraTests_1997c}.
-An important construct in the semantics is \emph{atoms}, which are a string of all the primitive tests either in its positive or negative form: for \(B ≜ \{b₁, b₂, …, bₙ\}\)
-\[\At_B ≜ \{b₁' ⋅ b₂' ⋅ ⋯ ⋅ bₙ' ∣ bᵢ' ∈ \{bᵢ, \overline{bᵢ}\}\}.\]
+An important construct in the semantics is \emph{atoms}, which are conjunctions of all the primitive tests either in its positive or negative form: for \(B ≜ \{b₁, b₂, …, bₙ\}\)
+\[\At_B ≜ \{b₁' ∧ b₂' ∧ ⋯ ∧ bₙ' ∣ bᵢ' ∈ \{bᵢ, \overline{bᵢ}\}\}.\]
 Follow the conventional notation, we denote an atom using \(α, β\); and we sometimes omit the subscript \(B\) when no confusing can arise.
-Intuitively, atoms are truth assignments to each primitive tests, indicating which primitive tests is satisfied in the current program states.
+For a boolean expression \(b\), if \(α\) can be proven to be smaller than \(b\) using axioms of the Boolean Algebra, we will say \(b\) \emph{contains} \(α\), denoted as \(α ≤ b\).
+Alternatively, atoms can also be thought of as truth assignments to each primitive tests, indicating which primitive tests is satisfied in the current program states; and \(α ≤ b\) if and only if the truth assignment represented by \(α\) satisfies \(b\).
 
@@
 Furthermore, the finite trace equivalence between \(s ∈ S\) and \(u ∈ U\) can be computed by first normalizing \(S\) and \(U\), then decide whether there is a bisimulation on \(\norm(S)\) and \(\norm(U)\) that includes \((s, t)\).
-For a more detailed explanation on the finite trace semantics, we refer the reader to the work of \Citeauthor{smolka_GuardedKleeneAlgebra_2020}~\cite{smolka_GuardedKleeneAlgebra_2020}.
+For a more detailed explanation on the finite trace semantics, we refer the reader to the work of \Citeauthor{smolka_GuardedKleeneAlgebra_2020}~\cite{smolka_GuardedKleeneAlgebra_2020}, however we will recall two important theorems here.
+
+\begin{theorem}[Correctness~\cite{smolka_GuardedKleeneAlgebra_2020}]\label{thm:norm-bisim-correctness}
+    Given two states in two GKAT coalgebra \(s ∈ S\) and \(u ∈ U\), then there exists a bisimulation between normalized coalgebras \({∼} ⊆ \norm(S) × \norm(U)\) s.t.
+\(s ∼ u\) if and only if \(s\) and \(u\) are trace equivalent \(⟦s⟧_S = ⟦u⟧_U\)
+\end{theorem}
 
@@
     \end{tikzcd}\]
-    And because normalization operation preserves states in \(∼\), therefore \((s, u) ∈ \norm(∼)\), and because \(\norm(∼)\) is a bisimulation between the normalization of \(S\) and \(U\), therefore \(⟦s⟧_{S} = ⟦u⟧_{U}\)
+    Because \(s ∼ u\) and normalization operation preserves states in \(∼\), therefore \((s, u) ∈ \norm(∼)\), and because \(\norm(∼)\) is a bisimulation between the normalization of \(S\) and \(U\), therefore \(⟦s⟧_{S} = ⟦u⟧_{U}\) (\cref{thm:norm-bisim-correctness}).
 \end{proof}
@@
 
-\begin{theorem}[Recursive Construction]\label{thm:recursive-construction}
+\begin{lemma}\label{thm:recursive-construction-lemma}
     Given two GKAT coalgebra \(S\) and \(U\), and two of their elements \(s ∈ S\) and \(u ∈ U\),
@@
     \end{enumerate}
-\end{theorem}
+\end{lemma}
 
@@
     \begin{itemize}
-        \item If \(δ_S(s, α) = \accept\), then by condition~\labelcref{itm:acc-condition}, \(δ_U(u, α) = \accept\); therefore \[(δ_{\norm(S)})(s, α) = \norm(δ_{\norm(U)})(u, α) = \accept.\]
+        \item If \(δ_S(s, α) = \accept\), then by condition~\labelcref{itm:acc-condition}, \(δ_U(u, α) = \accept\); therefore \[δ_{\norm(S)}(s, α) = δ_{\norm(U)}(u, α) = \accept.\]
         \item If \(δ_S(s, α)\) transitions to a dead state or reject, then by condition~\labelcref{itm:rej-or-dead} \(δ_U(u, α)\) will also transition to a dead state or reject, then \[δ_{\norm(S)}(s, α) = δ_{\norm(U)}(u, α) = \reject.\]
@@
 
-\begin{corollary}[Correctness]\label{thm:recursive-construction-correctness}
-    For any two states in two GKAT coalgebra \(s ∈ S, u ∈ U\), \(s\) and \(u\) are finite-trace equivalent if and only if there exists a \emph{bisimulation equivalence} \(≃\) in \(\norm(⟨s⟩) + \norm(⟨u⟩)\), which is equivalent to all the following conditions hold:
+\begin{theorem}[Recursive Construction]\label{thm:recursive-construction}
+    For any two states in two GKAT coalgebra \(s ∈ S, u ∈ U\), \(s\) and \(u\) are finite-trace equivalent \(⟦s⟧_{S} = ⟦u⟧_{U}\) if and only if all the following conditions hold:
     \begin{enumerate}
@@
         \item \(s\) reject \(α\) or transition to a dead state via \(α\) if and only if \(u\) rejects \(α\) or transition to a dead state via \(α\);  
-        \item if \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', p)\), then there exists a bisimulation equivalence \({≃'}\) in \(\norm(⟨s'⟩) + \norm(⟨u'⟩)\), s.t. \(s' ≃' u'\);
+        \item if \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', p)\), then \(⟦s⟧_{S} = ⟦u⟧_{U}\);
         \item if \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', q)\), s.t. \(p ≠ q\), then both \(s'\) and \(t'\) are dead.
     \end{enumerate}
-    Notice that above condition is similar to those in~\cref{thm:recursive-construction}, except bisimulation is replaced with bisimulation \emph{equivalence}.
-\end{corollary}
-
-\begin{proof}
-    By the standard argument with normalization preserves sub-coalgebra (\Cref{thm:norm-sub-coalg}), we can obtain 
-    \(\norm(⟨s⟩) ⊑ \norm(S)\) and \(\norm(⟨u⟩) ⊑ \norm(U)\).
-    Because sub-coalgebra reflects and preserves bisimulation (\Cref{thm:sub-coalg-preserve-bisim}) and the correspondence between bisimulations and bisimulation equivalences in GKAT (\cref{thm:GKAT-bisim-iff-bisim-eq}): let \({∼_{s, u}} ⊆ \norm(⟨s⟩) × \norm(⟨u⟩)\) and \({∼} ⊆ \norm(S) × \norm(U)\) be bisimulations and \(≃\) be a bisimulation equivalence in \(\norm(⟨s⟩) + \norm(⟨u⟩)\):
-    \begin{align*}
-        & \text{Conditions in \cref{thm:recursive-construction} is satisfied} \\  
-        ⟺{}& \text{Conditions in current theorem is satisfied}\\
-        ⟺{}& ∃~ {≃}, s ≃ u 
-        ⟺ ∃~ {∼_{s, u}}, s ∼ u \\
-        ⟺{}& ∃~ {∼} , s ∼ u 
-        ⟺ ⟦s⟧_{S} = ⟦u⟧_{U}. \qedhere
+    Notice that above condition is similar to those in~\cref{thm:recursive-construction}, except bisimulation of \(s'\) and \(u'\) is replaced with trace equivalence.
+\end{theorem}
+
+\begin{proof}
+    By the standard argument with normalization preserves sub-coalgebra (\Cref{thm:norm-sub-coalg}), we can obtain for all \(s ∈ S\) and \(u ∈ U\), \(\norm(⟨s⟩) ⊑ \norm(S)\) and \(\norm(⟨u⟩) ⊑ \norm(U)\).
+    Therefore, because subcoalgebra preserve and reflect bisimulation (\Cref{thm:sub-coalg-preserve-bisim}) and the correctness of bisimulation on normalized coalgebra (\Cref{thm:norm-bisim-correctness}): for all \(s ∈ S\) and \(u ∈ U\),
+    \begin{align*}
+        & ∃~ {∼} ⊆ \norm(⟨s⟩) × \norm(⟨u⟩), s ∼ u \\
+        ⟺{}& ∃~ {∼} ⊆ \norm(S) × \norm(U), s ∼ u \\  
+        ⟺{}& ⟦s⟧_S = ⟦u⟧_U.
+    \end{align*}
+    We can instantiate the above equivalence to \(s', u'\) and \(s, u\) respectively: the instantiation to \(s', u'\) will give us the conditions in this theorem is equivalent to the ones in~\Cref{thm:recursive-construction-lemma}; and instantiation to \(s, u\) show that existence of bisimulation is equivalent to trace equivalence:
+    \begin{align*}
+        & \text{Conditions in \cref{thm:recursive-construction-lemma} hold} \\  
+        ⟺{}& \text{Conditions in the current theorem hold}\\
+        ⟺{}& ∃~ {∼} ⊆ \norm(⟨s⟩) × \norm(⟨u⟩), s ∼ u\\
+        ⟺{}& ⟦s⟧_{S} = ⟦u⟧_{U}. \qedhere
     \end{align*}
@@
 The above theorem already gives us an algorithm to recursively decide whether \(⟦s⟧_{S} = ⟦u⟧_U\), when \(⟨s⟩_S\) and \(⟨u⟩_U\) is finite.
-However, this algorithm can be further optimized in several ways: we will then derive that a dead state can never relate to live states.
-This means that when checking the bisimulation of states \(s\) and \(t\), if we already know one of them is dead, we only need to check whether the other is dead, instead of going through all the conditions in~\cref{thm:recursive-construction}.
+However, this algorithm can be further optimized: we will then derive that a dead state is only trace equivalent to other dead states.
+This means that when checking the trace-equivalence of states \(s\) and \(u\), if we already know one of them is dead, we only need to check whether the other is dead, instead of going through all the conditions in~\cref{thm:recursive-construction}.
 
 \begin{theorem}\label{thm:bisim-one-dead}
-    Given two states \(s ∈ S\) and \(t ∈ T\), if \(s\) is a dead state in \(S\), then there exists a bisimulation equivalence \(≃\) in \(S + T\) where \(s ≃ t\) if and only if \(t\) is dead.
-Similarly, for \(t ∈ T\).
+    Given two states \(s ∈ S\) and \(u ∈ U\), if \(s\) is a dead state in \(S\), then \(s\) and \(u\) is trace equivalent if and only if \(u\) is also dead.
 \end{theorem}
@@
 \begin{proof}
-    Notice because of the homomorphism \(\mathrm{inj_l}: S → S + T\) and homomorphism preserves liveness (\cref{thm:hom-preserve-liveness}), therefore \(s ∈ S\) is dead if and only if \(s ∈ S + T\) is dead; similarly for \(t ∈ T\).
-
-    If there exists a bisimulation equivalence \(≃\), s.t. \(s ≃ t\), because \(s\) is dead and bisimulation preserves liveness~\cref{thm:bisim-preserve-liveness}, then \(t\) is dead.
-
-    And if both \(t\) and \(s\) is dead, then a bisimulation between \(S\) and \(T\) can be constructed in~\cref{thm:bisim-between-dead}, which induces a bisimulation equivalence via~\Cref{thm:GKAT-bisim-iff-bisim-eq}.
+    TODO: I think we can refine \(⟹\) direction, but I am not sure...
+
+    For the \(⟸\) direction, we can construct the bisimulation as in~\cref{thm:bisim-between-dead}, which implies trace equivalence.
+    And the \(⟹\) direction, if \(s\) is dead, then by definition of normalization, it will be all rejecting, i.e. for all \(α ∈ \At\), \(δ_{\norm(S)}(s, α) = \reject\).
+    If \(s\) and \(u\) are trace equivalent, then there exists a bisimulation \({∼}: \norm(S) × \norm(U)\) s.t. \(s ∼ u\).
+    By unfolding the definition of a bisimulation, \(u\) also need to be all rejecting in \(\norm(U)\).
+    By definition of \(\norm\), for all \(α ∈ \At\), \(δ_U(u, α)\) either reject to go to a dead state, and because \(⟨u⟩_U\) is all the reachable state from \(u\), therefore 
+    \begin{align*}
+        ⟨u⟩_U = & ⋃ \{⟨u'⟩ ∣ ∃ α ∈ \At, p ∈ K, δ_U(u, α) = (u', p)\} \\
+        & ∪ \{u\}
+    \end{align*}
+    And because all of \(u'\) is dead, all of \(⟨u'⟩_U\) cannot be accepting (\Cref{thm:dead-iff-all-reachable-dead}), nor is \(u\) an accepting state, then \(⟨u⟩_U\) does not contain any accepting state, and by definition of dead state, \(u\) is dead.
 \end{proof}
@@
 
-Finally, we present our finite-trace equivalent checking algorithm as~\cref{alg:bisim}, where we search for a bisimulation equivalence \({≃}\) in \(\norm(⟨s⟩) + \norm(⟨u⟩)\), s.t.
-\(s ≃ u\), using the condition in~\Cref{thm:recursive-construction-correctness}.
-Seeking an equivalence relation allows us to organize our explored state pairs into equivalence classes, as opposed to a set of pairs, and enables us to use union-find data structures to avoid looking for two states that are already in the same equivalence class.
+Finally, we present our algorithm to decide trace-equivalence as~\cref{alg:bisim}, where we recursively check the condition in~\Cref{thm:recursive-construction}.
+Notice that trace equivalence is an equivalence relation, therefore we can organize the previously explored states into equivalent relation using union-find structure, instead of a set of state pairs.
 Specifically, we use the call \(\Call{union}{$s, u$}\) to union the equivalence class of \(s\) and \(u\), and use \(\Call{eq}{$s, u$}\) to check whether \(s\) and \(u\) is in the same equivalence class.
-In this algorithm, we first check if one of \(s\) and \(u\) is known to be dead, if so we only need to check whether the other is dead, because of~\Cref{thm:bisim-one-dead}.
-Otherwise, we will check the conditions in~\Cref{thm:recursive-construction-correctness}, which is a minor modification of the conditions in~\cref{thm:recursive-construction}.
+In this algorithm, we first check if one of \(s\) and \(u\) is known to be dead, if so we only need to check whether the other is dead, because of~\Cref{thm:bisim-one-dead}; otherwise, we will check the conditions in~\Cref{thm:recursive-construction}.
 
@@
 
-\section{Symbolic Algorithm}
-
-Given the alphabet \(K, B\), a \emph{symbolic GKAT coalgebra} \(Ŝ ≜ ⟨S, ϵ̂, δ̂⟩\) consists of a state set \(S\) and a accepting function \(ϵ̂\) and a transition function \(δ̂\):
+\section{Symbolic Coalgebra and Algorithm}
+
+Although \cref{alg:bisim} is on-the-fly, which do not require the construction of the entire coalgebra prior to running the bisimulation algorithm, it still uses GKAT coalgebra, which is known to be inefficient.
+Specifically, the transition function \(δ: S → \At_B → \{\accept, \reject\} + S × K\) requires computing the transition result for each \emph{atom}, and the number of atoms \(\At_B ≅ 2^{B}\) is exponential to the size of primitive tests in \(B\).
+
+In our symbolic GKAT coalgebra, instead of computing the behavior of each atom individually, we group them into boolean expressions, this leads to a much more space efficient coalgebra and an efficient equivalence checking algorithm with the help of off-the-shelf SAT solvers.
+Specifically, given the alphabet \(K, B\), a \emph{symbolic GKAT coalgebra} \(Ŝ ≜ (S, ϵ̂, δ̂)\) consists of a state set \(S\) and an accepting function \(ϵ̂\) and a transition function \(δ̂\):
 \begin{mathpar}
-    ϵ̂: S → 𝒫(\Bool_B), \and
-    δ̂: S → 𝒫(\Bool_B × S × K),
+    ϵ̂: S → 𝒫(\BExp_B), \and
+    δ̂: S → 𝒫(\BExp_B × S × K),
 \end{mathpar}
-where \(\Bool_B\) is the free boolean algebra over \(B\) (boolean expressions modulo boolean algebra axioms); for all states \(s ∈ S\), all the booleans are ``disjoint''; namely the conjunction of any two expression from the set \(\{ϵ̂(s)\} ∪ \{b ∣ ∃ (b, s', p) ∈ δ(s)\}\) are false.
-We will then use \(ρ̂(s): \Bool_B\) to denote the boolean expressions that contain all the atoms that the state \(s\) rejects, and \(ρ̂(s)\) can be computed as follows:
-\[ρ̂(s) ≜ ¬ ϵ̂(s) ∨ ¬ \left( ⋁_{(b, s', p) ∈ δ(s)} b \right)\]
-
-Instead of modeling each atom individually in the automata, we group them into boolean expressions, this leads to a much more space efficient automata, and enables efficient bisimulation algorithms using off-the-shelf SAT solvers.
-
-With the above intuition in mind, a symbolic GKAT coalgebra \(Ŝ ≜ ⟨S, ϵ̂, δ̂⟩\) can be lowered into a GKAT coalgebra \(⟨S, δ⟩\) in the following manner:
+where for all states \(s ∈ S\), all the boolean expressions are ``disjoint''; namely the conjunction of any two expressions from the set \(\{ϵ̂(s)\} ∪ \{b ∣ ∃ (b, s', p) ∈ δ(s)\}\) is equivalent to \(0\) under boolean algebra.
+
+Notice that we call \(ϵ̂\) the accepting function and \(δ̂\) the transition function; intuitively, a state \(s\) accepts an atom \(α\) if and only if there exists a \(b ∈ ϵ̂(s)\), s.t.
+\(α ≤ b\); similarly, the state \(s\) transitions to \(s'\) via atom \(α\) while executing \(p\) if and only if there exists \((b, s', p) ∈ δ̂(s)\) and \(α ≤ b\).
+With the above intuition in mind, a symbolic GKAT coalgebra \(Ŝ ≜ (S, ϵ̂, δ̂)\) can be lowered into a GKAT coalgebra \((S, δ)\) in the following manner:
 \begin{align}\label{cons:lowering}
 δ(s, α) ≜ \begin{cases}
-    \accept & ∃ b ∈ ϵ̂(s), α ≤ b \\  
-    (s', p) & ∃ b ∈ \Bool_B, α ≤ b \text{ and } δ(s, b) = (s', p) \\  
+    \accept & ∃ b ∈ ϵ̂(s), α ≤ b \\[5px]
+    (s', p) & 
+        \begin{aligned}
+            & ∃ b ∈ \BExp_B, α ≤ b \\
+            & \text{ and } δ(s, b) = (s', p)  
+        \end{aligned}\\[5px]
     \reject & \text{otherwise}
@@
 This is well-defined, i.e. no more than one clause can be satisfied precisely because the boolean expressions appear in \(ϵ̂\) and \(δ̂\) are disjoint.
-The trace semantics of a GKAT coalgebra \(⟨S, ϵ̂, δ̂⟩\) is then defined as the trace semantics of its lowering \(⟨S, δ⟩\).
-
-\begin{remark}[Canonicity]
-    Notice that symbolic GKAT coalgebra is not canonical, i.e. there exists two different symbolic GKAT colagebra with the same lowering, consider the state set \(S ≜ \{*\}\):
-    \[{δ̂}₁(*) ≜ \{b ↦ (*, p), ¬ b ↦ (*, p)\} \qquad 
-    {δ̂}₂(*) ≜ \{⊤ ↦ (*, p)\},\] 
-    and both \(ϵ̂\) will return constant \(⊥\).
-    These two symbolic GKAT coalgebra obviously have the same lowering hence behavior, yet, they are different.
+We usually use \(S\) to denote the lowering of \(Ŝ\).
+
+We will then use \(ρ̂(s): \BExp_B\) to represent the boolean expressions that contain all the atoms that the state \(s\) rejects, and \(ρ̂(s)\) can be computed as follows:
+\begin{align*}
+    ρ̂(s) ≜ ⋀ \{\overline{b} ∣{}
+        & ∃ s' ∈ S, p ∈ K, (b, s', p) ∈ δ̂(s) \\
+        & \text{ or } b ∈ ϵ̂(s)\}.
+\end{align*}
+The trace semantics of a GKAT coalgebra \((S, ϵ̂, δ̂)\) is then defined as the trace semantics of its lowering \((S, δ)\).
+
+\begin{remark}[Canonicity]\label{rem:canonicity}
+    Notice that symbolic GKAT coalgebra is not canonical, i.e. there exists two different symbolic GKAT colagebra with the same lowering, consider the state set \(S ≜ \{s\}\):
+    \begin{align*}
+        {δ̂}₁(s) & ≜ \{b ↦ (s, p), \overline{b} ↦ (s, p)\} \\
+        {δ̂}₂(s) & ≜ \{⊤ ↦ (s, p)\},
+    \end{align*} 
+    and both \(ϵ̂₁, ϵ̂₂\) will return constant \(0\).
+    These two symbolic GKAT coalgebra \(Ŝ₁ ≜ (S, δ̂₁, ϵ̂₁)\) and \(Ŝ₂ ≜ (S, δ̂₂, ϵ̂₂)\) have the same lowering and semantics, yet, they are different.
     There are other symbolic representation that will satisfy canonicity, yet we opt to use our current representation for ease of construction and computational efficiency.
@@
 \begin{proof}
-
-\end{proof}
+    Since \(Ŝ\) and \(Û\) have the same states as their lowering, therefore \(h: S → U\) is indeed a function, then we only need to verify the homomorphism condition on \(h\).
+    TODO: finish.
+\end{proof}
+
+The functoriality states that a homomorphism on two symbolic coalgebras implies a homomorphism of on their lowing; similarly, a bisimulation on symbolic GKAT coalgebra induces a bisimulation on their lowering.
+However, the converse is not true, precisely because of the canonicity problem noted in~\Cref{rem:canonicity}: take the \(Ŝ₁\) and \(Ŝ₂\) in~\Cref{rem:canonicity}, because they have the same lowering, therefore the identity homomorphism is a homomorphism on their lowerings, yet there is no homomorphism from \(Ŝ₁\) to \(Ŝ₂\).
 
@@
 
-\begin{theorem}[Symbolic Inductive Construction]\label{thm:symb-recursive-construction}
-    Given two symbolic GKAT coalgebra \(Ŝ = ⟨S, ϵ̂_S, δ̂_S⟩\) and \(T̂ = ⟨T, ϵ̂_T, δ̂_T⟩\) and two states \(s ∈ S\) and \(t ∈ T\), there exists a normalized bisimulation on the lowered coalgebra \({∼} ⊆ S × T\) s.t.
-\(s ∼ t\) if and only if all the following holds:
+\begin{theorem}[Symbolic Recursive Construction]\label{thm:symb-recursive-construction}
+    Given two symbolic GKAT coalgebra \(Ŝ = (S, ϵ̂_S, δ̂_S)\) and \(Û = (U, ϵ̂_U, δ̂_U)\) and two states \(s ∈ S\) and \(u ∈ U\), \(s\) and \(u\) are trace equivalent \(⟦s⟧_{Ŝ} = ⟦u⟧_{Û}\), if and only if all the following holds:
     \begin{itemize}
-        \item \(⋁ ϵ̂_S(s) ≡ ⋁ ϵ̂_T(t)\);
-        \item for all \((b, s', p) ∈ δ̂_S(s)\) and \((c, t', q) ∈ δ̂_T(t)\), if \(b ∧ c ≢ 0\) and \(p = q\) then here exists a normalized bisimulation \({∼_{s',t'}} ⊆ S × T\) s.t. \(s' ∼_{s',t'} t'\);
-        \item for all \((b, s', p) ∈ δ̂_S(s)\) and \((c, t', q) ∈ δ̂_T(t)\), if \(b ∧ c ≢ 0\) and \(p ≠ q\) then both \(s'\) and \(t'\) is dead;  
-        \item for all \((b, s', p) ∈ δ̂_S(s)\) and \(c ∈ ρ̂_T(t)\), if \(b ∧ c ≢ 0\), then \(s'\) is dead;
-        \item for all \(b ∈ ρ̂_S(s)\) and \((c, t', q) ∈ δ̂_T(t)\), if \(b ∧ c ≢ 0\), then \(t'\) is dead;
+        \item \(⋁ ϵ̂_S(s) ≡ ⋁ ϵ̂_U(u)\); 
+        \item for all \((b, s', p) ∈ δ̂_S(s)\) and \(c ∈ ρ̂_U(u)\), if \(b ∧ c ≢ 0\), then \(s'\) is dead;
+        \item for all \(b ∈ ρ̂_S(s)\) and \((c, u', q) ∈ δ̂_U(u)\), if \(b ∧ c ≢ 0\), then \(u'\) is dead;
+        \item for all \((b, s', p) ∈ δ̂_S(s)\) and \((c, u', q) ∈ δ̂_U(u)\), if \(b ∧ c ≢ 0\) and \(p ≠ q\) then both \(s'\) and \(u'\) is dead; 
+        \item for all \((b, s', p) ∈ δ̂_S(s)\) and \((c, u', q) ∈ δ̂_U(u)\), if \(b ∧ c ≢ 0\) and \(p = q\) then \(⟦s'⟧_{Ŝ} = ⟦u'⟧_{Û}\).
     \end{itemize}
@@
                 & ⋁ ϵ̂_S(s) ≡ ⋁ ϵ̂_T(t) \mathrel{\&\!\&} \\  
-                & \forall (b, s', p) ∈ δ̂_S(s), (c, t', q) ∈ δ̂_T(t), (b ∧ c) ≢ ⊥ ⟹ 
+                & ∀ (b, s', p) ∈ δ̂_S(s), (c, t', q) ∈ δ̂_T(t), (b ∧ c) ≢ 0 ⟹ 
                 \begin{cases}
@@
                 \end{cases} \mathrel{\&\!\&}\\
-                & \forall (b, s', p) ∈ δ̂_S(s), c ∈ ρ̂_T(t), (b ∧ c) ≢ ⊥ ⟹ \text{\Call{IsDead$_S$}{$s'$}} \mathrel{\&\!\&}\\
-                & \forall b ∈ ρ̂_S(s), (c, t', q) ∈ δ̂_T(t), (b ∧ c) ≢ ⊥ ⟹ \text{\Call{IsDead$_T$}{$t'$}}
+                & ∀ (b, s', p) ∈ δ̂_S(s), c ∈ ρ̂_T(t), (b ∧ c) ≢ 0 ⟹ \text{\Call{IsDead$_S$}{$s'$}} \mathrel{\&\!\&}\\
+                & ∀ b ∈ ρ̂_S(s), (c, t', q) ∈ δ̂_T(t), (b ∧ c) ≢ 0 ⟹ \text{\Call{IsDead$_T$}{$t'$}}
             \end{aligned}\)
@@
 
-\section{Construction of Symbolic GKAT Automata}
+\section{Symbolic GKAT Coalgebra Construction}
 
@@
 With these definitions in mind, we can define symbolic Thompson's construction inductively as in~\Cref{tab:symb-thomposon-construction}, where we let \((S₁, ϵ̂₁, δ̂₁)\) and \((S₂, ϵ̂₂, δ̂₂)\) to be result of Thompson's construction for \(e₁\) and \(e₂\) respectively.
-The \(S₁ + S₂\) is the disjoint union of \(S₁\) and \(S₂\), and for any two transition dynamics \(δ₁(s₁): 𝒫(\Bool × S₁ × K)\) and \(δ₂(s₂): 𝒫(\Bool × S₂ × K)\), then \(δ₁^*(s₁) + δ₂^*(s₂): 𝒫(\Bool × (S₁ + S₂) × K)\) is the bifunctorial lift via \(+\).
+The \(S₁ + S₂\) is the disjoint union of \(S₁\) and \(S₂\), and for any two transition dynamics \(δ₁(s₁): 𝒫(\BExp × S₁ × K)\) and \(δ₂(s₂): 𝒫(\BExp × S₂ × K)\), then \(δ₁^*(s₁) + δ₂^*(s₂): 𝒫(\BExp × (S₁ + S₂) × K)\) is the bifunctorial lift via \(+\).
 

### 2025-01-06 main.tex
update

@@
 
-\title{Symbolic On-the-fly Algorithms for GKAT Equivalences
+\title{On-the-fly Algorithms for GKAT Equivalences
 \thanks{Identify applicable funding agency here. If none, delete this.}
@@
         \If {\Call{eq}{$s, u$}} {\Return true} \EndIf
-        \If {\Call{knownDead\(_S\)}{s}} 
+        \If {\Call{knownDead\(_S\)}{$s$}} 
             \State\Return \Call{isDead\(_U\)}{$u$}
         \EndIf
-        \If {\Call{knownDead\(_U\)}{u}} 
+        \If {\Call{knownDead\(_U\)}{$u$}} 
             \State\Return \Call{isDead\(_S\)}{$s$}
@@
 \begin{algorithm*}
-    \caption{Symbolic On-the-fly bisimulation algorithm}\label{alg:symb-bisim}
+    \caption{Symbolic On-the-fly Bisimulation Algorithm}\label{alg:symb-bisim}
     \begin{algorithmic}
-        \Function{equiv}{$s ∈ S, t ∈ T$}
-        \If {\Call{eq}{$s, t$}} {\Return true}
-        \ElsIf {\(s ∈\) deadStates\(_S\)} {\Return \Call{isDead\(_T\)}{$t$}} 
-        \ElsIf {\(t ∈\) deadStates\(_T\)} {\Return \Call{isDead\(_S\)}{$s$}} 
+        \Function{equiv}{$s ∈ S, u ∈ U$}
+        \If {\Call{eq}{$s, u$}} {\Return true}
+        \ElsIf {\Call{knownDead\(_S\)}{$s$}} {\Return \Call{isDead\(_U\)}{$u$}} 
+        \ElsIf {\Call{knownDead\(_U\)}{$u$}} {\Return \Call{isDead\(_S\)}{$s$}} 
         \Else {}
@@
             \begin{aligned}
-                & ⋁ ϵ̂_S(s) ≡ ⋁ ϵ̂_T(t) \mathrel{\&\!\&} \\  
-                & ∀ (b, s', p) ∈ δ̂_S(s), (c, t', q) ∈ δ̂_T(t), (b ∧ c) ≢ 0 ⟹ 
+                & ⋁ ϵ̂_S(s) ≡ ⋁ ϵ̂_U(u) \mathrel{\&\!\&} \\  
+                & ∀ (b, s', p) ∈ δ̂_S(s), (c, u', q) ∈ δ̂_U(u), (b ∧ c) ≢ 0 ⟹ 
                 \begin{cases}
-                    \text{\Call{IsDead$_S$}{$s$}} ∧ \text{\Call{IsDead$_T$}{$t$}} & \text{if \(p ≠ q\)} \\
-                    \text{\Call{Union}{$s$, $t$}}; \text{\Call{Equiv}{$s', t'$}} & \text{if \(p = q\)}
+                    \text{\Call{isDead$_S$}{$s$}} ∧ \text{\Call{isDead$_U$}{$u$}} & \text{if \(p ≠ q\)} \\
+                    \text{\Call{Union}{$s$, $u$}}; \text{\Call{Equiv}{$s', u'$}} & \text{if \(p = q\)}
                 \end{cases} \mathrel{\&\!\&}\\
-                & ∀ (b, s', p) ∈ δ̂_S(s), c ∈ ρ̂_T(t), (b ∧ c) ≢ 0 ⟹ \text{\Call{IsDead$_S$}{$s'$}} \mathrel{\&\!\&}\\
-                & ∀ b ∈ ρ̂_S(s), (c, t', q) ∈ δ̂_T(t), (b ∧ c) ≢ 0 ⟹ \text{\Call{IsDead$_T$}{$t'$}}
+                & ∀ (b, s', p) ∈ δ̂_S(s), c ∈ ρ̂_U(u), (b ∧ c) ≢ 0 ⟹ \text{\Call{IsDead$_S$}{$s'$}} \mathrel{\&\!\&}\\
+                & ∀ b ∈ ρ̂_S(s), (c, u', q) ∈ δ̂_U(u), (b ∧ c) ≢ 0 ⟹ \text{\Call{IsDead$_U$}{$u'$}}
             \end{aligned}\)
@@
     \end{mathpar}
-    \caption{Symbolic Derivative of GKAT Automata.}\label{fig:derivatives-rules}
+    \caption{Symbolic Derivative}\label{fig:derivatives-rules}
 \end{figure*}
@@
 The symbolic derivative coalgebra \(D̂\), with expressions as states, is the least symbolic GKAT coalgebra (ordered by point-wise subset ordering on \(ϵ̂\) and \(δ̂\)) that satisfy the rules in~\Cref{fig:derivatives-rules}.
-Notice that the rules listed on~\Cref{fig:derivatives-rules} is very close to that of Schmid et al.~\cite{schmid_GuardedKleeneAlgebra_2021}.
+Notice that the rules listed on~\Cref{fig:derivatives-rules} is very close to that of~\citeauthor{schmid_GuardedKleeneAlgebra_2021}~\cite{schmid_GuardedKleeneAlgebra_2021}.
 This is no coincidence, as our definition exactly lowers to the definition of theirs.
 This fact can be proven by case analysis on the shape of the source expression, and forms a basis on our correctness argument.
-
 \begin{theorem}[Correctness]
     The lowering of \(D̂\) is exactly the derivative defined by Schmid et. al.~\cite{schmid_GuardedKleeneAlgebra_2021}.
+    Therefore, the semantics of the expression is equal to the semantics \(⟦e⟧ = ⟦e⟧_{D̂}\).
     TODO: unfold the statement.
@@
 
-% \begin{table*}
+% For the ease of implementation, we will concretely define the accepting function \(ϵ̂_D\) and transition function \(δ̂_D\) of the derivative coalgebra \(D̂\).
+% Prior to that we will introduce an operation to ``guarded'' a function by a set of boolean \(B\), denoted \(⟨B|\), and it can be applied to both the accepting dynamic \(ϵ̂(s)\) and transition dynamic \(δ̂(s)\):
+% \begin{align*}
+%     ⟨B|ϵ̂(s) & ≜ \{b ∧ c ∣ b ∈ B, c ∈ ϵ(s)\}; \\
+%     ⟨B|δ̂(s) & ≜ \{(b ∧ c, s', p) ∣ b ∈ B, (c, s', p) ∈ δ(s)\}.
+% \end{align*}
+
+% \begin{figure*}
 %     \centering
-%     \begin{tabular}{c||c|c|c|c|c|c}
-%         Exp & \(S\) 
-%         & \(ϵ^*\) & \(δ^*\) 
-%         & \(ϵ̂(s)\) & \(δ̂(s)\) \\\hline
-%         \(b\) & \(∅\) 
-%         & \{b\} & \(∅\) 
-%         & N/A & N/A \\  
-%         \(p\) & \(\{*\}\) 
-%         & \(∅\) & \(\{(1, *, p)\}\) 
-%         & \(\{1\}\) & \(∅\)\\  
-%         \(e₁ +_b e₂\) & \(S₁ + S₂\) &
-%         \(⟨\{b\}|~ϵ₁^* ∪ ⟨\{\overline{b}\}|~ϵ₂^*\) &
-%         \(⟨\{b\}|~δ₁^* + ⟨\{\overline{b}\}|~δ₂^*\) & 
-%         \(\begin{cases}
-%             ϵ̂₁(s) & s ∈ S₁\\
-%             ϵ̂₂(s) & s ∈ S₂\\
-%         \end{cases}\) & 
-%         \(\begin{cases}
-%             δ̂₁(s) & s ∈ S₁\\
-%             δ̂₂(s) & s ∈ S₂\\
-%         \end{cases}\) \\  
-%         \(e₁ ; e₂\) & \(S₁ + S₂\) & 
-%         \(⟨ϵ₁^*|~ϵ₂^*\) & \(δ₁^* + ⟨ϵ₁^*|~δ₂^*\) & 
-%         \(\begin{cases}
-%             ⟨ϵ̂₁(s)| ~ ϵ₂^*& s ∈ S₁ \\  
-%             ϵ̂₂(s) & s ∈ S₂
-%         \end{cases}\)& 
-%         \(\begin{cases}
-%             δ̂₁(s) ∪ ⟨ϵ̂(s)|~δ₂^* & s ∈ S₁ \\  
-%             δ̂2(s) & s ∈ S₂
-%         \end{cases}\) \\  
-%         \(e₁^{(b)}\) & \(S₁\) & 
-%         \(\{\overline{b}\}\) & \(⟨\{b\}|~δ₁^*\) & 
-%         \(⟨\{\overline{b}\}|~ϵ̂1(s)\) & 
-%         \( ⟨\{b\}|⟨ϵ̂1(s)|~δ₁^* ∪ δ₁(s)\)
-%     \end{tabular}
-%     \caption{Symbolic Thompson's Construction}\label{tab:symb-thomposon-construction}
-% \end{table*}
+%     \begin{align*}
+%         δ̂(b) & ≜ ∅ & ϵ̂(b) & ≜ \{b\} \\  
+%         δ̂(p) & ≜ \{(1, 1, p)\} & ϵ̂(p) ≜ ∅ \\  
+%         δ̂(e +_b f) & ≜ ⟨\{b\}| δ̂(e) ∪ ⟨\{\overline{b}\}| δ̂(f) & 
+%         ϵ̂(e +_b f) & ≜ ⟨\{b\}| ϵ̂(e) ∪ ⟨\{\overline{b}\}| ϵ̂(f)
+%     \end{align*}
+% \end{figure*}
 
@@
 
-Another way to construct an automaton is via Thompson's construction, we lift the original construction to the symbolic setting.
+Another way to construct a coalgebra from a GKAT expression is via Thompson's construction, we lift the original construction to the symbolic setting.
 A common expression to construct is a guard operation, denoted by \(⟨B|\), where \(B\) is a set of boolean expressions.
@@
 \end{align*}
-Notably, besides guarding transition and acceptance with different conditions, like in if statements, the guard expression can be used to simulate uniform continuation.
-Specifically, we can use \(⟨ϵ̂(s)|~δ(s)\) to connecting all the accepting state of \(s\) to the dynamic \(δ(s)\).
+Notably, besides guarding transition and acceptance with different conditions in if statements and while loops, the guard operator can also be used to simulate uniform continuation.
+Specifically, we can use \(⟨ϵ̂(s)|δ(s')\) to connecting all the accepting state of \(s\) to the dynamic \(δ(s')\).
 
 With these definitions in mind, we can define symbolic Thompson's construction inductively as in~\Cref{tab:symb-thomposon-construction}, where we let \((S₁, ϵ̂₁, δ̂₁)\) and \((S₂, ϵ̂₂, δ̂₂)\) to be result of Thompson's construction for \(e₁\) and \(e₂\) respectively.
-The \(S₁ + S₂\) is the disjoint union of \(S₁\) and \(S₂\), and for any two transition dynamics \(δ₁(s₁): 𝒫(\BExp × S₁ × K)\) and \(δ₂(s₂): 𝒫(\BExp × S₂ × K)\), then \(δ₁^*(s₁) + δ₂^*(s₂): 𝒫(\BExp × (S₁ + S₂) × K)\) is the bifunctorial lift via \(+\).
+The \(S₁ + S₂\) is the disjoint union of \(S₁\) and \(S₂\), and for any two transition dynamics \(δ₁(s₁): 𝒫(\BExp × S₁ × K)\) and \(δ₂(s₂): 𝒫(\BExp × S₂ × K)\), then \(δ₁(s₁) + δ₂(s₂): 𝒫(\BExp × (S₁ + S₂) × K)\) is defined as follows:
+\begin{align*}
+    δ₁(s₁) & + δ₂(s₂) ≜ \\
+        & \{(b, \mathrm{inj}ₗ(s₁'), p) ∣ (b, s₁', p) ∈ δ₁(s₁)\} ∪ \\
+        & \{(b, \mathrm{inj}ᵣ(s₂'), p) ∣ (b, s₂', p) ∈ δ₂(s₂)\}
+\end{align*}
+where \(\mathrm{inj}ₗ: S₁ → S₁ + S₂\) and \(\mathrm{inj}ᵣ: S₂ → S₁ + S₂\) are the canonical left/right injection of the coproduct.
 
 One notable difference between the original construction~\cite{smolka_GuardedKleeneAlgebra_2020} and our construction is that we use a start state \(s^* ∈ S\), instead of a start dynamics (or pseudo-state).
-This choice will make the proof slightly easier. 
+This choice will make the proof of~\Cref{thm:hom-thompson-derivative} slightly easier. 
 However, in~\Cref{sec:optimization-implementation}, we will explain that our implementation uses start dynamics instead of start state, to avoid unnecessary lookups and unreachable states.
@@
 We would like to explore several desirable theoretical properties of both derivatives and Thompson's construction.
-Specifically, the correctness, i.e.
-the semantics of the ``start state'' the both construction have the same preserves the trace semantics of the expression; finiteness, i.e.
-the coalgebra generated is always finite, which means that our equivalence algorithm will eventually terminate; and finally, how does the number of reachable state relate to the size of the expression, so that we can estimate the complexity of the equivalence checking algorithm.
-Turns out all of these questions can be answered by a connection by a homomorphism from symbolic Thompson's construction to the symbolic derivatives.
+Specifically, the \emph{correctness}, i.e.
+the semantics of the ``start state'' the both construction have the same preserves the trace semantics of the expression; \emph{finiteness}, i.e.
+the coalgebra generated is always finite, which means that our equivalence algorithm will eventually terminate; and finally, \emph{complexity}, the relationship between the number of reachable states and the size of the input expression, which serves as an estimated complexity of our equivalence checking algorithm.
+Turns out all of these questions can be answered by a homomorphism from symbolic Thompson's construction to the symbolic derivatives.
 
@@
     \begin{align*}
-        & s \transAcc{Ŝ_{e}}{c} \text{ iff }
-        s \transAcc{Ŝ_{e₁}}{c} \text{ iff }
-        h₁(s) \transAcc{D̂}{c} \text{ iff }
-        h(s) \transAcc{D̂}{c} \\ 
-        & s \transvia{c ∣ p}_{Ŝ_{e}} s' \text{ iff }
-        s \transvia{c ∣ p}_{Ŝ_{e₁}} s' \text{ iff }
-        h₁(s) \transvia{c ∣ p}_{D̂} h₁(s') \text{ iff }
-        h(s) \transvia{c ∣ p}_{D̂} h(s')  
+        & s \transAcc{Ŝ_{e}}{c} 
+        \text{ iff } s \transAcc{Ŝ_{e₁}}{c} \\*
+        & \text{ iff } h₁(s) \transAcc{D̂}{c} 
+        \text{ iff } h(s) \transAcc{D̂}{c}; \\ 
+        & s \transvia{c ∣ p}_{Ŝ_{e}} s' 
+        \text{ iff } s \transvia{c ∣ p}_{Ŝ_{e₁}} s' \\*
+        & \text{ iff } h₁(s) \transvia{c ∣ p}_{D̂} h₁(s') 
+        \text{ iff } h(s) \transvia{c ∣ p}_{D̂} h(s'). 
     \end{align*}
@@
         (∃ a, b ∧ a = c \text{ and } 
-        s₁^* \transAcc{Ŝ_{e₁}}{a})
-        \text{ or } 
+        s₁^* \transAcc{Ŝ_{e₁}}{a}) \\*
+        & \text{ or } 
         (∃ a, \overline{b} ∧ a = c \text{ and } 
@@
         (∃ a, b ∧ a = c \text{ and } 
-        h₁(s₁^*) \transAcc{D̂}{a})
-        \text{ or } 
+        h₁(s₁^*) \transAcc{D̂}{a}) \\*
+        & \text{ or } 
         (∃ a, \overline{b} ∧ a = c \text{ and } 
@@
         (∃ a, b ∧ a = c \text{ and } 
-        e₁ \transAcc{D̂}{a})
-        \text{ or } 
+        e₁ \transAcc{D̂}{a}) \\*
+        & \text{ or } 
         (∃ a, \overline{b} ∧ a = c \text{ and } 
@@
         (∃ a, b ∧ a = c \text{ and } 
-        s₁^* \transvia{a ∣ p}_{Ŝ_{e₁}} s')
-        \text{ or } 
+        s₁^* \transvia{a ∣ p}_{Ŝ_{e₁}} s') \\*
+        & \text{ or } 
         (∃ a, \overline{b} ∧ a = c \text{ and } 
@@
         (∃ a, b ∧ a = c \text{ and } 
-        h₁(s₁^*) \transvia{a ∣ p}_{D̂} h(s'))
-        \text{ or } 
+        h₁(s₁^*) \transvia{a ∣ p}_{D̂} h(s')) \\*
+        & \text{ or } 
         (∃ a, \overline{b} ∧ a = c \text{ and } 
@@
         (∃ a, b ∧ a = c \text{ and } 
-        e₁ \transvia{a ∣ p}_{D̂} h(s'))
-        \text{ or } 
+        e₁ \transvia{a ∣ p}_{D̂} h(s')) \\*
+        & \text{ or } 
         (∃ a, \overline{b} ∧ a = c \text{ and } 
@@
         \text{iff }& e₁ +_b e₂ \transvia{a ∣ p}_{D̂} h(s') \\ 
-        \text{iff }& h(s^*) \transvia{a ∣ p}_{D̂} h(s')
+        \text{iff }& h(s^*) \transvia{a ∣ p}_{D̂} h(s').
     \end{align*}
@@
     \begin{align*}
-        s \transAcc{Ŝ_{e}}{c} 
+        & s \transAcc{Ŝ_{e}}{c} \\
         \text{ iff } & 
@@
         \text{ iff } h(s) \transAcc{D̂}{c}.\\[5px]
-        s \transvia{c ∣ p}_{S_{e}} s'
+        & s \transvia{c ∣ p}_{S_{e}} s' \\
         \text{ iff } & 
-            (s \transvia{c ∣ p}_{Ŝ_{e₁}} s')
-            \text{ or }
+            \begin{aligned}[t]
                 (∃ a, b, a ∧ b = c
-                \text{ and }
-                s \transAcc{Ŝ_{e₁}}{a}
-                \text{ and }
-                s₂^* \transvia{b ∣ p}_{Ŝ_{e₂}} s') \\  
+                & \text{ and }
+                s \transAcc{Ŝ_{e₁}}{a} \\*
+                & \text{ and }
+                s₂^* \transvia{b ∣ p}_{Ŝ_{e₂}} s') 
+            \end{aligned} \\*
+            & \text{ or }
+            (s \transvia{c ∣ p}_{Ŝ_{e₁}} s') \\  
         \text{ iff } & 
-            (h₁(s) \transvia{c ∣ p}_{D̂} h₁(s')) 
-            \text{ or }
+            \begin{aligned}[t]
                 (∃ a, b, a ∧ b = c
-                \text{ and }
-                h₁(s) \transAcc{D̂}{a}
-                \text{ and }
-                e₂ \transvia{b ∣ p}_{D̂} h₂(s')) \\
+                & \text{ and }
+                h₁(s) \transAcc{D̂}{a} \\*
+                &\text{ and }
+                e₂ \transvia{b ∣ p}_{D̂} h₂(s'))
+            \end{aligned} \\*
+            & \text{ or }
+            (h₁(s) \transvia{c ∣ p}_{D̂} h₁(s')) \\
         \text{ iff } & h₁(s) \transvia{c ∣ p} h(s') 
@@
     \begin{align*}
-        & s \transAcc{Ŝ_{e}}{c} \text{ iff }
-        s \transAcc{Ŝ_{e₂}}{c} \text{ iff }
-        h₂(s) \transAcc{D̂}{c} \text{ iff }
-        h(s) \transAcc{D̂}{c}, \\ 
-        & s \transvia{c ∣ p}_{Ŝ_{e}} s' \text{ iff }
-        s \transvia{c ∣ p}_{Ŝ_{e₂}} s' \text{ iff }
-        h₂(s) \transvia{c ∣ p}_{D̂} h₂(s') \text{ iff }
-        h(s) \transvia{c ∣ p}_{D̂} h(s').
+        & s \transAcc{Ŝ_{e}}{c} 
+        \text{ iff } s \transAcc{Ŝ_{e₂}}{c} \\*
+        & \text{ iff } h₂(s) \transAcc{D̂}{c} 
+        \text{ iff } h(s) \transAcc{D̂}{c}, \\ 
+        & s \transvia{c ∣ p}_{Ŝ_{e}} s' 
+        \text{ iff } s \transvia{c ∣ p}_{Ŝ_{e₂}} s' \\*
+        & \text{ iff } h₂(s) \transvia{c ∣ p}_{D̂} h₂(s') 
+        \text{ iff } h(s) \transvia{c ∣ p}_{D̂} h(s').
     \end{align*}
@@
         (s^* \transAcc{Ŝ_e}{c})
-        \text{ iff } & (s^* \transAcc{Ŝ_e}{c} \text{ and } c = \overline{b})
-        \text{ iff } ({e₁}^{(b)} \transAcc{D̂}{c} \text{ and } c = \overline{b})
-        \text{ iff } (h(s^*) \transAcc{D̂}{c}) \\
+        \text{ iff } & (s^* \transAcc{Ŝ_e}{c} \text{ and } c = \overline{b}) \\*
+        \text{ iff } & ({e₁}^{(b)} \transAcc{D̂}{c} \text{ and } c = \overline{b}) \\*
+        \text{ iff } & (h(s^*) \transAcc{D̂}{c}); \\
         (s^* \transvia{c ∣ p}_{Ŝ_e} s')
@@
         \text{ iff } & 
-            (s \transvia{c ∣ p}_{Ŝ_{e₁}} s' 
-            \text{ or } 
-            ∃ a₁, a₂, 
-                b ∧ a₁ ∧ a₂ = c 
-                \text{ and } 
-                s \transAcc{Ŝ_{e₁}}{a₁} 
-                \text{ and } 
-                s₁^* \transvia{a₂ ∣ p}_{Ŝ_{e₁}} s') \\
+            (s \transvia{c ∣ p}_{Ŝ_{e₁}} s' \\
+            & \text{ or } 
+            \begin{aligned}[t]
+                ∃ a₁, a₂, {}
+                & b ∧ a₁ ∧ a₂ = c \\*
+                & \text{ and } 
+                s \transAcc{Ŝ_{e₁}}{a₁} \\*
+                & \text{ and } 
+                s₁^* \transvia{a₂ ∣ p}_{Ŝ_{e₁}} s')
+            \end{aligned} \\
         \text{ iff } & 
-            (h₁(s) \transvia{c ∣ p}_{D̂} h₁(s') 
-            \text{ or } 
-            ∃ a₁, a₂, b ∧ a₁ ∧ a₂ = c 
-                \text{ and } 
-                h₁(s) \transAcc{D̂}{a₁} 
-                \text{ and } 
-                e₁ \transvia{a₂ ∣ p}_{D̂} h₁(s'))\\
+            (h₁(s) \transvia{c ∣ p}_{D̂} h₁(s') \\*
+            & \text{ or } 
+            \begin{aligned}[t]
+                ∃ a₁, a₂, &{} b ∧ a₁ ∧ a₂ = c \\*
+                & \text{ and } 
+                h₁(s) \transAcc{D̂}{a₁} \\*
+                & \text{ and } 
+                e₁ \transvia{a₂ ∣ p}_{D̂} h₁(s')) 
+            \end{aligned}\\
         \text{ iff } & 
-            (h₁(s) \transvia{c ∣ p}_{D̂} h₁(s') 
-            \text{ or } 
-            ∃ a₁, a₂, b ∧ a₁ ∧ a₂ = c 
-                \text{ and } 
-                h₁(s) \transAcc{D̂}{a₁} 
-                \text{ and } 
+            (h₁(s) \transvia{c ∣ p}_{D̂} h₁(s')  \\
+            & \text{ or } 
+            \begin{aligned}[t]
+                ∃ a₁, a₂, &{} b ∧ a₁ ∧ a₂ = c \\
+                & \text{ and } 
+                h₁(s) \transAcc{D̂}{a₁} \\
+                &{} \text{ and } 
                 e₁^{(b)} \transvia{b ∧ a₂ ∣ p}_{D̂} h₁(s');e₁^{(b)}) \\ 
+            \end{aligned}\\
         \text{ iff } &
-            (h₁(s); e₁^{(b)} \transvia{c ∣ p}_{D̂} h₁(s'); e₁^{(b)}) 
-        \text{ iff } h(s) \transvia{c ∣ p}_{D̂} h(s')
+            (h₁(s); e₁^{(b)} \transvia{c ∣ p}_{D̂} h₁(s'); e₁^{(b)}) \\
+        \text{ iff } & h(s) \transvia{c ∣ p}_{D̂} h(s').
         \qedhere
@@
 \begin{corollary}[Correctness]
-    Given any expression \(e\) and its Thompson's coalgebra \(Ŝ_{e}\) with a start state \(s^* ∈ Ŝ_{e}\), then the semantics of the start state is equivalent to the semantics of \(e\): \(⟦s^*⟧_{Ŝ_{e}} = ⟦e⟧.\)
+    Given any expression \(e\) and its Thompson's coalgebra \(Ŝ_{e}\) with the start state \(s^* ∈ Ŝ_{e}\), then the semantics of the start state is equivalent to the semantics of \(e\): \(⟦s^*⟧_{Ŝ_{e}} = ⟦e⟧.\)
 \end{corollary}
@@
 \begin{proof}
-    We define \(h'\) to be point-wise equal to \(h\), i.e. \(h'(s) ≜ h(s)\). 
-    Then we need to show that \(h'\) is well-defined ans surjective, which is a consequence of homomorphic image preserves principle sub-coalgebra (\Cref{thm:homo-img-preserve-principle-sub-coalg}): \(h(⟨s^*⟩_{Ŝ_{e}}) = ⟨h(s)⟩_{D̂} = ⟨e⟩_{D̂}.\)
-    In other words, the image of \(h\) on \(⟨s^*⟩_{Ŝ_{e}}\) is equal to \(⟨e⟩_{D̂}\); thus, because \(h'\) is point-wise equal to \(h\) and is defined on \(⟨s^*⟩_{Ŝ_{e}}\), the range of \(h'\) contains its codomain \(⟨e⟩_{D̂}\), showing that \(h'\) is surjective.
-\end{proof}
-
-An important consequence of~\cref{thm:suj-hom-thompson-derivative} is that \(⟨s^*⟩_{Ŝ_{e}}\) will have no less state than \(⟨e⟩_{D̂}\).
-However, it is important to notice that this does not mean running bisimulation algorithm on
+    We define \(h'\) to be point-wise equal to \(h\), i.e. \(h'(s) ≜ h(s)\), i.e. \(h'\) is \(h\) restricted on the domain \(⟨s^*⟩_{Ŝ_{e}}\). 
+    We need to show that \(h'\) is well-defined and surjective, which is a consequence of homomorphic image preserves principle sub-coalgebra (\Cref{thm:homo-img-preserve-principle-sub-coalg}): \(h(⟨s^*⟩_{Ŝ_{e}}) = ⟨h(s)⟩_{D̂} = ⟨e⟩_{D̂}.\)
+    In other words, the image of \(h\) on \(⟨s^*⟩_{Ŝ_{e}}\) is equal to \(⟨e⟩_{D̂}\); thus, because \(h'\) the restriction of \(h\) on \(⟨s^*⟩_{Ŝ_{e}}\), the range of \(h'\) contains its codomain \(⟨e⟩_{D̂}\), showing that \(h'\) is surjective.
+\end{proof}
+
+Therefore, \(⟨s^*⟩_{Ŝ_{e}}\) will have no less state than \(⟨e⟩_{D̂}\) as a consequence of the surjectivity of \(h'\) in~\cref{thm:suj-hom-thompson-derivative}.
+Because \(Ŝ_{e}\) have more states than \(⟨s^*⟩_{Ŝ_{e}}\), which have more states than \(⟨e⟩_{D̂}\), and notice that the size of \(Ŝ_{e}\) is linear to the size of the input expression, therefore \(⟨e⟩_{D̂}\) is at most linear to the size of the expression.
+
+Although the size of the coalgebra generated by the derivative is smaller than Thompson's construction, this does not imply that the decision procedure based on derivative is always more efficient than those based on Thompson's construction.
+Crucially, the states in the derivative coalgebra \(D̂\) are expressions, and computing the next transition of the coalgebra can also be more expensive.
+Thompson's construction, on the other hand, only requires inductively going through the expression once to construct the entire coalgebra \(Ŝ_{e}\), and its states can be represented by more efficient constructs, like integers.
 

### 2025-01-06 main.tex
proof of correctness

@@
 This is well-defined, i.e. no more than one clause can be satisfied precisely because the boolean expressions appear in \(ϵ̂\) and \(δ̂\) are disjoint.
-We usually use \(S\) to denote the lowering of \(Ŝ\).
+We usually use \(S\) to denote the lowering of \(Ŝ\); and the semantics of a state \(s ∈ Ŝ\) is defined as its semantics in the lowering \(⟦s⟧_{Ŝ} ≜ ⟦s⟧_{S}.\)
 
@@
 
-\begin{theorem}[Functoriality]
+\begin{theorem}[Functoriality]\label{thm:lowering-functor}
     The lowering operation is a functor, given a symbolic GKAT coalgebra homomorphism \(h: Ŝ → Û\), then \(h\) is also a homomorphism \(h: S → U\).
@@
 This fact can be proven by case analysis on the shape of the source expression, and forms a basis on our correctness argument.
-\begin{theorem}[Correctness]
-    The lowering of \(D̂\) is exactly the derivative defined by Schmid et. al.~\cite{schmid_GuardedKleeneAlgebra_2021}.
-    Therefore, the semantics of the expression is equal to the semantics \(⟦e⟧ = ⟦e⟧_{D̂}\).
+\begin{theorem}[Correctness]\label{thm:derivative-correctness}
+    The lowering of \(D̂\), denoted \(D\), is exactly the derivative defined by~\citeauthor{schmid_GuardedKleeneAlgebra_2021}~\cite{schmid_GuardedKleeneAlgebra_2021}.
+    Therefore, the semantics of the expression is equal to the semantics generated by the derivative coalgebra, \(⟦e⟧ = ⟦e⟧_{D} = ⟦e⟧_{D̂}\).
+\end{theorem}
+\begin{proof}
     TODO: unfold the statement.
-\end{theorem}
+\end{proof}
 
@@
 
+\begin{proof}
+    By functoriality of lowering~\cref{thm:lowering-functor}, then \(h: S_{e} → D\) is a homomorphism on their lowerings. 
+    By standard theorem of universal coalgebra, homomorphism preserves the result to the final coalgebra~\cite{rutten_UniversalCoalgebraTheory_2000}, in particular, \(⟦s^*⟧^{ω}_{S_{e}} = ⟦h(s^*)⟧^{ω}_{D} = ⟦e⟧^{ω}_{D}\), where \(⟦-⟧^{ω}\) is the infinite trace semantics i.e.
+the unique map into the final GKAT coalgebra.
+
+    Finally, because infinite trace equivalence implies finite trace equivalence (\Cref{thm:inf-trace-equiv-implies-fin-trace-equiv}), \(⟦s^*⟧_{S_{e}} = ⟦e⟧_{D}\); then the correctness of derivative (\Cref{thm:derivative-correctness}) concludes our proof: \(⟦s^*⟧_{S_{e}} = ⟦e⟧_{D} = ⟦e⟧\).
+\end{proof}
+
 A not so obvious consequence of the homomorphism in~\Cref{thm:hom-thompson-derivative}, is the complexity of the algorithm based on derivatives.

### 2025-01-06 main.tex
minor clean up

@@
 
-The original algorithm for deciding GKAT equivalences~\cite{smolka_GuardedKleeneAlgebra_2020} requires the entire automaton to be known prior to the execution of the bisimulation algorithm; specifically, in order to compute the liveness of a state \(s\), it is necessary iterate through all its reachable states \(⟨s⟩\) to see if there are any accepting states within.
+The original algorithm for deciding GKAT equivalences~\cite{smolka_GuardedKleeneAlgebra_2020} requires the entire coalgebra to be known prior to the execution of the bisimulation algorithm; specifically, in order to compute the liveness of a state \(s\), it is necessary iterate through all its reachable states \(⟨s⟩\) to see if there are any accepting states within.
 This limitation poses challenges to design an efficient on-the-fly algorithm for GKAT.
-In order to make the decision procedure scalable, we will need to merge the normalization and bisimulation procedure, so that our algorithm can normalized the automaton only when we need to.
+In order to make the decision procedure scalable, we will need to merge the normalization and bisimulation procedure, so that our algorithm can normalize the coalgebra only when we need to.
 
 In this section, we introduce an algorithm that merges bisimulation and normalization where we only need to test the liveness of the state when a disparity in the bisimulation has been found.
-For example, when one automaton leads to reject where the other transition to a state, then we will need to verify whether that state is dead or not.
-
-This on-the-fly algorithm inherits the efficiency of the original algorithm~\cite{smolka_GuardedKleeneAlgebra_2020}, where the worst case will require two passes of the automaton, where one pass will try to establish a bisimulation, when failed the other pass will kick in and compute whether the failed states are dead.
-In some special case, the on-the-fly algorithm can even out perform the original algorithm; for example, when the two input automata are bisimular (even when they are not normal), the on-the-fly algorithm can skip the liveness checking, only performing the bisimulation.
+For example, when one state leads to reject where the other transition to another state, then we will need to verify whether the resulting state is dead or not.
+
+This on-the-fly algorithm inherits the efficiency of the original algorithm~\cite{smolka_GuardedKleeneAlgebra_2020}, where the worst case will require two passes of the coalgebra, where one pass will try to establish a bisimulation, when failed the other pass will kick in and compute whether the failed states are dead.
+In some special case, the on-the-fly algorithm can even out perform the original algorithm; for example, when the two input states are bisimular (even when they are not normal), the on-the-fly algorithm can skip the liveness checking, only performing the bisimulation.
 
@@
                 \end{cases} \mathrel{\&\!\&}\\
-                & ∀ (b, s', p) ∈ δ̂_S(s), c ∈ ρ̂_U(u), (b ∧ c) ≢ 0 ⟹ \text{\Call{IsDead$_S$}{$s'$}} \mathrel{\&\!\&}\\
-                & ∀ b ∈ ρ̂_S(s), (c, u', q) ∈ δ̂_U(u), (b ∧ c) ≢ 0 ⟹ \text{\Call{IsDead$_U$}{$u'$}}
+                & ∀ (b, s', p) ∈ δ̂_S(s), c ∈ ρ̂_U(u), (b ∧ c) ≢ 0 ⟹ \text{\Call{isDead$_S$}{$s'$}} \mathrel{\&\!\&}\\
+                & ∀ b ∈ ρ̂_S(s), (c, u', q) ∈ δ̂_U(u), (b ∧ c) ≢ 0 ⟹ \text{\Call{isDead$_U$}{$u'$}}
             \end{aligned}\)
@@
 Inspired by the syntax of Ocaml, the \(\mathrel{\&\!\&}\) is the logical-and operator on the language level, specifying that all four conditions in the return statements must be satisfied to return true.
-Notice just like the non-symbolic case, this algorithm can be modified to perform symbolic bisimulation of (non-normalized) GKAT automaton, which coincides with infinite trace equivalence~\cite{schmid_GuardedKleeneAlgebra_2021}, by letting \textsc{IsDead} always return false and keep deadStates empty.
+Notice just like the non-symbolic case, this algorithm can be modified to perform symbolic bisimulation of (non-normalized) GKAT coalgebra, which coincides with infinite trace equivalence~\cite{schmid_GuardedKleeneAlgebra_2021}, by letting \textsc{isDead} and \textsc{knownDead} to always return false.
 

### 2025-01-07 main.tex
pass on background

@@
 
-\paragraph{Notation: } In this paper, we will use un-curried notation to apply curried functions, for example, given a function \(δ: X → Y → Z\), we will write the function applications as follow \(δ(x): Y → Z\) and \(δ(x, y): Z\).
-And when drawing commutative diagram, we will leave function restriction implicit.
-Specifically given \(A' ⊆ A\), and a function \(h: A → B\), we will draw:
-\[
-    \begin{tikzcd}
-        A' \ar{r}{h} & B
-    \end{tikzcd}
-\]
-where the function \(h\) is implicitly restricted to \(A'\).
-For bifunctors like \((-) × (-)\), we will write function lifting by applying the bifunctors on these functions: for example, given \(h₁: A₁ → B₁\) and \(h₂: A₂ → B₂\), we will use \[h₁ × h₂: A₁ × A₂ → B₁ × B₂\] to denote the bifunctorial lift of \(h₁\) and \(h₂\) via product \((-) × (-)\).
+\paragraph{Notation:} In this paper, we will use un-curried notation to apply curried functions, for example, given a function \(δ: X → Y → Z\), we will write the function applications as \(δ(x): Y → Z\) and \(δ(x, y): Z\).
 
@@
 
-When we can restrict the homomorphism map into a inclusion map \(i: S' → S\) for \(S' ⊆ S\), then we say that \(S'\) is a \emph{sub-coalgebra} of \(S\), denoted as \(S' ⊑ S\).
+When we can restrict the homomorphism map into an inclusion map \(i: S' → S\) for \(S' ⊆ S\), or more generally any monomorphism, then we say that \(S'\) is a \emph{sub-coalgebra} of \(S\), denoted as \(S' ⊑ S\).
 Specifically, the following diagram commutes when \(S' ⊑ S\):
@@
 In fact, the function \(σ_{S'}\) is uniquely determined by the states \(S'\)~\cite[Proposition 6.1]{rutten_UniversalCoalgebraTheory_2000}.
-
-The sub-coalgebras are preserved under homomorphic images and pre-images: 
+Sub-coalgebras are also preserved under homomorphic images and pre-images: 
 \begin{lemma}[Theorem 6.3~\cite{rutten_UniversalCoalgebraTheory_2000}]\label{thm:hom-(pre)img-preserve-sub-coalg}
@@
 
-One particularly important sub-coalgebra of and coalgebra \(S\) is the least coalgebra generated by a single element \(s\). 
+One particularly important sub-coalgebra of a coalgebra \(S\) is the least sub-coalgebra that contains a state \(s\). 
 We will denote this sub-coalgebra as \(⟨s⟩_{S}\), and call it \emph{principle sub-coalgebra} generated by \(s\). 
 We sometimes omit the subscript \(S\) when it can be inferred from context or irrelevant.
-Intuitively, we usually think of principle sub-coalgebra \(⟨s⟩_S\) as the sub-coalgebra that is formed by all the ``reachable state'' form the state \(s\).
-This coalgebraic characterization of reachable state can allow us to avoid induction on the length of path from \(s\) to reach another state.
+Intuitively, we usually think of principle sub-coalgebra \(⟨s⟩_S\) as the sub-coalgebra that is formed by all the ``reachable state'' from \(s\).
+This coalgebraic characterization of reachable state can allow us to avoid induction on the length of path from \(s\) to a state in \(⟨s⟩_S\).
 
@@
     We will need to show that \(h(⟨s⟩_{S})\) is the smallest sub-GKAT coalgebra of \(S\) that contain \(h(s)\). 
-First by definition of image, \(h(s) ∈ h(⟨s⟩_{S})\); second by \cref{thm:hom-(pre)img-preserve-sub-coalg}, \(h(⟨s⟩_S) ⊑ U\).
-
-    Finally, take any \(U' ⊑ U\) and \(h(s) ∈ U'\), recall that by \cref{thm:hom-(pre)img-preserve-sub-coalg}, \(h^{-1}(U') ⊑ S\). We can then derive that \(h(⟨s⟩_S) ⊑ U'\): 
+    By definition of image, \(h(⟨s⟩_{S})\) indeed contain \(h(s)\). 
+    By \Cref{thm:hom-(pre)img-preserve-sub-coalg}, \(h(⟨s⟩_S) ⊑ U\), i.e. \(h(⟨s⟩_S)\) is a sub-coalgebra of\(U\).
+    Finally, take any sub-coalgebra \(U' ⊑ U\) s.t. \(h(s) ∈ U'\): 
     \begin{align*}
@@
         & ⟹ ⟨s⟩_S ⊑ h^{-1}(U') & \text{definition of \(⟨s⟩_S\)}\\  
-        & ⟹ h(⟨s⟩_S) ⊑ U' & \text{\cref{thm:hom-(pre)img-preserve-sub-coalg}}
+        & ⟹ h(⟨s⟩_S) ⊑ U' & \text{\Cref{thm:hom-(pre)img-preserve-sub-coalg}}
     \end{align*}
@@
 
-A \emph{final coalgebra} \(ℱ\) over a signature \(F\), sometimes called the \emph{behavior} of coalgebras over \(F\), is a \(F\)-coalgebra s.t.
-for all \(F\)-coalgebra \(S\), there exists a unique homomorphism \(⟦-⟧_S: S → ℱ\).
+A \emph{final coalgebra} \(ℱ\) over a signature \(F\), sometimes called the \emph{behavior} or \emph{semantics} of coalgebras over \(F\), is an \(F\)-coalgebra s.t.
+for all \(F\)-coalgebra \(S\), there exists a unique homomorphism \(\mathrm{beh}_S: S → ℱ\).
 
@@
 A relation \({∼} ⊆ S × U\) is called a \emph{bisimulation relation} if it forms an \(F\)-coalgebra: \[σ_{∼}: {∼} → F(∼),\] 
-And its projections \(π₁: S × U → S\) and \(π₂: S × U → U\) are both homomorphisms:
+and its projections \(π₁: {∼} → S\) and \(π₂: {∼} → U\) are both homomorphisms:
 \[
-    \begin{tikzcd}
+    \begin{tikzcd}[column sep=1.25cm]
         S \ar[swap]{d}{σ_S}
@@
 
-\begin{corollary}
-    Given two \(F\)-coalgebra \(S, U\) and a homomorphism \(h: S → U\), then all for all \(s\), \(⟦s⟧_{S} = ⟦h(s)⟧_{U}.\)
+When given a homomorphism \(h: S → U\), we can construct a bisimulation \({∼} ⊆ S × U\), where it relates all the states \(s ∈ S\) with \(h(s) ∈ U\); and each component of the pair will transition individually via their respective transition function \(δ_∼ ≜ δ_S × δ_U\), where \(×\) is the bifunctorial lift of product.
+\begin{corollary}\label{thm:hom-preserves-semantics}
+    Given a homomorphism \(h: S → U\) between two \(F\)-coalgebra \(S, U\), then all for all states \(s ∈ S\), \(\mathrm{beh}_S(s) = \mathrm{beh}_U(h(s)).\)
 \end{corollary}
 
-Finally, we are interested in bisimulation equivalence.
-Given a coalgebra \(S\), a bisimulation \({≃} ⊆ S × S\) is a bisimulation equivalence when \(≃\) is a bisimulation and also an equivalence relation.
-As it turns out, searching for bisimulation is equivalent to bisimulation equivalence. 
-This result is very important, as it allows us to leverage more efficient data structure, like union-find, to search for equivalence relation instead of relations in general.
-
-\begin{theorem}\label{thm:bisim-iff-bisim-eq}
-    For any coalgebra \(S\), there exists a bisimulation \({∼} ⊆ S × S\) s.t. \(s₁ ∼ s₂\) if and only if there exists a bisimulation equivalence \({≃} ⊆ S × S\) s.t. \(s₁ ≃ s₂\)
-\end{theorem}
-
-\begin{proof}
-    First show the \(⟹\) direction, we consider the maximal bisimulation between \(S\) and itself \({≡} ⊆ S × S\), which is known to be a bisimulation equivalence~\cite[Corollary 5.6]{rutten_UniversalCoalgebraTheory_2000}.
-    Because \(≡\) is maximal, therefore \({∼} ⊆ {≡}\) and \(s₁ ≡ s₂\).
-
-    Then the \(⟸\) direction can be proven by setting \({∼} ≜ {≃}\).
-\end{proof}
-
 \subsection{GKAT and Its Coalgebra}
@@
 Guarded Kleene Algebra with Tests, or GKAT~\cite{smolka_GuardedKleeneAlgebra_2020}, is a deterministic fragment of Kleene Algebra with Tests. 
-The syntax of GKAT over a set of primitive actions \(K\) and a set of primitive tests \(T\) can be defined in two sorts, boolean expressions \(\BExp\) and expressions \(\Exp\):
+The syntax of GKAT over a set of primitive actions \(K\) and a set of primitive tests \(T\) can be defined in two sorts, boolean expressions \(\BExp\) and GKAT expressions \(\Exp\):
 \begin{align*}
@@
 \end{align*}
-where \(e +_b f\) is the if statement with condition \(b\); \(e;f\) is the sequencing of expression \(e\) and \(f\); finally \(e^{(b)}\) is the while loop with body \(e\) and condition \(b\).
+where \(e +_b f\) is the if statement with condition \(b\), \(e;f\) is the sequencing of expression \(e\) and \(f\), and \(e^{(b)}\) is the while loop with body \(e\) and condition \(b\).
+We use notation like \(b ≤ c\), \(b ≡ c\), and \(b ≢ c\) for the usual order, equivalence, and inequivalence in Boolean Algebra.
 A GKAT expression can be unfolded into a KAT expression in the usual manner~\cite{kozen_KleeneAlgebraTests_1997c}:
@@
 Then the semantics of each expression \(⟦e⟧\) can be computed by the semantics of Kleene Algebra with tests~\cite{kozen_KleeneAlgebraTests_1997c}.
-An important construct in the semantics is \emph{atoms}, which are conjunctions of all the primitive tests either in its positive or negative form: for \(B ≜ \{b₁, b₂, …, bₙ\}\)
-\[\At_B ≜ \{b₁' ∧ b₂' ∧ ⋯ ∧ bₙ' ∣ bᵢ' ∈ \{bᵢ, \overline{bᵢ}\}\}.\]
-Follow the conventional notation, we denote an atom using \(α, β\); and we sometimes omit the subscript \(B\) when no confusing can arise.
-For a boolean expression \(b\), if \(α\) can be proven to be smaller than \(b\) using axioms of the Boolean Algebra, we will say \(b\) \emph{contains} \(α\), denoted as \(α ≤ b\).
+An important construct in the semantics is \emph{atoms}, which are conjunctions of all the primitive tests either in its positive or negative form: for \(T ≜ \{t₁, t₂, …, tₙ\}\)
+\[\At_T ≜ \{t₁' ∧ t₂' ∧ ⋯ ∧ tₙ' ∣ tᵢ' ∈ \{tᵢ, \overline{tᵢ}\}\}.\]
+Follow the conventional notation, we denote an atom using \(α, β\); and we sometimes omit the subscript \(T\) when no confusing can arise.
 Alternatively, atoms can also be thought of as truth assignments to each primitive tests, indicating which primitive tests is satisfied in the current program states; and \(α ≤ b\) if and only if the truth assignment represented by \(α\) satisfies \(b\).
@@
 For the sake of brevity, we omit the complete definition of GKAT and KAT semantics, we refer the reader to previous works~\cite{smolka_GuardedKleeneAlgebra_2020,schmid_GuardedKleeneAlgebra_2021,kozen_KleeneAlgebraTests_1997c}, which explains these semantics in detail.
-Our work avoids directly interact with the semantics by leveraging results in the coalgebraic theory of GKAT, which we will recap below.
+Our work avoids direct interaction with the semantics by leveraging prior results in the coalgebraic theory of GKAT, which we will recap below.
 
 Formally, GKAT coalgebras over primitive actions \(K\) and primitive tests \(T\) are coalgebras over the following functor:
-\[G(S) ≜ (\{\accept, \reject\} + S × K)^{\At_B}.\] 
-Intuitively, given a state \(s ∈ S\) and an atom \(α ∈ \At\), \(δ(s, α)\) will deterministically execute one of the following: reject \(α\), denoted as \(δ(s, α) = \reject\); accept \(α\), denoted \(δ(s, α) = \accept\); or transition to a state \(s' ∈ S\) and execute action \(p ∈ K\), denoted as \(δ(s, α) = (s', p)\).
+\[G(S) ≜ (\{\accept, \reject\} + S × K)^{\At_T}.\] 
+Intuitively, given a state \(s ∈ S\) and an atom \(α ∈ \At\), \(δ(s, α)\) will deterministically execute one of the following: reject \(α\) when \(δ(s, α) = \reject\); accept \(α\) when \(δ(s, α) = \accept\); or transition to a state \(s' ∈ S\) and execute action \(p ∈ K\) when \(δ(s, α) = (s', p)\).
 
@@
 As we will see later, the deterministic behavior of GKAT coalgebra not only enables a more versatile symbolic algorithm than KCT~\cite{pous_SymbolicAlgorithmsLanguage_2015}, but also present challenges.
-Specifically, GKAT coalgebra requires normalization to compute finite trace equivalences~\cite{smolka_GuardedKleeneAlgebra_2020}, where we will remove all the state that cannot lead to acceptance, which are called ``dead states''.
-
-In previous works, these dead states are detected after all the necessary state and transition is computed.
-This approach requires storing all the information of the coalgebra in memory, which do not allow on-the-fly computation, which can not only terminate whenever a counter-example is found, but can also erase past states from memory.
-
-However, to truly understand the on-the-fly algorithm, we will first need to define ``dead states'', and its role in defining the coalgebraic semantic of GKAT.
+Specifically, GKAT coalgebra requires normalization to compute finite trace equivalences~\cite{smolka_GuardedKleeneAlgebra_2020}, where we reroute all the transitions that cannot lead to acceptance immediately into rejection; and states that can never lead to acceptance is called \emph{dead states}.
+
+In previous works, these dead states are detected after the necessary coalgebra is computed and stored.
+This approach requires storing all the transition and states of coalgebra in memory, meaning that the algorithm does not short circuit even if a mismatch can be detected in the starting state; for example, when for some \(α ∈ \At\), \(δ_S(s, α) = \reject\) and \(δ_U(u, α) = \accept\).
+
+Our algorithm is lazy in the dead state detection i.e.
+the dead state are only checked when a mismatch requires it; even then, our dead state detection algorithm will only check the necessary states to compute the liveness of the states causing the mismatch.
+This not only avoids unnecessarily checking the liveness of a state, but also enables allows on-the-fly generation of the coalgebra, like using derivatives, where we can remove irrelevant states from memory after it has been explored.
+
+However, to truly understand our on-the-fly algorithms, we will first need to define ``dead states'', and its role in defining the coalgebraic semantic of GKAT.
 
@@
 
-% liveness
-
 Traditionally, live and dead states are defined by whether they can reach an accepting state~\cite{smolka_GuardedKleeneAlgebra_2020}. 
-However, it is straightforward to show that principle sub-coalgebra \(⟨s⟩_S\) exactly correspondents to reachable states of \(s\) in coalgebra \(S\).
+However, we can use induction on the length of the path to show that principle sub-coalgebra \(⟨s⟩_S\) contains exactly the reachable states of \(s\) in any GKAT coalgebra \(S\). 
 Thus, the classical definition is equivalent to the following:
 \begin{definition}[liveness of states]\label{def:liveness-of-states}
-    A state \(s\) is \emph{accepting} if there exists a \(α ∈ \At\) s.t.
-\(δ(s, α) = \accept\).
-A state \(s'\) is \emph{live} if there exists an accepting state \(s' ∈ ⟨s⟩\).
-A state \(s'\) is \emph{dead} if there is no accepting state in \(⟨s⟩\).
+    A state \(s\) is \emph{accepting} if there exists an atom \(α ∈ \At\) s.t.
+\(δ(s, α) = \accept\); a state \(s'\) is \emph{live} if there exists an accepting state \(s' ∈ ⟨s⟩\); and a state \(s'\) is \emph{dead} if there is no accepting state in \(⟨s⟩\).
 \end{definition}
-This alternative liveness definition can help us formally prove important theorems regarding reachability and liveness without performing induction on traces.
-We can show the following lemmas as examples:
+This alternative liveness definition can help us prove important theorems regarding reachability and liveness without explicitly performing induction on traces. 
+We show the following theorems as examples:
 \begin{lemma}\label{thm:dead-iff-all-reachable-dead}
@@
     \(⟹\) direction can be proven as follows.
-    Take \(s' ∈ ⟨s⟩\), then \(⟨s'⟩ ⊑ ⟨s⟩\) by definition. 
-    Since there is no accepting state in \(⟨s⟩\), thus there cannot be any accepting state in \(⟨s'⟩\), hence \(⟨s'⟩\) is also dead.
+    Take any \(s' ∈ ⟨s⟩\), then \(⟨s'⟩ ⊑ ⟨s⟩\) because \(s'\) is the minimal sub-coalgebra that contains \(s'\). 
+    Since there is no accepting state in \(⟨s⟩\), thus there cannot be any accepting state in \(⟨s'⟩\), hence \(s'\) is also dead.
 \end{proof}
@@
     \[s' ∈ ⟨s⟩_S ⟺ h(s') ∈ h(⟨s⟩_S) ⟺ h(s') ∈ ⟨h(s)⟩_U.\]
-    And because \(s'\) is accepting if and only if \(h(s')\) accepting by definition of homomorphism; then \(⟨s⟩_S\) contains an accepting state if and only if \(h(⟨s⟩_S) = ⟨h(s)⟩_U\) contains an accepting state.
+    And because \(s'\) is accepting if and only if \(h(s')\) accepting by definition of homomorphism; then \(⟨s⟩_S\) contains an accepting state if and only if \(⟨h(s)⟩_U\) contains an accepting state.
     Therefore, \(s\) is live in \(S\) if and only if \(h(s)\) is live in \(U\).
@@
 
-The above theorem then leads to several interesting liveness preservation properties for important structures on coalgebras, like sub-coalgebra and bisimulation.
+The above theorem then leads to several interesting liveness preservation properties for structures on coalgebras, like sub-coalgebra and bisimulation.
 
@@
 
-% trace semantics
 \subsection{Normalization And Semantics}
 
-(Possibly infinite) trace model \(𝒢_ω\) is the final coalgebra of GKAT coalgebras~\cite{schmid_GuardedKleeneAlgebra_2021}.
-The finality of the model implies that every state in any GKAT coalgebra \(S\) can be assigned a semantics under the unique homomorphism \(⟦-⟧^{ω}_{S}: S → 𝒢_ω\); and such semantic equivalences can indeed be identified by bisimulation~\cite{schmid_GuardedKleeneAlgebra_2021}:
+Infinite-trace model \(𝒢_ω\) is the final coalgebra of GKAT coalgebras~\cite{schmid_GuardedKleeneAlgebra_2021}.
+The finality of the model allows us to define the semantics of each state in a given GKAT coalgebra \(S\), which we will denote as \(⟦-⟧^{ω}_{S}: S → 𝒢_ω\), where the semantic equivalences can identified by bisimulation~\cite{schmid_GuardedKleeneAlgebra_2021}:
 \(⟦s⟧^{ω}_{S} = ⟦t⟧^{ω}_{T}\) if and only if there exists a bisimulation \({∼} ⊆ S × T\), s.t. \(s ∼ t\).
@@
 Fortunately every GKAT coalgebra can be normalized by rerouting all the transition from dead states to rejection.
-Concretely, given a GKAT coalgebra \(S ≜ (S, δ_S)\), we let \(δ_{\norm(S)} : S → G(S)\) is defined as \(δ_{\norm(S)}(s, α) ≜ \reject\) when \(δ_S(s, α) = (s', p)\) and \(s'\) is dead in \(S\); and \(δ_{\norm(S)}(s, α) ≜ δ_S(s, α)\) otherwise; then \((S, δ_{\norm(S)})\) is the \emph{normalized} coalgebra of \(S ≜ (S, δ_S)\) denoted as \(\norm(S)\).
-
-The finality of \(𝒢\) means that the finite trace semantics \(⟦-⟧\) is the unique coalgebra homomorphism \(\norm(S) → 𝒢\). 
-Furthermore, the finite trace equivalence between \(s ∈ S\) and \(u ∈ U\) can be computed by first normalizing \(S\) and \(U\), then decide whether there is a bisimulation on \(\norm(S)\) and \(\norm(U)\) that includes \((s, t)\).
-For a more detailed explanation on the finite trace semantics, we refer the reader to the work of \Citeauthor{smolka_GuardedKleeneAlgebra_2020}~\cite{smolka_GuardedKleeneAlgebra_2020}, however we will recall two important theorems here.
+Concretely, given a GKAT coalgebra \(S ≜ (S, δ_S)\), \(δ_{\norm(S)} : S → G(S)\) is defined as \(δ_{\norm(S)}(s, α) ≜ \reject\) when \(δ_S(s, α) = (s', p)\) and \(s'\) is dead in \(S\); and \(δ_{\norm(S)}(s, α) ≜ δ_S(s, α)\) otherwise.
+We call \(\norm(S) ≜ (S, δ_{\norm(S)})\) the \emph{normalized} coalgebra of \(S\).
+
+And we use \(⟦-⟧_S: \norm(S) → 𝒢\) to denote the finite trace semantics of GKAT coalgebra, which is the unique homomorphism into the final coalgebra \(𝒢\). 
+The finite trace equivalence between \(s ∈ S\) and \(u ∈ U\) can be decided by first normalizing \(S\) and \(U\) then deciding whether there is a bisimulation \({∼} ⊆ \norm(S) × \norm(U)\) s.t.
+\(s ∼ t\).
+For a more detailed explanation on the finite trace semantics, we refer the reader to the work of \Citeauthor{smolka_GuardedKleeneAlgebra_2020}~\cite{smolka_GuardedKleeneAlgebra_2020}, however we will recall the correctness theorem here.
 
@@
 Because of the functoriality, we can show that two states are infinite-trace equivalent implies these two states are finite-trance equivalent.
-This gives us more tool in proving semantic equivalence between two states in GKAT coalgebras: proving bisimulation in the GKAT coalgebra is already enough to obtain the semantic equivalence for two states.
+This gives us more tool in proving semantic equivalence between two states in GKAT coalgebras: proving bisimulation of these two states in the \emph{non-normalized} GKAT coalgebra can also obtain semantic equivalence for two states.
 
@@
     \end{tikzcd}\]
-    Because \(s ∼ u\) and normalization operation preserves states in \(∼\), therefore \((s, u) ∈ \norm(∼)\), and because \(\norm(∼)\) is a bisimulation between the normalization of \(S\) and \(U\), therefore \(⟦s⟧_{S} = ⟦u⟧_{U}\) (\cref{thm:norm-bisim-correctness}).
+    Because \(s ∼ u\) and normalization operation preserves states in \(∼\), therefore \((s, u) ∈ \norm(∼)\), and because \(\norm(∼)\) is a bisimulation between the normalization of \(S\) and \(U\), therefore \(⟦s⟧_{S} = ⟦u⟧_{U}\) (\Cref{thm:norm-bisim-correctness}).
 \end{proof}
@@
 Although \cref{alg:bisim} is on-the-fly, which do not require the construction of the entire coalgebra prior to running the bisimulation algorithm, it still uses GKAT coalgebra, which is known to be inefficient.
-Specifically, the transition function \(δ: S → \At_B → \{\accept, \reject\} + S × K\) requires computing the transition result for each \emph{atom}, and the number of atoms \(\At_B ≅ 2^{B}\) is exponential to the size of primitive tests in \(B\).
+Specifically, the transition function \(δ: S → \At_T → \{\accept, \reject\} + S × K\) requires computing the transition result for each \emph{atom}, and the number of atoms \(\At_T ≅ 2^{T}\) is exponential to the size of primitive tests in \(T\).
 
 In our symbolic GKAT coalgebra, instead of computing the behavior of each atom individually, we group them into boolean expressions, this leads to a much more space efficient coalgebra and an efficient equivalence checking algorithm with the help of off-the-shelf SAT solvers.
-Specifically, given the alphabet \(K, B\), a \emph{symbolic GKAT coalgebra} \(Ŝ ≜ (S, ϵ̂, δ̂)\) consists of a state set \(S\) and an accepting function \(ϵ̂\) and a transition function \(δ̂\):
+Specifically, given a set of primitive actions \(K\) and primitive tests \(T\), a \emph{symbolic GKAT coalgebra} \(Ŝ ≜ (S, ϵ̂, δ̂)\) consists of a state set \(S\) and an accepting function \(ϵ̂\) and a transition function \(δ̂\):
 \begin{mathpar}
-    ϵ̂: S → 𝒫(\BExp_B), \and
-    δ̂: S → 𝒫(\BExp_B × S × K),
+    ϵ̂: S → 𝒫(\BExp_T), \and
+    δ̂: S → 𝒫(\BExp_T × S × K),
 \end{mathpar}
@@
         \begin{aligned}
-            & ∃ b ∈ \BExp_B, α ≤ b \\
+            & ∃ b ∈ \BExp_T, α ≤ b \\
             & \text{ and } δ(s, b) = (s', p)  
@@
 
-We will then use \(ρ̂(s): \BExp_B\) to represent the boolean expressions that contain all the atoms that the state \(s\) rejects, and \(ρ̂(s)\) can be computed as follows:
+We will then use \(ρ̂(s): \BExp_T\) to represent the boolean expressions that contain all the atoms that the state \(s\) rejects, and \(ρ̂(s)\) can be computed as follows:
 \begin{align*}
@@
 \end{proof}
-
-% For the ease of implementation, we will concretely define the accepting function \(ϵ̂_D\) and transition function \(δ̂_D\) of the derivative coalgebra \(D̂\).
-% Prior to that we will introduce an operation to ``guarded'' a function by a set of boolean \(B\), denoted \(⟨B|\), and it can be applied to both the accepting dynamic \(ϵ̂(s)\) and transition dynamic \(δ̂(s)\):
-% \begin{align*}
-%     ⟨B|ϵ̂(s) & ≜ \{b ∧ c ∣ b ∈ B, c ∈ ϵ(s)\}; \\
-%     ⟨B|δ̂(s) & ≜ \{(b ∧ c, s', p) ∣ b ∈ B, (c, s', p) ∈ δ(s)\}.
-% \end{align*}
-
-% \begin{figure*}
-%     \centering
-%     \begin{align*}
-%         δ̂(b) & ≜ ∅ & ϵ̂(b) & ≜ \{b\} \\  
-%         δ̂(p) & ≜ \{(1, 1, p)\} & ϵ̂(p) ≜ ∅ \\  
-%         δ̂(e +_b f) & ≜ ⟨\{b\}| δ̂(e) ∪ ⟨\{\overline{b}\}| δ̂(f) & 
-%         ϵ̂(e +_b f) & ≜ ⟨\{b\}| ϵ̂(e) ∪ ⟨\{\overline{b}\}| ϵ̂(f)
-%     \end{align*}
-% \end{figure*}
 

### 2025-01-07 main.tex
checkpoint symbolic

@@
 
-When we can restrict the homomorphism map into an inclusion map \(i: S' → S\) for \(S' ⊆ S\), or more generally any monomorphism, then we say that \(S'\) is a \emph{sub-coalgebra} of \(S\), denoted as \(S' ⊑ S\).
+When we can restrict the homomorphism map into an inclusion map \(i: S' → S\) for \(S' ⊆ S\) then we say that \(S'\) is a \emph{sub-coalgebra} of \(S\), denoted as \(S' ⊑ S\).
 Specifically, the following diagram commutes when \(S' ⊑ S\):
@@
 
-This deterministic behavior contrast that of Kleene coalgebra with tests~\cite{kozen_CoalgebraicTheoryKleene_2017}, where for each atom, the state can accept or  reject the atom (but not both), yet the state can also non-deterministically transition to multiple different state via the same atom, while executing different actions.
-% TODO: I think this belongs in the intro
+In particular, \(G\) is a simple polynomial functor~\cite{jacobs_IntroductionCoalgebraMathematics_2016}, allowing the sub-coalgebra to preserve and reflect bisimulations.
+\begin{theorem}[sub-coalgebras perserve and reflect bisimulation]\label{thm:sub-coalg-preserve-bisim}
+    Given two states in sub-coalgebra \(s ∈ S' ⊑ S\) and \(u ∈ U' ⊑ U\), there exists a bisimulation \({∼'} ⊆ S' × U'\) s.t.
+\(s ∼' u\) if and only if there exists a bisimulation \({∼} ⊆ S × U\) s.t.
+\(s ∼ u\).
+\end{theorem}
+
+\begin{proof}
+    For \(⟹\) direction, we will show that if \({∼'} ⊆ S' × U'\) is a bisimulation, then \({∼'}\) is also a bisimulation between \(S\) and \(U\):
+    \[
+        \begin{tikzcd}
+            S \ar{d}{δ_S} & S' \ar[hook',swap]{l}{i} \ar{d}{δ_S}
+            & ∼ \ar[swap]{l}{π₁} \ar{r}{π₂} \ar{d}{δ_∼}
+            & U' \ar[hook]{r}{i} \ar{d}{δ_S} & U \ar{d}{δ_T}\\  
+            G(S) & G(S') \ar[hook',swap]{l}{G(i)} 
+            & G(∼) \ar[swap]{l}{G(π₁)} \ar{r}{G(π₂)} & U' \ar[hook]{r}{G(i)} & U 
+        \end{tikzcd}
+    \]
+    Because the inclusion homomorphism \(i\) doesn't change the input, we have:
+    \begin{align*}
+        & {∼} \xrightarrow{π₁} S' \xrightarrow{i} S = {∼} \xrightarrow{π₁} S; \\
+        & {∼} \xrightarrow{π₂} U' \xrightarrow{i} U = {∼} \xrightarrow{π₂} U.
+    \end{align*}
+
+    To prove the \(⟸\) we will show that if \({∼} ⊆ S × U\) is a bisimulation, then the restriction \({∼}' ≜ \{(s, u) ∈ {∼} ∣ s ∈ S', u ∈ U'\}\) is a bisimulation between \(S'\) and \(U'\).
+    We first realize that \(∼_{S', U'}\) is a pre-image of the maximal bisimulation \(≣_{S', U'}\) along the inclusion homomorphism \(i: {∼} → {≡_{S, U}}\).
+    This means that \(∼_{S', U'}\) can be formed by a pullback square:
+    \[
+        \begin{tikzcd}
+            ∼_{S', U'} \ar{r}{i} \ar[swap]{d}{i} \ar[phantom, very near start]{dr}{\scalebox{1.5}{\(\lrcorner\)}} & ≣_{S', U'} \ar{d}{i}\\ 
+            {∼} \ar[swap]{r}{i} & {≡_{S, U}}
+        \end{tikzcd}
+    \]
+    Recall that elementary polynomial functor~\cite{jacobs_IntroductionCoalgebraMathematics_2016} like \(G\) preserves pullback, hence the pullback also uniquely generates a GKAT coalgebra~\cite{rutten_UniversalCoalgebraTheory_2000}
+\end{proof}
+
+Besides nice property with bisimulation, GKAT coalgebra is also deterministic, unlike Kleene coalgebra with tests (KCT)~\cite{kozen_CoalgebraicTheoryKleene_2017}.
+For each atom, states in a KCT can accept or reject the atom (but not both), yet states can also non-deterministically transition to multiple different states via the same atom, while executing different actions.
 As we will see later, the deterministic behavior of GKAT coalgebra not only enables a more versatile symbolic algorithm than KCT~\cite{pous_SymbolicAlgorithmsLanguage_2015}, but also present challenges.
@@
 
-In this section, we introduce an algorithm that merges bisimulation and normalization where we only need to test the liveness of the state when a disparity in the bisimulation has been found.
-For example, when one state leads to reject where the other transition to another state, then we will need to verify whether the resulting state is dead or not.
-
-This on-the-fly algorithm inherits the efficiency of the original algorithm~\cite{smolka_GuardedKleeneAlgebra_2020}, where the worst case will require two passes of the coalgebra, where one pass will try to establish a bisimulation, when failed the other pass will kick in and compute whether the failed states are dead.
-In some special case, the on-the-fly algorithm can even out perform the original algorithm; for example, when the two input states are bisimular (even when they are not normal), the on-the-fly algorithm can skip the liveness checking, only performing the bisimulation.
-
-TODO: I think we should move the next couple theorem to the background.
-
-\begin{theorem}[sub-coalgebras perserve and reflect bisimulation]\label{thm:sub-coalg-preserve-bisim}
-    Given any sub-coalgebra \(S' ⊑ S\) and \(T' ⊑ T\),
-    \begin{itemize}
-        \item Given a bisimulation \(∼\) between \(S'\) and \(T'\), then \(∼\) is also a bisimulation between \(S\) and \(T\);
-        \item if there exists a bisimulation \(∼\) between \(S\) and \(T\), then the restriction 
-        \begin{mathpar}
-            ∼_{S', T'} ≜ \{(s, t) ∣ s ∈ S', t ∈ T', s ∼ t\}
-        \end{mathpar}
-        forms a bisimulation between \(S'\) and \(T'\).
-    \end{itemize}
-\end{theorem}
-
-\begin{proof}
-    To prove that bisimulation \(∼\) between \(S'\) and \(T'\) is also a bisimulation of \(S\) and \(T\), we can simply enlarge the diagram by the inclusion homomorphism
-    \[
-        \begin{tikzcd}
-            S \ar{d}{δ_S} & S' \ar[hook',swap]{l}{i} \ar{d}{δ_S}
-            & ∼ \ar[swap]{l}{π₁} \ar{r}{π₂} \ar{d}{δ_∼}
-            & T' \ar[hook]{r}{i} \ar{d}{δ_S} & T \ar{d}{δ_T}\\  
-            G(S) & G(S') \ar[hook',swap]{l}{G(i)} 
-            & G(∼) \ar[swap]{l}{G(π₁)} \ar{r}{G(π₂)} & T' \ar[hook]{r}{G(i)} & T \\  
-        \end{tikzcd}
-    \]
-    Because the inclusion homomorphism \(i\) doesn't change the input thus, we have:
-    \begin{mathpar}
-        {∼} \xrightarrow{π₁} S' \xrightarrow{i} S = {∼} \xrightarrow{π₁} S \and 
-        {∼} \xrightarrow{π₂} T' \xrightarrow{i} T = {∼} \xrightarrow{π₂} T
-    \end{mathpar}
-
-    To prove that the bisimulation can be restricted, we first realize that \(∼_{S', T'}\) is a pre-image of the maximal bisimulation \(≣_{S', T'}\) along the inclusion homomorphism \(i: {∼} → {≡_{S, T}}\).
-    This means that \(∼_{S', T'}\) can be formed by a pullback square:
-    \[
-        \begin{tikzcd}
-            ∼_{S', T'} \ar{r}{i} \ar[swap]{d}{i} \ar[phantom, very near start]{dr}{\scalebox{1.5}{\(\lrcorner\)}} & ≣_{S', T'} \ar{d}{i}\\ 
-            {∼} \ar[swap]{r}{i} & {≡_{S, T}}
-        \end{tikzcd}
-    \]
-    Recall that elementary polynomial functor~\cite{jacobs_IntroductionCoalgebraMathematics_2016} like \(G\) preserves pullback, hence the pullback also uniquely generates a GKAT coalgebra~\cite{rutten_UniversalCoalgebraTheory_2000}
-\end{proof}
-
-\Cref{thm:sub-coalg-preserve-bisim} allows us to only search for bisimulation on a sub-coalgebra, speeding up our search algorithm.
-Another way to speed up the algorithm is to use efficient data structures to find a bisimulation equivalence instead of a bisimulation relation.
-This optimization is a special case of the \emph{up-to technique}~\cite{bonchi_GeneralAccountCoinduction_2017a}.
-Specifically, we will extend~\Cref{thm:bisim-iff-bisim-eq} to a setting where the bisimulation is no longer over the same coalgebra.
-
-\begin{theorem}\label{thm:GKAT-bisim-iff-bisim-eq}
-    Given two states in two coalgebras \(s ∈ S, u ∈ U\), let \(S + U\) be the coproduct coalgebra of \(S\) and \(U\)~\cite{rutten_UniversalCoalgebraTheory_2000}. 
-    There exists a bisimulation \({∼} ⊆ S × U\) s.t. \(s ∼ u\) if and only if there exists a bisimulation equivalence on \({≃} ⊆ (S + U) × (S + U)\), s.t. \(s ≃ u\).
-\end{theorem}
-
-\begin{proof}
-    Notice that both \(S\) and \(U\) are sub-coalgebra of \(S + U\) witnessed by the canonical injection \(\mathrm{inj}ₗ: S → S + U\) and \(\mathrm{inj}ᵣ: U → S + U\).
-
-    Let \(∼\) be a bisimulation and \(≃\) be a bisimulation equivalence, by~\Cref{thm:sub-coalg-preserve-bisim} and~\Cref{thm:bisim-iff-bisim-eq}:
-    \begin{align*}
-        & ∃~ {∼} ⊆ S × U, s ∼ u \\
-        ⟺{}& ∃~ {∼} ⊆ (S+U) × (S+U), s ∼ u \\  
-        ⟺{}& ∃~ {≃} ⊆ (S+U) × (S+U), s ≃ u.
-        \qedhere
-    \end{align*}
-\end{proof}
-
-After we justify the above optimizations of the algorithm, we will show the core theorem that establishes the correctness of our algorithm.
+In this section, we introduce an algorithm that merges bisimulation and normalization, where we only test the liveness of states when a disparity is noticed by the bisimulation algorithm.
+For example, when deciding the trace equivalence between \(s ∈ S\) and \(u ∈ U\), if \(δ_S(s, α) = (s', p)\) and \(δ_U(u, α) = (u', p)\), we will proceed to recurse on \(s'\) and \(u'\), without checking the liveness of \(s'\) and \(u'\).
+However, if \(s\) rejects \(α\) instead of transitioning to \(s'\), we will then check whether \(u'\) is dead.
+
+We present several sound and complete conditions for finite-trace equivalence between states of GKAT coalgebras in~\Cref{thm:norm-bisim-correctness}, which also serves as the core correctness theorem for our equivalence-checking algorithm i.e.~\Cref{alg:bisim}.
+This algorithm mostly consists of recursively checking the conditions in~\Cref{thm:norm-bisim-correctness} with minor optimizations, which we will outline later.
 
@@
 
+With some slight tweak to \Cref{thm:recursive-construction-lemma}, it will be applicable to trace equivalence instead of bisimulation.
+The correctness of this modification is supported by the correspondence between trace equivalence and the existence of a bisimulation between normalized GKAT coalgebras, which is stated in~\Cref{thm:norm-bisim-correctness}.
+
 \begin{theorem}[Recursive Construction]\label{thm:recursive-construction}
@@
     \begin{align*}
-        & \text{Conditions in \cref{thm:recursive-construction-lemma} hold} \\  
+        & \text{Conditions in \Cref{thm:recursive-construction-lemma} hold} \\  
         ⟺{}& \text{Conditions in the current theorem hold}\\
@@
 
-The above theorem already gives us an algorithm to recursively decide whether \(⟦s⟧_{S} = ⟦u⟧_U\), when \(⟨s⟩_S\) and \(⟨u⟩_U\) is finite.
-However, this algorithm can be further optimized: we will then derive that a dead state is only trace equivalent to other dead states.
-This means that when checking the trace-equivalence of states \(s\) and \(u\), if we already know one of them is dead, we only need to check whether the other is dead, instead of going through all the conditions in~\cref{thm:recursive-construction}.
+The above construction theorem already gives us an algorithm to recursively decide whether \(⟦s⟧_{S} = ⟦u⟧_U\), when \(⟨s⟩_S\) and \(⟨u⟩_U\) is finite.
+However, this algorithm can be further optimized: we will derive that a dead state is only trace equivalent to other dead states. 
+This means that when checking the trace-equivalence of states \(s\) and \(u\), if we already know one of them is dead, we only need to check whether the other is dead, instead of going through all the conditions in~\Cref{thm:recursive-construction}.
 
@@
 
-Before we write down the optimized algorithm, we will first briefly sketch the dead state detection algorithm, which will use a DFS to check whether the there exists any accepting states in reachable states of \(s ∈ S\): if there exists any accepting state in \(⟨s⟩_S\), then the algorithm terminates immediately, and return that \(s\) is live, otherwise it would return all the states in \(⟨s⟩\), and by~\cref{thm:dead-iff-all-reachable-dead}, all the states in \(⟨s⟩\) is dead.
-
-We will cache all the known dead states from all the previous searches, we use the call \Call{knownDead\(_S\)}{$s$} to check whether the state \(s\) is in the cached dead states. 
-And \Call{isDead\(_S\)}{$s$} will first check if \(s\) is known to be dead, and invoke the depth-first-search algorithm in the coalgebra \(s\), if \(s\) is not in the cached dead states.
-Because the non-symbolic version of dead state checking is similar to the symbolic version, we only provide the symbolic version of this algorithm in appendix
+Before we introduce the optimized algorithm, we will first briefly sketch the liveness-check algorithm, which will use a depth-first search to check whether the there exists any accepting states in reachable states of \(s ∈ S\): if there exists any accepting state in \(⟨s⟩_S\), then the algorithm terminates immediately, and return that \(s\) is live, otherwise it would return all the states in \(⟨s⟩\), and by~\cref{thm:dead-iff-all-reachable-dead}, all the states in \(⟨s⟩\) is dead.
+In this case, we use depth-first search for its simplicity to implement, other search algorithm will also be sound.
+
+We will cache all the known dead states from the previous searches; whether a state \(s\) is in this cache can be checked by the function call \Call{knownDead\(_S\)}{$s$}. 
+The function \Call{isDead\(_S\)}{$s$} will first check if \(s\) is known to be dead, and invoke the search algorithm in the coalgebra \(⟨s⟩\), if \(s\) is not in the cached dead states.
+Because the non-symbolic version of liveness checking is similar to the symbolic version, we only provide the symbolic version of this algorithm in appendix
 TODO: give the link to the section.
@@
 
-TODO: I am not sure if we need to give some intuition for this algorithm here.
-
-Finally, we present our algorithm to decide trace-equivalence as~\cref{alg:bisim}, where we recursively check the condition in~\Cref{thm:recursive-construction}.
-Notice that trace equivalence is an equivalence relation, therefore we can organize the previously explored states into equivalent relation using union-find structure, instead of a set of state pairs.
-Specifically, we use the call \(\Call{union}{$s, u$}\) to union the equivalence class of \(s\) and \(u\), and use \(\Call{eq}{$s, u$}\) to check whether \(s\) and \(u\) is in the same equivalence class.
-In this algorithm, we first check if one of \(s\) and \(u\) is known to be dead, if so we only need to check whether the other is dead, because of~\Cref{thm:bisim-one-dead}; otherwise, we will check the conditions in~\Cref{thm:recursive-construction}.
+Finally, we present our equivalence-checking algorithm as~\cref{alg:bisim}, where we first check if one of \(s\) or \(u\) is known to be dead, if so we only need to check whether the other is dead, because of~\Cref{thm:bisim-one-dead}; otherwise, we will check the conditions in~\Cref{thm:recursive-construction}.
+Because trace equivalence is an equivalence relation, therefore we can organize the previously explored states into equivalent classes using an efficient union-find structure, instead of a set of state pairs.
+This union-find structure need to provide two functions: \(\Call{union}{$s, u$}\) unions the equivalence class of \(s\) and \(u\), whereas \(\Call{eq}{$s, u$}\) checks whether \(s\) and \(u\) is in the same equivalence class.
+
+Concerning the complexity, our algorithm have similar worst case complexity as the original algorithm.
+In particular, when deciding the trace equivalence of two states \(s ∈ S\) and \(u ∈ U\), the original algorithm~\cite{smolka_GuardedKleeneAlgebra_2020} requires one pass of \(⟨s⟩\) and \(⟨u⟩\) to normalize them, then the bisimulation will visit each pair of states in \(⟨s⟩ × ⟨u⟩\) at most once, which means that they will visit at most \(|⟨s⟩ + ⟨u⟩ + ⟨s⟩ × ⟨u⟩|\) number of states.
+
+Similarly, our equivalence-checking algorithm attempts to find a bisimulation first, which visits each pair in \(⟨s⟩ × ⟨u⟩\) at most once, and when a mismatch is found in the process, the liveness-checking algorithm is then invoked.
+Crucially, our algorithm satisfy the following property: if the liveness-checking algorithm find a live or accepting state, the entire equivalence checking algorithm will halt and return false.
+Thus, by caching all the known dead states, the liveness-checking only visits states in \(⟨s⟩\) and \(⟨u⟩\) at most once.
+Therefore, the worst case of our algorithm is also \(|⟨s⟩ × ⟨u⟩ + ⟨s⟩ + ⟨u⟩|\).
+
+However, the on-the-fly algorithm will always out-perform the original algorithm, in terms of number of states visited, because this algorithm only invoke liveness check when necessary.
+In the extreme case when the two input states are infinite-trace equivalent, the on-the-fly algorithm can skip liveness checking entirely.
 
@@
 
-Although \cref{alg:bisim} is on-the-fly, which do not require the construction of the entire coalgebra prior to running the bisimulation algorithm, it still uses GKAT coalgebra, which is known to be inefficient.
-Specifically, the transition function \(δ: S → \At_T → \{\accept, \reject\} + S × K\) requires computing the transition result for each \emph{atom}, and the number of atoms \(\At_T ≅ 2^{T}\) is exponential to the size of primitive tests in \(T\).
-
-In our symbolic GKAT coalgebra, instead of computing the behavior of each atom individually, we group them into boolean expressions, this leads to a much more space efficient coalgebra and an efficient equivalence checking algorithm with the help of off-the-shelf SAT solvers.
+Although \cref{alg:bisim} is on-the-fly, it still uses GKAT coalgebra, which contains exponentially many transitions with respect to the number of primitive tests \(|T|\).
+Concretely, the transition function \(δ: S → \At_T → \{\accept, \reject\} + S × K\) requires computing the transition result for each \emph{atom}, and the number of atoms \(\At_T ≅ 2^{T}\) is exponential to the size of primitive tests in \(T\).
+This is also why several GKAT-related complexity results only consider constant sized \(T\).
+
+Symbolic GKAT coalgebra, instead of computing the behavior of each atom individually, groups atoms into boolean expressions. 
+This optimization leads to space-efficient coalgebras and an equivalence checking algorithm making use of off-the-shelf SAT solvers.
 Specifically, given a set of primitive actions \(K\) and primitive tests \(T\), a \emph{symbolic GKAT coalgebra} \(Ŝ ≜ (S, ϵ̂, δ̂)\) consists of a state set \(S\) and an accepting function \(ϵ̂\) and a transition function \(δ̂\):

### 2025-01-07 main.tex
checkpoint symb bisim

@@
     ϵ̂: S → 𝒫(\BExp_T), \and
-    δ̂: S → 𝒫(\BExp_T × S × K),
+    δ̂: S → 𝒫(\BExp_T × S × K).
 \end{mathpar}
-where for all states \(s ∈ S\), all the boolean expressions are ``disjoint''; namely the conjunction of any two expressions from the set \(\{ϵ̂(s)\} ∪ \{b ∣ ∃ (b, s', p) ∈ δ(s)\}\) is equivalent to \(0\) under boolean algebra.
-
-Notice that we call \(ϵ̂\) the accepting function and \(δ̂\) the transition function; intuitively, a state \(s\) accepts an atom \(α\) if and only if there exists a \(b ∈ ϵ̂(s)\), s.t.
-\(α ≤ b\); similarly, the state \(s\) transitions to \(s'\) via atom \(α\) while executing \(p\) if and only if there exists \((b, s', p) ∈ δ̂(s)\) and \(α ≤ b\).
-With the above intuition in mind, a symbolic GKAT coalgebra \(Ŝ ≜ (S, ϵ̂, δ̂)\) can be lowered into a GKAT coalgebra \((S, δ)\) in the following manner:
+This coalgebra is also required to satisfy the disjointedness condition, i.e.
+for all states \(s ∈ S\), the conjunction of any two expressions from the set \(ϵ̂(s) ∪ \{b ∣ ∃ (b, s', p) ∈ δ(s)\}\) is equivalent to \(0\) under boolean algebra.
+
+We name \(ϵ̂\) the accepting function because intuitively a state \(s\) accepts an atom \(α\) when there exists a \(b ∈ ϵ̂(s)\), s.t.
+\(α ≤ b\); similarly, \(δ̂\) is called the transition function because a state \(s\) transitions to \(s'\) via atom \(α\) while executing \(p\) when there exists \((b, s', p) ∈ δ̂(s)\) and \(α ≤ b\).
+With the above intuition in mind, a symbolic GKAT coalgebra \(Ŝ ≜ (S, ϵ̂, δ̂)\) can be lowered into a GKAT coalgebra \(S ≜ (S, δ)\) in the following manner:
 \begin{align}\label{cons:lowering}
@@
 \end{align}
-This is well-defined, i.e. no more than one clause can be satisfied precisely because the boolean expressions appear in \(ϵ̂\) and \(δ̂\) are disjoint.
+This is well-defined, i.e. exactly one clause can be satisfied for any \(s ∈ S\) and \(α ∈ \At\), because of the disjointedness condition.
 We usually use \(S\) to denote the lowering of \(Ŝ\); and the semantics of a state \(s ∈ Ŝ\) is defined as its semantics in the lowering \(⟦s⟧_{Ŝ} ≜ ⟦s⟧_{S}.\)
 
-We will then use \(ρ̂(s): \BExp_T\) to represent the boolean expressions that contain all the atoms that the state \(s\) rejects, and \(ρ̂(s)\) can be computed as follows:
+We will then use \(ρ̂(s): \BExp_T\) to represent all the atoms that the state \(s\) rejects in the lowering, and \(ρ̂(s)\) is computed as follows:
 \begin{align*}
@@
 \end{align*}
-The trace semantics of a GKAT coalgebra \((S, ϵ̂, δ̂)\) is then defined as the trace semantics of its lowering \((S, δ)\).
 
 \begin{remark}[Canonicity]\label{rem:canonicity}
-    Notice that symbolic GKAT coalgebra is not canonical, i.e. there exists two different symbolic GKAT colagebra with the same lowering, consider the state set \(S ≜ \{s\}\):
+    Symbolic GKAT coalgebra is not canonical, i.e. there exists two different symbolic GKAT coalgebra with the same lowering, consider the state set \(S ≜ \{s\}\):
     \begin{align*}
@@
     These two symbolic GKAT coalgebra \(Ŝ₁ ≜ (S, δ̂₁, ϵ̂₁)\) and \(Ŝ₂ ≜ (S, δ̂₂, ϵ̂₂)\) have the same lowering and semantics, yet, they are different.
-    There are other symbolic representation that will satisfy canonicity, yet we opt to use our current representation for ease of construction and computational efficiency.
+    It is possible to construct symbolic representations that satisfies canonicity, yet we opt to use our current representation for ease of construction and computational efficiency.
 \end{remark}
@@
 \begin{theorem}[Functoriality]\label{thm:lowering-functor}
-    The lowering operation is a functor, given a symbolic GKAT coalgebra homomorphism \(h: Ŝ → Û\), then \(h\) is also a homomorphism \(h: S → U\).
+    The lowering operation is a functor, every symbolic GKAT coalgebra homomorphism \(h: Ŝ → Û\), is also a GKAT coalgebra homomorphism \(h: S → U\).
 \end{theorem}
@@
 
-The functoriality states that a homomorphism on two symbolic coalgebras implies a homomorphism of on their lowing; similarly, a bisimulation on symbolic GKAT coalgebra induces a bisimulation on their lowering.
-However, the converse is not true, precisely because of the canonicity problem noted in~\Cref{rem:canonicity}: take the \(Ŝ₁\) and \(Ŝ₂\) in~\Cref{rem:canonicity}, because they have the same lowering, therefore the identity homomorphism is a homomorphism on their lowerings, yet there is no homomorphism from \(Ŝ₁\) to \(Ŝ₂\).
-
-We can then migrate the normalized bisimulation algorithm to the symbolic setting, we will first prove an inductive construction theorem like~\cref{thm:recursive-construction}.
+The functoriality states that a homomorphism on two symbolic coalgebras induces a homomorphism of on their lowing; similarly, a bisimulation on symbolic GKAT coalgebra also induces a bisimulation on their lowering.
+However, the converse is not true, precisely because of the canonicity problem noted in~\Cref{rem:canonicity}: take the \(Ŝ₁\) and \(Ŝ₂\) in~\Cref{rem:canonicity}, because they have the same lowering, therefore the identity homomorphism is a homomorphism on their lowerings, but there is no homomorphism from \(Ŝ₁\) to \(Ŝ₂\).
+
+We can then define the symbolic equivalence algorithm, with its correctness stated in~\cref{thm:recursive-construction}.
 
@@
 \begin{proof}
-    Reduces to~\cref{thm:recursive-construction} i.e. all the above condition holds if and only if all the condition in~\cref{thm:recursive-construction} holds in the lowered coalgebra.
-\end{proof}
-
-Then for the algorithm, we can just recursively check all the conditions in~\cref{thm:symb-recursive-construction}.
+    Reduces to~\Cref{thm:recursive-construction} i.e. all the above condition holds if and only if all the condition in~\Cref{thm:recursive-construction} holds in the lowering.
+\end{proof}
+
+Similar to the non-symbolic version of the algorithm, we will first check if either of the input states is dead, then  recursively check all the conditions in~\cref{thm:symb-recursive-construction}.
+And the symbolic version of the liveness checking algorithm is also based on a graph search algorithm on all the reachable state, where we try to identify an accepting state by checking whether \(⋁ ϵ̂(s) ≡ 0\).
+The only deviation from the non-symbolic cases is that we do not follow empty transitions in the symbolic case, i.e.
+if \(δ̂(s) ≜ (b, s', p)\) and \(b ≡ 0\), we will not search \(s'\) to check the liveness of \(s'\).
+In~\Cref{sec:optimization-implementation}, we will outline a simple modification for the symbolic GKAT coalgebra generation algorithm in~\Cref{sec:symb-gkat-construction} to avoid generating empty transitions, allowing us to skip the emptiness check when deciding whether a state is dead.
 
@@
 
-Inspired by the syntax of Ocaml, the \(\mathrel{\&\!\&}\) is the logical-and operator on the language level, specifying that all four conditions in the return statements must be satisfied to return true.
-Notice just like the non-symbolic case, this algorithm can be modified to perform symbolic bisimulation of (non-normalized) GKAT coalgebra, which coincides with infinite trace equivalence~\cite{schmid_GuardedKleeneAlgebra_2021}, by letting \textsc{isDead} and \textsc{knownDead} to always return false.
-
-\section{Symbolic GKAT Coalgebra Construction}
+The symbolic trace-equivalence algorithm is presented as~\cref{alg:symb-bisim}, where \(\mathrel{\&\!\&}\) is the logical-and operator on the language level, specifying that all the conditions in~\cref{thm:symb-recursive-construction} must be satisfied for the algorithm to return true.
+
+\section{Symbolic GKAT Coalgebra Construction}\label{sec:symb-gkat-construction}
 

### 2025-01-08 main.tex
related works

@@
 
-We present several sound and complete conditions for finite-trace equivalence between states of GKAT coalgebras in~\Cref{thm:norm-bisim-correctness}, which also serves as the core correctness theorem for our equivalence-checking algorithm i.e.~\Cref{alg:bisim}.
-This algorithm mostly consists of recursively checking the conditions in~\Cref{thm:norm-bisim-correctness} with minor optimizations, which we will outline later.
+We present several sound and complete conditions for finite-trace equivalence between states of GKAT coalgebras in~\Cref{thm:recursive-construction}, which also serves as the core correctness theorem for our equivalence-checking algorithm i.e.~\Cref{alg:bisim}.
+This algorithm mostly consists of recursively checking the conditions in~\Cref{thm:recursive-construction} with minor optimizations, which we will outline later.
 
@@
         \item \(s\) reject \(α\) or transition to a dead state via \(α\) if and only if \(u\) rejects \(α\) or transition to a dead state via \(α\);  
-        \item if \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', p)\), then \(⟦s⟧_{S} = ⟦u⟧_{U}\);
+        \item if \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', p)\), then \(⟦s'⟧_{S} = ⟦u'⟧_{U}\);
         \item if \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', q)\), s.t. \(p ≠ q\), then both \(s'\) and \(t'\) are dead.
@@
 This coalgebra is also required to satisfy the disjointedness condition, i.e.
-for all states \(s ∈ S\), the conjunction of any two expressions from the set \(ϵ̂(s) ∪ \{b ∣ ∃ (b, s', p) ∈ δ(s)\}\) is equivalent to \(0\) under boolean algebra.
+for all states \(s ∈ S\), the boolean expressions that \(s\) accepts \(ϵ̂(s)\) and the boolean expressions that enables \(s\) to transition \(\{b ∣ ∃ (b, s', p) ∈ δ(s)\}\) is disjoint; also the conjunction of any two distinct expressions from the set \(ϵ̂(s) ∪ \{b ∣ ∃ (b, s', p) ∈ δ(s)\}\) is equivalent to \(0\) under boolean algebra.
 
@@
 \end{proof}
-
-Similar to the non-symbolic version of the algorithm, we will first check if either of the input states is dead, then  recursively check all the conditions in~\cref{thm:symb-recursive-construction}.
-And the symbolic version of the liveness checking algorithm is also based on a graph search algorithm on all the reachable state, where we try to identify an accepting state by checking whether \(⋁ ϵ̂(s) ≡ 0\).
-The only deviation from the non-symbolic cases is that we do not follow empty transitions in the symbolic case, i.e.
-if \(δ̂(s) ≜ (b, s', p)\) and \(b ≡ 0\), we will not search \(s'\) to check the liveness of \(s'\).
-In~\Cref{sec:optimization-implementation}, we will outline a simple modification for the symbolic GKAT coalgebra generation algorithm in~\Cref{sec:symb-gkat-construction} to avoid generating empty transitions, allowing us to skip the emptiness check when deciding whether a state is dead.
 
@@
 
-The symbolic trace-equivalence algorithm is presented as~\cref{alg:symb-bisim}, where \(\mathrel{\&\!\&}\) is the logical-and operator on the language level, specifying that all the conditions in~\cref{thm:symb-recursive-construction} must be satisfied for the algorithm to return true.
+Similar to the non-symbolic version of the algorithm, \cref{alg:symb-bisim} first check if either of the input states is dead, then  recursively check all the conditions in~\Cref{thm:symb-recursive-construction}, where \(\mathrel{\&\!\&}\) denotes the logical and operator on boolean data type.
+The symbolic version of the liveness checking algorithm is also based on a graph search algorithm on all the reachable state, where we try to identify an accepting state by checking whether \(⋁ ϵ̂(s) ≡ 0\).
+The only deviation from the non-symbolic cases is that we do not follow empty transitions in the symbolic case, i.e.
+if \(δ̂(s) ≜ (b, s', p)\) and \(b ≡ 0\), we will not search \(s'\) to check the liveness of \(s'\).
+In~\Cref{sec:optimization-implementation}, we will outline a simple modification for the symbolic GKAT coalgebra generation algorithm in~\Cref{sec:symb-gkat-construction} to avoid generating empty transitions, allowing us to skip the emptiness check when deciding whether a state is dead.
+TODO: I think we need to make sure that we are presenting the dead state checking algorithm ignoring the empty transition.
+After which we can comment it here, to claim that this is for generality.
 
@@
 
-The final piece of the puzzle is to convert any given expression into an ``equivalent'' Symbolic GKAT Automata.
-This goal can be achieved by lifting existent constructions like derivatives and Thompson's construction~\cite{schmid_GuardedKleeneAlgebra_2021,smolka_GuardedKleeneAlgebra_2020}.
-The correctness of these conversions is a consequence of  correctness of their non-symbolic counter-part, i.e.
-we will prove that the lowering as shown in~\labelcref{cons:lowering} of these constructions will yield the conventional derivative and Thompson's construction.
+The final piece of the puzzle is to convert any given expression into an \emph{equivalent} state in some symbolic GKAT coalgebra. 
+This goal can be achieved by lifting existing constructions like derivatives and Thompson's construction~\cite{schmid_GuardedKleeneAlgebra_2021,smolka_GuardedKleeneAlgebra_2020} to the symbolic setting.
+Then the correctness of non-symbolic version can be used to show the correctness of the symbolic construction i.e.
+we will prove that the lowering as shown in construction~\labelcref{cons:lowering} of these constructions will yield the conventional derivative.
+However, several other important properties, like finiteness of derivative coalgebra and the correctness of the Thompson's construction can be established using a symbolic GKAT coalgebra homomorphism from the Thompson's construction to the derivative.
+
+We introduce some new notations for the transitions of symbolic GKAT coalgebra: for a symbolic GKAT coalgebra \(Ŝ ≜ (S, δ̂, ϵ̂)\) and state \(s ∈ S\), we will use \(s \transAcc{S}{b}\) to denote \(b ∈ ϵ̂(s)\); and use \(s \transvia{b ∣ p}_S s'\) to denote \((b, s', p) ∈ δ̂(s)\).
 
@@
     \end{mathpar}
-    \caption{Symbolic Derivative}\label{fig:derivatives-rules}
+    \caption{Symbolic Derivative Coalgebra \(D̂\)}\label{fig:derivatives-rules}
 \end{figure*}
@@
 The symbolic derivative coalgebra \(D̂\), with expressions as states, is the least symbolic GKAT coalgebra (ordered by point-wise subset ordering on \(ϵ̂\) and \(δ̂\)) that satisfy the rules in~\Cref{fig:derivatives-rules}.
-Notice that the rules listed on~\Cref{fig:derivatives-rules} is very close to that of~\citeauthor{schmid_GuardedKleeneAlgebra_2021}~\cite{schmid_GuardedKleeneAlgebra_2021}.
+In more plain words, a transition is in \(D̂\) if and only if it is derivable by the rules in~\Cref{fig:derivatives-rules}.
+These rules are very close to the rules of GKAT derivative by~\citeauthor{schmid_GuardedKleeneAlgebra_2021}~\cite{schmid_GuardedKleeneAlgebra_2021}.
 This is no coincidence, as our definition exactly lowers to the definition of theirs.
@@
     \end{tabular}
-    \caption{Symbolic Thompson's Construction}\label{tab:symb-thomposon-construction}
+    \caption{Symbolic Thompson's Construction}\label{tab:symb-Thompson-construction}
 \end{table*}
@@
 Another way to construct a coalgebra from a GKAT expression is via Thompson's construction, we lift the original construction to the symbolic setting.
-A common expression to construct is a guard operation, denoted by \(⟨B|\), where \(B\) is a set of boolean expressions.
-TODO: define transition dynamics and accepting dynamics earlier.
-Concretely, this guard can be defined on both accepting dynamics and transition dynamics:
+A useful operation for Thompson's construction is the following guard operation, denoted by \(⟨B|\), where \(B\) is a set of boolean expressions: for a transition function \(δ̂\), a accepting function \(ϵ̂\), and a state \(s\),
 \begin{align*}
@@
 \end{align*}
-Notably, besides guarding transition and acceptance with different conditions in if statements and while loops, the guard operator can also be used to simulate uniform continuation.
-Specifically, we can use \(⟨ϵ̂(s)|δ(s')\) to connecting all the accepting state of \(s\) to the dynamic \(δ(s')\).
-
-With these definitions in mind, we can define symbolic Thompson's construction inductively as in~\Cref{tab:symb-thomposon-construction}, where we let \((S₁, ϵ̂₁, δ̂₁)\) and \((S₂, ϵ̂₂, δ̂₂)\) to be result of Thompson's construction for \(e₁\) and \(e₂\) respectively.
-The \(S₁ + S₂\) is the disjoint union of \(S₁\) and \(S₂\), and for any two transition dynamics \(δ₁(s₁): 𝒫(\BExp × S₁ × K)\) and \(δ₂(s₂): 𝒫(\BExp × S₂ × K)\), then \(δ₁(s₁) + δ₂(s₂): 𝒫(\BExp × (S₁ + S₂) × K)\) is defined as follows:
-\begin{align*}
-    δ₁(s₁) & + δ₂(s₂) ≜ \\
-        & \{(b, \mathrm{inj}ₗ(s₁'), p) ∣ (b, s₁', p) ∈ δ₁(s₁)\} ∪ \\
-        & \{(b, \mathrm{inj}ᵣ(s₂'), p) ∣ (b, s₂', p) ∈ δ₂(s₂)\}
+Besides guarding transition and acceptance with different conditions in if statements and while loops, the guard operator can also be used to simulate uniform continuation.
+Specifically, we can use \(⟨ϵ̂(s)|δ(s')\) to connecting all the accepting transition of \(s\) to the dynamic \(δ(s')\).
+
+With these definitions in mind, we can define symbolic Thompson's construction inductively as in~\Cref{tab:symb-Thompson-construction}, where we let \((S₁, ϵ̂₁, δ̂₁)\) and \((S₂, ϵ̂₂, δ̂₂)\) to be result of Thompson's construction for \(e₁\) and \(e₂\) respectively.
+In this table, \(S₁ + S₂\) denotes the disjoint union of \(S₁\) and \(S₂\), and for any two transition dynamics \(δ₁(s₁): 𝒫(\BExp × S₁ × K)\) and \(δ₂(s₂): 𝒫(\BExp × S₂ × K)\), we can also compose them in parallel as \(δ₁(s₁) + δ₂(s₂)\):
+\begin{align*}
+    δ₁(s₁) & + δ₂(s₂): 𝒫(\BExp × (S₁ + S₂) × K) \\*
+    δ₁(s₁) & + δ₂(s₂) ≜ \\*
+        & \{(b, \mathrm{inj}ₗ(s₁'), p) ∣ (b, s₁', p) ∈ δ₁(s₁)\} ∪ \\*
+        & \{(b, \mathrm{inj}ᵣ(s₂'), p) ∣ (b, s₂', p) ∈ δ₂(s₂)\},
 \end{align*}
@@
 
-One notable difference between the original construction~\cite{smolka_GuardedKleeneAlgebra_2020} and our construction is that we use a start state \(s^* ∈ S\), instead of a start dynamics (or pseudo-state).
+Our construction deviates from the original construction~\cite{smolka_GuardedKleeneAlgebra_2020} by using a start state \(s^* ∈ S\) instead of a start dynamics (or pseudo-state).
 This choice will make the proof of~\Cref{thm:hom-thompson-derivative} slightly easier. 
@@
 the semantics of the ``start state'' the both construction have the same preserves the trace semantics of the expression; \emph{finiteness}, i.e.
-the coalgebra generated is always finite, which means that our equivalence algorithm will eventually terminate; and finally, \emph{complexity}, the relationship between the number of reachable states and the size of the input expression, which serves as an estimated complexity of our equivalence checking algorithm.
-Turns out all of these questions can be answered by a homomorphism from symbolic Thompson's construction to the symbolic derivatives.
+the coalgebra generated is always finite, implying the termination of our equivalence algorithm; and finally, \emph{complexity}, i.e.
+the relationship between the number of reachable states and the size of the input expression, which serves as an estimated complexity of our equivalence checking algorithm.
+Turns out, all of these questions can be answered by a homomorphism from symbolic Thompson's construction to the symbolic derivatives.
 
@@
 
-\Cref{thm:hom-thompson-derivative} have several consequences, one of the more obvious one is that we can use the functoriality of the lowering operation to show the semantic equivalence of the start state in the thompson's construction and the expression in derivative.
+\Cref{thm:hom-thompson-derivative} have several consequences, one of the more obvious one is that we can use the functoriality of the lowering operation to show the semantic equivalence of the start state in the Thompson's construction and the expression in derivative.
 
@@
     By functoriality of lowering~\cref{thm:lowering-functor}, then \(h: S_{e} → D\) is a homomorphism on their lowerings. 
-    By standard theorem of universal coalgebra, homomorphism preserves the result to the final coalgebra~\cite{rutten_UniversalCoalgebraTheory_2000}, in particular, \(⟦s^*⟧^{ω}_{S_{e}} = ⟦h(s^*)⟧^{ω}_{D} = ⟦e⟧^{ω}_{D}\), where \(⟦-⟧^{ω}\) is the infinite trace semantics i.e.
-the unique map into the final GKAT coalgebra.
-
-    Finally, because infinite trace equivalence implies finite trace equivalence (\Cref{thm:inf-trace-equiv-implies-fin-trace-equiv}), \(⟦s^*⟧_{S_{e}} = ⟦e⟧_{D}\); then the correctness of derivative (\Cref{thm:derivative-correctness}) concludes our proof: \(⟦s^*⟧_{S_{e}} = ⟦e⟧_{D} = ⟦e⟧\).
+    Because homomorphism preserves semantics (\Cref{thm:hom-preserves-semantics}) \(⟦s^*⟧^{ω}_{S_{e}} = ⟦h(s^*)⟧^{ω}_{D} = ⟦e⟧^{ω}_{D}\), where \(⟦-⟧^{ω}\) is the infinite trace semantics i.e.
+the unique map into the final GKAT coalgebra \(𝒢_ω\).
+
+    Finally, because infinite trace equivalence implies finite trace equivalence (\Cref{thm:inf-trace-equiv-implies-fin-trace-equiv}) and the correctness of derivative (\Cref{thm:derivative-correctness}), we obtain the following chain of equalities:
+    \(⟦s^*⟧_{S_{e}} = ⟦e⟧_{D} = ⟦e⟧\).
 \end{proof}
@@
 
-Therefore, \(⟨s^*⟩_{Ŝ_{e}}\) will have no less state than \(⟨e⟩_{D̂}\) as a consequence of the surjectivity of \(h'\) in~\cref{thm:suj-hom-thompson-derivative}.
-Because \(Ŝ_{e}\) have more states than \(⟨s^*⟩_{Ŝ_{e}}\), which have more states than \(⟨e⟩_{D̂}\), and notice that the size of \(Ŝ_{e}\) is linear to the size of the input expression, therefore \(⟨e⟩_{D̂}\) is at most linear to the size of the expression.
-
-Although the size of the coalgebra generated by the derivative is smaller than Thompson's construction, this does not imply that the decision procedure based on derivative is always more efficient than those based on Thompson's construction.
-Crucially, the states in the derivative coalgebra \(D̂\) are expressions, and computing the next transition of the coalgebra can also be more expensive.
-Thompson's construction, on the other hand, only requires inductively going through the expression once to construct the entire coalgebra \(Ŝ_{e}\), and its states can be represented by more efficient constructs, like integers.
+Because of the surjectivity of \(h'\) in~\Cref{thm:suj-hom-thompson-derivative}, \(⟨s^*⟩_{Ŝ_{e}}\) has no fewer states than \(⟨e⟩_{D̂}\).
+And, the number of states in \(Ŝ_{e}\) is greater than \(⟨s^*⟩_{Ŝ_{e}}\), which is greater than \(⟨e⟩_{D̂}\); by induction, the number of states in \(Ŝ_{e}\) is linear to the size of the input expression, therefore the number of states in \(⟨e⟩_{D̂}\) is at most linear to the size of the expression \(e\).
+
+Although the size of the coalgebra generated by the derivative is smaller than Thompson's construction, the decision procedure based on derivative is not always more efficient than those based on Thompson's construction.
+Crucially, the states in the derivative coalgebra \(D̂\) are expressions, which is more expensive to store; and computing the next transition of the coalgebra can also be more computationally expensive.
+Whereas, Thompson's construction only requires inductively going through the expression once to construct the entire coalgebra \(Ŝ_{e}\), and its states can be represented by more efficient constructs, like integers.
 
@@
 
-\section{Future Work}
-
-Can weak symbolic coalgebra leads to a simpler completeness proof.
+\section{Related Works and Discussions}
+
+\subsection{Generic Symbolic Techniques}
+
+There are many studies that utilizes symbolic techniques to solve various problems surrounding (extended) regular expressions, and have found a wide range of real-world applications, including but not limited to low level program analysis~\cite{dallapreda_AbstractSymbolicAutomata_2015a}, list comprehension~\cite{saarikivi_FusingEffectfulComprehensions_2017}, constraint solving~\cite{stanford_SymbolicBooleanDerivatives_2021}, HTML decoding, malware fingerprinting, image blurring, location privacy~\cite{veanes_SymbolicFiniteState_2012}, regex processing, and string sanitizer~\cite{veanes_ApplicationsSymbolicFinite_2013}.
+
+Our study, on the other hand,  focus on GKAT and GKAT automata (represented as coalgebra throughout the paper). 
+Although previous works on deterministic symbolic transducer~\cite{saarikivi_FusingEffectfulComprehensions_2017,veanes_SymbolicFiniteState_2012} might seem similar to that of GKAT, the automata shape are subtly different, specifically, instead of having accept and rejecting states, states in GKAT automata will accept or reject its input.
+
+Another difference between aforementioned works and ours is that we utilize the coalgebraic theory of GKAT to streamline some of our proofs.
+In fact, we are not the first to look at symbolic transition system through the lens of coalgebra, \citeauthor{bonchi_CoalgebraicSymbolicSemantics_2009}~\cite{bonchi_CoalgebraicSymbolicSemantics_2009} have an elegant coalgebra theory surrounding symbolic transition system.
+However, instead of defining the symbolic semantics via the coalgebra theory, we opt to use the notion of lowering to connect symbolic GKAT coalgebra and GKAT coalgebra.
+This approach allows us to leverage previous correctness proofs like in~\Cref{thm:derivative-correctness}, and also enables a non-symbolic equivalence-checking algorithm as in~\cref{alg:bisim}.
+Another notable difference is that our equivalence checking also need to handle normalization, which is a unique property of GKAT not found in general coalgebra.
+
+\subsection{KAT and GKAT}
+
+GKAT is a guarded fragment of Kleene Algebra with Tests (KAT), which enjoys a symbolic algorithm~\cite{pous_SymbolicAlgorithmsLanguage_2015}.
+This algorithm by~\citeauthor{pous_SymbolicAlgorithmsLanguage_2015} generalizes the notion of derivatives also to boolean expressions, which uses binary decision diagram (BDD) as transition between symbolic expressions.
+Although some of our examples utilizes BDD to solve equivalence and inequivalence of boolean expressions, because of the structure of GKAT, our algorithm is not bounded by a particular boolean representation or solvers.
+In fact, we have demonstrated that in many scenarios, solvers like z3 and sentential decision diagram (SDD) can achieve better performance than BDD.
+
+In similar veins, KATch~\cite{moeller_KATchFastSymbolic_2024} is a symbolic solver for NetKAT~\cite{anderson_NetKATSemanticFoundations_2014} based on forwarding decision diagram, and achieved outstanding performance among state-of-the-art tools for reasoning about software-defined networks.
+GKAT also stem from the research of the research of software-defined networks~\cite{smolka_ScalableVerificationProbabilistic_2019,smolka_GuardedKleeneAlgebra_2020}, and quickly found applications in probabilistic verification~\cite{ro.zowski_ProbabilisticGuardedKAT_2023}, probabilistic program logic~\cite{gomes_KleeneAlgebraTests_2024}, networks~\cite{wasserstein_GUARDEDNETKATSOUNDNESS_2023}, and control-flow validation~\cite{zhang_CFGKATEfficientValidation_2025}.
+It would be interesting to see how our symbolic algorithm can be used to speed up these applications.
 

### 2025-01-10 main.tex
optimization

@@
 \section{Implementation}
+We implement our symbolic on-the-fly bisimulation algorithm~(\Cref{alg:symb-bisim}) in Rust along with derivative based~(\Cref{fig:derivatives-rules}) and Thompson's construction based~(\Cref{tab:symb-Thompson-construction}) algorithms for constructing symbolic GKAT coalgebras.
+The source code and benchmark suite is freely available in our repository TODO.
 
 \subsection{Optimization}\label{sec:optimization-implementation}
+\begin{definition}[blocked transition]
+    For any \(s \in S\), a transition \((b, s', p) \in \hat{\delta}_S(s)\) is blocked if \(b \equiv 0\).
+\end{definition}
+
+Notice that during bisimulation~(\Cref{alg:symb-bisim}) for some \(s \in S\) and \(u \in U\), if there is \((b,s',p) \in \hat{\delta}_S(s)\) where \(b\) is a semantically false boolean expression, then implications of form \((b \land \cdot) \not{\equiv} 0 \Rightarrow \cdot\) are trivially satisfied.
+In other words, \textit{blocked} transitions in \(S\) do not contribute to the result of bisimulation.
+The same observation holds true symmetrically for \(U\).
+So blocked transitions of a coalgebra can be safely pruned without impacting the result of bisimulation.
+
+While blocked transitions are irrelevant regarding the result of bisimulation, they can negatively impact the performance of bisimulation as the algorithm may perform many unneeded satisfiability checks.
+To prevent performance degradation due to blocked transitions, we remove them through an \textit{eager-pruning} optimization.
+Basically, in our coalgebra construction algorithms, \(\hat{\delta}\) is constructed with only transitions \((b, s, p)\) where \(b \not{\equiv} 0\).
+The coalgebras produced in this manner essentially have their blocked transitions pruned at the time of construction (eager).
+
+Eager-pruning also improves the efficiency of the Thompson's construction algorithm~(\Cref{tab:symb-Thompson-construction}) significantly.
+This is due to the fact that eager-pruning can reduce the size of \(\hat{\delta}'\) computed recursively for sub-expressions, which in turn makes computing \(\hat{\delta}\) for the overall expression faster.
 

### 2025-01-10 main.tex
merge some changes (proof of correctness in progress)

@@
 
-Besides giving us the finite-trace semantics, the normalization operation also connects the finite and infinite trace semantics, because it is an endofunctor on the category of GKAT coalgebra.
+Besides giving us the finite-trace semantics, the normalization operation also satisfy several nice properties.
+
+First, normalization preserves liveness, i.e.
+a state \(s ∈ S\) is dead (live) if and only if it is dead (live) in \(\norm(S)\), allowing us to perform liveness analysis in \(\norm(S)\) to obtain the liveness in \(S\), and vise versa.
+
+\begin{theorem}[normalization perserves liveness]
+    A state \(s ∈ S\) is dead (live) in \(S\) if and only if it is dead (live) in \(\norm(S)\).
+\end{theorem}
+
+\begin{proof}
+    By definition, every state is either dead of live, thus we will only need to show the \(⟹\) direction, and
+
+    This proof unfortunately requires us unfolding the path to accepting states.
+
+    If \(s\) is dead in \(S\), then by~\Cref{thm:dead-iff-all-reachable-dead}, for all \(α, p\), \(s \transvia{α ∣ p}_S s'\) implies \(s'\) is dead. 
+    Because \(s\) cannot accept any atom in \(S\), \(s\) will reject all atom in \(\norm(S)\); which implies that \(s\) is dead in \(\norm(S)\).
+
+    If \(s\) is live in \(S\), then there exists a walk \(s \transvia{b₁ ∣ p₁}_S s₁ \transvia{b₂ ∣ p₂}_S ⋯ \transvia{bₙ ∣ pₙ}_S sₙ\) s.t.
+\(sₙ\) is accepting, which implies every state \(sᵢ\) on the walk is live.
+    Hence, this walk also exists in \(\norm(S)\), and \(s\) is live in \(\norm(S)\).
+\end{proof}
+
+Second, normalization is an endofunctor in the category of GKAT coalgebra, this result can help us obtain sub-coalgebras of normalized coalgebra, and also connect the finite trace and infinite trace semantics.
 
@@
 
-The original algorithm for deciding GKAT equivalences~\cite{smolka_GuardedKleeneAlgebra_2020} requires the entire coalgebra to be known prior to the execution of the bisimulation algorithm; specifically, in order to compute the liveness of a state \(s\), it is necessary iterate through all its reachable states \(⟨s⟩\) to see if there are any accepting states within.
+The original algorithm for deciding GKAT equivalences~\cite{smolka_GuardedKleeneAlgebra_2020} requires the entire coalgebra to be known prior to the execution of the bisimulation algorithm; specifically, it is necessary to iterate through the entire coalgebra in order to identify the liveness of every single states, in order to perform the normalization operation.
 This limitation poses challenges to design an efficient on-the-fly algorithm for GKAT.
@@
 This algorithm mostly consists of recursively checking the conditions in~\Cref{thm:recursive-construction} with minor optimizations, which we will outline later.
+
+\begin{definition}[normalized bisimulation]\label{def:norm-bism}
+    A relation \({≃} ⊆ S × U\) on two GKAT coalgebra \(S\) and \(U\) is a normalized bisimulation if and only if for any two states \(s ≃ u\), all the following holds:
+    \begin{enumerate}
+        \item\label{itm:acc-condition} for all \(α ∈ \At\), \(δ_{S}(s, α) = \accept ⟺ δ_{U}(u, α) = \accept\);
+        \item\label{itm:rej-or-dead} if \(s\) reject \(α\) but \(u\) transitions to \(u'\), then \(u'\) is dead; similarly when \(u\) rejects \(α\);  
+        \item\label{itm:transition-bisim} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', p)\), then \(s' ≃ u'\).
+        \item\label{itm:transition-dead} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', q)\), s.t. \(p ≠ q\), then both \(s'\) and \(t'\) are dead.
+    \end{enumerate}
+\end{definition}
 
@@
     Given two GKAT coalgebra \(S\) and \(U\), and two of their elements \(s ∈ S\) and \(u ∈ U\),
-    there exists a bisimulation \({∼} ⊆ \norm(⟨s⟩) × \norm(⟨u⟩)\) s.t. \(s ∼ u\), if and only if all the following holds:
-    \begin{enumerate}
-        \item\label{itm:acc-condition} for all \(α ∈ \At\), \(δ_{S}(s, α) = \accept ⟺ δ_{U}(u, α) = \accept\);
-        \item\label{itm:rej-or-dead} \(s\) reject \(α\) or transition to a dead state via \(α\) if and only if \(u\) rejects \(α\) or transition to a dead state via \(α\);  
-        \item\label{itm:transition-bisim} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', p)\), then there exists a bisimulation \({∼'} ⊆ \norm(⟨s'⟩) × \norm(⟨u'⟩)\), s.t. \(s' ∼' u'\);
-        \item\label{itm:transition-dead} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', q)\), s.t. \(p ≠ q\), then both \(s'\) and \(t'\) are dead.
-    \end{enumerate}
+    there exists a bisimulation \({∼} ⊆ \norm(⟨s⟩) × \norm(⟨u⟩)\) s.t. \(s ∼ u\), if and only if there exists a normalized bisimulation \({≃} ⊆ ⟨s⟩ × ⟨u⟩\) s.t. \(s ≃ u\).
 \end{lemma}
@@
 \begin{proof}
-    We first prove \(⟹\) direction, recall that there exists a bisimulation \({∼} ⊆ \norm(⟨s⟩) × \norm(⟨u⟩)\) if and only if for all \(s₁ ∼ u₁\):
+    Recall that there exists a bisimulation \({∼} ⊆ \norm(⟨s⟩) × \norm(⟨u⟩)\) if and only if for all \(s₁ ∼ u₁\):
     \begin{itemize}
@@
 
-    The condition \labelcref{itm:acc-condition} holds:
+    We prove \(⟹\) direction by constructing the following normalized bisimulation from the bisimulation \({∼}\):
+    \[{≃} ≜ {∼} ∪ \{(s, u) ∣ \text{both \(s ∈ S, u ∈ U\) are dead}\}.\]
+    We then show that \({≃}\) satisfies all the condition in~\Cref{def:norm-bism}.
+
+    The condition \labelcref{itm:acc-condition} holds if \(s ∼ u\):
     \begin{align*}
@@
     \end{align*}
-
-    The condition~\labelcref{itm:rej-or-dead} holds: 
-    \begin{align*}
-        & δ_S(s, α) \text{ rejects or transition to dead states} \\
-        ⟺{}& δ_{\norm(S)}(s, α) = \reject \\
-        ⟺{}& δ_{\norm(U)}(u, α) = \reject \\
-        ⟺{}& δ_U(u, α) \text{ rejects or transition to dead states}.
-    \end{align*}
-
-    The condition~\labelcref{itm:transition-bisim} holds, by case analysis on the liveness of \(s'\) and \(u'\).
+    The condition \labelcref{itm:acc-condition} also holds if \(s, u\) are both dead, because both can never accepts any atoms.
+
+    The condition~\labelcref{itm:rej-or-dead} holds when \(s ∼ u\): 
+    \begin{align*}
+        & δ_S(s, α) = \reject \text{ and } δ_U(u, α) = (u', p)\\
+        ⟹{}& δ_{\norm(S)}(s, α) = \reject \text{ and } δ_U(u, α) = (u', p) \\
+        ⟹{}& δ_{\norm(U)}(u, α) = \reject \text{ and } δ_U(u, α) = (u', p)\\
+        ⟹{}& \text{\(u'\) is dead}
+    \end{align*}
+    Similarly, when \(u'\) reject \(α\) and \(s\) transitions to \(s'\).
+    The condition~\labelcref{itm:rej-or-dead} holds when \(s, u\) are both dead because both of them can only transition to dead state by~\Cref{thm:dead-iff-all-reachable-dead}.
+
+    The condition~\labelcref{itm:transition-bisim} holds when \(s ∼ u\).
     First note that \(s'\) and \(u'\) has to be both live or both dead: because \(δ_S(s, α) = (s', p)\), then \(\norm(δ_S)(s', α)\) can either be rejection or \((s',p)\), and so is \(\norm(δ_U)(u', α)\):

### 2025-01-13 main.tex
ok

@@
 Although the size of the coalgebra generated by the derivative is smaller than Thompson's construction, the decision procedure based on derivative is not always more efficient than those based on Thompson's construction.
-Crucially, the states in the derivative coalgebra \(D̂\) are expressions, which is more expensive to store; and computing the next transition of the coalgebra can also be more computationally expensive.
+Crucially, the states in the derivative coalgebra \(D̂\) are expressions, which are more expensive to store; and computing the next transition of the coalgebra can also be more computationally expensive.
 Whereas, Thompson's construction only requires inductively going through the expression once to construct the entire coalgebra \(Ŝ_{e}\), and its states can be represented by more efficient constructs, like integers.
@@
 This is due to the fact that eager-pruning can reduce the size of \(\hat{\delta}'\) computed recursively for sub-expressions, which in turn makes computing \(\hat{\delta}\) for the overall expression faster.
+
+As mentioned in \Cref{sec:symb-gkat-construction}, our implementation of Thompson's construction uses start dynamics (pseudo-state) instead of an explicit start state.
+This reduces the number of indirections needed to access the behaviors \(\hat{\epsilon}(s^*) \) and \(\hat{\delta}(s^*)\) of start state \(s^*\).
+More specifically, we track the behaviors \(\epsilon^*\) and \(\delta^*\) which characterize \(\hat{\epsilon}(s^*) \) and \(\hat{\delta}(s^*)\) respectively.
+The start state \(s^*\) does not appear explicitly during coalgebra construction.
+After the construction algorithm terminates, we generate a fresh state \(s^*\) and set \(\hat{\epsilon}(s^*) = \epsilon^*\) and \(\hat{\delta}(s^*) = \delta^*\).
+The final coalgebra that our implementation produces is equivalent to what would have been obtained from the original algorithm (\Cref{tab:symb-Thompson-construction}) modulo consistent renaming of states.
 

### 2025-01-13 main.tex
added performance evaluation

@@
 \subsection{Performance}\label{sec:performance-implementation}
+Our experiments are performed on a laptop with an Apple M4 Pro CPU and 24 GB RAM.
+\Cref{tab:benchmark} presents the results of equivalence checking based on derivatives (DRV), Thompson's construction (TC) and SymKAT~\cite{pous_SymbolicAlgorithmsLanguage_2015}.
+Each benchmark consists of 50 expression pairs with properties described by the benchmark's name.
+For example, expressions in E250B5P10RD have around 250 primitive actions in total (E250), a maximum boolean expression size of 5 (B5), 10 possible unique primitive actions (P10) and are completely random (RD).
+Benchmarks with the suffix EQ have expression pairs that are known to be equivalent.
+The time columns measure the total time used (in seconds) for checking all expression pairs in the benchmarks.
+The mem columns measure the peak memory used (in MB).
+
+It is clear from these results that our algorithms (DRV and TC) are more efficient than SymKAT in terms of both time and memory usage.
+Moreover, our Thompson's construction (TC) based algorithm is the only one which successfully completed all benchmarks without timeout.
+
+\begin{table*}
+\centering
+\begin{tabular}{l l l l l l l}
+    Benchmark & Time (DRV) & Time (TC) & Time (SK) & Mem (DRV) & Mem (TC) & Mem (SK) \\
+    \hline
+    E250B5P10RD    & 0.26    & 0.25  & 16.99   & 15.36  & 14.76  & 329.48  \\
+    E250B5P10EQ    & 0.28    & 0.25  & 2.99    & 15.56  & 14.95  & 100.48  \\
+    E500B5P50RD    & 0.30    & 0.28  & 113.88  & 16.26  & 15.49  & 1560.54 \\
+    E500B5P50EQ    & 0.35    & 0.28  & 14.06   & 17.97  & 15.54  & 546.91  \\
+    E1000B10P100RD & 0.43    & 0.37  & Timeout & 18.36  & 21.41  & N/A     \\
+    E1000B10P100EQ & 0.61    & 0.35  & 77.83   & 21.49  & 17.66  & 5822.66 \\
+    E2000B20P200RD & 3.25    & 2.62  & Timeout & 241.07 & 283.45 & N/A     \\
+    E2000B20P200EQ & Timeout & 46.45 & Timeout & N/A    & 638.00 & N/A     \\
+\end{tabular}
+\caption{Benchmarks}\label{tab:benchmark}
+\end{table*}
 

### 2025-01-14 main.tex
small change

@@
 Our experiments are performed on a laptop with an Apple M4 Pro CPU and 24 GB RAM.
-\Cref{tab:benchmark} presents the results of equivalence checking based on derivatives (DRV), Thompson's construction (TC) and SymKAT~\cite{pous_SymbolicAlgorithmsLanguage_2015}.
+\Cref{tab:benchmark} presents the results of equivalence checking based on derivatives (DRV), Thompson's construction (TC) and SymKAT~\cite{pous_SymbolicAlgorithmsLanguage_2015} (SK).
 Each benchmark consists of 50 expression pairs with properties described by the benchmark's name.

### 2025-01-15 main.tex
proof, WIP

@@
 \newcommand{\transAcc}[2]{⇒_{#1} #2}
+\newcommand{\transRej}[2]{↓_{#1} #2}
 
@@
 
-\begin{theorem}[normalization perserves liveness]
+\begin{theorem}[normalization perserves liveness]\label{thm:norm-perserve-liveness}
     A state \(s ∈ S\) is dead (live) in \(S\) if and only if it is dead (live) in \(\norm(S)\).
@@
 
-\begin{definition}[normalized bisimulation]\label{def:norm-bism}
-    A relation \({≃} ⊆ S × U\) on two GKAT coalgebra \(S\) and \(U\) is a normalized bisimulation if and only if for any two states \(s ≃ u\), all the following holds:
+\begin{definition}[normalizing bisimulation]\label{def:norm-bisim}
+    Given two GKAT coalgebra \(S\) and \(U\), and a bisimulation \({∼} ⊆ \norm(S) × \norm(U)\), a normalizing bisimulation \({≈} ⊆ S × U\) is a GKAT coalgebra that contains \(∼\) with the following transition function:
+    \[
+        δ_≈((s, u), α) ≜ \begin{cases}
+            \accept & \begin{aligned}
+                & \text{if } δ_S(s, α) = \accept \\[-3px]
+                & \text{and } δ_U(s, α) = \accept
+            \end{aligned}\\[5px]
+            ((s', u'), p) & 
+                \begin{aligned}
+                    & \text{if } s \transvia{α ∣ p}_{S} s' \\[-3px]
+                    & \text{and } u \transvia{α ∣ p}_{U} u' 
+                \end{aligned}\\[5px]
+            \reject & \text{otherwise}
+        \end{cases}
+    \]
+    We call \(∼\) \emph{the underlying bisimulation} of \(≈\).
+\end{definition}
+
+\begin{lemma}
+    Given a bisimulation \({∼} ⊆ \norm(S) × \norm(U)\), the following relation \({∼_d} ⊆ S × U\):
+    \[{∼_d} ≜ {∼} ∪ \{(s, u) ∣ \text{\(s\) and \(u\) are dead in \(S\) and \(U\)}\},\]
+    form a GKAT coalgebra the transition function \(δ_≈\) in~\Cref{def:norm-bisim}
+\end{lemma}
+
+\begin{proof}
+    We will need to show \(∼_d\) is closed under the transition of \(δ_{≈}\) in~\Cref{def:norm-bisim}.
+    Specifically, if \(s ∼_d u\) and \(s \transvia{α ∣ p}_{S} s'\) and \(u \transvia{α ∣ p} u'\), then \(s' ∼_d u'\).
+    By case analysis on the liveness of \(s'\) and \(u'\) in \(S\) and \(U\):
+    \begin{itemize}
+        \item If both \(s'\) and \(u'\) are dead, then \(s ∼_d u\) by definition of \(∼_d\).
+        \item If both \(s'\) and \(u'\) are live, then \(s, s'\) and \(u, u'\) are live in \(S\) and \(U\) respectively. 
+        Thus, \(s \transvia{α ∣ p}_{\norm(S)} s'\), and \(u \transvia{α ∣ p}_{\norm(U)} u'\).
+        Therefore, \(s' ∼ u'\) by definition of bisimulation on \(\norm(S)\) and \(\norm(U)\); and \(s ∼_d u'\).
+        \item If \(s'\) is live but \(u'\) is dead, then \(s\) is live, and \(s ∼ u\).
+        However, \(s \transvia{α ∣ p}_{\norm(S)} s'\) yet \(u \transRej{\norm(U)}{u}\), which violates the condition for bisimulation, thus this case cannot happen.
+        \item If \(u'\) is live but \(s'\) is dead, this case also cannot happen similar to the previous case.
+    \end{itemize}
+\end{proof}
+
+\begin{lemma}\label{thm:norm-bisim-exists-unique}
+    For any \(∼\) a normalizing bisimulation with \(∼\) as underlying bisimulation exists and is unique.
+\end{lemma}
+
+\begin{proof}
+    TODO: sub-coalgebra form a complete lattice.
+\end{proof}
+
+\begin{lemma}\label{thm:norm-bisim-not-bisim-then-component-dead}
+    Every state in \((s, u) ∈ {≈} ⊆ S × U\) but not in \({∼} ⊆ \norm(S) × \norm(U)\) then \(s\) is dead in \(S\), \(u\) is dead in \(U\).
+\end{lemma}
+
+\begin{proof}
+    Because \(∼_d\) is a GKAT coalgebra with transition \(δ_≈\) and also contains all the states in \(∼\), therefore \({≈} ⊑ {∼_d}\).
+    Thus, for all \((s, u) ∈ {≈}\) but not in \(∼\), \(s\) and \(u\) must be dead in \(S\) and \(U\).
+\end{proof}
+
+\begin{lemma}\label{thm:live-in-component-then-live-in-norm-bisim}
+    Given a normalizing bisimulation \(≈\), if a state \((s, u)\) is live in \(≈\), then \(s\) and \(u\) are live in \(S\) and \(U\) respectively, which also implies that \((s, u)\) is live in \(∼\).
+\end{lemma}
+
+\begin{proof}
+    TODO: by preservation of accepting state and liveness chain.
+\end{proof}
+
+\begin{lemma}
+    Every state in \((s, u) ∈ {≈} ⊆ S × U\) but not in \({∼} ⊆ \norm(S) × \norm(U)\) then \((s, u)\) is dead in \(≈\).
+\end{lemma}
+
+\begin{proof}
+    By~\Cref{thm:norm-bisim-not-bisim-then-component-dead}, \(s\) is dead in \(S\), then by contrapositive of~\Cref{thm:live-in-component-then-live-in-norm-bisim}, we obtain that \((s, u)\) is dead in \(≈\).
+\end{proof}
+
+\begin{theorem}[Correctness]
+    Given two states in GKAT coalgebras \(s ∈ S\) and \(u ∈ ,U\), there exists a normalizing bisimulation \({≈} ⊆ S × U\) s.t. \(s ≈ u\), if and only if \(⟦s⟧_S = ⟦u⟧_U\).
+\end{theorem}
+
+\begin{proof}
+    \(⟹\) direction, because normalization perserves carrier set, therefore \((s, u) ∈ \norm(≈)\).
+    And because \(\norm(≈)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\), \(⟦s⟧_S = ⟦u⟧_U\).
+
+    \(⟸\) direction, because \(⟦s⟧_S = ⟦u⟧_U\), then by~\cref{thm:norm-bisim-correctness}, then there a bisimulation \(∼ ⊆ \norm(S) × \norm(U)\) s.t. \(s ∼ u\). 
+    Because~\Cref{thm:norm-bisim-exists-unique}, therefore there exists a normalizing bisimulation \(≈\) with \(∼\) as its underlying bisimulation. 
+    Because the carriers \({∼} ⊆ {≈}\), therefore \(s ∼ u\) implies \(s ≈ u\).
+\end{proof}
+
+\begin{theorem}[Recursive Construction]
+    Given a relation \({≈} ⊆ S × U\), \(≈\) is a carrier of some normalizing bisimulation if and only if all the following condition are satisfied: for all pair of states \(s ≈ u\),
     \begin{enumerate}
-        \item\label{itm:acc-condition} for all \(α ∈ \At\), \(δ_{S}(s, α) = \accept ⟺ δ_{U}(u, α) = \accept\);
+        \item\label{itm:acc-condition} for all \(α ∈ \At\), 
+        \(s \transAcc{S}{α}\) if and only if \(u \transAcc{U}{α}\);
         \item\label{itm:rej-or-dead} if \(s\) reject \(α\) but \(u\) transitions to \(u'\), then \(u'\) is dead; similarly when \(u\) rejects \(α\);  
-        \item\label{itm:transition-bisim} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', p)\), then \(s' ≃ u'\).
-        \item\label{itm:transition-dead} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', q)\), s.t. \(p ≠ q\), then both \(s'\) and \(t'\) are dead.
+        \item\label{itm:transition-bisim} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', p)\), then \(s' ≈ u'\).
+        \item\label{itm:transition-dead} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', q)\), s.t. \(p ≠ q\), then both \(s'\) and \(u'\) are dead.
     \end{enumerate}
-\end{definition}
-
-\begin{lemma}[bisimulation between dead states]\label{thm:bisim-between-dead}
-    Given two dead states \(s ∈ S\) and \(u ∈ U\), then the singleton bisimulation \({∼} ⊆ \norm(S) × \norm(U)\):
-    \begin{mathpar}
-        {∼} ≜ \{(s,u)\} \and 
-        δ_{∼}((s,u), α) ≜ \reject
-    \end{mathpar}
-    is a bisimulation between \(\norm(S)\) and \(\norm(U)\).
-\end{lemma}
-
-\begin{proof}
-    By computation
-\end{proof}
-
-\begin{lemma}\label{thm:recursive-construction-lemma}
-    Given two GKAT coalgebra \(S\) and \(U\), and two of their elements \(s ∈ S\) and \(u ∈ U\),
-    there exists a bisimulation \({∼} ⊆ \norm(⟨s⟩) × \norm(⟨u⟩)\) s.t. \(s ∼ u\), if and only if there exists a normalized bisimulation \({≃} ⊆ ⟨s⟩ × ⟨u⟩\) s.t. \(s ≃ u\).
-\end{lemma}
-
-\begin{proof}
-    Recall that there exists a bisimulation \({∼} ⊆ \norm(⟨s⟩) × \norm(⟨u⟩)\) if and only if for all \(s₁ ∼ u₁\):
-    \begin{itemize}
-        \item for all results \(r ∈ \{\accept, \reject\}\): \(δ_{\norm(S)}(s₁, α) = r ⟺ δ_{\norm(U)}(u₂, α) = r\);
-        \item otherwise, let \((s₂, p) ≜ δ_{\norm(S)}(s₁, α)\) and \((u₂, q) ≜ δ_{\norm(u)}(u₁, α)\), then \(p = q\) and \(s₂ ∼ u₂\)
-    \end{itemize}
-
-    We prove \(⟹\) direction by constructing the following normalized bisimulation from the bisimulation \({∼}\):
-    \[{≃} ≜ {∼} ∪ \{(s, u) ∣ \text{both \(s ∈ S, u ∈ U\) are dead}\}.\]
-    We then show that \({≃}\) satisfies all the condition in~\Cref{def:norm-bism}.
-
-    The condition \labelcref{itm:acc-condition} holds if \(s ∼ u\):
+\end{theorem}
+
+\begin{proof}
+    To show the \(⟹\) direction, assuming there exists a normalizing bisimulation \(≈\), then it would satisfy all the conditions. 
+    We first note that the pairs \((s, u) ∈ {≈}\) can satisfy one of two properties (or both), either \(s\) and \(u\) are both dead in \(S\) and \(U\), or \((s, u)\) is in the underlying bisimulation \(∼\).
+    Then we will verify that all the condition is satisfied both when \(s ∼ u\) and when \(s, u\) are dead.
+
+    The condition~\labelcref{itm:acc-condition} holds if \(s ∼ u\):
     \begin{align*}
-        δ_S(s, α) = \accept 
-        ⟺{}& δ_{\norm(S)}(s, α) = \accept \\
-        ⟺{}& δ_∼((s,u), α) = \accept \\ 
-        ⟺{}& δ_{\norm(U)}(u, α) = \accept \\
-        ⟺{}& δ_U(u, α) = \accept
+        s \transAcc{S}{α}
+        \text{ iff }& s \transAcc{\norm(S)}{α} \\
+        \text{ iff }& (s, u) \transAcc{∼}{α} \\ 
+        \text{ iff }& u \transAcc{\norm(U)}{α} \\
+        \text{ iff }& u \transAcc{U}{α};
     \end{align*}
-    The condition \labelcref{itm:acc-condition} also holds if \(s, u\) are both dead, because both can never accepts any atoms.
+    condition~\labelcref{itm:acc-condition} also holds if \(s\) and \(u\) are dead in \(S\) and \(U\), because both can never accept any atoms.
 
@@
 
-    The condition~\labelcref{itm:transition-bisim} holds when \(s ∼ u\).
-    First note that \(s'\) and \(u'\) has to be both live or both dead: because \(δ_S(s, α) = (s', p)\), then \(\norm(δ_S)(s', α)\) can either be rejection or \((s',p)\), and so is \(\norm(δ_U)(u', α)\):
-    \begin{align*}
-        s' \text{ is live} 
-        & ⟺ δ_{\norm(S)}(s, α) = (s', p) \\
-        & ⟺ δ_{\norm(U)}(u, α) = (u', p) \\
-        & ⟺ u' \text{ is live}.
-    \end{align*}
+    The condition~\labelcref{itm:transition-bisim} holds because when \(s \transvia{α ∣ p}_S s'\) and \(u \transvia{α ∣ p}_U u'\), then \((s, u) \transvia{α ∣ p}_{≈} (s', u')\) by definition of normalizing bisimulation (\Cref{def:norm-bisim}).
+
+    The condition~\labelcref{itm:transition-dead} holds, let \(δ_S(s, α) = (s', p)\) and \(δ_U(u, α) = (u', q)\) and \(p ≠ q\). 
     \begin{itemize}
-        \item If both \(s'\) and \(u'\) are live, then \(s' ∼ u'\). By~\cref{thm:sub-coalg-preserve-bisim}, the bisimulation \(∼'\) is just \(∼\) restricted to \(⟨s'⟩\) and \(⟨u'⟩\).
-        \item If both \(s'\) and \(u'\) are dead, then \(∼'\) can just be the singleton relation, according to~\cref{thm:bisim-between-dead}.
+        \item If \(s ∼ u\), then we can show condition~\labelcref{itm:transition-dead} by case analysis on the liveness of \(s'\) and \(u'\):
+        if \(s', u'\) are not both dead, then the condition of \({∼} ⊆ \norm(S) × \norm(U)\) cannot be satisfied, hence \((s', u') ∉ {∼}\), which by~\Cref{thm:norm-bisim-not-bisim-then-component-dead} means that \(s'\) and \(u'\) are both dead, contradicting the premise.
+        Therefore, \(s'\) and \(u'\) needs to be both dead.
+        \item If \(s\) and \(u\) are both dead, then by~\Cref{thm:dead-iff-all-reachable-dead}, both \(s\) and \(u\) can only transition to dead state, hence \(s'\) and \(u'\) are both dead.
     \end{itemize}
 
-    The condition~\labelcref{itm:transition-dead} holds: by the proof of condition~\labelcref{itm:transition-bisim}, \(s'\) and \(u'\) has to be either both live or both dead; if they are both live, then there cannot be a element in \(G(∼)\) that can project to \((s', p)\) under \(π₁\) but projects to \((t', q)\) under \(π₂\).
-Thus both \(s'\) and \(t'\) has to be dead.
-
-    We then show the \(⟸\) direction, for arbitrary \(s' ∈ S\) and \(u' ∈ U\), we use \(≡_{s', u'}\) to denote the maximal bisimulation between \(\norm(⟨s'⟩)\) and \(\norm(⟨u'⟩)\).
-    \begin{align*}
-        {∼'} ≜ ⋃ \{≡_{s', u'} & ∣ ∃ α ∈ \At, p ∈ K, \\*
-            & δ_{\norm(S)}(s, α) = (s', p) \\*
-            & \text{ and } δ_{\norm(U)}(u, α) = (u', p)\}.
-    \end{align*}
-    For all the \(s'\) and \(u'\) in the above definition, \(⟨s'⟩ ⊑ ⟨s⟩\) and \(⟨u'⟩ ⊑ ⟨u⟩\), therefore by~\Cref{thm:norm-sub-coalg}, \(\norm(⟨s'⟩) ⊑ \norm(⟨s⟩)\) and \(\norm(⟨u'⟩) ⊑ \norm(⟨u⟩)\). 
-    By~\Cref{thm:sub-coalg-preserve-bisim}, every \(≡_{s', u'}\) is a bisimulation between \(\norm(⟨s⟩)\) and \(\norm(⟨t⟩)\), and because bisimulation is closed under arbitrary union~\cite{rutten_UniversalCoalgebraTheory_2000}, \(∼'\) is a bisimulation between \(\norm(⟨s⟩)\) and \(\norm(⟨t⟩)\).
-
-    To obtain the desired bisimulation \({∼}\) between \(\norm(⟨s⟩)\) and \(\norm(⟨u⟩)\), we add the pair \((s, t)\) to \(∼'\), 
-    \[{∼} ≜ {∼'} ∪ \{(s, u)\},\] 
-    with the following transition \(δ_{∼}\): for all \(α ∈ \At\),
-    \begin{itemize}[nosep]
-        \item if \(δ_{\norm(S)}(s, α) = δ_{\norm(U)}(u, α) = \accept\), then \(δ_{∼}((s, u), α) ≜ \accept\);
-        \item if \(δ_{\norm(S)}(s, α) = δ_{\norm(U)}(u, α) = \reject\), then \(δ_{∼}((s, u), α) ≜ \reject\);
-        \item if \(δ_{\norm(S)}(s, α) = (s', p)\) and \(δ_{\norm(U)}(u, α) = (u', p)\), then \(δ_{∼}((s, u), α) = ((s', u'), p)\);
-        \item for all \((s', u') ∈ {∼}'\) that is not equal to \((s, u)\), we let \(δ_∼\) inherits the transition of \(δ_{∼'}\), i.e. \(δ_{∼}((s', u'), α) = δ_{∼'}((s', u'), α)\)
+    Then we prove the \(⟸\) direction.
+    TODO: by constructing a \(≈\) and show that \(\norm(≈)\) is a bisimulation. Maybe using the maximal bisimulation?
+\end{proof}
+
+\begin{lemma}\label{thm:bisim-between-dead}
+    Given any two dead states \(s ∈ S\) and \(u ∈ U\), and the maximal bisimulation \({≡} ⊆ \norm(S) × \norm(U)\), then we have \(s ≡ u\).
+\end{lemma}
+
+\begin{theorem}[Equivalence Relation]
+    Given two states in GKAT coalgebra \(s ∈ S\) and \(u ∈ U\), there exists a normalizing bisimulation \(≈\) between \(S\) and \(U\) s.t.
+\(s ≈ u\), if and only if there exists a normalizing bisimulation that is an \emph{equivalence relation} \(≊\) in \(S + U\) s.t.
+\(s ≊ u\).
+\end{theorem}
+
+\begin{proof}
+    First, recall that by~\Cref{thm:} \(\norm(≈)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\).
+    Because \(S ⊑ S + U\) and \(U ⊑ S + U\) and normalization preserves sub-coalgebra~\Cref{thm:norm-sub-coalg}, we have \(\norm(S) ⊑ \norm(S + U)\) and \(\norm(U) ⊑ \norm(S + U)\); and by~\Cref{thm:sub-coalg-preserve-liveness}, \(s ∈ S\) is dead in \(S + U\) if and only if it is dead in \(S\), and similarly for \(u ∈ U\).
+
+    The \(⟹\) direction. 
+    Because sub-coalgebra preserves bisimulation (\Cref{thm:sub-coalg-preserve-bisim}), then \(\norm(≈)\) is a bisimulation between \(\norm(S + U)\) and itself.
+    We consider \(≡\) to be the maximal bisimulation between \(\norm(S + U)\) and itself, then \(≡\) is an equivalence relation~\cite[]{}.
+
+    Let \(≊\) be the normalizing bisimulation generated by \(≡\), then \(≊\) has the same carrier as \(≡\).
+    All the \((s, u) ∈ {≊}\) are either \(s ≡ u\) or \(s\) and \(u\) are dead in \(S\) and \(U\), which implies that both are dead in \(S + U\).
+    However, \(≡\) already contains all the dead pairs (\cref{thm:bisim-between-dead}), therefore for all \((s, u) ∈ {≊}\), they are also in \(≡\), implying \({≡} ⊇ {≊}\).
+    Finally, by definition of normalizing bisimulation (\Cref{def:norm-bisim}), \({≡} ⊆ {≊}\).
+    Therefore, \(≡\) and \(≊\) share the same carrier and \(≊\) is a normalizing bisimulation that is also an equivalence relation, which also contains \((s, u)\):
+    \[(s, u) ∈ {≈} = \norm(≈) ⊆ {≡} ⊆ {≊},\]
+    where all the \(=\) and \(⊆\) are orders on the carrier set, instead of the coalgebra.
+
+    The \(⟸\) direction can be split into two cases, by whether \((s, u)\) is in the underlying bisimulation of \(≊\), which we denote as \(≃\): 
+    \begin{itemize}
+        \item If \(s ≃ u\), then we can reflect \(≃\)  through \(\norm(S) ⊑ \norm(S + U)\) and \(\norm(U) ⊑ \norm(S + U)\) (\Cref{thm:sub-coalg-preserve-bisim}) to obtain a bisimulation \(∼ ⊆ \norm(S) × \norm(U)\) and \(s ∼ u\).
+        Taking the normalizing bisimulation over \(∼\) will yield a desired normalizing bisimulation \(≈\) s.t. \(s ≈ u\).
+        \item If \((s, u)\) are not in \(≃\), then they are both dead in \(S + U\) (\Cref{thm:norm-bisim-not-bisim-then-component-dead}), which implies \(s\) and \(u\) are respectively dead in \(S\) and \(U\) (\Cref{thm:sub-coalg-preserve-liveness}).
+        Then we take \(≈\) to be normalizing bisimulation generated form the maximal bisimulation between \(S\) and \(U\), which contains \(s\) and \(u\) (\Cref{thm:bisim-between-dead}).
+        \qedhere
     \end{itemize}
-    \emph{if \(δ_{∼}\) is well-defined}, then we can verify that \(∼\) is indeed a bisimulation between \(\norm(⟨s⟩)\) and \(\norm(⟨u⟩)\) where \(s ∼ u\). 
-    We show the slightly more complicated case: \(δ_{\norm(S)}(s, α) = (s', p)\) and \(δ_{\norm(U)}(u, α) = (u', p)\) implies \(s' ∼ u'\) as an example. 
-    By condition~\labelcref{itm:transition-bisim}, there exists a bisimulation \(∼_{s', u'} ⊆ \norm(⟨s'⟩) × \norm(⟨u'⟩)\), and because \(≡_{s', u'}\) is the maximal bisimulation between \(\norm(⟨s'⟩)\) and \(\norm(⟨u'⟩)\),
-    \[(s', u') ∈ {∼_{s', u'}} ⊆ {≡_{s', u'}} ⊆ {∼}.\]
-
-    Finally, we demonstrate that \(δ_∼\) is well-defined by leveraging the conditions in~\Cref{thm:recursive-construction}. 
-    Specifically, we will show that the definition of \(δ_∼\) covers all the possible cases, by case analysis on the result of \(δ_S\): for all \(α ∈ \At\),
-    \begin{itemize}
-        \item If \(δ_S(s, α) = \accept\), then by condition~\labelcref{itm:acc-condition}, \(δ_U(u, α) = \accept\); therefore \[δ_{\norm(S)}(s, α) = δ_{\norm(U)}(u, α) = \accept.\]
-        \item If \(δ_S(s, α)\) transitions to a dead state or reject, then by condition~\labelcref{itm:rej-or-dead} \(δ_U(u, α)\) will also transition to a dead state or reject, then \[δ_{\norm(S)}(s, α) = δ_{\norm(U)}(u, α) = \reject.\]
-        \item If \(δ_S(s, α) = (s', p)\) and \(s'\) is live, then \(δ_U(u, α) = (u', p)\) necessarily holds, otherwise it would violate one of conditions~\labelcref{itm:acc-condition,itm:rej-or-dead,itm:transition-dead}.
-
-        By condition~\labelcref{itm:transition-bisim}, there exists a bisimulation \(∼_{s', u'}\) between \(\norm(⟨s'⟩)\) and \(\norm(⟨u'⟩)\) s.t.
+\end{proof}
+
+% % \begin{definition}[Normalizable Bisimulation]
+% %     A relation \({≃} ⊆ S × U\) is a normalizable bisimulation if it has a GKAT coalgebra structure \(δ_≃\), and \(\norm(≃)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\).
+% % \end{definition}
+
+% % \begin{lemma}[Preservation of Liveness]
+% %     For any normalizable bisimulation \({≃} ⊆ S × U\) if \(s ≃ u\), then \(s\) is live in \(S\) if and only if \(u\) is live in \(U\).
+% % \end{lemma}
+
+% % \begin{proof}
+% %     Because \(\norm(≃)\) has two projection homomorphism to \(\norm(S)\) and \(\norm(U)\), then 
+% % \end{proof}
+
+% % \begin{theorem}[Correctness]
+% %     If there exists a normalizable bisimulation \({≃} ⊆ S × U\) s.t. \(s ∼ u\) for \(s ∈ S\) and \(u ∈ U\), then \(⟦s⟧_S = ⟦u⟧_U\).
+% % \end{theorem}
+
+% % \begin{theorem}[Construction]
+% %     Given two states \(s ∈ S, u ∈ U\) then there exists a normalizable bisimulation \({≃'} ⊆ S × U\) if and only if there exists a relation \({≃} ⊆ S × U\) that satisfy the following conditions:
+% %     \begin{enumerate}
+% %         \item\label{itm:acc-condition} for all \(α ∈ \At\), \(δ_{S}(s, α) = \accept ⟺ δ_{U}(u, α) = \accept\);
+% %         \item\label{itm:rej-or-dead} if \(s\) reject \(α\) but \(u\) transitions to \(u'\), then \(u'\) is dead; similarly when \(u\) rejects \(α\);  
+% %         \item\label{itm:transition-bisim} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', p)\), then \(s' ≃ u'\).
+% %         \item\label{itm:transition-dead} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', q)\), s.t. \(p ≠ q\), then both \(s'\) and \(t'\) are dead.
+% %     \end{enumerate}
+% % \end{theorem}
+
+% % \begin{proof}
+% %     The \(⟸\) direction can be shown that 
+% % \end{proof}
+
+% % \begin{definition}[normalizing bisimulation]\label{def:norm-bism}
+% %     A relation \({≃} ⊆ S × U\) on two GKAT coalgebra \(S\) and \(U\) is a normalizing bisimulation if and only if for any two states \(s ≃ u\), all the following holds:
+% %     \begin{enumerate}
+% %         \item\label{itm:acc-condition} for all \(α ∈ \At\), \(δ_{S}(s, α) = \accept ⟺ δ_{U}(u, α) = \accept\);
+% %         \item\label{itm:rej-or-dead} if \(s\) reject \(α\) but \(u\) transitions to \(u'\), then \(u'\) is dead; similarly when \(u\) rejects \(α\);  
+% %         \item\label{itm:transition-bisim} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', p)\), then \(s' ≃ u'\).
+% %         \item\label{itm:transition-dead} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', q)\), s.t. \(p ≠ q\), then both \(s'\) and \(t'\) are dead.
+% %     \end{enumerate}
+% %     And \(≃\) can be endowed with the following GKAT coalgebra structure:
+% %     \[
+% %         δ_≃((s, u), α) ≜ \begin{cases}
+% %             \accept & \begin{aligned}
+% %                 & \text{if } δ_S(s, α) = \accept \\[-3px]
+% %                 & \text{and } δ_U(s, α) = \accept
+% %             \end{aligned}\\[5px]
+% %             ((s', u'), p) & 
+% %                 \begin{aligned}
+% %                     & \text{if } s \transvia{α ∣ p}_{S} s' \\[-3px]
+% %                     & \text{and } u \transvia{α ∣ p}_{U} u' 
+% %                 \end{aligned}\\[5px]
+% %             \reject & \text{otherwise}
+% %         \end{cases}
+% %     \]
+% % \end{definition}
+
+% \begin{lemma}[Perservation of Liveness]\label{thm:norm-bisim-perserve-liveness}
+%     For any normalizing bisimulation \(≃\) and a pair of states \(s ≃ u\), \((s, u)\) is live in \(≃\) if and only if \(s\) is live in \(S\), and similarly for \(u\) in \(U\). 
+%     This implies that if \(s ≃ u\), then \(s\) is live in \(S\) if and only if \(u\) is live in \(U\).
+% \end{lemma}
+
+% \begin{proof}
+%     First by definition of normalizing bisimulation that a state \(s\) is accepting in \(S\) if and only if \((s, u)\) is accepting in \(≃\).
+
+%     Then, to prove the preservation of liveness, we prove the preservation of live chain, i.e.
+there exists a chain to accepting state in \(≃\): \((s₀, u₀) \transvia{α₁ ∣ p₁} (s₁, u₁) \transvia{α₂ ∣ p₂} (s₂, u₂) \transvia{α₂ ∣ p₂} ⋯ \transvia{αₙ ∣ pₙ} (sₙ, uₙ)\) if and only if there exists a similar accepting chain in \(S\): \(s₀ \transvia{α₁ ∣ p₁} s₁ \transvia{α₂ ∣ p₂} s₂ \transvia{α₂ ∣ p₂} ⋯ \transvia{αₙ ∣ pₙ} sₙ\).
+
+%     Because every state on an accepting chain is live, all we need to prove is that for two states \(s ≃ u\),
+%     \begin{enumerate}
+%         \item if \((s', u')\) is a live state in \(≃\) and \((s, u) \transvia{α ∣ p}_≃ (s', u')\), then \(s \transvia{α ∣ p}_S s'\)
+%         \item if \(s'\) is a live state in \(S\) and \(s \transvia{α ∣ p} s'\), then \((s, u) \transvia{α ∣ p} (s', u')\) for some \(u'\).
+%     \end{enumerate}
+%     The first goal can be shown by definition of \(δ_{≃}\). 
+%     Because \(s'\) in the second goal is live, therefore only condition~\labelcref{itm:transition-bisim} in~\Cref{def:norm-bism} matches the condition, thus implying \(u \transvia{α ∣ p} u'\) and \((s, u) \transvia{α ∣ p} (s', u')\).
+% \end{proof}
+
+% \begin{theorem}
+%     Given a normalizing bisimulation \(≃: S × U\), its normalization is indeed a bisimulation between \(\norm(S)\) and \(\norm(U)\)
+% \end{theorem}
+
+% \begin{proof}
+%     We will need to prove the two projection of \(π₁: {≃} → S\) and \(π₂: {≃} → U\) is a homomorphism, without loss of generality, we will only show that \(π₁\) is a homomorphism.
+
+%     First the accepting case, for all \(s ≃ u\):
+%     \begin{align*}
+%         δ_{\norm(≃)}((s, u), α) = \accept
+%         & ⟹ δ_{≃}((s, u), α) = \accept \\
+%         & ⟹ δ_{S}(s, α) = \accept \\ 
+%         & ⟹ δ_{\norm(S)}(s, α) = \accept \\ 
+%     \end{align*}
+
+%     Second the transition case, for \(s ≃ u\), because normalizing bisimulation preserves liveness (\Cref{thm:norm-bisim-perserve-liveness}):
+%     \begin{align*}
+%         & (s, u) \transvia{α ∣ p}_{\norm(≃)} (s', u') \\*
+%         & ⟹ (s, u) \transvia{α ∣ p}_{≃} (s', u')
+%             \text{ and \((s, u)\) is live in \(≃\)}   \\*  
+%         & ⟹ s \transvia{α ∣ p}_{S} s'
+%             \text{ and \(s\) is live in \(S\)} \\* 
+%         & ⟹ s \transvia{α ∣ p}_{\norm(S)} s'
+%     \end{align*}
+
+%     Finally, we show the rejection case. \((s, u) \transRej{\norm(≃)}{α}\) can be split into two cases, we will show both cases lead to \(s \transRej{\norm(S)}{α}\).
+%     \begin{itemize}
+%         \item The first case is that \(δ_{≃}((s, u), α) = \reject\). 
+%         By definition, this case can happen in following three scenarios: 
+%         \begin{itemize}
+%             \item when \(s \transvia{α ∣ p} s'\), \(u \transvia{α ∣ q} u'\), and \(p ≠ q\), which implies \(s'\) is dead by condition~\labelcref{itm:transition-dead};  
+%             \item when \(s \transvia{α ∣ p} s'\) and \(δ_U(u, α) = \reject\), which implies \(s'\) is dead by condition~\labelcref{itm:rej-or-dead}
+%             \item when \(u \transRej{S}{α}\), then \(u \transRej{\norm(S)}{α}\).
+%         \end{itemize}
+%         Thus, all scenarios cases lead to \(u \transRej{\norm(S)}{α}\).
+%         \item The second case is that \((s, u) \transvia{α ∣ p}_{≃} (s', u')\) but \((s', u')\) is dead.
+%         By definition \(s \transvia{α ∣ p}_{S} s'\). 
+%         By~\cref{thm:norm-bisim-perserve-liveness} and \((s', u')\) is dead, we conclude that \(s'\) is dead and therefore \(s' \transRej{\norm(S)}{α}\).
+%     \end{itemize}
+% \end{proof}
+
+% % \begin{lemma}[bisimulation between dead states]\label{thm:bisim-between-dead}
+% %     Given two dead states \(s ∈ S\) and \(u ∈ U\), then the singleton bisimulation \({∼} ⊆ \norm(S) × \norm(U)\):
+% %     \begin{mathpar}
+% %         {∼} ≜ \{(s,u)\} \and 
+% %         δ_{∼}((s,u), α) ≜ \reject
+% %     \end{mathpar}
+% %     is a bisimulation between \(\norm(S)\) and \(\norm(U)\).
+% % \end{lemma}
+
+% % \begin{proof}
+% %     By computation
+% % \end{proof}
+
+% \begin{lemma}[bisimulation between dead states]\label{thm:bisim-between-dead}
+%     Given two dead states \(s ∈ S\) and \(u ∈ U\) and the maximal bisimulation \(≡\) between \(\norm(S)\) and \(\norm(U)\), then \(s ≡ u\).
+% \end{lemma}
+
+% \begin{proof}
+%     We can construct the singleton bisimulation \({∼} ⊆ \norm(S) × \norm(U)\):
+%     \begin{mathpar}% \begin{lemma}[bisimulation between dead states]\label{thm:bisim-between-dead}
+%         %     Given two dead states \(s ∈ S\) and \(u ∈ U\), then the singleton bisimulation \({∼} ⊆ \norm(S) × \norm(U)\):
+%         %     \begin{mathpar}
+%         %         {∼} ≜ \{(s,u)\} \and 
+%         %         δ_{∼}((s,u), α) ≜ \reject
+%         %     \end{mathpar}
+%         %     is a bisimulation between \(\norm(S)\) and \(\norm(U)\).
+%         % \end{lemma}
+
+%         % \begin{proof}
+%         %     By computation
+%         % \end{proof}
+%         {∼} ≜ \{(s,u)\} \and 
+%         δ_{∼}((s,u), α) ≜ \reject,
+%     \end{mathpar}
+%     and verify that it is indeed a bisimulation between \(\norm(S)\) and \(\norm(U)\) by unfolding the definition.
+% \end{proof}
+
+% \begin{theorem}
+%     For any two GKAT coalgebra \(S\) and \(U\), the maximal bisimulation \(≡\) between \(\norm(S)\) and \(\norm(U)\) satisfy all the conditions in~\Cref{def:norm-bism}, i.e.
+the relation \(≡\) can be a normalizing bisimulation between \(S\) and \(U\).
+% \end{theorem}
+
+% \begin{proof}
+%     We only verify that \(≡\) satisfy the conditions in~\cref{def:norm-bism}.
+
+%     The condition~\labelcref{itm:transition-bisim} holds when \(s ≡ u\).
+%     We first prove that \(s'\) and \(u'\) has to be both live or both dead: because \(δ_S(s, α) = (s', p)\) and \(δ_U(u, α) = (u', p)\),
+%     \begin{align*}
+%         s' \text{ is live} 
+%         & ⟺ δ_{\norm(S)}(s, α) = (s', p) \\
+%         & ⟺ δ_{\norm(U)}(u, α) = (u', p) \\
+%         & ⟺ u' \text{ is live}.
+%     \end{align*}
+%     \begin{itemize}
+%         \item If both \(s'\) and \(u'\) are live in \(S\) and \(U\), then \(s' ≡ u'\) because \(≡\) is a bisimulation.
+%         \item If both \(s'\) and \(u'\) are dead in \(S\) and \(U\), then they are also dead in \(\norm(S)\) and \(\norm(U)\) (\cref{thm:norm-perserve-liveness}).
+%         Finally, by~\Cref{thm:bisim-between-dead}, \(s' ≡ u'\).
+%     \end{itemize}
+
+%     TODO, finish
+% \end{proof}
+
+% \begin{lemma}\label{thm:recursive-construction-lemma}
+%     Given two states in GKAT coalgebra \(s ∈ S, u ∈ U\), consider the maximal bisimulation \({≡} ⊆ \norm(S) × \norm(U)\), \(s ≡ u\) if and only if there exists a normalized bisimulation \({≃} ⊆ S × U\) s.t.
+\(s ≃ u\).
+% \end{lemma}
+
+% \begin{proof}
+%     Recall that a relation \({∼} ⊆ \norm(S) × \norm(U)\) is a bisimulation if and only if it has a normalized GKAT coalgebra structure \(δ_∼\), s.t. for all \(s₁ ∼ u₁\):
+%     \begin{itemize}
+%         \item for all results 
+%         \(r ∈ \{\accept, \reject\}\): if \(δ_∼((s₁, u₁), α) = r\), then \(δ_{\norm(S)}(s₁, α) = r\) and \(δ_{\norm(U)}(u₁, α) = r\);
+%         \item if \((s₁, u₁) \transvia{α ∣ p}_{∼} (s₂, u₂)\) then \(s₁ \transvia{α ∣ p}_{\norm(S)} s₂\) and \(u₁ \transvia{α ∣ p}_{\norm(U)} u₂\).
+%     \end{itemize}
+
+%     We prove \(⟹\) direction by constructing the following normalized bisimulation from the bisimulation \({≡}\) between \(\norm(S)\) and \(\norm(U)\):
+%     \[{≃} ≜ {≡} ∪ \{(s, u) ∣ \text{both \(s ∈ S, u ∈ U\) are dead}\}.\]
+%     We then show that \({≃}\) satisfies all the condition in~\Cref{def:norm-bism}.
+
+%     The condition \labelcref{itm:acc-condition} holds if \(s ∼ u\):
+%     \begin{align*}
+%         δ_S(s, α) = \accept 
+%         ⟺{}& δ_{\norm(S)}(s, α) = \accept \\
+%         ⟺{}& δ_∼((s,u), α) = \accept \\ 
+%         ⟺{}& δ_{\norm(U)}(u, α) = \accept \\
+%         ⟺{}& δ_U(u, α) = \accept
+%     \end{align*}
+%     The condition \labelcref{itm:acc-condition} also holds if \(s, u\) are both dead, because both can never accept any atoms.
+
+%     The condition~\labelcref{itm:rej-or-dead} holds when \(s ≡ u\): 
+%     \begin{align*}
+%         & δ_S(s, α) = \reject \text{ and } δ_U(u, α) = (u', p)\\
+%         ⟹{}& δ_{\norm(S)}(s, α) = \reject \text{ and } δ_U(u, α) = (u', p) \\
+%         ⟹{}& δ_{\norm(U)}(u, α) = \reject \text{ and } δ_U(u, α) = (u', p)\\
+%         ⟹{}& \text{\(u'\) is dead}
+%     \end{align*}
+%     Similarly, when \(u'\) reject \(α\) and \(s\) transitions to \(s'\).
+%     The condition~\labelcref{itm:rej-or-dead} holds when \(s, u\) are both dead because both of them can only transition to dead state by~\Cref{thm:dead-iff-all-reachable-dead}.
+
+%     The condition~\labelcref{itm:transition-bisim} holds when \(s ≡ u\).
+%     We first prove that \(s'\) and \(u'\) has to be both live or both dead: because \(δ_S(s, α) = (s', p)\) and \(δ_U(u, α) = (u', p)\),
+%     \begin{align*}
+%         s' \text{ is live} 
+%         & ⟺ δ_{\norm(S)}(s, α) = (s', p) \\
+%         & ⟺ δ_{\norm(U)}(u, α) = (u', p) \\
+%         & ⟺ u' \text{ is live}.
+%     \end{align*}
+%     \begin{itemize}
+%         \item If both \(s'\) and \(u'\) are live, then \(s' ≡ u'\), and \(s' ≃ u'\).
+%         \item If both \(s'\) and \(u'\) are dead, \(s' ≃ u'\).
+%     \end{itemize}
+
+%     The condition~\labelcref{itm:transition-dead} holds, let \(δ_S(s, α) = (s', p)\) and \(δ_U(u, α) = (u', q)\) and \(p ≠ q\). 
+%     If \(s' ≃ u'\), then we can show condition~\labelcref{itm:transition-dead} by case analysis on the liveness of \(s'\) and \(u'\).
+%     If \(s', u'\) are not both dead, then the condition of \({≡} ⊆ \norm(S) × \norm(U)\) cannot be satisfied, hence \((s', u') ∉ {≡}\), which means that \(s'\) and \(u'\) are both dead, contradicting the premise.
+%     Therefore, \(s'\) and \(u'\) needs to be both dead.
+
+%     Then we show the \(⟸\) direction by
+
+%     Then we show the \(⟸\) direction by showing \(≃\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\).
+%     Take any pair of states \(s₁ ≃ u₁\), we show that all the conditions of a bisimulation is satisfied.
+
+%     First, we show for all \(α ∈ \At, δ_{\norm(S)}(s₁, α) = \accept ⟺ δ_{\norm(U)}(u₁, α) = \accept\):
+%     \begin{align*}
+%         δ_{\norm(S)}(s₁, α) = \accept 
+%         & ⟺ δ_S(s₁, α) = \accept \\  
+%         & ⟺ δ_U(u₁, α) = \accept \\  
+%         & ⟺ δ_{\norm(U)})(u₁, α) = \accept
+%     \end{align*}
+
+%     Then, we show for all \(α ∈ \At, δ_{\norm(S)}(s₁, α) = \reject ⟺ δ_{\norm(U)}(u₁, α) = \reject\).
+%     Without loss of generality, we only show the \(⟹\) direction.
+%     If \(δ_{\norm(S)}(s₁, α) = \reject\), then either \(δ_S(s₁, α) = \reject\), or \(δ_S(s₁, α) = (s₂, p)\) and \(s₂\) is dead.
+%     \begin{itemize}
+%         \item We consider the case where \(δ_S(s₁, α) = \reject\).
+%         By condition~\labelcref{itm:acc-condition}, \(δ_U(u₁, α)\) cannot equal to accept. 
+%         If \(δ_U(u₁, α) = \reject\), then the implication holds. 
+%         If \(δ_U(u₁, α) = (u₂, q)\), then by condition~\cref{itm:rej-or-dead}, \(u₂\) is dead, which also makes the implication hold.
+%         \item We consider the case where \(δ_S(s₁, α) = (s₂, p)\) and \(s₂\) is dead in \(S\).
+%         By condition~\labelcref{itm:acc-condition}, \(δ_U(u₁, α)\) cannot equal to accept.
+%         If \(δ_U(u₁, α) = \reject\), then the implication holds.
+%         If \(δ_U(u₁, α) = (u₂, q)\) and \(p ≠ q\), then by condition~\labelcref{itm:transition-dead}, both \(u₁, u₂\) are dead, then by~\cref{thm:bisim-between-dead}, \(s₂ ≡ s\). 
+%         If \(δ_U(u₁, α) = (u₂, q)\) and \(p = q\), then by condition~\labelcref{itm:transition-bisim}, then \(s₂ ≃ u₂\)
+%     \end{itemize}
+
+%     To obtain the desired bisimulation \({∼}\) between \(\norm(S)\) and \(\norm(U)\), we add the pair \((s, t)\) to \(∼'\), 
+%     \[{∼} ≜ {∼'} ∪ \{(s, u)\},\] 
+%     with the following transition \(δ_{∼}\): for all \(α ∈ \At\),
+%     \begin{itemize}[nosep]
+%         \item if \(δ_{\norm(S)}(s, α) = δ_{\norm(U)}(u, α) = \accept\), then \(δ_{∼}((s, u), α) ≜ \accept\);
+%         \item if \(δ_{\norm(S)}(s, α) = δ_{\norm(U)}(u, α) = \reject\), then \(δ_{∼}((s, u), α) ≜ \reject\);
+%         \item if \(δ_{\norm(S)}(s, α) = (s', p)\) and \(δ_{\norm(U)}(u, α) = (u', p)\), then \(δ_{∼}((s, u), α) = ((s', u'), p)\);
+%         \item for all \((s', u') ∈ {∼}'\) that is not equal to \((s, u)\), we let \(δ_∼\) inherits the transition of \(δ_{∼'}\), i.e. \(δ_{∼}((s', u'), α) = δ_{∼'}((s', u'), α)\)
+%     \end{itemize}
+%     \emph{if \(δ_{∼}\) is well-defined}, then we can verify that \(∼\) is indeed a bisimulation between \(\norm(⟨s⟩)\) and \(\norm(⟨u⟩)\) where \(s ∼ u\). 
+%     We show the slightly more complicated case: \(δ_{\norm(S)}(s, α) = (s', p)\) and \(δ_{\norm(U)}(u, α) = (u', p)\) implies \(s' ∼ u'\) as an example. 
+%     By condition~\labelcref{itm:transition-bisim}, there exists a bisimulation \(∼_{s', u'} ⊆ \norm(⟨s'⟩) × \norm(⟨u'⟩)\), and because \(≡_{s', u'}\) is the maximal bisimulation between \(\norm(⟨s'⟩)\) and \(\norm(⟨u'⟩)\),
+%     \[(s', u') ∈ {∼_{s', u'}} ⊆ {≡_{s', u'}} ⊆ {∼}.\]
+
+%     Finally, we demonstrate that \(δ_∼\) is well-defined by leveraging the conditions in~\Cref{thm:recursive-construction}. 
+%     Specifically, we will show that the definition of \(δ_∼\) covers all the possible cases, by case analysis on the result of \(δ_S\): for all \(α ∈ \At\),
+%     \begin{itemize}
+%         \item If \(δ_S(s, α) = \accept\), then by condition~\labelcref{itm:acc-condition}, \(δ_U(u, α) = \accept\); therefore \[δ_{\norm(S)}(s, α) = δ_{\norm(U)}(u, α) = \accept.\]
+%         \item If \(δ_S(s, α)\) transitions to a dead state or reject, then by condition~\labelcref{itm:rej-or-dead} \(δ_U(u, α)\) will also transition to a dead state or reject, then \[δ_{\norm(S)}(s, α) = δ_{\norm(U)}(u, α) = \reject.\]
+%         \item If \(δ_S(s, α) = (s', p)\) and \(s'\) is live, then \(δ_U(u, α) = (u', p)\) necessarily holds, otherwise it would violate one of conditions~\labelcref{itm:acc-condition,itm:rej-or-dead,itm:transition-dead}.
+
+%         By condition~\labelcref{itm:transition-bisim}, there exists a bisimulation \(∼_{s', u'}\) between \(\norm(⟨s'⟩)\) and \(\norm(⟨u'⟩)\) s.t.
 \(s' ∼_{s', u'} u'\).
 Because bisimulation preserves liveness (\Cref{thm:bisim-preserve-liveness}), \(s', u'\) has to be both dead or live.
-        Finally, because \(s'\) is live, therefore \(u'\) is also live, and we obtain the final case in the definition of \(δ_{∼}\): \(δ_{\norm(S)}(s, α) = (s', p)\) and \(δ_{\norm(U)}(u, α) = (u', p)\).
-        \qedhere
-    \end{itemize}
-\end{proof}
-
-With some slight tweak to \Cref{thm:recursive-construction-lemma}, it will be applicable to trace equivalence instead of bisimulation.
-The correctness of this modification is supported by the correspondence between trace equivalence and the existence of a bisimulation between normalized GKAT coalgebras, which is stated in~\Cref{thm:norm-bisim-correctness}.
-
-\begin{theorem}[Recursive Construction]\label{thm:recursive-construction}
-    For any two states in two GKAT coalgebra \(s ∈ S, u ∈ U\), \(s\) and \(u\) are finite-trace equivalent \(⟦s⟧_{S} = ⟦u⟧_{U}\) if and only if all the following conditions hold:
-    \begin{enumerate}
-        \item for all \(α ∈ \At\), \(δ_{S}(s, α) = \accept ⟺ δ_{U}(u, α) = \accept\);
-        \item \(s\) reject \(α\) or transition to a dead state via \(α\) if and only if \(u\) rejects \(α\) or transition to a dead state via \(α\);  
-        \item if \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', p)\), then \(⟦s'⟧_{S} = ⟦u'⟧_{U}\);
-        \item if \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', q)\), s.t. \(p ≠ q\), then both \(s'\) and \(t'\) are dead.
-    \end{enumerate}
-    Notice that above condition is similar to those in~\cref{thm:recursive-construction}, except bisimulation of \(s'\) and \(u'\) is replaced with trace equivalence.
-\end{theorem}
-
-\begin{proof}
-    By the standard argument with normalization preserves sub-coalgebra (\Cref{thm:norm-sub-coalg}), we can obtain for all \(s ∈ S\) and \(u ∈ U\), \(\norm(⟨s⟩) ⊑ \norm(S)\) and \(\norm(⟨u⟩) ⊑ \norm(U)\).
-    Therefore, because subcoalgebra preserve and reflect bisimulation (\Cref{thm:sub-coalg-preserve-bisim}) and the correctness of bisimulation on normalized coalgebra (\Cref{thm:norm-bisim-correctness}): for all \(s ∈ S\) and \(u ∈ U\),
-    \begin{align*}
-        & ∃~ {∼} ⊆ \norm(⟨s⟩) × \norm(⟨u⟩), s ∼ u \\
-        ⟺{}& ∃~ {∼} ⊆ \norm(S) × \norm(U), s ∼ u \\  
-        ⟺{}& ⟦s⟧_S = ⟦u⟧_U.
-    \end{align*}
-    We can instantiate the above equivalence to \(s', u'\) and \(s, u\) respectively: the instantiation to \(s', u'\) will give us the conditions in this theorem is equivalent to the ones in~\Cref{thm:recursive-construction-lemma}; and instantiation to \(s, u\) show that existence of bisimulation is equivalent to trace equivalence:
-    \begin{align*}
-        & \text{Conditions in \Cref{thm:recursive-construction-lemma} hold} \\  
-        ⟺{}& \text{Conditions in the current theorem hold}\\
-        ⟺{}& ∃~ {∼} ⊆ \norm(⟨s⟩) × \norm(⟨u⟩), s ∼ u\\
-        ⟺{}& ⟦s⟧_{S} = ⟦u⟧_{U}. \qedhere
-    \end{align*}
-\end{proof}
-
-The above construction theorem already gives us an algorithm to recursively decide whether \(⟦s⟧_{S} = ⟦u⟧_U\), when \(⟨s⟩_S\) and \(⟨u⟩_U\) is finite.
-However, this algorithm can be further optimized: we will derive that a dead state is only trace equivalent to other dead states. 
-This means that when checking the trace-equivalence of states \(s\) and \(u\), if we already know one of them is dead, we only need to check whether the other is dead, instead of going through all the conditions in~\Cref{thm:recursive-construction}.
-
-\begin{theorem}\label{thm:bisim-one-dead}
-    Given two states \(s ∈ S\) and \(u ∈ U\), if \(s\) is a dead state in \(S\), then \(s\) and \(u\) is trace equivalent if and only if \(u\) is also dead.
-\end{theorem}
-
-\begin{proof}
-    TODO: I think we can refine \(⟹\) direction, but I am not sure...
-
-    For the \(⟸\) direction, we can construct the bisimulation as in~\cref{thm:bisim-between-dead}, which implies trace equivalence.
-    And the \(⟹\) direction, if \(s\) is dead, then by definition of normalization, it will be all rejecting, i.e. for all \(α ∈ \At\), \(δ_{\norm(S)}(s, α) = \reject\).
-    If \(s\) and \(u\) are trace equivalent, then there exists a bisimulation \({∼}: \norm(S) × \norm(U)\) s.t. \(s ∼ u\).
-    By unfolding the definition of a bisimulation, \(u\) also need to be all rejecting in \(\norm(U)\).
-    By definition of \(\norm\), for all \(α ∈ \At\), \(δ_U(u, α)\) either reject to go to a dead state, and because \(⟨u⟩_U\) is all the reachable state from \(u\), therefore 
-    \begin{align*}
-        ⟨u⟩_U = & ⋃ \{⟨u'⟩ ∣ ∃ α ∈ \At, p ∈ K, δ_U(u, α) = (u', p)\} \\
-        & ∪ \{u\}
-    \end{align*}
-    And because all of \(u'\) is dead, all of \(⟨u'⟩_U\) cannot be accepting (\Cref{thm:dead-iff-all-reachable-dead}), nor is \(u\) an accepting state, then \(⟨u⟩_U\) does not contain any accepting state, and by definition of dead state, \(u\) is dead.
-\end{proof}
+%         Finally, because \(s'\) is live, therefore \(u'\) is also live, and we obtain the final case in the definition of \(δ_{∼}\): \(δ_{\norm(S)}(s, α) = (s', p)\) and \(δ_{\norm(U)}(u, α) = (u', p)\).
+%     \end{itemize}
+%     Because \(≃ ⊆ \norm(S) × \norm(U)\) is a bisimulation s.t. \(s ≃ u\), and \(≡\) is the maximal bisimulation between \(\norm(S)\) and \(\norm(U)\), we will obtain our result \(s ≡ u\).
+% \end{proof}
+
+% % \begin{proof}
+% %     Recall that there exists a bisimulation \({∼} ⊆ \norm(⟨s⟩) × \norm(⟨u⟩)\) if and only if for all \(s₁ ∼ u₁\):
+% %     \begin{itemize}
+% %         \item for all results \(r ∈ \{\accept, \reject\}\): \(δ_{\norm(S)}(s₁, α) = r ⟺ δ_{\norm(U)}(u₂, α) = r\);
+% %         \item otherwise, let \((s₂, p) ≜ δ_{\norm(S)}(s₁, α)\) and \((u₂, q) ≜ δ_{\norm(u)}(u₁, α)\), then \(p = q\) and \(s₂ ∼ u₂\)
+% %     \end{itemize}
+
+% %     We prove \(⟹\) direction by constructing the following normalized bisimulation from the bisimulation \({∼}\):
+% %     \[{≃} ≜ {∼} ∪ \{(s, u) ∣ \text{both \(s ∈ S, u ∈ U\) are dead}\}.\]
+% %     We then show that \({≃}\) satisfies all the condition in~\Cref{def:norm-bism}.
+
+% %     The condition \labelcref{itm:acc-condition} holds if \(s ∼ u\):
+% %     \begin{align*}
+% %         δ_S(s, α) = \accept 
+% %         ⟺{}& δ_{\norm(S)}(s, α) = \accept \\
+% %         ⟺{}& δ_∼((s,u), α) = \accept \\ 
+% %         ⟺{}& δ_{\norm(U)}(u, α) = \accept \\
+% %         ⟺{}& δ_U(u, α) = \accept
+% %     \end{align*}
+% %     The condition \labelcref{itm:acc-condition} also holds if \(s, u\) are both dead, because both can never accept any atoms.
+
+% %     The condition~\labelcref{itm:rej-or-dead} holds when \(s ∼ u\): 
+% %     \begin{align*}
+% %         & δ_S(s, α) = \reject \text{ and } δ_U(u, α) = (u', p)\\
+% %         ⟹{}& δ_{\norm(S)}(s, α) = \reject \text{ and } δ_U(u, α) = (u', p) \\
+% %         ⟹{}& δ_{\norm(U)}(u, α) = \reject \text{ and } δ_U(u, α) = (u', p)\\
+% %         ⟹{}& \text{\(u'\) is dead}
+% %     \end{align*}
+% %     Similarly, when \(u'\) reject \(α\) and \(s\) transitions to \(s'\).
+% %     The condition~\labelcref{itm:rej-or-dead} holds when \(s, u\) are both dead because both of them can only transition to dead state by~\Cref{thm:dead-iff-all-reachable-dead}.
+
+% %     The condition~\labelcref{itm:transition-bisim} holds when \(s ∼ u\).
+% %     We first prove that \(s'\) and \(u'\) has to be both live or both dead: because \(δ_S(s, α) = (s', p)\) and \(δ_U(u, α) = (u', p)\),
+% %     \begin{align*}
+% %         s' \text{ is live} 
+% %         & ⟺ δ_{\norm(S)}(s, α) = (s', p) \\
+% %         & ⟺ δ_{\norm(U)}(u, α) = (u', p) \\
+% %         & ⟺ u' \text{ is live}.
+% %     \end{align*}
+% %     \begin{itemize}
+% %         \item If both \(s'\) and \(u'\) are live, then \(s' ∼ u'\). By~\cref{thm:sub-coalg-preserve-bisim}, the bisimulation \(∼'\) is just \(∼\) restricted to \(⟨s'⟩\) and \(⟨u'⟩\).
+% %         \item If both \(s'\) and \(u'\) are dead, then \(∼'\) can just be the singleton relation, according to~\cref{thm:bisim-between-dead}.
+% %     \end{itemize}
+
+% %     The condition~\labelcref{itm:transition-dead} holds: by the proof of condition~\labelcref{itm:transition-bisim}, \(s'\) and \(u'\) has to be either both live or both dead; if they are both live, then there cannot be a element in \(G(∼)\) that can project to \((s', p)\) under \(π₁\) but projects to \((t', q)\) under \(π₂\).
+Thus both \(s'\) and \(t'\) has to be dead.
+
+% %     We then show the \(⟸\) direction, for arbitrary \(s' ∈ S\) and \(u' ∈ U\), we use \(≡_{s', u'}\) to denote the maximal bisimulation between \(\norm(⟨s'⟩)\) and \(\norm(⟨u'⟩)\).
+% %     \begin{align*}
+% %         {∼'} ≜ ⋃ \{≡_{s', u'} & ∣ ∃ α ∈ \At, p ∈ K, \\*
+% %             & δ_{\norm(S)}(s, α) = (s', p) \\*
+% %             & \text{ and } δ_{\norm(U)}(u, α) = (u', p)\}.
+% %     \end{align*}
+% %     For all the \(s'\) and \(u'\) in the above definition, \(⟨s'⟩ ⊑ ⟨s⟩\) and \(⟨u'⟩ ⊑ ⟨u⟩\), therefore by~\Cref{thm:norm-sub-coalg}, \(\norm(⟨s'⟩) ⊑ \norm(⟨s⟩)\) and \(\norm(⟨u'⟩) ⊑ \norm(⟨u⟩)\).
+% %     By~\Cref{thm:sub-coalg-preserve-bisim}, every \(≡_{s', u'}\) is a bisimulation between \(\norm(⟨s⟩)\) and \(\norm(⟨t⟩)\), and because bisimulation is closed under arbitrary union~\cite{rutten_UniversalCoalgebraTheory_2000}, \(∼'\) is a bisimulation between \(\norm(⟨s⟩)\) and \(\norm(⟨t⟩)\).
+
+% %     To obtain the desired bisimulation \({∼}\) between \(\norm(⟨s⟩)\) and \(\norm(⟨u⟩)\), we add the pair \((s, t)\) to \(∼'\), 
+% %     \[{∼} ≜ {∼'} ∪ \{(s, u)\},\] 
+% %     with the following transition \(δ_{∼}\): for all \(α ∈ \At\),
+% %     \begin{itemize}[nosep]
+% %         \item if \(δ_{\norm(S)}(s, α) = δ_{\norm(U)}(u, α) = \accept\), then \(δ_{∼}((s, u), α) ≜ \accept\);
+% %         \item if \(δ_{\norm(S)}(s, α) = δ_{\norm(U)}(u, α) = \reject\), then \(δ_{∼}((s, u), α) ≜ \reject\);
+% %         \item if \(δ_{\norm(S)}(s, α) = (s', p)\) and \(δ_{\norm(U)}(u, α) = (u', p)\), then \(δ_{∼}((s, u), α) = ((s', u'), p)\);
+% %         \item for all \((s', u') ∈ {∼}'\) that is not equal to \((s, u)\), we let \(δ_∼\) inherits the transition of \(δ_{∼'}\), i.e. \(δ_{∼}((s', u'), α) = δ_{∼'}((s', u'), α)\)
+% %     \end{itemize}
+% %     \emph{if \(δ_{∼}\) is well-defined}, then we can verify that \(∼\) is indeed a bisimulation between \(\norm(⟨s⟩)\) and \(\norm(⟨u⟩)\) where \(s ∼ u\). 
+% %     We show the slightly more complicated case: \(δ_{\norm(S)}(s, α) = (s', p)\) and \(δ_{\norm(U)}(u, α) = (u', p)\) implies \(s' ∼ u'\) as an example. 
+% %     By condition~\labelcref{itm:transition-bisim}, there exists a bisimulation \(∼_{s', u'} ⊆ \norm(⟨s'⟩) × \norm(⟨u'⟩)\), and because \(≡_{s', u'}\) is the maximal bisimulation between \(\norm(⟨s'⟩)\) and \(\norm(⟨u'⟩)\),
+% %     \[(s', u') ∈ {∼_{s', u'}} ⊆ {≡_{s', u'}} ⊆ {∼}.\]
+
+% %     Finally, we demonstrate that \(δ_∼\) is well-defined by leveraging the conditions in~\Cref{thm:recursive-construction}. 
+% %     Specifically, we will show that the definition of \(δ_∼\) covers all the possible cases, by case analysis on the result of \(δ_S\): for all \(α ∈ \At\),
+% %     \begin{itemize}
+% %         \item If \(δ_S(s, α) = \accept\), then by condition~\labelcref{itm:acc-condition}, \(δ_U(u, α) = \accept\); therefore \[δ_{\norm(S)}(s, α) = δ_{\norm(U)}(u, α) = \accept.\]
+% %         \item If \(δ_S(s, α)\) transitions to a dead state or reject, then by condition~\labelcref{itm:rej-or-dead} \(δ_U(u, α)\) will also transition to a dead state or reject, then \[δ_{\norm(S)}(s, α) = δ_{\norm(U)}(u, α) = \reject.\]
+% %         \item If \(δ_S(s, α) = (s', p)\) and \(s'\) is live, then \(δ_U(u, α) = (u', p)\) necessarily holds, otherwise it would violate one of conditions~\labelcref{itm:acc-condition,itm:rej-or-dead,itm:transition-dead}.
+
+% %         By condition~\labelcref{itm:transition-bisim}, there exists a bisimulation \(∼_{s', u'}\) between \(\norm(⟨s'⟩)\) and \(\norm(⟨u'⟩)\) s.t.
+\(s' ∼_{s', u'} u'\).
+Because bisimulation preserves liveness (\Cref{thm:bisim-preserve-liveness}), \(s', u'\) has to be both dead or live.
+% %         Finally, because \(s'\) is live, therefore \(u'\) is also live, and we obtain the final case in the definition of \(δ_{∼}\): \(δ_{\norm(S)}(s, α) = (s', p)\) and \(δ_{\norm(U)}(u, α) = (u', p)\).
+% %         \qedhere
+% %     \end{itemize}
+% % \end{proof}
+
+% With some slight tweak to \Cref{thm:recursive-construction-lemma}, it will be applicable to trace equivalence instead of bisimulation.
+% The correctness of this modification is supported by the correspondence between trace equivalence and the existence of a bisimulation between normalized GKAT coalgebras, which is stated in~\Cref{thm:norm-bisim-correctness}.
+
+% \begin{theorem}[Recursive Construction]\label{thm:recursive-construction}
+%     For any two states in two GKAT coalgebra \(s ∈ S, u ∈ U\), \(s\) and \(u\) are finite-trace equivalent \(⟦s⟧_{S} = ⟦u⟧_{U}\) if and only if all the following conditions hold:
+%     \begin{enumerate}
+%         \item for all \(α ∈ \At\), \(δ_{S}(s, α) = \accept ⟺ δ_{U}(u, α) = \accept\);
+%         \item \(s\) reject \(α\) or transition to a dead state via \(α\) if and only if \(u\) rejects \(α\) or transition to a dead state via \(α\);  
+%         \item if \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', p)\), then \(⟦s'⟧_{S} = ⟦u'⟧_{U}\);
+%         \item if \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', q)\), s.t. \(p ≠ q\), then both \(s'\) and \(u'\) are dead.
+%     \end{enumerate}
+%     Notice that above condition is similar to those in~\cref{thm:recursive-construction}, except bisimulation of \(s'\) and \(u'\) is replaced with trace equivalence.
+% \end{theorem}
+
+% \begin{proof}
+%     By the standard argument with normalization preserves sub-coalgebra (\Cref{thm:norm-sub-coalg}), we can obtain for all \(s ∈ S\) and \(u ∈ U\), \(\norm(⟨s⟩) ⊑ \norm(S)\) and \(\norm(⟨u⟩) ⊑ \norm(U)\).
+%     Therefore, because subcoalgebra preserve and reflect bisimulation (\Cref{thm:sub-coalg-preserve-bisim}) and the correctness of bisimulation on normalized coalgebra (\Cref{thm:norm-bisim-correctness}): for all \(s ∈ S\) and \(u ∈ U\),
+%     \begin{align*}
+%         & ∃~ {∼} ⊆ \norm(⟨s⟩) × \norm(⟨u⟩), s ∼ u \\
+%         ⟺{}& ∃~ {∼} ⊆ \norm(S) × \norm(U), s ∼ u \\  
+%         ⟺{}& ⟦s⟧_S = ⟦u⟧_U.
+%     \end{align*}
+%     We can instantiate the above equivalence to \(s', u'\) and \(s, u\) respectively: the instantiation to \(s', u'\) will give us the conditions in this theorem is equivalent to the ones in~\Cref{thm:recursive-construction-lemma}; and instantiation to \(s, u\) show that existence of bisimulation is equivalent to trace equivalence:
+%     \begin{align*}
+%         & \text{Conditions in \Cref{thm:recursive-construction-lemma} hold} \\  
+%         ⟺{}& \text{Conditions in the current theorem hold}\\
+%         ⟺{}& ∃~ {∼} ⊆ \norm(⟨s⟩) × \norm(⟨u⟩), s ∼ u\\
+%         ⟺{}& ⟦s⟧_{S} = ⟦u⟧_{U}. \qedhere
+%     \end{align*}
+% \end{proof}
+
+% The above construction theorem already gives us an algorithm to recursively decide whether \(⟦s⟧_{S} = ⟦u⟧_U\), when \(⟨s⟩_S\) and \(⟨u⟩_U\) is finite.
+% However, this algorithm can be further optimized: we will derive that a dead state is only trace equivalent to other dead states. 
+% This means that when checking the trace-equivalence of states \(s\) and \(u\), if we already know one of them is dead, we only need to check whether the other is dead, instead of going through all the conditions in~\Cref{thm:recursive-construction}.
+
+% \begin{theorem}\label{thm:bisim-one-dead}
+%     Given two states \(s ∈ S\) and \(u ∈ U\), if \(s\) is a dead state in \(S\), then \(s\) and \(u\) is trace equivalent if and only if \(u\) is also dead.
+% \end{theorem}
+
+% \begin{proof}
+%     TODO: I think we can refine \(⟹\) direction, but I am not sure...
+
+%     For the \(⟸\) direction, we can construct the bisimulation as in~\cref{thm:bisim-between-dead}, which implies trace equivalence.
+%     And the \(⟹\) direction, if \(s\) is dead, then by definition of normalization, it will be all rejecting, i.e. for all \(α ∈ \At\), \(δ_{\norm(S)}(s, α) = \reject\).
+%     If \(s\) and \(u\) are trace equivalent, then there exists a bisimulation \({∼}: \norm(S) × \norm(U)\) s.t. \(s ∼ u\).
+%     By unfolding the definition of a bisimulation, \(u\) also need to be all rejecting in \(\norm(U)\).
+%     By definition of \(\norm\), for all \(α ∈ \At\), \(δ_U(u, α)\) either reject to go to a dead state, and because \(⟨u⟩_U\) is all the reachable state from \(u\), therefore 
+%     \begin{align*}
+%         ⟨u⟩_U = & ⋃ \{⟨u'⟩ ∣ ∃ α ∈ \At, p ∈ K, δ_U(u, α) = (u', p)\} \\
+%         & ∪ \{u\}
+%     \end{align*}
+%     And because all of \(u'\) is dead, all of \(⟨u'⟩_U\) cannot be accepting (\Cref{thm:dead-iff-all-reachable-dead}), nor is \(u\) an accepting state, then \(⟨u⟩_U\) does not contain any accepting state, and by definition of dead state, \(u\) is dead.
+% \end{proof}
 

### 2025-01-16 main.tex
epsilon collapsing optimization

@@
 \begin{align*}
-    ⟨B|ϵ̂(s) & ≜ \{b ∧ c ∣ b ∈ B, c ∈ ϵ(s)\}; \\
-    ⟨B|δ̂(s) & ≜ \{(b ∧ c, s', p) ∣ b ∈ B, (c, s', p) ∈ δ(s)\}.
+    ⟨B|ϵ̂(s) & ≜ \{b ∧ c ∣ b ∈ B, c ∈ \hat{\epsilon}(s)\}; \\
+    ⟨B|δ̂(s) & ≜ \{(b ∧ c, s', p) ∣ b ∈ B, (c, s', p) ∈ \hat{\delta}(s)\}.
 \end{align*}
@@
 
+The final optimization that we perform is collapsing the sets of boolean expressions of \(\hat{\epsilon}(s)\) into a single boolean expression.
+Basically, each \(\hat{\epsilon}(s)\) is represented by the boolean disjunction of its elements.
+Consider the guard operation defined in \Cref{sec:symb-gkat-construction}.
+\begin{align*}
+    ⟨\hat{\epsilon_1}(s_1)|\hat{\epsilon_2}(s_2) & ≜ \{b ∧ c ∣ b ∈ \hat{\epsilon_1}(s_1), c ∈ \hat{\epsilon_2}(s_2)\}
+\end{align*}
+If \(\hat{\epsilon_1}(s_1)\) and \(\hat{\epsilon_2}(s_2)\) are sets of boolean expressions, then the guard operation would produce a set whose size is proportional to \(|\hat{\epsilon_1}(s_1)| \times |\hat{\epsilon_2}(s_2)|\).
+However, if the collapsed representation is used, \(\hat{\epsilon_1}(s_1)\) and \(\hat{\epsilon_2}(s_2)\) become just boolean expressions instead of sets of expressions.
+The guard operation now produces a single boolean expression.
+The distributivity of boolean conjunction and disjunction ensures the correctness of this optimization.
+
 \subsection{Performance}\label{sec:performance-implementation}

### 2025-01-16 main.tex
dead state detection implementation

@@
 
-\newpage
+\clearpage
 \appendix
 
+\begin{onecolumn}
+\begin{lstlisting}[
+    caption=Dead State Detection Implementation,
+    label=lst:dead-state-detection
+]
+pub struct Solver {
+    // search states
+    pub(super) dead_states: HashSet<Exp>,
+    pub(super) explored: HashSet<Exp>,
+    pub(super) uf_table: HashMap<Exp, UnionFindNode<()>>,
+    // caching
+    pub(super) eps_cache: LruCache<Exp, BExp>,
+    pub(super) deriv_cache: LruCache<Exp, Vec<(BExp, Exp, u64)>>,
+}
+
+impl Solver {
+    pub fn is_dead(&mut self, gkat: &mut Gkat, exp: &Exp) -> bool {
+        let mut stack = Vec::new();
+        stack.push(exp.clone());
+        self.explored.clear();
+        while let Some(exp) = stack.pop() {
+            if self.dead_states.contains(&exp) || self.explored.contains(&exp) {
+                continue;
+            }
+            self.explored.insert(exp.clone());
+            let eps = self.epsilon(gkat, &exp);
+            if eps.is_false() {
+                for (_, e, _) in self.derivative(gkat, &exp) {
+                    stack.push(e);
+                }
+            } else {
+                return false;
+            }
+        }
+        self.dead_states.extend(self.explored.iter().cloned());
+        return true;
+    }
+}
+\end{lstlisting}
+\end{onecolumn}
+
 \end{document}

### 2025-01-16 main.tex
qcfu email
finish proof

--- 2025-01-15 main.tex
@@
 
-\begin{definition}[normalizing bisimulation]\label{def:norm-bisim}
-    Given two GKAT coalgebra \(S\) and \(U\), and a bisimulation \({∼} ⊆ \norm(S) × \norm(U)\), a normalizing bisimulation \({≈} ⊆ S × U\) is a GKAT coalgebra that contains \(∼\) with the following transition function:
+\begin{definition}[greedy bisimulation]\label{def:norm-bisim}
+    Given two GKAT coalgebra \(S\) and \(U\), and a bisimulation \({∼} ⊆ \norm(S) × \norm(U)\), a greedy bisimulation \({≈} ⊆ S × U\) is a GKAT coalgebra that contains \(∼\) with the following transition function:
     \[
@@
 
-\begin{lemma}
-    Given a bisimulation \({∼} ⊆ \norm(S) × \norm(U)\), the following relation \({∼_d} ⊆ S × U\):
+\begin{lemma}\label{thm:dead-construction-greedy-bisim}
+    Given any eager bisimulation \({≈} ⊆ S × U\) with the underlying bisimulation \({∼}\), the following relation \({∼_d} ⊆ S × U\):
     \[{∼_d} ≜ {∼} ∪ \{(s, u) ∣ \text{\(s\) and \(u\) are dead in \(S\) and \(U\)}\},\]
-    form a GKAT coalgebra the transition function \(δ_≈\) in~\Cref{def:norm-bisim}
+    form a GKAT coalgebra the transition function \(δ_≈\) in~\Cref{def:norm-bisim}, which implies \({≈} ⊑ \)
+    Furthermore, \(\norm(∼_d)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\).
 \end{lemma}
@@
     \end{itemize}
+
+    By~\Cref{thm:bisim-between-dead}, we can construct a bisimulation \(∼_{s,u} ≜ \{(s, u)\} ⊆ \norm(S) × \norm(U)\) between any two dead states \(s ∈ S\) and \(u ∈ U\).
+    Then because bisimulation is closed under union~\cite[Theorem 5.5]{rutten_UniversalCoalgebraTheory_2000}, \(\norm(∼_d) = {∼} ∪ ⋃ \{∼_{s, u} ∣ \text{\(s, u\) are dead}\}\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\).
+    TODO: bisimulation are uniquely determined by its carrier.
 \end{proof}
@@
 \begin{lemma}\label{thm:norm-bisim-exists-unique}
-    For any \(∼\) a normalizing bisimulation with \(∼\) as underlying bisimulation exists and is unique.
+    For any bisimulation between two normalized GKAT coalgebra \({∼} ⊆ \norm(S) × \norm(U)\), a greedy bisimulation \(≈\) with \(∼\) as underlying bisimulation exists and is unique.
 \end{lemma}
@@
 \begin{proof}
-    TODO: sub-coalgebra form a complete lattice.
+    Because sub-coalgebra form a complete lattice under union and intersection~\cite[Theorem 6.4]{rutten_UniversalCoalgebraTheory_2000}, then we can construct \(≈\) by taking the intersection of all the sub-coalgebra of \(∼_d\) in~\Cref{thm:dead-construction-greedy-bisim}.
 \end{proof}
@@
 
-\begin{lemma}\label{thm:live-in-component-then-live-in-norm-bisim}
-    Given a normalizing bisimulation \(≈\), if a state \((s, u)\) is live in \(≈\), then \(s\) and \(u\) are live in \(S\) and \(U\) respectively, which also implies that \((s, u)\) is live in \(∼\).
+\begin{theorem}[Perservation Of Liveness]\label{thm:greedy-bisim-perserve-liveness}
+    Given a pair of states in a greedy bisimulation \(s ≈ u\) where \({≈} ⊆ S × U\), then the following are equivalent
+    \begin{itemize}
+        \item \((s, u)\) is live (or dead) in \({≈}\);
+        \item \(s\) is live (or dead) in \(S\);  
+        \item \(u\) is live (or dead) in \(U\).
+    \end{itemize}
+    Furthermore, if \((s, u)\) is in the underlying bisimulation \(∼\) of \(≈\), then all the above is also equivalent to \((s, u)\) is live (or dead) in \(∼\).
+\end{theorem}
+
+\begin{proof}
+    Let \(∼\) by the underlying bisimulation of \(≈\), consider the \(∼_d\) in~\Cref{thm:dead-construction-greedy-bisim}, then we know that \({≈} ⊑ {∼_d}\).
+
+    Because \(\norm(∼_d)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\), then
+    \begin{align*}
+        & \text{\(s\) is live in \(S\)} \\
+        \text{ iff } & \text{\(s\) is live in \(\norm(S)\)} 
+            & \text{\Cref{thm:norm-perserve-liveness}}\\  
+        \text{ iff } & \text{\((s, u)\) is live in \(\norm(∼_d)\)} 
+            & \text{\Cref{thm:bisim-preserve-liveness}}\\
+        \text{ iff } & \text{\((s, u)\) is live in \(∼_d\)} 
+            & \text{\Cref{thm:norm-perserve-liveness}}\\
+        \text{ iff } & \text{\((s, u)\) is live in \(≈\)} 
+            & \text{\Cref{thm:sub-coalg-preserve-liveness}}
+    \end{align*}
+\end{proof}
+
+\begin{lemma}\label{thm:greedy-bisim-iff-norm-bisim}
+    Given two GKAT coalgebras \(S\) and \(U\), and a relational GKAT coalgebra \({≈} ⊆ S × U\) with the transition \(δ_≈\) as in~\cref{def:norm-bisim}.
+    \(≈\) is a greedy bisimulation if and only if \(\norm(≈)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\)
 \end{lemma}
@@
 \begin{proof}
-    TODO: by preservation of accepting state and liveness chain.
-\end{proof}
-
-\begin{lemma}
-    Every state in \((s, u) ∈ {≈} ⊆ S × U\) but not in \({∼} ⊆ \norm(S) × \norm(U)\) then \((s, u)\) is dead in \(≈\).
-\end{lemma}
-
-\begin{proof}
-    By~\Cref{thm:norm-bisim-not-bisim-then-component-dead}, \(s\) is dead in \(S\), then by contrapositive of~\Cref{thm:live-in-component-then-live-in-norm-bisim}, we obtain that \((s, u)\) is dead in \(≈\).
+    The \(⟹\) direction, we let \(∼_d\) be defined as in~\Cref{thm:dead-construction-greedy-bisim}.
+    Because \({≈} ⊑ {∼_d}\) and~\Cref{thm:norm-sub-coalg}, \(\norm(≈) ⊑ \norm(∼_d)\). 
+    Because \(\norm(∼_d)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\), then we can show that \(\norm(≈)\) is a bisimulation by factoring through \(\norm(∼_d)\), in the category of GKAT coalgebras:
+    \begin{align*}
+        & \norm(≈) 
+        \xrightarrow{i} \norm(∼_d) 
+        \xrightarrow{π₁} \norm(S) \\
+        & \norm(≈) 
+        \xrightarrow{i} \norm(∼_d) 
+        \xrightarrow{π₂} \norm(U) 
+    \end{align*}
+    because inclusion homomorphism \(i\) do not change its input, the two homomorphism obtained by bisimulation is indeed \(π₁: \norm(≈) → \norm(S)\) and \(π₂: \norm(≈) → \norm(U)\), implying both projections of \(\norm(≈)\) are homomorphism, meaning \(\norm(≈)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\).
+
+    The \(⟸\) direction can be proven by simply letting \(\norm(≈)\) be the underlying bisimulation of \(≈\), because \(\norm(≈)\) has the same carrier set of \(≈\), then \(≈\) has to be the least GKAT coalgebra with \(δ_≈\) that contains \(\norm(≈)\), satisfying all the conditions for greedy bisimulation.
 \end{proof}
@@
 \begin{theorem}[Correctness]
-    Given two states in GKAT coalgebras \(s ∈ S\) and \(u ∈ ,U\), there exists a normalizing bisimulation \({≈} ⊆ S × U\) s.t. \(s ≈ u\), if and only if \(⟦s⟧_S = ⟦u⟧_U\).
+    Given two states in GKAT coalgebras \(s ∈ S\) and \(u ∈ ,U\), there exists a greedy bisimulation \({≈} ⊆ S × U\) s.t. \(s ≈ u\), if and only if \(⟦s⟧_S = ⟦u⟧_U\).
 \end{theorem}
@@
     \(⟸\) direction, because \(⟦s⟧_S = ⟦u⟧_U\), then by~\cref{thm:norm-bisim-correctness}, then there a bisimulation \(∼ ⊆ \norm(S) × \norm(U)\) s.t. \(s ∼ u\). 
-    Because~\Cref{thm:norm-bisim-exists-unique}, therefore there exists a normalizing bisimulation \(≈\) with \(∼\) as its underlying bisimulation. 
+    Because~\Cref{thm:norm-bisim-exists-unique}, therefore there exists a greedy bisimulation \(≈\) with \(∼\) as its underlying bisimulation. 
     Because the carriers \({∼} ⊆ {≈}\), therefore \(s ∼ u\) implies \(s ≈ u\).
@@
 
-\begin{theorem}[Recursive Construction]
-    Given a relation \({≈} ⊆ S × U\), \(≈\) is a carrier of some normalizing bisimulation if and only if all the following condition are satisfied: for all pair of states \(s ≈ u\),
+\begin{theorem}[Recursive Construction]\label{thm:recursive-construction}
+    Given a relation \({≈} ⊆ S × U\), \(≈\) is a carrier of some greedy bisimulation if and only if all the following condition are satisfied: for all pair of states \(s ≈ u\),
     \begin{enumerate}
@@
 \begin{proof}
-    To show the \(⟹\) direction, assuming there exists a normalizing bisimulation \(≈\), then it would satisfy all the conditions. 
+    To show the \(⟹\) direction, assuming there exists a greedy bisimulation \(≈\), then it would satisfy all the conditions. 
     We first note that the pairs \((s, u) ∈ {≈}\) can satisfy one of two properties (or both), either \(s\) and \(u\) are both dead in \(S\) and \(U\), or \((s, u)\) is in the underlying bisimulation \(∼\).
@@
 
-    The condition~\labelcref{itm:transition-bisim} holds because when \(s \transvia{α ∣ p}_S s'\) and \(u \transvia{α ∣ p}_U u'\), then \((s, u) \transvia{α ∣ p}_{≈} (s', u')\) by definition of normalizing bisimulation (\Cref{def:norm-bisim}).
+    The condition~\labelcref{itm:transition-bisim} holds because when \(s \transvia{α ∣ p}_S s'\) and \(u \transvia{α ∣ p}_U u'\), then \((s, u) \transvia{α ∣ p}_{≈} (s', u')\) by definition of greedy bisimulation (\Cref{def:norm-bisim}).
 
@@
     Then we prove the \(⟸\) direction.
-    TODO: by constructing a \(≈\) and show that \(\norm(≈)\) is a bisimulation. Maybe using the maximal bisimulation?
+    If \(≈\) satisfy all the conditions, then by condition~\labelcref{itm:transition-bisim}, \(≈\) is closed under \(δ_≈\), hence is a GKAT coalgebra.
+    We let its underlying bisimulation be \(\norm(≈)\), which is sound if \(\norm(≈)\) is a bisimulation (\cref{thm:greedy-bisim-iff-norm-bisim}).
+    Thus, we only need to show that \(\norm(≈)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\). Without loss of generality, we will show that \(π₁: \norm(≈) → \norm(S)\) is a bisimulation.
+
+    For all pair of states \((s, u) ∈ \norm(≈)\) and any atom \(α ∈ \At\), we will verify all the conditions of a homomorphism.
+    For the accepting case, by the definition of \(δ_{≈}\):
+    \begin{align*}
+        (s, u) \transAcc{\norm(≈)}{α}
+        & \text{ iff } (s, u) \transAcc{≈}{α} \\  
+        & \text{ implies } s \transAcc{≈}{α}
+    \end{align*}
+    For the transition case, because \((s, u)\) is live in \(≈\) implies that \(s\) is live in \(S\) (\Cref{thm:greedy-bisim-perserve-liveness}):
+    \begin{align*}
+        & (s, u) \transvia{α ∣ p}_{\norm(S)} (s', u') \\
+        & \text{iff }
+            (s, u) \transvia{α ∣ p}_{S} (s', u')
+            \text{ and \(s, u\) is live in \(≈\)} \\  
+        & \text{implies }
+            s \transvia{α ∣ p}_S s' 
+            \text{ and \(s\) is live in \(S\)} \\  
+        & \text{implies }
+            s \transvia{α ∣ p}_{\norm(S)} s' 
+    \end{align*}
+    Finally, the rejection case: \((s, u) \transRej{\norm(≈)}{α}\) either when \((s, u) \transvia{α ∣ p}_{≈} (s', u')\) and \((s', u')\) is dead, or \((s, u) \transRej{≈}{α}\).
+    We will show that both cases will lead to \(s \transRej{\norm(S)}{α}\).
+    The \((s, u) \transvia{α ∣ p}_{≈} (s', u')\) case can be proven by the definition of \(δ_≈\) and~\Cref{thm:greedy-bisim-perserve-liveness}:
+    \begin{align*}
+        & (s, u) \transvia{α ∣ p}_{≈} (s', u') 
+            \text{ and \((s', u')\) is dead in \(≈\)} \\
+        \text{implies } & s \transvia{α ∣ p}_{S} s' 
+            \text{ and \(s'\) is dead in \(S\)} \\  
+        \text{implies } & s \transRej{\norm(S)}{α}.
+    \end{align*}
+    Then the hardest case is when \((s, u) \transRej{≈}{α}\) because by definition of \(δ_≈\) (\Cref{def:norm-bisim}) there are many cases that can lead to rejection.
+    We will need to eliminate irrelevant cases using the condition provided by the construction.
+    We proceed by case analysis on dynamics of \(s\) and \(u\) on \(α\); first listing some cases that contradicts the premise \((s, u) \transRej{≈}{α}\):
+    \begin{itemize}
+        \item If \(s\) accepts \(α\) in \(S\), then \(u\) necessarily accepts \(α\) in \(U\) by condition~\labelcref{itm:acc-condition}, which implies \((s, u) \transAcc{≈}{α}\), contradicting the premise.
+        \item Similarly, if \(u\) accepts \(α\) in \(U\), then \((s, u) \transAcc{≈}{α}\), contradicting the premise.
+        \item If \(s \transvia{α ∣ p}_S s'\) and \(u \transvia{α ∣ p}_U u'\) then \((s, u) \transvia{α ∣ p}_≈ (s', u;)\), contradicting the premise.
+    \end{itemize}
+    Then we list the other cases, which indeed satisfy \((s, u) \transRej{≈}{α}\), and show that all of them leads to \(s \transRej{\norm(S)}{α}\), using the condition in the theorem:
+    \begin{itemize}
+        \item If \(s \transRej{S}{α}\), then \(s \transRej{\norm(S)}{α}\).
+        \item If \(s \transvia{α ∣ p}_S s'\) and \(u \transRej{U}{α}\), then by condition~\labelcref{itm:rej-or-dead}, \(s'\) is dead, and \(s \transRej{\norm(S)}{α}\).
+        \item If \(s \transvia{α ∣ p}_S s'\) and \(u \transvia{α ∣ q}_U u'\), then by condition~\labelcref{itm:transition-dead}, both \(s'\) and \(u'\) are dead, and \(s \transRej{\norm(S)}{α}\).
+        \qedhere
+    \end{itemize}
 \end{proof}
@@
 \begin{lemma}\label{thm:bisim-between-dead}
-    Given any two dead states \(s ∈ S\) and \(u ∈ U\), and the maximal bisimulation \({≡} ⊆ \norm(S) × \norm(U)\), then we have \(s ≡ u\).
+    Given any two dead states \(s ∈ S\) and \(u ∈ U\), we can construct a bisimulation \(∼\) between \(\norm(S)\) and \(\norm(U)\):
+    \begin{align*}
+        ∼ &≜ \{(s, u)\}, & δ_∼((s, u)) & ≜ \reject.
+    \end{align*}
+    Therefore, for the maximal bisimulation \({≡} ⊆ \norm(S) × \norm(U)\), \(s ≡ u\).
 \end{lemma}
@@
 \begin{theorem}[Equivalence Relation]
-    Given two states in GKAT coalgebra \(s ∈ S\) and \(u ∈ U\), there exists a normalizing bisimulation \(≈\) between \(S\) and \(U\) s.t.
-\(s ≈ u\), if and only if there exists a normalizing bisimulation that is an \emph{equivalence relation} \(≊\) in \(S + U\) s.t.
+    Given two states in GKAT coalgebra \(s ∈ S\) and \(u ∈ U\), there exists a greedy bisimulation \(≈\) between \(S\) and \(U\) s.t.
+\(s ≈ u\), if and only if there exists a greedy bisimulation that is an \emph{equivalence relation} \(≊\) in \(S + U\) s.t.
 \(s ≊ u\).
@@
 \begin{proof}
-    First, recall that by~\Cref{thm:} \(\norm(≈)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\).
+    First, recall that by~\Cref{thm:greedy-bisim-iff-norm-bisim} \(\norm(≈)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\).
     Because \(S ⊑ S + U\) and \(U ⊑ S + U\) and normalization preserves sub-coalgebra~\Cref{thm:norm-sub-coalg}, we have \(\norm(S) ⊑ \norm(S + U)\) and \(\norm(U) ⊑ \norm(S + U)\); and by~\Cref{thm:sub-coalg-preserve-liveness}, \(s ∈ S\) is dead in \(S + U\) if and only if it is dead in \(S\), and similarly for \(u ∈ U\).
@@
     Because sub-coalgebra preserves bisimulation (\Cref{thm:sub-coalg-preserve-bisim}), then \(\norm(≈)\) is a bisimulation between \(\norm(S + U)\) and itself.
-    We consider \(≡\) to be the maximal bisimulation between \(\norm(S + U)\) and itself, then \(≡\) is an equivalence relation~\cite[]{}.
-
-    Let \(≊\) be the normalizing bisimulation generated by \(≡\), then \(≊\) has the same carrier as \(≡\).
+    We consider \(≡\) to be the maximal bisimulation between \(\norm(S + U)\) and itself, then \(≡\) is an equivalence relation~\cite[Corollary 5.6]{rutten_UniversalCoalgebraTheory_2000}.
+
+    Let \(≊\) be the greedy bisimulation generated by \(≡\), then \(≊\) has the same carrier as \(≡\).
     All the \((s, u) ∈ {≊}\) are either \(s ≡ u\) or \(s\) and \(u\) are dead in \(S\) and \(U\), which implies that both are dead in \(S + U\).
     However, \(≡\) already contains all the dead pairs (\cref{thm:bisim-between-dead}), therefore for all \((s, u) ∈ {≊}\), they are also in \(≡\), implying \({≡} ⊇ {≊}\).
-    Finally, by definition of normalizing bisimulation (\Cref{def:norm-bisim}), \({≡} ⊆ {≊}\).
-    Therefore, \(≡\) and \(≊\) share the same carrier and \(≊\) is a normalizing bisimulation that is also an equivalence relation, which also contains \((s, u)\):
+    Finally, by definition of greedy bisimulation (\Cref{def:norm-bisim}), \({≡} ⊆ {≊}\).
+    Therefore, \(≡\) and \(≊\) share the same carrier and \(≊\) is a greedy bisimulation that is also an equivalence relation, which also contains \((s, u)\):
     \[(s, u) ∈ {≈} = \norm(≈) ⊆ {≡} ⊆ {≊},\]
@@
     \begin{itemize}
-        \item If \(s ≃ u\), then we can reflect \(≃\)  through \(\norm(S) ⊑ \norm(S + U)\) and \(\norm(U) ⊑ \norm(S + U)\) (\Cref{thm:sub-coalg-preserve-bisim}) to obtain a bisimulation \(∼ ⊆ \norm(S) × \norm(U)\) and \(s ∼ u\).
-        Taking the normalizing bisimulation over \(∼\) will yield a desired normalizing bisimulation \(≈\) s.t. \(s ≈ u\).
+        \item If \(s ≃ u\), then we can reflect \(≃\)  through \(\norm(S) ⊑ \norm(S + U)\) and \(\norm(U) ⊑ \norm(S + U)\) (\Cref{thm:sub-coalg-preserve-bisim}) to obtain a bisimulation \({∼} ⊆ \norm(S) × \norm(U)\) and \(s ∼ u\).
+        Taking the greedy bisimulation over \(∼\) will yield a desired greedy bisimulation \(≈\) s.t. \(s ≈ u\).
         \item If \((s, u)\) are not in \(≃\), then they are both dead in \(S + U\) (\Cref{thm:norm-bisim-not-bisim-then-component-dead}), which implies \(s\) and \(u\) are respectively dead in \(S\) and \(U\) (\Cref{thm:sub-coalg-preserve-liveness}).
-        Then we take \(≈\) to be normalizing bisimulation generated form the maximal bisimulation between \(S\) and \(U\), which contains \(s\) and \(u\) (\Cref{thm:bisim-between-dead}).
+        Then we take \(≈\) to be greedy bisimulation generated form the maximal bisimulation between \(S\) and \(U\), which contains \(s\) and \(u\) (\Cref{thm:bisim-between-dead}).
         \qedhere
@@
 
-% % \begin{definition}[normalizing bisimulation]\label{def:norm-bism}
-% %     A relation \({≃} ⊆ S × U\) on two GKAT coalgebra \(S\) and \(U\) is a normalizing bisimulation if and only if for any two states \(s ≃ u\), all the following holds:
+% % \begin{definition}[greedy bisimulation]\label{def:norm-bism}
+% %     A relation \({≃} ⊆ S × U\) on two GKAT coalgebra \(S\) and \(U\) is a greedy bisimulation if and only if for any two states \(s ≃ u\), all the following holds:
 % %     \begin{enumerate}
@@
 % \begin{lemma}[Perservation of Liveness]\label{thm:norm-bisim-perserve-liveness}
-%     For any normalizing bisimulation \(≃\) and a pair of states \(s ≃ u\), \((s, u)\) is live in \(≃\) if and only if \(s\) is live in \(S\), and similarly for \(u\) in \(U\). 
+%     For any greedy bisimulation \(≃\) and a pair of states \(s ≃ u\), \((s, u)\) is live in \(≃\) if and only if \(s\) is live in \(S\), and similarly for \(u\) in \(U\). 
 %     This implies that if \(s ≃ u\), then \(s\) is live in \(S\) if and only if \(u\) is live in \(U\).
@@
 % \begin{proof}
-%     First by definition of normalizing bisimulation that a state \(s\) is accepting in \(S\) if and only if \((s, u)\) is accepting in \(≃\).
+%     First by definition of greedy bisimulation that a state \(s\) is accepting in \(S\) if and only if \((s, u)\) is accepting in \(≃\).
 
@@
 % \begin{theorem}
-%     Given a normalizing bisimulation \(≃: S × U\), its normalization is indeed a bisimulation between \(\norm(S)\) and \(\norm(U)\)
+%     Given a greedy bisimulation \(≃: S × U\), its normalization is indeed a bisimulation between \(\norm(S)\) and \(\norm(U)\)
 % \end{theorem}
@@
 
-%     Second the transition case, for \(s ≃ u\), because normalizing bisimulation preserves liveness (\Cref{thm:norm-bisim-perserve-liveness}):
+%     Second the transition case, for \(s ≃ u\), because greedy bisimulation preserves liveness (\Cref{thm:norm-bisim-perserve-liveness}):
 %     \begin{align*}
@@
 %     For any two GKAT coalgebra \(S\) and \(U\), the maximal bisimulation \(≡\) between \(\norm(S)\) and \(\norm(U)\) satisfy all the conditions in~\Cref{def:norm-bism}, i.e.
-the relation \(≡\) can be a normalizing bisimulation between \(S\) and \(U\).
+the relation \(≡\) can be a greedy bisimulation between \(S\) and \(U\).
 % \end{theorem}
@@
 % \end{proof}
+
+\begin{theorem}\label{thm:greedy-bisim-one-dead}
+
+\end{theorem}
 
@@
 
-Finally, we present our equivalence-checking algorithm as~\cref{alg:bisim}, where we first check if one of \(s\) or \(u\) is known to be dead, if so we only need to check whether the other is dead, because of~\Cref{thm:bisim-one-dead}; otherwise, we will check the conditions in~\Cref{thm:recursive-construction}.
+Finally, we present our equivalence-checking algorithm as~\cref{alg:bisim}, where we first check if one of \(s\) or \(u\) is known to be dead, if so we only need to check whether the other is dead, because of~\Cref{thm:greedy-bisim-one-dead}; otherwise, we will check the conditions in~\Cref{thm:recursive-construction}.
 Because trace equivalence is an equivalence relation, therefore we can organize the previously explored states into equivalent classes using an efficient union-find structure, instead of a set of state pairs.

### 2025-01-16 main.tex

@@
 Boston, USA \\
-email address or ORCID}
+qcfu@bu.edu}
 \and

### 2025-01-16 main.tex
minor fix

@@
     \[{∼_d} ≜ {∼} ∪ \{(s, u) ∣ \text{\(s\) and \(u\) are dead in \(S\) and \(U\)}\},\]
-    form a GKAT coalgebra the transition function \(δ_≈\) in~\Cref{def:norm-bisim}, which implies \({≈} ⊑ \)
+    form a GKAT coalgebra the transition function \(δ_≈\) in~\Cref{def:norm-bisim}, which implies \({≈} ⊑ {∼_d}\).
     Furthermore, \(\norm(∼_d)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\).
@@
 \begin{theorem}\label{thm:greedy-bisim-one-dead}
-
+    Consider two states in GKAT coalgebras \(s ∈ S, u ∈ U\) and an arbitrary bisimulation \(≈\), if \(s\) is dead, then 
+    \begin{itemize}
+        \item if \(u\) is not dead, then \((s, u) ∉ {≈}\).
+        \item if \(u\) is dead, 
+    \end{itemize}
 \end{theorem}

### 2025-01-16 main.tex
final proof

@@
 \begin{theorem}\label{thm:greedy-bisim-one-dead}
-    Consider two states in GKAT coalgebras \(s ∈ S, u ∈ U\) and an arbitrary bisimulation \(≈\), if \(s\) is dead, then 
-    \begin{itemize}
-        \item if \(u\) is not dead, then \((s, u) ∉ {≈}\).
-        \item if \(u\) is dead, 
-    \end{itemize}
+    Consider two states in GKAT coalgebras \(s ∈ S, u ∈ U\) and an arbitrary bisimulation \({∼} ⊆ \norm(S) × \norm(U)\), if \(s\) is dead, then 
+    \begin{enumerate}
+        \item\label{itm:greedy-bisim-one-dead-then-both-dead} if \(s ∼ u\), then \(u\) is dead.
+        \item\label{itm:remove-dead-pairs-sound} if \(u\) is dead, then \({∼} ∖ ⟨(s, u)⟩_∼\) is also a bisimulation between \(\norm(S)\) and \(\norm(U)\).
+    \end{enumerate}
 \end{theorem}
+
+\begin{proof}
+    \Cref{itm:greedy-bisim-one-dead-then-both-dead} is a consequence of bisimulation preserves liveness (\Cref{thm:bisim-preserve-liveness}) and normalization preserves liveness (\Cref{thm:norm-perserve-liveness}).
+
+    \Cref{itm:remove-dead-pairs-sound} can be proven by realizing that \((s, u)\) is a dead state in \(∼\) (\Cref{thm:bisim-preserve-liveness}) which means that all the states in \(⟨(s, u)⟩_∼\) are dead.
+    Because \(∼\) is a normalized GKAT coalgebra (\Cref{thm:}), therefore no state can transition into dead states like \(⟨(s, u)⟩_∼\), hence \({∼} ∖ ⟨(s, u)⟩_∼\) is still closed under its transition, and form a GKAT coalgebra.
+    And the fact that \({∼} ∖ ⟨(s, u)⟩_∼\) is also a bisimulation can be proven by factoring through \(∼\) via the inclusion homomorphism:
+    \begin{align*}
+        ({∼} ∖ ⟨(s, u)⟩_∼) \xrightarrow{i} {∼} \xrightarrow{π₁} \norm(S) \\
+        ({∼} ∖ ⟨(s, u)⟩_∼) \xrightarrow{i} {∼} \xrightarrow{π₂} \norm(U)
+    \end{align*}
+    TODO: extract the above to a lemma.
+\end{proof}
 

### 2025-01-16 main.tex
minor tweak

@@
 
-In this paper, we will make heavy use of coalgebraic theory, thus it is empirical for us to recall some notions and useful theorems in universal coalgebra.
+In this paper, we will make heavy use of coalgebraic theory, thus we will recall useful notions and theorems in universal coalgebra.
 Given a functor \(F\) on the category of set and functions, a \emph{coalgebra over \(F\)} or \emph{\(F\)-coalgebra} consists of a set \(S\) and a function \(σ_S: S → F(S)\).
@@
     By definition of image, \(h(⟨s⟩_{S})\) indeed contain \(h(s)\). 
-    By \Cref{thm:hom-(pre)img-preserve-sub-coalg}, \(h(⟨s⟩_S) ⊑ U\), i.e. \(h(⟨s⟩_S)\) is a sub-coalgebra of\(U\).
+    By \Cref{thm:hom-(pre)img-preserve-sub-coalg}, \(h(⟨s⟩_S) ⊑ U\), i.e. \(h(⟨s⟩_S)\) is a sub-coalgebra of \(U\).
     Finally, take any sub-coalgebra \(U' ⊑ U\) s.t. \(h(s) ∈ U'\): 
@@
 Traditionally, live and dead states are defined by whether they can reach an accepting state~\cite{smolka_GuardedKleeneAlgebra_2020}.
-However, we can use induction on the length of the path to show that principle sub-coalgebra \(⟨s⟩_S\) contains exactly the reachable states of \(s\) in any GKAT coalgebra \(S\). 
-Thus, the classical definition is equivalent to the following:
-\begin{definition}[liveness of states]\label{def:liveness-of-states}
-    A state \(s\) is \emph{accepting} if there exists an atom \(α ∈ \At\) s.t.
-\(δ(s, α) = \accept\); a state \(s'\) is \emph{live} if there exists an accepting state \(s' ∈ ⟨s⟩\); and a state \(s'\) is \emph{dead} if there is no accepting state in \(⟨s⟩\).
+\begin{definition}[Liveness]\label{def:liveness-of-states}
+    A state \(s\) is \emph{accepting} if there exists an atom \(α ∈ \At\) s.t. \(δ(s, α) = \accept\).
+    A state \(s₀\) is \emph{live} if there exists an accepting state \(sₙ\) s.t. there is a liveness chain:
+    \[s₀ \transvia{α₁ ∣ p₁} s₁ \transvia{α₂ ∣ p₂} s₂ ⋯ \transvia{αₙ ∣ pₙ} sₙ.\]
+    A state is \emph{dead} if it is not dead.
 \end{definition}
-This alternative liveness definition can help us prove important theorems regarding reachability and liveness without explicitly performing induction on traces. 
+However, besides directly working with this explicit definition, we can also give a more coalgebraic characterization of these properties.
+Specifically, by induction on the length of the path from one state to the next, we can show that the principle sub-coalgebra \(⟨s⟩_S\) exactly contains all the reachable state of \(s\) in the GKAT coalgebra \(S\).
+Thus we obtain the following coalgebraic characterization of liveness:
+\begin{theorem}\label{thm:liveness-principle-sub-coalg}
+    A state \(s\) is \emph{live} if there exists an accepting state \(s' ∈ ⟨s⟩\); and a state \(s\) is \emph{dead} if there is no accepting state in \(⟨s⟩\).
+\end{theorem}
+This alternative liveness definition can help us prove important theorems regarding reachability and liveness without explicitly performing induction on the liveness chain. 
 We show the following theorems as examples:

### 2025-01-16 main.tex
check point recursive construction

@@
 
-The original algorithm for deciding GKAT equivalences~\cite{smolka_GuardedKleeneAlgebra_2020} requires the entire coalgebra to be known prior to the execution of the bisimulation algorithm; specifically, it is necessary to iterate through the entire coalgebra in order to identify the liveness of every single states, in order to perform the normalization operation.
+The original algorithm for deciding GKAT equivalences~\cite{smolka_GuardedKleeneAlgebra_2020} requires the entire coalgebra to be known prior to the execution of the bisimulation algorithm; specifically, it is necessary to iterate through the entire coalgebra in order to identify the liveness of every single states and perform the normalization operation.
 This limitation poses challenges to design an efficient on-the-fly algorithm for GKAT.
-In order to make the decision procedure scalable, we will need to merge the normalization and bisimulation procedure, so that our algorithm can normalize the coalgebra only when we need to.
-
-In this section, we introduce an algorithm that merges bisimulation and normalization, where we only test the liveness of states when a disparity is noticed by the bisimulation algorithm.
+To make the equivalence-checking procedure scalable, we propose to greedily run the bisimulation algorithm and only invoke liveness detection when a discrepancy is found. 
 For example, when deciding the trace equivalence between \(s ∈ S\) and \(u ∈ U\), if \(δ_S(s, α) = (s', p)\) and \(δ_U(u, α) = (u', p)\), we will proceed to recurse on \(s'\) and \(u'\), without checking the liveness of \(s'\) and \(u'\).
@@
 
-We present several sound and complete conditions for finite-trace equivalence between states of GKAT coalgebras in~\Cref{thm:recursive-construction}, which also serves as the core correctness theorem for our equivalence-checking algorithm i.e.~\Cref{alg:bisim}.
-This algorithm mostly consists of recursively checking the conditions in~\Cref{thm:recursive-construction} with minor optimizations, which we will outline later.
-
-\begin{definition}[greedy bisimulation]\label{def:norm-bisim}
-    Given two GKAT coalgebra \(S\) and \(U\), and a bisimulation \({∼} ⊆ \norm(S) × \norm(U)\), a greedy bisimulation \({≈} ⊆ S × U\) is a GKAT coalgebra that contains \(∼\) with the following transition function:
+This approach offers several benefit: first, because the liveness detection is only called when necessary, thus it is unlikely to iterate through the entire coalgebra; second, the algorithm can short-circuit on a counter-example, allowing termination even before all the states are explored; finally, the algorithm supports on-the-fly constructions like derivatives, because the coalgebra no longer is required to be fully constructed.
+
+The correctness argument of this on-the-fly algorithm is centered around the following coalgebraic structure, which we named ``greedy bisimulation''.
+
+\begin{definition}[Greedy Bisimulation]\label{def:greedy-bisim}
+    Given two GKAT coalgebra \(S\) and \(U\), and a bisimulation \({∼} ⊆ \norm(S) × \norm(U)\), a greedy bisimulation \({≈} ⊆ S × U\) is the least GKAT coalgebra that contains \(∼\) with the following transition function:
     \[
@@
 
+Given any bisimulation \({∼} ⊆ S × U\), we can construct a GKAT coalgebra \(∼_d\) s.t.
+it is almost a greedy bisimulation, i.e.
+it is closed under \(δ_≈\) in~\Cref{def:greedy-bisim}, and contains \(∼\), except it is not the \emph{least} GKAT coalgebra that satisfy these properties.
+
 \begin{lemma}\label{thm:dead-construction-greedy-bisim}
@@
     \[{∼_d} ≜ {∼} ∪ \{(s, u) ∣ \text{\(s\) and \(u\) are dead in \(S\) and \(U\)}\},\]
-    form a GKAT coalgebra the transition function \(δ_≈\) in~\Cref{def:norm-bisim}, which implies \({≈} ⊑ {∼_d}\).
+    form a GKAT coalgebra the transition function \(δ_≈\) in~\Cref{def:greedy-bisim}, which implies \({≈} ⊑ {∼_d}\).
     Furthermore, \(\norm(∼_d)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\).
@@
 \begin{proof}
-    We will need to show \(∼_d\) is closed under the transition of \(δ_{≈}\) in~\Cref{def:norm-bisim}.
+    We will need to show \(∼_d\) is closed under the transition of \(δ_{≈}\) in~\Cref{def:greedy-bisim}.
     Specifically, if \(s ∼_d u\) and \(s \transvia{α ∣ p}_{S} s'\) and \(u \transvia{α ∣ p} u'\), then \(s' ∼_d u'\).
@@
 
-\begin{lemma}\label{thm:norm-bisim-exists-unique}
+Although the proof of~\Cref{thm:dead-construction-greedy-bisim} is basically by unfolding the definition, \(∼_d\) turns out to be a important structure. 
+Specifically, \(∼_d\) serves a connection between \(≈\) and its underlying bisimulation \(∼\), and the order \({∼} ⊆ {≈} ⊑ {~_d}\) will reveal many structural properties of \(≈\).
+We show the following corollary to demonstrate the usefulness of \(∼_d\).
+
+\begin{corollary}\label{thm:norm-bisim-exists-unique}
     For any bisimulation between two normalized GKAT coalgebra \({∼} ⊆ \norm(S) × \norm(U)\), a greedy bisimulation \(≈\) with \(∼\) as underlying bisimulation exists and is unique.
-\end{lemma}
+\end{corollary}
 
@@
 
-\begin{lemma}\label{thm:norm-bisim-not-bisim-then-component-dead}
-    Every state in \((s, u) ∈ {≈} ⊆ S × U\) but not in \({∼} ⊆ \norm(S) × \norm(U)\) then \(s\) is dead in \(S\), \(u\) is dead in \(U\).
-\end{lemma}
+\begin{corollary}\label{thm:greedy-bisim-dead-or-bisim}
+    For every pair of states \((s, u)\) in a greedy bisimulation \({≈} ⊆ S × U\), either \((s, u)\) is in the underlying bisimulation \(∼\), or both \(s\) and \(u\) are dead in \(S\) and \(U\).
+\end{corollary}
 
@@
 
-\begin{theorem}[Perservation Of Liveness]\label{thm:greedy-bisim-perserve-liveness}
+\begin{corollary}[Perservation Of Liveness]\label{thm:greedy-bisim-perserve-liveness}
     Given a pair of states in a greedy bisimulation \(s ≈ u\) where \({≈} ⊆ S × U\), then the following are equivalent
@@
     \end{itemize}
-    Furthermore, if \((s, u)\) is in the underlying bisimulation \(∼\) of \(≈\), then all the above is also equivalent to \((s, u)\) is live (or dead) in \(∼\).
-\end{theorem}
+    Furthermore, if \((s, u)\) is in the underlying bisimulation \(∼\) of \(≈\), then all of above are also equivalent to \((s, u)\) is live (or dead) in \(∼\).
+\end{corollary}
 
@@
 
-\begin{lemma}\label{thm:greedy-bisim-iff-norm-bisim}
-    Given two GKAT coalgebras \(S\) and \(U\), and a relational GKAT coalgebra \({≈} ⊆ S × U\) with the transition \(δ_≈\) as in~\cref{def:norm-bisim}.
+\begin{corollary}\label{thm:greedy-bisim-iff-norm-bisim}
+    Given two GKAT coalgebras \(S\) and \(U\), and a relational GKAT coalgebra \({≈} ⊆ S × U\) with the transition \(δ_≈\) as in~\cref{def:greedy-bisim}.
     \(≈\) is a greedy bisimulation if and only if \(\norm(≈)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\)
-\end{lemma}
+\end{corollary}
 
@@
     Because \({≈} ⊑ {∼_d}\) and~\Cref{thm:norm-sub-coalg}, \(\norm(≈) ⊑ \norm(∼_d)\). 
-    Because \(\norm(∼_d)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\), then we can show that \(\norm(≈)\) is a bisimulation by factoring through \(\norm(∼_d)\), in the category of GKAT coalgebras:
-    \begin{align*}
-        & \norm(≈) 
-        \xrightarrow{i} \norm(∼_d) 
-        \xrightarrow{π₁} \norm(S) \\
-        & \norm(≈) 
-        \xrightarrow{i} \norm(∼_d) 
-        \xrightarrow{π₂} \norm(U) 
-    \end{align*}
-    because inclusion homomorphism \(i\) do not change its input, the two homomorphism obtained by bisimulation is indeed \(π₁: \norm(≈) → \norm(S)\) and \(π₂: \norm(≈) → \norm(U)\), implying both projections of \(\norm(≈)\) are homomorphism, meaning \(\norm(≈)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\).
+    Because \(\norm(∼_d)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\), therefore \(\norm(≈)\) is a bisimulation by~\cref{thm:}
 
@@
 \end{proof}
+
+We then establish the correctness of greedy bisimulation, i.e. it is equivalent to trace equivalence.
 
@@
 \begin{proof}
-    \(⟹\) direction, because normalization perserves carrier set, therefore \((s, u) ∈ \norm(≈)\).
-    And because \(\norm(≈)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\), \(⟦s⟧_S = ⟦u⟧_U\).
+    \(⟹\) direction, because normalization preserves carrier set, therefore \((s, u) ∈ \norm(≈)\).
+    And because \(\norm(≈)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\) (\Cref{thm:greedy-bisim-iff-norm-bisim}), \(⟦s⟧_S = ⟦u⟧_U\).
 
@@
 \end{proof}
+
+Despite proving many desirable structural properties, the definition of greedy bisimulation is still coalgebraic.
+To design an equivalence-checking algorithm to construct a greedy bisimulation still requires turning the coalgebra into a set of concrete criteria, that can be checked by an algorithm.
+In the following theorem, we will present four conditions, and the algorithm will be largely based on recursively checking these conditions to establish a greedy bisimulation.
 
@@
 
-    The condition~\labelcref{itm:transition-bisim} holds because when \(s \transvia{α ∣ p}_S s'\) and \(u \transvia{α ∣ p}_U u'\), then \((s, u) \transvia{α ∣ p}_{≈} (s', u')\) by definition of greedy bisimulation (\Cref{def:norm-bisim}).
+    The condition~\labelcref{itm:transition-bisim} holds because when \(s \transvia{α ∣ p}_S s'\) and \(u \transvia{α ∣ p}_U u'\), then \((s, u) \transvia{α ∣ p}_{≈} (s', u')\) by definition of greedy bisimulation (\Cref{def:greedy-bisim}).
 
@@
         \item If \(s ∼ u\), then we can show condition~\labelcref{itm:transition-dead} by case analysis on the liveness of \(s'\) and \(u'\):
-        if \(s', u'\) are not both dead, then the condition of \({∼} ⊆ \norm(S) × \norm(U)\) cannot be satisfied, hence \((s', u') ∉ {∼}\), which by~\Cref{thm:norm-bisim-not-bisim-then-component-dead} means that \(s'\) and \(u'\) are both dead, contradicting the premise.
+        if \(s', u'\) are not both dead, then the condition of \({∼} ⊆ \norm(S) × \norm(U)\) cannot be satisfied, hence \((s', u') ∉ {∼}\), which by~\Cref{thm:greedy-bisim-dead-or-bisim} means that \(s'\) and \(u'\) are both dead, contradicting the premise.
         Therefore, \(s'\) and \(u'\) needs to be both dead.
@@
     \end{align*}
-    Then the hardest case is when \((s, u) \transRej{≈}{α}\) because by definition of \(δ_≈\) (\Cref{def:norm-bisim}) there are many cases that can lead to rejection.
+    Then the hardest case is when \((s, u) \transRej{≈}{α}\) because by definition of \(δ_≈\) (\Cref{def:greedy-bisim}) there are many cases that can lead to rejection.
     We will need to eliminate irrelevant cases using the condition provided by the construction.
@@
     First, recall that by~\Cref{thm:greedy-bisim-iff-norm-bisim} \(\norm(≈)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\).
-    Because \(S ⊑ S + U\) and \(U ⊑ S + U\) and normalization preserves sub-coalgebra~\Cref{thm:norm-sub-coalg}, we have \(\norm(S) ⊑ \norm(S + U)\) and \(\norm(U) ⊑ \norm(S + U)\); and by~\Cref{thm:sub-coalg-preserve-liveness}, \(s ∈ S\) is dead in \(S + U\) if and only if it is dead in \(S\), and similarly for \(u ∈ U\).
+    Because \(S ⊑ S + U\) and \(U ⊑ S + U\) and normalization preserves sub-coalgebra (\Cref{thm:norm-sub-coalg}), we have \(\norm(S) ⊑ \norm(S + U)\) and \(\norm(U) ⊑ \norm(S + U)\); and by~\Cref{thm:sub-coalg-preserve-liveness}, \(s ∈ S\) is dead in \(S + U\) if and only if it is dead in \(S\), and similarly for \(u ∈ U\).
 
@@
     However, \(≡\) already contains all the dead pairs (\cref{thm:bisim-between-dead}), therefore for all \((s, u) ∈ {≊}\), they are also in \(≡\), implying \({≡} ⊇ {≊}\).
-    Finally, by definition of greedy bisimulation (\Cref{def:norm-bisim}), \({≡} ⊆ {≊}\).
+    Finally, by definition of greedy bisimulation (\Cref{def:greedy-bisim}), \({≡} ⊆ {≊}\).
     Therefore, \(≡\) and \(≊\) share the same carrier and \(≊\) is a greedy bisimulation that is also an equivalence relation, which also contains \((s, u)\):
@@
         Taking the greedy bisimulation over \(∼\) will yield a desired greedy bisimulation \(≈\) s.t. \(s ≈ u\).
-        \item If \((s, u)\) are not in \(≃\), then they are both dead in \(S + U\) (\Cref{thm:norm-bisim-not-bisim-then-component-dead}), which implies \(s\) and \(u\) are respectively dead in \(S\) and \(U\) (\Cref{thm:sub-coalg-preserve-liveness}).
+        \item If \((s, u)\) are not in \(≃\), then they are both dead in \(S + U\) (\Cref{thm:greedy-bisim-dead-or-bisim}), which implies \(s\) and \(u\) are respectively dead in \(S\) and \(U\) (\Cref{thm:sub-coalg-preserve-liveness}).
         Then we take \(≈\) to be greedy bisimulation generated form the maximal bisimulation between \(S\) and \(U\), which contains \(s\) and \(u\) (\Cref{thm:bisim-between-dead}).
@@
 \end{proof}
-
-% % \begin{definition}[Normalizable Bisimulation]
-% %     A relation \({≃} ⊆ S × U\) is a normalizable bisimulation if it has a GKAT coalgebra structure \(δ_≃\), and \(\norm(≃)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\).
-% % \end{definition}
-
-% % \begin{lemma}[Preservation of Liveness]
-% %     For any normalizable bisimulation \({≃} ⊆ S × U\) if \(s ≃ u\), then \(s\) is live in \(S\) if and only if \(u\) is live in \(U\).
-% % \end{lemma}
-
-% % \begin{proof}
-% %     Because \(\norm(≃)\) has two projection homomorphism to \(\norm(S)\) and \(\norm(U)\), then 
-% % \end{proof}
-
-% % \begin{theorem}[Correctness]
-% %     If there exists a normalizable bisimulation \({≃} ⊆ S × U\) s.t. \(s ∼ u\) for \(s ∈ S\) and \(u ∈ U\), then \(⟦s⟧_S = ⟦u⟧_U\).
-% % \end{theorem}
-
-% % \begin{theorem}[Construction]
-% %     Given two states \(s ∈ S, u ∈ U\) then there exists a normalizable bisimulation \({≃'} ⊆ S × U\) if and only if there exists a relation \({≃} ⊆ S × U\) that satisfy the following conditions:
-% %     \begin{enumerate}
-% %         \item\label{itm:acc-condition} for all \(α ∈ \At\), \(δ_{S}(s, α) = \accept ⟺ δ_{U}(u, α) = \accept\);
-% %         \item\label{itm:rej-or-dead} if \(s\) reject \(α\) but \(u\) transitions to \(u'\), then \(u'\) is dead; similarly when \(u\) rejects \(α\);  
-% %         \item\label{itm:transition-bisim} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', p)\), then \(s' ≃ u'\).
-% %         \item\label{itm:transition-dead} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', q)\), s.t. \(p ≠ q\), then both \(s'\) and \(t'\) are dead.
-% %     \end{enumerate}
-% % \end{theorem}
-
-% % \begin{proof}
-% %     The \(⟸\) direction can be shown that 
-% % \end{proof}
-
-% % \begin{definition}[greedy bisimulation]\label{def:norm-bism}
-% %     A relation \({≃} ⊆ S × U\) on two GKAT coalgebra \(S\) and \(U\) is a greedy bisimulation if and only if for any two states \(s ≃ u\), all the following holds:
-% %     \begin{enumerate}
-% %         \item\label{itm:acc-condition} for all \(α ∈ \At\), \(δ_{S}(s, α) = \accept ⟺ δ_{U}(u, α) = \accept\);
-% %         \item\label{itm:rej-or-dead} if \(s\) reject \(α\) but \(u\) transitions to \(u'\), then \(u'\) is dead; similarly when \(u\) rejects \(α\);  
-% %         \item\label{itm:transition-bisim} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', p)\), then \(s' ≃ u'\).
-% %         \item\label{itm:transition-dead} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', q)\), s.t. \(p ≠ q\), then both \(s'\) and \(t'\) are dead.
-% %     \end{enumerate}
-% %     And \(≃\) can be endowed with the following GKAT coalgebra structure:
-% %     \[
-% %         δ_≃((s, u), α) ≜ \begin{cases}
-% %             \accept & \begin{aligned}
-% %                 & \text{if } δ_S(s, α) = \accept \\[-3px]
-% %                 & \text{and } δ_U(s, α) = \accept
-% %             \end{aligned}\\[5px]
-% %             ((s', u'), p) & 
-% %                 \begin{aligned}
-% %                     & \text{if } s \transvia{α ∣ p}_{S} s' \\[-3px]
-% %                     & \text{and } u \transvia{α ∣ p}_{U} u' 
-% %                 \end{aligned}\\[5px]
-% %             \reject & \text{otherwise}
-% %         \end{cases}
-% %     \]
-% % \end{definition}
-
-% \begin{lemma}[Perservation of Liveness]\label{thm:norm-bisim-perserve-liveness}
-%     For any greedy bisimulation \(≃\) and a pair of states \(s ≃ u\), \((s, u)\) is live in \(≃\) if and only if \(s\) is live in \(S\), and similarly for \(u\) in \(U\). 
-%     This implies that if \(s ≃ u\), then \(s\) is live in \(S\) if and only if \(u\) is live in \(U\).
-% \end{lemma}
-
-% \begin{proof}
-%     First by definition of greedy bisimulation that a state \(s\) is accepting in \(S\) if and only if \((s, u)\) is accepting in \(≃\).
-
-%     Then, to prove the preservation of liveness, we prove the preservation of live chain, i.e.
-there exists a chain to accepting state in \(≃\): \((s₀, u₀) \transvia{α₁ ∣ p₁} (s₁, u₁) \transvia{α₂ ∣ p₂} (s₂, u₂) \transvia{α₂ ∣ p₂} ⋯ \transvia{αₙ ∣ pₙ} (sₙ, uₙ)\) if and only if there exists a similar accepting chain in \(S\): \(s₀ \transvia{α₁ ∣ p₁} s₁ \transvia{α₂ ∣ p₂} s₂ \transvia{α₂ ∣ p₂} ⋯ \transvia{αₙ ∣ pₙ} sₙ\).
-
-%     Because every state on an accepting chain is live, all we need to prove is that for two states \(s ≃ u\),
-%     \begin{enumerate}
-%         \item if \((s', u')\) is a live state in \(≃\) and \((s, u) \transvia{α ∣ p}_≃ (s', u')\), then \(s \transvia{α ∣ p}_S s'\)
-%         \item if \(s'\) is a live state in \(S\) and \(s \transvia{α ∣ p} s'\), then \((s, u) \transvia{α ∣ p} (s', u')\) for some \(u'\).
-%     \end{enumerate}
-%     The first goal can be shown by definition of \(δ_{≃}\). 
-%     Because \(s'\) in the second goal is live, therefore only condition~\labelcref{itm:transition-bisim} in~\Cref{def:norm-bism} matches the condition, thus implying \(u \transvia{α ∣ p} u'\) and \((s, u) \transvia{α ∣ p} (s', u')\).
-% \end{proof}
-
-% \begin{theorem}
-%     Given a greedy bisimulation \(≃: S × U\), its normalization is indeed a bisimulation between \(\norm(S)\) and \(\norm(U)\)
-% \end{theorem}
-
-% \begin{proof}
-%     We will need to prove the two projection of \(π₁: {≃} → S\) and \(π₂: {≃} → U\) is a homomorphism, without loss of generality, we will only show that \(π₁\) is a homomorphism.
-
-%     First the accepting case, for all \(s ≃ u\):
-%     \begin{align*}
-%         δ_{\norm(≃)}((s, u), α) = \accept
-%         & ⟹ δ_{≃}((s, u), α) = \accept \\
-%         & ⟹ δ_{S}(s, α) = \accept \\ 
-%         & ⟹ δ_{\norm(S)}(s, α) = \accept \\ 
-%     \end{align*}
-
-%     Second the transition case, for \(s ≃ u\), because greedy bisimulation preserves liveness (\Cref{thm:norm-bisim-perserve-liveness}):
-%     \begin{align*}
-%         & (s, u) \transvia{α ∣ p}_{\norm(≃)} (s', u') \\*
-%         & ⟹ (s, u) \transvia{α ∣ p}_{≃} (s', u')
-%             \text{ and \((s, u)\) is live in \(≃\)}   \\*  
-%         & ⟹ s \transvia{α ∣ p}_{S} s'
-%             \text{ and \(s\) is live in \(S\)} \\* 
-%         & ⟹ s \transvia{α ∣ p}_{\norm(S)} s'
-%     \end{align*}
-
-%     Finally, we show the rejection case. \((s, u) \transRej{\norm(≃)}{α}\) can be split into two cases, we will show both cases lead to \(s \transRej{\norm(S)}{α}\).
-%     \begin{itemize}
-%         \item The first case is that \(δ_{≃}((s, u), α) = \reject\). 
-%         By definition, this case can happen in following three scenarios: 
-%         \begin{itemize}
-%             \item when \(s \transvia{α ∣ p} s'\), \(u \transvia{α ∣ q} u'\), and \(p ≠ q\), which implies \(s'\) is dead by condition~\labelcref{itm:transition-dead};  
-%             \item when \(s \transvia{α ∣ p} s'\) and \(δ_U(u, α) = \reject\), which implies \(s'\) is dead by condition~\labelcref{itm:rej-or-dead}
-%             \item when \(u \transRej{S}{α}\), then \(u \transRej{\norm(S)}{α}\).
-%         \end{itemize}
-%         Thus, all scenarios cases lead to \(u \transRej{\norm(S)}{α}\).
-%         \item The second case is that \((s, u) \transvia{α ∣ p}_{≃} (s', u')\) but \((s', u')\) is dead.
-%         By definition \(s \transvia{α ∣ p}_{S} s'\). 
-%         By~\cref{thm:norm-bisim-perserve-liveness} and \((s', u')\) is dead, we conclude that \(s'\) is dead and therefore \(s' \transRej{\norm(S)}{α}\).
-%     \end{itemize}
-% \end{proof}
-
-% % \begin{lemma}[bisimulation between dead states]\label{thm:bisim-between-dead}
-% %     Given two dead states \(s ∈ S\) and \(u ∈ U\), then the singleton bisimulation \({∼} ⊆ \norm(S) × \norm(U)\):
-% %     \begin{mathpar}
-% %         {∼} ≜ \{(s,u)\} \and 
-% %         δ_{∼}((s,u), α) ≜ \reject
-% %     \end{mathpar}
-% %     is a bisimulation between \(\norm(S)\) and \(\norm(U)\).
-% % \end{lemma}
-
-% % \begin{proof}
-% %     By computation
-% % \end{proof}
-
-% \begin{lemma}[bisimulation between dead states]\label{thm:bisim-between-dead}
-%     Given two dead states \(s ∈ S\) and \(u ∈ U\) and the maximal bisimulation \(≡\) between \(\norm(S)\) and \(\norm(U)\), then \(s ≡ u\).
-% \end{lemma}
-
-% \begin{proof}
-%     We can construct the singleton bisimulation \({∼} ⊆ \norm(S) × \norm(U)\):
-%     \begin{mathpar}% \begin{lemma}[bisimulation between dead states]\label{thm:bisim-between-dead}
-%         %     Given two dead states \(s ∈ S\) and \(u ∈ U\), then the singleton bisimulation \({∼} ⊆ \norm(S) × \norm(U)\):
-%         %     \begin{mathpar}
-%         %         {∼} ≜ \{(s,u)\} \and 
-%         %         δ_{∼}((s,u), α) ≜ \reject
-%         %     \end{mathpar}
-%         %     is a bisimulation between \(\norm(S)\) and \(\norm(U)\).
-%         % \end{lemma}
-
-%         % \begin{proof}
-%         %     By computation
-%         % \end{proof}
-%         {∼} ≜ \{(s,u)\} \and 
-%         δ_{∼}((s,u), α) ≜ \reject,
-%     \end{mathpar}
-%     and verify that it is indeed a bisimulation between \(\norm(S)\) and \(\norm(U)\) by unfolding the definition.
-% \end{proof}
-
-% \begin{theorem}
-%     For any two GKAT coalgebra \(S\) and \(U\), the maximal bisimulation \(≡\) between \(\norm(S)\) and \(\norm(U)\) satisfy all the conditions in~\Cref{def:norm-bism}, i.e.
-the relation \(≡\) can be a greedy bisimulation between \(S\) and \(U\).
-% \end{theorem}
-
-% \begin{proof}
-%     We only verify that \(≡\) satisfy the conditions in~\cref{def:norm-bism}.
-
-%     The condition~\labelcref{itm:transition-bisim} holds when \(s ≡ u\).
-%     We first prove that \(s'\) and \(u'\) has to be both live or both dead: because \(δ_S(s, α) = (s', p)\) and \(δ_U(u, α) = (u', p)\),
-%     \begin{align*}
-%         s' \text{ is live} 
-%         & ⟺ δ_{\norm(S)}(s, α) = (s', p) \\
-%         & ⟺ δ_{\norm(U)}(u, α) = (u', p) \\
-%         & ⟺ u' \text{ is live}.
-%     \end{align*}
-%     \begin{itemize}
-%         \item If both \(s'\) and \(u'\) are live in \(S\) and \(U\), then \(s' ≡ u'\) because \(≡\) is a bisimulation.
-%         \item If both \(s'\) and \(u'\) are dead in \(S\) and \(U\), then they are also dead in \(\norm(S)\) and \(\norm(U)\) (\cref{thm:norm-perserve-liveness}).
-%         Finally, by~\Cref{thm:bisim-between-dead}, \(s' ≡ u'\).
-%     \end{itemize}
-
-%     TODO, finish
-% \end{proof}
-
-% \begin{lemma}\label{thm:recursive-construction-lemma}
-%     Given two states in GKAT coalgebra \(s ∈ S, u ∈ U\), consider the maximal bisimulation \({≡} ⊆ \norm(S) × \norm(U)\), \(s ≡ u\) if and only if there exists a normalized bisimulation \({≃} ⊆ S × U\) s.t.
-\(s ≃ u\).
-% \end{lemma}
-
-% \begin{proof}
-%     Recall that a relation \({∼} ⊆ \norm(S) × \norm(U)\) is a bisimulation if and only if it has a normalized GKAT coalgebra structure \(δ_∼\), s.t. for all \(s₁ ∼ u₁\):
-%     \begin{itemize}
-%         \item for all results 
-%         \(r ∈ \{\accept, \reject\}\): if \(δ_∼((s₁, u₁), α) = r\), then \(δ_{\norm(S)}(s₁, α) = r\) and \(δ_{\norm(U)}(u₁, α) = r\);
-%         \item if \((s₁, u₁) \transvia{α ∣ p}_{∼} (s₂, u₂)\) then \(s₁ \transvia{α ∣ p}_{\norm(S)} s₂\) and \(u₁ \transvia{α ∣ p}_{\norm(U)} u₂\).
-%     \end{itemize}
-
-%     We prove \(⟹\) direction by constructing the following normalized bisimulation from the bisimulation \({≡}\) between \(\norm(S)\) and \(\norm(U)\):
-%     \[{≃} ≜ {≡} ∪ \{(s, u) ∣ \text{both \(s ∈ S, u ∈ U\) are dead}\}.\]
-%     We then show that \({≃}\) satisfies all the condition in~\Cref{def:norm-bism}.
-
-%     The condition \labelcref{itm:acc-condition} holds if \(s ∼ u\):
-%     \begin{align*}
-%         δ_S(s, α) = \accept 
-%         ⟺{}& δ_{\norm(S)}(s, α) = \accept \\
-%         ⟺{}& δ_∼((s,u), α) = \accept \\ 
-%         ⟺{}& δ_{\norm(U)}(u, α) = \accept \\
-%         ⟺{}& δ_U(u, α) = \accept
-%     \end{align*}
-%     The condition \labelcref{itm:acc-condition} also holds if \(s, u\) are both dead, because both can never accept any atoms.
-
-%     The condition~\labelcref{itm:rej-or-dead} holds when \(s ≡ u\): 
-%     \begin{align*}
-%         & δ_S(s, α) = \reject \text{ and } δ_U(u, α) = (u', p)\\
-%         ⟹{}& δ_{\norm(S)}(s, α) = \reject \text{ and } δ_U(u, α) = (u', p) \\
-%         ⟹{}& δ_{\norm(U)}(u, α) = \reject \text{ and } δ_U(u, α) = (u', p)\\
-%         ⟹{}& \text{\(u'\) is dead}
-%     \end{align*}
-%     Similarly, when \(u'\) reject \(α\) and \(s\) transitions to \(s'\).
-%     The condition~\labelcref{itm:rej-or-dead} holds when \(s, u\) are both dead because both of them can only transition to dead state by~\Cref{thm:dead-iff-all-reachable-dead}.
-
-%     The condition~\labelcref{itm:transition-bisim} holds when \(s ≡ u\).
-%     We first prove that \(s'\) and \(u'\) has to be both live or both dead: because \(δ_S(s, α) = (s', p)\) and \(δ_U(u, α) = (u', p)\),
-%     \begin{align*}
-%         s' \text{ is live} 
-%         & ⟺ δ_{\norm(S)}(s, α) = (s', p) \\
-%         & ⟺ δ_{\norm(U)}(u, α) = (u', p) \\
-%         & ⟺ u' \text{ is live}.
-%     \end{align*}
-%     \begin{itemize}
-%         \item If both \(s'\) and \(u'\) are live, then \(s' ≡ u'\), and \(s' ≃ u'\).
-%         \item If both \(s'\) and \(u'\) are dead, \(s' ≃ u'\).
-%     \end{itemize}
-
-%     The condition~\labelcref{itm:transition-dead} holds, let \(δ_S(s, α) = (s', p)\) and \(δ_U(u, α) = (u', q)\) and \(p ≠ q\). 
-%     If \(s' ≃ u'\), then we can show condition~\labelcref{itm:transition-dead} by case analysis on the liveness of \(s'\) and \(u'\).
-%     If \(s', u'\) are not both dead, then the condition of \({≡} ⊆ \norm(S) × \norm(U)\) cannot be satisfied, hence \((s', u') ∉ {≡}\), which means that \(s'\) and \(u'\) are both dead, contradicting the premise.
-%     Therefore, \(s'\) and \(u'\) needs to be both dead.
-
-%     Then we show the \(⟸\) direction by
-
-%     Then we show the \(⟸\) direction by showing \(≃\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\).
-%     Take any pair of states \(s₁ ≃ u₁\), we show that all the conditions of a bisimulation is satisfied.
-
-%     First, we show for all \(α ∈ \At, δ_{\norm(S)}(s₁, α) = \accept ⟺ δ_{\norm(U)}(u₁, α) = \accept\):
-%     \begin{align*}
-%         δ_{\norm(S)}(s₁, α) = \accept 
-%         & ⟺ δ_S(s₁, α) = \accept \\  
-%         & ⟺ δ_U(u₁, α) = \accept \\  
-%         & ⟺ δ_{\norm(U)})(u₁, α) = \accept
-%     \end{align*}
-
-%     Then, we show for all \(α ∈ \At, δ_{\norm(S)}(s₁, α) = \reject ⟺ δ_{\norm(U)}(u₁, α) = \reject\).
-%     Without loss of generality, we only show the \(⟹\) direction.
-%     If \(δ_{\norm(S)}(s₁, α) = \reject\), then either \(δ_S(s₁, α) = \reject\), or \(δ_S(s₁, α) = (s₂, p)\) and \(s₂\) is dead.
-%     \begin{itemize}
-%         \item We consider the case where \(δ_S(s₁, α) = \reject\).
-%         By condition~\labelcref{itm:acc-condition}, \(δ_U(u₁, α)\) cannot equal to accept. 
-%         If \(δ_U(u₁, α) = \reject\), then the implication holds. 
-%         If \(δ_U(u₁, α) = (u₂, q)\), then by condition~\cref{itm:rej-or-dead}, \(u₂\) is dead, which also makes the implication hold.
-%         \item We consider the case where \(δ_S(s₁, α) = (s₂, p)\) and \(s₂\) is dead in \(S\).
-%         By condition~\labelcref{itm:acc-condition}, \(δ_U(u₁, α)\) cannot equal to accept.
-%         If \(δ_U(u₁, α) = \reject\), then the implication holds.
-%         If \(δ_U(u₁, α) = (u₂, q)\) and \(p ≠ q\), then by condition~\labelcref{itm:transition-dead}, both \(u₁, u₂\) are dead, then by~\cref{thm:bisim-between-dead}, \(s₂ ≡ s\). 
-%         If \(δ_U(u₁, α) = (u₂, q)\) and \(p = q\), then by condition~\labelcref{itm:transition-bisim}, then \(s₂ ≃ u₂\)
-%     \end{itemize}
-
-%     To obtain the desired bisimulation \({∼}\) between \(\norm(S)\) and \(\norm(U)\), we add the pair \((s, t)\) to \(∼'\), 
-%     \[{∼} ≜ {∼'} ∪ \{(s, u)\},\] 
-%     with the following transition \(δ_{∼}\): for all \(α ∈ \At\),
-%     \begin{itemize}[nosep]
-%         \item if \(δ_{\norm(S)}(s, α) = δ_{\norm(U)}(u, α) = \accept\), then \(δ_{∼}((s, u), α) ≜ \accept\);
-%         \item if \(δ_{\norm(S)}(s, α) = δ_{\norm(U)}(u, α) = \reject\), then \(δ_{∼}((s, u), α) ≜ \reject\);
-%         \item if \(δ_{\norm(S)}(s, α) = (s', p)\) and \(δ_{\norm(U)}(u, α) = (u', p)\), then \(δ_{∼}((s, u), α) = ((s', u'), p)\);
-%         \item for all \((s', u') ∈ {∼}'\) that is not equal to \((s, u)\), we let \(δ_∼\) inherits the transition of \(δ_{∼'}\), i.e. \(δ_{∼}((s', u'), α) = δ_{∼'}((s', u'), α)\)
-%     \end{itemize}
-%     \emph{if \(δ_{∼}\) is well-defined}, then we can verify that \(∼\) is indeed a bisimulation between \(\norm(⟨s⟩)\) and \(\norm(⟨u⟩)\) where \(s ∼ u\). 
-%     We show the slightly more complicated case: \(δ_{\norm(S)}(s, α) = (s', p)\) and \(δ_{\norm(U)}(u, α) = (u', p)\) implies \(s' ∼ u'\) as an example. 
-%     By condition~\labelcref{itm:transition-bisim}, there exists a bisimulation \(∼_{s', u'} ⊆ \norm(⟨s'⟩) × \norm(⟨u'⟩)\), and because \(≡_{s', u'}\) is the maximal bisimulation between \(\norm(⟨s'⟩)\) and \(\norm(⟨u'⟩)\),
-%     \[(s', u') ∈ {∼_{s', u'}} ⊆ {≡_{s', u'}} ⊆ {∼}.\]
-
-%     Finally, we demonstrate that \(δ_∼\) is well-defined by leveraging the conditions in~\Cref{thm:recursive-construction}. 
-%     Specifically, we will show that the definition of \(δ_∼\) covers all the possible cases, by case analysis on the result of \(δ_S\): for all \(α ∈ \At\),
-%     \begin{itemize}
-%         \item If \(δ_S(s, α) = \accept\), then by condition~\labelcref{itm:acc-condition}, \(δ_U(u, α) = \accept\); therefore \[δ_{\norm(S)}(s, α) = δ_{\norm(U)}(u, α) = \accept.\]
-%         \item If \(δ_S(s, α)\) transitions to a dead state or reject, then by condition~\labelcref{itm:rej-or-dead} \(δ_U(u, α)\) will also transition to a dead state or reject, then \[δ_{\norm(S)}(s, α) = δ_{\norm(U)}(u, α) = \reject.\]
-%         \item If \(δ_S(s, α) = (s', p)\) and \(s'\) is live, then \(δ_U(u, α) = (u', p)\) necessarily holds, otherwise it would violate one of conditions~\labelcref{itm:acc-condition,itm:rej-or-dead,itm:transition-dead}.
-
-%         By condition~\labelcref{itm:transition-bisim}, there exists a bisimulation \(∼_{s', u'}\) between \(\norm(⟨s'⟩)\) and \(\norm(⟨u'⟩)\) s.t.
-\(s' ∼_{s', u'} u'\).
-Because bisimulation preserves liveness (\Cref{thm:bisim-preserve-liveness}), \(s', u'\) has to be both dead or live.
-%         Finally, because \(s'\) is live, therefore \(u'\) is also live, and we obtain the final case in the definition of \(δ_{∼}\): \(δ_{\norm(S)}(s, α) = (s', p)\) and \(δ_{\norm(U)}(u, α) = (u', p)\).
-%     \end{itemize}
-%     Because \(≃ ⊆ \norm(S) × \norm(U)\) is a bisimulation s.t. \(s ≃ u\), and \(≡\) is the maximal bisimulation between \(\norm(S)\) and \(\norm(U)\), we will obtain our result \(s ≡ u\).
-% \end{proof}
-
-% % \begin{proof}
-% %     Recall that there exists a bisimulation \({∼} ⊆ \norm(⟨s⟩) × \norm(⟨u⟩)\) if and only if for all \(s₁ ∼ u₁\):
-% %     \begin{itemize}
-% %         \item for all results \(r ∈ \{\accept, \reject\}\): \(δ_{\norm(S)}(s₁, α) = r ⟺ δ_{\norm(U)}(u₂, α) = r\);
-% %         \item otherwise, let \((s₂, p) ≜ δ_{\norm(S)}(s₁, α)\) and \((u₂, q) ≜ δ_{\norm(u)}(u₁, α)\), then \(p = q\) and \(s₂ ∼ u₂\)
-% %     \end{itemize}
-
-% %     We prove \(⟹\) direction by constructing the following normalized bisimulation from the bisimulation \({∼}\):
-% %     \[{≃} ≜ {∼} ∪ \{(s, u) ∣ \text{both \(s ∈ S, u ∈ U\) are dead}\}.\]
-% %     We then show that \({≃}\) satisfies all the condition in~\Cref{def:norm-bism}.
-
-% %     The condition \labelcref{itm:acc-condition} holds if \(s ∼ u\):
-% %     \begin{align*}
-% %         δ_S(s, α) = \accept 
-% %         ⟺{}& δ_{\norm(S)}(s, α) = \accept \\
-% %         ⟺{}& δ_∼((s,u), α) = \accept \\ 
-% %         ⟺{}& δ_{\norm(U)}(u, α) = \accept \\
-% %         ⟺{}& δ_U(u, α) = \accept
-% %     \end{align*}
-% %     The condition \labelcref{itm:acc-condition} also holds if \(s, u\) are both dead, because both can never accept any atoms.
-
-% %     The condition~\labelcref{itm:rej-or-dead} holds when \(s ∼ u\): 
-% %     \begin{align*}
-% %         & δ_S(s, α) = \reject \text{ and } δ_U(u, α) = (u', p)\\
-% %         ⟹{}& δ_{\norm(S)}(s, α) = \reject \text{ and } δ_U(u, α) = (u', p) \\
-% %         ⟹{}& δ_{\norm(U)}(u, α) = \reject \text{ and } δ_U(u, α) = (u', p)\\
-% %         ⟹{}& \text{\(u'\) is dead}
-% %     \end{align*}
-% %     Similarly, when \(u'\) reject \(α\) and \(s\) transitions to \(s'\).
-% %     The condition~\labelcref{itm:rej-or-dead} holds when \(s, u\) are both dead because both of them can only transition to dead state by~\Cref{thm:dead-iff-all-reachable-dead}.
-
-% %     The condition~\labelcref{itm:transition-bisim} holds when \(s ∼ u\).
-% %     We first prove that \(s'\) and \(u'\) has to be both live or both dead: because \(δ_S(s, α) = (s', p)\) and \(δ_U(u, α) = (u', p)\),
-% %     \begin{align*}
-% %         s' \text{ is live} 
-% %         & ⟺ δ_{\norm(S)}(s, α) = (s', p) \\
-% %         & ⟺ δ_{\norm(U)}(u, α) = (u', p) \\
-% %         & ⟺ u' \text{ is live}.
-% %     \end{align*}
-% %     \begin{itemize}
-% %         \item If both \(s'\) and \(u'\) are live, then \(s' ∼ u'\). By~\cref{thm:sub-coalg-preserve-bisim}, the bisimulation \(∼'\) is just \(∼\) restricted to \(⟨s'⟩\) and \(⟨u'⟩\).
-% %         \item If both \(s'\) and \(u'\) are dead, then \(∼'\) can just be the singleton relation, according to~\cref{thm:bisim-between-dead}.
-% %     \end{itemize}
-
-% %     The condition~\labelcref{itm:transition-dead} holds: by the proof of condition~\labelcref{itm:transition-bisim}, \(s'\) and \(u'\) has to be either both live or both dead; if they are both live, then there cannot be a element in \(G(∼)\) that can project to \((s', p)\) under \(π₁\) but projects to \((t', q)\) under \(π₂\).
-Thus both \(s'\) and \(t'\) has to be dead.
-
-% %     We then show the \(⟸\) direction, for arbitrary \(s' ∈ S\) and \(u' ∈ U\), we use \(≡_{s', u'}\) to denote the maximal bisimulation between \(\norm(⟨s'⟩)\) and \(\norm(⟨u'⟩)\).
-% %     \begin{align*}
-% %         {∼'} ≜ ⋃ \{≡_{s', u'} & ∣ ∃ α ∈ \At, p ∈ K, \\*
-% %             & δ_{\norm(S)}(s, α) = (s', p) \\*
-% %             & \text{ and } δ_{\norm(U)}(u, α) = (u', p)\}.
-% %     \end{align*}
-% %     For all the \(s'\) and \(u'\) in the above definition, \(⟨s'⟩ ⊑ ⟨s⟩\) and \(⟨u'⟩ ⊑ ⟨u⟩\), therefore by~\Cref{thm:norm-sub-coalg}, \(\norm(⟨s'⟩) ⊑ \norm(⟨s⟩)\) and \(\norm(⟨u'⟩) ⊑ \norm(⟨u⟩)\).
-% %     By~\Cref{thm:sub-coalg-preserve-bisim}, every \(≡_{s', u'}\) is a bisimulation between \(\norm(⟨s⟩)\) and \(\norm(⟨t⟩)\), and because bisimulation is closed under arbitrary union~\cite{rutten_UniversalCoalgebraTheory_2000}, \(∼'\) is a bisimulation between \(\norm(⟨s⟩)\) and \(\norm(⟨t⟩)\).
-
-% %     To obtain the desired bisimulation \({∼}\) between \(\norm(⟨s⟩)\) and \(\norm(⟨u⟩)\), we add the pair \((s, t)\) to \(∼'\), 
-% %     \[{∼} ≜ {∼'} ∪ \{(s, u)\},\] 
-% %     with the following transition \(δ_{∼}\): for all \(α ∈ \At\),
-% %     \begin{itemize}[nosep]
-% %         \item if \(δ_{\norm(S)}(s, α) = δ_{\norm(U)}(u, α) = \accept\), then \(δ_{∼}((s, u), α) ≜ \accept\);
-% %         \item if \(δ_{\norm(S)}(s, α) = δ_{\norm(U)}(u, α) = \reject\), then \(δ_{∼}((s, u), α) ≜ \reject\);
-% %         \item if \(δ_{\norm(S)}(s, α) = (s', p)\) and \(δ_{\norm(U)}(u, α) = (u', p)\), then \(δ_{∼}((s, u), α) = ((s', u'), p)\);
-% %         \item for all \((s', u') ∈ {∼}'\) that is not equal to \((s, u)\), we let \(δ_∼\) inherits the transition of \(δ_{∼'}\), i.e. \(δ_{∼}((s', u'), α) = δ_{∼'}((s', u'), α)\)
-% %     \end{itemize}
-% %     \emph{if \(δ_{∼}\) is well-defined}, then we can verify that \(∼\) is indeed a bisimulation between \(\norm(⟨s⟩)\) and \(\norm(⟨u⟩)\) where \(s ∼ u\). 
-% %     We show the slightly more complicated case: \(δ_{\norm(S)}(s, α) = (s', p)\) and \(δ_{\norm(U)}(u, α) = (u', p)\) implies \(s' ∼ u'\) as an example. 
-% %     By condition~\labelcref{itm:transition-bisim}, there exists a bisimulation \(∼_{s', u'} ⊆ \norm(⟨s'⟩) × \norm(⟨u'⟩)\), and because \(≡_{s', u'}\) is the maximal bisimulation between \(\norm(⟨s'⟩)\) and \(\norm(⟨u'⟩)\),
-% %     \[(s', u') ∈ {∼_{s', u'}} ⊆ {≡_{s', u'}} ⊆ {∼}.\]
-
-% %     Finally, we demonstrate that \(δ_∼\) is well-defined by leveraging the conditions in~\Cref{thm:recursive-construction}. 
-% %     Specifically, we will show that the definition of \(δ_∼\) covers all the possible cases, by case analysis on the result of \(δ_S\): for all \(α ∈ \At\),
-% %     \begin{itemize}
-% %         \item If \(δ_S(s, α) = \accept\), then by condition~\labelcref{itm:acc-condition}, \(δ_U(u, α) = \accept\); therefore \[δ_{\norm(S)}(s, α) = δ_{\norm(U)}(u, α) = \accept.\]
-% %         \item If \(δ_S(s, α)\) transitions to a dead state or reject, then by condition~\labelcref{itm:rej-or-dead} \(δ_U(u, α)\) will also transition to a dead state or reject, then \[δ_{\norm(S)}(s, α) = δ_{\norm(U)}(u, α) = \reject.\]
-% %         \item If \(δ_S(s, α) = (s', p)\) and \(s'\) is live, then \(δ_U(u, α) = (u', p)\) necessarily holds, otherwise it would violate one of conditions~\labelcref{itm:acc-condition,itm:rej-or-dead,itm:transition-dead}.
-
-% %         By condition~\labelcref{itm:transition-bisim}, there exists a bisimulation \(∼_{s', u'}\) between \(\norm(⟨s'⟩)\) and \(\norm(⟨u'⟩)\) s.t.
-\(s' ∼_{s', u'} u'\).
-Because bisimulation preserves liveness (\Cref{thm:bisim-preserve-liveness}), \(s', u'\) has to be both dead or live.
-% %         Finally, because \(s'\) is live, therefore \(u'\) is also live, and we obtain the final case in the definition of \(δ_{∼}\): \(δ_{\norm(S)}(s, α) = (s', p)\) and \(δ_{\norm(U)}(u, α) = (u', p)\).
-% %         \qedhere
-% %     \end{itemize}
-% % \end{proof}
-
-% With some slight tweak to \Cref{thm:recursive-construction-lemma}, it will be applicable to trace equivalence instead of bisimulation.
-% The correctness of this modification is supported by the correspondence between trace equivalence and the existence of a bisimulation between normalized GKAT coalgebras, which is stated in~\Cref{thm:norm-bisim-correctness}.
-
-% \begin{theorem}[Recursive Construction]\label{thm:recursive-construction}
-%     For any two states in two GKAT coalgebra \(s ∈ S, u ∈ U\), \(s\) and \(u\) are finite-trace equivalent \(⟦s⟧_{S} = ⟦u⟧_{U}\) if and only if all the following conditions hold:
-%     \begin{enumerate}
-%         \item for all \(α ∈ \At\), \(δ_{S}(s, α) = \accept ⟺ δ_{U}(u, α) = \accept\);
-%         \item \(s\) reject \(α\) or transition to a dead state via \(α\) if and only if \(u\) rejects \(α\) or transition to a dead state via \(α\);  
-%         \item if \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', p)\), then \(⟦s'⟧_{S} = ⟦u'⟧_{U}\);
-%         \item if \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', q)\), s.t. \(p ≠ q\), then both \(s'\) and \(u'\) are dead.
-%     \end{enumerate}
-%     Notice that above condition is similar to those in~\cref{thm:recursive-construction}, except bisimulation of \(s'\) and \(u'\) is replaced with trace equivalence.
-% \end{theorem}
-
-% \begin{proof}
-%     By the standard argument with normalization preserves sub-coalgebra (\Cref{thm:norm-sub-coalg}), we can obtain for all \(s ∈ S\) and \(u ∈ U\), \(\norm(⟨s⟩) ⊑ \norm(S)\) and \(\norm(⟨u⟩) ⊑ \norm(U)\).
-%     Therefore, because subcoalgebra preserve and reflect bisimulation (\Cref{thm:sub-coalg-preserve-bisim}) and the correctness of bisimulation on normalized coalgebra (\Cref{thm:norm-bisim-correctness}): for all \(s ∈ S\) and \(u ∈ U\),
-%     \begin{align*}
-%         & ∃~ {∼} ⊆ \norm(⟨s⟩) × \norm(⟨u⟩), s ∼ u \\
-%         ⟺{}& ∃~ {∼} ⊆ \norm(S) × \norm(U), s ∼ u \\  
-%         ⟺{}& ⟦s⟧_S = ⟦u⟧_U.
-%     \end{align*}
-%     We can instantiate the above equivalence to \(s', u'\) and \(s, u\) respectively: the instantiation to \(s', u'\) will give us the conditions in this theorem is equivalent to the ones in~\Cref{thm:recursive-construction-lemma}; and instantiation to \(s, u\) show that existence of bisimulation is equivalent to trace equivalence:
-%     \begin{align*}
-%         & \text{Conditions in \Cref{thm:recursive-construction-lemma} hold} \\  
-%         ⟺{}& \text{Conditions in the current theorem hold}\\
-%         ⟺{}& ∃~ {∼} ⊆ \norm(⟨s⟩) × \norm(⟨u⟩), s ∼ u\\
-%         ⟺{}& ⟦s⟧_{S} = ⟦u⟧_{U}. \qedhere
-%     \end{align*}
-% \end{proof}
-
-% The above construction theorem already gives us an algorithm to recursively decide whether \(⟦s⟧_{S} = ⟦u⟧_U\), when \(⟨s⟩_S\) and \(⟨u⟩_U\) is finite.
-% However, this algorithm can be further optimized: we will derive that a dead state is only trace equivalent to other dead states. 
-% This means that when checking the trace-equivalence of states \(s\) and \(u\), if we already know one of them is dead, we only need to check whether the other is dead, instead of going through all the conditions in~\Cref{thm:recursive-construction}.
-
-% \begin{theorem}\label{thm:bisim-one-dead}
-%     Given two states \(s ∈ S\) and \(u ∈ U\), if \(s\) is a dead state in \(S\), then \(s\) and \(u\) is trace equivalent if and only if \(u\) is also dead.
-% \end{theorem}
-
-% \begin{proof}
-%     TODO: I think we can refine \(⟹\) direction, but I am not sure...
-
-%     For the \(⟸\) direction, we can construct the bisimulation as in~\cref{thm:bisim-between-dead}, which implies trace equivalence.
-%     And the \(⟹\) direction, if \(s\) is dead, then by definition of normalization, it will be all rejecting, i.e. for all \(α ∈ \At\), \(δ_{\norm(S)}(s, α) = \reject\).
-%     If \(s\) and \(u\) are trace equivalent, then there exists a bisimulation \({∼}: \norm(S) × \norm(U)\) s.t. \(s ∼ u\).
-%     By unfolding the definition of a bisimulation, \(u\) also need to be all rejecting in \(\norm(U)\).
-%     By definition of \(\norm\), for all \(α ∈ \At\), \(δ_U(u, α)\) either reject to go to a dead state, and because \(⟨u⟩_U\) is all the reachable state from \(u\), therefore 
-%     \begin{align*}
-%         ⟨u⟩_U = & ⋃ \{⟨u'⟩ ∣ ∃ α ∈ \At, p ∈ K, δ_U(u, α) = (u', p)\} \\
-%         & ∪ \{u\}
-%     \end{align*}
-%     And because all of \(u'\) is dead, all of \(⟨u'⟩_U\) cannot be accepting (\Cref{thm:dead-iff-all-reachable-dead}), nor is \(u\) an accepting state, then \(⟨u⟩_U\) does not contain any accepting state, and by definition of dead state, \(u\) is dead.
-% \end{proof}
 
@@
     \Cref{itm:remove-dead-pairs-sound} can be proven by realizing that \((s, u)\) is a dead state in \(∼\) (\Cref{thm:bisim-preserve-liveness}) which means that all the states in \(⟨(s, u)⟩_∼\) are dead.
-    Because \(∼\) is a normalized GKAT coalgebra (\Cref{thm:}), therefore no state can transition into dead states like \(⟨(s, u)⟩_∼\), hence \({∼} ∖ ⟨(s, u)⟩_∼\) is still closed under its transition, and form a GKAT coalgebra.
-    And the fact that \({∼} ∖ ⟨(s, u)⟩_∼\) is also a bisimulation can be proven by factoring through \(∼\) via the inclusion homomorphism:
-    \begin{align*}
-        ({∼} ∖ ⟨(s, u)⟩_∼) \xrightarrow{i} {∼} \xrightarrow{π₁} \norm(S) \\
-        ({∼} ∖ ⟨(s, u)⟩_∼) \xrightarrow{i} {∼} \xrightarrow{π₂} \norm(U)
-    \end{align*}
-    TODO: extract the above to a lemma.
+    Because \(∼\) is a normalized GKAT coalgebra (\Cref{thm:}), therefore no state can transition into dead states like \(⟨(s, u)⟩_∼\), hence \({∼} ∖ ⟨(s, u)⟩_∼\) is still closed under its transition, and form a sub-coalgebra of \(∼\).
+    Finally because \(∼\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\), then so is \({∼} ∖ ⟨(s, u)⟩_∼\) (\Cref{thm:})
 \end{proof}

### 2025-01-17 main.tex
abstract

@@
 \begin{abstract}
-This document is a model and instructions for \LaTeX.
-This and the IEEEtran.cls file define the components of your paper [title, text, heads, etc.]. *CRITICAL: Do Not Use Symbols, Special Characters, Footnotes, 
-or Math in Paper Title or Abstract.
+We present several new algorithms to efficiently decide Guarded Kleene Algebra with Tests (GKAT) equivalences.
+Although the current algorithm scales nearly-linearly with the size of the expression, its performance is hindered by the need to compute the entire coalgebra for the sake of normalization and exponential performance degradation when expanding the set of primitive tests.
+To address these problems, we introduce two new algorithms for bisimulation checking: an on-the-fly bisimulation algorithm based on greedy bisimulation, which enables immediate termination upon discovering a counter-example and supports on-the-fly coalgebra generation with derivatives; and a symbolic algorithm that builds upon the on-the-fly algorithm while leveraging efficient SAT solvers to speed up boolean comparisons and compress the transition structures of GKAT coalgebra.
+To utilize the symbolic bisimulation checking, we also provide two methods to generate symbolic GKAT coalgebras from GKAT expressions, analogous to derivative and Thompson's construction in previous works.
+We have proven the correctness of our algorithms and conducted detailed complexity analyses.
+Finally, we implemented the symbolic algorithms in Rust, demonstrating significant performance improvements over state-of-the-art tools.
 \end{abstract}
@@
 \begin{IEEEkeywords}
-component, formatting, style, styling, insert.
+Guarded Kleene Algebra With Tests, Kleene Algebra, Coalgebra, Program Verification, Algorithm
 \end{IEEEkeywords}

### 2025-01-17 main.tex
minor tweaks

@@
 
-\begin{corollary}\label{thm:norm-bisim-exists-unique}
+\begin{corollary}\label{thm:greedy-bisim-exists-unique}
     For any bisimulation between two normalized GKAT coalgebra \({∼} ⊆ \norm(S) × \norm(U)\), a greedy bisimulation \(≈\) with \(∼\) as underlying bisimulation exists and is unique.
@@
     Because \({≈} ⊑ {∼_d}\) and~\Cref{thm:norm-sub-coalg}, \(\norm(≈) ⊑ \norm(∼_d)\). 
-    Because \(\norm(∼_d)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\), therefore \(\norm(≈)\) is a bisimulation by~\cref{thm:}
+    Because \(\norm(∼_d)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\), therefore \(\norm(≈)\) is a bisimulation (\Cref{thm:greedy-bisim-iff-norm-bisim}).
 
@@
 
-We then establish the correctness of greedy bisimulation, i.e. it is equivalent to trace equivalence.
-
-\begin{theorem}[Correctness]
-    Given two states in GKAT coalgebras \(s ∈ S\) and \(u ∈ ,U\), there exists a greedy bisimulation \({≈} ⊆ S × U\) s.t. \(s ≈ u\), if and only if \(⟦s⟧_S = ⟦u⟧_U\).
-\end{theorem}
-
-\begin{proof}
-    \(⟹\) direction, because normalization preserves carrier set, therefore \((s, u) ∈ \norm(≈)\).
-    And because \(\norm(≈)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\) (\Cref{thm:greedy-bisim-iff-norm-bisim}), \(⟦s⟧_S = ⟦u⟧_U\).
-
-    \(⟸\) direction, because \(⟦s⟧_S = ⟦u⟧_U\), then by~\cref{thm:norm-bisim-correctness}, then there a bisimulation \(∼ ⊆ \norm(S) × \norm(U)\) s.t. \(s ∼ u\). 
-    Because~\Cref{thm:norm-bisim-exists-unique}, therefore there exists a greedy bisimulation \(≈\) with \(∼\) as its underlying bisimulation. 
-    Because the carriers \({∼} ⊆ {≈}\), therefore \(s ∼ u\) implies \(s ≈ u\).
-\end{proof}
+% Similar to the correctness of bisimulation between normalized coalgebra (\Cref{thm:norm-bisim-correctness}), we also establish a correctness theorem for greedy bisimulation, i.e. there exists
+
+% \begin{theorem}[Correctness]
+%     Given two states in GKAT coalgebras \(s ∈ S\) and \(u ∈ ,U\), there exists a greedy bisimulation \({≈} ⊆ S × U\) s.t. \(s ≈ u\), if and only if \(⟦s⟧_S = ⟦u⟧_U\).
+% \end{theorem}
+
+% \begin{proof}
+%     \(⟹\) direction, because normalization preserves carrier set, therefore \((s, u) ∈ \norm(≈)\).
+%     And because \(\norm(≈)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\) (\Cref{thm:greedy-bisim-iff-norm-bisim}), \(⟦s⟧_S = ⟦u⟧_U\).
+
+%     \(⟸\) direction, because \(⟦s⟧_S = ⟦u⟧_U\), then by~\cref{thm:norm-bisim-correctness}, then there a bisimulation \(∼ ⊆ \norm(S) × \norm(U)\) s.t. \(s ∼ u\). 
+%     Because~\Cref{thm:greedy-bisim-exists-unique}, therefore there exists a greedy bisimulation \(≈\) with \(∼\) as its underlying bisimulation. 
+%     Because the carriers \({∼} ⊆ {≈}\), therefore \(s ∼ u\) implies \(s ≈ u\).
+% \end{proof}
 
 Despite proving many desirable structural properties, the definition of greedy bisimulation is still coalgebraic.
-To design an equivalence-checking algorithm to construct a greedy bisimulation still requires turning the coalgebra into a set of concrete criteria, that can be checked by an algorithm.
-In the following theorem, we will present four conditions, and the algorithm will be largely based on recursively checking these conditions to establish a greedy bisimulation.
+In order for our equivalence-checking algorithm to construct a greedy bisimulation, we need to turn the coalgebra into a set of concrete criteria, that can be checked by an algorithm.
+In the following theorem, we will present such necessary and sufficient conditions to construct a greedy bisimulation.
+Our algorithm will be largely based on recursively checking these conditions to establish a greedy bisimulation.
 

### 2025-01-17 main.tex
make on-the-fly algorithm more similar to the symbolic one

@@
         \EndIf
-        \For{\(α ∈ \At\)}{}{
-            \Match{$δ_{S}(s, α), δ_{U}(u, α)$}
-            \Case{\(\accept, \accept\)} {\Continue} \EndCase
-            \Case{\(\reject, \reject\)} {\Continue} \EndCase
-            \Case{\(\reject, (u', q)\)} 
-                \State\Return \Call{isDead\(_U\)}{$u'$} 
-            \EndCase
-            \Case{\((s', p), \reject\)}
-                \State\Return\Call{isDead\(_S\)}{$s'$}
-            \EndCase
-            \Case{\((s', p), (u', q)\)} {
-                \If {\(p = q\)}
-                    \State\Call{union}{$s, u$}
-                    \State\Return \Call{equiv}{$s, t$} 
-                \EndIf
-                \If{\Call{isDead\(_S\)}{$s$} and \Call{isDead\(_U\)}{$u$}}
-                    \State\Continue
-                \EndIf
-                \State\Return false
-            } \EndCase
-            \Default { \Return false } \EndDefault
-            \EndMatch
-        }\EndFor
-        \State\Return true
+        \State\Return \(∀ α ∈ \At\) \Comment{conditions in~\Cref{thm:recursive-construction}}
+        \\\vspace{5px}
+        \(\qquad\begin{aligned}
+            & s \transAcc{S}{α} \text{ iff } u \transAcc{U}{α} \mathrel{\&\!\&}\\  
+            & s \transRej{S}{α} \text{ and } u \transvia{α ∣ p}_U u' \text{ implies } \Call{isDead\(_U\)}{u'} \mathrel{\&\!\&}\\  
+            & s \transvia{α ∣ p}_S s' \text{ and } u \transRej{U}{α} \text{ implies } \Call{isDead\(_U\)}{s'} \mathrel{\&\!\&}\\  
+            & s \transvia{α ∣ p}_S s' \text{ and } u \transvia{α ∣ p}_U u' \text{ implies } \Call{union}{s, u}; \Call{equiv}{s', u'} \mathrel{\&\!\&}\\
+            & s \transvia{α ∣ p}_S s' \text{ and } u \transvia{α ∣ q}_U u' \text{ and } p ≠ q \text{ implies } \Call{isDead\(_S\)}{s'} \text{ and } \Call{isDead\(_U\)}{u'}
+        \end{aligned}\)
+        % \Forall{\(α ∈ \At\)}{}{
+        %     \Match{$δ_{S}(s, α), δ_{U}(u, α)$}
+        %     \Case{\(\accept, \accept\)} {\Continue} \EndCase
+        %     \Case{\(\reject, \reject\)} {\Continue} \EndCase
+        %     \Case{\(\reject, (u', q)\)} 
+        %         \State\Return \Call{isDead\(_U\)}{$u'$} 
+        %     \EndCase
+        %     \Case{\((s', p), \reject\)}
+        %         \State\Return\Call{isDead\(_S\)}{$s'$}
+        %     \EndCase
+        %     \Case{\((s', p), (u', q)\)} {
+        %         \If {\(p = q\)}
+        %             \State\Call{union}{$s, u$}
+        %             \State\Return \Call{equiv}{$s, t$} 
+        %         \EndIf
+        %         \If{\Call{isDead\(_S\)}{$s$} and \Call{isDead\(_U\)}{$u$}}
+        %             \State\Continue
+        %         \EndIf
+        %         \State\Return false
+        %     } \EndCase
+        %     \Default { \Return false } \EndDefault
+        %     \EndMatch
+        % }\EndFor
+        % \State\Return true
         \EndFunction
     \end{algorithmic}
-\end{algorithm}
+\end{algorithm*}
 
@@
         \Return {
-            \Comment{conditions of ~\cref{thm:symb-recursive-construction}}
+            \Comment{conditions of ~\Cref{thm:symb-recursive-construction}}
             \\\vspace{5px}
@@
                 & ⋁ ϵ̂_S(s) ≡ ⋁ ϵ̂_U(u) \mathrel{\&\!\&} \\  
-                & ∀ (b, s', p) ∈ δ̂_S(s), (c, u', q) ∈ δ̂_U(u), (b ∧ c) ≢ 0 ⟹ 
-                \begin{cases}
-                    \text{\Call{isDead$_S$}{$s$}} ∧ \text{\Call{isDead$_U$}{$u$}} & \text{if \(p ≠ q\)} \\
-                    \text{\Call{Union}{$s$, $u$}}; \text{\Call{Equiv}{$s', u'$}} & \text{if \(p = q\)}
-                \end{cases} \mathrel{\&\!\&}\\
-                & ∀ (b, s', p) ∈ δ̂_S(s), c ∈ ρ̂_U(u), (b ∧ c) ≢ 0 ⟹ \text{\Call{isDead$_S$}{$s'$}} \mathrel{\&\!\&}\\
-                & ∀ b ∈ ρ̂_S(s), (c, u', q) ∈ δ̂_U(u), (b ∧ c) ≢ 0 ⟹ \text{\Call{isDead$_U$}{$u'$}}
+                & s \transvia{b ∣ p} s' \text{ and } (b ∧ ρ̂_U(u)) ≢ 0 \text{ implies \Call{isDead$_S$}{$s'$}} \mathrel{\&\!\&}\\
+                & u \transvia{b ∣ p} u' \text{ and } (ρ̂_S(s) ∧ b) ≢ 0 \text{ implies \Call{isDead$_U$}{$u'$}} \mathrel{\&\!\&}\\
+                & s \transvia{b ∣ p} s' \text{ and } u \transvia{c ∣ p} u' \text{ implies } \text{\Call{Union}{$s$, $u$}}; \text{\Call{Equiv}{$s', u'$}} \mathrel{\&\!\&}\\
+                & s \transvia{b ∣ p} s' \text{ and } u \transvia{c ∣ q} u' \text{ and } p ≠ q \text{ implies \Call{isDead}{$s'$} and \Call{isDead}{$u'$}}
             \end{aligned}\)

### 2025-01-17 main.tex
changes to is_dead code

@@
 ]
+pub type Deriv = Vec<(BExp, Exp, u64)>;
+
 pub struct Solver {
     // search states
-    pub(super) dead_states: HashSet<Exp>,
-    pub(super) explored: HashSet<Exp>,
-    pub(super) uf_table: HashMap<Exp, UnionFindNode<()>>,
+    dead_states: HashSet<Exp>,
+    explored: HashSet<Exp>,
+    uf_table: HashMap<Exp, UnionFindNode<()>>,
     // caching
-    pub(super) eps_cache: LruCache<Exp, BExp>,
-    pub(super) deriv_cache: LruCache<Exp, Vec<(BExp, Exp, u64)>>,
+    eps_cache: LruCache<Exp, BExp>,
+    deriv_cache: LruCache<Exp, Deriv>,
 }
@@
 impl Solver {
+    #[inline]
+    pub fn known_dead(&self, exp: &Exp) -> bool {
+        self.dead_states.contains(&exp)
+    }
+
     pub fn is_dead(&mut self, gkat: &mut Gkat, exp: &Exp) -> bool {
@@
         while let Some(exp) = stack.pop() {
-            if self.dead_states.contains(&exp) || self.explored.contains(&exp) {
+            if self.known_dead(&exp) || self.explored.contains(&exp) {
                 continue;
@@
             if eps.is_false() {
-                for (_, e, _) in self.derivative(gkat, &exp) {
+                for (b, e, _) in self.derivative(gkat, &exp) {
+                    // check is not strictly needed due to eager-pruning
+                    if b.is_false() {
+                        continue;
+                    }
                     stack.push(e);

### 2025-01-17 main.tex
change implementation description to match algorithm

@@
 
-Notice that during bisimulation~(\Cref{alg:symb-bisim}) for some \(s \in S\) and \(u \in U\), if there is \((b,s',p) \in \hat{\delta}_S(s)\) where \(b\) is a semantically false boolean expression, then implications of form \((b \land \cdot) \not{\equiv} 0 \Rightarrow \cdot\) are trivially satisfied.
+Notice that during bisimulation~(\Cref{alg:symb-bisim}) for some \(s \in S\) and \(u \in U\), if there is \(s \transvia{b ∣ p} s'\), i.e.
+\((b,s',p) \in \hat{\delta}_S(s)\), where \(b\) is a semantically false boolean expression, then implications with premises of the form \((b \land \cdot) \not{\equiv} 0\) are trivially satisfied.
 In other words, \textit{blocked} transitions in \(S\) do not contribute to the result of bisimulation.

### 2025-01-18 main.tex
updated experimental results

@@
 Each benchmark consists of 50 expression pairs with properties described by the benchmark's name.
-For example, expressions in E250B5P10RD have around 250 primitive actions in total (E250), a maximum boolean expression size of 5 (B5), 10 possible unique primitive actions (P10) and are completely random (RD).
+For example, expressions in E250B5P10RD have around 250 primitive actions in total (E250), a maximum boolean expression size of 5 (B5), 10 possible unique primitive tests (P10) and are completely random (RD).
 Benchmarks with the suffix EQ have expression pairs that are known to be equivalent.
-The time columns measure the total time used (in seconds) for checking all expression pairs in the benchmarks.
-The mem columns measure the peak memory used (in MB).
-
-It is clear from these results that our algorithms (DRV and TC) are more efficient than SymKAT in terms of both time and memory usage.
-Moreover, our Thompson's construction (TC) based algorithm is the only one which successfully completed all benchmarks without timeout.
+The time columns measure the \textit{total time} used (in seconds) for checking all expression pairs in the benchmarks.
+The mem columns measure the \textit{peak memory} used (in MB).
+
+It is clear from these results that our algorithms (DRV and TC) can efficiently check even large programs in just a few seconds.
 
@@
 \centering
-\begin{tabular}{l l l l l l l}
+\begin{tabular}{l r r r r r r}
     Benchmark & Time (DRV) & Time (TC) & Time (SK) & Mem (DRV) & Mem (TC) & Mem (SK) \\
     \hline
-    E250B5P10RD    & 0.26    & 0.25  & 16.99   & 15.36  & 14.76  & 329.48  \\
-    E250B5P10EQ    & 0.28    & 0.25  & 2.99    & 15.56  & 14.95  & 100.48  \\
-    E500B5P50RD    & 0.30    & 0.28  & 113.88  & 16.26  & 15.49  & 1560.54 \\
-    E500B5P50EQ    & 0.35    & 0.28  & 14.06   & 17.97  & 15.54  & 546.91  \\
-    E1000B10P100RD & 0.43    & 0.37  & Timeout & 18.36  & 21.41  & N/A     \\
-    E1000B10P100EQ & 0.61    & 0.35  & 77.83   & 21.49  & 17.66  & 5822.66 \\
-    E2000B20P200RD & 3.25    & 2.62  & Timeout & 241.07 & 283.45 & N/A     \\
-    E2000B20P200EQ & Timeout & 46.45 & Timeout & N/A    & 638.00 & N/A     \\
+    E250B5P10RD    & 0.19 & 0.18 & 5.82    & 15.36  & 14.76  & 114.06  \\
+    E250B5P10EQ    & 0.21 & 0.18 & 2.83    & 15.56  & 14.95  & 100.48  \\
+    E500B5P50RD    & 0.23 & 0.22 & 37.28   & 16.26  & 15.49  & 524.89  \\
+    E500B5P50EQ    & 0.26 & 0.21 & 14.06   & 17.97  & 15.54  & 546.91  \\
+    E1000B10P100RD & 0.31 & 0.37 & Timeout & 18.36  & 21.41  & N/A     \\
+    E1000B10P100EQ & 0.37 & 0.28 & 77.83   & 20.41  & 17.66  & 5822.66 \\
+    E2000B20P200RD & 1.50 & 2.75 & Timeout & 239.71 & 283.45 & N/A     \\
+    E2000B20P200EQ & 3.34 & 3.03 & Timeout & 107.92 & 102.46 & N/A     \\
 \end{tabular}

### 2025-01-19 main.tex
update results

@@
 Our experiments are performed on a laptop with an Apple M4 Pro CPU and 24 GB RAM.
-\Cref{tab:benchmark} presents the results of equivalence checking based on derivatives (DRV), Thompson's construction (TC) and SymKAT~\cite{pous_SymbolicAlgorithmsLanguage_2015} (SK).
+\Cref{tab:benchmark-time,tab:benchmark-memory} present the results of equivalence checking based on derivatives (DV), Thompson's construction (TC) and SymKAT~\cite{pous_SymbolicAlgorithmsLanguage_2015} (SK).
 Each benchmark consists of 50 expression pairs with properties described by the benchmark's name.
@@
 
-\begin{table*}
+\begin{table}
+\footnotesize
 \centering
-\begin{tabular}{l r r r r r r}
-    Benchmark & Time (DRV) & Time (TC) & Time (SK) & Mem (DRV) & Mem (TC) & Mem (SK) \\
+\setlength{\tabcolsep}{3pt}
+\begin{tabular}{l r r r r r}
+    Benchmark & DV$_{\textsf{BDD}}$ & TC$_{\textsf{BDD}}$ & DV$_{\textsf{SAT}}$ & TC$_{\textsf{SAT}}$ & SK \\
     \hline
-    E250B5P10RD    & 0.19 & 0.18 & 5.82    & 15.36  & 14.76  & 114.06  \\
-    E250B5P10EQ    & 0.21 & 0.18 & 2.83    & 15.56  & 14.95  & 100.48  \\
-    E500B5P50RD    & 0.23 & 0.22 & 37.28   & 16.26  & 15.49  & 524.89  \\
-    E500B5P50EQ    & 0.26 & 0.21 & 14.06   & 17.97  & 15.54  & 546.91  \\
-    E1000B10P100RD & 0.31 & 0.37 & Timeout & 18.36  & 21.41  & N/A     \\
-    E1000B10P100EQ & 0.37 & 0.28 & 77.83   & 20.41  & 17.66  & 5822.66 \\
-    E2000B20P200RD & 1.50 & 2.75 & Timeout & 239.71 & 283.45 & N/A     \\
-    E2000B20P200EQ & 3.34 & 3.03 & Timeout & 107.92 & 102.46 & N/A     \\
+    e250b5p10rd    &   0.19 &   0.18 & 0.20 & 0.23 &    5.82 \\
+    e250b5p10eq    &   0.21 &   0.18 & 0.18 & 0.29 &    2.83 \\
+    e500b5p50rd    &   0.23 &   0.22 & 0.27 & 0.39 &   37.28 \\
+    e500b5p50eq    &   0.26 &   0.21 & 0.28 & 0.57 &   14.06 \\
+    e1000b10p100rd &   0.31 &   0.37 & 0.48 & 0.85 & timeout \\
+    e1000b10p100eq &   0.37 &   0.28 & 0.44 & 1.04 &   77.83 \\
+    e2000b20p200rd &   1.50 &   2.75 & 0.99 & 2.01 & timeout \\
+    e2000b20p200eq &   3.34 &   3.03 & 0.67 & 3.14 & timeout \\
+    e3000b30p200rd &   1.75 &  26.38 & 1.91 & 4.46 & timeout \\
+    e3000b30p200eq &  17.86 &  22.59 & 1.09 & 5.25 & timeout \\
+    degenerate     & 220.53 & 232.63 & 0.42 & 0.93 & timeout
 \end{tabular}
-\caption{Benchmarks}\label{tab:benchmark}
-\end{table*}
+\caption{Benchmark Time Usage}\label{tab:benchmark-time}
+\end{table}
+
+\begin{table}
+\footnotesize
+\centering
+\setlength{\tabcolsep}{3pt}
+\begin{tabular}{l r r r r r}
+    Benchmark & DV$_{\textsf{BDD}}$ & TC$_{\textsf{BDD}}$ & DV$_{\textsf{SAT}}$ & TC$_{\textsf{SAT}}$ & SK \\
+    \hline
+    e250b5p10rd    &  15.36 &   14.76 &  6.81 &  7.02 &  114.06 \\
+    e250b5p10eq    &  15.56 &   14.95 &  7.06 &  7.25 &  100.48 \\
+    e500b5p50rd    &  16.26 &   15.49 &  7.02 &  7.25 &  524.89 \\
+    e500b5p50eq    &  17.97 &   15.54 &  7.07 &  7.27 &  546.91 \\
+    e1000b10p100rd &  18.36 &   21.41 &  8.22 &  7.99 & timeout \\
+    e1000b10p100eq &  20.41 &   17.66 &  8.69 &  7.63 & 5822.66 \\
+    e2000b20p200rd & 239.71 &  283.45 & 13.33 & 11.92 & timeout \\
+    e2000b20p200eq & 107.92 &  102.46 & 13.20 & 14.02 & timeout \\
+    e3000b30p200rd & 112.59 & 1235.46 & 21.44 & 20.10 & timeout \\
+    e3000b30p200eq & 245.92 &  228.85 & 21.41 & 18.15 & timeout \\
+    degenerate     & 631.47 & 1229.30 & 19.62 & 18.43 & timeout
+\end{tabular}
+\caption{Benchmark Memory Usage}\label{tab:benchmark-memory}
+\end{table}
 
@@
 ]
-pub type Deriv = Vec<(BExp, Exp, u64)>;
-
-pub struct Solver {
+pub type Deriv<B> = Vec<(B, Exp<B>, u64)>;
+
+pub struct Solver<B> {
     // search states
-    dead_states: HashSet<Exp>,
-    explored: HashSet<Exp>,
-    uf_table: HashMap<Exp, UnionFindNode<()>>,
+    dead_states: HashSet<Exp<B>>,
+    explored: HashSet<Exp<B>>,
+    uf_table: HashMap<Exp<B>, UnionFindNode<()>>,
     // caching
-    eps_cache: LruCache<Exp, BExp>,
-    deriv_cache: LruCache<Exp, Deriv>,
+    eps_cache: HashMap<Exp<B>, B>,
+    drv_cache: HashMap<Exp<B>, Deriv<B>>,
 }
 
-impl Solver {
+impl<B: BExp> Solver<B> {
     #[inline]
-    pub fn known_dead(&self, exp: &Exp) -> bool {
+    pub fn known_dead(&self, exp: &Exp<B>) -> bool {
         self.dead_states.contains(&exp)
@@
 
-    pub fn is_dead(&mut self, gkat: &mut Gkat, exp: &Exp) -> bool {
+    pub fn is_dead<G: Gkat<B>>(&mut self, gkat: &mut G, exp: &Exp<B>) -> bool {
         let mut stack = Vec::new();
@@
             let eps = self.epsilon(gkat, &exp);
-            if eps.is_false() {
-                for (b, e, _) in self.derivative(gkat, &exp) {
-                    // check is not strictly needed due to eager-pruning
-                    if b.is_false() {
-                        continue;
-                    }
+            if gkat.is_false(&eps) {
+                for (_, e, _) in self.derivative(gkat, &exp) {
                     stack.push(e);

### 2025-01-19 main.tex
revise evaluation section

@@
 
-\section{Implementation}
+\section{Implementation and Evaluation}
 We implement our symbolic on-the-fly bisimulation algorithm~(\Cref{alg:symb-bisim}) in Rust along with derivative based~(\Cref{fig:derivatives-rules}) and Thompson's construction based~(\Cref{tab:symb-Thompson-construction}) algorithms for constructing symbolic GKAT coalgebras.
@@
 
-\subsection{Performance}\label{sec:performance-implementation}
+\subsection{Evaluation}\label{sec:performance-implementation}
+Due to the fact that our symbolic algorithms are generic regarding the choice of solvers for boolean satisfiability, we compare their performance using BDD to using a SAT solver (MiniSat~\cite{een_MINISAT_2004}).
 Our experiments are performed on a laptop with an Apple M4 Pro CPU and 24 GB RAM.
-\Cref{tab:benchmark-time,tab:benchmark-memory} present the results of equivalence checking based on derivatives (DV), Thompson's construction (TC) and SymKAT~\cite{pous_SymbolicAlgorithmsLanguage_2015} (SK).
+\Cref{tab:benchmark-time,tab:benchmark-memory} present the experimental results of equivalence checking based on derivatives (DV), Thompson's construction (TC) and SymKAT~\cite{pous_SymbolicAlgorithmsLanguage_2015} (SK).
+Note that SymKAT uses BDD internally for checking boolean satisfiability.
 Each benchmark consists of 50 expression pairs with properties described by the benchmark's name.
-For example, expressions in E250B5P10RD have around 250 primitive actions in total (E250), a maximum boolean expression size of 5 (B5), 10 possible unique primitive tests (P10) and are completely random (RD).
-Benchmarks with the suffix EQ have expression pairs that are known to be equivalent.
-The time columns measure the \textit{total time} used (in seconds) for checking all expression pairs in the benchmarks.
-The mem columns measure the \textit{peak memory} used (in MB).
-
-It is clear from these results that our algorithms (DRV and TC) can efficiently check even large programs in just a few seconds.
-
-\begin{table}
+For example, expressions in `e250b5p10eq' have approximately 250 primitive actions in total (e250), a maximum boolean expression size of 5 (b5), 10 possible unique primitive tests (p10) and are known to be equivalent (eq).
+Benchmarks with the suffix `ne' have expression pairs that are known to be non-equivalent.
+
+\begin{table}[H]
 \footnotesize
@@
     \hline
-    e250b5p10rd    &   0.19 &   0.18 & 0.20 & 0.23 &    5.82 \\
+    e250b5p10ne    &   0.19 &   0.18 & 0.20 & 0.23 &    5.82 \\
     e250b5p10eq    &   0.21 &   0.18 & 0.18 & 0.29 &    2.83 \\
-    e500b5p50rd    &   0.23 &   0.22 & 0.27 & 0.39 &   37.28 \\
+    e500b5p50ne    &   0.23 &   0.22 & 0.27 & 0.39 &   37.28 \\
     e500b5p50eq    &   0.26 &   0.21 & 0.28 & 0.57 &   14.06 \\
-    e1000b10p100rd &   0.31 &   0.37 & 0.48 & 0.85 & timeout \\
+    e1000b10p100ne &   0.31 &   0.37 & 0.48 & 0.85 & timeout \\
     e1000b10p100eq &   0.37 &   0.28 & 0.44 & 1.04 &   77.83 \\
-    e2000b20p200rd &   1.50 &   2.75 & 0.99 & 2.01 & timeout \\
+    e2000b20p200ne &   1.50 &   2.75 & 0.99 & 2.01 & timeout \\
     e2000b20p200eq &   3.34 &   3.03 & 0.67 & 3.14 & timeout \\
-    e3000b30p200rd &   1.75 &  26.38 & 1.91 & 4.46 & timeout \\
+    e3000b30p200ne &   1.75 &  26.38 & 1.91 & 4.46 & timeout \\
     e3000b30p200eq &  17.86 &  22.59 & 1.09 & 5.25 & timeout \\
@@
 \end{tabular}
-\caption{Benchmark Time Usage}\label{tab:benchmark-time}
+\caption{Total Time Usage}\label{tab:benchmark-time}
 \end{table}
 
-\begin{table}
+\begin{table}[H]
 \footnotesize
@@
     \hline
-    e250b5p10rd    &  15.36 &   14.76 &  6.81 &  7.02 &  114.06 \\
+    e250b5p10ne    &  15.36 &   14.76 &  6.81 &  7.02 &  114.06 \\
     e250b5p10eq    &  15.56 &   14.95 &  7.06 &  7.25 &  100.48 \\
-    e500b5p50rd    &  16.26 &   15.49 &  7.02 &  7.25 &  524.89 \\
+    e500b5p50ne    &  16.26 &   15.49 &  7.02 &  7.25 &  524.89 \\
     e500b5p50eq    &  17.97 &   15.54 &  7.07 &  7.27 &  546.91 \\
-    e1000b10p100rd &  18.36 &   21.41 &  8.22 &  7.99 & timeout \\
+    e1000b10p100ne &  18.36 &   21.41 &  8.22 &  7.99 & timeout \\
     e1000b10p100eq &  20.41 &   17.66 &  8.69 &  7.63 & 5822.66 \\
-    e2000b20p200rd & 239.71 &  283.45 & 13.33 & 11.92 & timeout \\
+    e2000b20p200ne & 239.71 &  283.45 & 13.33 & 11.92 & timeout \\
     e2000b20p200eq & 107.92 &  102.46 & 13.20 & 14.02 & timeout \\
-    e3000b30p200rd & 112.59 & 1235.46 & 21.44 & 20.10 & timeout \\
+    e3000b30p200ne & 112.59 & 1235.46 & 21.44 & 20.10 & timeout \\
     e3000b30p200eq & 245.92 &  228.85 & 21.41 & 18.15 & timeout \\
@@
 \end{tabular}
-\caption{Benchmark Memory Usage}\label{tab:benchmark-memory}
+\caption{Peak Memory Usage}\label{tab:benchmark-memory}
 \end{table}
+
+We observe that all variants of DV and TC perform better than SymKAT in terms of time and memory usage.
+When comparing BDD to SAT, notice that BDD is marginally faster than SAT for smaller benchmarks, however, its time and memory usage degrades for larger ones.
+On the other hand, the performance of SAT is stable for all benchmarks.
+Furthermore, we have encountered BDD-degenerative expressions (the `degenerate' benchmark) which greatly reduce BDD performance but can be efficiently solved by SAT.
+We did not encounter any SAT-degenerative expressions where BDD performed significantly better than SAT.
 

### 2025-01-19 main.tex
fix typo
small fix

@@
 
-Because of the functoriality, we can show that two states are infinite-trace equivalent implies these two states are finite-trance equivalent.
+Because of the functoriality, we can show that two states are infinite-trace equivalent implies these two states are finite-trace equivalent.
 This gives us more tool in proving semantic equivalence between two states in GKAT coalgebras: proving bisimulation of these two states in the \emph{non-normalized} GKAT coalgebra can also obtain semantic equivalence for two states.
@@
 
-The final optimization that we perform is collapsing the sets of boolean expressions of \(\hat{\epsilon}(s)\) into a single boolean expression.
+The final optimization that we perform is collapsing the boolean expression set \(\hat{\epsilon}(s)\) into a single boolean expression.
 Basically, each \(\hat{\epsilon}(s)\) is represented by the boolean disjunction of its elements.

### 2025-01-19 main.tex
remove mention of Z3
update units in table

@@
 \end{tabular}
-\caption{Total Time Usage}\label{tab:benchmark-time}
+\caption{Total Time Usage (seconds)}\label{tab:benchmark-time}
 \end{table}
@@
 \end{tabular}
-\caption{Peak Memory Usage}\label{tab:benchmark-memory}
+\caption{Peak Memory Usage (megabytes)}\label{tab:benchmark-memory}
 \end{table}
@@
 Although some of our examples utilizes BDD to solve equivalence and inequivalence of boolean expressions, because of the structure of GKAT, our algorithm is not bounded by a particular boolean representation or solvers.
-In fact, we have demonstrated that in many scenarios, solvers like z3 and sentential decision diagram (SDD) can achieve better performance than BDD.
+In fact, we have demonstrated that in many scenarios, SAT solvers like MiniSat~\cite{een_MINISAT_2004} can achieve better performance than BDD.
 

### 2025-01-19 main.tex
repo url
minor change to evaluation conclusion

@@
 We implement our symbolic on-the-fly bisimulation algorithm~(\Cref{alg:symb-bisim}) in Rust along with derivative based~(\Cref{fig:derivatives-rules}) and Thompson's construction based~(\Cref{tab:symb-Thompson-construction}) algorithms for constructing symbolic GKAT coalgebras.
-The source code and benchmark suite is freely available in our repository TODO.
+The source code and benchmark suite is freely available in our repository \url{https://anonymous.4open.science/r/rust-gkat-071E}.
 
@@
 We observe that all variants of DV and TC perform better than SymKAT in terms of time and memory usage.
-When comparing BDD to SAT, notice that BDD is marginally faster than SAT for smaller benchmarks, however, its time and memory usage degrades for larger ones.
+When comparing BDD to SAT, notice that BDD is marginally faster than SAT for smaller benchmarks, however, its time and memory usage increases for larger ones.
 On the other hand, the performance of SAT is stable for all benchmarks.
-Furthermore, we have encountered BDD-degenerative expressions (the `degenerate' benchmark) which greatly reduce BDD performance but can be efficiently solved by SAT.
-We did not encounter any SAT-degenerative expressions where BDD performed significantly better than SAT.
+Furthermore, we have encountered degenerative expression pairs (the `degenerate' benchmark) which greatly reduce BDD performance but can be efficiently checked by SAT.
+We did not encounter any expression pairs where BDD performed significantly better than SAT.
 

### 2025-01-19 main.tex
small change

@@
 We observe that all variants of DV and TC perform better than SymKAT in terms of time and memory usage.
-When comparing BDD to SAT, notice that BDD is marginally faster than SAT for smaller benchmarks, however, its time and memory usage increases for larger ones.
+When comparing the solver backends, BDD is marginally faster than SAT for smaller benchmarks, however, its time and memory usage increases for larger ones.
 On the other hand, the performance of SAT is stable for all benchmarks.

### 2025-01-20 main.tex
merge

@@
     \end{align*}
+\end{proof}
+
+\begin{corollary}\label{thm:max-bisim-is-max-greedy-bisim}
+    Given two GKAT coalgebra \(S\) and \(U\), carrier of the maximal bisimulation \({≡} ⊆ S × U\) is also the carrier of maximal greedy bisimulation.
+    We will abuse the notation \(≡\) to represent both the maximal bisimulation and maximal greedy bisimulation.
+\end{corollary}
+
+\begin{proof}
+    Consider the GKAT coalgebra \(≡_d\) as generated in~\cref{thm:dead-construction-greedy-bisim}, then the carrier of \(≡_d\) and \(≡\) coincide, because \(≡\) already contains all the dead state pairs, by~\Cref{thm:bisim-between-dead}.
+    Because the 
 \end{proof}
@@
 
-Finally, we present our equivalence-checking algorithm as~\cref{alg:bisim}, where we first check if one of \(s\) or \(u\) is known to be dead, if so we only need to check whether the other is dead, because of~\Cref{thm:greedy-bisim-one-dead}; otherwise, we will check the conditions in~\Cref{thm:recursive-construction}.
-Because trace equivalence is an equivalence relation, therefore we can organize the previously explored states into equivalent classes using an efficient union-find structure, instead of a set of state pairs.
-This union-find structure need to provide two functions: \(\Call{union}{$s, u$}\) unions the equivalence class of \(s\) and \(u\), whereas \(\Call{eq}{$s, u$}\) checks whether \(s\) and \(u\) is in the same equivalence class.
+Finally, we present our equivalence-checking algorithm as~\cref{alg:bisim}, which are mostly checking the conditions in~\cref{thm:recursive-construction}.  
+In the algorithm, we use the text ``implies'', ``iff'', ``and'' are logical operators that takes booleans (i.e. true or false) and return a boolean.
+Similarly, \(\&\!\&\) is the logical and operation, which is the same as the the written ``and''. 
+We opt to use \(\&\!\&\), instead of ``and'', to separate the conditions for clarity; formally \(\&\!\&\) is the same function as ``and'', but binds less tightly.
+
+Intuitively, starting with the pair of state \((s, u)\), it will explore following \(δ_≈\) in~\cref{def:greedy-bisim}.
 
@@
 
-\section{Implementation and Evaluation}
+\section{Implementation and Evaluation}\label{sec:implementation}
 We implement our symbolic on-the-fly bisimulation algorithm~(\Cref{alg:symb-bisim}) in Rust along with derivative based~(\Cref{fig:derivatives-rules}) and Thompson's construction based~(\Cref{tab:symb-Thompson-construction}) algorithms for constructing symbolic GKAT coalgebras.

### 2025-01-20 main.tex
update results

--- 2025-01-19 main.tex
@@
     \hline
-    e250b5p10ne    &   0.19 &   0.18 & 0.20 & 0.23 &    5.82 \\
-    e250b5p10eq    &   0.21 &   0.18 & 0.18 & 0.29 &    2.83 \\
-    e500b5p50ne    &   0.23 &   0.22 & 0.27 & 0.39 &   37.28 \\
-    e500b5p50eq    &   0.26 &   0.21 & 0.28 & 0.57 &   14.06 \\
-    e1000b10p100ne &   0.31 &   0.37 & 0.48 & 0.85 & timeout \\
-    e1000b10p100eq &   0.37 &   0.28 & 0.44 & 1.04 &   77.83 \\
-    e2000b20p200ne &   1.50 &   2.75 & 0.99 & 2.01 & timeout \\
-    e2000b20p200eq &   3.34 &   3.03 & 0.67 & 3.14 & timeout \\
-    e3000b30p200ne &   1.75 &  26.38 & 1.91 & 4.46 & timeout \\
-    e3000b30p200eq &  17.86 &  22.59 & 1.09 & 5.25 & timeout \\
-    degenerate     & 220.53 & 232.63 & 0.42 & 0.93 & timeout
+    e250b5p10ne    &   0.19 &   0.18 & 0.17 & 0.21 &    5.82 \\
+    e250b5p10eq    &   0.21 &   0.18 & 0.24 & 0.34 &    2.83 \\
+    e500b5p50ne    &   0.23 &   0.22 & 0.25 & 0.28 &   37.28 \\
+    e500b5p50eq    &   0.26 &   0.21 & 0.24 & 0.34 &   14.06 \\
+    e1000b10p100ne &   0.31 &   0.37 & 0.42 & 0.52 & timeout \\
+    e1000b10p100eq &   0.37 &   0.28 & 0.35 & 0.52 &   77.83 \\
+    e2000b20p200ne &   1.50 &   2.75 & 0.84 & 1.11 & timeout \\
+    e2000b20p200eq &   3.34 &   3.03 & 0.51 & 1.30 & timeout \\
+    e3000b30p200ne &   1.75 &  26.38 & 1.59 & 2.29 & timeout \\
+    e3000b30p200eq &  17.86 &  22.59 & 0.81 & 2.02 & timeout \\
+    degenerate     & 220.53 & 232.63 & 0.28 & 0.37 & timeout
 \end{tabular}
@@
     e250b5p10eq    &  15.56 &   14.95 &  7.06 &  7.25 &  100.48 \\
-    e500b5p50ne    &  16.26 &   15.49 &  7.02 &  7.25 &  524.89 \\
-    e500b5p50eq    &  17.97 &   15.54 &  7.07 &  7.27 &  546.91 \\
-    e1000b10p100ne &  18.36 &   21.41 &  8.22 &  7.99 & timeout \\
+    e500b5p50ne    &  16.26 &   15.49 &  7.02 &  6.99 &  524.89 \\
+    e500b5p50eq    &  17.97 &   15.54 &  7.07 &  7.01 &  546.91 \\
+    e1000b10p100ne &  18.36 &   21.41 &  8.43 &  8.07 & timeout \\
     e1000b10p100eq &  20.41 &   17.66 &  8.69 &  7.63 & 5822.66 \\
-    e2000b20p200ne & 239.71 &  283.45 & 13.33 & 11.92 & timeout \\
-    e2000b20p200eq & 107.92 &  102.46 & 13.20 & 14.02 & timeout \\
-    e3000b30p200ne & 112.59 & 1235.46 & 21.44 & 20.10 & timeout \\
-    e3000b30p200eq & 245.92 &  228.85 & 21.41 & 18.15 & timeout \\
-    degenerate     & 631.47 & 1229.30 & 19.62 & 18.43 & timeout
+    e2000b20p200ne & 239.71 &  283.45 & 13.54 & 12.27 & timeout \\
+    e2000b20p200eq & 107.92 &  102.46 & 13.20 & 13.84 & timeout \\
+    e3000b30p200ne & 112.59 & 1235.46 & 21.69 & 20.33 & timeout \\
+    e3000b30p200eq & 245.92 &  228.85 & 21.79 & 18.33 & timeout \\
+    degenerate     & 631.47 & 1229.30 & 19.75 & 18.57 & timeout
 \end{tabular}

### 2025-01-20 main.tex
minor change

@@
 \subsection{Evaluation}\label{sec:performance-implementation}
-Due to the fact that our symbolic algorithms are generic regarding the choice of solvers for boolean satisfiability, we compare their performance using BDD to using a SAT solver (MiniSat~\cite{een_MINISAT_2004}).
-Our experiments are performed on a laptop with an Apple M4 Pro CPU and 24 GB RAM.
-\Cref{tab:benchmark-time,tab:benchmark-memory} present the experimental results of equivalence checking based on derivatives (DV), Thompson's construction (TC) and SymKAT~\cite{pous_SymbolicAlgorithmsLanguage_2015} (SK).
-Note that SymKAT uses BDD internally for checking boolean satisfiability.
-Each benchmark consists of 50 expression pairs with properties described by the benchmark's name.
-For example, expressions in `e250b5p10eq' have approximately 250 primitive actions in total (e250), a maximum boolean expression size of 5 (b5), 10 possible unique primitive tests (p10) and are known to be equivalent (eq).
-Benchmarks with the suffix `ne' have expression pairs that are known to be non-equivalent.
-
-\begin{table}[H]
+\begin{table}
 \footnotesize
@@
 \end{table}
-
-\begin{table}[H]
+\begin{table}
 \footnotesize
@@
 \end{table}
+
+Due to the fact that our symbolic algorithms are generic regarding the choice of solvers for boolean satisfiability, we compare their performance using BDD to using a SAT solver (MiniSat~\cite{een_MINISAT_2004}).
+Our experiments are performed on a laptop with an Apple M4 Pro CPU and 24 GB RAM.
+\Cref{tab:benchmark-time,tab:benchmark-memory} present the experimental results of equivalence checking based on derivatives (DV), Thompson's construction (TC) and SymKAT~\cite{pous_SymbolicAlgorithmsLanguage_2015} (SK).
+Note that SymKAT uses BDD internally for checking boolean satisfiability.
+Each benchmark consists of 50 expression pairs with properties described by the benchmark's name.
+For example, expressions in `e250b5p10eq' have approximately 250 primitive actions in total (e250), a maximum boolean expression size of 5 (b5), 10 possible unique primitive tests (p10) and are known to be equivalent (eq).
+Benchmarks with the suffix `ne' have expression pairs that are known to be non-equivalent.
 

### 2025-01-20 main.tex
minor change

@@
 We observe that all variants of DV and TC perform better than SymKAT in terms of time and memory usage.
-When comparing the solver backends, BDD is marginally faster than SAT for smaller benchmarks, however, its time and memory usage increases for larger ones.
+When comparing the solver backends, BDD is marginally faster than SAT for smaller benchmarks, however, BDD's time and memory usage increases considerably for larger benchmarks.
 On the other hand, the performance of SAT is stable for all benchmarks.

### 2025-01-20 main.tex
first sketch of the proof

@@
 \lstset{language=caml, escapeinside={[*}{*]}}
+% Clever ref names
+\crefname{lstlisting}{listing}{listings}
+\Crefname{lstlisting}{Listing}{Listings}
 
@@
 
-\begin{corollary}\label{thm:max-bisim-is-max-greedy-bisim}
-    Given two GKAT coalgebra \(S\) and \(U\), carrier of the maximal bisimulation \({≡} ⊆ S × U\) is also the carrier of maximal greedy bisimulation.
-    We will abuse the notation \(≡\) to represent both the maximal bisimulation and maximal greedy bisimulation.
-\end{corollary}
-
-\begin{proof}
-    Consider the GKAT coalgebra \(≡_d\) as generated in~\cref{thm:dead-construction-greedy-bisim}, then the carrier of \(≡_d\) and \(≡\) coincide, because \(≡\) already contains all the dead state pairs, by~\Cref{thm:bisim-between-dead}.
-    Because the 
-\end{proof}
-
 \begin{corollary}\label{thm:greedy-bisim-iff-norm-bisim}
@@
 
-% Similar to the correctness of bisimulation between normalized coalgebra (\Cref{thm:norm-bisim-correctness}), we also establish a correctness theorem for greedy bisimulation, i.e. there exists
-
-% \begin{theorem}[Correctness]
-%     Given two states in GKAT coalgebras \(s ∈ S\) and \(u ∈ ,U\), there exists a greedy bisimulation \({≈} ⊆ S × U\) s.t. \(s ≈ u\), if and only if \(⟦s⟧_S = ⟦u⟧_U\).
-% \end{theorem}
-
-% \begin{proof}
-%     \(⟹\) direction, because normalization preserves carrier set, therefore \((s, u) ∈ \norm(≈)\).
-%     And because \(\norm(≈)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\) (\Cref{thm:greedy-bisim-iff-norm-bisim}), \(⟦s⟧_S = ⟦u⟧_U\).
-
-%     \(⟸\) direction, because \(⟦s⟧_S = ⟦u⟧_U\), then by~\cref{thm:norm-bisim-correctness}, then there a bisimulation \(∼ ⊆ \norm(S) × \norm(U)\) s.t. \(s ∼ u\). 
-%     Because~\Cref{thm:greedy-bisim-exists-unique}, therefore there exists a greedy bisimulation \(≈\) with \(∼\) as its underlying bisimulation. 
-%     Because the carriers \({∼} ⊆ {≈}\), therefore \(s ∼ u\) implies \(s ≈ u\).
-% \end{proof}
+\begin{corollary}\label{thm:max-bisim-is-max-greedy-bisim}
+    Given two GKAT coalgebra \(S\) and \(U\), carrier of the maximal bisimulation \({≡} ⊆ \norm(S) × \norm(U)\) is also the carrier of maximal greedy bisimulation, which we will denote as \(≊ ⊆ S × U\).
+    We will abuse the notation \(≡\) to represent both the maximal bisimulation and maximal greedy bisimulation.
+\end{corollary}
+
+\begin{proof}
+    by definition of greedy bisimulation, the carrier of \({≡}\) is a subset of the carrier of \({≊}\). 
+    Then because of~\Cref{thm:greedy-bisim-iff-norm-bisim}, \(\norm(≊)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\), which implies \(\norm(≡') ⊑ {≡}\).
+    Finally because \(\norm\) do not change the carrier set, therefore the carrier of \(≡\) and \(≊\) necessarily coincide.
+
+    Then we will need to show that \(≡\) with the \(δ_≈\) defined in~\Cref{def:greedy-bisim} is the maximal greedy bisimulation.
+    Take an arbitrary greedy bisimulation \({≈} ⊆ S × U\) with underlying bisimulation \(∼\), and a pair of states \(s ≈ u\), we will need to show \(s ≡ u\).
+    By~\cref{thm:greedy-bisim-dead-or-bisim}, we can case analysis on \(s\) and \(u\): if \(s ∼ u\), then \(s ≡ u\) because \(≡\) is the maximal bisimulation; and if \(s\) and \(u\) are both dead, then by~\cref{thm:bisim-between-dead}, \(s ≡ u\).
+\end{proof}
+
+\Cref{thm:max-bisim-is-max-greedy-bisim} allows us to take the maximal bisimulation as a greedy bisimulation, aiding several proofs.
+
+Similar to the correctness of bisimulation between normalized coalgebra (\Cref{thm:norm-bisim-correctness}), we also establish a correctness theorem for greedy bisimulation, i.e. there exists
+
+\begin{theorem}[Correctness]
+    Given two states in GKAT coalgebras \(s ∈ S\) and \(u ∈ ,U\), there exists a greedy bisimulation \({≈} ⊆ S × U\) s.t. \(s ≈ u\), if and only if \(⟦s⟧_S = ⟦u⟧_U\).
+\end{theorem}
+
+\begin{proof}
+    \(⟹\) direction, because normalization preserves carrier set, therefore \((s, u) ∈ \norm(≈)\).
+    And because \(\norm(≈)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\) (\Cref{thm:greedy-bisim-iff-norm-bisim}), \(⟦s⟧_S = ⟦u⟧_U\).
+
+    \(⟸\) direction, because \(⟦s⟧_S = ⟦u⟧_U\), then by~\cref{thm:norm-bisim-correctness}, then there a bisimulation \(∼ ⊆ \norm(S) × \norm(U)\) s.t. \(s ∼ u\). 
+    Because~\Cref{thm:greedy-bisim-exists-unique}, therefore there exists a greedy bisimulation \(≈\) with \(∼\) as its underlying bisimulation. 
+    Because the carriers \({∼} ⊆ {≈}\), therefore \(s ∼ u\) implies \(s ≈ u\).
+\end{proof}
 
@@
 
-\begin{theorem}[Equivalence Relation]
+\begin{theorem}[Equivalence Relation]\label{thm:greedy-bisim-iff-greedy-bisim-equiv}
     Given two states in GKAT coalgebra \(s ∈ S\) and \(u ∈ U\), there exists a greedy bisimulation \(≈\) between \(S\) and \(U\) s.t.
-\(s ≈ u\), if and only if there exists a greedy bisimulation that is an \emph{equivalence relation} \(≊\) in \(S + U\) s.t.
-\(s ≊ u\).
+\(s ≈ u\), if and only if there exists a greedy bisimulation that is an \emph{equivalence relation} \(≃\) in \(S + U\) s.t.
+\(s ≃ u\).
 \end{theorem}
@@
     The \(⟹\) direction. 
-    Because sub-coalgebra preserves bisimulation (\Cref{thm:sub-coalg-preserve-bisim}), then \(\norm(≈)\) is a bisimulation between \(\norm(S + U)\) and itself.
-    We consider \(≡\) to be the maximal bisimulation between \(\norm(S + U)\) and itself, then \(≡\) is an equivalence relation~\cite[Corollary 5.6]{rutten_UniversalCoalgebraTheory_2000}.
-
-    Let \(≊\) be the greedy bisimulation generated by \(≡\), then \(≊\) has the same carrier as \(≡\).
-    All the \((s, u) ∈ {≊}\) are either \(s ≡ u\) or \(s\) and \(u\) are dead in \(S\) and \(U\), which implies that both are dead in \(S + U\).
-    However, \(≡\) already contains all the dead pairs (\cref{thm:bisim-between-dead}), therefore for all \((s, u) ∈ {≊}\), they are also in \(≡\), implying \({≡} ⊇ {≊}\).
-    Finally, by definition of greedy bisimulation (\Cref{def:greedy-bisim}), \({≡} ⊆ {≊}\).
-    Therefore, \(≡\) and \(≊\) share the same carrier and \(≊\) is a greedy bisimulation that is also an equivalence relation, which also contains \((s, u)\):
-    \[(s, u) ∈ {≈} = \norm(≈) ⊆ {≡} ⊆ {≊},\]
-    where all the \(=\) and \(⊆\) are orders on the carrier set, instead of the coalgebra.
-
-    The \(⟸\) direction can be split into two cases, by whether \((s, u)\) is in the underlying bisimulation of \(≊\), which we denote as \(≃\): 
-    \begin{itemize}
-        \item If \(s ≃ u\), then we can reflect \(≃\)  through \(\norm(S) ⊑ \norm(S + U)\) and \(\norm(U) ⊑ \norm(S + U)\) (\Cref{thm:sub-coalg-preserve-bisim}) to obtain a bisimulation \({∼} ⊆ \norm(S) × \norm(U)\) and \(s ∼ u\).
-        Taking the greedy bisimulation over \(∼\) will yield a desired greedy bisimulation \(≈\) s.t. \(s ≈ u\).
-        \item If \((s, u)\) are not in \(≃\), then they are both dead in \(S + U\) (\Cref{thm:greedy-bisim-dead-or-bisim}), which implies \(s\) and \(u\) are respectively dead in \(S\) and \(U\) (\Cref{thm:sub-coalg-preserve-liveness}).
-        Then we take \(≈\) to be greedy bisimulation generated form the maximal bisimulation between \(S\) and \(U\), which contains \(s\) and \(u\) (\Cref{thm:bisim-between-dead}).
-        \qedhere
-    \end{itemize}
-\end{proof}
-
-\begin{theorem}\label{thm:greedy-bisim-one-dead}
-    Consider two states in GKAT coalgebras \(s ∈ S, u ∈ U\) and an arbitrary bisimulation \({∼} ⊆ \norm(S) × \norm(U)\), if \(s\) is dead, then 
-    \begin{enumerate}
-        \item\label{itm:greedy-bisim-one-dead-then-both-dead} if \(s ∼ u\), then \(u\) is dead.
-        \item\label{itm:remove-dead-pairs-sound} if \(u\) is dead, then \({∼} ∖ ⟨(s, u)⟩_∼\) is also a bisimulation between \(\norm(S)\) and \(\norm(U)\).
-    \end{enumerate}
-\end{theorem}
-
-\begin{proof}
-    \Cref{itm:greedy-bisim-one-dead-then-both-dead} is a consequence of bisimulation preserves liveness (\Cref{thm:bisim-preserve-liveness}) and normalization preserves liveness (\Cref{thm:norm-perserve-liveness}).
-
-    \Cref{itm:remove-dead-pairs-sound} can be proven by realizing that \((s, u)\) is a dead state in \(∼\) (\Cref{thm:bisim-preserve-liveness}) which means that all the states in \(⟨(s, u)⟩_∼\) are dead.
-    Because \(∼\) is a normalized GKAT coalgebra (\Cref{thm:}), therefore no state can transition into dead states like \(⟨(s, u)⟩_∼\), hence \({∼} ∖ ⟨(s, u)⟩_∼\) is still closed under its transition, and form a sub-coalgebra of \(∼\).
-    Finally because \(∼\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\), then so is \({∼} ∖ ⟨(s, u)⟩_∼\) (\Cref{thm:})
-\end{proof}
+    Because sub-coalgebra preserves bisimulation (\Cref{thm:sub-coalg-preserve-bisim}) and \(\norm(S) ⊑ \norm(S + U)\) and \(\norm(U) ⊑ \norm(S + U)\), we have \(\norm(≈)\) is a bisimulation between \(\norm(S + U)\) and itself.
+    let \(≡\) to be the maximal bisimulation within \(\norm(S + U)\), then \(≡\) is an equivalence relation~\cite[Corollary 5.6]{rutten_UniversalCoalgebraTheory_2000}, and so is the maximal greedy bisimulation \(≊\) (\Cref{thm:max-bisim-is-max-greedy-bisim}).
+    This gives us the following chain of set inclusions:
+    \[(s, u) ∈ {≈} = \norm(≈) ⊑ {≡} ⊆ {≊}.\]
+    Thus \(≊\) is a equivalence relation~\cite[Corollary 5.6]{rutten_UniversalCoalgebraTheory_2000} and a greedy bisimulation s.t. \(s ≡ u\).
+
+    The \(⟸\) direction can be shown by considering the maximal greedy bisimulation \(≊\) within \(S + U\) and maximal bisimulation \(≡\) within \(\norm(S + U)\).
+    By maximality and the carrier of \(≡\) coincide with \(≊\), we have the following set inclusions:
+    \[(s, u) ∈ {≃} ⊑ {≊} = {≡}.\]
+    Finally, we reflect \(≡\) through the inclusion homomorphisms \(\norm(S) ⊑ \norm(S + U)\) and \(\norm(U) ⊑ \norm(S + U)\) as in~\Cref{thm:sub-coalg-preserve-bisim} to obtain \({∼} ⊆ \norm(S) × \norm(U)\) and \(s ∼ u\).
+    Therefore \((s, u)\) is also in the greedy bisimulation \({≈} ⊆ \norm(S) × \norm(U)\), generated by \(∼\).
+\end{proof}
+
+% \begin{theorem}\label{thm:greedy-bisim-one-dead}
+%     Consider two states in GKAT coalgebras \(s ∈ S, u ∈ U\) and an arbitrary bisimulation \({∼} ⊆ \norm(S) × \norm(U)\), if \(s\) is dead, then 
+%     \begin{enumerate}
+%         \item\label{itm:greedy-bisim-one-dead-then-both-dead} if \(s ∼ u\), then \(u\) is dead.
+%         \item\label{itm:remove-dead-pairs-sound} if \(u\) is dead, then \({∼} ∖ ⟨(s, u)⟩_∼\) is also a bisimulation between \(\norm(S)\) and \(\norm(U)\).
+%     \end{enumerate}
+% \end{theorem}
+
+% \begin{proof}
+%     \Cref{itm:greedy-bisim-one-dead-then-both-dead} is a consequence of bisimulation preserves liveness (\Cref{thm:bisim-preserve-liveness}) and normalization preserves liveness (\Cref{thm:norm-perserve-liveness}).
+
+%     \Cref{itm:remove-dead-pairs-sound} can be proven by realizing that \((s, u)\) is a dead state in \(∼\) (\Cref{thm:bisim-preserve-liveness}) which means that all the states in \(⟨(s, u)⟩_∼\) are dead.
+%     Because \(∼\) is a normalized GKAT coalgebra (\Cref{thm:}), therefore no state can transition into dead states like \(⟨(s, u)⟩_∼\), hence \({∼} ∖ ⟨(s, u)⟩_∼\) is still closed under its transition, and form a sub-coalgebra of \(∼\).
+%     Finally because \(∼\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\), then so is \({∼} ∖ ⟨(s, u)⟩_∼\) (\Cref{thm:})
+% \end{proof}
 
@@
 The function \Call{isDead\(_S\)}{$s$} will first check if \(s\) is known to be dead, and invoke the search algorithm in the coalgebra \(⟨s⟩\), if \(s\) is not in the cached dead states.
-Because the non-symbolic version of liveness checking is similar to the symbolic version, we only provide the symbolic version of this algorithm in appendix
-TODO: give the link to the section.
+Because the non-symbolic version of liveness checking is similar to the symbolic version, we only provide the symbolic version of this algorithm in~\cref{lst:dead-state-detection}.
 
@@
 
-Intuitively, starting with the pair of state \((s, u)\), it will explore following \(δ_≈\) in~\cref{def:greedy-bisim}.
+Intuitively, starting with the pair of state \((s, u)\), it will explore all of its reachable state following \(δ_≈\) in~\cref{def:greedy-bisim}.
+In other word, if the algorithm terminates and returned true, we will have explored every state in \(⟨(s, u)⟩_{≊}\) where \(≊\) is the maximal greedy bisimulation between \(S\) and \(U\).
+However, our algorithm slightly deviates from the ideal scenario described above, specifically we implemented two optimization.
+First, we uses a union-find data structure to keep track of existing state, which organizes the explored states into equivalence classes.
+Indeed, by looking for a greedy bisimulation that is also equivalence relation, we can use more efficient structures like union-find, instead of set of pairs; and the correctness of such optimization is justified by~\Cref{thm:greedy-bisim-iff-greedy-bisim-equiv}.
+The second optimization is to terminate and return false as soon as we know one of \(s, u\) is dead and the other is not dead.
+Indeed, in order for \(s ≈ u\) for any greedy bisimulation \(≈\), \(s\) and \(u\) necessarily have the same liveness in their respective GKAT coalgebra.
+And in the other case, if both \(s\) and \(u\) are dead in their respective coalgebra, we will skip the check on all the reachable state of \((s, u)\) following \(δ_≈\), because a pair of dead state is necessarily in the maximal greedy bisimulation \(≊\) (\Cref{thm:bisim-between-dead,thm:max-bisim-is-max-greedy-bisim}).
 

### 2025-01-20 main.tex
setup proof at end

@@
 \usepackage[margin=2.5cm]{geometry}
+
+% move proofs to the end
+\usepackage[conf={restate,no link to proof}]{proof-at-the-end}
 
@@
 Sub-coalgebras are also preserved under homomorphic images and pre-images: 
-\begin{lemma}[Theorem 6.3~\cite{rutten_UniversalCoalgebraTheory_2000}]\label{thm:hom-(pre)img-preserve-sub-coalg}
+\begin{theoremEnd}{lemma}[Theorem 6.3~\cite{rutten_UniversalCoalgebraTheory_2000}]\label{thm:hom-(pre)img-preserve-sub-coalg}
     given a homomorphism \(h: S → U\), and sub-coalgebras \(S' ⊑ S\) and \(U' ⊑ U\), then 
     \[h(S') ⊑ U \text{ and } h^{-1}(U') ⊑ S.\]
-\end{lemma}
+\end{theoremEnd}
 
@@
 Similar to sub-coalgebra, principle sub-coalgebra is also preserved under homomorphic image:
-\begin{theorem}\label{thm:homo-img-preserve-principle-sub-coalg}
+\begin{theoremEnd}{theorem}\label{thm:homo-img-preserve-principle-sub-coalg}
     Homomorphic image preserves principle sub-GKAT coalgebra. Specifically, given a homomorphism \(h: S → U\):
     \[h(⟨s⟩_S) = ⟨h(s)⟩_U\]
-\end{theorem}
-
-\begin{proof}
+\end{theoremEnd}
+
+\begin{proofEnd}
     We will need to show that \(h(⟨s⟩_{S})\) is the smallest sub-GKAT coalgebra of \(S\) that contain \(h(s)\). 
@@
     Hence \(h(⟨s⟩_S)\) is the smallest sub-GKAT coalgebra of \(U\) that contains \(h(s)\).
-\end{proof}
+\end{proofEnd}
 
@@
 When given a homomorphism \(h: S → U\), we can construct a bisimulation \({∼} ⊆ S × U\), where it relates all the states \(s ∈ S\) with \(h(s) ∈ U\); and each component of the pair will transition individually via their respective transition function \(δ_∼ ≜ δ_S × δ_U\), where \(×\) is the bifunctorial lift of product.
-\begin{corollary}\label{thm:hom-preserves-semantics}
+\begin{theoremEnd}{corollary}\label{thm:hom-preserves-semantics}
     Given a homomorphism \(h: S → U\) between two \(F\)-coalgebra \(S, U\), then all for all states \(s ∈ S\), \(\mathrm{beh}_S(s) = \mathrm{beh}_U(h(s)).\)
-\end{corollary}
+\end{theoremEnd}
 
@@
 In particular, \(G\) is a simple polynomial functor~\cite{jacobs_IntroductionCoalgebraMathematics_2016}, allowing the sub-coalgebra to preserve and reflect bisimulations.
-\begin{theorem}[sub-coalgebras perserve and reflect bisimulation]\label{thm:sub-coalg-preserve-bisim}
+\begin{theoremEnd}{theorem}[sub-coalgebras perserve and reflect bisimulation]\label{thm:sub-coalg-preserve-bisim}
     Given two states in sub-coalgebra \(s ∈ S' ⊑ S\) and \(u ∈ U' ⊑ U\), there exists a bisimulation \({∼'} ⊆ S' × U'\) s.t.
@@
 \(s ∼ u\).
-\end{theorem}
-
-\begin{proof}
+\end{theoremEnd}
+
+\begin{proofEnd}
     For \(⟹\) direction, we will show that if \({∼'} ⊆ S' × U'\) is a bisimulation, then \({∼'}\) is also a bisimulation between \(S\) and \(U\):
@@
     Recall that elementary polynomial functor~\cite{jacobs_IntroductionCoalgebraMathematics_2016} like \(G\) preserves pullback, hence the pullback also uniquely generates a GKAT coalgebra~\cite{rutten_UniversalCoalgebraTheory_2000}
-\end{proof}
+\end{proofEnd}
 
@@
 Thus we obtain the following coalgebraic characterization of liveness:
-\begin{theorem}\label{thm:liveness-principle-sub-coalg}
+\begin{theoremEnd}{theorem}\label{thm:liveness-principle-sub-coalg}
     A state \(s\) is \emph{live} if there exists an accepting state \(s' ∈ ⟨s⟩\); and a state \(s\) is \emph{dead} if there is no accepting state in \(⟨s⟩\).
-\end{theorem}
+\end{theoremEnd}
 This alternative liveness definition can help us prove important theorems regarding reachability and liveness without explicitly performing induction on the liveness chain. 
 We show the following theorems as examples:
-\begin{lemma}\label{thm:dead-iff-all-reachable-dead}
+\begin{theoremEnd}{lemma}\label{thm:dead-iff-all-reachable-dead}
     A state \(s\) is dead if and only if all elements in \(⟨s⟩\) is dead.
-\end{lemma}
-\begin{proof}
+\end{theoremEnd}
+\begin{proofEnd}
     \(⟸\) direction is true, because \(s ∈ ⟨s⟩\): if all \(⟨s⟩\) is dead, then \(s\) is dead. 
@@
     Since there is no accepting state in \(⟨s⟩\), thus there cannot be any accepting state in \(⟨s'⟩\), hence \(s'\) is also dead.
-\end{proof}
-
-\begin{theorem}[homomorphism perserves liveness]\label{thm:hom-preserve-liveness}
+\end{proofEnd}
+
+\begin{theoremEnd}{theorem}[homomorphism perserves liveness]\label{thm:hom-preserve-liveness}
     Given a homomorphism \(h: S → U\) and a state \(s ∈ S\):
     \[\text{\(s\) is live} ⟺ \text{\(h(s)\) is live}\]
-\end{theorem}
-
-\begin{proof}
+\end{theoremEnd}
+
+\begin{proofEnd}
     Because homomorphic image preserves principle sub-GKAT coalgebra (\Cref{thm:homo-img-preserve-principle-sub-coalg})
@@
     Therefore, \(s\) is live in \(S\) if and only if \(h(s)\) is live in \(U\).
-\end{proof}
+\end{proofEnd}
 
@@
 
-\begin{corollary}[sub-coalgebra perserves liveness]\label{thm:sub-coalg-preserve-liveness}
+\begin{theoremEnd}{corollary}[sub-coalgebra perserves liveness]\label{thm:sub-coalg-preserve-liveness}
     For a sub-coalgebra \(S' ⊑ S\) and a state \(s ∈ S'\), \(s\) is live in \(S'\) if and only if \(s\) is live in \(S\).
-\end{corollary}
-
-\begin{proof}
+\end{theoremEnd}
+
+\begin{proofEnd}
     Let the homomorphism \(h\) in \cref{thm:hom-preserve-liveness} be the inclusion homomorphism \(i: S' → S\).
-\end{proof}
-
-\begin{corollary}[bisimulation preserves liveness]\label{thm:bisim-preserve-liveness}
+\end{proofEnd}
+
+\begin{theoremEnd}{corollary}[bisimulation preserves liveness]\label{thm:bisim-preserve-liveness}
     If there exists a bisimulation \(∼\) between GKAT coalgebra \(S\) and \(U\) s.t.
 \(s ∼ u\) for some states \(s ∈ S\) and \(u ∈ U\), then \(s\) and \(u\) has to be either both accepting, both live or both dead.
-\end{corollary}
-
-\begin{proof}
+\end{theoremEnd}
+
+\begin{proofEnd}
     Because for a \(∼\) is a bisimulation when both \(π₁: {∼} → S\) and \(π₂: {∼} → U\) are homomorphisms.
@@
     \end{align*}
-\end{proof}
+\end{proofEnd}
 
@@
 
-\begin{theorem}[Correctness~\cite{smolka_GuardedKleeneAlgebra_2020}]\label{thm:norm-bisim-correctness}
+\begin{theoremEnd}{theorem}[Correctness~\cite{smolka_GuardedKleeneAlgebra_2020}]\label{thm:norm-bisim-correctness}
     Given two states in two GKAT coalgebra \(s ∈ S\) and \(u ∈ U\), then there exists a bisimulation between normalized coalgebras \({∼} ⊆ \norm(S) × \norm(U)\) s.t.
 \(s ∼ u\) if and only if \(s\) and \(u\) are trace equivalent \(⟦s⟧_S = ⟦u⟧_U\)
-\end{theorem}
+\end{theoremEnd}
 
@@
 
-\begin{theorem}[normalization perserves liveness]\label{thm:norm-perserve-liveness}
+\begin{theoremEnd}{theorem}[normalization perserves liveness]\label{thm:norm-perserve-liveness}
     A state \(s ∈ S\) is dead (live) in \(S\) if and only if it is dead (live) in \(\norm(S)\).
-\end{theorem}
-
-\begin{proof}
+\end{theoremEnd}
+
+\begin{proofEnd}
     By definition, every state is either dead of live, thus we will only need to show the \(⟹\) direction, and
@@
     Hence, this walk also exists in \(\norm(S)\), and \(s\) is live in \(\norm(S)\).
-\end{proof}
+\end{proofEnd}
 
@@
 
-\begin{theorem}\label{thm:norm-functor}
+\begin{theoremEnd}{theorem}\label{thm:norm-functor}
     \(\norm\) is an endofunctor in the category GKAT coalgebra.
     More specifically, if \(h: S → U\) is a GKAT homomorphism, then \(h: \norm(S) → \norm(U)\) is also a homomorphism.
-\end{theorem}
-
-\begin{proof}
+\end{theoremEnd}
+
+\begin{proofEnd}
     Recall that \(h\) is a homomorphism if and only if for all \(s ∈ S\) and \(α ∈ \At\):
@@
     \end{align*}
-\end{proof}
-
-\begin{corollary}\label{thm:norm-sub-coalg}
+\end{proofEnd}
+
+\begin{theoremEnd}{corollary}\label{thm:norm-sub-coalg}
     Normalization preserves sub-coalgebra, i.e. if \(S' ⊑ S\) then \(\norm(S') ⊑ \norm(S)\).
-\end{corollary}
-
-\begin{proof}
+\end{theoremEnd}
+
+\begin{proofEnd}
     By letting the homomorphism in~\Cref{thm:norm-functor} to be the inclusion homomorphism \(i: S' → S\)
-\end{proof}
+\end{proofEnd}
 
@@
 
-\begin{corollary}\label{thm:inf-trace-equiv-implies-fin-trace-equiv}
+\begin{theoremEnd}{corollary}\label{thm:inf-trace-equiv-implies-fin-trace-equiv}
     Given two states in two GKAT coalgebra \(s ∈ S, u ∈ U\), \(⟦s⟧^{ω}_{S} = ⟦u⟧^{ω}_{U} ⟹ ⟦s⟧_{S} = ⟦u⟧_{U}\).
-\end{corollary}
-
-\begin{proof}
+\end{theoremEnd}
+
+\begin{proofEnd}
     Because \(⟦s⟧^{ω}_{S} = ⟦u⟧^{ω}_{U}\), there exists a bisimulation \({∼} ⊆ S × U\) s.t. \(s ∼ u\)~\cite{schmid_GuardedKleeneAlgebra_2021}.
@@
     Because \(s ∼ u\) and normalization operation preserves states in \(∼\), therefore \((s, u) ∈ \norm(∼)\), and because \(\norm(∼)\) is a bisimulation between the normalization of \(S\) and \(U\), therefore \(⟦s⟧_{S} = ⟦u⟧_{U}\) (\Cref{thm:norm-bisim-correctness}).
-\end{proof}
+\end{proofEnd}
 
@@
 
-\begin{lemma}\label{thm:dead-construction-greedy-bisim}
+\begin{theoremEnd}{lemma}\label{thm:dead-construction-greedy-bisim}
     Given any eager bisimulation \({≈} ⊆ S × U\) with the underlying bisimulation \({∼}\), the following relation \({∼_d} ⊆ S × U\):
@@
     Furthermore, \(\norm(∼_d)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\).
-\end{lemma}
-
-\begin{proof}
+\end{theoremEnd}
+
+\begin{proofEnd}
     We will need to show \(∼_d\) is closed under the transition of \(δ_{≈}\) in~\Cref{def:greedy-bisim}.
@@
     TODO: bisimulation are uniquely determined by its carrier.
-\end{proof}
+\end{proofEnd}
 
@@
 
-\begin{corollary}\label{thm:greedy-bisim-exists-unique}
+\begin{theoremEnd}{corollary}\label{thm:greedy-bisim-exists-unique}
     For any bisimulation between two normalized GKAT coalgebra \({∼} ⊆ \norm(S) × \norm(U)\), a greedy bisimulation \(≈\) with \(∼\) as underlying bisimulation exists and is unique.
-\end{corollary}
-
-\begin{proof}
+\end{theoremEnd}
+
+\begin{proofEnd}
     Because sub-coalgebra form a complete lattice under union and intersection~\cite[Theorem 6.4]{rutten_UniversalCoalgebraTheory_2000}, then we can construct \(≈\) by taking the intersection of all the sub-coalgebra of \(∼_d\) in~\Cref{thm:dead-construction-greedy-bisim}.
-\end{proof}
-
-\begin{corollary}\label{thm:greedy-bisim-dead-or-bisim}
+\end{proofEnd}
+
+\begin{theoremEnd}{corollary}\label{thm:greedy-bisim-dead-or-bisim}
     For every pair of states \((s, u)\) in a greedy bisimulation \({≈} ⊆ S × U\), either \((s, u)\) is in the underlying bisimulation \(∼\), or both \(s\) and \(u\) are dead in \(S\) and \(U\).
-\end{corollary}
-
-\begin{proof}
+\end{theoremEnd}
+
+\begin{proofEnd}
     Because \(∼_d\) is a GKAT coalgebra with transition \(δ_≈\) and also contains all the states in \(∼\), therefore \({≈} ⊑ {∼_d}\).
     Thus, for all \((s, u) ∈ {≈}\) but not in \(∼\), \(s\) and \(u\) must be dead in \(S\) and \(U\).
-\end{proof}
-
-\begin{corollary}[Perservation Of Liveness]\label{thm:greedy-bisim-perserve-liveness}
+\end{proofEnd}
+
+\begin{theoremEnd}{corollary}[Perservation Of Liveness]\label{thm:greedy-bisim-perserve-liveness}
     Given a pair of states in a greedy bisimulation \(s ≈ u\) where \({≈} ⊆ S × U\), then the following are equivalent
@@
     Furthermore, if \((s, u)\) is in the underlying bisimulation \(∼\) of \(≈\), then all of above are also equivalent to \((s, u)\) is live (or dead) in \(∼\).
-\end{corollary}
-
-\begin{proof}
+\end{theoremEnd}
+
+\begin{proofEnd}
     Let \(∼\) by the underlying bisimulation of \(≈\), consider the \(∼_d\) in~\Cref{thm:dead-construction-greedy-bisim}, then we know that \({≈} ⊑ {∼_d}\).
@@
     \end{align*}
-\end{proof}
-
-\begin{corollary}\label{thm:greedy-bisim-iff-norm-bisim}
+\end{proofEnd}
+
+\begin{theoremEnd}{corollary}\label{thm:greedy-bisim-iff-norm-bisim}
     Given two GKAT coalgebras \(S\) and \(U\), and a relational GKAT coalgebra \({≈} ⊆ S × U\) with the transition \(δ_≈\) as in~\cref{def:greedy-bisim}.
     \(≈\) is a greedy bisimulation if and only if \(\norm(≈)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\)
-\end{corollary}
-
-\begin{proof}
+\end{theoremEnd}
+
+\begin{proofEnd}
     The \(⟹\) direction, we let \(∼_d\) be defined as in~\Cref{thm:dead-construction-greedy-bisim}.
@@
     The \(⟸\) direction can be proven by simply letting \(\norm(≈)\) be the underlying bisimulation of \(≈\), because \(\norm(≈)\) has the same carrier set of \(≈\), then \(≈\) has to be the least GKAT coalgebra with \(δ_≈\) that contains \(\norm(≈)\), satisfying all the conditions for greedy bisimulation.
-\end{proof}
-
-\begin{corollary}\label{thm:max-bisim-is-max-greedy-bisim}
+\end{proofEnd}
+
+\begin{theoremEnd}{corollary}\label{thm:max-bisim-is-max-greedy-bisim}
     Given two GKAT coalgebra \(S\) and \(U\), carrier of the maximal bisimulation \({≡} ⊆ \norm(S) × \norm(U)\) is also the carrier of maximal greedy bisimulation, which we will denote as \(≊ ⊆ S × U\).
     We will abuse the notation \(≡\) to represent both the maximal bisimulation and maximal greedy bisimulation.
-\end{corollary}
-
-\begin{proof}
+\end{theoremEnd}
+
+\begin{proofEnd}
     by definition of greedy bisimulation, the carrier of \({≡}\) is a subset of the carrier of \({≊}\). 
@@
     By~\cref{thm:greedy-bisim-dead-or-bisim}, we can case analysis on \(s\) and \(u\): if \(s ∼ u\), then \(s ≡ u\) because \(≡\) is the maximal bisimulation; and if \(s\) and \(u\) are both dead, then by~\cref{thm:bisim-between-dead}, \(s ≡ u\).
-\end{proof}
+\end{proofEnd}
 
@@
 
-\begin{theorem}[Correctness]
+\begin{theoremEnd}{theorem}[Correctness]
     Given two states in GKAT coalgebras \(s ∈ S\) and \(u ∈ ,U\), there exists a greedy bisimulation \({≈} ⊆ S × U\) s.t. \(s ≈ u\), if and only if \(⟦s⟧_S = ⟦u⟧_U\).
-\end{theorem}
-
-\begin{proof}
+\end{theoremEnd}
+
+\begin{proofEnd}
     \(⟹\) direction, because normalization preserves carrier set, therefore \((s, u) ∈ \norm(≈)\).
@@
     Because the carriers \({∼} ⊆ {≈}\), therefore \(s ∼ u\) implies \(s ≈ u\).
-\end{proof}
+\end{proofEnd}
 
@@
 
-\begin{theorem}[Recursive Construction]\label{thm:recursive-construction}
+\begin{theoremEnd}{theorem}[Recursive Construction]\label{thm:recursive-construction}
     Given a relation \({≈} ⊆ S × U\), \(≈\) is a carrier of some greedy bisimulation if and only if all the following condition are satisfied: for all pair of states \(s ≈ u\),
@@
     \end{enumerate}
-\end{theorem}
-
-\begin{proof}
+\end{theoremEnd}
+
+\begin{proofEnd}
     To show the \(⟹\) direction, assuming there exists a greedy bisimulation \(≈\), then it would satisfy all the conditions. 
@@
     \end{itemize}
-\end{proof}
-
-\begin{lemma}\label{thm:bisim-between-dead}
+\end{proofEnd}
+
+\begin{theoremEnd}{lemma}\label{thm:bisim-between-dead}
     Given any two dead states \(s ∈ S\) and \(u ∈ U\), we can construct a bisimulation \(∼\) between \(\norm(S)\) and \(\norm(U)\):
@@
     Therefore, for the maximal bisimulation \({≡} ⊆ \norm(S) × \norm(U)\), \(s ≡ u\).
-\end{lemma}
-
-\begin{theorem}[Equivalence Relation]\label{thm:greedy-bisim-iff-greedy-bisim-equiv}
+\end{theoremEnd}
+
+\begin{theoremEnd}{theorem}[Equivalence Relation]\label{thm:greedy-bisim-iff-greedy-bisim-equiv}
     Given two states in GKAT coalgebra \(s ∈ S\) and \(u ∈ U\), there exists a greedy bisimulation \(≈\) between \(S\) and \(U\) s.t.
@@
 \(s ≃ u\).
-\end{theorem}
-
-\begin{proof}
+\end{theoremEnd}
+
+\begin{proofEnd}
     First, recall that by~\Cref{thm:greedy-bisim-iff-norm-bisim} \(\norm(≈)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\).
@@
     Therefore \((s, u)\) is also in the greedy bisimulation \({≈} ⊆ \norm(S) × \norm(U)\), generated by \(∼\).
-\end{proof}
-
-% \begin{theorem}\label{thm:greedy-bisim-one-dead}
+\end{proofEnd}
+
+% \begin{theoremEnd}{theorem}\label{thm:greedy-bisim-one-dead}
 %     Consider two states in GKAT coalgebras \(s ∈ S, u ∈ U\) and an arbitrary bisimulation \({∼} ⊆ \norm(S) × \norm(U)\), if \(s\) is dead, then 
@@
 %     \end{enumerate}
-% \end{theorem}
-
-% \begin{proof}
+% \end{theoremEnd}
+
+% \begin{proofEnd}
 %     \Cref{itm:greedy-bisim-one-dead-then-both-dead} is a consequence of bisimulation preserves liveness (\Cref{thm:bisim-preserve-liveness}) and normalization preserves liveness (\Cref{thm:norm-perserve-liveness}).
@@
 %     Finally because \(∼\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\), then so is \({∼} ∖ ⟨(s, u)⟩_∼\) (\Cref{thm:})
-% \end{proof}
+% \end{proofEnd}
 
@@
 
-\begin{theorem}[Functoriality]\label{thm:lowering-functor}
+\begin{theoremEnd}{theorem}[Functoriality]\label{thm:lowering-functor}
     The lowering operation is a functor, every symbolic GKAT coalgebra homomorphism \(h: Ŝ → Û\), is also a GKAT coalgebra homomorphism \(h: S → U\).
-\end{theorem}
-
-\begin{proof}
+\end{theoremEnd}
+
+\begin{proofEnd}
     Since \(Ŝ\) and \(Û\) have the same states as their lowering, therefore \(h: S → U\) is indeed a function, then we only need to verify the homomorphism condition on \(h\).
     TODO: finish.
-\end{proof}
+\end{proofEnd}
 
@@
 
-\begin{theorem}[Symbolic Recursive Construction]\label{thm:symb-recursive-construction}
+\begin{theoremEnd}{theorem}[Symbolic Recursive Construction]\label{thm:symb-recursive-construction}
     Given two symbolic GKAT coalgebra \(Ŝ = (S, ϵ̂_S, δ̂_S)\) and \(Û = (U, ϵ̂_U, δ̂_U)\) and two states \(s ∈ S\) and \(u ∈ U\), \(s\) and \(u\) are trace equivalent \(⟦s⟧_{Ŝ} = ⟦u⟧_{Û}\), if and only if all the following holds:
@@
     \end{itemize}
-\end{theorem}
-
-\begin{proof}
+\end{theoremEnd}
+
+\begin{proofEnd}
     Reduces to~\Cref{thm:recursive-construction} i.e. all the above condition holds if and only if all the condition in~\Cref{thm:recursive-construction} holds in the lowering.
-\end{proof}
+\end{proofEnd}
 
@@
 This fact can be proven by case analysis on the shape of the source expression, and forms a basis on our correctness argument.
-\begin{theorem}[Correctness]\label{thm:derivative-correctness}
+\begin{theoremEnd}{theorem}[Correctness]\label{thm:derivative-correctness}
     The lowering of \(D̂\), denoted \(D\), is exactly the derivative defined by~\citeauthor{schmid_GuardedKleeneAlgebra_2021}~\cite{schmid_GuardedKleeneAlgebra_2021}.
     Therefore, the semantics of the expression is equal to the semantics generated by the derivative coalgebra, \(⟦e⟧ = ⟦e⟧_{D} = ⟦e⟧_{D̂}\).
-\end{theorem}
-\begin{proof}
+\end{theoremEnd}
+\begin{proofEnd}
     TODO: unfold the statement.
-\end{proof}
+\end{proofEnd}
 
@@
 
-\begin{theorem}\label{thm:hom-thompson-derivative}
+\begin{theoremEnd}{theorem}\label{thm:hom-thompson-derivative}
     Given any GKAT expression \(e\), the resulting symbolic GKAT coalgebra from Thompson's construction \(Ŝ_e\) have a homomorphism to derivatives \(h: Ŝ_e → D̂\), s.t.
 for the start state \(s^* ∈ S, h(s^*) = e\).
-\end{theorem}
-
-\begin{proof}
+\end{theoremEnd}
+
+\begin{proofEnd}
     By induction on the structure of \(e\).
@@
 
-\begin{corollary}[Correctness]
+\begin{theoremEnd}{corollary}[Correctness]
     Given any expression \(e\) and its Thompson's coalgebra \(Ŝ_{e}\) with the start state \(s^* ∈ Ŝ_{e}\), then the semantics of the start state is equivalent to the semantics of \(e\): \(⟦s^*⟧_{Ŝ_{e}} = ⟦e⟧.\)
-\end{corollary}
-
-\begin{proof}
+\end{theoremEnd}
+
+\begin{proofEnd}
     By functoriality of lowering~\cref{thm:lowering-functor}, then \(h: S_{e} → D\) is a homomorphism on their lowerings. 
@@
     \(⟦s^*⟧_{S_{e}} = ⟦e⟧_{D} = ⟦e⟧\).
-\end{proof}
+\end{proofEnd}
 
@@
 
-\begin{corollary}\label{thm:suj-hom-thompson-derivative}
+\begin{theoremEnd}{corollary}\label{thm:suj-hom-thompson-derivative}
     There exists a surjective homomorphism \(h': ⟨s^*⟩_{Ŝ_{e}} → ⟨e⟩_{D̂}\). 
     Because the size of \(⟨s^*⟩_{Ŝ_{e}}\) is linear to \(e\), the size of \(⟨e⟩_{D̂}\) is at most linear to the size of expression \(e\).
-\end{corollary}
-
-\begin{proof}
+\end{theoremEnd}
+
+\begin{proofEnd}
     We define \(h'\) to be point-wise equal to \(h\), i.e. \(h'(s) ≜ h(s)\), i.e. \(h'\) is \(h\) restricted on the domain \(⟨s^*⟩_{Ŝ_{e}}\). 
@@
     In other words, the image of \(h\) on \(⟨s^*⟩_{Ŝ_{e}}\) is equal to \(⟨e⟩_{D̂}\); thus, because \(h'\) the restriction of \(h\) on \(⟨s^*⟩_{Ŝ_{e}}\), the range of \(h'\) contains its codomain \(⟨e⟩_{D̂}\), showing that \(h'\) is surjective.
-\end{proof}
+\end{proofEnd}
 
@@
 
+\section{Detailed Proof}
+\printProofs
+
 \begin{onecolumn}

### 2025-01-20 main.tex
update evaluation

@@
     \hline
-    e250b5p10ne    &   0.19 &   0.18 & 0.17 & 0.21 &    5.82 \\
-    e250b5p10eq    &   0.21 &   0.18 & 0.24 & 0.34 &    2.83 \\
-    e500b5p50ne    &   0.23 &   0.22 & 0.25 & 0.28 &   37.28 \\
-    e500b5p50eq    &   0.26 &   0.21 & 0.24 & 0.34 &   14.06 \\
-    e1000b10p100ne &   0.31 &   0.37 & 0.42 & 0.52 & timeout \\
-    e1000b10p100eq &   0.37 &   0.28 & 0.35 & 0.52 &   77.83 \\
-    e2000b20p200ne &   1.50 &   2.75 & 0.84 & 1.11 & timeout \\
-    e2000b20p200eq &   3.34 &   3.03 & 0.51 & 1.30 & timeout \\
-    e3000b30p200ne &   1.75 &  26.38 & 1.59 & 2.29 & timeout \\
-    e3000b30p200eq &  17.86 &  22.59 & 0.81 & 2.02 & timeout \\
-    degenerate     & 220.53 & 232.63 & 0.28 & 0.37 & timeout
+    e250b5p10ne    & 0.18   & 0.18   & 0.16 & 0.17 &    5.82 \\
+    e250b5p10eq    & 0.19   & 0.18   & 0.15 & 0.18 &    2.83 \\
+    e500b5p50ne    & 0.21   & 0.21   & 0.21 & 0.22 &   37.28 \\
+    e500b5p50eq    & 0.22   & 0.21   & 0.20 & 0.27 &   14.06 \\
+    e1000b10p100ne & 0.26   & 0.28   & 0.34 & 0.38 & timeout \\
+    e1000b10p100eq & 0.28   & 0.26   & 0.28 & 0.41 &   77.83 \\
+    e2000b20p200ne & 1.32   & 2.60   & 0.68 & 0.79 & timeout \\
+    e2000b20p200eq & 1.92   & 2.95   & 0.41 & 1.00 & timeout \\
+    e3000b30p200ne & 1.57   & 24.86  & 1.30 & 1.60 & timeout \\
+    e3000b30p200eq & 10.82  & 22.83  & 0.66 & 1.54 & timeout \\
+    degenerate     & 99.48  & 228.16 & 0.24 & 0.30 & timeout
 \end{tabular}
@@
     \hline
-    e250b5p10ne    &  15.36 &   14.76 &  6.81 &  7.02 &  114.06 \\
-    e250b5p10eq    &  15.56 &   14.95 &  7.06 &  7.25 &  100.48 \\
-    e500b5p50ne    &  16.26 &   15.49 &  7.02 &  6.99 &  524.89 \\
-    e500b5p50eq    &  17.97 &   15.54 &  7.07 &  7.01 &  546.91 \\
-    e1000b10p100ne &  18.36 &   21.41 &  8.43 &  8.07 & timeout \\
-    e1000b10p100eq &  20.41 &   17.66 &  8.69 &  7.63 & 5822.66 \\
-    e2000b20p200ne & 239.71 &  283.45 & 13.54 & 12.27 & timeout \\
-    e2000b20p200eq & 107.92 &  102.46 & 13.20 & 13.84 & timeout \\
-    e3000b30p200ne & 112.59 & 1235.46 & 21.69 & 20.33 & timeout \\
-    e3000b30p200eq & 245.92 &  228.85 & 21.79 & 18.33 & timeout \\
-    degenerate     & 631.47 & 1229.30 & 19.75 & 18.57 & timeout
+    e250b5p10ne    & 15.17  & 14.54   & 7.07  & 7.01  &  114.06 \\
+    e250b5p10eq    & 15.69  & 14.69   & 7.06  & 7.29  &  100.48 \\
+    e500b5p50ne    & 15.99  & 15.33   & 7.02  & 7.29  &  524.89 \\
+    e500b5p50eq    & 16.85  & 15.26   & 6.99  & 7.01  &  546.91 \\
+    e1000b10p100ne & 17.51  & 21.31   & 7.78  & 7.50  & timeout \\
+    e1000b10p100eq & 19.52  & 17.48   & 8.17  & 7.04  & 5822.66 \\
+    e2000b20p200ne & 237.17 & 280.70  & 12.95 & 11.43 & timeout \\
+    e2000b20p200eq & 107.61 & 102.18  & 12.76 & 13.25 & timeout \\
+    e3000b30p200ne & 112.06 & 1232.24 & 21.00 & 19.30 & timeout \\
+    e3000b30p200eq & 245.23 & 228.86  & 21.18 & 17.64 & timeout \\
+    degenerate     & 632.70 & 1225.24 & 19.13 & 17.94 & timeout
 \end{tabular}
 \caption{Peak Memory Usage (megabytes)}\label{tab:benchmark-memory}
+\vspace{-1em}
 \end{table}
@@
 
-We observe that all variants of DV and TC perform better than SymKAT in terms of time and memory usage.
-When comparing the solver backends, BDD is marginally faster than SAT for smaller benchmarks, however, BDD's time and memory usage increases considerably for larger benchmarks.
+We observe that DV performs better than TC consistently when both algorithms use the same solver backend and all variants of DV and TC perform better than SymKAT.
+When comparing the solver backends, BDD and SAT perform similarly for smaller benchmarks, however, BDD's time and memory usage increase considerably for larger benchmarks.
 On the other hand, the performance of SAT is stable for all benchmarks.

### 2025-01-21 main.tex
first intro draft

@@
 To utilize the symbolic bisimulation checking, we also provide two methods to generate symbolic GKAT coalgebras from GKAT expressions, analogous to derivative and Thompson's construction in previous works.
-We have proven the correctness of our algorithms and conducted detailed complexity analyses.
-Finally, we implemented the symbolic algorithms in Rust, demonstrating significant performance improvements over state-of-the-art tools.
+We have proven the correctness of our algorithms and conducted brief complexity analyses.
+Finally, we implemented the symbolic algorithms in Rust, demonstrating orders-of-magnitude performance improvements over existing tools.
 \end{abstract}
@@
 
-\paragraph{Notation:} In this paper, we will use un-curried notation to apply curried functions, for example, given a function \(δ: X → Y → Z\), we will write the function applications as \(δ(x): Y → Z\) and \(δ(x, y): Z\).
-
-\section{Background on Coalgebra and GKAT}
+Kleene Algebra with Tests (KAT) is a staple algebra system to reason about program equivalence, and have found numerous applications in numerous fields, including software-defined networks~\cite{anderson_NetKATSemanticFoundations_2014a,moeller_KATchFastSymbolic_2024,smolka_ScalableVerificationProbabilistic_2019,smolka_CantorMeetsScott_2017}, program transformations~\cite{kozen_BohmJacopiniTheorem_2008,angus_KleeneAlgebraTests_2001,kozen_NonlocalFlowControl_2008,grathwohl_KAT_2014,kozen_KleeneAlgebraTests_1996}, and program logics~\cite{antonopoulos_AlgebraAlignmentRelational_2023,zhang_DomainReasoningTopKAT_2024,zhang_IncorrectnessLogicKleene_2022c,kozen_HoareLogicKleene_2000,desharnais_ModalKleeneAlgebra_2004}.
+Beyond its massive theoretical achievement, Kleene Algebra with Tests has also inspired numerous practical tools for several domains.
+Interestingly, despite the PSPACE-complete complexity of KAT, many tools developed are able to match, or even exceed, the performance of state-of-the-art tools in their respective domains~\cite{moeller_KATchFastSymbolic_2024,smolka_ScalableVerificationProbabilistic_2019}.
+
+This surprising fact lead to the development of Guarded Kleene Algebra with Tests (GKAT)~\cite{smolka_GuardedKleeneAlgebra_2020}, a (co)algebraic system that seeks to understand this efficient fragment of KAT.
+Indeed, the determinism of GKAT gives it a different coalgebra structure as KAT, and allows for GKAT expression to directly be converted to a linear-sized coalgebra, as opposed to the exponential sized coalgebra produced KAT, in general~\cite{smolka_GuardedKleeneAlgebra_2020,kozen_CoalgebraicTheoryKleene_2017}.
+
+However, the desirable theoretical properties of GKAT does not currently translate to tangible real-world performance. 
+surprisingly, the decision procedures for GKAT can often be slower than that of KAT, primarily due to the following two reasons.
+
+First, the natural semantics of GKAT coalgebra is not the finite-trace model, but the infinite-trace model~\cite{smolka_GuardedKleeneAlgebra_2020,schmid_GuardedKleeneAlgebra_2021}.
+Thus, to determine finite-trace equivalence, it is necessary to \emph{normalize} the coalgebra first, requiring to compute the entire coalgebra ahead-of-time; and iterate through its entirety.
+Without the need to normalize prior to computing bisimulation, the decision procedure of KAT equivalence can terminate as soon as a counter-example has been found, sometimes leading to better performance than the original GKAT decision algorithm~\cite{smolka_GuardedKleeneAlgebra_2020}.
+
+The lack of symbolic algorithm also makes GKAT a less desirable foundation for performance-critical applications.
+Although the generated coalgebra from GKAT expression have linear number of states, the number of transitions is exponential to \emph{all} the primitive tests appeared across the entire expression.
+This inefficiency have been noted in several pervious literatures, and poses real limitation on real-world systems~\cite{smolka_GuardedKleeneAlgebra_2020,zhang_CFGKATEfficientValidation_2025}.
+On the other hand, several symbolic algorithm for KAT have been proven to be effective for generic equality solving and domain-specific applications~\cite{pous_SymbolicAlgorithmsLanguage_2015,moeller_KATchFastSymbolic_2024}.
+
+In this paper, we present two algorithms that address both shortcomings, and demonstrate how coalgebra theory of GKAT serves as the bedrock of the design and correctness of these algorithms.
+
+The first algorithm supports on-the-fly coalgebra generation by performing bisimulation in a greedy manner.
+Specifically, the liveness checking of states only kicks in when bisimulation algorithm encountered a disparity.
+We support the correctness of this algorithm by designing a coalgebraic concept which we named \emph{greedy bisimulation}.
+
+The second algorithm builds lifts the first into a symbolic setting, where our development is based on a extension of GKAT coalgebra, where we named \emph{symbolic GKAT coalgebra}.
+In contrast to the symbolic KAT derivative~\cite{pous_SymbolicAlgorithmsLanguage_2015}, the deterministic nature of GKAT coalgebra allows its symbolic variant to transition via boolean expressions, instead of primitive tests.
+This optimization not only allows more compact coalgebra representation, but also allows us to use arbitrary of-the-shelf SAT solvers to resolve boolean constraint, instead of the binary decision diagram presented by the symbolic algorithm of KAT~\cite{pous_SymbolicAlgorithmsLanguage_2015}.
+Finally, we lifted both the derivative construction~\cite{schmid_GuardedKleeneAlgebra_2021} and Thompson's construction~\cite{smolka_GuardedKleeneAlgebra_2020} to symbolic setting, and proved their correspondence, correctness, and finiteness.
+
+To evaluate the practical impact of our work, we have implemented our algorithm in rust. 
+Our implementation performs orders of magnitude better than existing implementation for KAT~\cite{pous_SymbolicAlgorithmsLanguage_2015},
+proving the theoretical elegance of GKAT indeed leads to significant real-world performance differences.
+
+We list our contributions and structures of this paper as follows:
+\begin{itemize}
+    \item In~\Cref{sec:background}, we recap the necessary concepts in universal coalgebra, and the coalgebraic theory of GKAT.
+In particular, we do not define the semantics of GKAT either algebraically or coalgebraically, as we can avoid directly interacting with the semantics by leveraging the coalgebraic theory of GKAT.
+    \item In~\Cref{sec:greedy-bisim}, we define \emph{greedy bisimulation} and developed its coalgebraic theory to serve as the ground for the correctness of our algorithm.
+    We also performs a brief complexity analysis, and argue that our algorithm is almost always more efficient than the original normalize and then bisimulate approach~\cite{smolka_GuardedKleeneAlgebra_2020}.
+    \item In~\Cref{sec:symb-gkat-greedy-bisim}, we introduced \emph{symbolic GKAT coalgebra}, and lifted the greedy bisimulation algorithm to the symbolic setting.
+    \item In~\Cref{sec:symb-gkat-construction}, we lifts the Thompson's construction and derivative construction to the symbolic setting.
+    Then we established their connection via a coalgebra homomorphism, which then allows us to derive the correctness and finiteness of one construction from the other.
+    \item Finally, in~\Cref{sec:implementation}, we described our implementation of symbolic derivable, symbolic Thompson's construction, and symbolic greedy bisimulation.
+    Then we conduct a performance analysis using different SAT solvers, and compared to previous implementations.
+    When using minisat as the SAT solver of choice~\cite{}, our implementation can solve most test cases in less than a second with minimal memory usage, whereas previous implementation~\cite{pous_SymbolicAlgorithmsLanguage_2015} almost always result in a timeout.
+\end{itemize}
+
+\section{Background on Coalgebra and GKAT}\label{sec:background}
 
@@
 
-\section{On-The-Fly Bisimulation}
+\section{On-The-Fly Bisimulation}\label{sec:greedy-bisim}
 
@@
 
-\section{Symbolic Coalgebra and Algorithm}
+\section{Symbolic Coalgebra and Algorithm}\label{sec:symb-gkat-greedy-bisim}
 

### 2025-01-22 main.tex
some fixes to the intro

--- 2025-01-20 main.tex
@@
 
-\paragraph{Notation:} In this paper, we will use un-curried notation to apply curried functions, for example, given a function \(δ: X → Y → Z\), we will write the function applications as \(δ(x): Y → Z\) and \(δ(x, y): Z\).
-
-\section{Background on Coalgebra and GKAT}
+Kleene Algebra with Tests (KAT) is a staple algebra system to reason about program equivalence, and have found applications in numerous fields, including software-defined networks~\cite{anderson_NetKATSemanticFoundations_2014a,moeller_KATchFastSymbolic_2024,smolka_ScalableVerificationProbabilistic_2019,smolka_CantorMeetsScott_2017}, program transformations~\cite{kozen_BohmJacopiniTheorem_2008,angus_KleeneAlgebraTests_2001,kozen_NonlocalFlowControl_2008,grathwohl_KAT_2014,kozen_KleeneAlgebraTests_1996}, and program logics~\cite{antonopoulos_AlgebraAlignmentRelational_2023,zhang_DomainReasoningTopKAT_2024,zhang_IncorrectnessLogicKleene_2022c,kozen_HoareLogicKleene_2000,desharnais_ModalKleeneAlgebra_2004}.
+Beyond its theoretical achievements, Kleene Algebra with Tests has also inspired several practical tools for specialized domains.
+Interestingly, despite the PSPACE-complete complexity of KAT, many tools developed are able to match, or even exceed, the performance of specialized state-of-the-art tools in their respective domains~\cite{moeller_KATchFastSymbolic_2024,smolka_ScalableVerificationProbabilistic_2019}.
+
+This surprising fact lead to the development of Guarded Kleene Algebra with Tests (GKAT)~\cite{smolka_GuardedKleeneAlgebra_2020}, a (co)algebraic system that seeks to encapsulate this efficient fragment of KAT.
+Indeed, the determinism of GKAT gives it a different coalgebra structure from KAT, and allows for a GKAT expression to be directly converted to a linear-sized coalgebra, as opposed to the exponential sized coalgebra produced KAT in general~\cite{smolka_GuardedKleeneAlgebra_2020,kozen_CoalgebraicTheoryKleene_2017}.
+
+However, as it current stands, the desirable theoretical properties of GKAT do not translate to tangible real-world performance. 
+The decision procedures for GKAT can often be slower than that of KAT, primarily due to the following two reasons.
+
+First, the natural semantics of GKAT coalgebra is not the finite-trace model, but the infinite-trace model~\cite{smolka_GuardedKleeneAlgebra_2020,schmid_GuardedKleeneAlgebra_2021}.
+Thus, to determine finite-trace equivalence, it is necessary to \emph{normalize} the coalgebra first, requiring one to compute the entire coalgebra ahead-of-time; and iterate through its entirety.
+Without the need to normalize prior to computing bisimulation, the decision procedure of KAT equivalence can terminate as soon as a counter-example has been found which can lead to better performance than the original GKAT decision algorithm~\cite{smolka_GuardedKleeneAlgebra_2020}.
+
+The lack of symbolic algorithms also makes GKAT a less desirable foundation for performance-critical applications.
+Although the generated coalgebra from GKAT expression have a linear number of states, the number of transitions is exponential to \emph{all} the primitive tests appearing in the entire expression.
+This inefficiency is noted in previous literature, and poses a real limitation on real-world systems~\cite{smolka_GuardedKleeneAlgebra_2020,zhang_CFGKATEfficientValidation_2025}.
+On the other hand, symbolic algorithms for KAT have been proven to be effective for generic equality solving and domain-specific applications~\cite{pous_SymbolicAlgorithmsLanguage_2015,moeller_KATchFastSymbolic_2024}.
+
+In this paper, we present two novel algorithms that address both shortcomings, and demonstrate how the coalgebra theory of GKAT serves as the foundation for the design and correctness of these algorithms.
+
+The first algorithm supports on-the-fly coalgebra generation by performing bisimulation in a greedy manner.
+Specifically, the liveness checking of states only kicks in when the bisimulation algorithm encounters a disparity.
+We show the correctness of this algorithm through a novel coalgebraic concept which we named \emph{greedy bisimulation}.
+
+The second algorithm lifts the first into a symbolic setting, where our development is based on an extension of GKAT coalgebra, which we named \emph{symbolic GKAT coalgebra}.
+In contrast to the symbolic KAT derivative~\cite{pous_SymbolicAlgorithmsLanguage_2015}, the deterministic nature of GKAT coalgebra allows its symbolic variant to transition via boolean expressions instead of primitive tests.
+This optimization not only allows more compact coalgebra representation, but also allows the usage of arbitrary off-the-shelf SAT solvers to resolve boolean constraints, instead of being limited to binary decision diagrams as utilized by the symbolic algorithm of KAT~\cite{pous_SymbolicAlgorithmsLanguage_2015}.
+Finally, we lift both the derivative construction~\cite{schmid_GuardedKleeneAlgebra_2021} and Thompson's construction~\cite{smolka_GuardedKleeneAlgebra_2020} to the symbolic setting, and prove their correspondence, correctness, and finiteness.
+
+To evaluate the practical impact of our work, we implement our algorithms in Rust. We show through experiments that
+our implementation performs orders of magnitude better than existing implementation for KAT~\cite{pous_SymbolicAlgorithmsLanguage_2015},
+confirming that the theoretical elegance of GKAT indeed leads to significant real-world performance improvements.
+
+We list our contributions and structures of this paper as follows:
+\begin{itemize}
+    \item In~\Cref{sec:background}, we recap the necessary concepts in universal coalgebra, and the coalgebraic theory of GKAT.
+In particular, we do not define the semantics of GKAT either algebraically or coalgebraically, as we can avoid directly interacting with the semantics by leveraging the coalgebraic theory of GKAT.
+    \item In~\Cref{sec:greedy-bisim}, we define \emph{greedy bisimulation} and developed its coalgebraic theory to serve as the ground for the correctness of our algorithm.
+    We also performs a brief complexity analysis, and argue that our algorithm is almost always more efficient than the original normalize-bisimulate approach~\cite{smolka_GuardedKleeneAlgebra_2020}.
+    \item In~\Cref{sec:symb-gkat-greedy-bisim}, we introduce \emph{symbolic GKAT coalgebra}, and lift the greedy bisimulation algorithm to the symbolic setting.
+    \item In~\Cref{sec:symb-gkat-construction}, we lift the Thompson's construction and derivative construction to the symbolic setting.
+    Then we establish their connection via a coalgebra homomorphism which allows us to derive the correctness and finiteness of one construction from the other.
+    \item Finally, in~\Cref{sec:implementation}, we describe our implementation of symbolic derivative, symbolic Thompson's construction, and symbolic greedy bisimulation.
+    We analyze algorithm performance under different satisfiability solver backends, and compare our work to previous implementations.
+    When using MiniSat~\cite{een_MINISAT_2004} as the backend, our implementation can solve most test cases in less than a second with minimal memory usage, whereas previous implementation~\cite{pous_SymbolicAlgorithmsLanguage_2015} almost always result in a timeout.
+\end{itemize}
+
+\section{Background on Coalgebra and GKAT}\label{sec:background}
 

### 2025-01-22 main.tex
fix unicode

@@
 \and
-\IEEEauthorblockN{Hang Ji}
+\IEEEauthorblockN{Hang Ji\textsuperscript{\textsection}}
 \IEEEauthorblockA{\textit{Department of Computer Science} \\
-\textit{Boston University}\\
-Boston, USA \\
-email address or ORCID}
+\textit{Cornell University}\\
+Ithaca, USA \\
+hj476@cornell.edu}
 \and
@@
 Boston, USA \\
-email address or ORCID}
+inessdv15@gmail.com}
 }
@@
 
-The lack of symbolic algorithms also makes GKAT a less desirable foundation for performance-critical applications.
+Secondly, the lack of symbolic algorithms makes GKAT a less desirable foundation for performance-critical applications.
 Although the generated coalgebra from GKAT expression have a linear number of states, the number of transitions is exponential to \emph{all} the primitive tests appearing in the entire expression.
@@
 Intuitively, we usually think of principle sub-coalgebra \(⟨s⟩_S\) as the sub-coalgebra that is formed by all the ``reachable state'' from \(s\).
-This coalgebraic characterization of reachable state can allow us to avoid induction on the length of path from \(s\) to a state in \(⟨s⟩_S\).
+This coalgebraic characterization of reachable state can allow us to avoid induction on the length of a path from \(s\) to a state in \(⟨s⟩_S\).
 
@@
 \end{align*}
-Then the semantics of each expression \(⟦e⟧\) can be computed by the semantics of Kleene Algebra with tests~\cite{kozen_KleeneAlgebraTests_1997c}.
+Then the semantics of each expression \(⟦e⟧\) can be computed by the semantics of Kleene Algebra with Tests~\cite{kozen_KleeneAlgebraTests_1997c}.
 An important construct in the semantics is \emph{atoms}, which are conjunctions of all the primitive tests either in its positive or negative form: for \(T ≜ \{t₁, t₂, …, tₙ\}\)
 \[\At_T ≜ \{t₁' ∧ t₂' ∧ ⋯ ∧ tₙ' ∣ tᵢ' ∈ \{tᵢ, \overline{tᵢ}\}\}.\]
-Follow the conventional notation, we denote an atom using \(α, β\); and we sometimes omit the subscript \(T\) when no confusing can arise.
+Following the conventional notation, we denote an atom using \(α, β\); and we sometimes omit the subscript \(T\) when no confusion can arise.
 Alternatively, atoms can also be thought of as truth assignments to each primitive tests, indicating which primitive tests is satisfied in the current program states; and \(α ≤ b\) if and only if the truth assignment represented by \(α\) satisfies \(b\).
 
-For the sake of brevity, we omit the complete definition of GKAT and KAT semantics, we refer the reader to previous works~\cite{smolka_GuardedKleeneAlgebra_2020,schmid_GuardedKleeneAlgebra_2021,kozen_KleeneAlgebraTests_1997c}, which explains these semantics in detail.
+For the sake of brevity, we omit the complete definition of GKAT and KAT semantics, we refer the reader to previous works~\cite{smolka_GuardedKleeneAlgebra_2020,schmid_GuardedKleeneAlgebra_2021,kozen_KleeneAlgebraTests_1997c}, which explain these semantics in detail.
 Our work avoids direct interaction with the semantics by leveraging prior results in the coalgebraic theory of GKAT, which we will recap below.
@@
 
-Besides nice property with bisimulation, GKAT coalgebra is also deterministic, unlike Kleene coalgebra with tests (KCT)~\cite{kozen_CoalgebraicTheoryKleene_2017}.
+Besides nice property with bisimulation, GKAT coalgebra is also deterministic, unlike Kleene Coalgebra with Tests (KCT)~\cite{kozen_CoalgebraicTheoryKleene_2017}.
 For each atom, states in a KCT can accept or reject the atom (but not both), yet states can also non-deterministically transition to multiple different states via the same atom, while executing different actions.
 As we will see later, the deterministic behavior of GKAT coalgebra not only enables a more versatile symbolic algorithm than KCT~\cite{pous_SymbolicAlgorithmsLanguage_2015}, but also present challenges.
-Specifically, GKAT coalgebra requires normalization to compute finite trace equivalences~\cite{smolka_GuardedKleeneAlgebra_2020}, where we reroute all the transitions that cannot lead to acceptance immediately into rejection; and states that can never lead to acceptance is called \emph{dead states}.
+Specifically, GKAT coalgebra requires normalization to compute finite trace equivalences~\cite{smolka_GuardedKleeneAlgebra_2020}, where we reroute all the transitions that cannot lead to acceptance immediately into rejection; and states that can never lead to acceptance are called \emph{dead states}.
 
 In previous works, these dead states are detected after the necessary coalgebra is computed and stored.
-This approach requires storing all the transition and states of coalgebra in memory, meaning that the algorithm does not short circuit even if a mismatch can be detected in the starting state; for example, when for some \(α ∈ \At\), \(δ_S(s, α) = \reject\) and \(δ_U(u, α) = \accept\).
-
-Our algorithm is lazy in the dead state detection i.e.
-the dead state are only checked when a mismatch requires it; even then, our dead state detection algorithm will only check the necessary states to compute the liveness of the states causing the mismatch.
-This not only avoids unnecessarily checking the liveness of a state, but also enables allows on-the-fly generation of the coalgebra, like using derivatives, where we can remove irrelevant states from memory after it has been explored.
-
-However, to truly understand our on-the-fly algorithms, we will first need to define ``dead states'', and its role in defining the coalgebraic semantic of GKAT.
+This approach requires storing all the transitions and states of coalgebra in memory, meaning that the algorithm does not short circuit even if a mismatch can be detected in the starting state; for example, when for some \(α ∈ \At\), \(δ_S(s, α) = \reject\) and \(δ_U(u, α) = \accept\).
+
+Our algorithm is lazy in the dead states detection i.e.
+the dead states are only checked when a mismatch requires it; even then, our dead state detection algorithm will only check the necessary states to compute the liveness of the states causing the mismatch.
+This not only avoids unnecessarily checking the liveness of a state, but also enables on-the-fly generation of the coalgebra, like using derivatives, where we can remove irrelevant states from memory after they have been explored.
+
+However, to truly understand our on-the-fly algorithms, we will first need to define ``dead states'', and their roles in defining the coalgebraic semantic of GKAT.
 
@@
 
-Traditionally, live and dead states are defined by whether they can reach an accepting state~\cite{smolka_GuardedKleeneAlgebra_2020}.
+Traditionally, live and dead states are defined by whether they can reach to an accepting state~\cite{smolka_GuardedKleeneAlgebra_2020}.
 \begin{definition}[Liveness]\label{def:liveness-of-states}
@@
 
-Infinite-trace model \(𝒢_ω\) is the final coalgebra of GKAT coalgebras~\cite{schmid_GuardedKleeneAlgebra_2021}.
-The finality of the model allows us to define the semantics of each state in a given GKAT coalgebra \(S\), which we will denote as \(⟦-⟧^{ω}_{S}: S → 𝒢_ω\), where the semantic equivalences can identified by bisimulation~\cite{schmid_GuardedKleeneAlgebra_2021}:
+Infinite-trace model \(\mathscr{G}_ω\) is the final coalgebra of GKAT coalgebras~\cite{schmid_GuardedKleeneAlgebra_2021}.
+The finality of the model allows us to define the semantics of each state in a given GKAT coalgebra \(S\), which we will denote as \(⟦-⟧^{ω}_{S}: S → \mathscr{G}_ω\), where the semantic equivalences can identified by bisimulation~\cite{schmid_GuardedKleeneAlgebra_2021}:
 \(⟦s⟧^{ω}_{S} = ⟦t⟧^{ω}_{T}\) if and only if there exists a bisimulation \({∼} ⊆ S × T\), s.t. \(s ∼ t\).
@@
 The infinite trace equivalences can be directly computed with bisimulation on derivative, which supports on-the-fly algorithm as demonstrated by similar systems~\cite{kozen_CoalgebraicTheoryKleene_2017,almeida_DecidingKATHoare_2012,pous_SymbolicAlgorithmsLanguage_2015}.
-However, the \emph{finite} trace model \(𝒢\) is the final coalgebra of GKAT coalgebras without dead states, which we call \emph{normal GKAT coalgebra}~\cite{smolka_GuardedKleeneAlgebra_2020}. 
-Fortunately every GKAT coalgebra can be normalized by rerouting all the transition from dead states to rejection.
+However, the \emph{finite} trace model \(\mathscr{G}\) is the final coalgebra of GKAT coalgebras without dead states, which we call \emph{normal GKAT coalgebra}~\cite{smolka_GuardedKleeneAlgebra_2020}.
+Fortunately every GKAT coalgebra can be normalized by rerouting all the transitions from dead states to rejection.
 Concretely, given a GKAT coalgebra \(S ≜ (S, δ_S)\), \(δ_{\norm(S)} : S → G(S)\) is defined as \(δ_{\norm(S)}(s, α) ≜ \reject\) when \(δ_S(s, α) = (s', p)\) and \(s'\) is dead in \(S\); and \(δ_{\norm(S)}(s, α) ≜ δ_S(s, α)\) otherwise.
@@
 
-And we use \(⟦-⟧_S: \norm(S) → 𝒢\) to denote the finite trace semantics of GKAT coalgebra, which is the unique homomorphism into the final coalgebra \(𝒢\). 
+And we use \(⟦-⟧_S: \norm(S) → \mathscr{G}\) to denote the finite trace semantics of GKAT coalgebra, which is the unique homomorphism into the final coalgebra \(\mathscr{G}\). 
 The finite trace equivalence between \(s ∈ S\) and \(u ∈ U\) can be decided by first normalizing \(S\) and \(U\) then deciding whether there is a bisimulation \({∼} ⊆ \norm(S) × \norm(U)\) s.t.
 \(s ∼ t\).
-For a more detailed explanation on the finite trace semantics, we refer the reader to the work of \Citeauthor{smolka_GuardedKleeneAlgebra_2020}~\cite{smolka_GuardedKleeneAlgebra_2020}, however we will recall the correctness theorem here.
+For a more detailed explanation on the finite trace semantics, we refer the reader to the work of \Citeauthor{smolka_GuardedKleeneAlgebra_2020}~\cite{smolka_GuardedKleeneAlgebra_2020}, whereas we will recall the correctness theorem here.
 
@@
 
-Besides giving us the finite-trace semantics, the normalization operation also satisfy several nice properties.
+Besides giving us the finite-trace semantics, the normalization operation also satisfies several nice properties.
 
@@
 
-Second, normalization is an endofunctor in the category of GKAT coalgebra, this result can help us obtain sub-coalgebras of normalized coalgebra, and also connect the finite trace and infinite trace semantics.
+Second, normalization is an endofunctor in the category of GKAT coalgebra.
+This result can help us obtain sub-coalgebras of normalized coalgebra, and also connect the finite trace with infinite trace semantics.
 
@@
 \begin{mathpar}
-    ϵ̂: S → 𝒫(\BExp_T), \and
-    δ̂: S → 𝒫(\BExp_T × S × K).
+    ϵ̂: S → \mathscr{P}(\BExp_T) \and
+    δ̂: S → \mathscr{P}(\BExp_T × S × K).
 \end{mathpar}
@@
 With these definitions in mind, we can define symbolic Thompson's construction inductively as in~\Cref{tab:symb-Thompson-construction}, where we let \((S₁, ϵ̂₁, δ̂₁)\) and \((S₂, ϵ̂₂, δ̂₂)\) to be result of Thompson's construction for \(e₁\) and \(e₂\) respectively.
-In this table, \(S₁ + S₂\) denotes the disjoint union of \(S₁\) and \(S₂\), and for any two transition dynamics \(δ₁(s₁): 𝒫(\BExp × S₁ × K)\) and \(δ₂(s₂): 𝒫(\BExp × S₂ × K)\), we can also compose them in parallel as \(δ₁(s₁) + δ₂(s₂)\):
-\begin{align*}
-    δ₁(s₁) & + δ₂(s₂): 𝒫(\BExp × (S₁ + S₂) × K) \\*
+In this table, \(S₁ + S₂\) denotes the disjoint union of \(S₁\) and \(S₂\), and for any two transition dynamics \(δ₁(s₁): \mathscr{P}(\BExp × S₁ × K)\) and \(δ₂(s₂): \mathscr{P}(\BExp × S₂ × K)\), we can also compose them in parallel as \(δ₁(s₁) + δ₂(s₂)\):
+\begin{align*}
+    δ₁(s₁) & + δ₂(s₂): \mathscr{P}(\BExp × (S₁ + S₂) × K) \\*
     δ₁(s₁) & + δ₂(s₂) ≜ \\*
@@
     Because homomorphism preserves semantics (\Cref{thm:hom-preserves-semantics}) \(⟦s^*⟧^{ω}_{S_{e}} = ⟦h(s^*)⟧^{ω}_{D} = ⟦e⟧^{ω}_{D}\), where \(⟦-⟧^{ω}\) is the infinite trace semantics i.e.
-the unique map into the final GKAT coalgebra \(𝒢_ω\).
+the unique map into the final GKAT coalgebra \(\mathscr{G}_ω\).
 

### 2025-01-22 main.tex
anonymize authors

@@
 \title{On-the-fly Algorithms for GKAT Equivalences
-\thanks{Identify applicable funding agency here. If none, delete this.}
+% \thanks{Identify applicable funding agency here. If none, delete this.}
 }
 
-\author{\IEEEauthorblockN{Cheng Zhang\textsuperscript{\textsection}}
-\IEEEauthorblockA{\textit{Department of Computer Science} \\
-\textit{University College London}\\
-London, United Kingdom \\
-0000-0002-8197-6181}
-\and
-\IEEEauthorblockN{Qiancheng Fu}
-\IEEEauthorblockA{\textit{Department of Computer Science} \\
-\textit{Boston University}\\
-Boston, USA \\
-qcfu@bu.edu}
-\and
-\IEEEauthorblockN{Hang Ji\textsuperscript{\textsection}}
-\IEEEauthorblockA{\textit{Department of Computer Science} \\
-\textit{Cornell University}\\
-Ithaca, USA \\
-hj476@cornell.edu}
-\and
-\IEEEauthorblockN{Ines Santacruz}
-\IEEEauthorblockA{\textit{Department of Computer Science} \\
-\textit{Boston University}\\
-Boston, USA \\
-email address or ORCID}
-\and
-\IEEEauthorblockN{Marco Gaboardi}
-\IEEEauthorblockA{\textit{Department of Computer Science} \\
-\textit{Boston University}\\
-Boston, USA \\
-inessdv15@gmail.com}
-}
+% anonymize authors
+%
+% \author{\IEEEauthorblockN{Cheng Zhang\textsuperscript{\textsection}}
+% \IEEEauthorblockA{\textit{Department of Computer Science} \\
+% \textit{University College London}\\
+% London, United Kingdom \\
+% 0000-0002-8197-6181}
+% \and
+% \IEEEauthorblockN{Qiancheng Fu}
+% \IEEEauthorblockA{\textit{Department of Computer Science} \\
+% \textit{Boston University}\\
+% Boston, USA \\
+% qcfu@bu.edu}
+% \and
+% \IEEEauthorblockN{Hang Ji\textsuperscript{\textsection}}
+% \IEEEauthorblockA{\textit{Department of Computer Science} \\
+% \textit{Cornell University}\\
+% Ithaca, USA \\
+% hj476@cornell.edu}
+% \and
+% \IEEEauthorblockN{Ines Santacruz}
+% \IEEEauthorblockA{\textit{Department of Computer Science} \\
+% \textit{Boston University}\\
+% Boston, USA \\
+% inessdv15@gmail.com}
+% \and
+% \IEEEauthorblockN{Marco Gaboardi}
+% \IEEEauthorblockA{\textit{Department of Computer Science} \\
+% \textit{Boston University}\\
+% Boston, USA \\
+% gaboardi@bu.edu}
+% }
 
 \maketitle
-\begingroup\renewcommand\thefootnote{\textsection}
-\footnotetext{Work largely performed at Boston University}
-\endgroup
+% \begingroup\renewcommand\thefootnote{\textsection}
+% \footnotetext{Work largely performed at Boston University}
+% \endgroup
 
@@
 \subsection{Evaluation}\label{sec:performance-implementation}
+Due to the fact that our symbolic algorithms are generic regarding the choice of solvers for boolean satisfiability, we compare their performance using BDD to using a SAT solver (MiniSat~\cite{een_MINISAT_2004}).
+Our experiments are performed on a laptop with an Apple M4 Pro CPU and 24 GB RAM.
+\Cref{tab:benchmark-time,tab:benchmark-memory} present the experimental results of equivalence checking based on derivatives (DV), Thompson's construction (TC) and SymKAT~\cite{pous_SymbolicAlgorithmsLanguage_2015} (SK).
+Note that SymKAT uses BDD internally for checking boolean satisfiability.
+Each benchmark consists of 50 expression pairs with properties described by the benchmark's name.
+For example, expressions in `e250b5p10eq' have approximately 250 primitive actions in total (e250), a maximum boolean expression size of 5 (b5), 10 possible unique primitive tests (p10) and are known to be equivalent (eq).
+Benchmarks with the suffix `ne' have expression pairs that are known to be non-equivalent.
+
 \begin{table}
@@
 
-Due to the fact that our symbolic algorithms are generic regarding the choice of solvers for boolean satisfiability, we compare their performance using BDD to using a SAT solver (MiniSat~\cite{een_MINISAT_2004}).
-Our experiments are performed on a laptop with an Apple M4 Pro CPU and 24 GB RAM.
-\Cref{tab:benchmark-time,tab:benchmark-memory} present the experimental results of equivalence checking based on derivatives (DV), Thompson's construction (TC) and SymKAT~\cite{pous_SymbolicAlgorithmsLanguage_2015} (SK).
-Note that SymKAT uses BDD internally for checking boolean satisfiability.
-Each benchmark consists of 50 expression pairs with properties described by the benchmark's name.
-For example, expressions in `e250b5p10eq' have approximately 250 primitive actions in total (e250), a maximum boolean expression size of 5 (b5), 10 possible unique primitive tests (p10) and are known to be equivalent (eq).
-Benchmarks with the suffix `ne' have expression pairs that are known to be non-equivalent.
-
 We observe that DV performs better than TC consistently when both algorithms use the same solver backend and all variants of DV and TC perform better than SymKAT.

### 2025-01-22 main.tex
edits to section 2

@@
 \begin{itemize}
-    \item In~\Cref{sec:background}, we recap the necessary concepts in universal coalgebra, and the coalgebraic theory of GKAT.
+    \item In~\Cref{sec:background}, we outline the necessary concepts in universal coalgebra, and the coalgebraic theory of GKAT.
 In particular, we do not define the semantics of GKAT either algebraically or coalgebraically, as we can avoid directly interacting with the semantics by leveraging the coalgebraic theory of GKAT.
-    \item In~\Cref{sec:greedy-bisim}, we define \emph{greedy bisimulation} and developed its coalgebraic theory to serve as the ground for the correctness of our algorithm.
-    We also performs a brief complexity analysis, and argue that our algorithm is almost always more efficient than the original normalize-bisimulate approach~\cite{smolka_GuardedKleeneAlgebra_2020}.
+    \item In~\Cref{sec:greedy-bisim}, we define \emph{greedy bisimulation} and develop its coalgebraic theory to serve as the ground for the correctness of our algorithm.
+    We also perform a brief complexity analysis and argue that our algorithm is almost always more efficient than the original normalize-bisimulate approach~\cite{smolka_GuardedKleeneAlgebra_2020}.
     \item In~\Cref{sec:symb-gkat-greedy-bisim}, we introduce \emph{symbolic GKAT coalgebra}, and lift the greedy bisimulation algorithm to the symbolic setting.
@@
 We typically call elements in \(S\) the \emph{states} of the coalgebra, and \(σ_S(s)\) the \emph{dynamic} of state \(s\).
-We will sometimes use the states \(S\) to denote the coalgebra, when no ambiguity can arise.
+We will sometimes use the states \(S\) to denote the coalgebra when no ambiguity can arise.
 
@@
 \]
-In a special case, when there exists a homomorphism \(h: S → U\), then we can simply pick \({∼} ⊆ S × U\) to be \(\{(s, h(s)) ∣ s ∈ S\}\), which gives us a bisimulation with the lift \(σ_{∼} ≜ σ_S × σ_U\).
-
-When given a homomorphism \(h: S → U\), we can construct a bisimulation \({∼} ⊆ S × U\), where it relates all the states \(s ∈ S\) with \(h(s) ∈ U\); and each component of the pair will transition individually via their respective transition function \(δ_∼ ≜ δ_S × δ_U\), where \(×\) is the bifunctorial lift of product.
+In a special case, when there exists a homomorphism \(h: S → U\), we can simply pick \({∼} ⊆ S × U\) to be \(\{(s, h(s)) ∣ s ∈ S\}\), which gives us a bisimulation with the lift \(σ_{∼} ≜ σ_S × σ_U\).
+
+When given a homomorphism \(h: S → U\), we can construct a bisimulation \({∼} ⊆ S × U\) where it relates all the states \(s ∈ S\) with \(h(s) ∈ U\).
+Each component of the pair will transition individually via their respective transition function \(δ_∼ ≜ δ_S × δ_U\) where \(×\) is the bifunctorial lift of product.
 \begin{theoremEnd}{corollary}\label{thm:hom-preserves-semantics}
-    Given a homomorphism \(h: S → U\) between two \(F\)-coalgebra \(S, U\), then all for all states \(s ∈ S\), \(\mathrm{beh}_S(s) = \mathrm{beh}_U(h(s)).\)
+    Given a homomorphism \(h: S → U\) between two \(F\)-coalgebra \(S, U\), then for all states \(s ∈ S\), \(\mathrm{beh}_S(s) = \mathrm{beh}_U(h(s)).\)
 \end{theoremEnd}
@@
 Following the conventional notation, we denote an atom using \(α, β\); and we sometimes omit the subscript \(T\) when no confusion can arise.
-Alternatively, atoms can also be thought of as truth assignments to each primitive tests, indicating which primitive tests is satisfied in the current program states; and \(α ≤ b\) if and only if the truth assignment represented by \(α\) satisfies \(b\).
+Alternatively, atoms can also be thought of as truth assignments to each primitive test, indicating which primitive tests are satisfied in the current program state; and \(α ≤ b\) if and only if the truth assignment represented by \(α\) satisfies \(b\).
 
 For the sake of brevity, we omit the complete definition of GKAT and KAT semantics, we refer the reader to previous works~\cite{smolka_GuardedKleeneAlgebra_2020,schmid_GuardedKleeneAlgebra_2021,kozen_KleeneAlgebraTests_1997c}, which explain these semantics in detail.
-Our work avoids direct interaction with the semantics by leveraging prior results in the coalgebraic theory of GKAT, which we will recap below.
+Our work avoids direct interaction with the semantics by leveraging prior results in the coalgebraic theory of GKAT, which we summarize below.
 
@@
 For each atom, states in a KCT can accept or reject the atom (but not both), yet states can also non-deterministically transition to multiple different states via the same atom, while executing different actions.
-As we will see later, the deterministic behavior of GKAT coalgebra not only enables a more versatile symbolic algorithm than KCT~\cite{pous_SymbolicAlgorithmsLanguage_2015}, but also present challenges.
+As we will see later, while the deterministic behavior of GKAT coalgebra enables a more versatile symbolic algorithm than KCT~\cite{pous_SymbolicAlgorithmsLanguage_2015}, it also comes with unique challenges.
 Specifically, GKAT coalgebra requires normalization to compute finite trace equivalences~\cite{smolka_GuardedKleeneAlgebra_2020}, where we reroute all the transitions that cannot lead to acceptance immediately into rejection; and states that can never lead to acceptance are called \emph{dead states}.
@@
 the dead states are only checked when a mismatch requires it; even then, our dead state detection algorithm will only check the necessary states to compute the liveness of the states causing the mismatch.
-This not only avoids unnecessarily checking the liveness of a state, but also enables on-the-fly generation of the coalgebra, like using derivatives, where we can remove irrelevant states from memory after they have been explored.
-
-However, to truly understand our on-the-fly algorithms, we will first need to define ``dead states'', and their roles in defining the coalgebraic semantic of GKAT.
+This not only avoids unnecessarily checking the liveness of a state, but also enables on-the-fly generation of the coalgebra similarly to using derivatives where we can remove irrelevant states from memory after they have been explored.
+
+However, to truly understand our on-the-fly algorithms, we will first need to rigorously define ``dead states'', and their roles in defining the coalgebraic semantic of GKAT.
 
@@
 
-Traditionally, live and dead states are defined by whether they can reach to an accepting state~\cite{smolka_GuardedKleeneAlgebra_2020}.
+Traditionally, live and dead states are defined by whether they can reach an accepting state~\cite{smolka_GuardedKleeneAlgebra_2020}.
 \begin{definition}[Liveness]\label{def:liveness-of-states}
@@
     \[s₀ \transvia{α₁ ∣ p₁} s₁ \transvia{α₂ ∣ p₂} s₂ ⋯ \transvia{αₙ ∣ pₙ} sₙ.\]
-    A state is \emph{dead} if it is not dead.
+    A state is \emph{dead} if it is not live.
 \end{definition}
-However, besides directly working with this explicit definition, we can also give a more coalgebraic characterization of these properties.
-Specifically, by induction on the length of the path from one state to the next, we can show that the principle sub-coalgebra \(⟨s⟩_S\) exactly contains all the reachable state of \(s\) in the GKAT coalgebra \(S\).
+However, besides directly working with this explicit definition, we can also give a coalgebraic characterization of these properties.
+Specifically, by induction on the length of the path from one state to the next, we can show that the principle sub-coalgebra \(⟨s⟩_S\) exactly contains all the reachable states of \(s\) in the GKAT coalgebra \(S\).
 Thus we obtain the following coalgebraic characterization of liveness:
@@
 \begin{theoremEnd}{lemma}\label{thm:dead-iff-all-reachable-dead}
-    A state \(s\) is dead if and only if all elements in \(⟨s⟩\) is dead.
+    A state \(s\) is dead if and only if all elements in \(⟨s⟩\) are dead.
 \end{theoremEnd}
@@
 Infinite-trace model \(\mathscr{G}_ω\) is the final coalgebra of GKAT coalgebras~\cite{schmid_GuardedKleeneAlgebra_2021}.
-The finality of the model allows us to define the semantics of each state in a given GKAT coalgebra \(S\), which we will denote as \(⟦-⟧^{ω}_{S}: S → \mathscr{G}_ω\), where the semantic equivalences can identified by bisimulation~\cite{schmid_GuardedKleeneAlgebra_2021}:
-\(⟦s⟧^{ω}_{S} = ⟦t⟧^{ω}_{T}\) if and only if there exists a bisimulation \({∼} ⊆ S × T\), s.t. \(s ∼ t\).
-
-The infinite trace equivalences can be directly computed with bisimulation on derivative, which supports on-the-fly algorithm as demonstrated by similar systems~\cite{kozen_CoalgebraicTheoryKleene_2017,almeida_DecidingKATHoare_2012,pous_SymbolicAlgorithmsLanguage_2015}.
-However, the \emph{finite} trace model \(\mathscr{G}\) is the final coalgebra of GKAT coalgebras without dead states, which we call \emph{normal GKAT coalgebra}~\cite{smolka_GuardedKleeneAlgebra_2020}.
-Fortunately every GKAT coalgebra can be normalized by rerouting all the transitions from dead states to rejection.
-Concretely, given a GKAT coalgebra \(S ≜ (S, δ_S)\), \(δ_{\norm(S)} : S → G(S)\) is defined as \(δ_{\norm(S)}(s, α) ≜ \reject\) when \(δ_S(s, α) = (s', p)\) and \(s'\) is dead in \(S\); and \(δ_{\norm(S)}(s, α) ≜ δ_S(s, α)\) otherwise.
+The finality of the model allows us to define the semantics of each state in a given GKAT coalgebra \(S\), which we denote as \(⟦-⟧^{ω}_{S}: S → \mathscr{G}_ω\).
+Semantic equivalence \(⟦s⟧^{ω}_{S} = ⟦t⟧^{ω}_{T}\) can be identified by bisimulation~\cite{schmid_GuardedKleeneAlgebra_2021} if and only if there exists a bisimulation \({∼} ⊆ S × T\), s.t.
+\(s ∼ t\).
+
+Infinite trace equivalences can be directly computed with bisimulation on derivatives, which supports on-the-fly algorithms as demonstrated by similar systems~\cite{kozen_CoalgebraicTheoryKleene_2017,almeida_DecidingKATHoare_2012,pous_SymbolicAlgorithmsLanguage_2015}.
+However, the \emph{finite} trace model \(\mathscr{G}\) is the final coalgebra of GKAT coalgebras without dead states~\cite{smolka_GuardedKleeneAlgebra_2020}.
+We call the finite trace model the \emph{normal GKAT coalgebra}.
+Fortunately, every GKAT coalgebra can be normalized by rerouting all the transitions from dead states to rejection.
+Given a GKAT coalgebra \(S ≜ (S, δ_S)\), \(δ_{\norm(S)} : S → G(S)\) is defined as \(δ_{\norm(S)}(s, α) ≜ \reject\) when \(δ_S(s, α) = (s', p)\) and \(s'\) is dead in \(S\); and \(δ_{\norm(S)}(s, α) ≜ δ_S(s, α)\) otherwise.
 We call \(\norm(S) ≜ (S, δ_{\norm(S)})\) the \emph{normalized} coalgebra of \(S\).
 
-And we use \(⟦-⟧_S: \norm(S) → \mathscr{G}\) to denote the finite trace semantics of GKAT coalgebra, which is the unique homomorphism into the final coalgebra \(\mathscr{G}\). 
+We use \(⟦-⟧_S: \norm(S) → \mathscr{G}\) to denote the finite trace semantics of GKAT coalgebra, which is the unique homomorphism into the final coalgebra \(\mathscr{G}\). 
 The finite trace equivalence between \(s ∈ S\) and \(u ∈ U\) can be decided by first normalizing \(S\) and \(U\) then deciding whether there is a bisimulation \({∼} ⊆ \norm(S) × \norm(U)\) s.t.
@@
 
-Because of the functoriality, we can show that two states are infinite-trace equivalent implies these two states are finite-trace equivalent.
-This gives us more tool in proving semantic equivalence between two states in GKAT coalgebras: proving bisimulation of these two states in the \emph{non-normalized} GKAT coalgebra can also obtain semantic equivalence for two states.
+Because of the functoriality, we can show that if two states are infinite-trace equivalent, then these two states are finite-trace equivalent.
+This gives us another tool in proving semantic equivalence between two states in GKAT coalgebras: proving bisimulation between two states in the \emph{non-normalized} GKAT coalgebra proves semantic equivalence for them.
 

### 2025-01-24 main.tex
on the fly bisimulation

@@
 To address these problems, we introduce two new algorithms for bisimulation checking: an on-the-fly bisimulation algorithm based on greedy bisimulation, which enables immediate termination upon discovering a counter-example and supports on-the-fly coalgebra generation with derivatives; and a symbolic algorithm that builds upon the on-the-fly algorithm while leveraging efficient SAT solvers to speed up boolean comparisons and compress the transition structures of GKAT coalgebra.
-To utilize the symbolic bisimulation checking, we also provide two methods to generate symbolic GKAT coalgebras from GKAT expressions, analogous to derivative and Thompson's construction in previous works.
+Additionally, to utilize the symbolic bisimulation checking, we provide two methods to generate symbolic GKAT coalgebras from GKAT expressions, analogous to derivative and Thompson's construction in previous works.
 We have proven the correctness of our algorithms and conducted brief complexity analyses.
@@
 
-Kleene Algebra with Tests (KAT) is a staple algebra system to reason about program equivalence, and have found applications in numerous fields, including software-defined networks~\cite{anderson_NetKATSemanticFoundations_2014a,moeller_KATchFastSymbolic_2024,smolka_ScalableVerificationProbabilistic_2019,smolka_CantorMeetsScott_2017}, program transformations~\cite{kozen_BohmJacopiniTheorem_2008,angus_KleeneAlgebraTests_2001,kozen_NonlocalFlowControl_2008,grathwohl_KAT_2014,kozen_KleeneAlgebraTests_1996}, and program logics~\cite{antonopoulos_AlgebraAlignmentRelational_2023,zhang_DomainReasoningTopKAT_2024,zhang_IncorrectnessLogicKleene_2022c,kozen_HoareLogicKleene_2000,desharnais_ModalKleeneAlgebra_2004}.
+Kleene Algebra with Tests (KAT) is a staple algebra system to reason about program equivalence.
+It has found applications in numerous fields including but not limited to software-defined networks~\cite{anderson_NetKATSemanticFoundations_2014a,moeller_KATchFastSymbolic_2024,smolka_ScalableVerificationProbabilistic_2019,smolka_CantorMeetsScott_2017}, program transformations~\cite{kozen_BohmJacopiniTheorem_2008,angus_KleeneAlgebraTests_2001,kozen_NonlocalFlowControl_2008,grathwohl_KAT_2014,kozen_KleeneAlgebraTests_1996}, and program logics~\cite{antonopoulos_AlgebraAlignmentRelational_2023,zhang_DomainReasoningTopKAT_2024,zhang_IncorrectnessLogicKleene_2022c,kozen_HoareLogicKleene_2000,desharnais_ModalKleeneAlgebra_2004}.
 Beyond its theoretical achievements, Kleene Algebra with Tests has also inspired several practical tools for specialized domains.
@@
 
-This surprising fact lead to the development of Guarded Kleene Algebra with Tests (GKAT)~\cite{smolka_GuardedKleeneAlgebra_2020}, a (co)algebraic system that seeks to encapsulate this efficient fragment of KAT.
-Indeed, the determinism of GKAT gives it a different coalgebra structure from KAT, and allows for a GKAT expression to be directly converted to a linear-sized coalgebra, as opposed to the exponential sized coalgebra produced KAT in general~\cite{smolka_GuardedKleeneAlgebra_2020,kozen_CoalgebraicTheoryKleene_2017}.
-
-However, as it current stands, the desirable theoretical properties of GKAT do not translate to tangible real-world performance. 
-The decision procedures for GKAT can often be slower than that of KAT, primarily due to the following two reasons.
+This surprising fact lead to the development of Guarded Kleene Algebra with Tests (GKAT)~\cite{smolka_GuardedKleeneAlgebra_2020}, a (co)algebraic system that seeks to characterize this efficient fragment of KAT.
+The determinism of GKAT gives it a different coalgebra structure from KAT, and allows for a GKAT expression to be directly converted to a linear-sized coalgebra, as opposed to the exponential sized coalgebra produced by KAT expressions~\cite{smolka_GuardedKleeneAlgebra_2020,kozen_CoalgebraicTheoryKleene_2017}.
+
+However, as it currently stands, the desirable theoretical properties of GKAT do not translate to tangible real-world performance. 
+The decision procedures for GKAT can often be slower than that of KAT; primarily due to the following two reasons.
 
 First, the natural semantics of GKAT coalgebra is not the finite-trace model, but the infinite-trace model~\cite{smolka_GuardedKleeneAlgebra_2020,schmid_GuardedKleeneAlgebra_2021}.
-Thus, to determine finite-trace equivalence, it is necessary to \emph{normalize} the coalgebra first, requiring one to compute the entire coalgebra ahead-of-time; and iterate through its entirety.
-Without the need to normalize prior to computing bisimulation, the decision procedure of KAT equivalence can terminate as soon as a counter-example has been found which can lead to better performance than the original GKAT decision algorithm~\cite{smolka_GuardedKleeneAlgebra_2020}.
+Thus, to determine finite-trace equivalence, it is necessary to \emph{normalize} the coalgebra first. 
+This requires one to compute the coalgebra ahead-of-time and iterate through its entirety.
+Without the need to normalize prior to computing bisimulation, the decision procedure of KAT equivalence can terminate as soon as a counter-example has been found, potentially leading to better performance than the original GKAT decision algorithm~\cite{smolka_GuardedKleeneAlgebra_2020}.
 
@@
 Although the generated coalgebra from GKAT expression have a linear number of states, the number of transitions is exponential to \emph{all} the primitive tests appearing in the entire expression.
-This inefficiency is noted in previous literature, and poses a real limitation on real-world systems~\cite{smolka_GuardedKleeneAlgebra_2020,zhang_CFGKATEfficientValidation_2025}.
+This inefficiency is noted in previous literature, and poses a real limitation on practical systems~\cite{smolka_GuardedKleeneAlgebra_2020,zhang_CFGKATEfficientValidation_2025}.
 On the other hand, symbolic algorithms for KAT have been proven to be effective for generic equality solving and domain-specific applications~\cite{pous_SymbolicAlgorithmsLanguage_2015,moeller_KATchFastSymbolic_2024}.
@@
 
-The second algorithm lifts the first into a symbolic setting, where our development is based on an extension of GKAT coalgebra, which we named \emph{symbolic GKAT coalgebra}.
+The second algorithm lifts the first into a symbolic setting, where our development is based on a generalization of GKAT coalgebra, which we named \emph{symbolic GKAT coalgebra}.
 In contrast to the symbolic KAT derivative~\cite{pous_SymbolicAlgorithmsLanguage_2015}, the deterministic nature of GKAT coalgebra allows its symbolic variant to transition via boolean expressions instead of primitive tests.
@@
 To evaluate the practical impact of our work, we implement our algorithms in Rust. We show through experiments that
-our implementation performs orders of magnitude better than existing implementation for KAT~\cite{pous_SymbolicAlgorithmsLanguage_2015},
-confirming that the theoretical elegance of GKAT indeed leads to significant real-world performance improvements.
+our implementation performs orders of magnitude better than existing implementation for KAT~\cite{pous_SymbolicAlgorithmsLanguage_2015}, confirming that the theoretical elegance of GKAT indeed leads to significant real-world performance improvements.
 
@@
     \item In~\Cref{sec:symb-gkat-construction}, we lift the Thompson's construction and derivative construction to the symbolic setting.
-    Then we establish their connection via a coalgebra homomorphism which allows us to derive the correctness and finiteness of one construction from the other.
+    Then we establish their connection via a coalgebra homomorphism, which allows us to derive the correctness and finiteness of one construction from the other.
     \item Finally, in~\Cref{sec:implementation}, we describe our implementation of symbolic derivative, symbolic Thompson's construction, and symbolic greedy bisimulation.
@@
 We typically call elements in \(S\) the \emph{states} of the coalgebra, and \(σ_S(s)\) the \emph{dynamic} of state \(s\).
-We will sometimes use the states \(S\) to denote the coalgebra when no ambiguity can arise.
-
-A homomorphism between two \(F\)-coalgebra \(S\) and \(U\) is a map \(h: S → U\) that preserves the function \(σ\); diagrammatically, the following diagram commutes:
+We will sometimes overload \(S\) to denote both the state set and the coalgebra, when no ambiguity can arise.
+
+A homomorphism between two \(F\)-coalgebra \(S\) and \(U\) is a map \(h: S → U\) that preserves the function \(σ\):
 \[
@@
 When we can restrict the homomorphism map into an inclusion map \(i: S' → S\) for \(S' ⊆ S\) then we say that \(S'\) is a \emph{sub-coalgebra} of \(S\), denoted as \(S' ⊑ S\).
-Specifically, the following diagram commutes when \(S' ⊑ S\):
-\[
-    \begin{tikzcd}
-        S' \ar[hook]{r}{i} \ar[swap]{d}{σ_{S'}} & S \ar{d}{σ_S} \\  
-        F(S') \ar[hook]{r}{F(i)} & F(S)
-    \end{tikzcd}    
-\]
 In fact, the function \(σ_{S'}\) is uniquely determined by the states \(S'\)~\cite[Proposition 6.1]{rutten_UniversalCoalgebraTheory_2000}.
@@
 We sometimes omit the subscript \(S\) when it can be inferred from context or irrelevant.
-Intuitively, we usually think of principle sub-coalgebra \(⟨s⟩_S\) as the sub-coalgebra that is formed by all the ``reachable state'' from \(s\).
-This coalgebraic characterization of reachable state can allow us to avoid induction on the length of a path from \(s\) to a state in \(⟨s⟩_S\).
-
+Intuitively, we usually think of principle sub-coalgebra \(⟨s⟩_S\) as the sub-coalgebra that is formed by all the ``reachable states'' from \(s\).
 For all coalgebra \(S\) and a state \(s ∈ S\), principle sub-coalgebra \(⟨s⟩_S\) always exists and is unique, because sub-coalgebra of any coalgebra forms a complete lattice~\cite[theorem 6.4]{rutten_UniversalCoalgebraTheory_2000}; thus taking the meet of all the sub-coalgebra that contains \(s\) will yield \(⟨s⟩_S\).
@@
 \]
-In a special case, when there exists a homomorphism \(h: S → U\), we can simply pick \({∼} ⊆ S × U\) to be \(\{(s, h(s)) ∣ s ∈ S\}\), which gives us a bisimulation with the lift \(σ_{∼} ≜ σ_S × σ_U\).
+We sometimes use \({∼} ⊆ S × U\) to specify that \(∼\) is a bisimulation between \(S\) and \(U\).
 
 When given a homomorphism \(h: S → U\), we can construct a bisimulation \({∼} ⊆ S × U\) where it relates all the states \(s ∈ S\) with \(h(s) ∈ U\).
-Each component of the pair will transition individually via their respective transition function \(δ_∼ ≜ δ_S × δ_U\) where \(×\) is the bifunctorial lift of product.
+Each component of the pair will transition individually via their respective transition function \(δ_∼ ≜ δ_S × δ_U\).
 \begin{theoremEnd}{corollary}\label{thm:hom-preserves-semantics}
@@
 
-Guarded Kleene Algebra with Tests, or GKAT~\cite{smolka_GuardedKleeneAlgebra_2020}, is a deterministic fragment of Kleene Algebra with Tests. 
+Guarded Kleene Algebra with Tests, or GKAT~\cite{smolka_GuardedKleeneAlgebra_2020}, is a deterministic fragment of Kleene Algebra with Tests (KAT). 
 The syntax of GKAT over a set of primitive actions \(K\) and a set of primitive tests \(T\) can be defined in two sorts, boolean expressions \(\BExp\) and GKAT expressions \(\Exp\):
@@
 Then the semantics of each expression \(⟦e⟧\) can be computed by the semantics of Kleene Algebra with Tests~\cite{kozen_KleeneAlgebraTests_1997c}.
-An important construct in the semantics is \emph{atoms}, which are conjunctions of all the primitive tests either in its positive or negative form: for \(T ≜ \{t₁, t₂, …, tₙ\}\)
+An important construct in the semantics is \emph{atoms}, which are conjunctions of all the primitive tests either in its positive or negative form: for \(T ≜ \{t₁, t₂, …, tₙ\}\),
 \[\At_T ≜ \{t₁' ∧ t₂' ∧ ⋯ ∧ tₙ' ∣ tᵢ' ∈ \{tᵢ, \overline{tᵢ}\}\}.\]
 Following the conventional notation, we denote an atom using \(α, β\); and we sometimes omit the subscript \(T\) when no confusion can arise.
-Alternatively, atoms can also be thought of as truth assignments to each primitive test, indicating which primitive tests are satisfied in the current program state; and \(α ≤ b\) if and only if the truth assignment represented by \(α\) satisfies \(b\).
+Alternatively, atoms can also be thought of as truth assignments to each primitive test, indicating which primitive tests are satisfied in the current program state; and \(α ≤ b\) in Boolean Algebra if and only if the truth assignment represented by \(α\) satisfies \(b\).
 
@@
 \[G(S) ≜ (\{\accept, \reject\} + S × K)^{\At_T}.\] 
-Intuitively, given a state \(s ∈ S\) and an atom \(α ∈ \At\), \(δ(s, α)\) will deterministically execute one of the following: reject \(α\) when \(δ(s, α) = \reject\); accept \(α\) when \(δ(s, α) = \accept\); or transition to a state \(s' ∈ S\) and execute action \(p ∈ K\) when \(δ(s, α) = (s', p)\).
+Intuitively, given a state \(s ∈ S\) and an atom \(α ∈ \At\), \(δ_S(s, α)\) will deterministically execute one of the following: reject \(α\) when \(δ(s, α) = \reject\), denoted as \(s \transRej{S}{α}\); accept \(α\) when \(δ(s, α) = \accept\), denoted \(s \transAcc{S}{α}\); or transition to a state \(s' ∈ S\) and execute action \(p ∈ K\) when \(δ(s, α) = (s', p)\), denoted \(s \transvia{α ∣ p}_S s'\).
 
@@
 
-Besides nice property with bisimulation, GKAT coalgebra is also deterministic, unlike Kleene Coalgebra with Tests (KCT)~\cite{kozen_CoalgebraicTheoryKleene_2017}.
+Besides the nice property of bisimulation, GKAT coalgebra is also deterministic, unlike Kleene Coalgebra with Tests (KCT)~\cite{kozen_CoalgebraicTheoryKleene_2017}.
 For each atom, states in a KCT can accept or reject the atom (but not both), yet states can also non-deterministically transition to multiple different states via the same atom, while executing different actions.
@@
 the dead states are only checked when a mismatch requires it; even then, our dead state detection algorithm will only check the necessary states to compute the liveness of the states causing the mismatch.
-This not only avoids unnecessarily checking the liveness of a state, but also enables on-the-fly generation of the coalgebra similarly to using derivatives where we can remove irrelevant states from memory after they have been explored.
+This not only avoids unnecessarily checking the liveness of a state, but also enables on-the-fly generation of the coalgebra, where we can remove irrelevant states from memory after they have been explored.
 
@@
 We show the following theorems as examples:
-\begin{theoremEnd}{lemma}\label{thm:dead-iff-all-reachable-dead}
+\begin{theoremEnd}[normal]{lemma}\label{thm:dead-iff-all-reachable-dead}
     A state \(s\) is dead if and only if all elements in \(⟨s⟩\) are dead.
@@
 
-\begin{theoremEnd}{theorem}[homomorphism perserves liveness]\label{thm:hom-preserve-liveness}
+\begin{theoremEnd}[normal]{theorem}[homomorphism perserves liveness]\label{thm:hom-preserve-liveness}
     Given a homomorphism \(h: S → U\) and a state \(s ∈ S\):
@@
 
-Infinite-trace model \(\mathscr{G}_ω\) is the final coalgebra of GKAT coalgebras~\cite{schmid_GuardedKleeneAlgebra_2021}.
-The finality of the model allows us to define the semantics of each state in a given GKAT coalgebra \(S\), which we denote as \(⟦-⟧^{ω}_{S}: S → \mathscr{G}_ω\).
-Semantic equivalence \(⟦s⟧^{ω}_{S} = ⟦t⟧^{ω}_{T}\) can be identified by bisimulation~\cite{schmid_GuardedKleeneAlgebra_2021} if and only if there exists a bisimulation \({∼} ⊆ S × T\), s.t.
-\(s ∼ t\).
-
-Infinite trace equivalences can be directly computed with bisimulation on derivatives, which supports on-the-fly algorithms as demonstrated by similar systems~\cite{kozen_CoalgebraicTheoryKleene_2017,almeida_DecidingKATHoare_2012,pous_SymbolicAlgorithmsLanguage_2015}.
-However, the \emph{finite} trace model \(\mathscr{G}\) is the final coalgebra of GKAT coalgebras without dead states~\cite{smolka_GuardedKleeneAlgebra_2020}.
-We call the finite trace model the \emph{normal GKAT coalgebra}.
+Infinite-trace model \(𝒢_ω\) is the final coalgebra of GKAT coalgebras~\cite{schmid_GuardedKleeneAlgebra_2021}.
+Thus, the unique homomorphism from any coalgebra \(S\) to \(𝒢_ω\) will assign each states in \(S\) a semantics.
+We denote this unique homomorphism as \(⟦-⟧^{ω}_{S}: S → 𝒢_ω\), where the semantic equivalences can be identified by bisimulation:
+\(⟦s⟧^{ω}_{S} = ⟦t⟧^{ω}_{T}\) if and only if there exists a bisimulation \({∼} ⊆ S × T\), s.t. \(s ∼ t\).
+
+Infinite-trace equivalences can be directly computed with bisimulation on derivatives, which supports on-the-fly algorithms as demonstrated by similar systems~\cite{kozen_CoalgebraicTheoryKleene_2017,almeida_DecidingKATHoare_2012,pous_SymbolicAlgorithmsLanguage_2015}.
+However, the \emph{finite} trace model \(\mathscr{G}\) is the final coalgebra of \emph{normal GKAT coalgebras}, where no state can transition to a dead state~\cite{smolka_GuardedKleeneAlgebra_2020}. 
 Fortunately, every GKAT coalgebra can be normalized by rerouting all the transitions from dead states to rejection.
-Given a GKAT coalgebra \(S ≜ (S, δ_S)\), \(δ_{\norm(S)} : S → G(S)\) is defined as \(δ_{\norm(S)}(s, α) ≜ \reject\) when \(δ_S(s, α) = (s', p)\) and \(s'\) is dead in \(S\); and \(δ_{\norm(S)}(s, α) ≜ δ_S(s, α)\) otherwise.
+Given a GKAT coalgebra \(S ≜ (S, δ_S)\), the transition of its normalized coalgebra \(δ_{\norm(S)} : S → G(S)\) is defined as \(δ_{\norm(S)}(s, α) ≜ \reject\) when \(δ_S(s, α) = (s', p)\) and \(s'\) is dead in \(S\); and \(δ_{\norm(S)}(s, α) ≜ δ_S(s, α)\) otherwise.
 We call \(\norm(S) ≜ (S, δ_{\norm(S)})\) the \emph{normalized} coalgebra of \(S\).
@@
 
-Because of the functoriality, we can show that if two states are infinite-trace equivalent, then these two states are finite-trace equivalent.
-This gives us another tool in proving semantic equivalence between two states in GKAT coalgebras: proving bisimulation between two states in the \emph{non-normalized} GKAT coalgebra proves semantic equivalence for them.
+Because of the functoriality, we can show that two states that are infinite-trace equivalent implies these two states are finite-trace equivalent.
+This gives us more tool in proving semantic equivalence between two states in GKAT coalgebras: proving bisimulation of these two states in the \emph{non-normalized} GKAT coalgebra can also obtain finite-trance equivalence for the two states.
 
@@
 The original algorithm for deciding GKAT equivalences~\cite{smolka_GuardedKleeneAlgebra_2020} requires the entire coalgebra to be known prior to the execution of the bisimulation algorithm; specifically, it is necessary to iterate through the entire coalgebra in order to identify the liveness of every single states and perform the normalization operation.
-This limitation poses challenges to design an efficient on-the-fly algorithm for GKAT.
+This limitation poses challenges to the design of an efficient on-the-fly algorithm for GKAT.
 To make the equivalence-checking procedure scalable, we propose to greedily run the bisimulation algorithm and only invoke liveness detection when a discrepancy is found. 
-For example, when deciding the trace equivalence between \(s ∈ S\) and \(u ∈ U\), if \(δ_S(s, α) = (s', p)\) and \(δ_U(u, α) = (u', p)\), we will proceed to recurse on \(s'\) and \(u'\), without checking the liveness of \(s'\) and \(u'\).
-However, if \(s\) rejects \(α\) instead of transitioning to \(s'\), we will then check whether \(u'\) is dead.
-
-This approach offers several benefit: first, because the liveness detection is only called when necessary, thus it is unlikely to iterate through the entire coalgebra; second, the algorithm can short-circuit on a counter-example, allowing termination even before all the states are explored; finally, the algorithm supports on-the-fly constructions like derivatives, because the coalgebra no longer is required to be fully constructed.
-
-The correctness argument of this on-the-fly algorithm is centered around the following coalgebraic structure, which we named ``greedy bisimulation''.
+For example, when constructing the bisimulation between \(s ∈ S\) and \(u ∈ U\), if \(s \transvia{α ∣ p}_S s'\) and \(u \transvia{α ∣ p}_U u'\), we will proceed to recurse on \(s'\) and \(u'\), without checking the liveness of \(s'\) and \(u'\).
+Whereas, if \(s \transRej{S}{α}\) but \(u \transvia{α ∣ p}_U u'\), then the transition of \(s\) and \(u\) disagree, requiring \(u'\) to be dead.
+
+This approach offers several benefit: first, the liveness detection is only called when necessary, thus unlikely to iterate through the entire coalgebra; second, the algorithm can short-circuit on a counter-example, allowing termination before all the states are explored; finally, the algorithm supports on-the-fly coalgebra constructions like derivatives, as the coalgebra is no longer required to be fully constructed before bisimulation.
+
+The intuition of this on-the-fly algorithm can be expressed as the following coalgebraic structure, which we named ``greedy bisimulation''.
 
@@
 it is almost a greedy bisimulation, i.e.
-it is closed under \(δ_≈\) in~\Cref{def:greedy-bisim}, and contains \(∼\), except it is not the \emph{least} GKAT coalgebra that satisfy these properties.
+it is closed under \(δ_≈\) in~\Cref{def:greedy-bisim}, and contains \(∼\), except it is not necessarily the \emph{least} GKAT coalgebra that satisfies these properties.
 
@@
     \[{∼_d} ≜ {∼} ∪ \{(s, u) ∣ \text{\(s\) and \(u\) are dead in \(S\) and \(U\)}\},\]
-    form a GKAT coalgebra the transition function \(δ_≈\) in~\Cref{def:greedy-bisim}, which implies \({≈} ⊑ {∼_d}\).
+    forms a GKAT coalgebra with the transition function \(δ_≈\) in~\Cref{def:greedy-bisim}, which implies \({≈} ⊑ {∼_d}\).
     Furthermore, \(\norm(∼_d)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\).
@@
 
-Although the proof of~\Cref{thm:dead-construction-greedy-bisim} is basically by unfolding the definition, \(∼_d\) turns out to be a important structure. 
-Specifically, \(∼_d\) serves a connection between \(≈\) and its underlying bisimulation \(∼\), and the order \({∼} ⊆ {≈} ⊑ {~_d}\) will reveal many structural properties of \(≈\).
-We show the following corollary to demonstrate the usefulness of \(∼_d\).
-
-\begin{theoremEnd}{corollary}\label{thm:greedy-bisim-exists-unique}
-    For any bisimulation between two normalized GKAT coalgebra \({∼} ⊆ \norm(S) × \norm(U)\), a greedy bisimulation \(≈\) with \(∼\) as underlying bisimulation exists and is unique.
+Although \Cref{thm:dead-construction-greedy-bisim} can be proved by unfolding the definition, \(∼_d\) turns out to be an important structure. 
+Specifically, \(∼_d\) serves a connection between \(≈\) and its underlying bisimulation \(∼\), and the order \({∼} ⊆ {≈} ⊑ {∼_d}\) will reveal many structural properties of \(≈\).
+
+\begin{theoremEnd}[normal]{corollary}\label{thm:greedy-bisim-exists-unique}
+    For any bisimulation between two normalized GKAT coalgebras \({∼} ⊆ \norm(S) × \norm(U)\), a greedy bisimulation \(≈\) with \(∼\) as underlying bisimulation exists and is unique.
 \end{theoremEnd}
@@
 
-\begin{theoremEnd}{corollary}\label{thm:greedy-bisim-dead-or-bisim}
+\begin{theoremEnd}[normal]{corollary}\label{thm:greedy-bisim-dead-or-bisim}
     For every pair of states \((s, u)\) in a greedy bisimulation \({≈} ⊆ S × U\), either \((s, u)\) is in the underlying bisimulation \(∼\), or both \(s\) and \(u\) are dead in \(S\) and \(U\).
@@
 \begin{proofEnd}
-    Because \(∼_d\) is a GKAT coalgebra with transition \(δ_≈\) and also contains all the states in \(∼\), therefore \({≈} ⊑ {∼_d}\).
-    Thus, for all \((s, u) ∈ {≈}\) but not in \(∼\), \(s\) and \(u\) must be dead in \(S\) and \(U\).
-\end{proofEnd}
-
-\begin{theoremEnd}{corollary}[Perservation Of Liveness]\label{thm:greedy-bisim-perserve-liveness}
+    Because \({≈} ⊑ {∼_d}\), and \(∼_d\) only contains pairs in \(∼\) or pairs of dead-state, then so is \(≈\).
+\end{proofEnd}
+
+\begin{theoremEnd}[normal]{corollary}[Perservation Of Liveness]\label{thm:greedy-bisim-perserve-liveness}
     Given a pair of states in a greedy bisimulation \(s ≈ u\) where \({≈} ⊆ S × U\), then the following are equivalent
@@
 \begin{proofEnd}
-    Let \(∼\) by the underlying bisimulation of \(≈\), consider the \(∼_d\) in~\Cref{thm:dead-construction-greedy-bisim}, then we know that \({≈} ⊑ {∼_d}\).
-
-    Because \(\norm(∼_d)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\), then
+    Let \(∼\) by the underlying bisimulation of \(≈\), consider the \(∼_d\) in~\Cref{thm:dead-construction-greedy-bisim}.
+    Because \({≈} ⊑ {∼_d}\) and \(\norm(∼_d) ⊆ \norm(S) × \norm(U)\) is a bisimulation,
     \begin{align*}
@@
 
-\begin{theoremEnd}{corollary}\label{thm:greedy-bisim-iff-norm-bisim}
+\begin{theoremEnd}[normal]{corollary}\label{thm:greedy-bisim-iff-norm-bisim}
     Given two GKAT coalgebras \(S\) and \(U\), and a relational GKAT coalgebra \({≈} ⊆ S × U\) with the transition \(δ_≈\) as in~\cref{def:greedy-bisim}.
@@
 
-\begin{theoremEnd}{corollary}\label{thm:max-bisim-is-max-greedy-bisim}
-    Given two GKAT coalgebra \(S\) and \(U\), carrier of the maximal bisimulation \({≡} ⊆ \norm(S) × \norm(U)\) is also the carrier of maximal greedy bisimulation, which we will denote as \(≊ ⊆ S × U\).
-    We will abuse the notation \(≡\) to represent both the maximal bisimulation and maximal greedy bisimulation.
+Finally, with all the structural properties in place, we can derive that the maximal greedy bisimulation actually corresponds to maximal bisimulation.
+
+\begin{theoremEnd}{theorem}\label{thm:max-bisim-is-max-greedy-bisim}
+    Given two GKAT coalgebra \(S\) and \(U\), carrier of the maximal bisimulation \({≡} ⊆ \norm(S) × \norm(U)\) is also the carrier of maximal greedy bisimulation, which we will denote as \({≊} ⊆ S × U\).
 \end{theoremEnd}
@@
     by definition of greedy bisimulation, the carrier of \({≡}\) is a subset of the carrier of \({≊}\). 
-    Then because of~\Cref{thm:greedy-bisim-iff-norm-bisim}, \(\norm(≊)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\), which implies \(\norm(≡') ⊑ {≡}\).
-    Finally because \(\norm\) do not change the carrier set, therefore the carrier of \(≡\) and \(≊\) necessarily coincide.
-
-    Then we will need to show that \(≡\) with the \(δ_≈\) defined in~\Cref{def:greedy-bisim} is the maximal greedy bisimulation.
-    Take an arbitrary greedy bisimulation \({≈} ⊆ S × U\) with underlying bisimulation \(∼\), and a pair of states \(s ≈ u\), we will need to show \(s ≡ u\).
-    By~\cref{thm:greedy-bisim-dead-or-bisim}, we can case analysis on \(s\) and \(u\): if \(s ∼ u\), then \(s ≡ u\) because \(≡\) is the maximal bisimulation; and if \(s\) and \(u\) are both dead, then by~\cref{thm:bisim-between-dead}, \(s ≡ u\).
-\end{proofEnd}
-
-\Cref{thm:max-bisim-is-max-greedy-bisim} allows us to take the maximal bisimulation as a greedy bisimulation, aiding several proofs.
-
-Similar to the correctness of bisimulation between normalized coalgebra (\Cref{thm:norm-bisim-correctness}), we also establish a correctness theorem for greedy bisimulation, i.e. there exists
-
-\begin{theoremEnd}{theorem}[Correctness]
+    Then because of~\Cref{thm:greedy-bisim-iff-norm-bisim}, \(\norm(≊)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\), which implies \(\norm(≊) ⊑ {≡}\).
+    Finally, because \(\norm\) do not change the carrier set, therefore the carrier of \(≡\) and \(≊\) necessarily coincide.
+\end{proofEnd}
+
+Similarly to the correctness of bisimulation between normalized coalgebra (\Cref{thm:norm-bisim-correctness}), we also establish a correctness theorem for greedy bisimulation, i.e.
+there exists a greedy bisimulation between two states if and only if they are semantically equivalent.
+
+\begin{theoremEnd}{theorem}[Correctness]\label{thm:greedy-bisim-correctness}
     Given two states in GKAT coalgebras \(s ∈ S\) and \(u ∈ ,U\), there exists a greedy bisimulation \({≈} ⊆ S × U\) s.t. \(s ≈ u\), if and only if \(⟦s⟧_S = ⟦u⟧_U\).
@@
 
-Despite proving many desirable structural properties, the definition of greedy bisimulation is still coalgebraic.
-In order for our equivalence-checking algorithm to construct a greedy bisimulation, we need to turn the coalgebra into a set of concrete criteria, that can be checked by an algorithm.
+Despite satisfying many desirable structural properties, the definition of greedy bisimulation is still coalgebraic.
+To aid the design of an algorithm based on greedy bisimulation, we turn the coalgebra into a set of concrete criteria, that can be checked by an algorithm.
 In the following theorem, we will present such necessary and sufficient conditions to construct a greedy bisimulation.
-Our algorithm will be largely based on recursively checking these conditions to establish a greedy bisimulation.
-
 \begin{theoremEnd}{theorem}[Recursive Construction]\label{thm:recursive-construction}
@@
 \end{theoremEnd}
+
+\begin{proof}[Sketch]
+    The \(⟹\) direction is by unfolding the definitions; and the \(⟸\) direction is proven by demonstrating \(\norm(≈) ⊆ \norm(S) × \norm(U)\) is a bisimulation, then \(≈\) is a greedy bisimulation because of~\Cref{thm:greedy-bisim-iff-norm-bisim}.
+\end{proof}
 
@@
 
-\begin{theoremEnd}{lemma}\label{thm:bisim-between-dead}
+\begin{theoremEnd}[all end]{lemma}\label{thm:bisim-between-dead}
     Given any two dead states \(s ∈ S\) and \(u ∈ U\), we can construct a bisimulation \(∼\) between \(\norm(S)\) and \(\norm(U)\):
@@
 \end{theoremEnd}
+
+Besides the construction, a classical optimization of the bisimulation-construction algorithm is to look for a equivalence relation instead of a relation~\cite{smolka_GuardedKleeneAlgebra_2020,hopcroft_LinearAlgorithmTesting_1971}.
+This technique also applies to greedy bisimulation.
 
@@
 
-Before we introduce the optimized algorithm, we will first briefly sketch the liveness-check algorithm, which will use a depth-first search to check whether the there exists any accepting states in reachable states of \(s ∈ S\): if there exists any accepting state in \(⟨s⟩_S\), then the algorithm terminates immediately, and return that \(s\) is live, otherwise it would return all the states in \(⟨s⟩\), and by~\cref{thm:dead-iff-all-reachable-dead}, all the states in \(⟨s⟩\) is dead.
-In this case, we use depth-first search for its simplicity to implement, other search algorithm will also be sound.
-
-We will cache all the known dead states from the previous searches; whether a state \(s\) is in this cache can be checked by the function call \Call{knownDead\(_S\)}{$s$}. 
-The function \Call{isDead\(_S\)}{$s$} will first check if \(s\) is known to be dead, and invoke the search algorithm in the coalgebra \(⟨s⟩\), if \(s\) is not in the cached dead states.
+Before we introduce the algorithm, we will first briefly sketch the liveness-check algorithm.
+It uses depth-first search to check whether there exists any accepting states in the reachable states of \(s ∈ S\): if there exists an accepting state in \(⟨s⟩_S\), then the algorithm terminates immediately, and returns that \(s\) is live, otherwise it would return all the states in \(⟨s⟩\), and by~\cref{thm:dead-iff-all-reachable-dead}, all the states in \(⟨s⟩\) are dead.
+Note that we use depth-first search here for its simplicity, other search algorithms will also be sound.
+
+We cache all the known dead states from the previous searches; whether a state \(s\) is in this cache can be checked by the function call \Call{knownDead\(_S\)}{$s$}. 
+The function \Call{isDead\(_S\)}{$s$} will first check if \(s\) is known to be dead, and invoke the search algorithm in the coalgebra \(⟨s⟩\) if \(s\) is not in the cached dead states.
 Because the non-symbolic version of liveness checking is similar to the symbolic version, we only provide the symbolic version of this algorithm in~\cref{lst:dead-state-detection}.
@@
         \end{aligned}\)
-        % \Forall{\(α ∈ \At\)}{}{
-        %     \Match{$δ_{S}(s, α), δ_{U}(u, α)$}
-        %     \Case{\(\accept, \accept\)} {\Continue} \EndCase
-        %     \Case{\(\reject, \reject\)} {\Continue} \EndCase
-        %     \Case{\(\reject, (u', q)\)} 
-        %         \State\Return \Call{isDead\(_U\)}{$u'$} 
-        %     \EndCase
-        %     \Case{\((s', p), \reject\)}
-        %         \State\Return\Call{isDead\(_S\)}{$s'$}
-        %     \EndCase
-        %     \Case{\((s', p), (u', q)\)} {
-        %         \If {\(p = q\)}
-        %             \State\Call{union}{$s, u$}
-        %             \State\Return \Call{equiv}{$s, t$} 
-        %         \EndIf
-        %         \If{\Call{isDead\(_S\)}{$s$} and \Call{isDead\(_U\)}{$u$}}
-        %             \State\Continue
-        %         \EndIf
-        %         \State\Return false
-        %     } \EndCase
-        %     \Default { \Return false } \EndDefault
-        %     \EndMatch
-        % }\EndFor
-        % \State\Return true
         \EndFunction
@@
 
-Finally, we present our equivalence-checking algorithm as~\cref{alg:bisim}, which are mostly checking the conditions in~\cref{thm:recursive-construction}.  
-In the algorithm, we use the text ``implies'', ``iff'', ``and'' are logical operators that takes booleans (i.e. true or false) and return a boolean.
-Similarly, \(\&\!\&\) is the logical and operation, which is the same as the the written ``and''. 
-We opt to use \(\&\!\&\), instead of ``and'', to separate the conditions for clarity; formally \(\&\!\&\) is the same function as ``and'', but binds less tightly.
-
-Intuitively, starting with the pair of state \((s, u)\), it will explore all of its reachable state following \(δ_≈\) in~\cref{def:greedy-bisim}.
-In other word, if the algorithm terminates and returned true, we will have explored every state in \(⟨(s, u)⟩_{≊}\) where \(≊\) is the maximal greedy bisimulation between \(S\) and \(U\).
-However, our algorithm slightly deviates from the ideal scenario described above, specifically we implemented two optimization.
-First, we uses a union-find data structure to keep track of existing state, which organizes the explored states into equivalence classes.
-Indeed, by looking for a greedy bisimulation that is also equivalence relation, we can use more efficient structures like union-find, instead of set of pairs; and the correctness of such optimization is justified by~\Cref{thm:greedy-bisim-iff-greedy-bisim-equiv}.
-The second optimization is to terminate and return false as soon as we know one of \(s, u\) is dead and the other is not dead.
-Indeed, in order for \(s ≈ u\) for any greedy bisimulation \(≈\), \(s\) and \(u\) necessarily have the same liveness in their respective GKAT coalgebra.
-And in the other case, if both \(s\) and \(u\) are dead in their respective coalgebra, we will skip the check on all the reachable state of \((s, u)\) following \(δ_≈\), because a pair of dead state is necessarily in the maximal greedy bisimulation \(≊\) (\Cref{thm:bisim-between-dead,thm:max-bisim-is-max-greedy-bisim}).
-
-Concerning the complexity, our algorithm have similar worst case complexity as the original algorithm.
-In particular, when deciding the trace equivalence of two states \(s ∈ S\) and \(u ∈ U\), the original algorithm~\cite{smolka_GuardedKleeneAlgebra_2020} requires one pass of \(⟨s⟩\) and \(⟨u⟩\) to normalize them, then the bisimulation will visit each pair of states in \(⟨s⟩ × ⟨u⟩\) at most once, which means that they will visit at most \(|⟨s⟩ + ⟨u⟩ + ⟨s⟩ × ⟨u⟩|\) number of states.
-
-Similarly, our equivalence-checking algorithm attempts to find a bisimulation first, which visits each pair in \(⟨s⟩ × ⟨u⟩\) at most once, and when a mismatch is found in the process, the liveness-checking algorithm is then invoked.
-Crucially, our algorithm satisfy the following property: if the liveness-checking algorithm find a live or accepting state, the entire equivalence checking algorithm will halt and return false.
+Finally, we present our equivalence-checking algorithm as~\Cref{alg:bisim} which consists of recursively checking the conditions in~\Cref{thm:recursive-construction} with minor optimizations.
+In the algorithm, ``implies'', ``iff'', ``and'' are logical operators that take booleans (i.e.
+true or false) and return a boolean.
+Additionally, \(\&\!\&\) is the logical conjunction operation.
+It is the same function as the the written ``and''.
+We use \(\&\!\&\), instead of ``and'', to separate the conditions for clarity; logically \(\&\!\&\) is the same function as ``and'', but syntactically binds less tightly.
+
+Intuitively, the algorithm returns true if and only if a greedy bisimulation \(≈\) can be constructed s.t. \(s ≈ u\).
+Note the existence of such \(≈\) is not only equivalent to the trace equivalence of \(s\) and \(u\) (\Cref{thm:greedy-bisim-correctness}), but also equivalent to \(s ≊ u\), where \({≊} ⊆ S × U\) is the maximal bisimulation.
+However, our algorithm slightly deviates from the ideal scenario described above. Specifically, we implement two optimizations.
+First, optimization is to terminate and return false as soon as we know one of \(s, u\) is dead and the other is not.
+This is because in order to have \(s ≈ u\) for any greedy bisimulation \(≈\), \(s\) and \(u\) must have the same liveness in their respective GKAT coalgebra (\Cref{thm:greedy-bisim-perserve-liveness}).
+In the case where both \(s\) and \(u\) are dead in their respective coalgebra, we will immediate return true, because any dead states pair are necessarily in \(≊\) (\Cref{thm:bisim-between-dead,thm:max-bisim-is-max-greedy-bisim}).
+Then, we use a union-find data structure to organize the explored states into equivalence classes, which is justified by~\Cref{thm:greedy-bisim-iff-greedy-bisim-equiv}.
+By looking for a greedy bisimulation that is also an equivalence relation, we can use more efficient structures like union-find, instead of set of pairs; the correctness of this optimization is justified by.
+
+Concerning the complexity, our algorithm has similar worst case complexity as the original algorithm~\cite{smolka_GuardedKleeneAlgebra_2020}.
+In particular, when deciding the trace equivalence of two states \(s ∈ S\) and \(u ∈ U\), the original algorithm requires one pass of \(⟨s⟩\) and \(⟨u⟩\) to normalize them, and then visits each pair of states in \(⟨s⟩ × ⟨u⟩\) at most once during bisimulation.
+This means that the algorithm will visit at most \(|⟨s⟩ + ⟨u⟩ + ⟨s⟩ × ⟨u⟩|\) number of states in total.
+
+Our equivalence-checking algorithm attempts to find a bisimulation first, which visits each pair in \(⟨s⟩ × ⟨u⟩\) at most once. When a mismatch is found, the liveness-checking algorithm is invoked.
+Crucially, our algorithm satisfies the following property: if the liveness-checking algorithm finds a live state, the entire equivalence-checking algorithm will halt and return false.
 Thus, by caching all the known dead states, the liveness-checking only visits states in \(⟨s⟩\) and \(⟨u⟩\) at most once.
@@
 
-However, the on-the-fly algorithm will always out-perform the original algorithm, in terms of number of states visited, because this algorithm only invoke liveness check when necessary.
-In the extreme case when the two input states are infinite-trace equivalent, the on-the-fly algorithm can skip liveness checking entirely.
+Although both algorithms have similar worst case complexity, the on-the-fly algorithm will always out-perform the original algorithm in non-worst case situations.
+This is because the on-the-fly algorithm only invokes liveness checking when necessary.
+In the extreme case when the two input states are infinite-trace equivalent, the on-the-fly algorithm can even skip liveness checking entirely.
 
@@
 
-Although \cref{alg:bisim} is on-the-fly, it still uses GKAT coalgebra, which contains exponentially many transitions with respect to the number of primitive tests \(|T|\).
+Although \Cref{alg:bisim} is on-the-fly, it still uses GKAT coalgebra, which contains exponentially many transitions with respect to the number of primitive tests \(|T|\).
 Concretely, the transition function \(δ: S → \At_T → \{\accept, \reject\} + S × K\) requires computing the transition result for each \emph{atom}, and the number of atoms \(\At_T ≅ 2^{T}\) is exponential to the size of primitive tests in \(T\).
@@
 Symbolic GKAT coalgebra, instead of computing the behavior of each atom individually, groups atoms into boolean expressions. 
-This optimization leads to space-efficient coalgebras and an equivalence checking algorithm making use of off-the-shelf SAT solvers.
+This optimization leads to space-efficient coalgebras and an equivalence checking algorithm that can make use of off-the-shelf SAT solvers.
 Specifically, given a set of primitive actions \(K\) and primitive tests \(T\), a \emph{symbolic GKAT coalgebra} \(Ŝ ≜ (S, ϵ̂, δ̂)\) consists of a state set \(S\) and an accepting function \(ϵ̂\) and a transition function \(δ̂\):
@@
 \end{mathpar}
-This coalgebra is also required to satisfy the disjointedness condition, i.e.
-for all states \(s ∈ S\), the boolean expressions that \(s\) accepts \(ϵ̂(s)\) and the boolean expressions that enables \(s\) to transition \(\{b ∣ ∃ (b, s', p) ∈ δ(s)\}\) is disjoint; also the conjunction of any two distinct expressions from the set \(ϵ̂(s) ∪ \{b ∣ ∃ (b, s', p) ∈ δ(s)\}\) is equivalent to \(0\) under boolean algebra.
+This coalgebra is required to satisfy the disjointedness condition, i.e.
+for all states \(s ∈ S\), the boolean expressions that \(s\) accepts \(ϵ̂(s)\) and the boolean expressions that enables \(s\) to transition \(\{b ∣ ∃ (b, s', p) ∈ δ(s)\}\) are disjoint; also the conjunction of any two distinct expressions from the set \(ϵ̂(s) ∪ \{b ∣ ∃ (b, s', p) ∈ δ(s)\}\) is equivalent to \(0\) under boolean algebra.
 
@@
     These two symbolic GKAT coalgebra \(Ŝ₁ ≜ (S, δ̂₁, ϵ̂₁)\) and \(Ŝ₂ ≜ (S, δ̂₂, ϵ̂₂)\) have the same lowering and semantics, yet, they are different.
-    It is possible to construct symbolic representations that satisfies canonicity, yet we opt to use our current representation for ease of construction and computational efficiency.
+    It is possible to construct symbolic representations that satisfy canonicity, yet we opt to use our current representation for ease of construction and computational efficiency.
 \end{remark}
@@
 
-The functoriality states that a homomorphism on two symbolic coalgebras induces a homomorphism of on their lowing; similarly, a bisimulation on symbolic GKAT coalgebra also induces a bisimulation on their lowering.
+Functoriality states that a homomorphism on two symbolic coalgebras induces a homomorphism of on their lowering; similarly, a bisimulation on symbolic GKAT coalgebra also induces a bisimulation on their lowering.
 However, the converse is not true, precisely because of the canonicity problem noted in~\Cref{rem:canonicity}: take the \(Ŝ₁\) and \(Ŝ₂\) in~\Cref{rem:canonicity}, because they have the same lowering, therefore the identity homomorphism is a homomorphism on their lowerings, but there is no homomorphism from \(Ŝ₁\) to \(Ŝ₂\).
@@
 
-Similar to the non-symbolic version of the algorithm, \cref{alg:symb-bisim} first check if either of the input states is dead, then  recursively check all the conditions in~\Cref{thm:symb-recursive-construction}, where \(\mathrel{\&\!\&}\) denotes the logical and operator on boolean data type.
-The symbolic version of the liveness checking algorithm is also based on a graph search algorithm on all the reachable state, where we try to identify an accepting state by checking whether \(⋁ ϵ̂(s) ≡ 0\).
-The only deviation from the non-symbolic cases is that we do not follow empty transitions in the symbolic case, i.e.
+Similar to the non-symbolic version of the algorithm, \Cref{alg:symb-bisim} first checks if either of the input states is dead, then  recursively checks all the conditions in~\Cref{thm:symb-recursive-construction}, where \(\mathrel{\&\!\&}\) denotes the logical conjunction operator on booleans.
+The symbolic version of the liveness checking algorithm is also based on a graph search algorithm on all the reachable state where we try to identify an accepting state by checking whether \(⋁ ϵ̂(s) ≡ 0\).
+The only deviation from the non-symbolic case is that we do not follow empty transitions in the symbolic case, i.e.
 if \(δ̂(s) ≜ (b, s', p)\) and \(b ≡ 0\), we will not search \(s'\) to check the liveness of \(s'\).
-In~\Cref{sec:optimization-implementation}, we will outline a simple modification for the symbolic GKAT coalgebra generation algorithm in~\Cref{sec:symb-gkat-construction} to avoid generating empty transitions, allowing us to skip the emptiness check when deciding whether a state is dead.
-TODO: I think we need to make sure that we are presenting the dead state checking algorithm ignoring the empty transition.
-After which we can comment it here, to claim that this is for generality.
+In~\Cref{sec:optimization-implementation}, we will outline a simple modification for the symbolic GKAT coalgebra generation algorithm in~\Cref{sec:symb-gkat-construction} to avoid generating empty transitions, allowing us to skip the emptiness check when deciding whether a state is dead or not.
+% TODO: I think we need to make sure that we are presenting the dead state checking algorithm ignoring the empty transition.
+% After which we can comment it here, to claim that this is for generality.
 
@@
 This goal can be achieved by lifting existing constructions like derivatives and Thompson's construction~\cite{schmid_GuardedKleeneAlgebra_2021,smolka_GuardedKleeneAlgebra_2020} to the symbolic setting.
-Then the correctness of non-symbolic version can be used to show the correctness of the symbolic construction i.e.
-we will prove that the lowering as shown in construction~\labelcref{cons:lowering} of these constructions will yield the conventional derivative.
-However, several other important properties, like finiteness of derivative coalgebra and the correctness of the Thompson's construction can be established using a symbolic GKAT coalgebra homomorphism from the Thompson's construction to the derivative.
+Then the correctness of the non-symbolic construction can be used to show the correctness of the symbolic construction i.e.
+we will prove that the lowering as shown in construction~\labelcref{cons:lowering} will yield the conventional derivative.
+Other important properties, like finiteness of derivative coalgebra and the correctness of the Thompson's construction can be established using a symbolic GKAT coalgebra homomorphism from the Thompson's construction to the derivative.
 
@@
 The symbolic derivative coalgebra \(D̂\), with expressions as states, is the least symbolic GKAT coalgebra (ordered by point-wise subset ordering on \(ϵ̂\) and \(δ̂\)) that satisfy the rules in~\Cref{fig:derivatives-rules}.
-In more plain words, a transition is in \(D̂\) if and only if it is derivable by the rules in~\Cref{fig:derivatives-rules}.
+Basically, a transition is in \(D̂\) if and only if it is derivable by the rules in~\Cref{fig:derivatives-rules}.
 These rules are very close to the rules of GKAT derivative by~\citeauthor{schmid_GuardedKleeneAlgebra_2021}~\cite{schmid_GuardedKleeneAlgebra_2021}.
-This is no coincidence, as our definition exactly lowers to the definition of theirs.
-This fact can be proven by case analysis on the shape of the source expression, and forms a basis on our correctness argument.
+This is no coincidence as our definition exactly lowers to the definition of theirs.
+This fact can be proven by case analysis on the shape of the source expression, and forms a basis for our correctness argument.
 \begin{theoremEnd}{theorem}[Correctness]\label{thm:derivative-correctness}
@@
 
-Another way to construct a coalgebra from a GKAT expression is via Thompson's construction, we lift the original construction to the symbolic setting.
+Another way to construct a coalgebra from a GKAT expression is via Thompson's construction. We lift the original construction to the symbolic setting.
 A useful operation for Thompson's construction is the following guard operation, denoted by \(⟨B|\), where \(B\) is a set of boolean expressions: for a transition function \(δ̂\), a accepting function \(ϵ̂\), and a state \(s\),
@@
 \end{align*}
-Besides guarding transition and acceptance with different conditions in if statements and while loops, the guard operator can also be used to simulate uniform continuation.
-Specifically, we can use \(⟨ϵ̂(s)|δ(s')\) to connecting all the accepting transition of \(s\) to the dynamic \(δ(s')\).
-
-With these definitions in mind, we can define symbolic Thompson's construction inductively as in~\Cref{tab:symb-Thompson-construction}, where we let \((S₁, ϵ̂₁, δ̂₁)\) and \((S₂, ϵ̂₂, δ̂₂)\) to be result of Thompson's construction for \(e₁\) and \(e₂\) respectively.
+Besides guarding transition and acceptance with different conditions in if-statements and while-loops, the guard operator can also be used to simulate uniform continuation.
+Specifically, we can use \(⟨ϵ̂(s)|δ(s')\) to connect all the accepting transition of \(s\) to the dynamic \(δ(s')\).
+
+With these definitions in mind, we can define symbolic Thompson's construction inductively as in~\Cref{tab:symb-Thompson-construction}, where we let \((S₁, ϵ̂₁, δ̂₁)\) and \((S₂, ϵ̂₂, δ̂₂)\) to be results of Thompson's construction for \(e₁\) and \(e₂\) respectively.
 In this table, \(S₁ + S₂\) denotes the disjoint union of \(S₁\) and \(S₂\), and for any two transition dynamics \(δ₁(s₁): \mathscr{P}(\BExp × S₁ × K)\) and \(δ₂(s₂): \mathscr{P}(\BExp × S₂ × K)\), we can also compose them in parallel as \(δ₁(s₁) + δ₂(s₂)\):
@@
 This choice will make the proof of~\Cref{thm:hom-thompson-derivative} slightly easier. 
-However, in~\Cref{sec:optimization-implementation}, we will explain that our implementation uses start dynamics instead of start state, to avoid unnecessary lookups and unreachable states.
+However, as explained in~\Cref{sec:optimization-implementation}, our implementation uses start dynamics instead of start state to avoid unnecessary lookups.
 
 We would like to explore several desirable theoretical properties of both derivatives and Thompson's construction.
-Specifically, the \emph{correctness}, i.e.
-the semantics of the ``start state'' the both construction have the same preserves the trace semantics of the expression; \emph{finiteness}, i.e.
+Specifically, \emph{correctness}, i.e.
+the semantics of the ``start state'' of both constructions preserve the trace semantics of the expression; \emph{finiteness}, i.e.
 the coalgebra generated is always finite, implying the termination of our equivalence algorithm; and finally, \emph{complexity}, i.e.
 the relationship between the number of reachable states and the size of the input expression, which serves as an estimated complexity of our equivalence checking algorithm.
-Turns out, all of these questions can be answered by a homomorphism from symbolic Thompson's construction to the symbolic derivatives.
+It turns out, all of these questions can be answered by a homomorphism from symbolic Thompson's construction to the symbolic derivatives.
 
@@
 
-\Cref{thm:hom-thompson-derivative} have several consequences, one of the more obvious one is that we can use the functoriality of the lowering operation to show the semantic equivalence of the start state in the Thompson's construction and the expression in derivative.
+\Cref{thm:hom-thompson-derivative} has several consequences, one of the more obvious ones is that we can use the functoriality of the lowering operation to show the semantic equivalence of the start state in the Thompson's construction and the expression in derivative.
+% TODO: what is "in derivative"?
 
@@
 \begin{proofEnd}
-    By functoriality of lowering~\cref{thm:lowering-functor}, then \(h: S_{e} → D\) is a homomorphism on their lowerings. 
+    By functoriality of lowering~\Cref{thm:lowering-functor}, then \(h: S_{e} → D\) is a homomorphism on their lowerings. 
     Because homomorphism preserves semantics (\Cref{thm:hom-preserves-semantics}) \(⟦s^*⟧^{ω}_{S_{e}} = ⟦h(s^*)⟧^{ω}_{D} = ⟦e⟧^{ω}_{D}\), where \(⟦-⟧^{ω}\) is the infinite trace semantics i.e.
@@
 
-A not so obvious consequence of the homomorphism in~\Cref{thm:hom-thompson-derivative}, is the complexity of the algorithm based on derivatives.
+A not so obvious consequence of the homomorphism in~\Cref{thm:hom-thompson-derivative} is the complexity of the algorithm based on derivatives.
 Our bisimulation algorithm (\Cref{alg:symb-bisim}) only explores the principle sub-coalgebra of the start state, i.e.
-\(s^*\) in the Thompson's construction \(Ŝ_{e}\) or \(e\) in the derivative \(D̂\); thus, deducing an upper bound on the size of the principle sub-coalgebras \(⟨s^*⟩_{Ŝ_{e}}\) and \(⟨e⟩_{D̂}\) are crucial to our complexity analysis.
+\(s^*\) in the Thompson's construction \(Ŝ_{e}\) or \(e\) in the derivative \(D̂\); thus, deducing upper bounds on the size of the principle sub-coalgebras \(⟨s^*⟩_{Ŝ_{e}}\) and \(⟨e⟩_{D̂}\) is crucial to our complexity analysis.
 An upper bound on \(⟨s^*⟩_{Ŝ_{e}}\) is easy to obtain, as the size of \(Ŝ_{e}\), which subsumes the states of \(⟨s^*⟩_{Ŝ_{e}}\), is linear to the size of expression \(e\); therefore \(⟨s^*⟩_{Ŝ_{e}}\) is at most linear to the size of the expression \(e\).
-On the other hand the size of \(⟨e⟩_{D̂}\) can, again, be derived from the homomorphism in~\cref{thm:hom-thompson-derivative}.
+On the other hand the size of \(⟨e⟩_{D̂}\) can, again, be derived from the homomorphism in~\Cref{thm:hom-thompson-derivative}.
 
@@
 
-Although the size of the coalgebra generated by the derivative is smaller than Thompson's construction, the decision procedure based on derivative is not always more efficient than those based on Thompson's construction.
-Crucially, the states in the derivative coalgebra \(D̂\) are expressions, which are more expensive to store; and computing the next transition of the coalgebra can also be more computationally expensive.
-Whereas, Thompson's construction only requires inductively going through the expression once to construct the entire coalgebra \(Ŝ_{e}\), and its states can be represented by more efficient constructs, like integers.
+Although the size of the coalgebra generated by the derivative is smaller than Thompson's construction, the states in the derivative coalgebra \(D̂\) are expressions which can be more expensive to store.
+Computing the next transition of the derivative coalgebra can also take longer.
+On the other hand, the states of Thompson's construction only requires inductively going through the expression once to construct the entire coalgebra \(Ŝ_{e}\), and its states can be represented by more efficient constructs, like integers.
+In practice, we utilize techniques such as hashconsing and memoization to address the memory and time usage problems of derivative construction respectively.
 

### 2025-01-24 main.tex
symbolic coalgebra

@@
 
-% \section{The Algorithm}
-
-% In this section we will present the pseudo-code for our on-the-fly algorithm. 
-% In order to implement the the inductive construction theorem (\cref{thm:recursive-construction}), we will need to determine the liveness of the state.
-This can be simply computed via a DFS from the state being checked.
-
-% TODO: we should merge the two so that it is easier to 
-% \begin{algorithm}
-%     \caption{Check whether a state \(s\) is dead}\label{alg:check-dead-main}
-%     \begin{algorithmic}
-%         \Function{isDeadLoop}{$s ∈ S$, explored}
-%         \If {\(s ∈\) explored} {\Return explored} 
-%         \Else { 
-%             \For{\(α ∈ \At\)}{}{
-%                 \Match{\(δ_{S}(s, α)\)}
-%                 \Case{\(\accept\)} {\Return none} \Comment{\(s\) transition to accept}
-%                 \EndCase
-%                 \Case{\(\reject\)} {\texttt{continue}}
-%                 \Comment{skip if \(s\) transition to reject}
-%                 \EndCase
-%                 \Case{($s', p$)}{ 
-%                     \If{\Call{IsDeadLoop}{$s'$} = none} {\Return none} \Comment{\(s\) transitions to a live state \(s'\)}
-%                     \Else { explored \(←\) (explored \(∪\) \Call{isDeadLoop}{$s'$, explored}) } \EndIf
-%                 } \EndCase
-%                 \EndMatch
-%             }\EndFor}
-%         \EndIf
-%         \State {\Return explored}    
-%         \EndFunction
-%     \end{algorithmic}
-% \end{algorithm}
-
-% By~\cref{thm:dead-iff-all-reachable-dead}, if \(s\) is dead then all the reachable states of \(s\) (denoted by \(⟨s⟩\)).
-Then by returning all the reachable states of \(s\), we can cache these states to avoid checking them again.
-To encapsulate the caching, we have the following function, which we will actually use in our bisimulation algorithm.
-
-% \begin{algorithm}
-%     \caption{A cached algorithm to check whether a state is dead}\label{alg:is-dead}
-%     \begin{algorithmic}
-%         \State{deadStates \(← ∅\)}
-
-%         \Function{isDead}{$s ∈ S$}
-%         \If {\(s ∈\) deadStates} {\Return true} 
-%         \ElsIf {\Call{isDeadLoop}{$s, ∅$} = none} {\Return false}
-%         \Else 
-%             \State {deadStates \(←\) (deadStates \(∪\) \Call{isDeadLoop}{$s, ∅$})}
-%             \State {\Return {true}}
-%         \EndIf
-%         \EndFunction
-%     \end{algorithmic}
-% \end{algorithm}
-
-% Given the direct correspondence between bisimulation and bisimulation equivalence and bisimulation in sub-algebra:
-% \begin{align*}
-%     & ∃ \text{ bisimulation } {∼} ⊆ ⟨s⟩ × ⟨t⟩ \text{ s.t. } s ∼ t \\
-%     & ⟺ ∃ \text{ bisimulation } {∼} ⊆ (⟨s⟩∪⟨t⟩) × (⟨s⟩∪⟨t⟩) \text{ s.t. } s ∼ t 
-%         & \text{\cref{thm:sub-coalg-preserve-bisim}}\\  
-%     & ⟺ ∃ \text{ bisimulation equivalence } {≃} ⊆ (⟨s⟩∪⟨t⟩) × (⟨s⟩∪⟨t⟩) \text{ s.t. } s ≃ t 
-%         & \text{\cref{thm:bisim-iff-bisim-equiv}}
-% \end{align*}
-% we can safely replace the bisimulation in inductive construction (\cref{thm:recursive-construction}) with bisimulation equivalence. 
-% Dealing with equivalence relations allows us to leverage efficient data structures like union find in our bisimulation algorithm.
-
-% We will use \(\Call{union}{$s,t$}\) to denote the operation to equate \(s\) and \(t\) in a union-find, and use \(\Call{eq}{$s,t$}\) to check if \(s\) and \(t\) belongs to the same equivalence class, i.e.
-share the same representative.
-% Specifically, we will use the union-find structures to keep track of the equivalence classes that we are in the process of checking, hence avoiding repeatedly checking the same pair of states to remove infinite loops.
-
-% Our on-the-fly bisimulation algorithm will decide whether there exists a bisimulation relation in \(⟨s⟩ ∪ ⟨t⟩\) s.t.
-\(s ∼ t\).
-This algorithm generally reproduce the setting of inductive construction theorem~\cref{thm:recursive-construction};
-% except by~\cref{thm:bisim-one-dead}, in the special case where \(s\) or \(t\) is dead, then we will only need to check whether the other is dead.
-
-% Because the dead state detection algorithm is coalgebra-specific, we use a subscript on ``deadStates'' and ``\textsc{IsDead}'' to indicate the coalgebra. 
-% The soundness and completeness of~\cref{alg:bisim} can be observed by the fact that \emph{when the algorithm terminate}, the algorithm returns true if and only if there exists a bisimulation between \(⟨s⟩\) and \(⟨t⟩\) s.t.
-\(s ∼ t\), which is then logically equivalent to trace equivalence.
-% Such equivalence is a direct consequence of~\cref{thm:bisim-one-dead,thm:recursive-construction}.
-
-% \begin{remark}
-%     The caching of dead state and the shortcut to check whether \(s\) is dead when \(t\) is dead and vise versa, is not essential to the soundness and completeness of algorithm, they are here to trade speed with memory.
-%     In a memory-constraint situation, the ``\textnormal{deadStates}'' variable can be cleared periodically to save memory.
-% \end{remark}
-
 \section{Symbolic Coalgebra and Algorithm}\label{sec:symb-gkat-greedy-bisim}
@@
 This optimization leads to space-efficient coalgebras and an equivalence checking algorithm that can make use of off-the-shelf SAT solvers.
-Specifically, given a set of primitive actions \(K\) and primitive tests \(T\), a \emph{symbolic GKAT coalgebra} \(Ŝ ≜ (S, ϵ̂, δ̂)\) consists of a state set \(S\) and an accepting function \(ϵ̂\) and a transition function \(δ̂\):
+Specifically, given a set of primitive actions \(K\) and primitive tests \(T\), a \emph{symbolic GKAT coalgebra} \(Ŝ ≜ (S, ϵ̂, δ̂)\) consists of a state set \(S\), an accepting function \(ϵ̂\), and a transition function \(δ̂\):
 \begin{mathpar}
@@
 \end{mathpar}
-This coalgebra is required to satisfy the disjointedness condition, i.e.
-for all states \(s ∈ S\), the boolean expressions that \(s\) accepts \(ϵ̂(s)\) and the boolean expressions that enables \(s\) to transition \(\{b ∣ ∃ (b, s', p) ∈ δ(s)\}\) are disjoint; also the conjunction of any two distinct expressions from the set \(ϵ̂(s) ∪ \{b ∣ ∃ (b, s', p) ∈ δ(s)\}\) is equivalent to \(0\) under boolean algebra.
+This coalgebra is required to satisfy the \emph{disjointedness condition}, i.e.
+for all states \(s ∈ S\), the boolean expressions that \(s\) accepts \(ϵ̂(s)\) and the boolean expressions that enables \(s\) to transition \(\{b ∣ ∃ (b, s', p) ∈ δ(s)\}\) are disjoint; i.e.
+the conjunction of any two distinct expressions from the set \(ϵ̂(s) ∪ \{b ∣ ∃ (b, s', p) ∈ δ(s)\}\) is equivalent to \(0\) under Boolean Algebra.
+Like the non-symbolic case, we use \(s \transAcc{Ŝ}{b}\) to indicate \(b ∈ ϵ̂(s)\) and \(s \transvia{b ∣ p}_{Ŝ} s'\) to indicate \((b, p, s') ∈ δ̂(s)\).
 
@@
 \end{align}
-This is well-defined, i.e. exactly one clause can be satisfied for any \(s ∈ S\) and \(α ∈ \At\), because of the disjointedness condition.
+The function \(δ\) is well-defined, i.e. exactly one clause can be satisfied for any \(s ∈ S\) and \(α ∈ \At\), because of the disjointedness condition.
 We usually use \(S\) to denote the lowering of \(Ŝ\); and the semantics of a state \(s ∈ Ŝ\) is defined as its semantics in the lowering \(⟦s⟧_{Ŝ} ≜ ⟦s⟧_{S}.\)
@@
     \end{align*} 
-    and both \(ϵ̂₁, ϵ̂₂\) will return constant \(0\).
+    and both \(ϵ̂₁, ϵ̂₂\) return constant \(0\).
     These two symbolic GKAT coalgebra \(Ŝ₁ ≜ (S, δ̂₁, ϵ̂₁)\) and \(Ŝ₂ ≜ (S, δ̂₂, ϵ̂₂)\) have the same lowering and semantics, yet, they are different.
@@
 \begin{theoremEnd}{theorem}[Symbolic Recursive Construction]\label{thm:symb-recursive-construction}
-    Given two symbolic GKAT coalgebra \(Ŝ = (S, ϵ̂_S, δ̂_S)\) and \(Û = (U, ϵ̂_U, δ̂_U)\) and two states \(s ∈ S\) and \(u ∈ U\), \(s\) and \(u\) are trace equivalent \(⟦s⟧_{Ŝ} = ⟦u⟧_{Û}\), if and only if all the following holds:
+    Given two symbolic GKAT coalgebra \(Ŝ = (S, ϵ̂_S, δ̂_S)\) and \(Û = (U, ϵ̂_U, δ̂_U)\) and two states \(s ∈ S\) and \(u ∈ U\), \(s\) and \(u\) are trace equivalent \(⟦s⟧_{Ŝ} = ⟦u⟧_{Û}\), if and only there exists a relation \({≈} ⊆ Ŝ × Û\) s.t.
+the following holds:
     \begin{itemize}
         \item \(⋁ ϵ̂_S(s) ≡ ⋁ ϵ̂_U(u)\); 
-        \item for all \((b, s', p) ∈ δ̂_S(s)\) and \(c ∈ ρ̂_U(u)\), if \(b ∧ c ≢ 0\), then \(s'\) is dead;
-        \item for all \(b ∈ ρ̂_S(s)\) and \((c, u', q) ∈ δ̂_U(u)\), if \(b ∧ c ≢ 0\), then \(u'\) is dead;
-        \item for all \((b, s', p) ∈ δ̂_S(s)\) and \((c, u', q) ∈ δ̂_U(u)\), if \(b ∧ c ≢ 0\) and \(p ≠ q\) then both \(s'\) and \(u'\) is dead; 
-        \item for all \((b, s', p) ∈ δ̂_S(s)\) and \((c, u', q) ∈ δ̂_U(u)\), if \(b ∧ c ≢ 0\) and \(p = q\) then \(⟦s'⟧_{Ŝ} = ⟦u'⟧_{Û}\).
+        \item if \(s \transvia{b ∣ p}_{Ŝ} s'\) and \(b ∧ ρ̂_U(u) ≢ 0\), then \(s'\) is dead;
+        \item if \(u \transvia{c ∣ p}_{Û} u'\) and \(ρ̂_S(s) ∧ c ≢ 0\), then \(u'\) is dead;
+        \item if \(s \transvia{b ∣ p}_{Ŝ} s'\), \(u \transvia{c ∣ p}_{Û} u'\), and \(b ∧ c ≢ 0\), then \(s' ≈ u'\); 
+        \item if \(s \transvia{b ∣ p}_{Ŝ} s'\), \(u \transvia{c ∣ q}_{Û} u'\), \(p ≠ q\), and \(b ∧ c ≢ 0\), then \(s'\) and \(u'\) are both dead.
     \end{itemize}
@@
 
-Similar to the non-symbolic version of the algorithm, \Cref{alg:symb-bisim} first checks if either of the input states is dead, then  recursively checks all the conditions in~\Cref{thm:symb-recursive-construction}, where \(\mathrel{\&\!\&}\) denotes the logical conjunction operator on booleans.
+Similar to the non-symbolic version of the algorithm, \Cref{alg:symb-bisim} first checks if either of the input states is dead, then recursively checks all the conditions in~\Cref{thm:symb-recursive-construction}, and organize the explored state in equivalence classes for efficiency.
+The operator ``and'', ``implies'', and \(\mathrel{\&\!\&}\) denotes the corresponding logical operator as in~\Cref{alg:bisim}.
 The symbolic version of the liveness checking algorithm is also based on a graph search algorithm on all the reachable state where we try to identify an accepting state by checking whether \(⋁ ϵ̂(s) ≡ 0\).

### 2025-01-24 main.tex
automata construction

@@
 
-The final piece of the puzzle is to convert any given expression into an \emph{equivalent} state in some symbolic GKAT coalgebra. 
+The final piece of the puzzle is to convert any GKAT expression into an \emph{equivalent} state in some symbolic GKAT coalgebra. 
 This goal can be achieved by lifting existing constructions like derivatives and Thompson's construction~\cite{schmid_GuardedKleeneAlgebra_2021,smolka_GuardedKleeneAlgebra_2020} to the symbolic setting.
 Then the correctness of the non-symbolic construction can be used to show the correctness of the symbolic construction i.e.
-we will prove that the lowering as shown in construction~\labelcref{cons:lowering} will yield the conventional derivative.
+we will prove that the lowering of symbolic derivatives will yield the conventional derivative.
 Other important properties, like finiteness of derivative coalgebra and the correctness of the Thompson's construction can be established using a symbolic GKAT coalgebra homomorphism from the Thompson's construction to the derivative.
-
-We introduce some new notations for the transitions of symbolic GKAT coalgebra: for a symbolic GKAT coalgebra \(Ŝ ≜ (S, δ̂, ϵ̂)\) and state \(s ∈ S\), we will use \(s \transAcc{S}{b}\) to denote \(b ∈ ϵ̂(s)\); and use \(s \transvia{b ∣ p}_S s'\) to denote \((b, s', p) ∈ δ̂(s)\).
 
@@
 
-The symbolic derivative coalgebra \(D̂\), with expressions as states, is the least symbolic GKAT coalgebra (ordered by point-wise subset ordering on \(ϵ̂\) and \(δ̂\)) that satisfy the rules in~\Cref{fig:derivatives-rules}.
-Basically, a transition is in \(D̂\) if and only if it is derivable by the rules in~\Cref{fig:derivatives-rules}.
-These rules are very close to the rules of GKAT derivative by~\citeauthor{schmid_GuardedKleeneAlgebra_2021}~\cite{schmid_GuardedKleeneAlgebra_2021}.
-This is no coincidence as our definition exactly lowers to the definition of theirs.
+The symbolic derivative coalgebra \(D̂\), with expressions as states, is the symbolic GKAT coalgebra generated by the rules in~\Cref{fig:derivatives-rules} i.e.
+a transition is in \(D̂\) if and only if it is derivable.
+Our symbolic derivative resembles the GKAT derivative by~\citeauthor{schmid_GuardedKleeneAlgebra_2021}~\cite{schmid_GuardedKleeneAlgebra_2021}.
+This similarity is no coincidence, as our definition exactly lowers to the definition of theirs.
 This fact can be proven by case analysis on the shape of the source expression, and forms a basis for our correctness argument.
@@
 \end{theoremEnd}
-\begin{proofEnd}
-    TODO: unfold the statement.
-\end{proofEnd}
+% \begin{proofEnd}
+%     TODO: unfold the statement.
+% \end{proofEnd}
 
@@
 
-Another way to construct a coalgebra from a GKAT expression is via Thompson's construction. We lift the original construction to the symbolic setting.
-A useful operation for Thompson's construction is the following guard operation, denoted by \(⟨B|\), where \(B\) is a set of boolean expressions: for a transition function \(δ̂\), a accepting function \(ϵ̂\), and a state \(s\),
+Similarly, we also lift the Thompson's construction~\cite{smolka_GuardedKleeneAlgebra_2020} to the symbolic setting.
+A useful operation for Thompson's construction is the following guard operation, denoted by \(⟨B|\), where \(B\) is a set of boolean expressions: for a transition function \(δ̂\), an accepting function \(ϵ̂\), and a state \(s\),
 \begin{align*}
@@
 \end{align*}
-Besides guarding transition and acceptance with different conditions in if-statements and while-loops, the guard operator can also be used to simulate uniform continuation.
+Besides guarding transition and acceptance with conditions for if-statements and while-loops, the guard operator can also be used to simulate uniform continuation. 
 Specifically, we can use \(⟨ϵ̂(s)|δ(s')\) to connect all the accepting transition of \(s\) to the dynamic \(δ(s')\).
@@
 This choice will make the proof of~\Cref{thm:hom-thompson-derivative} slightly easier. 
-However, as explained in~\Cref{sec:optimization-implementation}, our implementation uses start dynamics instead of start state to avoid unnecessary lookups.
+However, as explained in~\Cref{sec:optimization-implementation}, our implementation still utilizes start dynamics to avoid unnecessary lookups.
 
@@
 
-\Cref{thm:hom-thompson-derivative} has several consequences, one of the more obvious ones is that we can use the functoriality of the lowering operation to show the semantic equivalence of the start state in the Thompson's construction and the expression in derivative.
+\Cref{thm:hom-thompson-derivative} can be proven by unfolding the definition, and implies several properties we desire. 
+One of the more obvious ones is the semantic equivalence of the start state in the Thompson's construction and the expression in the derivative coalgebra. 
 % TODO: what is "in derivative"?
@@
 
-Because of the surjectivity of \(h'\) in~\Cref{thm:suj-hom-thompson-derivative}, \(⟨s^*⟩_{Ŝ_{e}}\) has no fewer states than \(⟨e⟩_{D̂}\).
-And, the number of states in \(Ŝ_{e}\) is greater than \(⟨s^*⟩_{Ŝ_{e}}\), which is greater than \(⟨e⟩_{D̂}\); by induction, the number of states in \(Ŝ_{e}\) is linear to the size of the input expression, therefore the number of states in \(⟨e⟩_{D̂}\) is at most linear to the size of the expression \(e\).
-
-Although the size of the coalgebra generated by the derivative is smaller than Thompson's construction, the states in the derivative coalgebra \(D̂\) are expressions which can be more expensive to store.
+Because of the surjectivity of \(h'\), \(⟨s^*⟩_{Ŝ_{e}}\) has no fewer states than \(⟨e⟩_{D̂}\).
+Because the number of states in \(⟨s^*⟩_{Ŝ_{e}}\) is bounded by \(Ŝ_{e}\), and \(|Ŝ_{e}|\) is linear to the size of the input expression, the number of states in \(⟨e⟩_{D̂}\) is at most linear to the size of the expression \(e\).
+
+% Although the size of the coalgebra generated by the derivative is smaller than Thompson's construction, the states in the derivative coalgebra \(D̂\) are expressions which can be more expensive to store.
 Computing the next transition of the derivative coalgebra can also take longer.

### 2025-01-24 main.tex
implementation

@@
 \section{Implementation and Evaluation}\label{sec:implementation}
-We implement our symbolic on-the-fly bisimulation algorithm~(\Cref{alg:symb-bisim}) in Rust along with derivative based~(\Cref{fig:derivatives-rules}) and Thompson's construction based~(\Cref{tab:symb-Thompson-construction}) algorithms for constructing symbolic GKAT coalgebras.
+We implement our symbolic on-the-fly bisimulation algorithm~(\Cref{alg:symb-bisim}) in Rust along with derivative-based~(\Cref{fig:derivatives-rules}) and Thompson's-construction-based~(\Cref{tab:symb-Thompson-construction}) algorithms for constructing symbolic GKAT coalgebras.
 The source code and benchmark suite is freely available in our repository \url{https://anonymous.4open.science/r/rust-gkat-071E}.
@@
 
-Notice that during bisimulation~(\Cref{alg:symb-bisim}) for some \(s \in S\) and \(u \in U\), if there is \(s \transvia{b ∣ p} s'\), i.e.
-\((b,s',p) \in \hat{\delta}_S(s)\), where \(b\) is a semantically false boolean expression, then implications with premises of the form \((b \land \cdot) \not{\equiv} 0\) are trivially satisfied.
+Notice that during bisimulation~(\Cref{alg:symb-bisim}) for some \(s \in S\) and \(u \in U\), if \(s \transvia{b ∣ p} s'\) and \(b\) is equivalent to \(0\), then implications with premises of the form \((b \land \cdot) \not{\equiv} 0\) are trivially satisfied.
 In other words, \textit{blocked} transitions in \(S\) do not contribute to the result of bisimulation.
@@
 
-As mentioned in \Cref{sec:symb-gkat-construction}, our implementation of Thompson's construction uses start dynamics (pseudo-state) instead of an explicit start state.
-This reduces the number of indirections needed to access the behaviors \(\hat{\epsilon}(s^*) \) and \(\hat{\delta}(s^*)\) of start state \(s^*\).
-More specifically, we track the behaviors \(\epsilon^*\) and \(\delta^*\) which characterize \(\hat{\epsilon}(s^*) \) and \(\hat{\delta}(s^*)\) respectively.
-The start state \(s^*\) does not appear explicitly during coalgebra construction.
-After the construction algorithm terminates, we generate a fresh state \(s^*\) and set \(\hat{\epsilon}(s^*) = \epsilon^*\) and \(\hat{\delta}(s^*) = \delta^*\).
-The final coalgebra that our implementation produces is equivalent to what would have been obtained from the original algorithm (\Cref{tab:symb-Thompson-construction}) modulo consistent renaming of states.
+Our implementation of Thompson's construction also uses start dynamics to reduces the number of lookups when accessing the dynamics \(\hat{\epsilon}(s^*) \) and \(\hat{\delta}(s^*)\) of start state \(s^*\).
+Concretely, we use two extra fields \(\epsilon^*\) and \(\delta^*\) to track \(\hat{\epsilon}(s^*) \) and \(\hat{\delta}(s^*)\) respectively. 
+By tracking \(ϵ^*\) and \(δ^*\), we can remove \(s^*\) from the coalgebra construction, only adding \(s^*\) back as a fresh state after the construction terminates. 
+The resulting coalgebra is equivalent \(⟨s^*⟩_{Ŝ_{e}}\) from \Cref{tab:symb-Thompson-construction}.
 
@@
 If \(\hat{\epsilon_1}(s_1)\) and \(\hat{\epsilon_2}(s_2)\) are sets of boolean expressions, then the guard operation would produce a set whose size is proportional to \(|\hat{\epsilon_1}(s_1)| \times |\hat{\epsilon_2}(s_2)|\).
-However, if the collapsed representation is used, \(\hat{\epsilon_1}(s_1)\) and \(\hat{\epsilon_2}(s_2)\) become just boolean expressions instead of sets of expressions.
-The guard operation now produces a single boolean expression.
-The distributivity of boolean conjunction and disjunction ensures the correctness of this optimization.
+However, if the collapsed representation is used, \(⟨\hat{\epsilon_1}(s_1)|\hat{\epsilon_2}(s_2)\) will simply be a conjunction of \(\hat{\epsilon_1}(s_1)\) and \(\hat{\epsilon_2}(s_2)\).
+This optimization is indeed sound because of the distributivity of boolean algebra.
 
@@
 \Cref{tab:benchmark-time,tab:benchmark-memory} present the experimental results of equivalence checking based on derivatives (DV), Thompson's construction (TC) and SymKAT~\cite{pous_SymbolicAlgorithmsLanguage_2015} (SK).
-Note that SymKAT uses BDD internally for checking boolean satisfiability.
 Each benchmark consists of 50 expression pairs with properties described by the benchmark's name.

### 2025-01-24 main.tex
remove todo and go through related works

@@
     Then because bisimulation is closed under union~\cite[Theorem 5.5]{rutten_UniversalCoalgebraTheory_2000}, \(\norm(∼_d) = {∼} ∪ ⋃ \{∼_{s, u} ∣ \text{\(s, u\) are dead}\}\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\).
-    TODO: bisimulation are uniquely determined by its carrier.
+    % TODO: bisimulation are uniquely determined by its carrier.
 \end{proofEnd}
@@
 
-\begin{proofEnd}
-    Since \(Ŝ\) and \(Û\) have the same states as their lowering, therefore \(h: S → U\) is indeed a function, then we only need to verify the homomorphism condition on \(h\).
-    TODO: finish.
-\end{proofEnd}
+% \begin{proofEnd}
+%     Since \(Ŝ\) and \(Û\) have the same states as their lowering, therefore \(h: S → U\) is indeed a function, then we only need to verify the homomorphism condition on \(h\).
+%     TODO: finish.
+% \end{proofEnd}
 
@@
 Our study, on the other hand,  focus on GKAT and GKAT automata (represented as coalgebra throughout the paper). 
-Although previous works on deterministic symbolic transducer~\cite{saarikivi_FusingEffectfulComprehensions_2017,veanes_SymbolicFiniteState_2012} might seem similar to that of GKAT, the automata shape are subtly different, specifically, instead of having accept and rejecting states, states in GKAT automata will accept or reject its input.
+Although previous works on deterministic symbolic transducer~\cite{saarikivi_FusingEffectfulComprehensions_2017,veanes_SymbolicFiniteState_2012} might seem similar to that of GKAT, the automata shape are subtly different, specifically, instead of having accept and rejecting states, a state in GKAT automata can accept or reject based on the input atoms.
 
@@
 This approach allows us to leverage previous correctness proofs like in~\Cref{thm:derivative-correctness}, and also enables a non-symbolic equivalence-checking algorithm as in~\cref{alg:bisim}.
-Another notable difference is that our equivalence checking also need to handle normalization, which is a unique property of GKAT not found in general coalgebra.
+Another notable difference is that our equivalence checking also need to handle normalization, which is a unique property of GKAT not found in general coalgebra or automata.
 
@@
 GKAT is a guarded fragment of Kleene Algebra with Tests (KAT), which enjoys a symbolic algorithm~\cite{pous_SymbolicAlgorithmsLanguage_2015}.
-This algorithm by~\citeauthor{pous_SymbolicAlgorithmsLanguage_2015} generalizes the notion of derivatives also to boolean expressions, which uses binary decision diagram (BDD) as transition between symbolic expressions.
-Although some of our examples utilizes BDD to solve equivalence and inequivalence of boolean expressions, because of the structure of GKAT, our algorithm is not bounded by a particular boolean representation or solvers.
+This algorithm by~\citeauthor{pous_SymbolicAlgorithmsLanguage_2015} generalizes the notion of derivatives, and utilizes binary decision diagram (BDD) to form the desired coalgebras.
+Although some of our examples leverage BDD to solve equivalence and inequivalence of boolean expressions, our algorithm is not bounded by a particular boolean representation or solvers.
 In fact, we have demonstrated that in many scenarios, SAT solvers like MiniSat~\cite{een_MINISAT_2004} can achieve better performance than BDD.
@@
 In similar veins, KATch~\cite{moeller_KATchFastSymbolic_2024} is a symbolic solver for NetKAT~\cite{anderson_NetKATSemanticFoundations_2014} based on forwarding decision diagram, and achieved outstanding performance among state-of-the-art tools for reasoning about software-defined networks.
-GKAT also stem from the research of the research of software-defined networks~\cite{smolka_ScalableVerificationProbabilistic_2019,smolka_GuardedKleeneAlgebra_2020}, and quickly found applications in probabilistic verification~\cite{ro.zowski_ProbabilisticGuardedKAT_2023}, probabilistic program logic~\cite{gomes_KleeneAlgebraTests_2024}, networks~\cite{wasserstein_GUARDEDNETKATSOUNDNESS_2023}, and control-flow validation~\cite{zhang_CFGKATEfficientValidation_2025}.
-It would be interesting to see how our symbolic algorithm can be used to speed up these applications.
+The system of GKAT also stem from the research of the research of software-defined networks~\cite{smolka_ScalableVerificationProbabilistic_2019,smolka_GuardedKleeneAlgebra_2020}, and quickly found applications in probabilistic verification~\cite{ro.zowski_ProbabilisticGuardedKAT_2023}, probabilistic program logic~\cite{gomes_KleeneAlgebraTests_2024}, networks~\cite{wasserstein_GUARDEDNETKATSOUNDNESS_2023}, and control-flow validation~\cite{zhang_CFGKATEfficientValidation_2025}.
+It would be interesting to see how our symbolic algorithm can be used to speed up these variants of GKAT.
 

### 2025-01-24 main.tex
misc fixes

@@
 
-Our algorithm is lazy in the dead states detection i.e.
+Our algorithm is lazy in the dead-state detection i.e.
 the dead states are only checked when a mismatch requires it; even then, our dead state detection algorithm will only check the necessary states to compute the liveness of the states causing the mismatch.
@@
 Specifically, by induction on the length of the path from one state to the next, we can show that the principle sub-coalgebra \(⟨s⟩_S\) exactly contains all the reachable states of \(s\) in the GKAT coalgebra \(S\).
-Thus we obtain the following coalgebraic characterization of liveness:
+Thus, we obtain the following coalgebraic characterization of liveness:
 \begin{theoremEnd}{theorem}\label{thm:liveness-principle-sub-coalg}
@@
 
-This approach offers several benefit: first, the liveness detection is only called when necessary, thus unlikely to iterate through the entire coalgebra; second, the algorithm can short-circuit on a counter-example, allowing termination before all the states are explored; finally, the algorithm supports on-the-fly coalgebra constructions like derivatives, as the coalgebra is no longer required to be fully constructed before bisimulation.
+This approach offers several benefits: first, the liveness detection is only called when necessary, thus unlikely to iterate through the entire coalgebra; second, the algorithm can short-circuit on a counter-example, allowing termination before all the states are explored; finally, the algorithm supports on-the-fly coalgebra constructions like derivatives, as the coalgebra is no longer required to be fully constructed before bisimulation.
 
@@
 \begin{theoremEnd}{lemma}\label{thm:dead-construction-greedy-bisim}
-    Given any eager bisimulation \({≈} ⊆ S × U\) with the underlying bisimulation \({∼}\), the following relation \({∼_d} ⊆ S × U\):
+    For any eager bisimulation \({≈} ⊆ S × U\) with the underlying bisimulation \({∼}\), the following relation \({∼_d} ⊆ S × U\):
     \[{∼_d} ≜ {∼} ∪ \{(s, u) ∣ \text{\(s\) and \(u\) are dead in \(S\) and \(U\)}\},\]
@@
 \begin{proofEnd}
-    by definition of greedy bisimulation, the carrier of \({≡}\) is a subset of the carrier of \({≊}\). 
+    By definition of greedy bisimulation, the carrier of \({≡}\) is a subset of the carrier of \({≊}\). 
     Then because of~\Cref{thm:greedy-bisim-iff-norm-bisim}, \(\norm(≊)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\), which implies \(\norm(≊) ⊑ {≡}\).
@@
 
-Besides the construction, a classical optimization of the bisimulation-construction algorithm is to look for a equivalence relation instead of a relation~\cite{smolka_GuardedKleeneAlgebra_2020,hopcroft_LinearAlgorithmTesting_1971}.
+Besides the construction, a classical optimization of the bisimulation-construction algorithm is to look for an equivalence relation instead of a relation~\cite{smolka_GuardedKleeneAlgebra_2020,hopcroft_LinearAlgorithmTesting_1971}.
 This technique also applies to greedy bisimulation.
@@
     Because sub-coalgebra preserves bisimulation (\Cref{thm:sub-coalg-preserve-bisim}) and \(\norm(S) ⊑ \norm(S + U)\) and \(\norm(U) ⊑ \norm(S + U)\), we have \(\norm(≈)\) is a bisimulation between \(\norm(S + U)\) and itself.
-    let \(≡\) to be the maximal bisimulation within \(\norm(S + U)\), then \(≡\) is an equivalence relation~\cite[Corollary 5.6]{rutten_UniversalCoalgebraTheory_2000}, and so is the maximal greedy bisimulation \(≊\) (\Cref{thm:max-bisim-is-max-greedy-bisim}).
+    Let \(≡\) to be the maximal bisimulation within \(\norm(S + U)\), then \(≡\) is an equivalence relation~\cite[Corollary 5.6]{rutten_UniversalCoalgebraTheory_2000}, and so is the maximal greedy bisimulation \(≊\) (\Cref{thm:max-bisim-is-max-greedy-bisim}).
     This gives us the following chain of set inclusions:
     \[(s, u) ∈ {≈} = \norm(≈) ⊑ {≡} ⊆ {≊}.\]
-    Thus \(≊\) is a equivalence relation~\cite[Corollary 5.6]{rutten_UniversalCoalgebraTheory_2000} and a greedy bisimulation s.t. \(s ≡ u\).
+    Thus, \(≊\) is an equivalence relation~\cite[Corollary 5.6]{rutten_UniversalCoalgebraTheory_2000} and a greedy bisimulation s.t. \(s ≡ u\).
 
@@
     Finally, we reflect \(≡\) through the inclusion homomorphisms \(\norm(S) ⊑ \norm(S + U)\) and \(\norm(U) ⊑ \norm(S + U)\) as in~\Cref{thm:sub-coalg-preserve-bisim} to obtain \({∼} ⊆ \norm(S) × \norm(U)\) and \(s ∼ u\).
-    Therefore \((s, u)\) is also in the greedy bisimulation \({≈} ⊆ \norm(S) × \norm(U)\), generated by \(∼\).
+    Therefore, \((s, u)\) is also in the greedy bisimulation \({≈} ⊆ \norm(S) × \norm(U)\), generated by \(∼\).
 \end{proofEnd}
@@
 Additionally, \(\&\!\&\) is the logical conjunction operation.
-It is the same function as the the written ``and''.
+It is the same function as the written ``and''.
 We use \(\&\!\&\), instead of ``and'', to separate the conditions for clarity; logically \(\&\!\&\) is the same function as ``and'', but syntactically binds less tightly.
@@
 However, our algorithm slightly deviates from the ideal scenario described above. Specifically, we implement two optimizations.
-First, optimization is to terminate and return false as soon as we know one of \(s, u\) is dead and the other is not.
-This is because in order to have \(s ≈ u\) for any greedy bisimulation \(≈\), \(s\) and \(u\) must have the same liveness in their respective GKAT coalgebra (\Cref{thm:greedy-bisim-perserve-liveness}).
-In the case where both \(s\) and \(u\) are dead in their respective coalgebra, we will immediate return true, because any dead states pair are necessarily in \(≊\) (\Cref{thm:bisim-between-dead,thm:max-bisim-is-max-greedy-bisim}).
+First optimization is to terminate and return false as soon as we know one of \(s, u\) is dead, and the other is not.
+This early termination is sound, because in order for \(s ≈ u\) for any greedy bisimulation, \(s\) and \(u\) need to have the same liveness (\Cref{thm:greedy-bisim-perserve-liveness}).
+In the case where both \(s\) and \(u\) are dead in their respective coalgebra, we will immediately return true, because all dead-state pairs are in \(≊\) (\Cref{thm:bisim-between-dead,thm:max-bisim-is-max-greedy-bisim}).
 Then, we use a union-find data structure to organize the explored states into equivalence classes, which is justified by~\Cref{thm:greedy-bisim-iff-greedy-bisim-equiv}.

### 2025-01-24 main.tex
making table less crammed
minor fix

@@
 \usepackage[margin=2.5cm]{geometry}
+\usepackage{tabularx}
 
@@
 
-This surprising fact lead to the development of Guarded Kleene Algebra with Tests (GKAT)~\cite{smolka_GuardedKleeneAlgebra_2020}, a (co)algebraic system that seeks to characterize this efficient fragment of KAT.
+This surprising fact leads to the development of Guarded Kleene Algebra with Tests (GKAT)~\cite{smolka_GuardedKleeneAlgebra_2020}, a (co)algebraic system that seeks to characterize this efficient fragment of KAT.
 The determinism of GKAT gives it a different coalgebra structure from KAT, and allows for a GKAT expression to be directly converted to a linear-sized coalgebra, as opposed to the exponential sized coalgebra produced by KAT expressions~\cite{smolka_GuardedKleeneAlgebra_2020,kozen_CoalgebraicTheoryKleene_2017}.
@@
 \begin{proofEnd}
-    Because homomorphic image preserves principle sub-GKAT coalgebra (\Cref{thm:homo-img-preserve-principle-sub-coalg})
-    \[h(⟨s⟩_S) = ⟨h(s)⟩_U;\]
-    therefore for any state \(s' ∈ S\):
+    Because homomorphic image preserves principle sub-GKAT coalgebra (\Cref{thm:homo-img-preserve-principle-sub-coalg}) \(h(⟨s⟩_S) = ⟨h(s)⟩_U;\) therefore for any state \(s' ∈ S\):
     \[s' ∈ ⟨s⟩_S ⟺ h(s') ∈ h(⟨s⟩_S) ⟺ h(s') ∈ ⟨h(s)⟩_U.\]
@@
 \begin{theoremEnd}{corollary}[bisimulation preserves liveness]\label{thm:bisim-preserve-liveness}
-    If there exists a bisimulation \(∼\) between GKAT coalgebra \(S\) and \(U\) s.t.
-\(s ∼ u\) for some states \(s ∈ S\) and \(u ∈ U\), then \(s\) and \(u\) has to be either both accepting, both live or both dead.
+    For two states in GKAT coalgebras \(s ∈ S\) and \(u ∈ U\), with a bisimulation \({∼} ⊆ S × U\) s.t.
+\(s ∼ u\), then \(s\), \(u\), and \((s, u)\) has the same liveness, i.e.
+needs to be all accepting, all live, or all dead, in \(S\), \(U\), and \(∼\) respectively.
 \end{theoremEnd}
@@
     \centering
+    \setlength{\extrarowheight}{3px}
     \begin{tabular}{c||c|c|l|l}

### 2025-03-17 main.tex
tweaks

@@
 %
-% \author{\IEEEauthorblockN{Cheng Zhang\textsuperscript{\textsection}}
-% \IEEEauthorblockA{\textit{Department of Computer Science} \\
-% \textit{University College London}\\
-% London, United Kingdom \\
-% 0000-0002-8197-6181}
-% \and
-% \IEEEauthorblockN{Qiancheng Fu}
-% \IEEEauthorblockA{\textit{Department of Computer Science} \\
-% \textit{Boston University}\\
-% Boston, USA \\
-% qcfu@bu.edu}
-% \and
-% \IEEEauthorblockN{Hang Ji\textsuperscript{\textsection}}
-% \IEEEauthorblockA{\textit{Department of Computer Science} \\
-% \textit{Cornell University}\\
-% Ithaca, USA \\
-% hj476@cornell.edu}
-% \and
-% \IEEEauthorblockN{Ines Santacruz}
-% \IEEEauthorblockA{\textit{Department of Computer Science} \\
-% \textit{Boston University}\\
-% Boston, USA \\
-% inessdv15@gmail.com}
-% \and
-% \IEEEauthorblockN{Marco Gaboardi}
-% \IEEEauthorblockA{\textit{Department of Computer Science} \\
-% \textit{Boston University}\\
-% Boston, USA \\
-% gaboardi@bu.edu}
-% }
+\author{\IEEEauthorblockN{Cheng Zhang\textsuperscript{\textsection}}
+\IEEEauthorblockA{\textit{Department of Computer Science} \\
+\textit{University College London}\\
+London, United Kingdom \\
+0000-0002-8197-6181}
+\and
+\IEEEauthorblockN{Qiancheng Fu}
+\IEEEauthorblockA{\textit{Department of Computer Science} \\
+\textit{Boston University}\\
+Boston, USA \\
+qcfu@bu.edu}
+\and
+\IEEEauthorblockN{Hang Ji\textsuperscript{\textsection}}
+\IEEEauthorblockA{\textit{Department of Computer Science} \\
+\textit{Cornell University}\\
+Ithaca, USA \\
+hj476@cornell.edu}
+\and
+\IEEEauthorblockN{Ines Santacruz}
+\IEEEauthorblockA{\textit{Department of Computer Science} \\
+\textit{Boston University}\\
+Boston, USA \\
+inessdv15@gmail.com}
+\and
+\IEEEauthorblockN{Marco Gaboardi}
+\IEEEauthorblockA{\textit{Department of Computer Science} \\
+\textit{Boston University}\\
+Boston, USA \\
+gaboardi@bu.edu}
+}
 
 \maketitle
-% \begingroup\renewcommand\thefootnote{\textsection}
-% \footnotetext{Work largely performed at Boston University}
-% \endgroup
+\begingroup\renewcommand\thefootnote{\textsection}
+\footnotetext{Work largely performed at Boston University}
+\endgroup
 
@@
     given a homomorphism \(h: S → U\), and sub-coalgebras \(S' ⊑ S\) and \(U' ⊑ U\), then 
-    \[h(S') ⊑ U \text{ and } h^{-1}(U') ⊑ S.\]
+    \(h(S') ⊑ U \text{ and } h^{-1}(U') ⊑ S.\)
 \end{theoremEnd}
@@
     Homomorphic image preserves principle sub-GKAT coalgebra. Specifically, given a homomorphism \(h: S → U\):
-    \[h(⟨s⟩_S) = ⟨h(s)⟩_U\]
+    \(h(⟨s⟩_S) = ⟨h(s)⟩_U\)
 \end{theoremEnd}
@@
 \]
-We sometimes use \({∼} ⊆ S × U\) to specify that \(∼\) is a bisimulation between \(S\) and \(U\).
+We sometimes abuse the notation \({∼} ⊆ S × U\) to specify that \(∼\) is a bisimulation between \(S\) and \(U\).
 
@@
 Besides the nice property of bisimulation, GKAT coalgebra is also deterministic, unlike Kleene Coalgebra with Tests (KCT)~\cite{kozen_CoalgebraicTheoryKleene_2017}.
-For each atom, states in a KCT can accept or reject the atom (but not both), yet states can also non-deterministically transition to multiple different states via the same atom, while executing different actions.
+For each atom, states in a KCT can non-deterministically accept, reject (but not both), or transition to multiple different states while executing different actions.
 As we will see later, while the deterministic behavior of GKAT coalgebra enables a more versatile symbolic algorithm than KCT~\cite{pous_SymbolicAlgorithmsLanguage_2015}, it also comes with unique challenges.
@@
 In previous works, these dead states are detected after the necessary coalgebra is computed and stored.
-This approach requires storing all the transitions and states of coalgebra in memory, meaning that the algorithm does not short circuit even if a mismatch can be detected in the starting state; for example, when for some \(α ∈ \At\), \(δ_S(s, α) = \reject\) and \(δ_U(u, α) = \accept\).
+This approach requires storing all the transitions and states of coalgebra in memory, meaning that the algorithm does not short circuit even if a mismatch can be detected in the starting state; for example, when for some \(α ∈ \At\), \(δ_S(s, α) = \reject\) and \(δ_U(u, α) = \accept\), then we can conclude \(s, u\) have different trace semantics, without needing to compute the rest of the coalgebra.
 
@@
 \end{definition}
-However, besides directly working with this explicit definition, we can also give a coalgebraic characterization of these properties.
-Specifically, by induction on the length of the path from one state to the next, we can show that the principle sub-coalgebra \(⟨s⟩_S\) exactly contains all the reachable states of \(s\) in the GKAT coalgebra \(S\).
+Besides directly working with this explicit definition, we can also give a coalgebraic characterization of liveness.
+Specifically, by induction on the length of the path from state \(s\) to an arbitrary reachable state from \(s\), we can show that the principle sub-coalgebra \(⟨s⟩_S\) exactly contains all the reachable states of \(s\) in the GKAT coalgebra \(S\).
 Thus, we obtain the following coalgebraic characterization of liveness:
@@
     \(⟸\) direction is true, because \(s ∈ ⟨s⟩\): if all \(⟨s⟩\) is dead, then \(s\) is dead. 
-    \(⟹\) direction can be proven as follows.
-    Take any \(s' ∈ ⟨s⟩\), then \(⟨s'⟩ ⊑ ⟨s⟩\) because \(s'\) is the minimal sub-coalgebra that contains \(s'\). 
-    Since there is no accepting state in \(⟨s⟩\), thus there cannot be any accepting state in \(⟨s'⟩\), hence \(s'\) is also dead.
+    \(⟹\) direction can be proven by taking any \(s' ∈ ⟨s⟩\); then we notice \(⟨s'⟩ ⊑ ⟨s⟩\) because \(s'\) is the minimal sub-coalgebra that contains \(s'\). 
+    Since there is no accepting state in \(⟨s⟩\), there cannot be any accepting state in \(⟨s'⟩\), hence \(s'\) is also dead.
 \end{proofEnd}
@@
 \(s ∼ t\).
-For a more detailed explanation on the finite trace semantics, we refer the reader to the work of \Citeauthor{smolka_GuardedKleeneAlgebra_2020}~\cite{smolka_GuardedKleeneAlgebra_2020}, whereas we will recall the correctness theorem here.
+For a more detailed explanation on the finite trace semantics, we refer the reader to the work of \Citeauthor{smolka_GuardedKleeneAlgebra_2020}~\cite{smolka_GuardedKleeneAlgebra_2020}, we will only recall the correctness theorem here.
 
@@
 Because of the functoriality, we can show that two states that are infinite-trace equivalent implies these two states are finite-trace equivalent.
-This gives us more tool in proving semantic equivalence between two states in GKAT coalgebras: proving bisimulation of these two states in the \emph{non-normalized} GKAT coalgebra can also obtain finite-trance equivalence for the two states.
+This gives us more tool in proving semantic equivalence between two states in GKAT coalgebras: proving bisimulation of these two states in the \emph{non-normalized} GKAT coalgebra is sufficient (but not necessary) to obtain finite-trance equivalence for the two states.
 
@@
 Given any bisimulation \({∼} ⊆ S × U\), we can construct a GKAT coalgebra \(∼_d\) s.t.
-it is almost a greedy bisimulation, i.e.
-it is closed under \(δ_≈\) in~\Cref{def:greedy-bisim}, and contains \(∼\), except it is not necessarily the \emph{least} GKAT coalgebra that satisfies these properties.
+it is \emph{almost} a greedy bisimulation, i.e.
+it is closed under \(δ_≈\) in~\Cref{def:greedy-bisim}, and contains \(∼\), except it is not necessarily the least GKAT coalgebra that satisfies these properties.
 
 \begin{theoremEnd}{lemma}\label{thm:dead-construction-greedy-bisim}
-    For any eager bisimulation \({≈} ⊆ S × U\) with the underlying bisimulation \({∼}\), the following relation \({∼_d} ⊆ S × U\):
+    For any greedy bisimulation \({≈} ⊆ S × U\) with the underlying bisimulation \({∼}\), the following relation \({∼_d} ⊆ S × U\):
     \[{∼_d} ≜ {∼} ∪ \{(s, u) ∣ \text{\(s\) and \(u\) are dead in \(S\) and \(U\)}\},\]
@@
 \begin{theoremEnd}[normal]{corollary}\label{thm:greedy-bisim-exists-unique}
-    For any bisimulation between two normalized GKAT coalgebras \({∼} ⊆ \norm(S) × \norm(U)\), a greedy bisimulation \(≈\) with \(∼\) as underlying bisimulation exists and is unique.
-\end{theoremEnd}
-
-\begin{proofEnd}
-    Because sub-coalgebra form a complete lattice under union and intersection~\cite[Theorem 6.4]{rutten_UniversalCoalgebraTheory_2000}, then we can construct \(≈\) by taking the intersection of all the sub-coalgebra of \(∼_d\) in~\Cref{thm:dead-construction-greedy-bisim}.
+    For any bisimulation \({∼} ⊆ \norm(S) × \norm(U)\), a greedy bisimulation \(≈\) with \(∼\) as underlying bisimulation exists and is unique.
+\end{theoremEnd}
+
+\begin{proofEnd}
+    Because sub-coalgebra form a complete lattice under union and intersection~\cite[Theorem 6.4]{rutten_UniversalCoalgebraTheory_2000}, then \(≈\) is the intersection of all the sub-coalgebra of \(∼_d\) (\Cref{thm:dead-construction-greedy-bisim}) containing \(∼\).
 \end{proofEnd}
@@
 \begin{theoremEnd}[normal]{corollary}[Perservation Of Liveness]\label{thm:greedy-bisim-perserve-liveness}
-    Given a pair of states in a greedy bisimulation \(s ≈ u\) where \({≈} ⊆ S × U\), then the following are equivalent
-    \begin{itemize}
-        \item \((s, u)\) is live (or dead) in \({≈}\);
-        \item \(s\) is live (or dead) in \(S\);  
-        \item \(u\) is live (or dead) in \(U\).
-    \end{itemize}
-    Furthermore, if \((s, u)\) is in the underlying bisimulation \(∼\) of \(≈\), then all of above are also equivalent to \((s, u)\) is live (or dead) in \(∼\).
+    Given a pair of states in a greedy bisimulation \(s ≈ u\) where \({≈} ⊆ S × U\), then \((s, u)\), \(s\), \(u\) all have the same liveness in \(≈\), \(S\), and \(U\) respectively.
 \end{theoremEnd}
@@
 \begin{theoremEnd}[normal]{corollary}\label{thm:greedy-bisim-iff-norm-bisim}
-    Given two GKAT coalgebras \(S\) and \(U\), and a relational GKAT coalgebra \({≈} ⊆ S × U\) with the transition \(δ_≈\) as in~\cref{def:greedy-bisim}.
+    Given two GKAT coalgebras \(S\) and \(U\), and a relational GKAT coalgebra \({≈} ⊆ S × U\) with the transition \(δ_≈\) as in~\Cref{def:greedy-bisim}.
     \(≈\) is a greedy bisimulation if and only if \(\norm(≈)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\)
@@
     Because \({≈} ⊑ {∼_d}\) and~\Cref{thm:norm-sub-coalg}, \(\norm(≈) ⊑ \norm(∼_d)\). 
-    Because \(\norm(∼_d)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\), therefore \(\norm(≈)\) is a bisimulation (\Cref{thm:greedy-bisim-iff-norm-bisim}).
+    Because \(\norm(∼_d)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\), therefore \(\norm(≈)\) is also a bisimulation (\Cref{thm:greedy-bisim-iff-norm-bisim}).
 
@@
 Despite satisfying many desirable structural properties, the definition of greedy bisimulation is still coalgebraic.
-To aid the design of an algorithm based on greedy bisimulation, we turn the coalgebra into a set of concrete criteria, that can be checked by an algorithm.
+To aid the design of an algorithm based on greedy bisimulation, we establish the correspondence of its definition with a set of concrete criteria, that can be checked by an algorithm.
 In the following theorem, we will present such necessary and sufficient conditions to construct a greedy bisimulation.
@@
 
-% \begin{theoremEnd}{theorem}\label{thm:greedy-bisim-one-dead}
-%     Consider two states in GKAT coalgebras \(s ∈ S, u ∈ U\) and an arbitrary bisimulation \({∼} ⊆ \norm(S) × \norm(U)\), if \(s\) is dead, then 
-%     \begin{enumerate}
-%         \item\label{itm:greedy-bisim-one-dead-then-both-dead} if \(s ∼ u\), then \(u\) is dead.
-%         \item\label{itm:remove-dead-pairs-sound} if \(u\) is dead, then \({∼} ∖ ⟨(s, u)⟩_∼\) is also a bisimulation between \(\norm(S)\) and \(\norm(U)\).
-%     \end{enumerate}
-% \end{theoremEnd}
-
-% \begin{proofEnd}
-%     \Cref{itm:greedy-bisim-one-dead-then-both-dead} is a consequence of bisimulation preserves liveness (\Cref{thm:bisim-preserve-liveness}) and normalization preserves liveness (\Cref{thm:norm-perserve-liveness}).
-
-%     \Cref{itm:remove-dead-pairs-sound} can be proven by realizing that \((s, u)\) is a dead state in \(∼\) (\Cref{thm:bisim-preserve-liveness}) which means that all the states in \(⟨(s, u)⟩_∼\) are dead.
-%     Because \(∼\) is a normalized GKAT coalgebra (\Cref{thm:}), therefore no state can transition into dead states like \(⟨(s, u)⟩_∼\), hence \({∼} ∖ ⟨(s, u)⟩_∼\) is still closed under its transition, and form a sub-coalgebra of \(∼\).
-%     Finally because \(∼\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\), then so is \({∼} ∖ ⟨(s, u)⟩_∼\) (\Cref{thm:})
-% \end{proofEnd}
-
 Before we introduce the algorithm, we will first briefly sketch the liveness-check algorithm.
@@
 We cache all the known dead states from the previous searches; whether a state \(s\) is in this cache can be checked by the function call \Call{knownDead\(_S\)}{$s$}. 
-The function \Call{isDead\(_S\)}{$s$} will first check if \(s\) is known to be dead, and invoke the search algorithm in the coalgebra \(⟨s⟩\) if \(s\) is not in the cached dead states.
-Because the non-symbolic version of liveness checking is similar to the symbolic version, we only provide the symbolic version of this algorithm in~\cref{lst:dead-state-detection}.
+On the other hand, the function \Call{isDead\(_S\)}{$s$} will first check if \(s\) is known to be dead, and invoke the search algorithm in the coalgebra \(⟨s⟩\) if \(s\) is not in the cached dead states.
+Because the non-symbolic version of liveness checking is similar to the symbolic version, we only present the symbolic version of this algorithm in~\cref{lst:dead-state-detection}.
 
@@
 
-Finally, we present our equivalence-checking algorithm as~\Cref{alg:bisim} which consists of recursively checking the conditions in~\Cref{thm:recursive-construction} with minor optimizations.
+Finally, we present our equivalence-checking algorithm as~\Cref{alg:bisim} which consists of recursively checking the conditions in~\Cref{thm:recursive-construction} with two optimization: early termination with known dead states, and the use of union-find data structure; both of which we will justify later.
 In the algorithm, ``implies'', ``iff'', ``and'' are logical operators that take booleans (i.e.
 true or false) and return a boolean.
-Additionally, \(\&\!\&\) is the logical conjunction operation.
-It is the same function as the written ``and''.
-We use \(\&\!\&\), instead of ``and'', to separate the conditions for clarity; logically \(\&\!\&\) is the same function as ``and'', but syntactically binds less tightly.
+We use \(\&\!\&\), instead of ``and'', to separate the conditions for clarity; logically \(\&\!\&\) is the same function as ``and'', but syntactically \(\&\!\&\) binds less tightly than ``and''.
 
 Intuitively, the algorithm returns true if and only if a greedy bisimulation \(≈\) can be constructed s.t. \(s ≈ u\).
-Note the existence of such \(≈\) is not only equivalent to the trace equivalence of \(s\) and \(u\) (\Cref{thm:greedy-bisim-correctness}), but also equivalent to \(s ≊ u\), where \({≊} ⊆ S × U\) is the maximal bisimulation.
-However, our algorithm slightly deviates from the ideal scenario described above. Specifically, we implement two optimizations.
-First optimization is to terminate and return false as soon as we know one of \(s, u\) is dead, and the other is not.
-This early termination is sound, because in order for \(s ≈ u\) for any greedy bisimulation, \(s\) and \(u\) need to have the same liveness (\Cref{thm:greedy-bisim-perserve-liveness}).
-In the case where both \(s\) and \(u\) are dead in their respective coalgebra, we will immediately return true, because all dead-state pairs are in \(≊\) (\Cref{thm:bisim-between-dead,thm:max-bisim-is-max-greedy-bisim}).
-Then, we use a union-find data structure to organize the explored states into equivalence classes, which is justified by~\Cref{thm:greedy-bisim-iff-greedy-bisim-equiv}.
-By looking for a greedy bisimulation that is also an equivalence relation, we can use more efficient structures like union-find, instead of set of pairs; the correctness of this optimization is justified by.
+Note the existence of such \(≈\) is not only equivalent to the trace equivalence of \(s\) and \(u\) (\Cref{thm:greedy-bisim-correctness}), but is also equivalent to \(s ≊ u\), where \({≊} ⊆ S × U\) is the maximal greedy bisimulation.
+Our algorithm deviates from the ideal scenario by implementing two algorithm mentioned above. 
+First optimization is to terminate early when one state is known to be dead. 
+If the other state is live, then the algorithm return false, because \(s ≈ u\) implies \(s\) and \(u\) have the same liveness (\Cref{thm:greedy-bisim-perserve-liveness}); if both input state are dead, because all dead state pairs are in \(≊\) (\Cref{thm:bisim-between-dead,thm:max-bisim-is-max-greedy-bisim}).
+Then, we use a union-find data structure to organize the explored states into equivalence classes, which is justified by finding a equivalence relation (\Cref{thm:greedy-bisim-iff-greedy-bisim-equiv}).
 
@@
 
-Although both algorithms have similar worst case complexity, the on-the-fly algorithm will always out-perform the original algorithm in non-worst case situations.
+Although both algorithms have similar worst case complexity, the on-the-fly algorithm will almost always out-perform the original algorithm. 
 This is because the on-the-fly algorithm only invokes liveness checking when necessary. 
@@
 This coalgebra is required to satisfy the \emph{disjointedness condition}, i.e.
-for all states \(s ∈ S\), the boolean expressions that \(s\) accepts \(ϵ̂(s)\) and the boolean expressions that enables \(s\) to transition \(\{b ∣ ∃ (b, s', p) ∈ δ(s)\}\) are disjoint; i.e.
-the conjunction of any two distinct expressions from the set \(ϵ̂(s) ∪ \{b ∣ ∃ (b, s', p) ∈ δ(s)\}\) is equivalent to \(0\) under Boolean Algebra.
-Like the non-symbolic case, we use \(s \transAcc{Ŝ}{b}\) to indicate \(b ∈ ϵ̂(s)\) and \(s \transvia{b ∣ p}_{Ŝ} s'\) to indicate \((b, p, s') ∈ δ̂(s)\).
+for all states \(s ∈ S\), the boolean expressions that \(s\) accepts \(ϵ̂(s)\) and the boolean expressions that enables \(s\) to transition \(\{b ∣ ∃ (b, s', p) ∈ δ(s)\}\) are disjoint.
+Formally, the disjointedness condition states the conjunction of any two distinct expressions from the set \(ϵ̂(s) ∪ \{b ∣ ∃ (b, s', p) ∈ δ(s)\}\) is equivalent to \(0\) under Boolean Algebra.
+Similar to the non-symbolic case, we use \(s \transAcc{Ŝ}{b}\) to indicate \(b ∈ ϵ̂(s)\) and \(s \transvia{b ∣ p}_{Ŝ} s'\) to indicate \((b, p, s') ∈ δ̂(s)\).
 

### 2025-03-18 rebuttal.md
first draft of rebuttal

# Overall Response

We would like to express our sincere gratitude for the constructive feedback received from the reviewer. We will carefully address the points raised and implement the necessary changes to our paper.

First, we address the efficiency of our lazy normalization approach. 
The classical algorithm guarantees a full traversal of the coalgebra during liveness checking prior to bisimulation checking, our algorithm invokes bisimulation and only run traverses the reachable states of the discrepancy state.
Therefore, our lazy liveness checking consistently traverses fewer states, even for positive results. 
Additionally, lazy liveness checking enables on-the-fly coalgebra generation using methods like derivatives, which the original algorithm does not support.

Second, we address the theoretical contribution of our work. 
While the result may seem practically intuitive, we maintain that our proof is not trivial. 
We are able to provide straightforward and concise proofs because we introduced a middle ground between coalgebraic theory and practical algorithm, namely greedy bisimulation. 
Without this framework, verifying Theorem 22 and Theorem 23 would be more challenging, especially given that neither the input coalgebra nor their greedy bisimulation are assumed to be finite.

Finally, we emphasize that our contribution extends beyond bisimulation of machines (or coalgebras).
In Section V, we present two methods for generating coalgebras from expressions and prove their correctness, finiteness, and correspondence.
The finiteness result, in particular, arises naturally from the correspondence result. 
The simplicity of this result is noted by one of the reviewers.

Regarding the practical aspects of our work, specifically the experiments, we are committed to revising the tests to better represent real-world scenarios before the deadline. 
However, we also argue that our theoretical results have already demonstrated improvements over existing methods. 
Plus, trivial test cases do not provide any performance benefit for our algorithm based on Thompson's construction, and dead test cases also do not favor our algorithm against other tools. 
Although we will improve the balance and coverage of our test cases, we believe the current test cases are sufficient in demonstrating the efficiency of our implementation.

# Reviewer Questions

### Reviewer 1

> Where are the benchmarks from?

The benchmark is generated by qcheck library of Ocaml, using the module QCheck2. 
The generation is based on algebraic rules of GKAT (except the fixpoint rules) with reflexivity and symmetry rule.
We will include the generation script in the final submission, and provide documentation on test generation

> Unnecessary but harmful generalization

We will specify in the background section that we assume the function `F` to be simple polynomial functors.

### Reviewer 2

Our approach to symbolic transducers diverges from previous works in several key aspects:
- Our on-the-fly method is tailored to handle normalization, a unique challenge in achieving finite trace equivalence for GKAT. 
- The signature of GKAT differs from traditional symbolic transducers. Unlike typical accepting or rejecting states, a GKAT state can accept or reject based on different input atoms. This results in a distinct semantics where a trace's end is marked by an atom, representing the output program state.

While symbolic bisimulation is pivotal for the efficiency of our equivalence algorithm, it is not our sole theoretical contribution. 
We have also developed:
- An on-the-fly algorithm for bisimulation, and proved its correctness using greedy bisimulation, a novel coalgebra construction.
- Two distinct methods to construct symbolic GKAT coalgebras, while also proved their correctness, finiteness, and correspondence.

# Detailed Response

### Reviewer 1

> why is it relevant to check whether two GKAT expressions are equivalent?

GKAT has been used in several important verification tasks in the domain of networks, control-flow transformation, probabilistic computation, and program logic.
We mentioned this fact at the end of the related work section. 
We will also mention the practical application of GKAT in the introduction section.

> it's unclear to me what section III is for if you have section IV

The correctness result of section IV is based on the correctness result in section III. 
Besides, there are also cases where the number of primitive tests are small and the algorithm in section III can potentially be more efficient.

### Reviewer 2

> there can still be dead states in a normalized coalgebra

Dead states are indeed preserved in the normalized coalgebra.
This approach makes the normalized coalgebra share the same carrier as the original coalgebra, allowing us to define the semantics of dead states also via the unique map into final coalgebra.

> can you say a bit more about SymKAT and why it might perform worse

SymKAT is a symbolic equivalence checker for Kleene Algebra with Tests (KAT).
Our algorithm is specialized to GKAT, thus able to leverage the deterministic structure to produce smaller automata (coalgebra) and use existing SAT solver to resolve boolean equivalence/inequivalence.
Our contribution, to the best of our knowledge, is both a theoretical and practical demonstration that GKAT equivalence can be checked more efficiently using a specialized solver.

### Reviewer 3

> Theorem 23 is true, but Theorem 23 is not the proper tool to justify the correctness of the algorithm as mentioned on p7c2

A union-find structure represents an equivalence relation. 
By employing union-find instead of a set of equivalent pairs, algorithm will search for a greedy bisimulation that is also an equivalence relation. 
While the correctness of greedy bisimulation is established (Theorem 21), there is no inherent justification for why finding a greedy bisimulation that is also an equivalence relation is equivalent to finding just any greedy bisimulation. 
To address this, we will apply Theorem 23 to demonstrate that there exists a greedy bisimulation that is also an equivalence relation if and only if two states are trace equivalent.

> p5c2 "This approach offers several benefits: first, the liveness detection is only called when necessary, thus unlikely to iterate through the entire coalgebra; second, the algorithm can short-circuit on a counter-example, finally [...]""
> -> The first two are the same benefit.

The first benefit stems from the laziness of our liveness checking algorithm, allowing us to iterate through fewer states in the classical algorithm (see overall response and the next comment).  
This benefit is apparent no matter the result or the generation method of the coalgebra. 
The second benefit stems from the on-the-fly nature of our algorithm.
Specifically, our algorithm can terminate without generating the rest of the coalgebra, when a counter-example is encountered while applying an on-the-fly generation method (such as derivatives).
In contrast, the original algorithm requires the complete generation of the coalgebra in order to perform liveness checking first.

> p7c2 "In the extreme case when the two input states are infinite-trace equivalent, the on-the-fly algorithm can even skip liveness checking entirely." 
> -> yes, but in that case the liveness analysis would be faster (linear instead of almost linear)

We are uncertain why the original algorithm might be faster in this specific case. 
The original algorithm requires determining the liveness of every single state before performing bisimulation. 
In contrast, our algorithm only conducts liveness checking on a few states when the bisimulation algorithm detects a discrepancy.
Consider two bisimilar states (not just finite-trace equivalent) in a GKAT coalgebra composed entirely of live states. 
The classical algorithm would first perform a liveness check, iterating through all the states, and then proceed with bisimulation, iterating through all the states again. 
Our algorithm, however, simply performs the bisimulation and confirms there is no discrepancy, effectively bypassing the liveness check altogether.

# !!!!!!! NOTES, DO NOT SUBMIT !!!!!!!
1. Practicality of the tool
    <!-- - Do people care about GKAT 
        - not a lot of application of GKAT, because there is no evidence that GKAT is faster than KAT -->
    - Speed up of normalization 
        - discussion at the end of section III;
        - better sketch of liveness checking.
    - Trivial/Dead Equivalence
2. Comparison with existing work
    - Intuitive, but coalgebraic characterization is novel
    - coalgebraic characterization enables straightforward proofs
    - Theorem 22 is non-trivial because there is no assumption of finiteness
3. additional results besides automata
    - We also presented two ways to generate automata in Section V and proved them correct and equivalent.
    - Automata shape difference, allow acceptation or rejection based on an atom.
3. Tests generation
    - We generate the tests roughly based on the shape of real-world program input 
        - more sequencing than while etc.
    <!-- - "trivial" tests does not impact the performance evaluation for Thompson
    - "dead" tests does not impact the performance evaluation -->
    - negative cases are random expression, but they still need to do liveness check
    - We will rework the tests to 
        - reduce the number of trivial and dead tests, 
        - make sure that the random expression do go into dead state detection immediately
4. Writing
    - More about the efficiency of laziness

### 2025-03-18 rebuttal.md
comment out the notes in case it is submitted

@@
 
-# !!!!!!! NOTES, DO NOT SUBMIT !!!!!!!
+<!-- 
+!!!!!!! NOTES, DO NOT SUBMIT !!!!!!!
 1. Practicality of the tool
-    <!-- - Do people care about GKAT 
-        - not a lot of application of GKAT, because there is no evidence that GKAT is faster than KAT -->
     - Speed up of normalization 
@@
         - more sequencing than while etc.
-    <!-- - "trivial" tests does not impact the performance evaluation for Thompson
-    - "dead" tests does not impact the performance evaluation -->
     - negative cases are random expression, but they still need to do liveness check
@@
     - More about the efficiency of laziness 
+-->

### 2025-03-19 rebuttal-600words.md
Exactly 600!

# Overall Response

We express our gratitude for the valuable feedback received from reviewers. 
We will carefully address the points raised and implement the necessary changes to our paper.

First, we address concerns surrounding lazy normalization. 
The classical algorithm guarantees a full traversal of the coalgebra during liveness checking prior to bisimulation checking, our algorithm invokes bisimulation and only traverses the reachable states of the discrepancies.
Therefore, lazy liveness checking consistently visits fewer states, even for positive results. 
Additionally, lazy liveness checking enables on-the-fly coalgebra generation using derivatives.

While our algorithms seem intuitive, our proofs are not trivial. 
The simple proofs are enabled by the design of greedy bisimulation. 
Without this middle ground, verifying Theorem 22 would be challenging, especially given that neither the input coalgebras nor their bisimulation are assumed to be finite.

Finally, we emphasize that our contribution extends beyond bisimulation of coalgebras.
We also presented two methods for generating coalgebras from expressions and proved their correctness, finiteness, and correspondence.

Regarding our experiments, we have revised our tests in our repository. 
We also stress that trivial tests do not provide any performance benefit for algorithms based on Thompson's construction, and dead tests do not favor our algorithm. 
As mentioned before, we have conceptual demonstrations of performance improvements in the submitted version.

# Reviewer Questions

### Reviewer 1

> Where are the benchmarks from?

It is generated by https://github.com/c-cube/qcheck, based on algebraic rules of GKAT (excluding fixpoint rule) with reflexivity, congruence, and symmetry rules.
We will include the generation script and provide documentation.

> Unnecessary but harmful generalization

We will clarify that the functor `F` satisfies necessary properties.

### Reviewer 2

Our approach diverges from previous works on symbolic transducers in several aspects:
- Our on-the-fly method is tailored for normalization, a unique challenge in achieving finite trace equivalence for GKAT. 
- The signature of GKAT differs from traditional symbolic transducers. Unlike typical accepting or rejecting states, a GKAT state can accept or reject based on different input atoms. This results in a semantics where a trace ends in an atom, representing the output program state.

While symbolic bisimulation is a valuable contribution, it is not our sole contribution. 
We spent most of the paper developing an on-the-fly bisimulation algorithm and two distinct method to construct symbolic GKAT coalgebra from expressions, and proved their desirable properties.

# Detailed Response

### Reviewer 1

> it's unclear to me what section III is for

The correctness result of section IV is based on that of section III.

### Reviewer 2

> can you say a bit more about SymKAT

SymKAT is a symbolic equivalence checker for KAT; on the other hand, our algorithm focuses on the GKAT fragment.
This specialization allows us to produce smaller coalgebra and utilize efficient SAT solver.
Our comparison demonstrated that GKAT equivalence can be checked more efficiently using our specialized solver.

### Reviewer 3

> Theorem 23 is not the proper tool to justify the correctness of the algorithm

Union-find structures represent equivalence relations. 
By employing union-find instead of a set of pairs, our algorithm searches for a greedy bisimulation that is also an equivalence relation. 
While the correctness of greedy bisimulation is established (Theorem 21), Theorem 23 justifies finding a greedy bisimulation that is also an equivalence relation is correct.

> p7c2 [...] the liveness analysis would be faster

Consider two bisimilar states in a GKAT coalgebra with no dead state. 
The classical algorithm would perform a liveness check, iterating through all the states; then proceed with bisimulation, iterating through all the states again. 
Our algorithm simply performs the bisimulation, bypassing the liveness check altogether.

<!-- > p5c2 [...] The first two are the same benefit.

The first benefit stems from the laziness of our liveness checking algorithm, allowing us to iterate through fewer states in the classical algorithm (see overall response and the next comment).  
This benefit is apparent no matter the result or the generation method of the coalgebra. 
The second benefit stems from the on-the-fly nature of our algorithm.
Specifically, our algorithm can terminate without generating the rest of the coalgebra, when a counter-example is encountered while applying an on-the-fly generation method (such as derivatives).
In contrast, the original algorithm requires the complete generation of the coalgebra in order to perform liveness checking first. -->

<!-- 
!!!!!!! NOTES, DO NOT SUBMIT !!!!!!!
1. Practicality of the tool
    - Speed up of normalization 
        - discussion at the end of section III;
        - better sketch of liveness checking.
    - Trivial/Dead Equivalence
2. Comparison with existing work
    - Intuitive, but coalgebraic characterization is novel
    - coalgebraic characterization enables straightforward proofs
    - Theorem 22 is non-trivial because there is no assumption of finiteness
3. additional results besides automata
    - We also presented two ways to generate automata in Section V and proved them correct and equivalent.
    - Automata shape difference, allow acceptation or rejection based on an atom.
3. Tests generation
    - We generate the tests roughly based on the shape of real-world program input 
        - more sequencing than while etc.
    - negative cases are random expression, but they still need to do liveness check
    - We will rework the tests to 
        - reduce the number of trivial and dead tests, 
        - make sure that the random expression do go into dead state detection immediately
4. Writing
    - More about the efficiency of laziness 
-->

### 2025-03-19 rebuttal-600words.md
stronger comment on trivial tests

@@
 Regarding our experiments, we have revised our tests in our repository. 
-We also stress that trivial tests do not provide any performance benefit for algorithms based on Thompson's construction, and dead tests do not favor our algorithm. 
+We also emphasize that neither trivial nor dead tests provide performance benefit in our current implementation, so submitted tests are reflective of the general cases. 
 As mentioned before, we have conceptual demonstrations of performance improvements in the submitted version.
@@
 Our algorithm simply performs the bisimulation, bypassing the liveness check altogether.
-
-<!-- > p5c2 [...] The first two are the same benefit.
-
-The first benefit stems from the laziness of our liveness checking algorithm, allowing us to iterate through fewer states in the classical algorithm (see overall response and the next comment).  
-This benefit is apparent no matter the result or the generation method of the coalgebra. 
-The second benefit stems from the on-the-fly nature of our algorithm.
-Specifically, our algorithm can terminate without generating the rest of the coalgebra, when a counter-example is encountered while applying an on-the-fly generation method (such as derivatives).
-In contrast, the original algorithm requires the complete generation of the coalgebra in order to perform liveness checking first. -->
-
-<!-- 
-!!!!!!! NOTES, DO NOT SUBMIT !!!!!!!
-1. Practicality of the tool
-    - Speed up of normalization 
-        - discussion at the end of section III;
-        - better sketch of liveness checking.
-    - Trivial/Dead Equivalence
-2. Comparison with existing work
-    - Intuitive, but coalgebraic characterization is novel
-    - coalgebraic characterization enables straightforward proofs
-    - Theorem 22 is non-trivial because there is no assumption of finiteness
-3. additional results besides automata
-    - We also presented two ways to generate automata in Section V and proved them correct and equivalent.
-    - Automata shape difference, allow acceptation or rejection based on an atom.
-3. Tests generation
-    - We generate the tests roughly based on the shape of real-world program input 
-        - more sequencing than while etc.
-    - negative cases are random expression, but they still need to do liveness check
-    - We will rework the tests to 
-        - reduce the number of trivial and dead tests, 
-        - make sure that the random expression do go into dead state detection immediately
-4. Writing
-    - More about the efficiency of laziness 
--->

### 2025-03-20 rebuttal.md
minor fix
a longer draft, changes incoming!

@@
 
-We would like to express our sincere gratitude for the constructive feedback received from the reviewer. We will carefully address the points raised and implement the necessary changes to our paper.
+We would like to express our sincere gratitude for the constructive feedback received from the reviewer. 
+We will carefully address the points raised and implement the necessary changes to our paper.
 
 First, we address the efficiency of our lazy normalization approach. 
-The classical algorithm guarantees a full traversal of the coalgebra during liveness checking prior to bisimulation checking, our algorithm invokes bisimulation and only run traverses the reachable states of the discrepancy state.
-Therefore, our lazy liveness checking consistently traverses fewer states, even for positive results. 
+The classical algorithm always performs a full traversal of the coalgebra during liveness checking prior to bisimulation checking, our algorithm invokes bisimulation and only run traverses the reachable states of the discrepancy state.
+Therefore, our algorithm consistently traverses fewer states for liveness checking, even for positive results. 
 Additionally, lazy liveness checking enables on-the-fly coalgebra generation using methods like derivatives, which the original algorithm does not support.
@@
 Second, we address the theoretical contribution of our work. 
-While the result may seem practically intuitive, we maintain that our proof is not trivial. 
-We are able to provide straightforward and concise proofs because we introduced a middle ground between coalgebraic theory and practical algorithm, namely greedy bisimulation. 
-Without this framework, verifying Theorem 22 and Theorem 23 would be more challenging, especially given that neither the input coalgebra nor their greedy bisimulation are assumed to be finite.
+While the result may seem intuitive, we maintain that our proof is not trivial. 
+In particular, directly connecting the algorithm with bisimulation on the normalized coalgebra will be challenging (similar to Theorem 20), since we assume neither the input coalgebra nor their bisimulation is finite.
+We have reworked the framework several times to pin-down the intuitive and concise coalgebra structure "greedy bisimulation" to sever as a middle ground, so that most of the proof are either unfolding definitions or results of standard coalgebra manipulations.
 
@@
 
-Regarding the practical aspects of our work, specifically the experiments, we are committed to revising the tests to better represent real-world scenarios before the deadline. 
-However, we also argue that our theoretical results have already demonstrated improvements over existing methods. 
+Regarding experiments in our work, we have updated our test cases with all the "trivial" or dead test cases removed, this development can be found in our repository: https://anonymous.4open.science/r/rust-gkat-071E/ .
+Evidently, our performance are minimally impacted by the new test cases. 
+We hypothesize that because our implementation is so fast, non-negligible amount of time are spent on parsing and hash-consing expressions.
 Plus, trivial test cases do not provide any performance benefit for our algorithm based on Thompson's construction, and dead test cases also do not favor our algorithm against other tools.
-Although we will improve the balance and coverage of our test cases, we believe the current test cases are sufficient in demonstrating the efficiency of our implementation.
 
@@
 
-The benchmark is generated by qcheck library of Ocaml, using the module QCheck2. 
-The generation is based on algebraic rules of GKAT (except the fixpoint rules) with reflexivity and symmetry rule.
+It is generated by https://github.com/c-cube/qcheck, based on congruence, symmetry, and some algebraic rules of GKAT (excluding rules that will be "trivial" tests, as named by Reviewer 3).
 We will include the generation script in the final submission, and provide documentation on test generation
@@
 The correctness result of section IV is based on the correctness result in section III. 
-Besides, there are also cases where the number of primitive tests are small and the algorithm in section III can potentially be more efficient.
+Besides, there can be edge cases (depending on the application) where the number of primitive tests are so small that the algorithm in section III can potentially be more efficient.
 
@@
 
-Dead states are indeed preserved in the normalized coalgebra.
-This approach makes the normalized coalgebra share the same carrier as the original coalgebra, allowing us to define the semantics of dead states also via the unique map into final coalgebra.
+As in all the works in GKAT, dead states are indeed preserved in the normalized coalgebra.
+This approach makes the normalized coalgebra share the carrier as the original coalgebra, allowing the semantics of states to be defined via the unique map into final coalgebra, as it is standard.
 
@@
 Our algorithm is specialized to GKAT, thus able to leverage the deterministic structure to produce smaller automata (coalgebra) and use existing SAT solver to resolve boolean equivalence/inequivalence.
-Our contribution, to the best of our knowledge, is both a theoretical and practical demonstration that GKAT equivalence can be checked more efficiently using a specialized solver.
+Our work is both the first theoretical and practical demonstration that GKAT equivalence can be checked more efficiently than KAT, by using a specialized solver like we designed in this paper.
+
+> GKAT is a bit artificial
+
+GKAT semantics corresponds to trace semantics, a standard semantics of programming language and control-flow analysis.
+Indeed, as you specified, trace equivalence is a stronger equivalence than input-output equivalence.
+However, trace equivalence turns out to be useful in various applications, as we listed in the last paragraph of the related works section, and in response to reviewer 1.
 
@@
 A union-find structure represents an equivalence relation. 
-By employing union-find instead of a set of equivalent pairs, algorithm will search for a greedy bisimulation that is also an equivalence relation. 
+By employing union-find instead of a set of state pairs, our algorithm will effectively search for a greedy bisimulation that is also an equivalence relation. 
 While the correctness of greedy bisimulation is established (Theorem 21), there is no inherent justification for why finding a greedy bisimulation that is also an equivalence relation is equivalent to finding just any greedy bisimulation.
@@
 
-<!-- 
-!!!!!!! NOTES, DO NOT SUBMIT !!!!!!!
-1. Practicality of the tool
-    - Speed up of normalization 
-        - discussion at the end of section III;
-        - better sketch of liveness checking.
-    - Trivial/Dead Equivalence
-2. Comparison with existing work
-    - Intuitive, but coalgebraic characterization is novel
-    - coalgebraic characterization enables straightforward proofs
-    - Theorem 22 is non-trivial because there is no assumption of finiteness
-3. additional results besides automata
-    - We also presented two ways to generate automata in Section V and proved them correct and equivalent.
-    - Automata shape difference, allow acceptation or rejection based on an atom.
-3. Tests generation
-    - We generate the tests roughly based on the shape of real-world program input 
-        - more sequencing than while etc.
-    - negative cases are random expression, but they still need to do liveness check
-    - We will rework the tests to 
-        - reduce the number of trivial and dead tests, 
-        - make sure that the random expression do go into dead state detection immediately
-4. Writing
-    - More about the efficiency of laziness 
--->
+> say somewhere that these are more Antimorov' partial derivatives than Brzozowski's derivatives
+
+Although we do have power set in the signature of symbolic GKAT coalgebra, our derivative do not transition in a non-deterministic manner.
+In fact, we are able to out-perform state-of-the-art tools for KAT equivalence precisely because the GKAT fragment is deterministic.
+We believe that our derivative is closer to Brozozowski style derivative, as we will transition to (at most) a single expression given an input expression, as opposed to a set of expression.

### 2025-03-20 rebuttal.md
revision

@@
 First, we address the efficiency of our lazy normalization approach. 
-The classical algorithm always performs a full traversal of the coalgebra during liveness checking prior to bisimulation checking, our algorithm invokes bisimulation and only run traverses the reachable states of the discrepancy state.
-Therefore, our algorithm consistently traverses fewer states for liveness checking, even for positive results. 
+The classical algorithm performs a full traversal of the coalgebra during liveness checking before bisimulation checking. 
+In contrast, our algorithm invokes bisimulation and only traverses the reachable states of the discrepancy. 
+This results in fewer states being traversed during liveness checking, even for positive results. 
 Additionally, lazy liveness checking enables on-the-fly coalgebra generation using methods like derivatives, which the original algorithm does not support.
@@
 Second, we address the theoretical contribution of our work. 
-While the result may seem intuitive, we maintain that our proof is not trivial. 
-In particular, directly connecting the algorithm with bisimulation on the normalized coalgebra will be challenging (similar to Theorem 20), since we assume neither the input coalgebra nor their bisimulation is finite.
-We have reworked the framework several times to pin-down the intuitive and concise coalgebra structure "greedy bisimulation" to sever as a middle ground, so that most of the proof are either unfolding definitions or results of standard coalgebra manipulations.
+While the result may seem intuitive, our proof is not trivial. 
+Directly connecting the algorithm with bisimulation on the normalized coalgebra, similar to Theorem 21, is challenging; partially because we do not assume that the input coalgebra or bisimulation are finite.
+Our development leads to a more general result than the finitary case. 
+We have reworked the framework several times to pin-down the intuitive notion of "greedy bisimulation", leading to modular results that build up to the final correctness result. 
+The effectiveness of greedy bisimulation as a theoretical concept is demonstrated by the simplicity of most proofs, which are either unfoldings of definitions or standard exercises in coalgebra.
 
-Finally, we emphasize that our contribution extends beyond bisimulation of machines (or coalgebras).
+Finally, we emphasize that our contribution extends beyond bisimulation of coalgebras (or machines).
 In Section V, we present two methods for generating coalgebras from expressions and prove their correctness, finiteness, and correspondence. 
-The finiteness result, in particular, arises naturally from the correspondence result. 
-The simplicity of this result is noted by one of the reviewers.
+The finiteness result elegantly arises from the correspondence result, a point noted by one of the reviewers.
 
-Regarding experiments in our work, we have updated our test cases with all the "trivial" or dead test cases removed, this development can be found in our repository: https://anonymous.4open.science/r/rust-gkat-071E/ .
-Evidently, our performance are minimally impacted by the new test cases. 
-We hypothesize that because our implementation is so fast, non-negligible amount of time are spent on parsing and hash-consing expressions.
-Plus, trivial test cases do not provide any performance benefit for our algorithm based on Thompson's construction, and dead test cases also do not favor our algorithm against other tools.
+Regarding experiments in our work, We have reworked our test cases, removing all "trivial" or dead test cases. 
+Evidently, our performance is minimally impacted by the new test cases. 
+We hypothesize that our implementation is so fast that a non-negligible amount of time is spent on parsing and hash-consing expressions, causing the performance impact of "trivial" tests to be minimal.
+Additionally, "trivial" tests do not provide any performance benefit for the algorithm based on Thompson's construction, and dead test cases do not favor our algorithms against other tools.
 
@@
 As in all the works in GKAT, dead states are indeed preserved in the normalized coalgebra.
-This approach makes the normalized coalgebra share the carrier as the original coalgebra, allowing the semantics of states to be defined via the unique map into final coalgebra, as it is standard.
+This approach allows the semantics of dead states to be defined in the standard manner, i.e. via the unique map into final coalgebra.
 
@@
 SymKAT is a symbolic equivalence checker for Kleene Algebra with Tests (KAT). 
-Our algorithm is specialized to GKAT, thus able to leverage the deterministic structure to produce smaller automata (coalgebra) and use existing SAT solver to resolve boolean equivalence/inequivalence.
-Our work is both the first theoretical and practical demonstration that GKAT equivalence can be checked more efficiently than KAT, by using a specialized solver like we designed in this paper.
+Our algorithm is specialized for GKAT, leveraging its deterministic structure to produce smaller automata (coalgebras) and utilizing existing SAT solvers to resolve boolean equivalence and inequivalence.
+This work represents the first theoretical and practical demonstration that GKAT equivalence can be checked more efficiently than KAT by employing a specialized solver, as designed in this paper.
 
@@
 
-GKAT semantics corresponds to trace semantics, a standard semantics of programming language and control-flow analysis.
-Indeed, as you specified, trace equivalence is a stronger equivalence than input-output equivalence.
-However, trace equivalence turns out to be useful in various applications, as we listed in the last paragraph of the related works section, and in response to reviewer 1.
+GKAT semantics is based on trace semantics, a standard semantics used in programming languages and control-flow analysis. 
+As you noted, trace equivalence is a stronger equivalence than input-output equivalence. 
+However, trace equivalence proves to be useful in various applications, as we detailed in the last paragraph of the related work section and in response to Reviewer 1.
+In fact, we foresee that our works here can be adapted to these applications and greatly improve their performance.
 
@@
 The first benefit stems from the laziness of our liveness checking algorithm, allowing us to iterate through fewer states in the classical algorithm (see overall response and the next comment).  
-This benefit is apparent no matter the result or the generation method of the coalgebra. 
+This benefit exists regardless of the result of the algorithm or the generation method of the coalgebra. 
 The second benefit stems from the on-the-fly nature of our algorithm.
-Specifically, our algorithm can terminate without generating the rest of the coalgebra, when a counter-example is encountered while applying an on-the-fly generation method (such as derivatives).
+Specifically, while we encounter a counter-example while using derivative to generate the coalgebra on-the-fly, our algorithm can terminate without generating the rest of the coalgebra.
 In contrast, the original algorithm requires the complete generation of the coalgebra in order to perform liveness checking first.
@@
 We are uncertain why the original algorithm might be faster in this specific case. 
-The original algorithm requires determining the liveness of every single state before performing bisimulation. 
-In contrast, our algorithm only conducts liveness checking on a few states when the bisimulation algorithm detects a discrepancy.
-Consider two bisimilar states (not just finite-trace equivalent) in a GKAT coalgebra composed entirely of live states. 
-The classical algorithm would first perform a liveness check, iterating through all the states, and then proceed with bisimulation, iterating through all the states again. 
-Our algorithm, however, simply performs the bisimulation and confirms there is no discrepancy, effectively bypassing the liveness check altogether.
+Consider two bisimilar states (infinite-trace equivalent) in a GKAT coalgebra with no dead state. 
+The classical algorithm would perform a liveness check, iterating through all the states; then proceed with bisimulation, iterating through all the states again. 
+Our algorithm simply performs the bisimulation; because these two states are bisimilar, our algorithm will bypass the liveness check altogether.
 
@@
 
-Although we do have power set in the signature of symbolic GKAT coalgebra, our derivative do not transition in a non-deterministic manner.
-In fact, we are able to out-perform state-of-the-art tools for KAT equivalence precisely because the GKAT fragment is deterministic.
-We believe that our derivative is closer to Brozozowski style derivative, as we will transition to (at most) a single expression given an input expression, as opposed to a set of expression.
+While we employ the power set in the signature of the symbolic GKAT coalgebra, our derivatives do not transition in a non-deterministic manner. 
+Specifically, when given an GKAT expression and a (valid) boolean expression, derivative will deterministically transition to a GKAT expression.
+In this sense, our notion of derivative is closer to Brzozowski style than that of Antimorov.

### 2025-03-20 main.tex
minor bug
typo fix

@@
 Although the current algorithm scales nearly-linearly with the size of the expression, its performance is hindered by the need to compute the entire coalgebra for the sake of normalization and exponential performance degradation when expanding the set of primitive tests.
-To address these problems, we introduce two new algorithms for bisimulation checking: an on-the-fly bisimulation algorithm based on greedy bisimulation, which enables immediate termination upon discovering a counter-example and supports on-the-fly coalgebra generation with derivatives; and a symbolic algorithm that builds upon the on-the-fly algorithm while leveraging efficient SAT solvers to speed up boolean comparisons and compress the transition structures of GKAT coalgebra.
+To address these problems, we introduce two new algorithms for bisimulation checking: an on-the-fly bisimulation algorithm based on greedy bisimulation, a coalgebriac representation of our algorithm; and its symbolic version.
+Our algorithm enables immediate termination upon discovering a counter-example and supports on-the-fly coalgebra generation with derivatives; and a symbolic algorithm that builds upon the on-the-fly algorithm while leveraging efficient SAT solvers to speed up boolean comparisons and compress the transition structures of GKAT coalgebra.
 Additionally, to utilize the symbolic bisimulation checking, we provide two methods to generate symbolic GKAT coalgebras from GKAT expressions, analogous to derivative and Thompson's construction in previous works.
@@
 
-Given any bisimulation \({∼} ⊆ S × U\), we can construct a GKAT coalgebra \(∼_d\) s.t.
+Given any bisimulation \({∼} ⊆ \norm(S) × \norm(U)\), we can construct a GKAT coalgebra \(∼_d\) s.t.
 it is \emph{almost} a greedy bisimulation, i.e.
@@
             & ∃ b ∈ \BExp_T, α ≤ b \\
-            & \text{ and } δ(s, b) = (s', p)  
+            & \text{ and } δ̂(s, b) = (s', p)  
         \end{aligned}\\[5px]

### 2025-03-20 rebuttal.md
more response to reviewer 2

@@
 
+> "least GKAT coalgebra" doesn't make sense.
+
+GKAT coalgebra is naturally ordered with sub-GKAT coalgebra relation; then a least GKAT coalgebra satisfying a certain property is a sub-GKAT coalgebra of all the GKAT coalgebra satisfying said property.
+Note that we do not assume such GKAT coalgebra always exists for all properties, the existence and uniqueness of greedy bisimulation is proved later.
+
+> It should be least relation containing ~ and closed under the transition relation.
+
+This also would lead to the same coalgebra. 
+Treating greedy bisimulation itself as a GKAT coalgebra allows us to reuse several prior results, like homomorphism preserves liveness.
+In fact, we believe that showing the existence of such ~ showed also require a theorem similar to Lemma 15, and then use the fact that subsets of ~_d form a complete lattice.
+The proof above is essentially the same as our existence proof.
+
 > GKAT is a bit artificial
@@
 In fact, we foresee that our works here can be adapted to these applications and greatly improve their performance.
+
+> Overall, I didn't quite understand, do you need to use bisimulations, or could you have start states and just compare traces directly?
+
+The algebraic treatment of bisimulation should coincide the normal automata theoretical intuition, at least in the finite case.
+However, by treating (greedy) bisimulation as coalgebras, we are able to use theorems like homomorphism preserves liveness or the functoriality of normalization.
+The coalgebraic treatment makes our proofs more modular, and each proof more concise and straightforward, which we thought LICS appreciates.
 

### 2025-03-20 rebuttal.md
add couple sentences

@@
 This results in a distinct semantics where a trace's end is marked by an atom, representing the output program state.
+In general, we feel our work is a novel contribution, independent of the prior works on symbolic transducers or symbolic coalgebra.
 
-While symbolic bisimulation is pivotal for the efficiency of our equivalence algorithm, it is not our sole theoretical contribution. 
+While symbolic algorithm is pivotal for scaling our equivalence algorithm to more primitive tests, it is not our sole contribution. 
 We have also developed:

### 2025-03-20 rebuttal.md
revise the rebuttal

@@
 
-We would like to express our sincere gratitude for the constructive feedback received from the reviewer. 
+We would like to express our sincere gratitude for the constructive feedback received from the reviewers. 
 We will carefully address the points raised and implement the necessary changes to our paper.
@@
 First, we address the efficiency of our lazy normalization approach. 
-The classical algorithm performs a full traversal of the coalgebra during liveness checking before bisimulation checking. 
-In contrast, our algorithm invokes bisimulation and only traverses the reachable states of the discrepancy. 
-This results in fewer states being traversed during liveness checking, even for positive results. 
+The classical algorithm performs a full traversal of the coalgebra during liveness checking before bisimulation. 
+In contrast, our algorithm invokes bisimulation and only traverses the reachable states of the discrepancies. 
+This change results in fewer states being traversed during liveness checking, even for positive results. 
 Additionally, lazy liveness checking enables on-the-fly coalgebra generation using methods like derivatives, which the original algorithm does not support.
@@
 Second, we address the theoretical contribution of our work. 
-While the result may seem intuitive, our proof is not trivial. 
-Directly connecting the algorithm with bisimulation on the normalized coalgebra, similar to Theorem 21, is challenging; partially because we do not assume that the input coalgebra or bisimulation are finite.
-Our development leads to a more general result than the finitary case. 
-We have reworked the framework several times to pin-down the intuitive notion of "greedy bisimulation", leading to modular results that build up to the final correctness result. 
-The effectiveness of greedy bisimulation as a theoretical concept is demonstrated by the simplicity of most proofs, which are either unfoldings of definitions or standard exercises in coalgebra.
+While our algorithm may seem intuitive, our proof is not trivial. 
+Directly connecting the algorithm with bisimulation on the normalized coalgebra, as in Theorem 21, is challenging; partially because we do not assume that the input coalgebra or bisimulation are finite.
+We have reworked the framework several times to pin-down the concise notion of "greedy bisimulation", leading to modular results that build up to the final correctness result. 
+The effectiveness of greedy bisimulation as a theoretical tool is demonstrated by the generality of our result, and the simplicity of most proofs, which are either unfoldings of definitions or standard exercises in coalgebra.
 
@@
 Evidently, our performance is minimally impacted by the new test cases. 
-We hypothesize that our implementation is so fast that a non-negligible amount of time is spent on parsing and hash-consing expressions, causing the performance impact of "trivial" tests to be minimal.
-Additionally, "trivial" tests do not provide any performance benefit for the algorithm based on Thompson's construction, and dead test cases do not favor our algorithms against other tools.
+By analyzing the algorithm, we can see that "trivial" tests do not provide any performance benefit for the algorithm based on Thompson's construction, and dead test cases do not favor our algorithms against other tools.
 
@@
 
-It is generated by https://github.com/c-cube/qcheck, based on congruence, symmetry, and some algebraic rules of GKAT (excluding rules that will be "trivial" tests, as named by Reviewer 3).
+Benchmarks are generated by https://github.com/c-cube/qcheck, based on congruence, symmetry, and some algebraic rules of GKAT (excluding rules that will be "trivial" tests, as named by Reviewer 3).
 We will include the generation script in the final submission, and provide documentation on test generation
@@
 We have also developed:
-- An on-the-fly algorithm for bisimulation, and proved its correctness using greedy bisimulation, a novel coalgebra construction.
+- An on-the-fly algorithm for bisimulation, and proved its correctness using greedy bisimulation, a novel and clean coalgebra construction.
 - Two distinct methods to construct symbolic GKAT coalgebras, while also proved their correctness, finiteness, and correspondence.
@@
 
-GKAT has been used in several important verification tasks in the domain of networks, control-flow transformation, probabilistic computation, and program logic.
+GKAT has been used in several important verification tasks for networks, control-flow transformation, probabilistic computation, and program logic.
 We mentioned this fact at the end of the related work section. 
@@
 Besides, there can be edge cases (depending on the application) where the number of primitive tests are so small that the algorithm in section III can potentially be more efficient. 
+We will make sure that the structure of the paper is more evident in the introduction section and the first paragraph of each section.
 
@@
 This approach allows the semantics of dead states to be defined in the standard manner, i.e. via the unique map into final coalgebra.
+We will provide better explanations in the background sections.
 
@@
 This work represents the first theoretical and practical demonstration that GKAT equivalence can be checked more efficiently than KAT by employing a specialized solver, as designed in this paper.
+We will make the distinction and the contribution more clear in the final version.
 
@@
 
-GKAT coalgebra is naturally ordered with sub-GKAT coalgebra relation; then a least GKAT coalgebra satisfying a certain property is a sub-GKAT coalgebra of all the GKAT coalgebra satisfying said property.
+GKAT coalgebra is naturally ordered with sub-GKAT coalgebra relation; then the least GKAT coalgebra satisfying a certain property is a sub-GKAT coalgebra of all the GKAT coalgebra satisfying said property.
+We will provide this definition after the definition of greedy bisimulation.
 Note that we do not assume such GKAT coalgebra always exists for all properties, the existence and uniqueness of greedy bisimulation is proved later.
@@
 In fact, we foresee that our works here can be adapted to these applications and greatly improve their performance.
+We will add a sentence stating the relevant applications of GKAT in the introduction section.
 
@@
 
-The algebraic treatment of bisimulation should coincide the normal automata theoretical intuition, at least in the finite case.
+The algebraic treatment of bisimulation should coincide with the normal automata theoretical intuition, at least in the finite case.
 However, by treating (greedy) bisimulation as coalgebras, we are able to use theorems like homomorphism preserves liveness or the functoriality of normalization.
 The coalgebraic treatment makes our proofs more modular, and each proof more concise and straightforward, which we thought LICS appreciates. 
+We will more clearly detail that our approach serves as a different approach than the automata theoretical intuition, while generating similar algorithms.
+
+> For on-the-fly generally, do you want to cite some earlier work, e.g. Fernandez and Mounier, CAV 1991?
+
+As the core challenge of our on-the-fly algorithm is normalization, we see our work as orthogonal to the work of Fernandez and Mounier.
+However, we see that them as valuable references. 
+In the final version, we will provide comparison between their work and ours in the related works section.
 
@@
 To address this, we will apply Theorem 23 to demonstrate that there exists a greedy bisimulation that is also an equivalence relation if and only if two states are trace equivalent.
+We will add more detailed explanation of how Theorem 23 and 21 interacts with each other.
 
@@
 In contrast, the original algorithm requires the complete generation of the coalgebra in order to perform liveness checking first.
+In the final version, we will provide more detail and examples for these two benefits.
 
@@
 Our algorithm simply performs the bisimulation; because these two states are bisimilar, our algorithm will bypass the liveness check altogether.
+Additionally, the main performance bottleneck of our algorithm would be boolean (in)equivalence checks.
+These checks are present in both liveness checking and bisimulation, and overshadows the time spent to look up in a union-find structure in practice.
 
@@
 Specifically, when given an GKAT expression and a (valid) boolean expression, derivative will deterministically transition to a GKAT expression.
-In this sense, our notion of derivative is closer to Brzozowski style than that of Antimorov.
+In this sense, our notion of derivative is closer to Brzozowski style than that of Antimorov, which we will detail in the final version

### 2025-04-14 LICS25/rebuttal-600words.md
move past to LICS 2025 folder

--- 2025-03-19 rebuttal-600words.md
@@
 While our algorithms seem intuitive, our proofs are not trivial. 
-The simple proofs are enabled by the design of greedy bisimulation. 
+Specifically, the simple proofs are enabled by the design of greedy bisimulation, which took several iterations.
 Without this middle ground, verifying Theorem 22 would be challenging, especially given that neither the input coalgebras nor their bisimulation are assumed to be finite.
@@
 Finally, we emphasize that our contribution extends beyond bisimulation of coalgebras.
-We also presented two methods for generating coalgebras from expressions and proved their correctness, finiteness, and correspondence.
+We also presented two methods for generating coalgebras from expressions; and proved their correctness, finiteness, and correspondence.
 
-Regarding our experiments, we have revised our tests in our repository. 
-We also emphasize that neither trivial nor dead tests provide performance benefit in our current implementation, so submitted tests are reflective of the general cases. 
-As mentioned before, we have conceptual demonstrations of performance improvements in the submitted version.
+Regarding our experiments, we have uploaded new tests to our repository, and will provide the code and documentation for test generation.
+The performance preserves the efficiency of previous versions, as neither trivial nor dead tests provide performance benefit in our implementation. 
+Additionally, we already have several conceptual demonstrations of performance improvements in the paper.
 
@@
 
-It is generated by https://github.com/c-cube/qcheck, based on algebraic rules of GKAT (excluding fixpoint rule) with reflexivity, congruence, and symmetry rules.
-We will include the generation script and provide documentation.
-
-> Unnecessary but harmful generalization
-
-We will clarify that the functor `F` satisfies necessary properties.
+It is generated by https://github.com/c-cube/qcheck, based on congruence, symmetry, and some algebraic rules of GKAT (excluding rules that will be "trivial" tests, as named by Reviewer 3).
 
@@
 Our approach diverges from previous works on symbolic transducers in several aspects:
-- Our on-the-fly method is tailored for normalization, a unique challenge in achieving finite trace equivalence for GKAT. 
+- Our on-the-fly method is tailored for normalization, a challenge that is not shared by symbolic transducers. 
 - The signature of GKAT differs from traditional symbolic transducers.

### 2025-04-14 LICS25/rebuttal.md
move past to LICS 2025 folder

--- 2025-03-20 rebuttal.md
@@
 Regarding experiments in our work, We have reworked our test cases, removing all "trivial" or dead test cases. 
-Evidently, our performance is minimally impacted by the new test cases. 
-By analyzing the algorithm, we can see that "trivial" tests do not provide any performance benefit for the algorithm based on Thompson's construction, and dead test cases do not favor our algorithms against other tools.
+Our performance is minimally impacted by the new test cases. 
+By analyzing the algorithm, we can also see that "trivial" tests do not provide any performance benefit for the algorithm based on Thompson's construction, and dead test cases do not favor our algorithms against other tools.
 

### 2025-04-14 POPL26/paper.tex
add ACM template

%%
%% This is file `sample-sigconf-lualatex.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `all,proceedings,bibtex,sigconf')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigconf-lualatex.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%%
%% Commands for TeXCount
%TC:macro \cite [option:text,text]
%TC:macro \citep [option:text,text]
%TC:macro \citet [option:text,text]
%TC:envir table 0 1
%TC:envir table* 0 1
%TC:envir tabular [ignore] word
%TC:envir displaymath 0 word
%TC:envir math 0 word
%TC:envir comment 0 0
%%
%% The first command in your LaTeX source must be the \documentclass
%% command.
%%
%% For submission and review of your manuscript please change the
%% command to \documentclass[manuscript, screen, review]{acmart}.
%%
%% When submitting camera ready or to TAPS, please change the command
%% to \documentclass[sigconf]{acmart} or whichever template is required
%% for your publication.
%%
%%
\documentclass[manuscript,screen,review,anonymous]{acmart}
%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    Bib\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmlicensed}
\copyrightyear{2018}
\acmYear{2018}
\acmDOI{XXXXXXX.XXXXXXX}
%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[Conference acronym 'XX]{Make sure to enter the correct
  conference title from your rights confirmation email}{June 03--05,
  2018}{Woodstock, NY}
%%
%%  Uncomment \acmBooktitle if the title of the proceedings is different
%%  from ``Proceedings of ...''!
%%
%%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
%%  June 03--05, 2018, Woodstock, NY}
\acmISBN{978-1-4503-XXXX-X/2018/06}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% For managing citations, it is recommended to use bibliography
%% files in BibTeX format.
%%
%% You can then either use BibTeX with the ACM-Reference-Format style,
%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
%% support for advanced citation of software artefact from the
%% biblatex-software package, also separately available on CTAN.
%%
%% Look at the sample-*-biblatex.tex files for templates showcasing
%% the biblatex styles.
%%

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{Efficient Algorithms For Variants of GKAT}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{TODO}
% \author{Ben Trovato}
% \authornote{Both authors contributed equally to this research.}
% \email{trovato@corporation.com}
% \orcid{1234-5678-9012}
% \author{G.K.M. Tobin}
% \authornotemark[1]
% \email{webmaster@marysville-ohio.com}
% \affiliation{%
%   \institution{Institute for Clarity in Documentation}
%   \city{Dublin}
%   \state{Ohio}
%   \country{USA}
% }

% \author{Lars Th{\o}rv{\"a}ld}
% \affiliation{%
%   \institution{The Th{\o}rv{\"a}ld Group}
%   \city{Hekla}
%   \country{Iceland}}
% \email{larst@affiliation.org}

% \author{Valerie B\'eranger}
% \affiliation{%
%   \institution{Inria Paris-Rocquencourt}
%   \city{Rocquencourt}
%   \country{France}
% }

% \author{Aparna Patel}
% \affiliation{%
%  \institution{Rajiv Gandhi University}
%  \city{Doimukh}
%  \state{Arunachal Pradesh}
%  \country{India}}

% \author{Huifen Chan}
% \affiliation{%
%   \institution{Tsinghua University}
%   \city{Haidian Qu}
%   \state{Beijing Shi}
%   \country{China}}

% \author{Charles Palmer}
% \affiliation{%
%   \institution{Palmer Research Laboratories}
%   \city{San Antonio}
%   \state{Texas}
%   \country{USA}}
% \email{cpalmer@prl.com}

% \author{John Smith}
% \affiliation{%
%   \institution{The Th{\o}rv{\"a}ld Group}
%   \city{Hekla}
%   \country{Iceland}}
% \email{jsmith@affiliation.org}

% \author{Julius P. Kumquat}
% \affiliation{%
%   \institution{The Kumquat Consortium}
%   \city{New York}
%   \country{USA}}
% \email{jpkumquat@consortium.net}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Trovato et al.}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  TODO
\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>00000000.0000000.0000000</concept_id>
  <concept_desc>Do Not Use This Code, Generate the Correct Terms for Your Paper</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>00000000.00000000.00000000</concept_id>
  <concept_desc>Do Not Use This Code, Generate the Correct Terms for Your Paper</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>00000000.00000000.00000000</concept_id>
  <concept_desc>Do Not Use This Code, Generate the Correct Terms for Your Paper</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>00000000.00000000.00000000</concept_id>
  <concept_desc>Do Not Use This Code, Generate the Correct Terms for Your Paper</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Do Not Use This Code~Generate the Correct Terms for Your Paper}
\ccsdesc[300]{Do Not Use This Code~Generate the Correct Terms for Your Paper}
\ccsdesc{Do Not Use This Code~Generate the Correct Terms for Your Paper}
\ccsdesc[100]{Do Not Use This Code~Generate the Correct Terms for Your Paper}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{Do, Not, Us, This, Code, Put, the, Correct, Terms, for,
  Your, Paper}
%% A "teaser" image appears between the author and affiliation
%% information and the body of the document, and typically spans the
%% page.

\received{20 February 2007}
\received[revised]{12 March 2009}
\received[accepted]{5 June 2009}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\section{Introduction}

\section{Control-Flow GKAT}

\section{Guarded NetKAT}

\end{document}
\endinput
%%
%% End of file `sample-sigconf-lualatex.tex'.

### 2025-04-14 POPL26/paper.tex
add bib

@@
 %%
-\documentclass[manuscript,screen,review,anonymous]{acmart}
+\documentclass[
+  manuscript,screen,review,anonymous,
+  natbib=false]{acmart}
 %%
@@
 
+% Biblatex
+\RequirePackage[
+  datamodel=acmdatamodel,
+  style=acmnumeric,
+  ]{biblatex}
+%% Declare bibliography sources (one \addbibresource command per source)
+\addbibresource{refs.bib}
+
+% For adding inline comments in the text.
+\usepackage[margin=false,inline=true]{fixme}
+\FXRegisterAuthor{aaa}{anaaa}{\color{cyan}AAA}
+\FXRegisterAuthor{mg}{anmg}{\color{red}MG}
+\FXRegisterAuthor{cz}{ancz}{\color{orange}CZ}
+% \newcommand{\aaa}[1]{\aaanote{#1}}
+% \newcommand{\mg}[1]{\mgnote{#1}}
+% \newcommand{\cz}[1]{\cznote{#1}}
+\newcommand{\aaa}[1]{}
+\newcommand{\mg}[1]{}
+\newcommand{\cz}[1]{}
+
+\usepackage{stmaryrd}
+
+\usepackage{stackengine}
+\usepackage{mathrsfs}
+\usepackage{braket}
+\usepackage{annotate-equations}
+\usepackage{scalerel}
+
+% graphs
+\usepackage{tikz}
+\usetikzlibrary{arrows,automata,positioning}
+
+% commutative diagram
+\usepackage{tikz-cd}
+
+% ref
+\usepackage{hyperref}
+\usepackage{cleveref}
+
+% inference rule
+\usepackage{mathpartir}
+% cross-referencing infer rule
+% based on https://tex.stackexchange.com/questions/340788/cross-referencing-inference-rules
+\makeatletter
+\let\originferrule\inferrule
+\DeclareDocumentCommand \inferrule { s O {} m m}{%
+  \IfBooleanTF{#1}%
+  {%
+    \mpr@inferstar[#2]{#3}{#4}%
+  }{%
+    \mpr@inferrule[#2]{#3}{#4}%
+  }%
+  \IfValueT{#2}%
+  {%
+    \my@name@inferrule{#2}%
+  }%
+}
+\NewDocumentCommand \my@name@inferrule { m }{%
+  \def\@currentlabelname{\textsc{#1}}%
+}
+\makeatother
+
+% item spacing
+\usepackage{enumitem}
+
+% for code
+\usepackage{listings}
+\lstset{basicstyle=\footnotesize\ttfamily,breaklines=true}
+\lstset{framextopmargin=50pt,frame=bottomline}
+\lstset{language=caml, escapeinside={[*}{*]}}
+% Clever ref names
+\crefname{lstlisting}{listing}{listings}
+\Crefname{lstlisting}{Listing}{Listings}
+
+% for algorithm
+\usepackage{algorithm}
+\usepackage[noend]{algpseudocode}
+% switch statement for pattern matching
+\algnewcommand\algorithmicmatch{\textbf{match}}
+\algnewcommand\algorithmicwith{\textbf{with}}
+\algnewcommand\algorithmiccase{\textbf{case}}
+\algnewcommand\algorithmicdefault{\textbf{default}}
+\algnewcommand\Continue{\textbf{continue}}
+\algdef{SE}[MATCH]{Match}{EndMatch}[1]{\algorithmicmatch\ #1\ \algorithmicwith}{\algorithmicend\ \algorithmicmatch}%
+\algdef{SE}[CASE]{Case}{EndCase}[1]{\algorithmiccase\ #1 \algorithmicthen}{\algorithmicend\ \algorithmiccase}%
+\algdef{SE}[DEFAULT]{Default}{EndDefault}{\hskip\algorithmicindent\algorithmicdefault}{\algorithmicend\algorithmicdefault}%
+\algtext*{EndMatch}%
+\algtext*{EndCase}%
+\algtext*{EndDefault}%
+
+% for better table
+\usepackage{booktabs}
+
+% subcaption
+\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
+                        %% http://ctan.org/pkg/subcaption
+
+% unicode math symbols
+\usepackage{unicode-math}
+% support for hat, overline, underline, vec, and sim combining charactors
+\protected\def\afteracc{\directlua{
+    local nest = tex.nest[tex.nest.ptr]
+    local last = nest.tail
+    if not (last and last.id == 18) then
+      error'I can only put accents on simple noads.'
+    end
+    if last.sub or last.sup then
+      error'If you want accents on a superscript or subscript, please use braces.'
+    end
+    local acc = node.new(21, 1)
+    acc.nucleus = last.nucleus
+    last.nucleus = nil
+    local is_bottom = token.scan_keyword'bot' and 'bot_accent' or 'accent'
+    acc[is_bottom] = node.new(23)
+    acc[is_bottom].fam, acc[is_bottom].char = 0, token.scan_int()
+    nest.head = node.insert_after(node.remove(nest.head, last), nil, acc)
+    nest.tail = acc
+    node.flush_node(last)
+  }}
+\AtBeginDocument{
+\begingroup
+  \def\UnicodeMathSymbol#1#2#3#4{%
+    \ifx#3\mathaccent
+      \def\mytmpmacro{\afteracc#1 }%
+      \global\letcharcode#1=\mytmpmacro
+      \global\mathcode#1="8000
+    \else\ifx#3\mathbotaccentwide
+      \def\mytmpmacro{\afteracc bot#1 }%
+      \global\letcharcode#1=\mytmpmacro
+      \global\mathcode#1="8000
+    \fi\fi
+  }
+  \input{unicode-math-table}
+\endgroup
+}
+
+% math font, this is needed to render \setminus command
+\setmathfont{latinmodern-math}
+\setmathfont[range=\setminus]{STIX Two Math}
+\setmathfont[range=\similarrightarrow]{STIX Two Math}
+
+\newtheorem{theorem}{Theorem}
+\newtheorem{corollary}[theorem]{Corollary}
+\newtheorem{lemma}[theorem]{Lemma}
+
+\newtheorem{definition}{Definition}
+\newtheorem*{definition*}{Definition}
+\newtheorem{example}[definition]{Example}
+\newtheorem{remark}[definition]{Remark}
+
+%%%% Macros %%%%%
+
+% Math?
+\newcommand{\true}{\mathrm{true}}
+\newcommand{\false}{\mathrm{false}}
+\newcommand{\At}{\mathbf{At}}
+
+% operators
+\newcommand{\dom}[1]{\mathrm{dom}(#1)}
+\newcommand{\cod}[1]{\mathrm{cod}(#1)}
+\DeclareMathOperator{\post}{\mathrm{post}}
+\newcommand{\reject}{\mathinner{\mathrm{rej}}}
+\newcommand{\accept}{\mathinner{\mathrm{acc}}}
+\DeclareMathOperator*{\bigplus}{\scalerel*{+}{\sum}}
+\newcommand{\clos}[1]{\mathrel{\overline{#1}}}
+\DeclareMathOperator{\norm}{\mathrm{norm}}
+\DeclareMathOperator{\dead}{\mathrm{dead}}
+\DeclareMathOperator{\symb}{\mathrm{symb}}
+\DeclareMathOperator{\unsymb}{\symb^{-1}}
+
+% commands 
+\newcommand{\command}[1]{{\mathtt{#1}}}
+\newcommand{\comAssume}[1]{\command{assume}~#1}
+\newcommand{\comITE}[3]{\command{if}~#1~\command{then}~#2~\command{else}~#3}
+\newcommand{\comWhile}[2]{\command{while}~#1~\command{do}~#2}
+
+% set of models
+\newcommand{\theoryOf}[1]{\ensuremath{\mathsf{#1}}}
+\newcommand{\Exp}{\theoryOf{Exp}}
+\newcommand{\BExp}{\theoryOf{BExp}}
+\DeclareMathOperator{\GS}{\mathrm{GS}}
+
+\newcommand\altxrightarrow[2][0pt]{\mathrel{\ensurestackMath{\stackengine%
+  {\dimexpr#1-7.5pt}{\xrightarrow{\phantom{#2}}}{\scriptstyle\!#2\,}%
+  {O}{c}{F}{F}{S}}}}
+\newcommand{\transvia}[1]{
+    \mathrel{\raisebox{-2px}{\(\altxrightarrow[-2px]{#1}\)}}
+}
+\newcommand{\transAcc}[2]{⇒_{#1} #2}
+\newcommand{\transRej}[2]{↓_{#1} #2}
+
+%%%%%%%%%%
 %% Rights management information.  This information is sent to you
@@
 
+Control-Flow GKAT (or CF-GKAT) extends Guarded Kleene Algebra with tests with common control-flow operators, including indicator variable, break, continue, and goto's.
+
+Similar to GKAT, the syntax of CF-GKAT is also two sorted:
+
 \section{Guarded NetKAT}

### 2025-04-14 POPL26/paper.tex
accept bibtex as our lord and savor

@@
 %%
-\documentclass[
-  manuscript,screen,review,anonymous,
-  natbib=false]{acmart}
+\documentclass[manuscript,screen,review,anonymous]{acmart}
 %%
@@
     Bib\TeX}}}
-
-% Biblatex
-\RequirePackage[
-  datamodel=acmdatamodel,
-  style=acmnumeric,
-  ]{biblatex}
-%% Declare bibliography sources (one \addbibresource command per source)
-\addbibresource{refs.bib}
 
@@
 
-Control-Flow GKAT (or CF-GKAT) extends Guarded Kleene Algebra with tests with common control-flow operators, including indicator variable, break, continue, and goto's.
+Control-Flow GKAT (or CF-GKAT)~\cite{zhang_CFGKATEfficientValidation_2025} extends Guarded Kleene Algebra with tests with common control-flow operators, including indicator variable, break, continue, and goto's.
 
 Similar to GKAT, the syntax of CF-GKAT is also two sorted:
-
+\begin{align*}
+  \BExp & ≜ t ∈ T ∣ 
+\end{align*}
 \section{Guarded NetKAT}
 
+\bibliographystyle{ACM-Reference-Format}
+\bibliography{refs}
 \end{document}

### 2025-04-15 POPL26/paper.tex
derivative for Symbolic Guarded NetKAT

@@
 
+% commands
+\newcommand{\command}[1]{\texttt{#1}}
+\newcommand{\comAssert}[1]{\command{assert}~#1}
+\newcommand{\comSkip}{\command{skip}}
+\newcommand{\comITE}[3]{\command{if}~#1~\command{then}~#2~\command{else}~#3}
+\newcommand{\comIT}[2]{\command{if}~#1~\command{then}~#2}
+\newcommand{\comWhile}[2]{\command{while}~#1~\command{do}~#2}
+\newcommand{\comLabel}[1]{\command{label}~#1}
+\newcommand{\comBrk}{\command{break}}
+\newcommand{\comCont}{\command{continue}}
+\newcommand{\comRet}{\command{return}}
+\newcommand{\comGoto}[1]{\command{goto}~#1}
+\newcommand{\comDup}{\command{dup}}
+
 % operators
@@
 
-% commands 
-\newcommand{\command}[1]{{\mathtt{#1}}}
-\newcommand{\comAssume}[1]{\command{assume}~#1}
-\newcommand{\comITE}[3]{\command{if}~#1~\command{then}~#2~\command{else}~#3}
-\newcommand{\comWhile}[2]{\command{while}~#1~\command{do}~#2}
+% continuations
+\newcommand{\conti}[1]{\mathbf{#1}}
+\newcommand{\acc}[1]{{\conti{acc} ~ #1}}  % accept with an output indicator value
+\newcommand{\setc}[1]{{\conti{set} ~ #1}}  % set variables in NetKAT
+\newcommand{\retc}{{\conti{ret}}}  % return, terminate the program
+\newcommand{\brkc}[1]{\conti{brk} ~ #1}  % break with an output indicator value
+\newcommand{\jmpc}[1]{{\conti{jmp} ~ #1}} % goto a label
 
@@
 \newcommand{\Exp}{\theoryOf{Exp}}
+\newcommand{\Prog}{\theoryOf{Prog}}
 \newcommand{\BExp}{\theoryOf{BExp}}
@@
 }
-\newcommand{\transAcc}[2]{⇒_{#1} #2}
+\newcommand{\transOut}[2]{⇒_{#1} #2}
 \newcommand{\transRej}[2]{↓_{#1} #2}
@@
 %% allowing the author to define a "short title" to be used in page headers.
-\title{Efficient Algorithms For Variants of GKAT}
+\title{SyGO: Generic On-the-fly Symbolic Algorithms For Variants of GKAT}
 
@@
 
+\section{Overview}
+
+\section{GKAT with Output}
+
+Symbolic GKAT with Output (SyGO) automata extends GKAT automata, where instead of only emitting accepts, it allows to produce a finite set of outputs \(O\).
+In the case of GKAT and Guarded NetKAT, the output set \(O\) is a singleton set \(\acc{}\); whereas in CF-GKAT, the outcome can be any continuations.
+
+Formally, a SyGO Automata is defined over a set of tests \(\BExp\), a set of outputs \(O\), and a set of state \(S\);  
+with a start state \(ŝ ∈ S\), and two functions:
+\begin{align*}
+  ϵ & : S → 𝒫(\BExp × O) & 
+  δ & : S → 𝒫(\BExp × S × Σ).
+\end{align*}
+Intuitively, \(ϵ(s, o)\) returns the condition required to immediately return the output \(o\) when at the state \(s\); and if \((b, s') ∈ δ(s)\) then it means that from state \(s\), when condition \(b\) is met, the machine will transition to \(s'\).
+
+To ensure the determinism of SyGO automata, we will impose a \emph{disjointness condition}, i.e. all the conditions starting from \(s\) will be disjoint.
+Formally, fix any \(s ∈ S\):
+\begin{itemize}
+  \item if \((b₁, o₁) ∈ ϵ(s)\) and \((b₂, o₂) ∈ ϵ(s)\), then \(o₁ ≠ o₂\) implies \(b₁\) is disjoint with \(b₂\);
+  \item if \((b₁, s₁, p) ∈ δ(s)\) and \((b₂, s₂, q) ∈ δ(s)\), then \(s₁ ≠ s₂\) or \(p ≠ q\) implies \(b₁\) is disjoint with \(b₂\);
+  \item for all \(ϵ(s, o)\) is disjoint with all \(b\) s.t. there exists \(s'\) where \((b, s') ∈ δ(s)\).
+\end{itemize}
+
+We list our bisimulation algorithm in~\Cref{alg:symb-bisim}.
+
+\begin{algorithm*}
+  \caption{Symbolic On-the-fly Bisimulation Algorithm}\label{alg:symb-bisim}
+  \begin{algorithmic}
+      \Function{equiv}{$s ∈ S, u ∈ U$}
+      \If {\Call{eq}{$s, u$}} {\Return true}
+      \ElsIf {\Call{knownDead\(_S\)}{$s$}} {\Return \Call{isDead\(_U\)}{$u$}} 
+      \ElsIf {\Call{knownDead\(_U\)}{$u$}} {\Return \Call{isDead\(_S\)}{$s$}} 
+      \Else {}
+      \Return {
+          %\Comment{conditions of ~\Cref{thm:symb-recursive-construction}}
+          \\\vspace{5px}
+          \(\qquad
+          \begin{aligned}
+              & ∀ o ∈ O, ϵ(s, o) ≡ ⋁ ϵ(u, o) \mathrel{\&\!\&} \\  
+              & s \transvia{b ∣ p} s' \text{ and } (b ∧ ρ_U(u)) ≢ 0 \text{ implies \Call{isDead$_S$}{$s'$}} \mathrel{\&\!\&}\\
+              & u \transvia{b ∣ p} u' \text{ and } (ρ_S(s) ∧ b) ≢ 0 \text{ implies \Call{isDead$_U$}{$u'$}} \mathrel{\&\!\&}\\
+              & s \transvia{b ∣ p} s' \text{ and } u \transvia{c ∣ p} u' \text{ implies } \text{\Call{Union}{$s$, $u$}}; \text{\Call{Equiv}{$s', u'$}} \mathrel{\&\!\&}\\
+              & s \transvia{b ∣ p} s' \text{ and } u \transvia{c ∣ q} u' \text{ and } p ≠ q \text{ implies \Call{isDead}{$s'$} and \Call{isDead}{$u'$}}
+          \end{aligned}\)
+      }
+      \EndIf
+      \EndFunction
+  \end{algorithmic}
+\end{algorithm*}
+
+\section{Guarded NetKAT}
+
+Guarded NetKAT~\cite{wasserstein_GUARDEDNETKATSOUNDNESS_2023} is a variant of NetKAT that disallows non-deterministic assignments. 
+For Guarded NetKAT, the possible outputs \(O\) contains acceptance \(\acc{}\) and \(\setc{σ}\), where \(σ: X ↛ V\) is a partial function that assign \(σ(x) ∈ V\) to some variable \(x ∈ X\), when \(x\) is in the domain of \(σ\).
+And we can extend \(σ: X ↛ V\):
+\[σ[x ↦ v](y) ≜ \begin{cases}
+  v & \text{if } x = y \\  
+  σ(x) & \text{otherwise}
+\end{cases}\]
+
+Here is the syntax for Guarded NetKAT over a set of values \(V\) and variable \(X\):
+\begin{align*}
+  \BExp ∋ b, c & ≜ 
+    \false 
+    ∣ \true 
+    ∣ {\color{blue} x = v~(x ∈ X, v ∈ V)} 
+    ∣ b ∧ c 
+    ∣ b ∨ c 
+    ∣ ¬ b 
+    ∣ {\color{blue} b[σ]}\\
+  \Exp ∋ e, f & ≜ 
+    b ∈ \BExp
+    ∣ p ∈ \{\comDup\}
+    ∣ {\color{blue}x := v~(x ∈ X, v ∈ V)}
+    ∣ e +_b f 
+    ∣ e; f 
+    ∣ e^{(b)}
+\end{align*}
+Note that we added a boolean expression \(b[σ]\), but following the syntax of original Guarded NetKAT~\cite{wasserstein_GUARDEDNETKATSOUNDNESS_2023}, we disallow this construct from NetKAT programs. 
+\(b[σ]\) generates a new predicate by plugging in the value of \(σ: X ↛ V\) into the boolean expression \(b\).
+Formally \(b[σ]\) is defined inductively as follows:
+\begin{align*}
+  \false[σ] & ≜ \false & 
+  \true[σ] & ≜ \true \\
+  (x = i)[σ] & ≜ \begin{cases}
+    x = i & x ∉ \dom{σ} \\
+    \true & σ(x) = i \\  
+    \false & σ(x) ≠ i 
+  \end{cases} & 
+  (b ∧ c)[σ] & ≜ b[σ] ∧ c[σ] \\
+  (b ∨ c)[σ] & ≜ b[σ] ∨ c[σ] &
+  (¬ b)[σ] & ≜ ¬ (b[σ])
+\end{align*}
+
+The state in the symbolic Guarded NetKAT automata has two components: \(σ: X ↛ V\) indicating the current known assignment to variables, and \(e: \Exp\) representing the rest of the expression
+\begin{mathpar}
+  \inferrule
+  {\\}
+  {(σ, \comAssert{b})\transOut{b[σ]}{\acc{}}}
+  \and  
+  \inferrule
+  {\\}
+  {(σ, x := v)\transOut{\true}{\setc{σ[x ↦ v]}}}
+  \and  
+  \inferrule
+  {\\}
+  {(σ,\comDup)\transvia{b_σ ∣ \comDup} (σ,\comSkip)}
+  \and  
+  \inferrule
+  {(σ, e) \transOut{c}{o} \\ o ∈ O}
+  {(σ, e +_b f) \transOut{b[σ] ∧ c}{o}}
+  \and  
+  \inferrule
+  {(σ, f) \transOut{c}{o} \\ o ∈ O}
+  {(σ, e +_b f) \transOut{¬b[σ] ∧ c}{o}}
+  \and
+  \inferrule
+  {(σ, e) \transvia{c ∣ p} (σ', e')}
+  {(σ, e +_b f) \transvia{b[σ] ∧ c ∣ p} (σ', e')}
+  \and  
+  \inferrule
+  {(σ, f) \transvia{c ∣ p} (σ', f')}
+  {(σ, e +_b f) \transvia{b[σ] ∧ c ∣ p} (σ', f')}
+  \and  
+  \inferrule
+  {(σ, e) \transOut{b}{\acc{}} \\ (σ, f) \transOut{c}{o} \\ o ∈ O}
+  {(σ, e; f) \transOut{b ∧ c}{o}}
+  \and  
+  \inferrule
+  {(σ, e) \transOut{b}{\acc{}} \\ (σ, f) \transvia{b ∣ p} (σ', f')}
+  {(σ, e; f) \transvia{b ∧ c ∣ p} (σ, f')}
+  \and  
+  \inferrule
+  {(σ, e) \transOut{b}{\setc{σ'}} \\ (σ', f) \transOut{c}{o} \\ o ∈ O}
+  {(σ, e; f) \transOut{b ∧ c}{o}}
+  \and  
+  \inferrule
+  {(σ₁, e) \transOut{b}{\setc{σ₂}} \\ (σ₂, f) \transvia{b ∣ p} (σ₃, f')}
+  {(σ₁, e; f) \transvia{b ∧ c ∣ p} (σ₃, f')}
+  \and  
+  \inferrule
+  {(σ, e) \transvia{b ∣ p} (σ', e')}
+  {(σ, e;f) \transvia{b ∣ p} (σ', e';f)}
+  \and  
+  \inferrule
+  {\\}
+  {(σ, e^{(b)}) \transOut{¬b[σ]}{\acc{}}}
+  \and  
+  \inferrule
+  {(σ₁, e) \transOut{c}{\setc{σ₂}} \\  (σ₂, e^{(b)}) \transvia{a ∣ p} (σ₃, f)}
+  {(σ₁, e^{(b)}) \transvia{c ∧ a ∣ p} (σ₃, f)}
+  \and  
+  \inferrule
+  {(σ, e) \transvia{c ∣ p} (σ', e')}
+  {(σ, e^{(b)}) \transvia{b ∧ c ∣ p} (σ', e'; e^{(b)})}
+\end{mathpar}
+
 \section{Control-Flow GKAT}
+
+\subsection{Syntax}
 
@@
 
-Similar to GKAT, the syntax of CF-GKAT is also two sorted:
+Similar to GKAT, the syntax of CF-GKAT is also two sorted, we highlight the newly added :
 \begin{align*}
-  \BExp & ≜ t ∈ T ∣ 
+  \BExp ∋ b, c ≜{}&
+  \false
+  ∣ \true
+  ∣ t  ∈  T
+  ∣ {\color{blue}x = i\ (i ∈ I)}
+  ∣ b  \lor  c
+  ∣ b  \land  c
+  ∣ ¬ b \\
+  \Exp ∋ e, f ≜{}&
+    \comAssert{b}
+    ∣ p  ∈  Σ
+    ∣ {\color{blue}x := i\ (i ∈ I)}
+    ∣ {\color{blue}e ⋆ f}
+    ∣ e; f
+    ∣ e +_b f
+    ∣ \\
+  &
+    e^{(b)}
+    ∣ {\color{blue} \comBrk}
+    ∣ {\color{blue} \comCont}
+    ∣ {\color{blue} \comRet}
+    ∣ {\color{blue} \comGoto{l}\ (l ∈  L)}
+    ∣ {\color{blue} \comLabel{l}\ (l ∈ L)}
 \end{align*}
-\section{Guarded NetKAT}
+Additionally, for a CF-GKAT expression \(e\) to be a CF-GKAT \emph{program} (or \emph{well-formed expressions}), we enforce the following conditions:
+\begin{itemize}
+  \item If \(\comLabel{l}\) does not occur in \(e\), then so is \(\comGoto{l}\).
+  \item Each \(\comLabel{l}\) only appear once in the program \(e\).
+  \item No \(\comBrk\) or \(\comRet\) outside a while-loop.  
+  \item All programs need to end with a return, i.e. \(e\) need to be of the form \(f; \comRet\) where \(f ∈ \Exp\).
+  \item The \(⋆\) operator cannot appear in the program. 
+\end{itemize}
+In addition to the syntax introduced in the original paper~\cite{zhang_CFGKATEfficientValidation_2025}, we added two additional constructions: \(\comCont\), a common control-flow manipulating structure; and the \emph{unfolding star} \(e ⋆ f\) is an expression that is generated from loop unfolding, hence cannot appear in any CF-GKAT programs.
+We use the common shorthand \(\comSkip\) for \(\comAssert{\true}\).
+On the other hand, \(\comAssert{\false}\) is the divergent or constant rejection program; in fact, it is equivalent to \(\comSkip^{(\true)}\).
+
+\subsection{Label Extraction}
+
+First, we define the label extraction process, where we extract the meaning of each label syntactically: given a label \(l ∈ L\), \((-)_l\) will take an input program, and give the expression that corresponds to the meaning of \(l\) in the input program:
+\begin{align*}
+  (-)_{l}&: \Prog → \Exp \\  
+  (\comLabel{l'})_{l} & ≜ \begin{cases}
+    \comSkip & l = l' \\  
+    \comAssert{\false} & l ≠ l'
+  \end{cases} & 
+  (e +_b f)_{l} & ≜ \begin{cases}
+    (e)_{l} & \text{if \(\comLabel{l}\) occurs in \(e\)}\\
+    (f)_{l} & \text{if \(\comLabel{l}\) occurs in \(f\)}
+  \end{cases} \\  
+  (e; f)_{l} & ≜ \begin{cases}
+    (e)_{l}; f & \text{if \(\comLabel{l}\) occurs in \(e\)}\\
+    (f)_{l} & \text{if \(\comLabel{l}\) occurs in \(f\)}
+  \end{cases} & 
+  (e ⋆ f)_{l} & ≜ \begin{cases}
+    (e)_{l} ⋆ f & \text{if \(\comLabel{l}\) occurs in \(e\)}\\
+    (f)_{l} & \text{if \(\comLabel{l}\) occurs in \(f\)}
+  \end{cases} \\
+  (e^{(b)})_{l} & ≜ 
+    (e)_l ⋆ e^{(b)}
+\end{align*}
+And the label extraction \((e)_l\) for every the other program \(e\) is defined to be \(\comAssert{\false}\).
+Notice the \(⋆\) operation is produced in the loop case, this serves as an example of loop unfolding.
+
+\subsection{Boolean Semantics}
+
+For a boolean expression \(b ∈ \BExp\), the semantics is
+
+\subsection{Derivatives and Loop Resolution}
+
+The symbolic derivatives of CF-GKAT can be defined as follows:
 

### 2025-04-16 POPL26/paper.tex
simplify semantics

@@
 Guarded NetKAT~\cite{wasserstein_GUARDEDNETKATSOUNDNESS_2023} is a variant of NetKAT that disallows non-deterministic assignments. 
-For Guarded NetKAT, the possible outputs \(O\) contains acceptance \(\acc{}\) and \(\setc{σ}\), where \(σ: X ↛ V\) is a partial function that assign \(σ(x) ∈ V\) to some variable \(x ∈ X\), when \(x\) is in the domain of \(σ\).
+For Guarded NetKAT, the possible outputs \(O\) are \emph{partial assignments} \(σ: X ↛ V\), which are partial functions that encodes all the \emph{known} variable assignment at this stage of the execution.
 And we can extend \(σ: X ↛ V\):
@@
   {\\}
-  {(σ, \comAssert{b})\transOut{b[σ]}{\acc{}}}
+  {(σ, \comAssert{b})\transOut{b[σ]}{σ}}
   \and  
@@
   {\\}
-  {(σ, x := v)\transOut{\true}{\setc{σ[x ↦ v]}}}
+  {(σ, x := v)\transOut{\true}{σ[x ↦ v]}}
   \and  
@@
   \inferrule
-  {(σ, e) \transOut{c}{o} \\ o ∈ O}
-  {(σ, e +_b f) \transOut{b[σ] ∧ c}{o}}
+  {(σ, e) \transOut{c}{σ'}}
+  {(σ, e +_b f) \transOut{b[σ] ∧ c}{σ'}}
   \and  
   \inferrule
-  {(σ, f) \transOut{c}{o} \\ o ∈ O}
-  {(σ, e +_b f) \transOut{¬b[σ] ∧ c}{o}}
+  {(σ, f) \transOut{c}{σ'}}
+  {(σ, e +_b f) \transOut{¬b[σ] ∧ c}{σ'}}
   \and
@@
   \inferrule
-  {(σ, e) \transOut{b}{\acc{}} \\ (σ, f) \transOut{c}{o} \\ o ∈ O}
-  {(σ, e; f) \transOut{b ∧ c}{o}}
+  {(σ₁, e) \transOut{b}{σ₂} \\ (σ₂, f) \transOut{c}{σ₃}}
+  {(σ₁, e; f) \transOut{b ∧ c}{σ₃}}
   \and  
   \inferrule
-  {(σ, e) \transOut{b}{\acc{}} \\ (σ, f) \transvia{b ∣ p} (σ', f')}
-  {(σ, e; f) \transvia{b ∧ c ∣ p} (σ, f')}
-  \and  
-  \inferrule
-  {(σ, e) \transOut{b}{\setc{σ'}} \\ (σ', f) \transOut{c}{o} \\ o ∈ O}
-  {(σ, e; f) \transOut{b ∧ c}{o}}
-  \and  
-  \inferrule
-  {(σ₁, e) \transOut{b}{\setc{σ₂}} \\ (σ₂, f) \transvia{b ∣ p} (σ₃, f')}
+  {(σ₁, e) \transOut{b}{σ₂} \\ (σ₂, f) \transvia{b ∣ p} (σ₃, f')}
   {(σ₁, e; f) \transvia{b ∧ c ∣ p} (σ₃, f')}
@@
   {\\}
-  {(σ, e^{(b)}) \transOut{¬b[σ]}{\acc{}}}
+  {(σ, e^{(b)}) \transOut{¬b[σ]}{σ}}
   \and  
   \inferrule
-  {(σ₁, e) \transOut{c}{\setc{σ₂}} \\  (σ₂, e^{(b)}) \transvia{a ∣ p} (σ₃, f)}
-  {(σ₁, e^{(b)}) \transvia{c ∧ a ∣ p} (σ₃, f)}
+  {(σ₁, e) \transOut{c}{σ₂} \\  (σ₂, e^{(b)}) \transOut{a}{σ₃}}
+  {(σ₁, e^{(b)}) \transOut{b[σ] ∧ c ∧ a}{σ₃}}
+  \and  
+  \inferrule
+  {(σ₁, e) \transOut{c}{σ₂} \\  (σ₂, e^{(b)}) \transvia{a ∣ p} (σ₃, f)}
+  {(σ₁, e^{(b)}) \transvia{b[σ] ∧ c ∧ a ∣ p} (σ₃, f)}
   \and  
@@
   {(σ, e) \transvia{c ∣ p} (σ', e')}
-  {(σ, e^{(b)}) \transvia{b ∧ c ∣ p} (σ', e'; e^{(b)})}
+  {(σ, e^{(b)}) \transvia{b[σ] ∧ c ∣ p} (σ', e'; e^{(b)})}
 \end{mathpar}
+Notice that when \((σ₁, e) \transOut{c}{σ₂}\), the derivative rule with the source \((σ₁, e^{(b)})\) appears to be self-referential. 
+However, since there are only finitely many partial assignments \(σ: X ↛ V\), for any finite set \(X\) and \(V\); there are only finitely many cases to check.
+Practically, we use the following recursive algorithm to compute \(ϵ(σ, e^{(b)})\), where \(Ψ\) denotes the set of all explored \(σ\).
+\begin{align*}
+  ϵ̂(σ, e^{(b)}, Ψ) & ≜ \begin{cases}
+    ∅ & \text{if } σ ∈ Ψ \\  
+    \begin{aligned}
+      & \{(¬b[σ], σ)\} ∪ \\
+      & \{(b[σ] ∧ c ∧ a, σ₂) ∣ (a, σ₂) ∈ ϵ̂(σ₁, e^{(b)}, Ψ ∪ \{σ\})\}
+    \end{aligned}
+     & \text{else, if } (σ, e)\transOut{c}{σ₁}\\
+  \end{cases}\\
+  ϵ(σ, e^{(b)}) & ≜ ϵ̂(σ, e^{(b)}, ∅)
+\end{align*}
 
